### 2013-01-28 1_introduction/introduction.tex
Adding LaTeX files

% this file is called up by thesis.tex
% content in this file will be fed into the main document

%: ----------------------- introduction file header -----------------------
\chapter{Introduction}

% the code below specifies where the figures are stored
\ifpdf
    \graphicspath{{1_introduction/figures/PNG/}{1_introduction/figures/PDF/}{1_introduction/figures/}}
\else
    \graphicspath{{1_introduction/figures/EPS/}{1_introduction/figures/}}
\fi

% ----------------------------------------------------------------------
%: ----------------------- introduction content ----------------------- 
% ----------------------------------------------------------------------

%: ----------------------- HELP: latex document organisation
% the commands below help you to subdivide and organise your thesis
%    \chapter{}       = level 1, top level
%    \section{}       = level 2
%    \subsection{}    = level 3
%    \subsubsection{} = level 4
% note that everything after the percentage sign is hidden from output

\section{Signal pre-processing and sensor fusion for timed patterns} % section 
%headings are printed smaller than chapter names
% intro

\section{Temporal segmentation}
This section will give an introduction and in-depth analysis of temporal 
segmentation. test. nog een test. test

\subsection{Aims of segmentation}
When processing and analyzing time series of data, e.g. motion measurements, 
stock market fluctuations or natural language, first a low-level division 
between the discriminative parts of the stream must be made. One can view this 
as splitting the series into the \emph{atoms}, which are the building blocks 
of the total stream. These building blocks will be the aggregation of 
non-overlapping, internally homogeneous segments \cite{himberg2001time}. This 
means that the data points inside a segment should have some resemblance 
relation to each other and their difference lies between some boundary. The 
process of segmenting can be viewed as a subproblem to context analysis of 
time series. Temporal segmentation is closely related to temporal clustering, 
although it is a stricter, and simpler, process. Whereby clustering only 
restricts the data points on their distance relation (as used in a Voronoi 
diagram), within a segment the data points must also be contiguous.

The task of segmentation can be performed in a manual 
matter, by cutting and labeling parts of the stream into coherent parts. This 
would require human (expert) knowledge and does not yield a clear cut because 
of ambiguity. With increasing storage abilities and easier motion capture 
systems, there is a desire for automated systems which perform the 
segmentation task unsupervised. Some algorithms used have the (often desired) 
side-effect of also clustering the segments, such that classes of segments can 
be discovered in the time series. These algorithms would not only be able to 
make a distinction between walking, sitting and walking, but would also 
recognize the reappearance of the walking activity.

\subsection{Formal definition}
Formally, temporal segmentation is dividing a time series $s$, which consists 
of $N$ samples $\mathbf{x}(1),\mathbf{x}(2),\dots,\mathbf{x}(N)$ from 
$\mathbf{R}^d$. Individual \emph{segments} are referenced by $s(a,b)$, 
consisting of the consecutive samples 
$\mathbf{x}(a),\mathbf{x}(a+1),\dots,\mathbf{x}(b)$, $a \le b$. Let $s_1 = 
s(a,b)$ and $s_2 = s(b+1,c)$ be two segments, then their concatenation is 
$s_1s_2 = s(a,c)$. A segmentation $S$ of $s$ consists of a sequence of $k$ 
non-empty segments $s_1s_2 \dots s_k = s$.
This notation has close resemblance to the notation as used in 
\cite{himberg2001time}.

As stated, informally each segment should be internally homogeneous. This can 
formally be measured with an cost function $F$, indication the heterogeneity 
of a segment. The overall aim is to minimize the cost $F$. The cost of a 
segment is a function from the data points and the number of data points $n = 
b - a + 1$ and is expressed as
\begin{equation} \label{eq:segment_cost}
\mathrm{cost}_F (s(a,b)) = F(\mathbf{x};n|\mathbf{x} \in s(a,b))
\end{equation}
The cost of a \emph{k-segmentation} $S$ is the summation of the costs of the 
$k$ segments:
\begin{equation} \label{eq:segmentation_cost}
\mathrm{Cost}_F (s_1 s_2 \dots s_k) = \sum_{i=1}^{k} \mathrm{cost}_F (s_k)
\end{equation}
With the objective of minimizing the cost function, the optimal 
$k$-segmentation $S_F^\mathit{opt}(s;k)$ is the segmentation with minimal 
$\mathrm{Cost}_F(s_1 s_2 \dots s_k)$ over all possible $k$-segmentations.

The cost function, to calculate the heterogeneity of a (set of) segment(s), 
can 
be any function. A simple and natural function would be the sum of variances 
of the segments. The overall cost function would then be
\begin{equation} \label{eq:cost_variances}
\mathrm{Cost}_V = \frac{1}{N} \sum_{i=i}^{k} \sum_{j=c_{i-1}+1}^{c_i} \| 
\mathbf{x}(j) - \mu_i \|^2
\end{equation} 
where $\mu_i$ is the mean vector of data points in segment $s_i$.

\subsection{Application in research fields}

[CHARACTERISTICS OF HUMAN MOTION]
temporal variability, invariance over time, metrics over actions.

[COMPUTER VISION]

[GRAPHICS/VIDEO]

[DATA-MINING]

[MODEL BASED]

To analyze time series it is often preferred to divide the stream in segments 
of correlated data. After dividing, each segment represent a period in time in 
which the same activity is performed. Or, stated otherwise, it results in 
transitions moments between activities.

Many fields of research have been active in the unsupervised segmentation of 
data. Many authors rely on a form of Principal Component Analysis (PCA), as 
used a.o. in \cite{barbivc2004segmenting}. Often PCA is used to reduce the 
dimensionality of the data being processed [REFERENCE] by only using the top 
$r$ dimensions to describe the data set. It is observed that data series of 
simple motions have a lower dimensionality then complexer motions. When a 
simple (repetitive) motion is about to end and fluently transforms in a new 
motion, there will be a window of time in which a high dimensionality will be 
present, due to the new motion. After this period of transition, the 
dimensionality will decrease, since only the new simple motion is present in 
the window of time. The first algorithm of \cite{barbivc2004segmenting} is 
based on this principle.

Given a set of data points, a lower dimensional hyperplane can by constructed 
to which the data points can be projected. This projection on a lower 
dimension introduces a error to the original position. When the error is 
fixed, less dimensions are needed for simple motions in which movements of 
body parts are highly correlated. For segments in which the data points are 
lesser correlated, e.g. because of transition state, a higher degree of 
dimensions of the hyperplane is needed to represent the data with equal error 
degree.

[SKIPPING]
By analyzing the derivative of the error rate, the algorithm is capable of 
selecting transition points and thus segmenting the data. When a data point 
$d_i$ is more than $k_\sigma=3$ times the standard deviation from the average 
over the previous data points $d_j, j<i$, then a cut is assigned to that frame.

A second approach in \cite{barbivc2004segmenting} uses the probabilistic 
variant of PCA (PPCA) to model the data set as a Gaussian distribution instead 
of ignoring the frames which do not fit in the subspace. Over windows of 
frames the mean and variance are calculated. In a forward manner the 
Mahalanobis distance of a new window of frames is calculated, which represents 
the likelihood of the new window belonging to the same segment as the original 
widow. When the distance decreases, the likelihood increases which happens 
when the motions in the becomes more homogeneous. When a peak in the distance 
is reached, the new window of frames indicates a heterogeneous collection of 
motions in the window and thus a low likelihood of membership and a indication 
of a transition. In order to distinct activities and subactivities (which 
require a subset of motions is a distinct activity) the algorithm is also 
processed backward over the data series.

The third algorithm in \cite{barbivc2004segmenting} is based on the 
observation that data points (frames) tend to form clusters in the space. 
These clusters are represented by $k$ Gaussian distributions for which each 
the Expectation-Maximization (EM) algorithm estimates the mean $m_j$, 
covariance matrix $\sum_{j}$ and prior $\pi_j$. With all the Gaussian 
distributions estimated, the data points are assigned to the cluster with the 
highest membership likelihood. When two consecutive frames $x_i$ and $x_{i+1}$ 
belong to different clusters, a transition of activities is recognized. Note 
that this algorithm succeeds in segmenting the data and also labels the 
similar simple activities.

A drawback in this system, and many others which implement a variant of the 
$k$-means algorithm, is that the number of clusters $k$ need to be 
predetermined. To cope with this, often the algorithm is performed multiple 
times for different values of $k$. Using some criteria, e.g. the Bayesian 
Information Criterion \cite{pelleg2000x} or the Davies-Bouldin Index which 
guides $k$-means clustering as used in \cite{krause2003unsupervised}.

\subsection{Principal Component Analysis}

\subsection{Segmentation as clustering}
In the previous section the discussed methods all relied on the stream of data 
points and tried to find cuts the discriminate between successive different 
type of activities. An other approach is to consider the tasks of segmentation 
as a type of clustering \cite{zhou2008aligned}. In clustering the objective is 
to assign labels, or classes, to all the data points indication a similar type 
of activity. A clustering $\mathcal{L}$ is thereby more informative then a 
segmentation but is also harder to produce.

A clustering $\mathcal{L}$ is generated from a sequence of elements 
$\mathbf{X}$ which is decomposed in $m$ disjoint segments, each belonging to 
one of the $k$ classes. A segment $\mathbf{Y}_i \hat{=} 
\mathbf{X}_{[s_i,s_{i+1})}$ is composed of frames from position $s_i$ to 
$s_{i+1}$. A vector $g_{ci} = 1$ indicates class membership if $\mathbf{Y}_i$ 
belongs to class $c$, otherwise $g_{ci} = 0$.

When regarding segmentation of human motion as a task of clustering the 
difficulty is to model the temporal variability of actions 
and defining a robust metric between temporal actions. To overcome this, 
\cite{zhou2008aligned} introduces Aligned Cluster Analysis (ACA), by minimizing
\begin{equation} \label{eq:ACA}
J_{\mathit{ACA}}(\mathbf{G},\mathbf{s}) = 
\sum_{c=i}^{K} \sum_{i=1}^{m} g_{ci} \mathit{dist}_c 
(\mathbf{X}_{[s_i,s_{i+1})})
\end{equation}

The characteristic of ACA is that is enables segments to span over different 
number of data points, whereas the standard kernel $k$-means algorithm results 
in equally sized segments. [!!! TRUE?!] The second difference is that the 
kernel used in $\mathit{dist}_c$ to measure the distance from a segment to the 
class which it is assigned to uses the Dynamic Time Alignment Kernel 
[REFERENCE?] to measure between time series.

\section{Temporal pattern recognition}

\subsection{Dynamic Time Warping}
Used to measure similarity between time sequences. Exact matching is 
high-cost, so approximations such as Minimum Bounding Rectangles are used.

\subsection{$k$-means clustering}
[EXPLAIN] divide data set $n$ into $k$ clusters.

Among many unsupervised clustering techniques, $k$-means is successfully 
applied to large data sets. It is simple to implement and linear in time 
complexity so computationally attractive \cite{jain1999data}. A drawback of 
the method is that the results of the algorithm greatly depends on the initial 
configuration (the data points which will act as centroids) and the number of 
cluster $k$ must be determined beforehand.

Generally, the $k$-means methods will minimize the squared error for a 
clustering $\mathcal{L}$ criterion which is defined as the distance from the 
data points the centroid for each cluster in $\mathcal{K}$. This is expressed 
as optimizing to a local optimum the energy function
\begin{equation} \label{eq:k-means energy}
e^2(\mathcal{K},\mathcal{L}) = 
\sum_{j=i}^{K}\sum_{i=1}^{n_j}\|\mathbf{x}_i^{(j)} - 
\mathbf{c}_j\|^2
\end{equation}

There are several limitation on the $k$-means method. One of these is that 
only spherical shapes of cluster can be generated. One of the extensions is 
kernel $k$-means \cite{scholkopf1998nonlinear}, which implicitly projects the 
data points to a higher dimension and thereby is able to form irregular shaped 
cluster.

\subsection{Self-organizing Map}

\subsection{Support Vector Machine}

\subsection{Na\"{i}ve Bayes}

\section{Unsupervised clustering of temporal patterns}

%: ----------------------- HELP: special characters
% above you can see how special characters are coded; e.g. $\alpha$
% below are the most frequently used codes:
%$\alpha$  $\beta$  $\gamma$  $\delta$

%$^{chars to be superscripted}$  OR $^x$ (for a single character)
%$_{chars to be suberscripted}$  OR $_x$

%>  $>$  greater,  <  $<$  less
%≥  $\ge$  greater than or equal, ≤  $\ge$  lesser than or equal
%~  $\sim$  similar to

%$^{\circ}$C   ° as in degree C
%±  \pm     plus/minus sign

%$\AA$     produces  Å (Angstrom)

% dextran, starch, glycogen continued
%Starch of plants and glycogen of animals consists of $\alpha$-1,4-glycosidic 
%glucose polymers \cite{lastname07}. See figure \ref{largepotato} for %a 
%%comparison of glucose polymer structure and chemistry.

%Two references can be placed separated by a comma \cite{lastname07,name06}.

%: ----------------------- HELP: references
% References can be links to figures, tables, sections, or references.
% For figures, tables, and text you define the target of the link with \label{XYZ}. Then you call cross-link with the command \ref{XYZ}, as above
% Citations are bound in a very similar way with \cite{XYZ}. You store your references in a BibTex file with a programme like BibDesk.

%\figuremacro{largepotato}{A common glucose polymers}{The figure shows starch 
%granules in potato cells, taken from 
%%%\href{http://molecularexpressions.com/micro/gallery/burgersnfries/burgersnfries4.html}{Molecular
%% Expressions}.}

%: ----------------------- HELP: adding figures with macros
% This template provides a very convenient way to add figures with minimal code.
% \figuremacro{1}{2}{3}{4} calls up a series of commands formating your image.
% 1 = name of the file without extension; PNG, JPEG is ok; GIF doesn't work
% 2 = title of the figure AND the name of the label for cross-linking
% 3 = caption text for the figure

%: ----------------------- HELP: www links
% You can also see above how, www links are placed
% \href{http://www.something.net}{link text}

%\figuremacroW{lar gepotato}{Title}{Caption}{0.8}
% variation of the above macro with a width setting
% \figuremacroW{1}{2}{3}{4}
% 1-3 as above
% 4 = size relative to text width which is 1; use this to reduce figures

%: ----------------------- HELP: lists
% This is how you generate lists in LaTeX.
% If you replace {itemize} by {enumerate} you get a numbered list.

%: ----------------------- HELP: tables
% Directly coding tables in latex is tiresome. See below.
% I would recommend using a converter macro that allows you to make the table in Excel and convert them into latex code which you can then paste into your doc.
% This is the link: http://www.softpedia.com/get/Office-tools/Other-Office-Tools/Excel2Latex.shtml
% It's a Excel template file containing a macro for the conversion.

%\begin{table}[htdp]
%\centering
%\begin{tabular}{ccc} % ccc means 3 columns, all centered; alternatives are l, 
%%r
%
%{\bf Gene} & {\bf GeneID} & {\bf Length} \\ 
%% & denotes the end of a cell/column, \\ changes to next table row
%\hline % draws a line under the column headers
%
%human latexin & 1234 & 14.9 kbps \\
%mouse latexin & 2345 & 10.1 kbps \\
%rat latexin   & 3456 & 9.6 kbps \\
%% Watch out. Every line must have 3 columns = 2x &. 
%% Otherwise you will get an error.
%
%\end{tabular}
%\caption[title of table]{\textbf{title of table} - Overview of latexin genes.}
%% You only need to write the title twice if you don't want it to appear in 
%%bold in the list of tables.
%\label{latexin_genes} % label for cross-links with \ref{latexin_genes}
%\end{table}

% There you go. You already know the most important things.

% ----------------------------------------------------------------------

### 2013-01-30 Chapters/Chapter1.tex
New template, merged old text

% Chapter 1

\chapter{Introduction} % Main chapter title

\label{Chapter1} % For referencing the chapter elsewhere, use \ref{Chapter1}

\lhead{Chapter 1. \emph{Introduction}} % This is for the header on each page - 
%perhaps a shortened title

%----------------------------------------------------------------------------------------

\section{Problem statement}
*** Give context problems, means to reach (e.g. monitoring of patients, 
getting conclusions over health and activity).
%----------------------------------------------------------------------------------------

\section{Literature review}
Look into earlier application mentioned in the literature to this kind of 
problems. Look for similarities in the problem and address where the 
techniques used fail or are not applicable.

%----------------------------------------------------------------------------------------

\section{Learning \LaTeX{}}

### 2013-01-30 Chapters/Chapter2.tex
New template, merged old text

--- 2013-01-28 1_introduction/introduction.tex
@@
-% this file is called up by thesis.tex
-% content in this file will be fed into the main document
+% Chapter 2
 
-%: ----------------------- introduction file header -----------------------
-\chapter{Introduction}
+\chapter{Techniques} % Main chapter title
 
-% the code below specifies where the figures are stored
-\ifpdf
-    \graphicspath{{1_introduction/figures/PNG/}{1_introduction/figures/PDF/}{1_introduction/figures/}}
-\else
-    \graphicspath{{1_introduction/figures/EPS/}{1_introduction/figures/}}
-\fi
+\label{Chapter2} % For referencing the chapter elsewhere, use \ref{Chapter1}
 
-% ----------------------------------------------------------------------
-%: ----------------------- introduction content ----------------------- 
-% ----------------------------------------------------------------------
+\lhead{Chapter 2. \emph{Techniques}} % This is for the header on each 
+%page - 
+%perhaps a shortened title
 
-%: ----------------------- HELP: latex document organisation
-% the commands below help you to subdivide and organise your thesis
-%    \chapter{}       = level 1, top level
-%    \section{}       = level 2
-%    \subsection{}    = level 3
-%    \subsubsection{} = level 4
-% note that everything after the percentage sign is hidden from output
+%----------------------------------------------------------------------------------------
 
-\section{Signal pre-processing and sensor fusion for timed patterns} % section 
-%headings are printed smaller than chapter names
-% intro
+\section{Signal pro-processing and sensor fusion for timed patterns}
 
-\section{Temporal segmentation}
+\section{Temporal Segmentation}
+
 This section will give an introduction and in-depth analysis of temporal 
@@
 
+\section{Clustering}
+
 \section{Temporal pattern recognition}
@@
 \section{Unsupervised clustering of temporal patterns}
-
-%: ----------------------- HELP: special characters
-% above you can see how special characters are coded; e.g. $\alpha$
-% below are the most frequently used codes:
-%$\alpha$  $\beta$  $\gamma$  $\delta$
-
-%$^{chars to be superscripted}$  OR $^x$ (for a single character)
-%$_{chars to be suberscripted}$  OR $_x$
-
-%>  $>$  greater,  <  $<$  less
-%≥  $\ge$  greater than or equal, ≤  $\ge$  lesser than or equal
-%~  $\sim$  similar to
-
-%$^{\circ}$C   ° as in degree C
-%±  \pm     plus/minus sign
-
-%$\AA$     produces  Å (Angstrom)
-
-% dextran, starch, glycogen continued
-%Starch of plants and glycogen of animals consists of $\alpha$-1,4-glycosidic 
-%glucose polymers \cite{lastname07}. See figure \ref{largepotato} for %a 
-%%comparison of glucose polymer structure and chemistry.
-
-%Two references can be placed separated by a comma \cite{lastname07,name06}.
-
-%: ----------------------- HELP: references
-% References can be links to figures, tables, sections, or references.
-% For figures, tables, and text you define the target of the link with \label{XYZ}. Then you call cross-link with the command \ref{XYZ}, as above
-% Citations are bound in a very similar way with \cite{XYZ}. You store your references in a BibTex file with a programme like BibDesk.
-
-%\figuremacro{largepotato}{A common glucose polymers}{The figure shows starch 
-%granules in potato cells, taken from 
-%%%\href{http://molecularexpressions.com/micro/gallery/burgersnfries/burgersnfries4.html}{Molecular
-%% Expressions}.}
-
-%: ----------------------- HELP: adding figures with macros
-% This template provides a very convenient way to add figures with minimal code.
-% \figuremacro{1}{2}{3}{4} calls up a series of commands formating your image.
-% 1 = name of the file without extension; PNG, JPEG is ok; GIF doesn't work
-% 2 = title of the figure AND the name of the label for cross-linking
-% 3 = caption text for the figure
-
-%: ----------------------- HELP: www links
-% You can also see above how, www links are placed
-% \href{http://www.something.net}{link text}
-
-%\figuremacroW{lar gepotato}{Title}{Caption}{0.8}
-% variation of the above macro with a width setting
-% \figuremacroW{1}{2}{3}{4}
-% 1-3 as above
-% 4 = size relative to text width which is 1; use this to reduce figures
-
-%: ----------------------- HELP: lists
-% This is how you generate lists in LaTeX.
-% If you replace {itemize} by {enumerate} you get a numbered list.
-
-%: ----------------------- HELP: tables
-% Directly coding tables in latex is tiresome. See below.
-% I would recommend using a converter macro that allows you to make the table in Excel and convert them into latex code which you can then paste into your doc.
-% This is the link: http://www.softpedia.com/get/Office-tools/Other-Office-Tools/Excel2Latex.shtml
-% It's a Excel template file containing a macro for the conversion.
-
-%\begin{table}[htdp]
-%\centering
-%\begin{tabular}{ccc} % ccc means 3 columns, all centered; alternatives are l, 
-%%r
-%
-%{\bf Gene} & {\bf GeneID} & {\bf Length} \\ 
-%% & denotes the end of a cell/column, \\ changes to next table row
-%\hline % draws a line under the column headers
-%
-%human latexin & 1234 & 14.9 kbps \\
-%mouse latexin & 2345 & 10.1 kbps \\
-%rat latexin   & 3456 & 9.6 kbps \\
-%% Watch out. Every line must have 3 columns = 2x &. 
-%% Otherwise you will get an error.
-%
-%\end{tabular}
-%\caption[title of table]{\textbf{title of table} - Overview of latexin genes.}
-%% You only need to write the title twice if you don't want it to appear in 
-%%bold in the list of tables.
-%\label{latexin_genes} % label for cross-links with \ref{latexin_genes}
-%\end{table}
-
-% There you go. You already know the most important things.
-
-% ----------------------------------------------------------------------

### 2013-01-30 Chapters/Chapter3.tex
New template, merged old text

% Chapter 3

\chapter{Experiments} % Main chapter title

\label{Chapter3} % For referencing the chapter elsewhere, use \ref{Chapter1}

\lhead{Chapter 3. \emph{Experiments}} % This is for the header on each 
%page - 
%perhaps a shortened title

%----------------------------------------------------------------------------------------

\section{}

### 2013-01-30 main.tex
New template, merged old text

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Thesis 
% LaTeX Template
% Version 1.3 (21/12/12)
%
% This template has been downloaded from:
% http://www.latextemplates.com
%
% Original authors:
% Steven Gunn 
% http://users.ecs.soton.ac.uk/srg/softwaretools/document/templates/
% and
% Sunil Patel
% http://www.sunilpatel.co.uk/thesis-template/
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
% Note:
% Make sure to edit document variables in the Thesis.cls file
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[11pt, a4paper, oneside]{Thesis} % Paper size, default font size and one-sided paper

\graphicspath{{./Pictures/}} % Specifies the directory where pictures are stored

\usepackage[square, numbers, comma, sort&compress]{natbib} % Use the natbib reference package - read up on this to edit the reference style; if you want text (e.g. Smith et al., 2012) for the in-text references (instead of numbers), remove 'numbers' 
\hypersetup{urlcolor=blue, colorlinks=true} % Colors hyperlinks in blue - change to black if annoying
\title{\ttitle} % Defines the thesis title - don't touch this

\begin{document}

\frontmatter % Use roman page numbering style (i, ii, iii, iv...) for the pre-content pages

\setstretch{1.3} % Line spacing of 1.3

% Define the page headers using the FancyHdr package and set up for one-sided printing
\fancyhead{} % Clears all page headers and footers
\rhead{\thepage} % Sets the right side header to show the page number
\lhead{} % Clears the left side page header

\pagestyle{fancy} % Finally, use the "fancy" page style to implement the FancyHdr headers

\newcommand{\HRule}{\rule{\linewidth}{0.5mm}} % New command to make the lines in the title page

% PDF meta-data
\hypersetup{pdftitle={\ttitle}}
\hypersetup{pdfsubject=\subjectname}
\hypersetup{pdfauthor=\authornames}
\hypersetup{pdfkeywords=\keywordnames}

%----------------------------------------------------------------------------------------
%	TITLE PAGE
%----------------------------------------------------------------------------------------

\begin{titlepage}
\begin{center}

\textsc{\LARGE \univname}\\[1.5cm] % University name
\textsc{\Large Doctoral Thesis}\\[0.5cm] % Thesis type

\HRule \\[0.4cm] % Horizontal line
{\huge \bfseries \ttitle}\\[0.4cm] % Thesis title
\HRule \\[1.5cm] % Horizontal line

\begin{minipage}{0.4\textwidth}
\begin{flushleft} \large
\emph{Author:}\\
\href{http://www.johnsmith.com}{\authornames} % Author name - remove the \href bracket to remove the link
\end{flushleft}
\end{minipage}
\begin{minipage}{0.4\textwidth}
\begin{flushright} \large
\emph{Supervisor:} \\
\href{http://www.jamessmith.com}{\supname} % Supervisor name - remove the \href bracket to remove the link  
\end{flushright}
\end{minipage}\\[3cm]

\large \textit{A thesis submitted in fulfilment of the requirements\\ for the degree of \degreename}\\[0.3cm] % University requirement text
\textit{in the}\\[0.4cm]
\groupname\\\deptname\\[2cm] % Research group name and department name

{\large \today}\\[4cm] % Date
%\includegraphics{Logo} % University/department logo - uncomment to place it

\vfill
\end{center}

\end{titlepage}

%----------------------------------------------------------------------------------------
%	DECLARATION PAGE
%	Your institution may give you a different text to place here
%----------------------------------------------------------------------------------------

\Declaration{

\addtocontents{toc}{\vspace{1em}} % Add a gap in the Contents, for aesthetics

I, \authornames, declare that this thesis titled, '\ttitle' and the work presented in it are my own. I confirm that:

\begin{itemize} 
\item[\tiny{$\blacksquare$}] This work was done wholly or mainly while in candidature for a research degree at this University.
\item[\tiny{$\blacksquare$}] Where any part of this thesis has previously been submitted for a degree or any other qualification at this University or any other institution, this has been clearly stated.
\item[\tiny{$\blacksquare$}] Where I have consulted the published work of others, this is always clearly attributed.
\item[\tiny{$\blacksquare$}] Where I have quoted from the work of others, the source is always given. With the exception of such quotations, this thesis is entirely my own work.
\item[\tiny{$\blacksquare$}] I have acknowledged all main sources of help.
\item[\tiny{$\blacksquare$}] Where the thesis is based on work done by myself jointly with others, I have made clear exactly what was done by others and what I have contributed myself.\\
\end{itemize}

Signed:\\
\rule[1em]{25em}{0.5pt} % This prints a line for the signature

Date:\\
\rule[1em]{25em}{0.5pt} % This prints a line to write the date
}

\clearpage % Start a new page

%----------------------------------------------------------------------------------------
%	QUOTATION PAGE
%----------------------------------------------------------------------------------------

\pagestyle{empty} % No headers or footers for the following pages

\null\vfill % Add some space to move the quote down the page a bit

\textit{``Thanks to my solid academic training, today I can write hundreds of words on virtually any topic without possessing a shred of information, which is how I got a good job in journalism."}

\begin{flushright}
Dave Barry
\end{flushright}

\vfill\vfill\vfill\vfill\vfill\vfill\null % Add some space at the bottom to position the quote just right

\clearpage % Start a new page

%----------------------------------------------------------------------------------------
%	ABSTRACT PAGE
%----------------------------------------------------------------------------------------

\addtotoc{Abstract} % Add the "Abstract" page entry to the Contents

\abstract{\addtocontents{toc}{\vspace{1em}} % Add a gap in the Contents, for aesthetics

The Thesis Abstract is written here (and usually kept to just this page). The page is kept centered vertically so can expand into the blank space above the title too\ldots
}

\clearpage % Start a new page

%----------------------------------------------------------------------------------------
%	ACKNOWLEDGEMENTS
%----------------------------------------------------------------------------------------

\setstretch{1.3} % Reset the line-spacing to 1.3 for body text (if it has changed)

\acknowledgements{\addtocontents{toc}{\vspace{1em}} % Add a gap in the Contents, for aesthetics

The acknowledgements and the people to thank go here, don't forget to include your project advisor\ldots
}
\clearpage % Start a new page

%----------------------------------------------------------------------------------------
%	LIST OF CONTENTS/FIGURES/TABLES PAGES
%----------------------------------------------------------------------------------------

\pagestyle{fancy} % The page style headers have been "empty" all this time, now use the "fancy" headers as defined before to bring them back

\lhead{\emph{Contents}} % Set the left side page header to "Contents"
\tableofcontents % Write out the Table of Contents

\lhead{\emph{List of Figures}} % Set the left side page header to "List of Figures"
\listoffigures % Write out the List of Figures

\lhead{\emph{List of Tables}} % Set the left side page header to "List of Tables"
\listoftables % Write out the List of Tables

%----------------------------------------------------------------------------------------
%	ABBREVIATIONS
%----------------------------------------------------------------------------------------

\clearpage % Start a new page

\setstretch{1.5} % Set the line spacing to 1.5, this makes the following tables easier to read

\lhead{\emph{Abbreviations}} % Set the left side page header to "Abbreviations"
\listofsymbols{ll} % Include a list of Abbreviations (a table of two columns)
{
\textbf{LAH} & \textbf{L}ist \textbf{A}bbreviations \textbf{H}ere \\
%\textbf{Acronym} & \textbf{W}hat (it) \textbf{S}tands \textbf{F}or \\
}

%----------------------------------------------------------------------------------------
%	PHYSICAL CONSTANTS/OTHER DEFINITIONS
%----------------------------------------------------------------------------------------

\clearpage % Start a new page

\lhead{\emph{Physical Constants}} % Set the left side page header to "Physical Constants"

\listofconstants{lrcl} % Include a list of Physical Constants (a four column table)
{
Speed of Light & $c$ & $=$ & $2.997\ 924\ 58\times10^{8}\ \mbox{ms}^{-\mbox{s}}$ (exact)\\
% Constant Name & Symbol & = & Constant Value (with units) \\
}

%----------------------------------------------------------------------------------------
%	SYMBOLS
%----------------------------------------------------------------------------------------

\clearpage % Start a new page

\lhead{\emph{Symbols}} % Set the left side page header to "Symbols"

\listofnomenclature{lll} % Include a list of Symbols (a three column table)
{
$a$ & distance & m \\
$P$ & power & W (Js$^{-1}$) \\
% Symbol & Name & Unit \\

& & \\ % Gap to separate the Roman symbols from the Greek

$\omega$ & angular frequency & rads$^{-1}$ \\
% Symbol & Name & Unit \\
}

%----------------------------------------------------------------------------------------
%	DEDICATION
%----------------------------------------------------------------------------------------

\setstretch{1.3} % Return the line spacing back to 1.3

\pagestyle{empty} % Page style needs to be empty for this page

\dedicatory{For/Dedicated to/To my\ldots} % Dedication text

\addtocontents{toc}{\vspace{2em}} % Add a gap in the Contents, for aesthetics

%----------------------------------------------------------------------------------------
%	THESIS CONTENT - CHAPTERS
%----------------------------------------------------------------------------------------

\mainmatter % Begin numeric (1,2,3...) page numbering

\pagestyle{fancy} % Return the page headers back to the "fancy" style

% Include the chapters of the thesis as separate files from the Chapters folder
% Uncomment the lines as you write the chapters

\input{./Chapters/Chapter1}
\input{./Chapters/Chapter2} 
\input{./Chapters/Chapter3}
\input{./Chapters/Chapter4} 
\input{./Chapters/Chapter5} 
\input{./Chapters/Chapter6} 
%\input{./Chapters/Chapter7}

%----------------------------------------------------------------------------------------
%	THESIS CONTENT - APPENDICES
%----------------------------------------------------------------------------------------

\addtocontents{toc}{\vspace{2em}} % Add a gap in the Contents, for aesthetics

\appendix % Cue to tell LaTeX that the following 'chapters' are Appendices

% Include the appendices of the thesis as separate files from the Appendices folder
% Uncomment the lines as you write the Appendices

\input{./Appendices/AppendixA}
%\input{./Appendices/AppendixB}
%\input{./Appendices/AppendixC}

\addtocontents{toc}{\vspace{2em}} % Add a gap in the Contents, for aesthetics

\backmatter

%----------------------------------------------------------------------------------------
%	BIBLIOGRAPHY
%----------------------------------------------------------------------------------------

\label{Bibliography}

\lhead{\emph{Bibliography}} % Change the page header to say "Bibliography"

\bibliographystyle{unsrtnat} % Use the "unsrtnat" BibTeX style for formatting the Bibliography

\bibliography{Bibliography} % The references (bibliography) information are stored in the file named "Bibliography.bib"

\end{document}

### 2013-01-30 Chapters/Chapter2.tex
Working on clustering section

@@
 
+%----------------------------------------------------------------------------------------
+
 \section{Clustering}
+
+This section will give an introduction and overview to clustering of data. 
+When processing time series of data the line of distinction between 
+segmentation and clustering is very fine. This section will introduce 
+clustering for the purpose as used in this project.
+
+Data clustering can be used with two different applications, exploratory or 
+confirmatory \cite{jain1999data}. In the former, the purpose is to discover 
+clusters in a point cloud of data points. Here the cluster is defined as a 
+group of homogeneous data points measured by some coherence measure, often a 
+distance function with the data points being defined as vectors. The produced 
+clusters, which are a representation of the data examined, can be used in the 
+latter application to assign new provided data points to a cluster and thereby 
+classify them.
+
+Depended on the precise setup, clustering can be used as a successive step 
+after or an implementation of temporal segmentation. When temporal 
+segmentation is used to create successive coherent data points, clustering can 
+be used to recognize the same activities in different points of time. The data 
+points provided to the clustering will be some representation of each segment 
+and each segment will be labeled with a activity. An other mechanism could be 
+to extract features from the raw data points and use these to create clusters 
+directly *** [temporal modifications]. Both mechanisms requires the 
+exploratory and confirmatory phases.
+
+The exploratory phase roughly consists of the following five steps 
+\cite{jain1999data}
+\begin{enumerate}
+	\item data point representation,
+	\item definition of coherence,
+	\item grouping data points,
+	\item cluster abstraction,
+	\item output qualification
+\end{enumerate}
+
+The first step is to find a \emph{representation} of the data points and 
+clusters, considering the number, type and scale of the features and the 
+number of desired classes to cluster in. When dealing with high-dimensional 
+data points a \emph{feature selection} can be performed to use only the most 
+discriminating features. When the raw features are not useful enough, 
+\emph{extraction} can be used to create new synthetic features. *** 
+[qualitative versus quantitative/conceptual]
+
+When the data points are represented in a meaningful way, the measure of 
+\emph{coherence} between pairs of points must be defined. Often this is 
+implemented as the Euclidean 
+or Mahalanobis distance when the data points are represented as vectors or 
+other similarity measures for conceptual patterns.
+
+The most critical step, the \emph{grouping},
+
+%-------------------
+
+\subsection{Formal definition}
+
+%-------------------
+
+\subsection{Application is research fields}
+
+%----------------------------------------------------------------------------------------
 

### 2013-01-30 Chapters/Chapter2.tex
Added text on clustering

@@
 distance function with the data points being defined as vectors. The produced 
-clusters, which are a representation of the data examined, can be used in the 
+clusters, which together form a representation of the data examined, can be 
+used in the 
 latter application to assign new provided data points to a cluster and thereby 
@@
 points provided to the clustering will be some representation of each segment 
-and each segment will be labeled with a activity. An other mechanism could be 
+and each segment will be labeled with a activity. Another mechanism could be 
 to extract features from the raw data points and use these to create clusters 
@@
 
-The most critical step, the \emph{grouping},
+In the most critical step, the \emph{grouping}, the data points with the 
+highest coherence are linked together to form a cluster. The result can be a 
+hard partition over the data or a fuzzy membership degree to each cluster for 
+each data point. When applying rules for merging and splitting cluster, a 
+hierarchical partition is constructed. Partitional clustering algorithms 
+assign data points to clusters by optimizing some criterion, e.g. the mean 
+squared distance from each data point to the clusters' centroid.
+
+Especially in case of large data sets, the resulting clusters will be defined 
+by many data points. To form a compacter and simpler representation, 
+\emph{abstraction} can be used. This will simply analysis by humans or 
+automated systems. A common abstraction is to use the clusters' centroid 
+\cite{diday1976clustering} or parameters of Gaussian patterns.
+
+The final step which can be applied to the resulting partition is some 
+\emph{qualification}. The quality of multiple partitions can be compared and 
+the validity of the partition can be determined. A partition is valid if 
+reasonably it could not been constructed by change or as some random process. 
+External references of models can be used to compare, or the internal data 
+points can be examined. *** [better wordings]
 

### 2013-01-30 Chapters/Chapter2.tex
Setup for PCA

@@
 \subsection{Principal Component Analysis}
+When working with simple 2- or 3-dimensional data it is often, for humans, 
+easy to discover patterns in the set. The data can be plotted and the lines or 
+planes among which the data points lie gives an indication of the pattern. 
+With Principal Component Analysis (PCA) this process is also easy for 
+automated systems. With this method the overall form of a set of data points 
+can be represented, the most discriminative dimensions can be found, the 
+dimensionality of a set can be reduced (which yield in data compression) or 
+the result can be used to characterize and differentiate sets of data points.
+
+The workings of the PCA \cite{smith2002tutorial} relies on the concepts of 
+standard deviation, 
+covariance, eigenvectors and eigenvalues of matrices. These concepts will be 
+discussed very briefly. The standard deviation of a set is an measure for the 
+spread around the mean for a single dimension, or feature. The covariance 
+between two features indicates how they are related; when positive the two 
+features will increase together; when negative one will decrease when the 
+other increases and when zero they are unrelated.
+*** [eigenvectors]
+*** [eigenvalues]
+
+***
+[purpose: dimensionality reduction, most discriminative dimensions. Difference 
+in dimensionality.]
 

### 2013-01-31 Chapters/Chapter2.tex
More on pca

@@
 planes among which the data points lie gives an indication of the pattern. 
-With Principal Component Analysis (PCA) this process is also easy for 
+With Principal Component Analysis (PCA) this process is also possible for 
 automated systems. With this method the overall form of a set of data points 
@@
 covariance, eigenvectors and eigenvalues of matrices. These concepts will be 
-discussed very briefly. The standard deviation of a set is an measure for the 
+discussed very briefly \footnote{For a more in-depth discussion we would like 
+to refer the reader to \cite{jolliffe2005principal}}. The standard deviation 
+of a set is an measure for the 
 spread around the mean for a single dimension, or feature. The covariance 
-between two features indicates how they are related; when positive the two 
+between two features (dimensions of the data points) indicates how they are 
+related; when positive the two 
 features will increase together; when negative one will decrease when the 
-other increases and when zero they are unrelated.
-*** [eigenvectors]
-*** [eigenvalues]
+other increases and when zero they are unrelated. The covariance matrix gives 
+all the covariances for all pairs of features. This matrix is symmetrical 
+about the diagonal and the values on the diagonal are the variances for each 
+feature. For this matrix the eigenvectors can be computed. Eigenvectors have 
+the characteristic that when they multiply a matrix, the resulting vector is a 
+multiple of the eigenvector. The amount by which it is the multiple is the 
+Eigenvalue. This is illustrated in formulae \ref{eq:no-eigenvector} and 
+\ref{eq:eigenvector}. The second formula shows an Eigenvector and its 
+associated Eigenvalue, 4. The vector in the first formula is not an 
+Eigenvector.
+
+\begin{equation} \label{eq:no-eigenvector}
+	\begin{pmatrix} 2 & 3 \\ 2 & 1 \end{pmatrix}
+\times
+	\begin{pmatrix} 1 \\ 3 \end{pmatrix}
+=
+	\begin{pmatrix} 2 \cdot 1 + 3 \cdot 3 \\ 2 \cdot 1 + 1 \cdot 3 
+	\end{pmatrix}
+=
+	\begin{pmatrix} 11 \\ 5 \end{pmatrix}
+\end{equation}
+
+\begin{equation} \label{eq:eigenvector}
+	\begin{pmatrix} 2 & 3 \\ 2 & 1 \end{pmatrix}
+\times
+	\begin{pmatrix} 3 \\ 2 \end{pmatrix}
+=
+	\begin{pmatrix} 2 \cdot 3 + 3 \cdot 2 \\ 2 \cdot 3 + 1 \cdot 2  
+	\end{pmatrix}
+=
+	\begin{pmatrix} 12 \\ 8 \end{pmatrix}
+=
+	4 \times \begin{pmatrix} 3 \\ 2 \end{pmatrix}
+\end{equation}
+
+It are these Eigenvectors and Eigenvalues that makes PCA possible. Given a 
+matrix of $n \times n$ (only square matrices have Eigenvectors), $n$ 
+Eigenvectors can be found. Each Eigenvector is perpendicular, or orthogonal, 
+to the others. The Eigenvectors combined describes the lines over which the 
+data is plotted. The Eigenvector with the largest Eigenvalue is the 
+\emph{principal} component of the data set. It is said that it is the most 
+significant dimension. Thus the Eigenvectors, and thereby the features or 
+dimensions of the data set, can be sorted on significance by the Eigenvalues.
+
+With this ordering there are a few applications possible. The first is to just 
+make a (comprehensible) representation of the data. The Eigenvectors describe 
+the cloud of data points and thus a
+
+*** [graph, a bit of formulae with sets and dimensions]
 

### 2013-02-01 Chapters/Chapter2.tex
Finished text on PCA. Figures still needed

@@
 This section will give an introduction and in-depth analysis of temporal 
-segmentation. test. nog een test. test
+segmentation.
 
@@
 
+The following steps are performed, which are explained below and illustrated 
+in figure *** [add figure with plot of sets] \ref{}:
+\begin{enumerate}
+	\item \textbf{Gather data} and represent them in a chosen number of 
+	features,
+	\item \textbf{Subtract mean} to center the data around the axis. The new 
+	data set will have mean zero,
+	\item \textbf{Generate covariance matrix} by calculating all pairwise 
+	feature variances,
+	\item \textbf{Calculate Eigenvectors and Eigenvalues} of the covariance 
+	matrix. The Eigenvectors are then normalized to make further calculations 
+	easier,
+	\item \textbf{Generate feature vector} which indicated which features are 
+	characteristic or meant to keep in de data set, by comparing the 
+	Eigenvalues. 
+\end{enumerate}
+
 The workings of the PCA \cite{smith2002tutorial} relies on the concepts of 
-standard deviation, 
+standard deviation, variance,
 covariance, eigenvectors and eigenvalues of matrices. These concepts will be 
@@
 to refer the reader to \cite{jolliffe2005principal}}. The standard deviation 
-of a set is an measure for the 
+and variance
+of a set are measures for the 
 spread around the mean for a single dimension, or feature. The covariance 
@@
 the characteristic that when they multiply a matrix, the resulting vector is a 
-multiple of the eigenvector. The amount by which it is the multiple is the 
+multiple of the Eigenvector. The amount by which it is the multiple is the 
 Eigenvalue. This is illustrated in formulae \ref{eq:no-eigenvector} and 
-\ref{eq:eigenvector}. The second formula shows an Eigenvector and its 
+\ref{eq:eigenvector}. The second formula shows an Eigenvector 
+$ \left( \begin{smallmatrix} 
+3 \\ 2 \end{smallmatrix} \right)$ and its 
 associated Eigenvalue, 4. The vector in the first formula is not an 
@@
 It are these Eigenvectors and Eigenvalues that makes PCA possible. Given a 
-matrix of $n \times n$ (only square matrices have Eigenvectors), $n$ 
+matrix of size $n \times n$ (only square matrices have Eigenvectors), $n$ 
 Eigenvectors can be found. Each Eigenvector is perpendicular, or orthogonal, 
@@
 \emph{principal} component of the data set. It is said that it is the most 
-significant dimension. Thus the Eigenvectors, and thereby the features or 
+significant feature. Thus the Eigenvectors, and thereby the features or 
 dimensions of the data set, can be sorted on significance by the Eigenvalues.
@@
 make a (comprehensible) representation of the data. The Eigenvectors describe 
-the cloud of data points and thus a
+the cloud of data points and thus are a compressed representation of the data. 
+When the original data points are to be compressed, it is possible to remove 
+the least significant features and reconstruct the data points from the 
+resulting Eigenvectors. This will yield in a lossy compression. *** [figure to 
+illustrate]. An extension of this application is to determine the number of 
+features needed to keep the compression within a certain error criterion, as 
+used in \cite{barbivc2004segmenting}. Sets of data can then be distinguished 
+by the number of features needed to describe the points.
 
 *** [graph, a bit of formulae with sets and dimensions]
-
-***
-[purpose: dimensionality reduction, most discriminative dimensions. Difference 
-in dimensionality.]
 

### 2013-02-01 main.tex
Finished text on PCA. Figures still needed

@@
 
-\usepackage[square, numbers, comma, sort&compress]{natbib} % Use the natbib reference package - read up on this to edit the reference style; if you want text (e.g.
-Smith et al., 2012) for the in-text references (instead of numbers), remove 'numbers'
-\hypersetup{urlcolor=blue, colorlinks=true} % Colors hyperlinks in blue - change to black if annoying
+\usepackage[square, numbers, comma, sort&compress]{natbib} % Use the natbib 
+%reference package - read up on this to edit the reference style; if you want 
+%text (e.g. Smith et al., 2012) for the in-text references (instead of 
+%numbers), remove 'numbers' 
+\hypersetup{urlcolor=blue, colorlinks=true} % Colors hyperlinks in blue - 
+%change to black if annoying
 \title{\ttitle} % Defines the thesis title - don't touch this

### 2013-02-01 Chapters/Chapter2.tex
Fix typs
Detailed method of 'Segmenting Motion Capture Data into Distinct Behaviors'

@@
 non-empty segments $s_1s_2 \dots s_k = s$.
-This notation has close resemblance to the notation as used in 
-\cite{himberg2001time}.
+This notation is adopted from \cite{himberg2001time}.
 
@@
 
+\subsection{PCA Based Methods}
+
 Many fields of research have been active in the unsupervised segmentation of 
@@
 dimensions of the hyperplane is needed to represent the data with equal error 
-degree.
+degree. When the dimensionality is reduced from $d$ to $r \le d$, the ratio of 
+error 
+$E_r$ can be calculated as
 
-[SKIPPING]
-By analyzing the derivative of the error rate, the algorithm is capable of 
-selecting transition points and thus segmenting the data. When a data point 
-$d_i$ is more than $k_\sigma=3$ times the standard deviation from the average 
-over the previous data points $d_j, j<i$, then a cut is assigned to that frame.
+\begin{equation}
+	E_r = \frac{\sum_{j=1}^{r} \sigma_j^2}{\sum_{j=1}^{d} \sigma_j^2}
+\end{equation}
+
+where $\sigma_j$ are the singular values as a result from Singular Value 
+Decomposition (SVD), which is closely related to PCA \cite{shlens2005tutorial}.
+
+In \cite{barbivc2004segmenting} the stream of frames is analyzed on cut-frames 
+to find transitions between action. First, for a number of $k$ frames (e.g. 
+the 
+equivalent of 2.5 seconds) the required dimensionality $r$ to keep the error 
+$E_r$ 
+below some threshold $\tau$ is calculated. This will yield in an error $e_i$ 
+for the first 
+$i$ frames. When more frames are added to the window, the error will increase 
+with a low constant when it is still in the same activity, due to noise in the 
+activity. When a new activity 
+starts, e.g. at frame $j$, the error at frame $j$ will start increasing 
+faster. This can be expressed by the derivative of the error rate $d_i = e_i - 
+e_{i-l}$, where $l$ is a constant to remove noise. From this derivative the 
+mean and standard deviation can be calculated, for each point. When a 
+derivative $d_j$ rises more than a factor $k_\sigma = 3$ standard deviations 
+from the mean, a transition point is encountered. The previous frames are then 
+cut from the sequence (as a segment) and the algorithm starts over.
+
+*** [figure to illustrate derivative and standard deviation]
 
@@
 processed backward over the data series.
+
+*** [figure to illustrate mahalanobis distance measure and peaks]
 

### 2013-02-01 Chapters/Chapter2.tex
Explaining more

@@
 
+\subsection{Statistical methods}
+An method to process data, or create a basis for further processing, is by 
+analyzing the statistical properties of a data set. When there is a continuous 
+stream of data, it is possible to extract events by comparing the statistical 
+properties of sliding time windows. By definition, events have a different 
+statistical profile then their background. When two consecutive actions are 
+regarded as the background, then a transition is an event between them.
+
+\subsection{Hidden Markov Models Based Methods}
+
+\subsection{Bayesian Methods}
+
 \subsection{Principal Component Analysis}
@@
 *** [graph, a bit of formulae with sets and dimensions]
+
+\subsection{Hidden Markov Models}
+Viterbi algorithm / path; find the most likely sequence of cause-states for 
+the observed measures.
 

### 2013-02-04 Chapters/Chapter1.tex
Added gitignore file

@@
 
+*** Most of the activity recognition techniques can be classified in two 
+different types. Some rely on a state-space model, in which the activities to 
+be recognized are represented in a statistical manner. Bayesian networks, 
+Finite State Machines and Hidden Markov Models can be considered among those. 
+Other techniques process the directly as a pattern recognition task, such as 
+Support Vector Machines, Neural Networks, Dynamic Time Warping, and Bayes and 
+K-means clustering.
+
 %----------------------------------------------------------------------------------------

### 2013-02-04 Chapters/Chapter2.tex
Added gitignore file

@@
 
+%-------------------
+
 \subsection{Hidden Markov Models}
+Many dynamic systems can be viewed as some sequence of consecutive states 
+which are observable by the world. These kind of systems are known as Markov 
+Chains. When the state in which the system is can not be observed with 
+complete certainty, a system known as a Hidden Markov Model (HMM) can be 
+constructed. It has been shown that HMMs are effective on in the task of 
+speech recognition \cite{rabiner1989tutorial}. Because of some resemblance 
+between natural speech construction and human activities, modifications have 
+been able to classify human activities \cite{guenterberg2009distributed} and 
+extensions are implemented which create layers of HMMs to handle different 
+granularity over time and reduce retraining 
+\cite{oliver2002layered,perdikis2008recognition}. A HMM is a state-space model 
+representing a single model for each label which needs to be classified.
+
+Formally, a Hidden Markov Model is characterized by the following elements 
+\cite{rabiner1989tutorial}:
+\begin{enumerate}
+	\item A number of $N$ states in the model. These states are a 
+	representation of the underlying mechanism of the system and can not be 
+	directly observed from outside (hence the \emph{hidden} model). Individual 
+	states are referenced by $S = \{S_1, S_2, \dots, S_N \}$ and a single 
+	state in time $t$ is denoted by $q_t$.
+	\item An alphabet of $M$ discrete observable symbols for each state. An 
+	observed symbol is the only indication of the systems current situation. 
+	Each individual symbol is referenced by $V = \{ v_1, v_2, \dots, v_m \}$.
+	\item A state transition probability distribution $A = \{ a_{ij} \}$ 
+	indication the probability of being in state $j$ at time $t+1$ when the 
+	state on time $t$ was $i$:
+	\begin{eqnarray}
+		a_{ij} = P [ q_{t+1} = S_j | q_t = S_i ], & 1 \le i, j \le N
+	\end{eqnarray}
+	Note that by setting the probability $a_{ij}=0$ for some $i$ and $j$ 
+	indicates that $j$ is not reachable from state $i$.
+	\item A symbol observation probability distribution in state $j$, $B = \{ 
+	b_j(k) \}$, where
+	\begin{eqnarray}
+	b_j(k) = P[ v_k\ \mathrm{at} \ t | q_t = S_j], & 1 \le j \le N \nonumber \\
+	& 1 \le k \le M
+	\end{eqnarray}
+	Note that every symbol can be observed in every state.
+	\item The initial state distribution in which the system starts $\pi= \{ 
+	\pi_i \}$ where
+	\begin{eqnarray}
+	\pi_i = P[q_1 = S_i], & 1 \le i \le N
+	\end{eqnarray}
+\end{enumerate}
+This system is often abbreviated to
+\begin{equation}
+	\lambda = \{ \pi_i, a_{ij}, b_j(k)\}
+\end{equation}
+The HMM will act as a generator giving a sequence of $T$ observations, each 
+being a symbol from $V$:
+\begin{equation}
+O = O_1, O_2 \dots O_T
+\end{equation}
+
+Given the characteristics of a specific HMM, there are three major application 
+which serve the classification of an activity. The first application is, given 
+a sequence of observations $O = O_1,O_2,\dots,O_T$, to label this sequence 
+with the activity with 
+the highest associated probability for that label, e.g. compute $P(O|\lambda)$ 
+for each model $\lambda$. The second task a HMM can perform is to 
+\emph{explain} a observation sequence $O$, by choosing an optimal 
+corresponding state sequence $Q = q_1,q_2,\dots,q_T$. The last application is 
+more of a problem statement, being the task to train the HMM to adjust the 
+parameters $\lambda = \{ \pi_i, a_{ij}, b_j(k)\}$ to maximize $P(O|\lambda)$. 
+In the overall task of activity recognition, one would first use the last 
+application to train a set of HMMs with the provided and labeled training 
+data. The second application can be used to further optimize the HMM (set the 
+number of states, adjust vector codebook etc.) and improve the capability of 
+modeling the activities. Finally, when the system is provided with unlabeled 
+data the first application is used to identify the observed sequence with 
+trained examples, by comparing the probability result of multiple HMMs.
+
 Viterbi algorithm / path; find the most likely sequence of cause-states for 
 the observed measures.
+
+%-------------------
 

### 2013-02-04 Chapters/Chapter2.tex
Worked on HMM

@@
 
+Generally a HMM can be fully connected, or ergodic model. Other configurations 
+are also possible, of which the left-right model, or Bakis model, is popular 
+for temporal applications \cite{rabiner1989tutorial}. This because the 
+structure of the state probability distribution resembles the construction of 
+possible states over time. This left-right model can be generalized by setting
+\begin{eqnarray}
+a_{ij} = 0, & j > i + \Delta
+\end{eqnarray}
+for some $\Delta$, indicating the possible \emph{distance} of the jumps over 
+states.
+
+*** [INSERT GRAPHICS SHOWING STRUCTURE OF HMM]
+
+In \cite{oliver2002layered} it is stated that HMMs are robust mechanisms with 
+respect to changes in temporal segmentation of observations, although they 
+suffer from a few problems when applied to reason about longer and complexer 
+sequences over time. They state that, especially with limited training data, 
+the HMM tend to be overfitted and tend to lack structure and have an excess of 
+parameters. The overcome this limitation, they introduce a Layered Hidden 
+Markov Model (LHHM) which they show is more robust to temporal changes in 
+observations. They make use of the hypothesis from the field of psychology, 
+that human behavior is hierarchically structured \cite{zacks2001event}.
+
 Viterbi algorithm / path; find the most likely sequence of cause-states for 

### 2013-02-05 Chapters/Chapter2.tex
More on HMM

@@
 \section{Signal pro-processing and sensor fusion for timed patterns}
+
+*** [fusion of accelerometer data, gyroscopes, magnetic field sensors, 
+pressure, microphones, etc]
 
@@
 regarded as the background, then a transition is an event between them.
+
+In \cite{guenterberg2009automatic} a automatic system is used to segment a 
+continuous stream of data. By calculating features as activity level from the 
+standard deviation of a sliding window and applying Adaptive Thresholds, the 
+data can be segmented in active and rest parts. There is no model 
+representation.
 
@@
 
+*** [discrete and continuous HHMs]
+
+*** Applications of HHM: \cite{shi2009towards} uses 10 coefficients extracted 
+with FFT from 6 axis (3 times accelerometer and 3 times gyroscope). Example of 
+continuous HMM (?). Recognition rates fo 50\%-100\%.
+
 In \cite{oliver2002layered} it is stated that HMMs are robust mechanisms with 
@@
 observations. They make use of the hypothesis from the field of psychology, 
-that human behavior is hierarchically structured \cite{zacks2001event}.
+that human behavior is hierarchically structured \cite{zacks2001event}. In 
+\cite{perdikis2008recognition} a layered construction of HMMs is used which 
+feeds the low-level data from a sliding window to the lower layer. This layer 
+will recognize simple motions, such as moving the arm in a pre-defined 
+segment. The results from these lower HMMs are then fed to a higher-level 
+layer of HMMs which are able to recognize abstract motions, such as adjusting 
+a monitor and picking up a pen. The results showed that the layered system was 
+better at recognition then a similar single-layered setup. It was also robust 
+in the sense that when the environment was changed, only the relatively simple 
+low-level HMMs needed to be retrained.
 
-Viterbi algorithm / path; find the most likely sequence of cause-states for 
+*** [DIAGRAM WITH EXAMPLE OF LOW- AND HIGH-LEVEL HMMs, SUCH AS MOVING ARM, 
+PICKING UP PEN]
+
+*** Viterbi algorithm / path; find the most likely sequence of cause-states 
+for 
 the observed measures.

### 2013-02-07 Chapters/Chapter2.tex
Added intended thesis outline

@@
 % Chapter 2
+% !TEX root = ../main.tex
 
@@
 The following steps are performed, which are explained below and illustrated
-in figure *** [add figure with plot of sets] \ref{}:
+in figure *** [add figure with plot of sets]:
+
 \begin{enumerate}
@@
 
-\subsection{$k$-means clustering}
+\subsection{k-means clustering}
 [EXPLAIN] divide data set $n$ into $k$ clusters.

### 2013-02-07 intended_contents.tex
Added intended thesis outline

\documentclass[a4paper,10pt]{extarticle}
\usepackage[english]{babel}
\usepackage{fullpage}
\usepackage{hyperref}
\author{R.Q. Vlasveld}
\title{Intended Table of Contents Thesis}

\begin{document}
\maketitle
\tableofcontents

\section{Introduction}
This section introduces the subject, outlines the global goal of the thesis
and explains the problem. The main research-questions are introduced. In the
subsections the main components (global, not as an implementation) are
introduced.

  \subsection{Thesis goal}
  The goal of the thesis will be discussed, e.g. "precise unsupervised on-line human activity recognition". Mention application to show why it is useful.

  \subsection{Results of earlier research}
  First, results of earlier project are outlined. Look at the data being used, the precision and error rates (shortly, no details) and look at the different concepts being discussed. This will give reason for the following subsections.

\section{Concepts}

  \subsection{Temporal pattern recognition}
  This will explain about what patterns are and what kind of patterns need to be recognized (long-term, short-term repetitive, etc)

	\subsection{Signal pre-processing}
  Describe how data is collected and processed. Include feature extraction, like PCA, FFT. Mention only shortly with meaning, no further explanation.

	\subsection{Temporal Segmentation}
  What is segmentation, the purpose and what kind of methods are there. E.g. model-based, statistics-based. Diagram with types. Supervised / unsupervised. HMMs, Statistical, Bayesian. K-means, SOM

  \subsection{Temporal clustering}
  When is segmenting clustering? Explain difference, mention techniques (diagrams), Dynamic Time Warping, k-means (and variations), SOM, SVM, Naive Bayes, Neural Networks.

\section{Problem statement}
This section proposes a setup to improve (or: answer) the research-questions. It will discuss techniques used in other projects and how implemented here. In the subsections each used technique is further explained. (Only explained when really used)

  \subsection{Proposed method}
  Give an overview of the "pipeline" of the implementation. Give a diagram how data is retrieved, (pre-)processed, analyzed and conclusions over data are drawn.

  \subsection{PCA}
  Something about the workings of PCA and why it is relevant and used. Discuss how it is used in other projects and how it will be used here.

  \subsection{Hidden Markov Models}
  Something about the workings of HMMs and why it is relevant and used.

  \subsection{Dynamic Time Warping}
  if used

  \subsection{k-Means clustering}
  if used

  \subsection{Self-organizing maps}
  if used

\section{Implementation}
This section will discuss implementation details. Hardware platform (if any), software used. Complexity, precision of metrics etc.

\section{Experiments, results}
This section will outline the precise experiments and gives the results of these.

\section{Discussion}
This section will discuss the results of the previous section. It will relate the results to previous work.

\section{Conclusion}
Conclusion (answer research question), summary of results, feature work.

\end{document}

### 2013-02-08 Chapters/Chapter2.tex
K-means graph
Removed linebreaks

@@
 
-\label{Chapter2} % For referencing the chapter elsewhere, use \ref{Chapter1}
+\label{Chapter2} % For referencing the chapter elsewhere, use~\ref{Chapter1}
 
@@
 
-*** [fusion of accelerometer data, gyroscopes, magnetic field sensors,
-pressure, microphones, etc]
+*** [fusion of accelerometer data, gyroscopes, magnetic field sensors, pressure, microphones, etc]
 
 \section{Temporal Segmentation}
-
-This section will give an introduction and in-depth analysis of temporal
-segmentation.
+This section will give an introduction and in-depth analysis of temporal segmentation.
 
 \subsection{Aims of segmentation}
-When processing and analyzing time series of data, e.g. motion measurements,
-stock market fluctuations or natural language, first a low-level division
-between the discriminative parts of the stream must be made. One can view this
-as splitting the series into the \emph{atoms}, which are the building blocks
-of the total stream. These building blocks will be the aggregation of
-non-overlapping, internally homogeneous segments \cite{himberg2001time}. This
-means that the data points inside a segment should have some resemblance
-relation to each other and their difference lies between some boundary. The
-process of segmenting can be viewed as a subproblem to context analysis of
-time series. Temporal segmentation is closely related to temporal clustering,
-although it is a stricter, and simpler, process. Whereby clustering only
-restricts the data points on their distance relation (as used in a Voronoi
-diagram), within a segment the data points must also be contiguous.
+When processing and analyzing time series of data, e.g.
+motion measurements, stock market fluctuations or natural language, first a low-level division between the discriminative parts of the stream must be made.
+One can view this as splitting the series into the \emph{atoms}, which are the building blocks of the total stream.
+These building blocks will be the aggregation of non-overlapping, internally homogeneous segments~\cite{himberg2001time}.
+This means that the data points inside a segment should have some resemblance relation to each other and their difference lies between some boundary.
+The process of segmenting can be viewed as a subproblem to context analysis of time series.
+Temporal segmentation is closely related to temporal clustering, although it is a stricter, and simpler, process.
+Whereby clustering only restricts the data points on their distance relation (as used in a Voronoi diagram), within a segment the data points must also be contiguous.
 
-The task of segmentation can be performed in a manual
-matter, by cutting and labeling parts of the stream into coherent parts. This
-would require human (expert) knowledge and does not yield a clear cut because
-of ambiguity. With increasing storage abilities and easier motion capture
-systems, there is a desire for automated systems which perform the
-segmentation task unsupervised. Some algorithms used have the (often desired)
-side-effect of also clustering the segments, such that classes of segments can
-be discovered in the time series. These algorithms would not only be able to
-make a distinction between walking, sitting and walking, but would also
-recognize the reappearance of the walking activity.
+The task of segmentation can be performed in a manual matter, by cutting and labeling parts of the stream into coherent parts.
+This would require human (expert) knowledge and does not yield a clear cut because of ambiguity.
+With increasing storage abilities and easier motion capture systems, there is a desire for automated systems which perform the segmentation task unsupervised.
+Some algorithms used have the (often desired) side-effect of also clustering the segments, such that classes of segments can be discovered in the time series.
+These algorithms would not only be able to make a distinction between walking, sitting and walking, but would also recognize the reappearance of the walking activity.
 
 \subsection{Formal definition}
-Formally, temporal segmentation is dividing a time series $s$, which consists
-of $N$ samples $\mathbf{x}(1),\mathbf{x}(2),\dots,\mathbf{x}(N)$ from
-$\mathbf{R}^d$. Individual \emph{segments} are referenced by $s(a,b)$,
-consisting of the consecutive samples
-$\mathbf{x}(a),\mathbf{x}(a+1),\dots,\mathbf{x}(b)$, $a \le b$. Let $s_1 =
-s(a,b)$ and $s_2 = s(b+1,c)$ be two segments, then their concatenation is
-$s_1s_2 = s(a,c)$. A segmentation $S$ of $s$ consists of a sequence of $k$
-non-empty segments $s_1s_2 \dots s_k = s$.
-This notation is adopted from \cite{himberg2001time}.
+Formally, temporal segmentation is dividing a time series $s$, which consists of $N$ samples $\mathbf{x}(1),\mathbf{x}(2),\dots,\mathbf{x}(N)$ from $\mathbf{R}^d$.
+Individual \emph{segments} are referenced by $s(a,b)$, consisting of the consecutive samples $\mathbf{x}(a),\mathbf{x}(a+1),\dots,\mathbf{x}(b)$, $a \le b$.
+Let $s_1 = s(a,b)$ and $s_2 = s(b+1,c)$ be two segments, then their concatenation is $s_1s_2 = s(a,c)$.
+A segmentation $S$ of $s$ consists of a sequence of $k$ non-empty segments $s_1s_2 \dots s_k = s$.
+This notation is adopted from~\cite{himberg2001time}.
 
-As stated, informally each segment should be internally homogeneous. This can
-formally be measured with an cost function $F$, indication the heterogeneity
-of a segment. The overall aim is to minimize the cost $F$. The cost of a
-segment is a function from the data points and the number of data points $n =
-b - a + 1$ and is expressed as
-\begin{equation} \label{eq:segment_cost}
+As stated, informally each segment should be internally homogeneous.
+This can formally be measured with an cost function $F$, indication the heterogeneity of a segment.
+The overall aim is to minimize the cost $F$. The cost of a segment is a function from the data points and the number of data points
+$n = b - a + 1$ and is expressed as
+\begin{equation}
+	\label{eq:segment_cost}
 	\mathrm{cost}_F (s(a,b)) = F(\mathbf{x};n|\mathbf{x} \in s(a,b))
 \end{equation}
-The cost of a \emph{k-segmentation} $S$ is the summation of the costs of the
-$k$ segments:
-\begin{equation} \label{eq:segmentation_cost}
+The cost of a \emph{k-segmentation} $S$ is the summation of the costs of the $k$ segments:
+\begin{equation}
+	\label{eq:segmentation_cost}
 	\mathrm{Cost}_F (s_1 s_2 \dots s_k) = \sum_{i=1}^{k} \mathrm{cost}_F (s_k)
 \end{equation}
-With the objective of minimizing the cost function, the optimal
-$k$-segmentation $S_F^\mathit{opt}(s;k)$ is the segmentation with minimal
-$\mathrm{Cost}_F(s_1 s_2 \dots s_k)$ over all possible $k$-segmentations.
+With the objective of minimizing the cost function, the optimal $k$-segmentation $S_F^\mathit{opt}(s;k)$ is the segmentation with minimal $\mathrm{Cost}_F(s_1 s_2 \dots s_k)$ over all possible $k$-segmentations.
 
-The cost function, to calculate the heterogeneity of a (set of) segment(s),
-can
-be any function. A simple and natural function would be the sum of variances
-of the segments. The overall cost function would then be
-\begin{equation} \label{eq:cost_variances}
+The cost function, to calculate the heterogeneity of a (set of) segment(s), can be any function.
+A simple and natural function would be the sum of variances of the segments.
+The overall cost function would then be
+\begin{equation}
+	\label{eq:cost_variances}
 	\mathrm{Cost}_V = \frac{1}{N} \sum_{i=i}^{k} \sum_{j=c_{i-1}+1}^{c_i} \|
@@
 
-To analyze time series it is often preferred to divide the stream in segments
-of correlated data. After dividing, each segment represent a period in time in
-which the same activity is performed. Or, stated otherwise, it results in
-transitions moments between activities.
+To analyze time series it is often preferred to divide the stream in segments of correlated data.
+After dividing, each segment represent a period in time in which the same activity is performed.
+Or, stated otherwise, it results in transitions moments between activities.
 
@@
 
-Many fields of research have been active in the unsupervised segmentation of
-data. Many authors rely on a form of Principal Component Analysis (PCA), as
-used a.o. in \cite{barbivc2004segmenting}. Often PCA is used to reduce the
-dimensionality of the data being processed [REFERENCE] by only using the top
-$r$ dimensions to describe the data set. It is observed that data series of
-simple motions have a lower dimensionality then complexer motions. When a
-simple (repetitive) motion is about to end and fluently transforms in a new
-motion, there will be a window of time in which a high dimensionality will be
-present, due to the new motion. After this period of transition, the
-dimensionality will decrease, since only the new simple motion is present in
-the window of time. The first algorithm of \cite{barbivc2004segmenting} is
-based on this principle.
+Many fields of research have been active in the unsupervised segmentation of data.
+Many authors rely on a form of Principal Component Analysis (PCA), as used a.o. in~\cite{barbivc2004segmenting}.
+Often PCA is used to reduce the dimensionality of the data being processed [REFERENCE] by only using the top $r$ dimensions to describe the data set.
+It is observed that data series of simple motions have a lower dimensionality then complexer motions.
+When a simple (repetitive) motion is about to end and fluently transforms in a new motion, there will be a window of time in which a high dimensionality will be present, due to the new motion.
+After this period of transition, the dimensionality will decrease, since only the new simple motion is present in the window of time.
+The first algorithm of~\cite{barbivc2004segmenting} is based on this principle.
 
-Given a set of data points, a lower dimensional hyperplane can by constructed
-to which the data points can be projected. This projection on a lower
-dimension introduces a error to the original position. When the error is
-fixed, less dimensions are needed for simple motions in which movements of
-body parts are highly correlated. For segments in which the data points are
-lesser correlated, e.g. because of transition state, a higher degree of
-dimensions of the hyperplane is needed to represent the data with equal error
-degree. When the dimensionality is reduced from $d$ to $r \le d$, the ratio of
-error
-$E_r$ can be calculated as
-
+Given a set of data points, a lower dimensional hyperplane can by constructed to which the data points can be projected.
+This projection on a lower dimension introduces a error to the original position.
+When the error is fixed, less dimensions are needed for simple motions in which movements of body parts are highly correlated.
+For segments in which the data points are lesser correlated, e.g. because of transition state, a higher degree of dimensions of the hyperplane is needed to represent the data with equal error degree.
+When the dimensionality is reduced from $d$ to $r \le d$, the ratio of error $E_r$ can be calculated as
 \begin{equation}
@@
 
-where $\sigma_j$ are the singular values as a result from Singular Value
-Decomposition (SVD), which is closely related to PCA \cite{shlens2005tutorial}.
+where $\sigma_j$ are the singular values as a result from Singular Value Decomposition (SVD), which is closely related to PCA~\cite{shlens2005tutorial}.
 
-In \cite{barbivc2004segmenting} the stream of frames is analyzed on cut-frames
-to find transitions between action. First, for a number of $k$ frames (e.g.
-the
-equivalent of 2.5 seconds) the required dimensionality $r$ to keep the error
-$E_r$
-below some threshold $\tau$ is calculated. This will yield in an error $e_i$
-for the first
-$i$ frames. When more frames are added to the window, the error will increase
-with a low constant when it is still in the same activity, due to noise in the
-activity. When a new activity
-starts, e.g. at frame $j$, the error at frame $j$ will start increasing
-faster. This can be expressed by the derivative of the error rate $d_i = e_i -
-e_{i-l}$, where $l$ is a constant to remove noise. From this derivative the
-mean and standard deviation can be calculated, for each point. When a
-derivative $d_j$ rises more than a factor $k_\sigma = 3$ standard deviations
-from the mean, a transition point is encountered. The previous frames are then
-cut from the sequence (as a segment) and the algorithm starts over.
+In~\cite{barbivc2004segmenting} the stream of frames is analyzed on cut-frames to find transitions between action.
+First, for a number of $k$ frames (e.g. the equivalent of 2.5 seconds) the required dimensionality $r$ to keep the error $E_r$ below some threshold $\tau$ is calculated.
+This will yield in an error $e_i$ for the first $i$ frames.
+When more frames are added to the window, the error will increase with a low constant when it is still in the same activity, due to noise in the activity.
+When a new activity starts, e.g. at frame $j$, the error at frame $j$ will start increasing faster.
+This can be expressed by the derivative of the error rate $d_i = e_i - e_{i-l}$, where $l$ is a constant to remove noise.
+From this derivative the mean and standard deviation can be calculated, for each point.
+When a derivative $d_j$ rises more than a factor $k_\sigma = 3$ standard deviations from the mean, a transition point is encountered.
+The previous frames are then cut from the sequence (as a segment) and the algorithm starts over.
 
@@
 
-A second approach in \cite{barbivc2004segmenting} uses the probabilistic
-variant of PCA (PPCA) to model the data set as a Gaussian distribution instead
-of ignoring the frames which do not fit in the subspace. Over windows of
-frames the mean and variance are calculated. In a forward manner the
-Mahalanobis distance of a new window of frames is calculated, which represents
-the likelihood of the new window belonging to the same segment as the original
-widow. When the distance decreases, the likelihood increases which happens
-when the motions in the becomes more homogeneous. When a peak in the distance
-is reached, the new window of frames indicates a heterogeneous collection of
-motions in the window and thus a low likelihood of membership and a indication
-of a transition. In order to distinct activities and subactivities (which
-require a subset of motions is a distinct activity) the algorithm is also
-processed backward over the data series.
+A second approach in~\cite{barbivc2004segmenting} uses the probabilistic variant of PCA (PPCA) to model the data set as a Gaussian distribution instead of ignoring the frames which do not fit in the subspace.
+Over windows of frames the mean and variance are calculated.
+In a forward manner the Mahalanobis distance of a new window of frames is calculated, which represents the likelihood of the new window belonging to the same segment as the original widow.
+When the distance decreases, the likelihood increases which happens when the motions in the becomes more homogeneous.
+When a peak in the distance is reached, the new window of frames indicates a heterogeneous collection of motions in the window and thus a low likelihood of membership and a indication of a transition.
+In order to distinct activities and subactivities (which require a subset of motions is a distinct activity) the algorithm is also processed backward over the data series.
 
@@
 
-The third algorithm in \cite{barbivc2004segmenting} is based on the
-observation that data points (frames) tend to form clusters in the space.
-These clusters are represented by $k$ Gaussian distributions for which each
-the Expectation-Maximization (EM) algorithm estimates the mean $m_j$,
-covariance matrix $\sum_{j}$ and prior $\pi_j$. With all the Gaussian
-distributions estimated, the data points are assigned to the cluster with the
-highest membership likelihood. When two consecutive frames $x_i$ and $x_{i+1}$
-belong to different clusters, a transition of activities is recognized. Note
-that this algorithm succeeds in segmenting the data and also labels the
-similar simple activities.
+The third algorithm in~\cite{barbivc2004segmenting} is based on the observation that data points (frames) tend to form clusters in the space.
+These clusters are represented by $k$ Gaussian distributions for which each the Expectation-Maximization (EM) algorithm estimates the mean $m_j$, covariance matrix $\sum_{j}$ and prior $\pi_j$.
+With all the Gaussian distributions estimated, the data points are assigned to the cluster with the highest membership likelihood.
+When two consecutive frames $x_i$ and $x_{i+1}$ belong to different clusters, a transition of activities is recognized.
+Note that this algorithm succeeds in segmenting the data and also labels the similar simple activities.
 
-A drawback in this system, and many others which implement a variant of the
-$k$-means algorithm, is that the number of clusters $k$ need to be
-predetermined. To cope with this, often the algorithm is performed multiple
-times for different values of $k$. Using some criteria, e.g. the Bayesian
-Information Criterion \cite{pelleg2000x} or the Davies-Bouldin Index which
-guides $k$-means clustering as used in \cite{krause2003unsupervised}.
+A drawback in this system, and many others which implement a variant of the $k$-means algorithm, is that the number of clusters $k$ need to be predetermined.
+To cope with this, often the algorithm is performed multiple times for different values of $k$.
+Using some criteria, e.g. the Bayesian Information Criterion~\cite{pelleg2000x} or the Davies-Bouldin Index which guides $k$-means clustering as used in~\cite{krause2003unsupervised}.
 
 \subsection{Statistical methods}
-An method to process data, or create a basis for further processing, is by
-analyzing the statistical properties of a data set. When there is a continuous
-stream of data, it is possible to extract events by comparing the statistical
-properties of sliding time windows. By definition, events have a different
-statistical profile then their background. When two consecutive actions are
-regarded as the background, then a transition is an event between them.
+An method to process data, or create a basis for further processing, is by analyzing the statistical properties of a data set.
+When there is a continuous stream of data, it is possible to extract events by comparing the statistical properties of sliding time windows.
+By definition, events have a different statistical profile then their background.
+When two consecutive actions are regarded as the background, then a transition is an event between them.
 
-In \cite{guenterberg2009automatic} a automatic system is used to segment a
-continuous stream of data. By calculating features as activity level from the
-standard deviation of a sliding window and applying Adaptive Thresholds, the
-data can be segmented in active and rest parts. There is no model
-representation.
+In~\cite{guenterberg2009automatic} a automatic system is used to segment a continuous stream of data.
+By calculating features as activity level from the standard deviation of a sliding window and applying Adaptive Thresholds, the data can be segmented in active and rest parts.
+There is no model representation.
 
@@
 \subsection{Principal Component Analysis}
-When working with simple 2- or 3-dimensional data it is often, for humans,
-easy to discover patterns in the set. The data can be plotted and the lines or
-planes among which the data points lie gives an indication of the pattern.
-With Principal Component Analysis (PCA) this process is also possible for
-automated systems. With this method the overall form of a set of data points
-can be represented, the most discriminative dimensions can be found, the
-dimensionality of a set can be reduced (which yield in data compression) or
-the result can be used to characterize and differentiate sets of data points.
+When working with simple 2- or 3-dimensional data it is often, for humans, easy to discover patterns in the set.
+The data can be plotted and the lines or planes among which the data points lie gives an indication of the pattern.
+With Principal Component Analysis (PCA) this process is also possible for automated systems.
+With this method the overall form of a set of data points can be represented, the most discriminative dimensions can be found, the dimensionality of a set can be reduced (which yield in data compression) or the result can be used to characterize and differentiate sets of data points.
 
-The following steps are performed, which are explained below and illustrated
-in figure *** [add figure with plot of sets]:
+The following steps are performed, which are explained below and illustrated in figure *** [add figure with plot of sets]:
 
 \begin{enumerate}
-	\item \textbf{Gather data} and represent them in a chosen number of
-	features,
-	\item \textbf{Subtract mean} to center the data around the axis. The new
-	data set will have mean zero,
-	\item \textbf{Generate covariance matrix} by calculating all pairwise
-	feature variances,
-	\item \textbf{Calculate Eigenvectors and Eigenvalues} of the covariance
-	matrix. The Eigenvectors are then normalized to make further calculations
-	easier,
-	\item \textbf{Generate feature vector} which indicated which features are
-	characteristic or meant to keep in de data set, by comparing the
-	Eigenvalues.
+	\item \textbf{Gather data} and represent them in a chosen number of features,
+	\item \textbf{Subtract mean} to center the data around the axis. The new data set will have mean zero,
+	\item \textbf{Generate covariance matrix} by calculating all pairwise feature variances,
+	\item \textbf{Calculate Eigenvectors and Eigenvalues} of the covariance matrix.
+	The Eigenvectors are then normalized to make further calculations easier,
+	\item \textbf{Generate feature vector} which indicated which features are characteristic or meant to keep in de data set, by comparing the Eigenvalues.
 \end{enumerate}
 
-The workings of the PCA \cite{smith2002tutorial} relies on the concepts of
-standard deviation, variance,
-covariance, eigenvectors and eigenvalues of matrices. These concepts will be
-discussed very briefly \footnote{For a more in-depth discussion we would like
-to refer the reader to \cite{jolliffe2005principal}}. The standard deviation
-and variance
-of a set are measures for the
-spread around the mean for a single dimension, or feature. The covariance
-between two features (dimensions of the data points) indicates how they are
-related; when positive the two
-features will increase together; when negative one will decrease when the
-other increases and when zero they are unrelated. The covariance matrix gives
-all the covariances for all pairs of features. This matrix is symmetrical
-about the diagonal and the values on the diagonal are the variances for each
-feature. For this matrix the eigenvectors can be computed. Eigenvectors have
-the characteristic that when they multiply a matrix, the resulting vector is a
-multiple of the Eigenvector. The amount by which it is the multiple is the
-Eigenvalue. This is illustrated in formulae \ref{eq:no-eigenvector} and
-\ref{eq:eigenvector}. The second formula shows an Eigenvector
-$ \left( \begin{smallmatrix}
-3 \\ 2 \end{smallmatrix} \right)$ and its
-associated Eigenvalue, 4. The vector in the first formula is not an
-Eigenvector.
+The workings of the PCA~\cite{smith2002tutorial} relies on the concepts of standard deviation, variance, covariance, eigenvectors and eigenvalues of matrices.
+These concepts will be discussed very briefly \footnote{For a more in-depth discussion we would like to refer the reader to~\cite{jolliffe2005principal}}.
+The standard deviation and variance of a set are measures for the spread around the mean for a single dimension, or feature.
+The covariance between two features (dimensions of the data points) indicates how they are related; when positive the two features will increase together; when negative one will decrease when the other increases and when zero they are unrelated.
+The covariance matrix gives all the covariances for all pairs of features.
+This matrix is symmetrical about the diagonal and the values on the diagonal are the variances for each feature.
+For this matrix the eigenvectors can be computed. Eigenvectors have the characteristic that when they multiply a matrix, the resulting vector is a multiple of the Eigenvector.
+The amount by which it is the multiple is the Eigenvalue.
+This is illustrated in formulae~\ref{eq:no-eigenvector} and \ref{eq:eigenvector}.
+The second formula shows an Eigenvector $ \left( \begin{smallmatrix} 3 \\ 2 \end{smallmatrix} \right)$ and its associated Eigenvalue, 4.
+The vector in the first formula is not an Eigenvector.
 
-\begin{equation} \label{eq:no-eigenvector}
+\begin{equation}
+	\label{eq:no-eigenvector}
 	\begin{pmatrix} 2 & 3 \\ 2 & 1 \end{pmatrix}
@@
 
-\begin{equation} \label{eq:eigenvector}
+\begin{equation}
+	\label{eq:eigenvector}
 	\begin{pmatrix} 2 & 3 \\ 2 & 1 \end{pmatrix}
@@
 
-It are these Eigenvectors and Eigenvalues that makes PCA possible. Given a
-matrix of size $n \times n$ (only square matrices have Eigenvectors), $n$
-Eigenvectors can be found. Each Eigenvector is perpendicular, or orthogonal,
-to the others. The Eigenvectors combined describes the lines over which the
-data is plotted. The Eigenvector with the largest Eigenvalue is the
-\emph{principal} component of the data set. It is said that it is the most
-significant feature. Thus the Eigenvectors, and thereby the features or
-dimensions of the data set, can be sorted on significance by the Eigenvalues.
+ It are these Eigenvectors and Eigenvalues that makes PCA possible.
+ Given a matrix of size $n \times n$ (only square matrices have Eigenvectors), $n$ Eigenvectors can be found.
+ Each Eigenvector is perpendicular, or orthogonal, to the others.
+ The Eigenvectors combined describes the lines over which the data is plotted.
+ The Eigenvector with the largest Eigenvalue is the \emph{principal} component of the data set.
+ It is said that it is the most significant feature.
+ Thus the Eigenvectors, and thereby the features or dimensions of the data set, can be sorted on significance by the Eigenvalues.
 
-With this ordering there are a few applications possible. The first is to just
-make a (comprehensible) representation of the data. The Eigenvectors describe
-the cloud of data points and thus are a compressed representation of the data.
-When the original data points are to be compressed, it is possible to remove
-the least significant features and reconstruct the data points from the
-resulting Eigenvectors. This will yield in a lossy compression. *** [figure to
-illustrate]. An extension of this application is to determine the number of
-features needed to keep the compression within a certain error criterion, as
-used in \cite{barbivc2004segmenting}. Sets of data can then be distinguished
-by the number of features needed to describe the points.
+With this ordering there are a few applications possible.
+The first is to just make a (comprehensible) representation of the data.
+The Eigenvectors describe the cloud of data points and thus are a compressed representation of the data.
+When the original data points are to be compressed, it is possible to remove the least significant features and reconstruct the data points from the resulting Eigenvectors.
+This will yield in a lossy compression. *** [figure to illustrate].
+An extension of this application is to determine the number of features needed to keep the compression within a certain error criterion, as used in~\cite{barbivc2004segmenting}.
+Sets of data can then be distinguished by the number of features needed to describe the points.
 
@@
 \subsection{Hidden Markov Models}
-Many dynamic systems can be viewed as some sequence of consecutive states
-which are observable by the world. These kind of systems are known as Markov
-Chains. When the state in which the system is can not be observed with
-complete certainty, a system known as a Hidden Markov Model (HMM) can be
-constructed. It has been shown that HMMs are effective on in the task of
-speech recognition \cite{rabiner1989tutorial}. Because of some resemblance
-between natural speech construction and human activities, modifications have
-been able to classify human activities \cite{guenterberg2009distributed} and
-extensions are implemented which create layers of HMMs to handle different
-granularity over time and reduce retraining
-\cite{oliver2002layered,perdikis2008recognition}. A HMM is a state-space model
-representing a single model for each label which needs to be classified.
+Many dynamic systems can be viewed as some sequence of consecutive states which are observable by the world.
+These kind of systems are known as Markov Chains.
+When the state in which the system is can not be observed with complete certainty, a system known as a Hidden Markov Model (HMM) can be constructed.
+It has been shown that HMMs are effective on in the task of speech recognition~\cite{rabiner1989tutorial}.
+Because of some resemblance between natural speech construction and human activities, modifications have been able to classify human activities~\cite{guenterberg2009distributed} and extensions are implemented which create layers of HMMs to handle different granularity over time and reduce retraining \cite{oliver2002layered,perdikis2008recognition}.
+A HMM is a state-space model representing a single model for each label which needs to be classified.
 
@@
 \begin{enumerate}
-	\item A number of $N$ states in the model. These states are a
-	representation of the underlying mechanism of the system and can not be
-	directly observed from outside (hence the \emph{hidden} model). Individual
-	states are referenced by $S = \{S_1, S_2, \dots, S_N \}$ and a single
-	state in time $t$ is denoted by $q_t$.
-	\item An alphabet of $M$ discrete observable symbols for each state. An
-	observed symbol is the only indication of the systems current situation.
+	\item
+		A number of $N$ states in the model.
+		These states are a representation of the underlying mechanism of the system and can not be directly observed from outside (hence the \emph{hidden} model).
+		Individual states are referenced by $S = \{S_1, S_2, \dots, S_N \}$ and a single state in time $t$ is denoted by $q_t$.
+	\item
+		An alphabet of $M$ discrete observable symbols for each state.
+		An observed symbol is the only indication of the systems current situation.
 		Each individual symbol is referenced by $V = \{ v_1, v_2, \dots, v_m \}$.
-	\item A state transition probability distribution $A = \{ a_{ij} \}$
-	indication the probability of being in state $j$ at time $t+1$ when the
-	state on time $t$ was $i$:
+	\item
+		A state transition probability distribution $A = \{ a_{ij} \}$ indication the probability of being in state $j$ at time $t+1$ when the state on time $t$ was $i$:
 		\begin{eqnarray}
@@
 		\end{eqnarray}
-	Note that by setting the probability $a_{ij}=0$ for some $i$ and $j$
-	indicates that $j$ is not reachable from state $i$.
-	\item A symbol observation probability distribution in state $j$, $B = \{
-	b_j(k) \}$, where
+		Note that by setting the probability $a_{ij}=0$ for some $i$ and $j$ indicates that $j$ is not reachable from state $i$.
+	\item
+		A symbol observation probability distribution in state $j$, $B = \{b_j(k) \}$, where
 		\begin{eqnarray}
@@
 		Note that every symbol can be observed in every state.
-	\item The initial state distribution in which the system starts $\pi= \{
-	\pi_i \}$ where
+	\item
+	The initial state distribution in which the system starts $\pi= \{ \pi_i \}$ where
 	\begin{eqnarray}
@@
 \end{enumerate}
+
 This system is often abbreviated to
@@
 \end{equation}
-The HMM will act as a generator giving a sequence of $T$ observations, each
-being a symbol from $V$:
+The HMM will act as a generator giving a sequence of $T$ observations, each being a symbol from $V$:
 \begin{equation}
@@
 
-Given the characteristics of a specific HMM, there are three major application
-which serve the classification of an activity. The first application is, given
-a sequence of observations $O = O_1,O_2,\dots,O_T$, to label this sequence
-with the activity with
-the highest associated probability for that label, e.g. compute $P(O|\lambda)$
-for each model $\lambda$. The second task a HMM can perform is to
-\emph{explain} a observation sequence $O$, by choosing an optimal
-corresponding state sequence $Q = q_1,q_2,\dots,q_T$. The last application is
-more of a problem statement, being the task to train the HMM to adjust the
-parameters $\lambda = \{ \pi_i, a_{ij}, b_j(k)\}$ to maximize $P(O|\lambda)$.
-In the overall task of activity recognition, one would first use the last
-application to train a set of HMMs with the provided and labeled training
-data. The second application can be used to further optimize the HMM (set the
-number of states, adjust vector codebook etc.) and improve the capability of
-modeling the activities. Finally, when the system is provided with unlabeled
-data the first application is used to identify the observed sequence with
-trained examples, by comparing the probability result of multiple HMMs.
+Given the characteristics of a specific HMM, there are three major application which serve the classification of an activity.
+The first application is, given a sequence of observations $O = O_1,O_2,\dots,O_T$, to label this sequence with the activity with the highest associated probability for that label, e.g.
+compute $P(O|\lambda)$ for each model $\lambda$.
+The second task a HMM can perform is to \emph{explain} a observation sequence $O$, by choosing an optimal corresponding state sequence $Q = q_1,q_2,\dots,q_T$.
+The last application is more of a problem statement, being the task to train the HMM to adjust the parameters $\lambda = \{ \pi_i, a_{ij}, b_j(k)\}$ to maximize $P(O|\lambda)$.
+In the overall task of activity recognition, one would first use the last application to train a set of HMMs with the provided and labeled training data.
+The second application can be used to further optimize the HMM (set the number of states, adjust vector codebook etc.) and improve the capability of modeling the activities.
+Finally, when the system is provided with unlabeled data the first application is used to identify the observed sequence with trained examples, by comparing the probability result of multiple HMMs.
 
-Generally a HMM can be fully connected, or ergodic model. Other configurations
-are also possible, of which the left-right model, or Bakis model, is popular
-for temporal applications \cite{rabiner1989tutorial}. This because the
-structure of the state probability distribution resembles the construction of
-possible states over time. This left-right model can be generalized by setting
+Generally a HMM can be fully connected, or ergodic model.
+Other configurations are also possible, of which the left-right model, or Bakis model, is popular for temporal applications~\cite{rabiner1989tutorial}.
+This because the structure of the state probability distribution resembles the construction of possible states over time.
+This left-right model can be generalized by setting
 \begin{eqnarray}
@@
 \end{eqnarray}
-for some $\Delta$, indicating the possible \emph{distance} of the jumps over
-states.
+for some $\Delta$, indicating the possible \emph{distance} of the jumps over states.
 
@@
 
-*** Applications of HHM: \cite{shi2009towards} uses 10 coefficients extracted
-with FFT from 6 axis (3 times accelerometer and 3 times gyroscope). Example of
-continuous HMM (?). Recognition rates fo 50\%-100\%.
+*** Applications of HHM:~\cite{shi2009towards} uses 10 coefficients extracted with FFT from 6 axis (3 times accelerometer and 3 times gyroscope).
+Example of continuous HMM (?).
+Recognition rates fo 50\%-100\%.
 
-In \cite{oliver2002layered} it is stated that HMMs are robust mechanisms with
-respect to changes in temporal segmentation of observations, although they
-suffer from a few problems when applied to reason about longer and complexer
-sequences over time. They state that, especially with limited training data,
-the HMM tend to be overfitted and tend to lack structure and have an excess of
-parameters. The overcome this limitation, they introduce a Layered Hidden
-Markov Model (LHHM) which they show is more robust to temporal changes in
-observations. They make use of the hypothesis from the field of psychology,
-that human behavior is hierarchically structured \cite{zacks2001event}. In
-\cite{perdikis2008recognition} a layered construction of HMMs is used which
-feeds the low-level data from a sliding window to the lower layer. This layer
-will recognize simple motions, such as moving the arm in a pre-defined
-segment. The results from these lower HMMs are then fed to a higher-level
-layer of HMMs which are able to recognize abstract motions, such as adjusting
-a monitor and picking up a pen. The results showed that the layered system was
-better at recognition then a similar single-layered setup. It was also robust
-in the sense that when the environment was changed, only the relatively simple
-low-level HMMs needed to be retrained.
+In~\cite{oliver2002layered} it is stated that HMMs are robust mechanisms with respect to changes in temporal segmentation of observations, although they suffer from a few problems when applied to reason about longer and complexer sequences over time.
+They state that, especially with limited training data, the HMM tend to be overfitted and tend to lack structure and have an excess of parameters.
+The overcome this limitation, they introduce a Layered Hidden Markov Model (LHHM) which they show is more robust to temporal changes in observations.
+They make use of the hypothesis from the field of psychology, that human behavior is hierarchically structured~\cite{zacks2001event}.
+In \cite{perdikis2008recognition} a layered construction of HMMs is used which feeds the low-level data from a sliding window to the lower layer.
+This layer will recognize simple motions, such as moving the arm in a pre-defined segment.
+The results from these lower HMMs are then fed to a higher-level layer of HMMs which are able to recognize abstract motions, such as adjusting a monitor and picking up a pen.
+The results showed that the layered system was better at recognition then a similar single-layered setup.
+It was also robust in the sense that when the environment was changed, only the relatively simple low-level HMMs needed to be retrained.
 
-*** [DIAGRAM WITH EXAMPLE OF LOW- AND HIGH-LEVEL HMMs, SUCH AS MOVING ARM,
-PICKING UP PEN]
+*** [DIAGRAM WITH EXAMPLE OF LOW- AND HIGH-LEVEL HMMs, SUCH AS MOVING ARM, PICKING UP PEN]
 
-*** Viterbi algorithm / path; find the most likely sequence of cause-states
-for
-the observed measures.
+*** Viterbi algorithm / path; find the most likely sequence of cause-states for the observed measures.
 
@@
 \subsection{Segmentation as clustering}
-In the previous section the discussed methods all relied on the stream of data
-points and tried to find cuts the discriminate between successive different
-type of activities. An other approach is to consider the tasks of segmentation
-as a type of clustering \cite{zhou2008aligned}. In clustering the objective is
-to assign labels, or classes, to all the data points indication a similar type
-of activity. A clustering $\mathcal{L}$ is thereby more informative then a
-segmentation but is also harder to produce.
+In the previous section the discussed methods all relied on the stream of data points and tried to find cuts the discriminate between successive different type of activities.
+An other approach is to consider the tasks of segmentation as a type of clustering~\cite{zhou2008aligned}.
+In clustering the objective is to assign labels, or classes, to all the data points indication a similar type of activity.
+A clustering $\mathcal{L}$ is thereby more informative then a segmentation but is also harder to produce.
 
-A clustering $\mathcal{L}$ is generated from a sequence of elements
-$\mathbf{X}$ which is decomposed in $m$ disjoint segments, each belonging to
-one of the $k$ classes. A segment $\mathbf{Y}_i \hat{=}
-\mathbf{X}_{[s_i,s_{i+1})}$ is composed of frames from position $s_i$ to
-$s_{i+1}$. A vector $g_{ci} = 1$ indicates class membership if $\mathbf{Y}_i$
-belongs to class $c$, otherwise $g_{ci} = 0$.
+A clustering $\mathcal{L}$ is generated from a sequence of elements $\mathbf{X}$ which is decomposed in $m$ disjoint segments, each belonging to one of the $k$ classes.
+A segment $\mathbf{Y}_i \hat{=} \mathbf{X}_{[s_i,s_{i+1})}$ is composed of frames from position $s_i$ to $s_{i+1}$.
+A vector $g_{ci} = 1$ indicates class membership if $\mathbf{Y}_i$ belongs to class $c$, otherwise $g_{ci} = 0$.
 
-When regarding segmentation of human motion as a task of clustering the
-difficulty is to model the temporal variability of actions
-and defining a robust metric between temporal actions. To overcome this,
-\cite{zhou2008aligned} introduces Aligned Cluster Analysis (ACA), by minimizing
-\begin{equation} \label{eq:ACA}
-J_{\mathit{ACA}}(\mathbf{G},\mathbf{s}) =
-\sum_{c=i}^{K} \sum_{i=1}^{m} g_{ci} \mathit{dist}_c
-(\mathbf{X}_{[s_i,s_{i+1})})
+When regarding segmentation of human motion as a task of clustering the difficulty is to model the temporal variability of actions and defining a robust metric between temporal actions.
+To overcome this, \cite{zhou2008aligned} introduces Aligned Cluster Analysis (ACA), by minimizing
+\begin{equation}
+	\label{eq:ACA}
+	J_{\mathit{ACA}}(\mathbf{G},\mathbf{s}) = \sum_{c=i}^{K} \sum_{i=1}^{m} g_{ci} \mathit{dist}_c (\mathbf{X}_{[s_i,s_{i+1})})
 \end{equation}
 
-The characteristic of ACA is that is enables segments to span over different
-number of data points, whereas the standard kernel $k$-means algorithm results
-in equally sized segments. [!!! TRUE?!] The second difference is that the
-kernel used in $\mathit{dist}_c$ to measure the distance from a segment to the
-class which it is assigned to uses the Dynamic Time Alignment Kernel
-[REFERENCE?] to measure between time series.
+The characteristic of ACA is that is enables segments to span over different number of data points, whereas the standard kernel $k$-means algorithm results in equally sized segments.
+[!!!
+TRUE?!] The second difference is that the kernel used in $\mathit{dist}_c$ to measure the distance from a segment to the class which it is assigned to uses the Dynamic Time Alignment Kernel [REFERENCE?] to measure between time series.
 
@@
 This section will give an introduction and overview to clustering of data.
-When processing time series of data the line of distinction between
-segmentation and clustering is very fine. This section will introduce
-clustering for the purpose as used in this project.
+When processing time series of data the line of distinction between segmentation and clustering is very fine.
+This section will introduce clustering for the purpose as used in this project.
 
-Data clustering can be used with two different applications, exploratory or
-confirmatory \cite{jain1999data}. In the former, the purpose is to discover
-clusters in a point cloud of data points. Here the cluster is defined as a
-group of homogeneous data points measured by some coherence measure, often a
-distance function with the data points being defined as vectors. The produced
-clusters, which together form a representation of the data examined, can be
-used in the
-latter application to assign new provided data points to a cluster and thereby
-classify them.
+Data clustering can be used with two different applications, exploratory or confirmatory~\cite{jain1999data}.
+In the former, the purpose is to discover clusters in a point cloud of data points.
+Here the cluster is defined as a group of homogeneous data points measured by some coherence measure, often a distance function with the data points being defined as vectors.
+The produced clusters, which together form a representation of the data examined, can be used in the latter application to assign new provided data points to a cluster and thereby classify them.
 
-Depended on the precise setup, clustering can be used as a successive step
-after or an implementation of temporal segmentation. When temporal
-segmentation is used to create successive coherent data points, clustering can
-be used to recognize the same activities in different points of time. The data
-points provided to the clustering will be some representation of each segment
-and each segment will be labeled with a activity. Another mechanism could be
-to extract features from the raw data points and use these to create clusters
-directly *** [temporal modifications]. Both mechanisms requires the
-exploratory and confirmatory phases.
+Depended on the precise setup, clustering can be used as a successive step after or an implementation of temporal segmentation.
+When temporal segmentation is used to create successive coherent data points, clustering can be used to recognize the same activities in different points of time.
+The data points provided to the clustering will be some representation of each segment and each segment will be labeled with a activity.
+Another mechanism could be to extract features from the raw data points and use these to create clusters directly *** [temporal modifications].
+Both mechanisms requires the exploratory and confirmatory phases.
 
@@
 
-The first step is to find a \emph{representation} of the data points and
-clusters, considering the number, type and scale of the features and the
-number of desired classes to cluster in. When dealing with high-dimensional
-data points a \emph{feature selection} can be performed to use only the most
-discriminating features. When the raw features are not useful enough,
-\emph{extraction} can be used to create new synthetic features. ***
-[qualitative versus quantitative/conceptual]
+The first step is to find a \emph{representation} of the data points and clusters, considering the number, type and scale of the features and the number of desired classes to cluster in.
+When dealing with high-dimensional data points a \emph{feature selection} can be performed to use only the most discriminating features.
+When the raw features are not useful enough, \emph{extraction} can be used to create new synthetic features. *** [qualitative versus quantitative/conceptual]
 
-When the data points are represented in a meaningful way, the measure of
-\emph{coherence} between pairs of points must be defined. Often this is
-implemented as the Euclidean
-or Mahalanobis distance when the data points are represented as vectors or
-other similarity measures for conceptual patterns.
+When the data points are represented in a meaningful way, the measure of \emph{coherence} between pairs of points must be defined.
+Often this is implemented as the Euclidean or Mahalanobis distance when the data points are represented as vectors or other similarity measures for conceptual patterns.
 
-In the most critical step, the \emph{grouping}, the data points with the
-highest coherence are linked together to form a cluster. The result can be a
-hard partition over the data or a fuzzy membership degree to each cluster for
-each data point. When applying rules for merging and splitting cluster, a
-hierarchical partition is constructed. Partitional clustering algorithms
-assign data points to clusters by optimizing some criterion, e.g. the mean
-squared distance from each data point to the clusters' centroid.
+In the most critical step, the \emph{grouping}, the data points with the highest coherence are linked together to form a cluster.
+The result can be a hard partition over the data or a fuzzy membership degree to each cluster for each data point.
+When applying rules for merging and splitting cluster, a hierarchical partition is constructed.
+Partitional clustering algorithms assign data points to clusters by optimizing some criterion, e.g. the mean squared distance from each data point to the clusters' centroid.
 
-Especially in case of large data sets, the resulting clusters will be defined
-by many data points. To form a compacter and simpler representation,
-\emph{abstraction} can be used. This will simply analysis by humans or
-automated systems. A common abstraction is to use the clusters' centroid
-\cite{diday1976clustering} or parameters of Gaussian patterns.
+Especially in case of large data sets, the resulting clusters will be defined by many data points.
+To form a compacter and simpler representation, \emph{abstraction} can be used.
+This will simply analysis by humans or automated systems.
+A common abstraction is to use the clusters' centroid \cite{diday1976clustering} or parameters of Gaussian patterns.
 
-The final step which can be applied to the resulting partition is some
-\emph{qualification}. The quality of multiple partitions can be compared and
-the validity of the partition can be determined. A partition is valid if
-reasonably it could not been constructed by change or as some random process.
-External references of models can be used to compare, or the internal data
-points can be examined. *** [better wordings]
+The final step which can be applied to the resulting partition is some \emph{qualification}.
+The quality of multiple partitions can be compared and the validity of the partition can be determined.
+A partition is valid if reasonably it could not been constructed by change or as some random process.
+External references of models can be used to compare, or the internal data points can be examined. *** [better wordings]
 
@@
 \subsection{Dynamic Time Warping}
-Used to measure similarity between time sequences. Exact matching is
-high-cost, so approximations such as Minimum Bounding Rectangles are used.
+Used to measure similarity between time sequences.
+Exact matching is high-cost, so approximations such as Minimum Bounding Rectangles are used.
 
 \subsection{k-means clustering}
-[EXPLAIN] divide data set $n$ into $k$ clusters.
+[EXPLAIN] divide data set $n$ into $k$ clusters. Follows the outline in figure~\ref{fig:k_means}
 
-Among many unsupervised clustering techniques, $k$-means is successfully
-applied to large data sets. It is simple to implement and linear in time
-complexity so computationally attractive \cite{jain1999data}. A drawback of
-the method is that the results of the algorithm greatly depends on the initial
-configuration (the data points which will act as centroids) and the number of
-cluster $k$ must be determined beforehand.
+Among many unsupervised clustering techniques, $k$-means is successfully applied to large data sets.
+It is simple to implement and linear in time complexity so computationally attractive~\cite{jain1999data}.
+A drawback of the method is that the results of the algorithm greatly depends on the initial configuration (the data points which will act as centroids) and the number of cluster $k$ must be determined beforehand.
 
-Generally, the $k$-means methods will minimize the squared error for a
-clustering $\mathcal{L}$ criterion which is defined as the distance from the
-data points the centroid for each cluster in $\mathcal{K}$. This is expressed
-as optimizing to a local optimum the energy function
-\begin{equation} \label{eq:k-means energy}
+\begin{figure}[htbp]
+	\centering
+		\includegraphics{./Figures/graphs/k_means.png}
+		\rule{35em}{0.5pt}
+	\caption[K-means]{Outline of the $k$-means algorithm.}
+	\label{fig:k_means}
+\end{figure}
+
+Generally, the $k$-means methods will minimize the squared error for a clustering $\mathcal{L}$ criterion which is defined as the distance from thedata points the centroid for each cluster in $\mathcal{K}$.
+This is expressed as optimizing to a local optimum the energy function
+\begin{equation}
+	\label{eq:k-means energy}
 	e^2(\mathcal{K},\mathcal{L}) =
@@
 
-There are several limitation on the $k$-means method. One of these is that
-only spherical shapes of cluster can be generated. One of the extensions is
-kernel $k$-means \cite{scholkopf1998nonlinear}, which implicitly projects the
-data points to a higher dimension and thereby is able to form irregular shaped
-cluster.
+There are several limitation on the $k$-means method.
+One of these is that only spherical shapes of cluster can be generated.
+One of the extensions is kernel $k$-means~\cite{scholkopf1998nonlinear}, which implicitly projects the data points to a higher dimension and thereby is able to form irregular shaped cluster.
 

### 2013-02-08 Chapters/Chapter2.tex
Cleanup. Extented intended contents

@@
 When a peak in the distance is reached, the new window of frames indicates a heterogeneous collection of motions in the window and thus a low likelihood of membership and a indication of a transition.
-In order to distinct activities and subactivities (which require a subset of motions is a distinct activity) the algorithm is also processed backward over the data series.
+In order to distinct activities and sub-activities (which require a subset of motions is a distinct activity) the algorithm is also processed backward over the data series.
 

### 2013-02-08 intended_contents.tex
Added drafts for some diagrams
Removed linebreaks
Cleanup. Extented intended contents

@@
 \section{Introduction}
-This section introduces the subject, outlines the global goal of the thesis
-and explains the problem. The main research-questions are introduced. In the
-subsections the main components (global, not as an implementation) are
-introduced.
+\label{sec:Introduction}
+This section introduces the subject, outlines the global goal of the thesis and explains the problem.
+The main research-questions are introduced.
+In the subsections the main components (global, not as an implementation) are introduced.
+Outline of the thesis is presented.
+Chapter~\ref{sec:Concepts} introduces the big parts of the puzzle.
+Chapter~\ref{sec:Setup} gives details about the algorithms used.
+Chapter~\ref{sec:Implementation} will give the implementation details.
+Chapter~\ref{sec:Experiments} will show how data is gathered and shows the data.
+Chapter~\ref{sec:Discussion} will discuss and relate the results
+Chapter~\ref{sec:Conclusion} will conclude and give final remarks.
 
   \subsection{Thesis goal}
-  The goal of the thesis will be discussed, e.g. "precise unsupervised on-line human activity recognition". Mention application to show why it is useful.
+  The goal of the thesis will be discussed, e.g. "precise unsupervised on-line human activity recognition".
+  Mention application to show why it is useful, such as medical and other health (losing weight, fitness) problems.
+  The main research questions will be formulated here.
 
   \subsection{Results of earlier research}
-  First, results of earlier project are outlined.
-Look at the data being used, the precision and error rates (shortly, no details) and look at the different concepts being discussed.
+  First, results of earlier project are outlined. Look at the data being used, the precision and error rates (shortly, no details) and look at the different concepts being discussed.
   This will give reason for the following subsections.
@@
 \section{Concepts}
+\label{sec:Concepts}
+This chapter will introduce the abstract concepts which will be used.
+There will be some formulas, but minimal implementations.
+The focus is on the meaning of the terms and what they do and how they do it.
+Real implementations will be discussed in chapter~\ref{sec:Setup}.
 
   \subsection{Temporal pattern recognition}
-  This will explain about what patterns are and what kind of patterns need to be recognized (long-term, short-term repetitive, etc)
+  This will explain about what patterns are and what kind of patterns need to be recognized (long-term, short-term repetitive, etc).
+  This is not specific about segmentation of clustering, but more as the ultimate goal.
 
 	\subsection{Signal pre-processing}
-  Describe how data is collected and processed. Include feature extraction, like PCA, FFT. Mention only shortly with meaning, no further explanation.
+  Describe how data is collected and processed, the form of the data (continuous to discrete, amplitude, energy, frequencies, etc).
+  Include feature extraction, like PCA, FFT, energy, mean, but discuss it further in subsection.
+  Only the concepts will be explained and the scheme of the computations (or simply formulas), no precise implementations.
+  Mention only shortly with meaning, no further explanation.
+
+    \subsubsection{Signal fusion}
+    What, why and how.
+
+    \subsubsection{Signal smoothing}
+    What, why and how.
+
+    \subsubsection{Feature extraction}
+    What, why and how.
 
 	\subsection{Temporal Segmentation}
-  What is segmentation, the purpose and what kind of methods are there. E.g. model-based, statistics-based. Diagram with types. Supervised / unsupervised. HMMs, Statistical, Bayesian. K-means, SOM
+  What is segmentation, the purpose and what kind of methods are there.
+  Disctinction with clustering.
+  E.g. model-based, statistics-based.
+  Diagram with types.
+  Supervised / unsupervised.
+
+    \subsubsection{Methods}
+    Describe methods, such as HMMs, Statistical, Bayesian. K-means, SOM
 
   \subsection{Temporal clustering}
-  When is segmenting clustering? Explain difference, mention techniques (diagrams), Dynamic Time Warping, k-means (and variations), SOM, SVM, Naive Bayes, Neural Networks.
+  When is segmenting clustering?
+  Explain difference of point of view/mechanism, mention techniques (diagrams).
+  This is the unsupervised part.
 
-\section{Problem statement}
+    \subsubsection{Methods}
+    Explain difference in techniques and give schematic overview of inner workings.
+    Dynamic Time Warping, k-means (and variations), SOM, SVM, Naive Bayes, Neural Networks.
+
+  \subsection{Matching}
+  Not really sure about this section yet.
+  It would be about matching new on-line gathered data with already learned patterns.
+  But maybe it is more of a on-line combination of segmentation and clustering.
+
+\section{Algorithms setup}
+\label{sec:Setup}
 This section proposes a setup to improve (or: answer) the research-questions.
 It will discuss techniques used in other projects and how implemented here.
-In the subsections each used technique is further explained.
-(Only explained when really used)
+In the subsections each used technique is further explained. (Only explained when really used)
 
   \subsection{Proposed method}
-  Give an overview of the "pipeline" of the implementation. Give a diagram how data is retrieved, (pre-)processed, analyzed and conclusions over data are drawn.
+  Give an overview of the "pipeline" of the implementation.
+  Give a diagram how data is retrieved, (pre-)processed, analyzed and conclusions over data are drawn.
 
   \subsection{PCA}
-  Something about the workings of PCA and why it is relevant and used. Discuss how it is used in other projects and how it will be used here.
+  Something about the workings of PCA and why it is relevant and used.
+  Discuss how it is used in other projects and how it will be used here.
 
@@
 \section{Implementation}
-This section will discuss implementation details. Hardware platform (if any), software used. Complexity, precision of metrics etc.
+\label{sec:Implementation}
+This section will discuss implementation details.
+Hardware platform (if any), software used. Complexity, precision of metrics etc.
+Data-linking between algorithms, pipeline construction.
 
 \section{Experiments, results}
+\label{sec:Experiments}
 This section will outline the precise experiments and gives the results of these.
@@
 \section{Discussion}
-This section will discuss the results of the previous section. It will relate the results to previous work.
+\label{sec:Discussion}
+This section will discuss the results of the previous section.
+It will relate the results to previous work.
 
 \section{Conclusion}
+\label{sec:Conclusion}
 Conclusion (answer research question), summary of results, feature work.

### 2013-02-08 Chapters/Chapter2/Section_signals.tex
Diagrams for sensor processing

--- 2013-02-07 intended_contents.tex
@@
-\documentclass[a4paper,10pt]{extarticle}
-\usepackage[english]{babel}
-\usepackage{fullpage}
-\usepackage{hyperref}
-\author{R.Q. Vlasveld}
-\title{Intended Table of Contents Thesis}
+% Section Signal Pre-processing
+% !TEX root = ../../main.tex
 
-\begin{document}
-\maketitle
-\tableofcontents
+\section{Signal pro-processing and sensor fusion for timed patterns}
+  Describe how data is collected and processed, the form of the data (continuous to discrete, amplitude, energy, frequencies, etc).
+  Include feature extraction, like PCA, FFT, energy, mean, but discuss it further in subsection.
+  Only the concepts will be explained and the scheme of the computations (or simply formulas), no precise implementations.
+  Mention only shortly with meaning, no further explanation.
 
-\section{Introduction}
-This section introduces the subject, outlines the global goal of the thesis
-and explains the problem. The main research-questions are introduced. In the
-subsections the main components (global, not as an implementation) are
-introduced.
+    \subsubsection{Signal fusion}
+    What, why and how.
 
-  \subsection{Thesis goal}
-  The goal of the thesis will be discussed, e.g. "precise unsupervised on-line human activity recognition". Mention application to show why it is useful.
+    \subsubsection{Signal smoothing}
+    What, why and how.
 
-  \subsection{Results of earlier research}
-  First, results of earlier project are outlined.
-Look at the data being used, the precision and error rates (shortly, no details) and look at the different concepts being discussed.
-This will give reason for the following subsections.
-
-\section{Concepts}
-
-  \subsection{Temporal pattern recognition}
-  This will explain about what patterns are and what kind of patterns need to be recognized (long-term, short-term repetitive, etc)
-
-	\subsection{Signal pre-processing}
-  Describe how data is collected and processed. Include feature extraction, like PCA, FFT. Mention only shortly with meaning, no further explanation.
-
-	\subsection{Temporal Segmentation}
-  What is segmentation, the purpose and what kind of methods are there. E.g. model-based, statistics-based. Diagram with types. Supervised / unsupervised. HMMs, Statistical, Bayesian. K-means, SOM
-
-  \subsection{Temporal clustering}
-  When is segmenting clustering? Explain difference, mention techniques (diagrams), Dynamic Time Warping, k-means (and variations), SOM, SVM, Naive Bayes, Neural Networks.
-
-\section{Problem statement}
-This section proposes a setup to improve (or: answer) the research-questions.
-It will discuss techniques used in other projects and how implemented here.
-In the subsections each used technique is further explained.
-(Only explained when really used)
-
-  \subsection{Proposed method}
-  Give an overview of the "pipeline" of the implementation. Give a diagram how data is retrieved, (pre-)processed, analyzed and conclusions over data are drawn.
-
-  \subsection{PCA}
-  Something about the workings of PCA and why it is relevant and used. Discuss how it is used in other projects and how it will be used here.
-
-  \subsection{Hidden Markov Models}
-  Something about the workings of HMMs and why it is relevant and used.
-
-  \subsection{Dynamic Time Warping}
-  if used
-
-  \subsection{k-Means clustering}
-  if used
-
-  \subsection{Self-organizing maps}
-  if used
-
-\section{Implementation}
-This section will discuss implementation details. Hardware platform (if any), software used. Complexity, precision of metrics etc.
-
-\section{Experiments, results}
-This section will outline the precise experiments and gives the results of these.
-
-\section{Discussion}
-This section will discuss the results of the previous section. It will relate the results to previous work.
-
-\section{Conclusion}
-Conclusion (answer research question), summary of results, feature work.
-
-\end{document}
+    \subsubsection{Feature extraction}
+    What, why and how.

### 2013-02-08 Chapters/Chapter2/Section_signals.tex
Beginning text of sensor fusion

@@
 
-    \subsubsection{Signal fusion}
+  --------
+
+    \subsection{Signal fusion}
+    Signal fusion is a broad field of research and application and many different interpretations exist.
+    The function and interpretation used in this thesis follows the definition from \cite{elmenreich2001introduction}, stating:
+    \begin{center}
+      \textbf{Sensor fusion} is the combining of sensory data or data derived from sensory data such that the resulting information is in some sense better than would be possible when these sources were used individually.
+    \end{center}
+    This can essentialy be considered as a form of synergy, in which the combination of the sensors provide more information than the sum of the raw output.
+    For the application of this thesis we will focus on direct fusion of raw data, which can be applied to both homogeneous and heterogeneous sensors.
+    The fusion stage will combine the data into a single representation, which can be interpreted by the controlling system.
+    This differs from mere multisensor integration, in which the controlling system is responsible for processing and interpreting the data from different sources.
+    Figure \ref{fig:sensor_fusion} shows the difference.
+
+    \begin{figure}[htbp]
+      \centering
+        \includegraphics{./Figures/graphs/sensor_fusion.png}
+        % \rule{35em}{0.5pt}
+      \caption[K-means]{On the left \emph{Sensor Fusion} and on the right \emph{Multisensor integration}}
+      \label{fig:sensor_fusion}
+    \end{figure}
+
+    Using the single representation, the computational and system complexity can be reduced.
+    The representation forms a standardized and normalized interface from the sensors to the system, which create independence and modularity of sensors.
+    Other benefits as a result from sensor fusion include \cite{elmenreich2001introduction}:
+    \begin{itemize}
+      \item \textbf{Robustness en reliability:} When the system incorporates multiple sensors as redundancy, the system becomes more stable in case of partial failure.
+      \item \textbf{Extended spatial and temporal coverage:} With multiple sensors a broader spectrum can be analyzed, both in the case of homogeneous and heterogeneous combinations.
+      \item \textbf{Increases precision:} Sensors typically have a range and domain of measurements.
+      By combining a set of heterogeneous sensors, the overall range of observervation is increased.
+      \item \textbf{Reduced ambiguity and uncertainty:} In case of high uncertainty about a observation, the observation of other sensors can confirm it.
+    \end{itemize}
+
+    Sensor fusion can be applied on multiple levels. A rough diversion over three levels is:
+    \begin{itemize}
+      \item \textbf{Low-level:} Raw sensor data is combined which is considered to be more informative than the original sources.
+      \item \textbf{Intermediate-level:} Also known as \emph{feature fusion} combines the features which can be extracted from the data.
+      See also section \ref{sec:feature_extraction}.
+      This data would be applicable to segmentation **[???].
+      \item \textbf{High-level:} Fusion of this level will support decision making, optionally followed by actions taken by the system to the environment or conclusion being drawn from the environment.
+    \end{itemize}
+
+    \subsection{Signal smoothing}
     What, why and how.
 
-    \subsubsection{Signal smoothing}
+    \subsection{Feature extraction}
+    \label{sec:feature_extraction}
     What, why and how.
-
-    \subsubsection{Feature extraction}
-    What, why and how.

### 2013-02-28 Chapters/Chapter1.tex
More on HMM
Removed linebreaks
Extend Literature review
Working on the literature review. For now discussed a few papers quite
elaborate. Goal is to discuss more papers and group them.

--- 2013-01-30 Chapters/Chapter1.tex
@@
 \section{Problem statement}
-*** Give context problems, means to reach (e.g. monitoring of patients, 
-getting conclusions over health and activity).
+*** Give context problems, means to reach (e.g. monitoring of patients,  getting conclusions over health and activity).
 %----------------------------------------------------------------------------------------
 
-\section{Literature review}
-Look into earlier application mentioned in the literature to this kind of 
-problems. Look for similarities in the problem and address where the 
-techniques used fail or are not applicable.
+\input{./Chapters/Chapter1/literature_review}
 
@@
 
-\section{Learning \LaTeX{}}
+\section{System architecture}
+
+*** [GRAPHICS diagram: sensors --> signal pre-processing --> feature extraction --> analysis [* segmentation, * labeling] --> conclusions]
+
+%----------------------------------------------------------------------------------------

### 2013-02-28 Chapters/Chapter1/literature_review.tex
Extend Literature review
Working on the literature review. For now discussed a few papers quite
elaborate. Goal is to discuss more papers and group them.

% Section Signal Pre-processing
% !TEX root = ../../main.tex

% \section{Literature review}
% Look into earlier application mentioned in the literature to this kind of  problems. Look for similarities in the problem and address where the  techniques used fail or are not applicable.

% *** Most of the activity recognition techniques can be classified in two  different types.
% Some rely on a state-space model, in which the activities to be recognized are represented in a statistical manner.
% Bayesian networks, Finite State Machines and Hidden Markov Models can be considered among those.
% Other techniques process the directly as a pattern recognition task, such as  Support Vector Machines, Neural Networks, Dynamic Time Warping, and Bayes and K-means clustering.

% \subsection{Segmentation}
% Describe the different approaches to segmentation.
% Group different viewpoints, e.g. cut-points, change detection, segmentation, etc.

\section{Literature review}
This section will provide a review of the existing literature and research of the different subjects discussed in this thesis.
First the different approaches to mere segmentation of signals is discussed, followed by *** other techniques. ***

\subsection{Segmentation}
One of the first stages in analyzing the sensor data time series is to segment the signal.
The goal is to set cut-points in the signal such that all the data points in between consecutive cut-points have some homogeneous property among them. *** "organizing data into homogeneous groups where the within-group-object similarity is minimized and the between-group-object dissimilarity is maximized." ***
This results in the more abstract notion of change (point) detection.
The general concern is to detect the occurrence and time identification of changes in the signal, often expressed as changes in the probability distribution of the time series.
This literature review will discuss the different types of change detection as applied in the research.
To each applications question about the methods, complexity and usability etc. are asked.
On account of usability, it is desired to have a successful segmentation algorithm with minimal user interaction, e.g. defining thresholds and error rates.
** something about what is considered, e.g. complexity, measurements, e.g. ***

Following the organization of \cite{warren2005clustering}, research can be structured in three groups.
The data being processed determines the classification, which can be the raw data, indirectly via features extracted from the raw data or indirectly via models extracted from the raw data.

One of the more simple and directs methods is to Piecewise Linear Represent the signal, while keeping the approximation error below a certain value, as discussed in \cite{keogh2001online}.
As with many algorithms, PLR can be applied to a batch of or real time stream input data (also known as off-line and on-line analyzes, respectively).
Keogh et al. propose a method which combines these different techniques to form a Sliding Window and Bottom-up (SWAB) analyzes.
The algorithm works by first, per window of data points, creating for every two data points a segment.
Neighboring segments are joined together, as long as the approximation error is below the provided value.
This analyzes can be used as a starting point to further analyze the data segments.
Although Keogh et al. apply the SWAB algorithm only direct on the raw data, it is easily imagined how to apply it to extracted features of the signal, e.g. the energy or mean.
In its default application, the method yields little information about each segment; it is merely a linear representation.
The SWAB algorithm works by segmenting parts of the data series by linear approximation while keeping the error below a certain rate.
This rate needs to be user defined (e.g. by an expert), which makes it not possible to apply directly on any signal.
This method only divides the signal into segments, which representations can be processed further to form clusters or related segments.
The complexity of this method is relatively low, $O(Ln)$, where $L$ is the expected segment length and $n$ the number of data points.

The task of segmenting any form of data is made easier when the number of segments is beforehand known.
Many authors apply the well-known \emph{k-means} clustering *** refs needed *** on this type of problems.
An incremental error-minimizing version can determine the optimal number of segments (or actually, clusters in this case).
The proposed method of \cite{himberg2001time}, Global Iterative Replacement, resembles the method of Keogh et al., in the sense that it merges segments.
The algorithm is applied on two different data sets.
The first data sets can be considered as extracted features from sensor data, whilst the second data set works on model properties and thus can give a richer contextual segmentation.
For both types of data sets a cost function is defined, which represents the internal heterogeneity of a segment (e.g. the variance of the data within a segment).
The difference with the method of Keogh et al. is that with GIR the number of segments is known beforehand and during the algorithm this number is, through splits and merges, kept constant.
By running multiple instances of the GIR algorithm with each an increasing number of $k$ segments (bounded by user provided values), the optimal number and location of cut points is determined.

Other approaches form a higher level analyses of the raw data.
As used in \cite{barbivc2004segmenting}, the dimensionality of the data can be used to distinguish between actions.
Using Principal Component Analyses, Barbi{\v{c}} et al. use the dimensionality required to project the data points on a hyperplane while keeping the error below a certain rate as a measure of similarity.
The error threshold is fixed as a range of three times the standard deviation from the mean of the error.
In an extension of their method, using Probabilistic PCA \cite{tipping1999probabilistic}, the Mahanalobis distance \cite{duda1995pattern} for each frame represents similarity of concurrent motions.
In a third method the authors use a Gaussian Mixture Model *** elaborate on this one ***.
With the PPCA method the authors where able to get the best results, with precision of 92\% and recall of 95\%.
The test subjects wore 14 sensors on joints of the body and each frame is represented as a 56-dimensional vector.

*** Elaborate on measures, e.g. precision, recall, confusion matrix etc.  ***
*** Precision is defined as the ratio of reported correct cuts versus the total number of reported cuts. Recall is defined as the ratio of reported correct cuts versus the total number of correct cuts. The closer precision and recall are to 1, the more accurate the algorithm is. ***

### 2013-02-28 Chapters/Chapter1/literature_review.tex
Extend Literature Review

@@
 In an extension of their method, using Probabilistic PCA \cite{tipping1999probabilistic}, the Mahanalobis distance \cite{duda1995pattern} for each frame represents similarity of concurrent motions.
-In a third method the authors use a Gaussian Mixture Model *** elaborate on this one ***.
-With the PPCA method the authors where able to get the best results, with precision of 92\% and recall of 95\%.
+By traversing the Mahanalobis distance as a function of the frames, cut-points are placed on specific points of interest.
+*** NOTE to self: thus this method can segment (and group) unknown and new activities? ***
+In a third method the authors use a Gaussian Mixture Model *** elaborate on this one [does it segment of direct classify?] ***.
+Using this segmenting, and comparing the segments with earlier gathered representations of motions, the PPCA method gets the best results, with precision of 92\% and recall of 95\%.
 The test subjects wore 14 sensors on joints of the body and each frame is represented as a 56-dimensional vector.
+The test considered seven simple human motions: walking, running, sitting down, forward jumping, climbing, arm stretching and punching.
 

### 2013-03-01 Chapters/Chapter1/literature_review.tex
Work of 28 feb
Make list of papers and short summary.

@@
 As with many algorithms, PLR can be applied to a batch of or real time stream input data (also known as off-line and on-line analyzes, respectively).
-Keogh et al. propose a method which combines these different techniques to form a Sliding Window and Bottom-up (SWAB) analyzes.
+Keogh et al. \cite{keogh2001online} propose a method which combines these different techniques to form a Sliding Window and Bottom-up (SWAB) analyzes.
 The algorithm works by first, per window of data points, creating for every two data points a segment.
@@
 Other approaches form a higher level analyses of the raw data.
-As used in \cite{barbivc2004segmenting}, the dimensionality of the data can be used to distinguish between actions.
+As used in \cite{barbivc2004segmenting}, from the field of computer graphics, the dimensionality of the data can be used to distinguish between actions.
 Using Principal Component Analyses, Barbi{\v{c}} et al.
@@
 
+It is very common the use a higher level of features extracted from the raw data.
+The approach of Zhou et al. \cite{zhou2008aligned} maps the data points to a higher dimension by using a kernel function.
+The proposed method, Aligned Cluster Analysis, extends kernel $k$-means clustering on this higher dimensional data set.
+Using the Dynamic Time Aligned Kernel (DTAK) metric, similarity between segments is measured.
+Due to the properties of $k$-means clustering, discovering an unknown number of segments in the data set requires and iterative approach.
+
+*** motifs ***
+
+*** HMM ***
+
+*** Finite mixture models ***
+
+*** Infinite mixture models ***
+
+*****
+List of all papers, shortly categorized
+\begin{itemize}
+  \item ``Discovering characteristic actions from on-body sensor data'' \cite{minnen2006discovering}, cited: 63. Motifs, HHM, DTW, 87\% accuracy. Known motifs.
+  \item ``Recognition of human activities using layered hidden Markov models'' \cite{perdikis2008recognition}, cited: 3.
+HMM, layerd (primitive and abstract actions).
+Uses vision for workplace-activities.
+Direct classification
+  \item ``Accelerometer-Based Gait Analysis, A survey'', \cite{derawi2010accelerometer}, cited: 4.
+Compares methods to distinguish walking of normal, fast, slow.
+Focus on gait, but compares classification methods such as SVM, PCA, KSOM.
+  \item ``An Automatic Segmentation Technique in Body Sensor Networks based on Signal Energy'', \cite{guenterberg2009automatic}, cited: 13.
+Automatic segmenting, adaptive threshold.
+Pure segmenting, no classification.
+Nice and clean method?
+Weakness: single action segmented as multiple.
+Used multiple sensors.
+  \item ``Towards HMM based Human Motion Recognition using MEMS Inertial Sensors'', \cite{shi2009towards}, cited: 13.
+Uses HHM, Fourier transform for features.
+5 fixed activities, trained HMM.
+Correct rates from 90\% to 100\%.
+Classification, not just segmentation
+  \item ``Change-Point Detection in Time-Series Data by Direct Density-Ratio Estimation'', \cite{kawahara2009change}, cited: 41.
+Non-parametric approach, online method, no strong model assumptions.
+Uses multiple datasets, origin referenced.
+Much follow-up work done.
+  \item ``Unsupervised, Dynamic Identification of Physiological and Activity Context in Wearable Computing'', \cite{krause2003unsupervised}, cited: 106(!).
+Combination of KSOM, iterative K-means clustering, transient states removal (using markov model).
+Online algorithm.
+  \item ``Activity Recognition using Cell Phone Accelerometers'', \cite{kwapisz2011activity}, cited: 122 (!).
+\end{itemize}
+
 *** Elaborate on measures, e.g. precision, recall, confusion matrix etc.  ***

### 2013-03-04 Chapters/Chapter1/literature_review.tex
Literature review

@@
 Combination of KSOM, iterative K-means clustering, transient states removal (using markov model).
-Online algorithm.
-  \item ``Activity Recognition using Cell Phone Accelerometers'', \cite{kwapisz2011activity}, cited: 122 (!).
+On-line algorithm.
+  \item ``Activity Recognition using Cell Phone Accelerometers'', \cite{kwapisz2011activity}, cited: 122 (!), 2011.
+Uses WEKA: decision trees, logistic regression and multilayer neural networks.
+Results up to 98\% (jogging).
+Much difficulty with stairs up and down.
+Single accelerometer.
+Offline method.
+Many recent references.
+  \item ``Activity recognition from accelerometer data'', \cite{ravi2005activity}, cited: 434 (!), 2005.
+Compares 18 (offline) classifiers, base level (WEKA) and meta-level.
+Single device, pelvic region.
+  \item ``Activity recognition from user-annotated acceleration data'', \cite{bao2004activity}, cited: 1008 (!!), 2004.
+Five bi-axial sensors.
+Decision tables, instance-based learning, C4.5 (trees, highest accuracy 89.30\%) and naive bayes (weka), twenty activities.
+  \item ``Using Hierarchical Clustering Methods to Classify Motor Activities of COPD Patients from Wearable Sensor Data'', \cite{sherril2005using}, cited: none, 2005.
+Uses Linear Discriminant Analysis for cluster classification (using Cluster Quality Index to determine k) and simple rule-based separation for high level ambulatory.
+Hierarchical Dendogram to merge clusters when similarity to high.
+  \item ``Non-Parametric Bayesian Human Motion Recognition Using a Single MEMS Tri-Axial Accelerometer'', \cite{ahmed2012non}, cited: none, 2012.
+Recognizes the number of human activities, single sensor on the chest.
+No training data.
+Infinite Gaussian Mixture Model, collapsed Gibbs sampler.
+Compares with parametric Fuzzy C-mean (data point belongs to multiple clusters), unsupervised K-means, non-parametric mean-shift.
+Outperforms all significantly, high hit rate and low false alarms.
+  \item ``An Online Algorithm for Segmenting Time Series'', \cite{keogh2001online}, cited: 487, 2001.
+Reviews algorithms to get a piecewise linear representation, proposes a novel sliding-window and bottom up approach, on-line.
+Mere segmentation.
+No classification; only approximation of signal.
+(Dataset available).
+  \item ``Segmenting Motion Capture Data into Distinct Behaviors'', \cite{barbivc2004segmenting}, cited: 242, 2004.
+On-line methods: cut-points on increased intrinsic dimensionality, distribution of poses is observed to change.
+Batch: cut when consecutive frames belong to different Gaussian mixture models.
+Fixed fourteen motions, compared with manual segmentation.
+Uses PCA and Probabilistic PCA (models the non-pc subspaces as noise).
+PPCA is the best.
+(paper anne)
+  \item ``Aligned Cluster Analysis for Temporal Segmentation of Human Motion'', \cite{zhou2008aligned}, cited: 54, 2008.
+Uses extension on kernel k-means clustering and Dynamic Time Alignment Kernel (kernel of DTW) for temporal invariance, robust temporal matching metric.
+Coordinate descent algorithm solves ACA.
+Fixed number of clusters, trapped in local minima.
+(paper anne)
+  \item ``Bayesian Nonparametric Methods for Learning Markov Switching Processes'', \cite{fox2010bayesian}, cited: 10, 2010.
+Uses HMM for state-space model for segmentation, Markov Jump Linear systems.
+Unbounded number of Markov modes (parameters).
+Number of models is fixed, though?
+(paper anne).
+  \item ``Layered Representations for Human Activity Recognition'', \cite{oliver2002layered}, cited: 215, 2002.
+Layers of Hidden Markov Models HMM, multiple levels of granularity (Based on intuition) and context.
+Less retraining, only lower models.
+Classified also.
+  \item ``Time series segmentation for context recognition in mobile devices'', \cite{himberg2001time}, cited: 151, 2001.
+Splits and merges segments, while keeping k constant using cost function on segments for internal heterogeneity.
+Multiple instances determine number of k.
+  \item ``A practical approach to recognizing physical activities'', \cite{lester2006practical}, cited: 259, 2006.
+Uses method of \cite{lester2005hybrid}, activity classification algorithm.
+Selects most useful features and then recognizes walking, sitting, etc.
+First layer static classifier on features or data (energy, mean, variance, correlation, etc), then a layer of HMM to estimate activity.
+All offline.
+Shows trained model is robust to location of wearing on body.
+  \item ``A hybrid discriminative/generative approach for modeling human activities'', \cite{lester2005hybrid}, cited: 252, 2005.
+Used in method above.
+Combines boosting to select and reduce useful features and learn static classifiers with HMM to capture regularities and smooth activities (quick switching is unlikely).
+Single sensor, multiple measures (accelerometer, audio, etc).
+  \item ``Using acceleration measurements for activity recognition: An effective learning algorithm for constructing neural classifiers'', \cite{yang2008using}, cited: 67, 2008.
+Separates dynamic from static activities.
+Uses multilayer feedforward neural networks to generate complex discriminating surfaces as activity classifiers.
+Feature subset selection approach is developed.
+Neural pre-classifier with constant threshold criterion.
+Uses Common Principal Component Analysis.
+Recognizes eight activities with 95\% overall accuracy.
+  \item ``Single-accelerometer-based daily physical activity classification'', \cite{long2009single}, cited: 46, 2009.
+Uses Naive Bayes classifier, sensor worn on wrist.
+Five activities.
+PCA to reduce dimensions and create independence for NB.
+Comparable results as Decision Trees, but with flexibility to add activities.
+  \item ``Distributed Continuous Action Recognition Using a Hidden Markov Model in Body Sensor Networks''.
+Single HMM, sensor network cluster movements, HMM constructs continuous actions using postures and actions.*** explain more on this one ***.
 \end{itemize}

### 2013-03-05 Chapters/Chapter1/literature_review.tex
Small review per paper

@@
 Comparable results as Decision Trees, but with flexibility to add activities.
-  \item ``Distributed Continuous Action Recognition Using a Hidden Markov Model in Body Sensor Networks''.
-Single HMM, sensor network cluster movements, HMM constructs continuous actions using postures and actions.*** explain more on this one ***.
+  \item ``Distributed Continuous Action Recognition Using a Hidden Markov Model in Body Sensor Networks'', cited: 17, 2009.
+Single HMM, sensor network cluster movements, HMM constructs continuous actions using postures and actions, merges left-to-right HMM for each action to a single.
+Transcript generation per sensor uses Gaussian Mixed Model, multiple models with $m$ mixtures are trained with Expectation-Maximization (EM), best is chosen.
+  \item ``Offline and online activity recognition on mobile devices using accelerometer data'', \cite{duque2012offline}, cited: none, 2012.
+Compares offline and online methods, cellphone in pocket.
+Reduces feature extraction (mean and standard deviation (for resources of cellphone).
+Training is done by database and labeling, classification with K-nearest neighbors, decision trees (C4.5) and decision rules.
+Online and offline KNN-3 is best, 99\% and 97\%.
+Six daily activities.
+  \item ``Recognizing human activities user-independently on smartphones based on accelerometer data'', \cite{siirtola2012recognizing}, cited: none, 2012.
+Classifiers K-nearest neighbors and QDA (quadratic discriminant analysis).
+Online regocnition implemented on phone.
+five daily activities.
+Model training by decision tree (active/inactive).
+Results for QDA, online: 95\%.
+Activities performed outside.
+  \item ``Activity recognition from acceleration data based on discrete consine transform and SVM'', \cite{he2009activity}, cited: 25, 2009.
+Uses Discrete Cosine Transform (DCT) to get features, PCA to reduce them an multiple SVMs to classify windows.
+Four daily activities, 97\% precision.
+  \item ``Physical Activity Recognition Using a Single Tri-Axis Accelerometer'', \cite{lee2178physical}, cited: 16, 2009.
+Five daily activities, FFT, fuzzy c means classification, offline.
+99\% accuracy.
+No smartphone, other device.
 \end{itemize}

### 2013-03-05 Chapters/Chapter1/literature_review.tex
Literature review segmentation and clustering. Sent to Gerard and Anne

--- 2013-03-01 Chapters/Chapter1/literature_review.tex
@@
 
+% \section{Literature review}
+% This section will provide a review of the existing literature and research of the different subjects discussed in this thesis.
+% First the different approaches to mere segmentation of signals is discussed, followed by *** other techniques. ***
+
+% \subsection{Segmentation}
+% One of the first stages in analyzing the sensor data time series is to segment the signal.
+% The goal is to set cut-points in the signal such that all the data points in between consecutive cut-points have some homogeneous property among them.
+*** "organizing data into homogeneous groups where the within-group-object similarity is minimized and the between-group-object dissimilarity is maximized." ***
+% This results in the more abstract notion of change (point) detection.
+% The general concern is to detect the occurrence and time identification of changes in the signal, often expressed as changes in the probability distribution of the time series.
+% This literature review will discuss the different types of change detection as applied in the research.
+% To each applications question about the methods, complexity and usability etc. are asked.
+% On account of usability, it is desired to have a successful segmentation algorithm with minimal user interaction, e.g. defining thresholds and error rates.
+% ** something about what is considered, e.g. complexity, measurements, e.g. ***
+
+% Following the organization of \cite{warren2005clustering}, research can be structured in three groups.
+% The data being processed determines the classification, which can be the raw data, indirectly via features extracted from the raw data or indirectly via models extracted from the raw data.
+
+% One of the more simple and directs methods is to Piecewise Linear Represent the signal, while keeping the approximation error below a certain value, as discussed in \cite{keogh2001online}.
+% As with many algorithms, PLR can be applied to a batch of or real time stream input data (also known as off-line and on-line analyzes, respectively).
+% Keogh et al. \cite{keogh2001online} propose a method which combines these different techniques to form a Sliding Window and Bottom-up (SWAB) analyzes.
+% The algorithm works by first, per window of data points, creating for every two data points a segment.
+% Neighboring segments are joined together, as long as the approximation error is below the provided value.
+% This analyzes can be used as a starting point to further analyze the data segments.
+% Although Keogh et al. apply the SWAB algorithm only direct on the raw data, it is easily imagined how to apply it to extracted features of the signal, e.g. the energy or mean.
+% In its default application, the method yields little information about each segment; it is merely a linear representation.
+% The SWAB algorithm works by segmenting parts of the data series by linear approximation while keeping the error below a certain rate.
+% This rate needs to be user defined (e.g. by an expert), which makes it not possible to apply directly on any signal.
+% This method only divides the signal into segments, which representations can be processed further to form clusters or related segments.
+% The complexity of this method is relatively low, $O(Ln)$, where $L$ is the expected segment length and $n$ the number of data points.
+
+% The task of segmenting any form of data is made easier when the number of segments is beforehand known.
+% Many authors apply the well-known \emph{k-means} clustering *** refs needed *** on this type of problems.
+% An incremental error-minimizing version can determine the optimal number of segments (or actually, clusters in this case).
+% The proposed method of \cite{himberg2001time}, Global Iterative Replacement, resembles the method of Keogh et al., in the sense that it merges segments.
+% The algorithm is applied on two different data sets.
+% The first data sets can be considered as extracted features from sensor data, whilst the second data set works on model properties and thus can give a richer contextual segmentation.
+% For both types of data sets a cost function is defined, which represents the internal heterogeneity of a segment (e.g. the variance of the data within a segment).
+% The difference with the method of Keogh et al. is that with GIR the number of segments is known beforehand and during the algorithm this number is, through splits and merges, kept constant.
+% By running multiple instances of the GIR algorithm with each an increasing number of $k$ segments (bounded by user provided values), the optimal number and location of cut points is determined.
+
+% Other approaches form a higher level analyses of the raw data.
+% As used in \cite{barbivc2004segmenting}, from the field of computer graphics, the dimensionality of the data can be used to distinguish between actions.
+% Using Principal Component Analyses, Barbi{\v{c}} et al.
+use the dimensionality required to project the data points on a hyperplane while keeping the error below a certain rate as a measure of similarity.
+% The error threshold is fixed as a range of three times the standard deviation from the mean of the error.
+% In an extension of their method, using Probabilistic PCA \cite{tipping1999probabilistic}, the Mahanalobis distance \cite{duda1995pattern} for each frame represents similarity of concurrent motions.
+% By traversing the Mahanalobis distance as a function of the frames, cut-points are placed on specific points of interest.
+% *** NOTE to self: thus this method can segment (and group) unknown and new activities? ***
+% In a third method the authors use a Gaussian Mixture Model *** elaborate on this one [does it segment of direct classify?] ***.
+% Using this segmenting, and comparing the segments with earlier gathered representations of motions, the PPCA method gets the best results, with precision of 92\% and recall of 95\%.
+% The test subjects wore 14 sensors on joints of the body and each frame is represented as a 56-dimensional vector.
+% The test considered seven simple human motions: walking, running, sitting down, forward jumping, climbing, arm stretching and punching.
+
+% It is very common the use a higher level of features extracted from the raw data.
+% The approach of Zhou et al. \cite{zhou2008aligned} maps the data points to a higher dimension by using a kernel function.
+% The proposed method, Aligned Cluster Analysis, extends kernel $k$-means clustering on this higher dimensional data set.
+% Using the Dynamic Time Aligned Kernel (DTAK) metric, similarity between segments is measured.
+% Due to the properties of $k$-means clustering, discovering an unknown number of segments in the data set requires and iterative approach.
+
+% *** motifs ***
+
+% *** HMM ***
+
+% *** Finite mixture models ***
+
+% *** Infinite mixture models ***
+
 \section{Literature review}
-This section will provide a review of the existing literature and research of the different subjects discussed in this thesis.
-First the different approaches to mere segmentation of signals is discussed, followed by *** other techniques. ***
+This section will provide a literature review per coherent subject of this thesis.
+In subsection \ref{sec:lit_review_segmentation} the research in segmentation and classification of human activities will be discussed.
 
-\subsection{Segmentation}
-One of the first stages in analyzing the sensor data time series is to segment the signal.
-The goal is to set cut-points in the signal such that all the data points in between consecutive cut-points have some homogeneous property among them.
-*** "organizing data into homogeneous groups where the within-group-object similarity is minimized and the between-group-object dissimilarity is maximized." ***
-This results in the more abstract notion of change (point) detection.
-The general concern is to detect the occurrence and time identification of changes in the signal, often expressed as changes in the probability distribution of the time series.
-This literature review will discuss the different types of change detection as applied in the research.
-To each applications question about the methods, complexity and usability etc. are asked.
-On account of usability, it is desired to have a successful segmentation algorithm with minimal user interaction, e.g. defining thresholds and error rates.
-** something about what is considered, e.g. complexity, measurements, e.g. ***
+\subsection{Segmentation and classification}\label{sec:lit_review_segmentation}
+In recent years much research is done in the field of segmentation and classification of signals into human activities.
+Due to the nature of many applied algorithms and the field of research (computer vision, data mining, robotics), often there is no clear distinction between the segmentation and classification phase.
+As a result, these two subjects will be discussed together although, where possible, the different techniques for the two will be mentioned.
+All the reviewed papers have a global setup in common; a single- or multiple- sensory device is worn by subjects.
+From the raw signals features are extracted which are used to train some models.
+During the testing phase the same extracted features are compared with the trained models to classify the activity.
+Some methods rely on clear segmentations of the signals, whilst other ignore this possible feature and classify on sliding windows.
 
-Following the organization of \cite{warren2005clustering}, research can be structured in three groups.
-The data being processed determines the classification, which can be the raw data, indirectly via features extracted from the raw data or indirectly via models extracted from the raw data.
+During the last ten years much research is performed, as a result of smaller and cheaper measuring devices. *** cite needed ***.
+With the introduction of smartphones with sensory hardware the research has shifted to these devices, as they create a more natural and ubiquitous setting for the test subject. *** cite needed ***.
 
-One of the more simple and directs methods is to Piecewise Linear Represent the signal, while keeping the approximation error below a certain value, as discussed in \cite{keogh2001online}.
-As with many algorithms, PLR can be applied to a batch of or real time stream input data (also known as off-line and on-line analyzes, respectively).
-Keogh et al. \cite{keogh2001online} propose a method which combines these different techniques to form a Sliding Window and Bottom-up (SWAB) analyzes.
-The algorithm works by first, per window of data points, creating for every two data points a segment.
-Neighboring segments are joined together, as long as the approximation error is below the provided value.
-This analyzes can be used as a starting point to further analyze the data segments.
-Although Keogh et al. apply the SWAB algorithm only direct on the raw data, it is easily imagined how to apply it to extracted features of the signal, e.g. the energy or mean.
-In its default application, the method yields little information about each segment; it is merely a linear representation.
-The SWAB algorithm works by segmenting parts of the data series by linear approximation while keeping the error below a certain rate.
-This rate needs to be user defined (e.g. by an expert), which makes it not possible to apply directly on any signal.
-This method only divides the signal into segments, which representations can be processed further to form clusters or related segments.
-The complexity of this method is relatively low, $O(Ln)$, where $L$ is the expected segment length and $n$ the number of data points.
+A distinction can be made in the sensor setup, like wearing multiple sensors, a single sensor and the location of the sensors.
+Research have been performed with multiple sensors, refered to as Body Sensor Networks, as done in \cite{guenterberg2009automatic, guenterberg2009distributed, bao2004activity, sherril2005using}.
+Locations used include the upper arm, waist, thigh, wrist and ankle.
+Single sensors have be worn at different locations, including the upper arm \cite{krause2003unsupervised}, the front leg pocket \cite{kwapisz2011activity, duque2012offline, siirtola2012recognizing, he2009activity}, the waist \cite{ravi2005activity, lester2006practical, lee2178physical}, the chest \cite{ahmed2012non, himberg2001time}, the shoulder \cite{lester2005hybrid} and the dominant wrist \cite{yang2008using, long2009single}.
+Other approaches have not used body-worn sensors at all, but instead relied on other meaures, such as motion capture data \cite{barbivc2004segmenting, zhou2008aligned} and visual \cite{perdikis2008recognition} and acoustic data in office surroundings \cite{oliver2002layered}
 
-The task of segmenting any form of data is made easier when the number of segments is beforehand known.
-Many authors apply the well-known \emph{k-means} clustering *** refs needed *** on this type of problems.
-An incremental error-minimizing version can determine the optimal number of segments (or actually, clusters in this case).
-The proposed method of \cite{himberg2001time}, Global Iterative Replacement, resembles the method of Keogh et al., in the sense that it merges segments.
-The algorithm is applied on two different data sets.
-The first data sets can be considered as extracted features from sensor data, whilst the second data set works on model properties and thus can give a richer contextual segmentation.
-For both types of data sets a cost function is defined, which represents the internal heterogeneity of a segment (e.g. the variance of the data within a segment).
-The difference with the method of Keogh et al. is that with GIR the number of segments is known beforehand and during the algorithm this number is, through splits and merges, kept constant.
-By running multiple instances of the GIR algorithm with each an increasing number of $k$ segments (bounded by user provided values), the optimal number and location of cut points is determined.
+In \cite{bao2004activity} the effectiveness of different body locations for sensors is discussed.
+It is shown that using two locations (the hip and wrist or thigh and wrist) instead of five does not significantly reduce the activity recognition.
+The (dominant) wrist is the preferred location when only a single sensor is used.
 
-Other approaches form a higher level analyses of the raw data.
-As used in \cite{barbivc2004segmenting}, from the field of computer graphics, the dimensionality of the data can be used to distinguish between actions.
-Using Principal Component Analyses, Barbi{\v{c}} et al.
-use the dimensionality required to project the data points on a hyperplane while keeping the error below a certain rate as a measure of similarity.
-The error threshold is fixed as a range of three times the standard deviation from the mean of the error.
-In an extension of their method, using Probabilistic PCA \cite{tipping1999probabilistic}, the Mahanalobis distance \cite{duda1995pattern} for each frame represents similarity of concurrent motions.
-By traversing the Mahanalobis distance as a function of the frames, cut-points are placed on specific points of interest.
-*** NOTE to self: thus this method can segment (and group) unknown and new activities? ***
-In a third method the authors use a Gaussian Mixture Model *** elaborate on this one [does it segment of direct classify?] ***.
-Using this segmenting, and comparing the segments with earlier gathered representations of motions, the PPCA method gets the best results, with precision of 92\% and recall of 95\%.
-The test subjects wore 14 sensors on joints of the body and each frame is represented as a 56-dimensional vector.
-The test considered seven simple human motions: walking, running, sitting down, forward jumping, climbing, arm stretching and punching.
+Some experiments have only been performed in laboratory settings \cite{yang2008using}, thus probably resulting in less robust trained models.
+In \cite{siirtola2012recognizing} the test subjects have performed the testing phase outside, and thus the method is tested in a more natural environment.
+\\ *** Elaborate on lab/outside ***.
 
-It is very common the use a higher level of features extracted from the raw data.
-The approach of Zhou et al. \cite{zhou2008aligned} maps the data points to a higher dimension by using a kernel function.
-The proposed method, Aligned Cluster Analysis, extends kernel $k$-means clustering on this higher dimensional data set.
-Using the Dynamic Time Aligned Kernel (DTAK) metric, similarity between segments is measured.
-Due to the properties of $k$-means clustering, discovering an unknown number of segments in the data set requires and iterative approach.
+A more objective comparison can be made based on the segmentation and classification technique used.
+In \cite{bao2004activity} and \cite{ravi2005activity} extensive comparisons are made between much used algorithms which are widely implemented, such as in the WEKA toolkit \cite{hall2009weka}.
+These include the base-classifiers Decision Trees (also used in \cite{kwapisz2011activity,duque2012offline}), Decision Tables, K-nearest neighbors (also used \cite{duque2012offline, siirtola2012recognizing}), Support Vector Machines (also used in \cite{derawi2010accelerometer,he2009activity, he2008activity}) and Naive Bayes (also used in \cite{long2009single}).
+The research of \cite{bao2004activity} also implemented a number of meta-level classifiers, of which Plurality Voting resulted in the highest accuracy.
+A very well known and applied algorithm from the field of data mining is $k$-means clustering.
+A drawback of this method is that the number of expected clusters needs to be predetermined.
+To overcome this shortcoming an interative implementation can be used, which selects the best clustering.
+A variant is known as Fuzzy $c$-means clustering, in which a data point can belong to one or more clusters, each with a degree of membership.
+The method of $k$-means clustering (or a modified version) is used in \cite{krause2003unsupervised, zhou2008aligned, lee2178physical}.
 
-*** motifs ***
+Very much used in machine learning problems are neural networks. *** cite needed ***
+Although these kind of networks typically perform well on spatial-typed problems *** cite needed ***, adjustments have been made to apply them for temporal classification.
+The methods of \cite{yang2008using, kwapisz2011activity} create a multi-level neural network to generate complex discriminators as activity classifiers.
+A variant of neural networks known as Kohonen Self Organizing Maps is used in \cite{krause2003unsupervised} to create a vector codebook, which is further processed by a $k$-means clustering.
 
-*** HMM ***
+Inspired on the problems for continuous speech recognition have Hidden Markov Models been applied to activity recognition tasks.
+A layered approach proved to be robust for small changes in \cite{perdikis2008recognition}.
+In \cite{shi2009towards} five activities have been modeled as HMMs.
+The method of \cite{fox2010bayesian} introduces HMMs with Markov Jump Linear Systems to segment a signal, with a fixed number of models.
+In \cite{lester2006practical} a layer of HMMs of on top of a layer of static classifiers to estimate the current activity.
+This has a commonality with \cite{lester2005hybrid} in which HMMs are used to capture regularities and smooth activity transitions.
+The application of \cite{guenterberg2009distributed} constructs separate HHMs for each activity and then joins them in a single model to estimate the likelihood of an activity.
 
-*** Finite mixture models ***
+A common property of the above mentioned classification algorithms is that there is no explicit segmentation phase.
+Data points can be segmented a posteriori, by joining adjacent data points which have the same label.
+An approach in which the data points are processed per segment include \cite{minnen2006discovering}, in which known motifs in the sensor signals are discovered.
+In \cite{guenterberg2009automatic} the signal is segmented (without classification) by applying adaptive threshold values on the energy of the signal.
+The estimates of the direct density-ratio for probability distributions are compared to create change-points in \cite{kawahara2009change}.
+By making minimal model assumptions a non-parametric segmentation is obtained, without classification.
+A further application of probability distributions is discussed in \cite{ahmed2012non}, which uses an non-parametric Bayesian method to create an Infinite Gaussian Mixture Model to represent activities.
+Very interesting is the ability to recognize past and new activities without training.
+The method of \cite{barbivc2004segmenting} considers the intrinsic dimensionality of poses by using Probabilistic Principal Component Analysis and creates cut-points.
+This method also is independent of trained models and thus also is a non-parametric approach.
+The method of \cite{keogh2001online} creates a piecewise linear representation of the signal as thus approximates it.
+This does not result in a segmentation in the sense of some homogeneity over the data points, but it can support one.
+In contrast, the method of \cite{himberg2001time} splits and merges segments the signal in a constant number $k$ segments using a cost function to minimize the heterogeity of a segment.
 
-*** Infinite mixture models ***
-
-*****
-List of all papers, shortly categorized
-\begin{itemize}
-  \item ``Discovering characteristic actions from on-body sensor data'' \cite{minnen2006discovering}, cited: 63. Motifs, HHM, DTW, 87\% accuracy. Known motifs.
-  \item ``Recognition of human activities using layered hidden Markov models'' \cite{perdikis2008recognition}, cited: 3.
-HMM, layerd (primitive and abstract actions).
-Uses vision for workplace-activities.
-Direct classification
-  \item ``Accelerometer-Based Gait Analysis, A survey'', \cite{derawi2010accelerometer}, cited: 4.
-Compares methods to distinguish walking of normal, fast, slow.
-Focus on gait, but compares classification methods such as SVM, PCA, KSOM.
-  \item ``An Automatic Segmentation Technique in Body Sensor Networks based on Signal Energy'', \cite{guenterberg2009automatic}, cited: 13.
-Automatic segmenting, adaptive threshold.
-Pure segmenting, no classification.
-Nice and clean method?
-Weakness: single action segmented as multiple.
-Used multiple sensors.
-  \item ``Towards HMM based Human Motion Recognition using MEMS Inertial Sensors'', \cite{shi2009towards}, cited: 13.
-Uses HHM, Fourier transform for features.
-5 fixed activities, trained HMM.
-Correct rates from 90\% to 100\%.
-Classification, not just segmentation
-  \item ``Change-Point Detection in Time-Series Data by Direct Density-Ratio Estimation'', \cite{kawahara2009change}, cited: 41.
-Non-parametric approach, online method, no strong model assumptions.
-Uses multiple datasets, origin referenced.
-Much follow-up work done.
-  \item ``Unsupervised, Dynamic Identification of Physiological and Activity Context in Wearable Computing'', \cite{krause2003unsupervised}, cited: 106(!).
-Combination of KSOM, iterative K-means clustering, transient states removal (using markov model).
-Online algorithm.
-  \item ``Activity Recognition using Cell Phone Accelerometers'', \cite{kwapisz2011activity}, cited: 122 (!).
-\end{itemize}
+*** TABLE *** \\
+*** Create a table of all the papers mentioned? Columns would be:\\
+Refs | single/multiple | body locations | activities | lab/outside | training/testing method | result ***
 
@@
 ***
+
+\input{./Chapters/Chapter1/paper_list}

### 2013-03-05 Chapters/Chapter1/paper_list.tex
Literature review segmentation and clustering. Sent to Gerard and Anne

% !TEX root = ../../main.tex

*****
\subsection{List of papers}
List of all papers, shortly categorized
\begin{itemize}
  \item ``Discovering characteristic actions from on-body sensor data'' \cite{minnen2006discovering}, cited: 63. Motifs, HHM, DTW, 87\% accuracy. Known motifs. Worn on wrist, single.
  \item ``Recognition of human activities using layered hidden Markov models'' \cite{perdikis2008recognition}, cited: 3. HMM, layerd (primitive and abstract actions). Uses vision for workplace-activities. Direct classification. No sensors
  \item ``Accelerometer-Based Gait Analysis, A survey'', \cite{derawi2010accelerometer}, cited: 4. Compares methods to distinguish walking of normal, fast, slow. Focus on gait, but compares classification methods such as SVM, PCA, KSOM.
  \item ``An Automatic Segmentation Technique in Body Sensor Networks based on Signal Energy'', \cite{guenterberg2009automatic}, cited: 13. Automatic segmenting, adaptive threshold. Pure segmenting, no classification. Nice and clean method? Weakness: single action segmented as multiple. Used multiple sensors.
  \item ``Towards HMM based Human Motion Recognition using MEMS Inertial Sensors'', \cite{shi2009towards}, cited: 13. Uses HHM, Fourier transform for features. 5 fixed activities, trained HMM. Correct rates from 90\% to 100\%. Classification, not just segmentation. Single sensor.
  \item ``Change-Point Detection in Time-Series Data by Direct Density-Ratio Estimation'', \cite{kawahara2009change}, cited: 41. Non-parametric approach, online method, no strong model assumptions. Uses multiple datasets, origin referenced. Much follow-up work done. Artificial datasets. (no sensors)
  \item ``Unsupervised, Dynamic Identification of Physiological and Activity Context in Wearable Computing'', \cite{krause2003unsupervised}, cited: 106(!). Combination of KSOM, iterative K-means clustering, transient states removal (using markov model). On-line algorithm. Single sensor on upper arm.
  \item ``Activity Recognition using Cell Phone Accelerometers'', \cite{kwapisz2011activity}, cited: 122 (!), 2011. Uses WEKA: decision trees, logistic regression and multilayer neural networks. Results up to 98\% (jogging). Much difficulty with stairs up and down. Single accelerometer, front leg pocket. Offline method. Many recent references.
  \item ``Activity recognition from accelerometer data'', \cite{ravi2005activity}, cited: 434 (!), 2005. Compares 18 (offline) classifiers, base level (WEKA) and meta-level. Single device, pelvic region.
  \item ``Activity recognition from user-annotated acceleration data'', \cite{bao2004activity}, cited: 1008 (!!), 2004. Five bi-axial sensors. Decision tables, instance-based learning, C4.5 (trees, highest accuracy 89.30\%) and naive bayes (weka), twenty activities.
  \item ``Using Hierarchical Clustering Methods to Classify Motor Activities of COPD Patients from Wearable Sensor Data'', \cite{sherril2005using}, cited: none, 2005. Uses Linear Discriminant Analysis for cluster classification (using Cluster Quality Index to determine k) and simple rule-based separation for high level ambulatory. Hierarchical Dendogram to merge clusters when similarity to high. Sensors on forearms and legs.
  \item ``Non-Parametric Bayesian Human Motion Recognition Using a Single MEMS Tri-Axial Accelerometer'', \cite{ahmed2012non}, cited: none, 2012. Recognizes the number of human activities, single sensor on the chest. No training data. Infinite Gaussian Mixture Model, collapsed Gibbs sampler. Compares with parametric Fuzzy C-mean (data point belongs to multiple clusters), unsupervised K-means, non-parametric mean-shift. Outperforms all significantly, high hit rate and low false alarms.
  \item ``An Online Algorithm for Segmenting Time Series'', \cite{keogh2001online}, cited: 487, 2001. Reviews algorithms to get a piecewise linear representation, proposes a novel sliding-window and bottom up approach, on-line. Mere segmentation. No classification; only approximation of signal. (Dataset available).
  \item ``Segmenting Motion Capture Data into Distinct Behaviors'', \cite{barbivc2004segmenting}, cited: 242, 2004. On-line methods: cut-points on increased intrinsic dimensionality, distribution of poses is observed to change. Batch: cut when consecutive frames belong to different Gaussian mixture models. Fixed fourteen motions, compared with manual segmentation. Uses PCA and Probabilistic PCA (models the non-pc subspaces as noise). PPCA is the best. (paper anne)
  \item ``Aligned Cluster Analysis for Temporal Segmentation of Human Motion'', \cite{zhou2008aligned}, cited: 54, 2008. Uses extension on kernel k-means clustering and Dynamic Time Alignment Kernel (kernel of DTW) for temporal invariance, robust temporal matching metric. Coordinate descent algorithm solves ACA. Fixed number of clusters, trapped in local minima. (paper anne)
  \item ``Bayesian Nonparametric Methods for Learning Markov Switching Processes'', \cite{fox2010bayesian}, cited: 10, 2010. Uses HMM for state-space model for segmentation, Markov Jump Linear systems. Unbounded number of Markov modes (parameters). Number of models is fixed, though? (paper anne). No sensors.
  \item ``Layered Representations for Human Activity Recognition'', \cite{oliver2002layered}, cited: 215, 2002. Layers of Hidden Markov Models HMM, multiple levels of granularity (Based on intuition) and context. Less retraining, only lower models. Classified also. No sensors, visual and akoustic data in office surroundings.
  \item ``Time series segmentation for context recognition in mobile devices'', \cite{himberg2001time}, cited: 151, 2001. Splits and merges segments, while keeping k constant using cost function on segments for internal heterogeneity. Multiple instances determine number of k. Microphone and accelerometer data, in front of chest.
  \item ``A practical approach to recognizing physical activities'', \cite{lester2006practical}, cited: 259, 2006. Uses method of \cite{lester2005hybrid}, activity classification algorithm. Selects most useful features and then recognizes walking, sitting, etc. First layer static classifier on features or data (energy, mean, variance, correlation, etc), then a layer of HMM to estimate activity. All offline. Shows trained model is robust to location of wearing on body.
  \item ``A hybrid discriminative/generative approach for modeling human activities'', \cite{lester2005hybrid}, cited: 252, 2005. Used in method above. Combines boosting to select and reduce useful features and learn static classifiers with HMM to capture regularities and smooth activities (quick switching is unlikely). Single sensor, multiple measures (accelerometer, audio, etc), worn on the shoulder.
  \item ``Using acceleration measurements for activity recognition: An effective learning algorithm for constructing neural classifiers'', \cite{yang2008using}, cited: 67, 2008. Separates dynamic from static activities. Uses multilayer feedforward neural networks to generate complex discriminating surfaces as activity classifiers. Feature subset selection approach is developed. Neural pre-classifier with constant threshold criterion. Uses Common Principal Component Analysis. Recognizes eight activities with 95\% overall accuracy. Laboratory environment. Worn on dominant wrist.
  \item ``Single-accelerometer-based daily physical activity classification'', \cite{long2009single}, cited: 46, 2009. Uses Naive Bayes classifier, sensor worn on wrist. Five activities. PCA to reduce dimensions and create independence for NB. Comparable results as Decision Trees, but with flexibility to add activities.
  \item ``Distributed Continuous Action Recognition Using a Hidden Markov Model in Body Sensor Networks'', \cite{guenterberg2009distributed} cited: 17, 2009. Single HMM, sensor network cluster movements, HMM constructs continuous actions using postures and actions, merges left-to-right HMM for each action to a single. Transcript generation per sensor uses Gaussian Mixed Model, multiple models with $m$ mixtures are trained with Expectation-Maximization (EM), best is chosen.
  \item ``Offline and online activity recognition on mobile devices using accelerometer data'', \cite{duque2012offline}, cited: none, 2012. Compares offline and online methods, cellphone in pocket. Reduces feature extraction (mean and standard deviation (for resources of cellphone). Training is done by database and labeling, classification with K-nearest neighbors, decision trees (C4.5) and decision rules. Online and offline KNN-3 is best, 99\% and 97\%. Six daily activities. Trousers pocket.
  \item ``Recognizing human activities user-independently on smartphones based on accelerometer data'', \cite{siirtola2012recognizing}, cited: none, 2012. Classifiers K-nearest neighbors and QDA (quadratic discriminant analysis). Online regocnition implemented on phone. five daily activities. Model training by decision tree (active/inactive). Results for QDA, online: 95\%. Activities performed outside. Trousers pocket.
  \item ``Activity recognition from acceleration data based on discrete consine transform and SVM'', \cite{he2009activity}, cited: 25, 2009. Uses Discrete Cosine Transform (DCT) to get features, PCA to reduce them an multiple SVMs to classify windows. Four daily activities, 97\% precision. Trousers pocket.
  \item ``Physical Activity Recognition Using a Single Tri-Axis Accelerometer'', \cite{lee2178physical}, cited: 16, 2009. Five daily activities, FFT, fuzzy c means classification, offline. 99\% accuracy. No smartphone, other device.
\end{itemize}

### 2013-03-08 Chapters/Chapter1/literature_review.tex
Add laboratory / outside settings

@@
 
-Some experiments have only been performed in laboratory settings \cite{yang2008using}, thus probably resulting in less robust trained models.
-In \cite{siirtola2012recognizing} the test subjects have performed the testing phase outside, and thus the method is tested in a more natural environment.
-\\ *** Elaborate on lab/outside ***.
+Some experiments have only been performed in an artificial or laboratory setting \cite{minnen2006discovering, yang2008using, perdikis2008recognition, guenterberg2009automatic, ravi2005activity, bao2004activity, sherril2005using, he2009activity}, thus probably resulting in less robust trained models.
+In \cite{siirtola2012recognizing, kwapisz2011activity} the experiments are performed in a uncontrolled environment.
+This results in a more natural setting with more noise due to free movement.
+It also introduces transition-states in the data, because there is no clear distinction between activities.
+The method of \cite{krause2003unsupervised} explicitly filters the signals for transitions.
+Others have provided the test subjects with a measuring device for a longer period of time, some up to a few days \cite{krause2003unsupervised, long2009single}.
 
@@
 A drawback of this method is that the number of expected clusters needs to be predetermined.
-To overcome this shortcoming an interative implementation can be used, which selects the best clustering.
+To overcome this shortcoming an iterative implementation can be used, which selects the best clustering.
 A variant is known as Fuzzy $c$-means clustering, in which a data point can belong to one or more clusters, each with a degree of membership.
@@
 
-Inspired on the problems for continuous speech recognition have Hidden Markov Models been applied to activity recognition tasks.
+Due similarity with the problem of continuous speech recognition, Hidden Markov Models have been applied to activity recognition tasks.
 A layered approach proved to be robust for small changes in \cite{perdikis2008recognition}.
@@
 Data points can be segmented a posteriori, by joining adjacent data points which have the same label.
-An approach in which the data points are processed per segment include \cite{minnen2006discovering}, in which known motifs in the sensor signals are discovered.
+An approach in which the data points are processed per segment include \cite{minnen2006discovering}, in which motifs over sensor data from actions are discovered.
 In \cite{guenterberg2009automatic} the signal is segmented (without classification) by applying adaptive threshold values on the energy of the signal.
@@
 This does not result in a segmentation in the sense of some homogeneity over the data points, but it can support one.
-In contrast, the method of \cite{himberg2001time} splits and merges segments the signal in a constant number $k$ segments using a cost function to minimize the heterogeity of a segment.
+In contrast, the method of \cite{himberg2001time} splits and merges segments the signal in a constant number $k$ segments using a cost function to minimize the heterogeneity of a segment.
 

### 2013-03-11 Chapters/Chapter1/literature_review.tex
Finish literature review about segmentation and classification. Beginning of Table of Papers

@@
 Some methods rely on clear segmentations of the signals, whilst other ignore this possible feature and classify on sliding windows.
+The content of this section is summarized in table \ref{table:papers_segmentation_classification}.
 
@@
 
-*** TABLE *** \\
-*** Create a table of all the papers mentioned? Columns would be:\\
-Refs | single/multiple | body locations | activities | lab/outside | training/testing method | result ***
+Each research outcome is defined in some kind of correctness to segment and classify the signal in the correct and preferred manner.
+In many methods an error was measured as the difference between the algorithms outcome and a domain-experts ground truth.
+In case of segmenting the time series, the method of \cite{guenterberg2009automatic} measures the delayed number of seconds before a new segment is created.
+For classification the measure takes on many forms and names, such as recognition rates \cite{bao2004activity} or hit precision \cite{ahmed2012non}.
+Confusion matrices often are used to give insight in which activities are hard to discriminate from each other, as for instance in \cite{kwapisz2011activity}.
+The term false alarm rate is often used to indicate incorrectly new introduced elements, such as activities in \cite{ahmed2012non} and \cite{kawahara2009change}.
 
-*** Elaborate on measures, e.g. precision, recall, confusion matrix etc.  ***
-*** Precision is defined as the ratio of reported correct cuts versus the total number of reported cuts.
-Recall is defined as the ratio of reported correct cuts versus the total number of correct cuts.
-The closer precision and recall are to 1, the more accurate the algorithm is.
+The overall results of many methods are fairly high and promising.
+Many methods, for a fixed number of activities or an iterative approach, obtain precision up to the range from 95\% to 100\% \cite{minnen2006discovering, shi2009towards, kwapisz2011activity, duque2012offline, he2009activity, lee2178physical, siirtola2012recognizing}.
+Problems for which the number of activities is unknown perform up to 95\% \cite{ahmed2012non, barbivc2004segmenting}.
+The most problems arise from discriminating between similar activities, such as taking the stairs up or down and walking when the movement of legs or waist is considered \cite{kwapisz2011activity, duque2012offline}.
+
+Some researchers have made their used data set explicitly public available and other data sets are just available, providing valuable labeled activity patterns.
+*** This should be in experiments setup?
 ***
+*** List here the datasets used ***
+
+\begin{table}
+\tiny
+\centering
+\begin{tabular}{ | l | c | p{1.5cm} | p{1.5cm} | p{1.5cm} | p{2.5cm} | p{3cm} | }
+  \hline
+  Refs. & \# & Location & Activities & Env. & Algorithms & Results \\
+  \hline
+  \cite{minnen2006discovering} & Single & Wrist & Signal motifs & Artificial data & Motif discovery, HHMs, DTW & overall 87\% \\
+  \hline
+  \cite{guenterberg2009automatic} & Multiple & wrist, legs, waist & 12 daily routine & Controlled & Adaptive threshold, segmenting & 85\% \\
+  \hline
+  \cite{shi2009towards} & Single & Unknown & walking, upstairs, falling, running, standing & Controlled & FFT, HMMs & 90\% - 100\% \\
+  \hline
+  \cite{kawahara2009change} & - & - & - & Artificial data & Non-parametric Density-Ratio estimation & ??? \\
+  \hline
+\end{tabular}
+\caption{Overview of researches with the focus on segmentation and classification techniques.}
+\label{table:papers_segmentation_classification}
+\end{table}
 

### 2013-03-11 Chapters/Chapter1/paper_list.tex
Add laboratory / outside settings
Finish literature review about segmentation and classification. Beginning of Table of Papers
Minor changes lit review

@@
 \subsection{List of papers}
-List of all papers, shortly categorized
+The following is a list of used papers.
+This list is for my own personal convenience.
+It is possible this will be in tabular form in the final thesis document.
 \begin{itemize}
-  \item ``Discovering characteristic actions from on-body sensor data'' \cite{minnen2006discovering}, cited: 63. Motifs, HHM, DTW, 87\% accuracy. Known motifs. Worn on wrist, single.
+  \item ``Discovering characteristic actions from on-body sensor data'' \cite{minnen2006discovering}, cited: 63.
+Motifs, HHM, DTW, 96\%, overall 87\% accuracy.
+Discovers motif (actions) from data.
+Worn on wrist, single.
+Environment artificial.
   \item ``Recognition of human activities using layered hidden Markov models'' \cite{perdikis2008recognition}, cited: 3.
-HMM, layerd (primitive and abstract actions).
+HMM, layered (primitive and abstract actions).
 Uses vision for workplace-activities.
 Direct classification.
-No sensors
+No sensors.
+Artificial environment for region labeling.
   \item ``Accelerometer-Based Gait Analysis, A survey'', \cite{derawi2010accelerometer}, cited: 4.
@@
   \item ``Unsupervised, Dynamic Identification of Physiological and Activity Context in Wearable Computing'', \cite{krause2003unsupervised}, cited: 106(!).
-Combination of KSOM, iterative K-means clustering, transient states removal (using markov model).
+Combination of KSOM, iterative K-means clustering, transient states removal (using Markov model).
 On-line algorithm.
 Single sensor on upper arm.
+Tested for a few days and controlled environment.
   \item ``Activity Recognition using Cell Phone Accelerometers'', \cite{kwapisz2011activity}, cited: 122 (!), 2011.
@@
 Many recent references.
+Controlled environment.
   \item ``Activity recognition from accelerometer data'', \cite{ravi2005activity}, cited: 434 (!), 2005.
@@
 Single device, pelvic region.
+Controlled setting.
   \item ``Activity recognition from user-annotated acceleration data'', \cite{bao2004activity}, cited: 1008 (!!), 2004.
 Five bi-axial sensors.
-Decision tables, instance-based learning, C4.5 (trees, highest accuracy 89.30\%) and naive bayes (weka), twenty activities.
+Decision tables, instance-based learning, C4.5 (trees, highest accuracy 89.30\%) and naive Bayes (weka), twenty activities.
+Controlled environment, common room.
   \item ``Using Hierarchical Clustering Methods to Classify Motor Activities of COPD Patients from Wearable Sensor Data'', \cite{sherril2005using}, cited: none, 2005.
@@
 Sensors on forearms and legs.
+Controlled clinical environment.
   \item ``Non-Parametric Bayesian Human Motion Recognition Using a Single MEMS Tri-Axial Accelerometer'', \cite{ahmed2012non}, cited: none, 2012.
@@
 Fixed number of clusters, trapped in local minima.
-(paper anne)
+(paper anne).
+Uses motion capture data.
   \item ``Bayesian Nonparametric Methods for Learning Markov Switching Processes'', \cite{fox2010bayesian}, cited: 10, 2010.
@@
 Microphone and accelerometer data, in front of chest.
+ARtificial data
   \item ``A practical approach to recognizing physical activities'', \cite{lester2006practical}, cited: 259, 2006.
@@
 Comparable results as Decision Trees, but with flexibility to add activities.
+Performed outside laboratory, 10 hours long.
   \item ``Distributed Continuous Action Recognition Using a Hidden Markov Model in Body Sensor Networks'', \cite{guenterberg2009distributed} cited: 17, 2009.
@@
 Classifiers K-nearest neighbors and QDA (quadratic discriminant analysis).
-Online regocnition implemented on phone.
+Online recognition implemented on phone.
 five daily activities.
@@
 Trousers pocket.
-  \item ``Activity recognition from acceleration data based on discrete consine transform and SVM'', \cite{he2009activity}, cited: 25, 2009.
+  \item ``Activity recognition from acceleration data based on discrete cosine transform and SVM'', \cite{he2009activity}, cited: 25, 2009.
 Uses Discrete Cosine Transform (DCT) to get features, PCA to reduce them an multiple SVMs to classify windows.
@@
 Trousers pocket.
+Laboratory setting.
   \item ``Physical Activity Recognition Using a Single Tri-Axis Accelerometer'', \cite{lee2178physical}, cited: 16, 2009.
@@
 No smartphone, other device.
+Outside, although restricted (standing, sitting, lying, walking, running)
 \end{itemize}

### 2013-03-14 Chapters/Chapter1.tex
Work on lit review non parametric clustering

@@
 % Chapter 1
+% !TEX root = ../main.tex
 
@@
 
-\input{./Chapters/Chapter1/literature_review}
+\section{Literature review}
+This section will provide a literature review per coherent subject of this thesis.
+In subsection \ref{sec:lit_review_segmentation} the research in segmentation and classification of human activities will be discussed.
+Subsection \ref{sec:lit_review_nonparametric} will take a look on applied non-parametric methods in clustering (time series) of data.
+
+\input{./Chapters/Chapter1/literature_review_segmentation_and_classification}
+
+\input{./Chapters/Chapter1/literature_review_nonparametric_clustering}
+
+\input{./Chapters/Chapter1/paper_list}
 

### 2013-03-20 Chapters/Chapter1/paper_list.tex
Temporal segmentation

@@
 
-*****
 \subsection{List of papers}
@@
 \end{itemize}
+
+\subsubsection{Pure segmentation}
+\begin{itemize}
+  \item ``Bayesian Online Changepoint Detection'', \cite{adams2007bayesian}, cited: 67, 2007.
+Uses a priori and posteriori distribution prediction of the current run-length.
+Online.
+Uses well-drilling data, Dow Jones return value and coal-mine accidents rate, which are all used by others.
+Does not compare results.
+Decouples algorithm from model (?).
+  \item CUSUM Method: piece wise segments of gaussian mean with noise. More refs needed.
+  \item ``An Automatic Segmentation Technique in Body Sensor Networks based on Signal Energy'', \cite{guenterberg2009automatic}, cited: 13.
+Automatic segmenting, adaptive threshold.
+Pure segmenting, no classification.
+Nice and clean method?
+Weakness: single action segmented as multiple.
+Used multiple sensors.
+  \item ``Segmenting Motion Capture Data into Distinct Behaviors'', \cite{barbivc2004segmenting}, cited: 242, 2004.
+On-line methods: cut-points on increased intrinsic dimensionality, distribution of poses is observed to change.
+Batch: cut when consecutive frames belong to different Gaussian mixture models.
+Fixed fourteen motions, compared with manual segmentation.
+Uses PCA and Probabilistic PCA (models the non-pc subspaces as noise).
+PPCA is the best.
+(paper anne)
+\end{itemize}

### 2013-03-21 Chapters/Chapter1/paper_list.tex
Work of wednesday 20 march

@@
 Microphone and accelerometer data, in front of chest.
-ARtificial data
+Artificial data
   \item ``A practical approach to recognizing physical activities'', \cite{lester2006practical}, cited: 259, 2006.
@@
   \item ``An Automatic Segmentation Technique in Body Sensor Networks based on Signal Energy'', \cite{guenterberg2009automatic}, cited: 13.
-Automatic segmenting, adaptive threshold.
+Automatic segmenting, adaptive threshold, standard deviation, niet bayesian.
 Pure segmenting, no classification.
@@
 PPCA is the best.
-(paper anne)
+(paper anne).
+From computer graphics / motion capture data.
+  \item ``Activity recognition using inertial sensing for healthcare, wellbeing and sports applications: A survey'', \cite{avci2010activity}, cited: 22, 2010.
+Discusses segmentation as form of sliding window, top-down, bottom-up or swab.
+Interesting overview of five steps: preprocessing, segmentation, feature extraction, dimensionality reduction and classifiers.
+Classifiers only with fixed number of activities (no non-parametric).
+  \item ``Event Detection From Time Series Data'', \cite{guralnik1999event}, cited: 291, 1999.
+Iterative (online, incremental) method to fit model to segment, likelihood to determine if partitioned further (number of change points), when batch.
+Maximum likelihood methods.
+Also model selection, loss and risk functions.
+With incremental (online) method, every new data point had likelihood criteria of being a new segment or current one.
+If no segment for long time, computations become increasingly expensive.
+Solution: sliding window for last $w$ points only.
+  \item ``An adaptive approach for online segmentation of multi-dimensional mobile data'', \cite{guo2012adaptive}, cited: 1, 2012.
+Adaptive model for selecting suitable dimensions to build model and select model (linear, quadratic, cubic).
+Multiple metric for results evaluation; traditional regression error, information retrieval (precision, recall, F-measure) and segmentation delay.
+Uses three-axis accelerometer data.
+Extends the SWAB method.
+Natural setting, no restriction on body location.
+  \item ``Activity Recognition on 3D Accelerometer Data (Technical Report)'', \cite{bernecker2012activity}, cited: 0, 2012.
+First makes distinction between periodic and non-periodic activities, directly on accelerometer data.
+Autocorrelation as a measure for self-similarity (periodicy).
+After segmentation, features are extracted for clustering.
+Then dimensionality reduction and (re)classification.
+Interesting post-processing; small segments with neighbors of the same activity are merged.
+Removes small mis-activities.
+  \item ``Change-Point Detection in Time-Series Data by Direct Density-Ratio Estimation'', \cite{kawahara2009change}, cited: 41.
+Non-parametric approach, online method, no strong model assumptions.
+Uses multiple datasets, origin referenced.
+Much follow-up work done.
+Artificial datasets.
+(no sensors) Uses likelihood/density ratios estimator, not the densities themselves.
+Probability distributions that generate the time series.
+  \item ``Adaptive Filtering and Change Detection'' (book), \cite{gustafsson2000adaptive}, cited: 682 (!!), 2000. Lots of algorithms, hard to read.
+  \item ``Online segmentation of time series based on polynomial least-squares approximations'', \cite{fuchs2010online}, cited: 17, 2010.
+Creates higher-order polynomial model of data, sliding and growing window.
+Works well for traditional time series (stocks), but not for non-stationary such as daily life.
+  \item ``Novel online methods for time series segmentation'', \cite{liu2008novel}, cited: 17, 2008.
+(Stepwise) Feasable Space Window.
+Compares with SWAB.
+Referenced in \cite{guo2012adaptive}, but that one choses SWAB because it should be lower time complexity and fast.
+(S)FSW creates less segments, while SWAB has lower error.
+FSW is fastest and reliable (due to this paper self).
+  \item ``An adaptive cusum-based test for signal change detection'', \cite{alippi2006adaptive} cited: 16, 2006.
+CUSUM-test with adaptive parameter settings.
+Can detect non-stationary processes by drift of signal.
+Applicable for extracted features, e.g.
+mean.
 \end{itemize}

### 2013-03-21 Chapters/Chapter1.tex
Temporal segmentation
Outline of temporal segmentation

--- 2013-02-28 Chapters/Chapter1.tex
@@
+% !TEX root = ../main.tex
 % Chapter 1
-
-\chapter{Introduction} % Main chapter title
-
-\label{Chapter1} % For referencing the chapter elsewhere, use \ref{Chapter1}
-
-\lhead{Chapter 1. \emph{Introduction}} % This is for the header on each page -
-%perhaps a shortened title
-
 %----------------------------------------------------------------------------------------
 
+%perhaps a shortened title
+*** [GRAPHICS diagram: sensors --> signal pre-processing --> feature extraction --> analysis [* segmentation, * labeling] --> conclusions]
+*** Give context problems, means to reach (e.g. monitoring of patients,  getting conclusions over health and activity).
+\chapter{Introduction} % Main chapter title
+\label{Chapter1} % For referencing the chapter elsewhere, use \ref{Chapter1}
+\lhead{Chapter 1. \emph{Introduction}} % This is for the header on each page -
+
+\section{Literature review}
+This section will provide a literature review per coherent subject of this thesis.
+
+% \input{./Chapters/Chapter1/literature_review_nonparametric_clustering}
+\input{./Chapters/Chapter1/lit_rev_segm_and_class}
+\input{./Chapters/Chapter1/lit_rev_temp_segm}
+\input{./Chapters/Chapter1/paper_list}
+
 \section{Problem statement}
-*** Give context problems, means to reach (e.g. monitoring of patients,  getting conclusions over health and activity).
-%----------------------------------------------------------------------------------------
 
-\input{./Chapters/Chapter1/literature_review}
-
-%----------------------------------------------------------------------------------------
-
-\section{System architecture}
-
-*** [GRAPHICS diagram: sensors --> signal pre-processing --> feature extraction --> analysis [* segmentation, * labeling] --> conclusions]
-
-%----------------------------------------------------------------------------------------
+% \section{System architecture}
+% In subsection \ref{sec:lit_review_segmentation} the research in segmentation and classification of human activities will be discussed.
+% Subsection \ref{sec:lit_review_nonparametric} will take a look on applied non-parametric methods in clustering (time series) of data.

### 2013-03-21 Chapters/Chapter1/lit_rev_temp_segm.tex
Outline of temporal segmentation

% !TEX root = ../../main.tex
% Section Literature Review Segmentation

\subsection{Temporal Segmentation}\label{sec:lit_review_temporal_segmentation}
Many different approaches have been applied to segmenting time series data.
Due to the nature of this research, only segmenting techniques with an on-line approach (or easy translation to one ***?***) will be considered.
Interesting are algorithms which have been applied to multi-dimensional data, since we are considering segmenting accelerometer data provided by three-axial sensors.
Nonetheless, online segmenting algorithms on single dimensional data will also be considered.

*** Piecewise linear approximation ***

Segmentation methods can roughly be categorized in three methods in the way the data is processed, as discussed by Avci et al \cite{avci2010activity}:
\begin{itemize}
  \item \textbf{Top-down} methods iteratively divide the signal in segments by splitting at the best location.
  The algorithm starts with two segments and completes when a certain condition is met, such as when an error value or number of segments $k$ is reached.
  These methods process the data points recursively, which results in a complexity of $O(n^2k)$.
  \item \textbf{Bottom-up} methods are the natural complement to top-down methods.
  They start with creating $n/2$ segments and iteratively join adjacent segments while the value of a cost function for that operation is below a certain value.
  Given the average segment length $L$, the complexity of this method is $O(nL)$.
  \item \textbf{Sliding-window} methods are simple and intuitive for online segmenting purposes.
  It starts with a small initial subsequence of the time series.
  New data points are joined in the segment until the fit-error is above a threshold.
  Because of the simple approax, the data is only processed very locally which can yield in poor results \cite{keogh2001online}.
  The complexity is equal to the bottom-up approach, $O(nL)$, where $L$ is the average segment length.
  \item \textbf{Sliding Window And Bottom-up} (SWAB), as introduced by Keogh et al. \cite{keogh2001online}, joins the ability of the sliding window mechanism to process time series online and the bottom-up approach the create superior segments in terms of fit-error.
  The algorithm processes the data in two stages.
  The first stage is to join new data points in the current segment created by a sliding window, and pass this to a buffer with space for a few segments.
  The buffer then processes the data Bottom-up and returns the first (left-most) segment as final segment.
  Because this second stage retains some (semi-)global view of the data, the results are comparative with normal Bottom-up.
  It is stated by Keogh et al. that the complexity of SWAB is a small constant factor worse than that of regular Bottom-up.
\end{itemize}
It is clear that for the application of this research Sliding-window and preferably SWAB-based algorithms should be considered.

The SWAB method proposed by Keogh et al. \cite{keogh2001online} is dependent on an user setting, providing the maximum error when performing both stages.
Each segment is approximated by using Piecewise Linear Representation, an often used method.
The user provided error threshold controls the granularity and number of segments.
Other methods have been proposed, such as an adaptive threshold based on the signal energy by Guenterberg et al. \cite{guenterberg2009automatic} and an adaptive CUSUM-based test by Alippi et al. \cite{alippi2006adaptive}, in order to eliminate this user-dependency.
The latter of these methods is able to process the accelerometer values directly, although better results are obtained when features of the signal are processed, as done in the former method.
Here the signal energy, mean and standard deviation are used to segment activities and by adding all the axial time series together, the Signal-To-Noise ration is increased, resulting in a robuster method.

The method of Guenterberg at al. extracts features from the raw sensor signal to base the segmentation on other properties than the pure values.
The method of Bernecker et al. \cite{bernecker2012activity} uses other statistical properties, namely autocorrelation, to distinguish periodic from non-periodic segments.
Using the SWAB method the self-similarity of a one-dimensional signal is obtained.
The authors claim that only a slight modification is needed to perform the method on multi-dimensional data.
After the segmentation phase, the method of Bernecker et al. extracts other statistical features which are used in the classification phase.

The proposal of Guo et al. \cite{guo2012adaptive} dynamically determines which features should be used for the segmentation and simultaneously determines the best model to fit the segment.
For each of the three dimensions features such as the mean, variance, covariance, correlation, energy and entropy are calculated.
By extending the SWAB method, for every frame a feature set is selected, using an enhanced version of Principal Component Analysis (PCA).
Whereas the before mentioned algorithms use a linear representation, this methods considers linear, quadratic and cubical representations for each segment.
This differs from other methods where the model is fixed for the whole time series, such as \cite{fuchs2010online}, which is stated to perform inferior on non-stationary time series such as daily life.

The time series data from a sensor can be considered as a stochastic process *** is this a correct term? ***.
Probabilistic models can be constructed on that signal, yielding in probabilistic and Bayesian based segmentation methods.
The CUSUM-method relies on the log-likelihood ratio to measure the difference between two distributions.
To calculate the ratio, the probability density functions need to be calculated.
The method of Kawahara et al. \cite{kawahara2009change} proposes to estimate the ratio of probability densities (known as the \emph{importance}) directly, without explicit estimation of the densities.
They claim this results in a robuster approach for real-world scenarios.
Although this is a model-based method, no strong assumptions (parameter settings) are made on the models.

An other application of PCA is to characterize the data by determining the dimensionality of a sequence of data points.
The proposed method of Berbi\v{c} et al. \cite{barbivc2004segmenting} determines the number of dimensions (features) needed to approximate a sequence within a specified error.
With the observation that more dimensions are needed to keep the error below the threshold when transitions between actions occur, cut-points can be located and segments will be created.
The superior extension of their approach uses a Probabilistic PCA algorithm to incorporate the dimensions outside the selected set as noise.

The method of Adams and MacKay \cite{adams2007bayesian} builds a probabilistic model on the segment run length, given the observed data so far.
Instead of modeling the values of the data points as a probabilistic distribution, the length of segments as a function if time is modeled by calculating the posterior probability.
It uses a prior estimate for the run length and a predictive distribution for newly-observed data, given the data since the last change point.
This method contrasts with the approach of Guralnik and Srivastava \cite{guralnik1999event} in which change points are detected by a change in the (parameters of an) underlying, observed, model.
For each new data point, the likelihoods of being a change point and part of the current segment are calculated, without a prior model (and thus is a non-Bayesian approach).
It is observed that when no change point is detected for a long period of time, the computational complexity increases significantly.

### 2013-03-22 Chapters/Chapter1/lit_rev_temp_segm.tex
Add marco for et al. abbrevation

@@
 
-Segmentation methods can roughly be categorized in three methods in the way the data is processed, as discussed by Avci et al \cite{avci2010activity}:
+Segmentation methods can roughly be categorized in three methods in the way the data is processed, as discussed by Avci \etal \cite{avci2010activity}:
 \begin{itemize}
@@
   The complexity is equal to the bottom-up approach, $O(nL)$, where $L$ is the average segment length.
-  \item \textbf{Sliding Window And Bottom-up} (SWAB), as introduced by Keogh et al.
-\cite{keogh2001online}, joins the ability of the sliding window mechanism to process time series online and the bottom-up approach the create superior segments in terms of fit-error.
+  \item \textbf{Sliding Window And Bottom-up} (SWAB), as introduced by Keogh \etal \cite{keogh2001online}, joins the ability of the sliding window mechanism to process time series online and the bottom-up approach the create superior segments in terms of fit-error.
   The algorithm processes the data in two stages.
@@
   Because this second stage retains some (semi-)global view of the data, the results are comparative with normal Bottom-up.
-  It is stated by Keogh et al. that the complexity of SWAB is a small constant factor worse than that of regular Bottom-up.
+  It is stated by Keogh \etal that the complexity of SWAB is a small constant factor worse than that of regular Bottom-up.
 \end{itemize}
@@
 
-The SWAB method proposed by Keogh et al. \cite{keogh2001online} is dependent on an user setting, providing the maximum error when performing both stages.
+The SWAB method proposed by Keogh \etal \cite{keogh2001online} is dependent on an user setting, providing the maximum error when performing both stages.
 Each segment is approximated by using Piecewise Linear Representation, an often used method.
 The user provided error threshold controls the granularity and number of segments.
-Other methods have been proposed, such as an adaptive threshold based on the signal energy by Guenterberg et al.
-\cite{guenterberg2009automatic} and an adaptive CUSUM-based test by Alippi et al.
-\cite{alippi2006adaptive}, in order to eliminate this user-dependency.
+Other methods have been proposed, such as an adaptive threshold based on the signal energy by Guenterberg \etal \cite{guenterberg2009automatic} and an adaptive CUSUM-based test by Alippi \etal \cite{alippi2006adaptive}, in order to eliminate this user-dependency.
 The latter of these methods is able to process the accelerometer values directly, although better results are obtained when features of the signal are processed, as done in the former method.
@@
 
-The method of Guenterberg at al. extracts features from the raw sensor signal to base the segmentation on other properties than the pure values.
-The method of Bernecker et al. \cite{bernecker2012activity} uses other statistical properties, namely autocorrelation, to distinguish periodic from non-periodic segments.
+The method of Guenterberg \etal extracts features from the raw sensor signal to base the segmentation on other properties than the pure values.
+The method of Bernecker \etal \cite{bernecker2012activity} uses other statistical properties, namely autocorrelation, to distinguish periodic from non-periodic segments.
 Using the SWAB method the self-similarity of a one-dimensional signal is obtained.
 The authors claim that only a slight modification is needed to perform the method on multi-dimensional data.
-After the segmentation phase, the method of Bernecker et al. extracts other statistical features which are used in the classification phase.
+After the segmentation phase, the method of Bernecker \etal extracts other statistical features which are used in the classification phase.
 
-The proposal of Guo et al. \cite{guo2012adaptive} dynamically determines which features should be used for the segmentation and simultaneously determines the best model to fit the segment.
+The proposal of Guo \etal \cite{guo2012adaptive} dynamically determines which features should be used for the segmentation and simultaneously determines the best model to fit the segment.
 For each of the three dimensions features such as the mean, variance, covariance, correlation, energy and entropy are calculated.
@@
 To calculate the ratio, the probability density functions need to be calculated.
-The method of Kawahara et al. \cite{kawahara2009change} proposes to estimate the ratio of probability densities (known as the \emph{importance}) directly, without explicit estimation of the densities.
+The method of Kawahara \etal \cite{kawahara2009change} proposes to estimate the ratio of probability densities (known as the \emph{importance}) directly, without explicit estimation of the densities.
 They claim this results in a robuster approach for real-world scenarios.
@@
 An other application of PCA is to characterize the data by determining the dimensionality of a sequence of data points.
-The proposed method of Berbi\v{c} et al. \cite{barbivc2004segmenting} determines the number of dimensions (features) needed to approximate a sequence within a specified error.
+The proposed method of Berbi\v{c} \etal \cite{barbivc2004segmenting} determines the number of dimensions (features) needed to approximate a sequence within a specified error.
 With the observation that more dimensions are needed to keep the error below the threshold when transitions between actions occur, cut-points can be located and segments will be created.

### 2013-03-22 Chapters/Chapter1/lit_rev_temp_segm.tex
Fix small typo

@@
 \end{itemize}
-It is clear that for the application of this research Sliding-window and preferably SWAB-based algorithms should be considered.
+It is clear that for the application of this research sliding-window and preferably SWAB-based algorithms should be considered.
 
 The SWAB method proposed by Keogh \etal \cite{keogh2001online} is dependent on an user setting, providing the maximum error when performing both stages.
-Each segment is approximated by using Piecewise Linear Representation, an often used method.
+Each segment is approximated by using piecewise linear representation (PLR), an often used method.
 The user provided error threshold controls the granularity and number of segments.
@@
 
-An other application of PCA is to characterize the data by determining the dimensionality of a sequence of data points.
+Another application of PCA is to characterize the data by determining the dimensionality of a sequence of data points.
 The proposed method of Berbi\v{c} \etal \cite{barbivc2004segmenting} determines the number of dimensions (features) needed to approximate a sequence within a specified error.

### 2013-05-08 programming/change_detection/run_test.m
Remove programming implementations from git

@@ -1,47 +1,0 @@
-# Test the bottom-up / swab algorithm with stock data
-clear all
-source('swab.m');
-
-% data = load('stock_data.dat');
-% data = load('t_body_acc_energy_x.dat');
-data = load('body_acc_x_train_first_column_head.dat');
-y = data(1:end,1)';
-% x = [1:length(y)];
-
-clf
-figure("Position",[0,0,1600,800]);
-hold on;
-
-segments = bottom_up(y, 0.6);
-
-% segments
-
-function draw_segments(segments)
-  % x = [1:length(segments)];
-  x_running = 1;
-  colors = "gbmck";
-  for i = 1 : length(segments)
-    segment = segments(i){1,1};
-    segment_length = length(segment);
-    x_segment = [x_running:x_running+segment_length-1];
-
-    p = polyfit(x_segment, segment, 1);
-    yfit = polyval(p,x_segment);
-
-    line_segment = plot(x_segment, yfit, 'r-');
-    points = plot(x_segment,segment, 'b.+');
-
-    set( [line_segment, points] , 'color', colors(mod(i - 1,length(colors)) + 1));
-
-    x_running += segment_length;
-  endfor
-endfunction
-
-% segments
-
-% best_line(0.05)
-
-% segments = swab(0.8, 10);
-draw_segments(segments);
-printf("Number of final segments: %i\n", length(segments));
-% segments

### 2013-05-08 programming/change_detection/swab.m

@@ -1,170 +1,0 @@
-## Octave script for SWAB: Sliding Window And Buttom-up
-## Performs segmentation of signal, using linear regression
-##
-## Based on: An Online Algorithm for Segmenting Time Series
-## Eamonn Keogh Selina Chu David Hart Michael Pazzani
-
-1;
-
-% Return a segments for a time series based on bottom-up approach
-function segments = bottom_up( series, max_error)
-
-  segments = {};
-
-  # Create initial fine segments
-  for i = 1 : 2 : length(series) - 1
-    segments(end+1) = series(i:i+1);
-  endfor
-
-
-  merge_cost = [];
-  for i = 1 : length(segments) - 1
-    merged_segment = [segments(i){1,1}, segments(i+1){1,1}];
-    merge_cost(i) = calculate_error(merged_segment);
-  endfor
-
-  % As long as we can merge a segment and stay below the error
-  while (min(merge_cost) < max_error && length(segments) > 1)
-    [minimum, index] = min(merge_cost);
-
-    # Merge segment with lowest cost
-    segments(index) = [segments(index){1,1} segments(index+1){1,1}];
-    segments(index+1) = [];
-
-    % Calculate new merge cost for current with next segment ...
-    if index < (length(segments) - 1)
-      % There is a next segment
-      merged_new_with_next = [segments(index){1,1} segments(index+1){1,1}];
-      merge_cost(index) = calculate_error( merged_new_with_next);
-      merge_cost(index+1) = [];
-    else
-      % This was the last segment, no next merging
-      merge_cost(index) = [];
-    end
-
-    # ... and current with previous segment
-    if index > 1
-      merged_new_with_previous = [segments(index-1){1,1} segments(index){1,1}];
-      merge_cost(index-1) = calculate_error( merged_new_with_previous);
-    endif
-  endwhile
-endfunction
-
-% Return the error for a segment, with applying linear regression
-% As used in http://www.mathworks.nl/help/matlab/data_analysis/linear-regression.html
-function error = calculate_error( segment )
-
-  x = [1:length(segment)];
-  y = segment;
-
-  % p = polyfit(x,y,1);
-  % yfit = polyval(p,x);
-  % yresid = y - yfit;
-  % SSresid = sum(yresid.^2);
-  % SStotal = (length(y)-1) * var(y);
-  % error = SSresid / SStotal;
-
-
[…]

### 2013-05-08 Chapters/Chapter1.tex
Create outlines for new focus: temporal segmentation

--- 2013-02-28 Chapters/Chapter1.tex
@@
+% !TEX root = ../main.tex
 % Chapter 1
 
-\chapter{Introduction} % Main chapter title
+\chapter{Introduction}
 
@@
 
-\lhead{Chapter 1. \emph{Introduction}} % This is for the header on each page -
-%perhaps a shortened title
+\lhead{Chapter 1. \emph{Introduction}} % This is for the header on each page - perhaps a shortened title
 
@@
 
-\section{Problem statement}
-*** Give context problems, means to reach (e.g. monitoring of patients,  getting conclusions over health and activity).
-%----------------------------------------------------------------------------------------
-
-\input{./Chapters/Chapter1/literature_review}
-
-%----------------------------------------------------------------------------------------
-
-\section{System architecture}
-
-*** [GRAPHICS diagram: sensors --> signal pre-processing --> feature extraction --> analysis [* segmentation, * labeling] --> conclusions]
-
-%----------------------------------------------------------------------------------------
+\section{Outline}
+\begin{itemize}
+  \item Context of research (human activity recognition), real-world applications
+  \item Current methods, wrapper vs. filter methods
+  \item Problem statement with current filter methods (which follows from Chapter \ref{Chapter3} which goes in-depth with methods).
+  \item Purpose of this research. E.g. "Find a better algorithm for short-activity segmentation"
+  \item Relate to real-world applications
+\end{itemize}

### 2013-05-08 Chapters/Chapter2.tex
Extended bib with sensor papers
Create outlines for new focus: temporal segmentation

--- 2013-02-07 Chapters/Chapter2.tex
@@
+% !TEX root = ../main.tex
 % Chapter 2
-% !TEX root = ../main.tex
 
-\chapter{Techniques} % Main chapter title
+\chapter{Literature review}
 
-\label{Chapter2} % For referencing the chapter elsewhere, use \ref{Chapter1}
+\label{Chapter2} % For referencing the chapter elsewhere, use \ref{Chapter2}
 
-\lhead{Chapter 2. \emph{Techniques}} % This is for the header on each
-%page -
-%perhaps a shortened title
+\lhead{Chapter 2. \emph{Literature review}} % This is for the header on each page - perhaps a shortened title
 
 %----------------------------------------------------------------------------------------
-
-\section{Signal pro-processing and sensor fusion for timed patterns}
-
-*** [fusion of accelerometer data, gyroscopes, magnetic field sensors,
-pressure, microphones, etc]
-
-\section{Temporal Segmentation}
-
-This section will give an introduction and in-depth analysis of temporal
-segmentation.
-
-\subsection{Aims of segmentation}
-When processing and analyzing time series of data, e.g. motion measurements,
-stock market fluctuations or natural language, first a low-level division
-between the discriminative parts of the stream must be made. One can view this
-as splitting the series into the \emph{atoms}, which are the building blocks
-of the total stream. These building blocks will be the aggregation of
-non-overlapping, internally homogeneous segments \cite{himberg2001time}. This
-means that the data points inside a segment should have some resemblance
-relation to each other and their difference lies between some boundary. The
-process of segmenting can be viewed as a subproblem to context analysis of
-time series. Temporal segmentation is closely related to temporal clustering,
-although it is a stricter, and simpler, process. Whereby clustering only
-restricts the data points on their distance relation (as used in a Voronoi
-diagram), within a segment the data points must also be contiguous.
-
-The task of segmentation can be performed in a manual
-matter, by cutting and labeling parts of the stream into coherent parts. This
-would require human (expert) knowledge and does not yield a clear cut because
-of ambiguity. With increasing storage abilities and easier motion capture
-systems, there is a desire for automated systems which perform the
-segmentation task unsupervised. Some algorithms used have the (often desired)
-side-effect of also clustering the segments, such that classes of segments can
-be discovered in the time series. These algorithms would not only be able to
-make a distinction between walking, sitting and walking, but would also
-recognize the reappearance of the walking activity.
-
-\subsection{Formal definition}
-Formally, temporal segmentation is dividing a time series $s$, which consists
-of $N$ samples $\mathbf{x}(1),\mathbf{x}(2),\dots,\mathbf{x}(N)$ from
-$\mathbf{R}^d$. Individual \emph{segments} are referenced by $s(a,b)$,
-consisting of the consecutive samples
-$\mathbf{x}(a),\mathbf{x}(a+1),\dots,\mathbf{x}(b)$, $a \le b$. Let $s_1 =
-s(a,b)$ and $s_2 = s(b+1,c)$ be two segments, then their concatenation is
-$s_1s_2 = s(a,c)$. A segmentation $S$ of $s$ consists of a sequence of $k$
-non-empty segments $s_1s_2 \dots s_k = s$.
-This notation is adopted from \cite{himberg2001time}.
-
-As stated, informally each segment should be internally homogeneous. This can
-formally be measured with an cost function $F$, indication the heterogeneity
-of a segment. The overall aim is to minimize the cost $F$. The cost of a
-segment is a function from the data points and the number of data points $n =
-b - a + 1$ and is expressed as
-\begin{equation} \label{eq:segment_cost}
-\mathrm{cost}_F (s(a,b)) = F(\mathbf{x};n|\mathbf{x} \in s(a,b))
-\end{equation}
-The cost of a \emph{k-segmentation} $S$ is the summation of the costs of the
-$k$ segments:
-\begin{equation} \label{eq:segmentation_cost}
-\mathrm{Cost}_F (s_1 s_2 \dots s_k) = \sum_{i=1}^{k} \mathrm{cost}_F (s_k)
-\end{equation}
-With the objective of minimizing the cost function, the optimal
-$k$-segmentation $S_F^\mathit{opt}(s;k)$ is the segmentation with minimal
-$\mathrm{Cost}_F(s_1 s_2 \dots s_k)$ over all possible $k$-segmentations.
-
-The cost function, to calculate the heterogeneity of a (set of) segment(s),
-can
-be any function. A simple and natural function would be the sum of variances
-of the segments. The overall cost function would then be
-\begin{equation} \label{eq:cost_variances}
-\mathrm{Cost}_V = \frac{1}{N} \sum_{i=i}^{k} \sum_{j=c_{i-1}+1}^{c_i} \|
-\mathbf{x}(j) - \mu_i \|^2
-\end{equation}
-where $\mu_i$ is the mean vector of data points in segment $s_i$.
-
-\subsection{Application in research fields}
-
-[CHARACTERISTICS OF HUMAN MOTION]
-temporal variability, invariance over time, metrics over actions.
-
-[COMPUTER VISION]
-
-[GRAPHICS/VIDEO]
-
-[DATA-MINING]
-
-[MODEL BASED]
-
-To analyze time series it is often preferred to divide the stream in segments
-of correlated data. After dividing, each segment represent a period in time in
-which the same activity is performed. Or, stated otherwise, it results in
-transitions moments between activities.
-
-\subsection{PCA Based Methods}
-
-Many fields of research have been active in the unsupervised segmentation of
-data. Many authors rely on a form of Principal Component Analysis (PCA), as
-used a.o. in \cite{barbivc2004segmenting}. Often PCA is used to reduce the
-dimensionality of the data being processed [REFERENCE] by only using the top
-$r$ dimensions to describe the data set. It is observed that data series of
-simple motions have a lower dimensionality then complexer motions. When a
-simple (repetitive) motion is about to end and fluently transforms in a new
-motion, there will be a window of time in which a high dimensionality will be
-present, due to the new motion. After this period of transition, the
-dimensionality will decrease, since only the new simple motion is present in
-the window of time. The first algorithm of \cite{barbivc2004segmenting} is
-based on this principle.
-
-Given a set of data points, a lower dimensional hyperplane can by constructed
-to which the data points can be projected. This projection on a lower
-dimension introduces a error to the original position. When the error is
-fixed, less dimensions are needed for simple motions in which movements of
-body parts are highly correlated. For segments in which the data points are
-lesser correlated, e.g. because of transition state, a higher degree of
-dimensions of the hyperplane is needed to represent the data with equal error
-degree. When the dimensionality is reduced from $d$ to $r \le d$, the ratio of
-error
-$E_r$ can be calculated as
-
-\begin{equation}
-	E_r = \frac{\sum_{j=1}^{r} \sigma_j^2}{\sum_{j=1}^{d} \sigma_j^2}
-\end{equation}
-
-where $\sigma_j$ are the singular values as a result from Singular Value
-Decomposition (SVD), which is closely related to PCA \cite{shlens2005tutorial}.
-
-In \cite{barbivc2004segmenting} the stream of frames is analyzed on cut-frames
-to find transitions between action. First, for a number of $k$ frames (e.g.
-the
-equivalent of 2.5 seconds) the required dimensionality $r$ to keep the error
-$E_r$
-below some threshold $\tau$ is calculated. This will yield in an error $e_i$
-for the first
-$i$ frames. When more frames are added to the window, the error will increase
-with a low constant when it is still in the same activity, due to noise in the
-activity. When a new activity
-starts, e.g. at frame $j$, the error at frame $j$ will start increasing
-faster. This can be expressed by the derivative of the error rate $d_i = e_i -
-e_{i-l}$, where $l$ is a constant to remove noise. From this derivative the
-mean and standard deviation can be calculated, for each point. When a
-derivative $d_j$ rises more than a factor $k_\sigma = 3$ standard deviations
-from the mean, a transition point is encountered. The previous frames are then
-cut from the sequence (as a segment) and the algorithm starts over.
-
-*** [figure to illustrate derivative and standard deviation]
-
-A second approach in \cite{barbivc2004segmenting} uses the probabilistic
-variant of PCA (PPCA) to model the data set as a Gaussian distribution instead
-of ignoring the frames which do not fit in the subspace. Over windows of
-frames the mean and variance are calculated. In a forward manner the
-Mahalanobis distance of a new window of frames is calculated, which represents
-the likelihood of the new window belonging to the same segment as the original
-widow. When the distance decreases, the likelihood increases which happens
-when the motions in the becomes more homogeneous. When a peak in the distance
-is reached, the new window of frames indicates a heterogeneous collection of
-motions in the window and thus a low likelihood of membership and a indication
-of a transition. In order to distinct activities and subactivities (which
-require a subset of motions is a distinct activity) the algorithm is also
-processed backward over the data series.
-
-*** [figure to illustrate mahalanobis distance measure and peaks]
-
-The third algorithm in \cite{barbivc2004segmenting} is based on the
-observation that data points (frames) tend to form clusters in the space.
-These clusters are represented by $k$ Gaussian distributions for which each
-the Expectation-Maximization (EM) algorithm estimates the mean $m_j$,
-covariance matrix $\sum_{j}$ and prior $\pi_j$. With all the Gaussian
-distributions estimated, the data points are assigned to the cluster with the
-highest membership likelihood. When two consecutive frames $x_i$ and $x_{i+1}$
-belong to different clusters, a transition of activities is recognized. Note
-that this algorithm succeeds in segmenting the data and also labels the
-similar simple activities.
-
-A drawback in this system, and many others which implement a variant of the
-$k$-means algorithm, is that the number of clusters $k$ need to be
-predetermined. To cope with this, often the algorithm is performed multiple
-times for different values of $k$. Using some criteria, e.g. the Bayesian
-Information Criterion \cite{pelleg2000x} or the Davies-Bouldin Index which
-guides $k$-means clustering as used in \cite{krause2003unsupervised}.
-
-\subsection{Statistical methods}
-An method to process data, or create a basis for further processing, is by
-analyzing the statistical properties of a data set. When there is a continuous
-stream of data, it is possible to extract events by comparing the statistical
-properties of sliding time windows. By definition, events have a different
-statistical profile then their background. When two consecutive actions are
-regarded as the background, then a transition is an event between them.
-
-In \cite{guenterberg2009automatic} a automatic system is used to segment a
-continuous stream of data. By calculating features as activity level from the
-standard deviation of a sliding window and applying Adaptive Thresholds, the
-data can be segmented in active and rest parts. There is no model
-representation.
-
-\subsection{Hidden Markov Models Based Methods}
-
-\subsection{Bayesian Methods}
-
-\subsection{Principal Component Analysis}
-When working with simple 2- or 3-dimensional data it is often, for humans,
-easy to discover patterns in the set. The data can be plotted and the lines or
-planes among which the data points lie gives an indication of the pattern.
-With Principal Component Analysis (PCA) this process is also possible for
-automated systems. With this method the overall form of a set of data points
-can be represented, the most discriminative dimensions can be found, the
-dimensionality of a set can be reduced (which yield in data compression) or
-the result can be used to characterize and differentiate sets of data points.
-
-The following steps are performed, which are explained below and illustrated
-in figure *** [add figure with plot of sets]:
-
-\begin{enumerate}
-	\item \textbf{Gather data} and represent them in a chosen number of
-	features,
-	\item \textbf{Subtract mean} to center the data around the axis. The new
-	data set will have mean zero,
-	\item \textbf{Generate covariance matrix} by calculating all pairwise
-	feature variances,
-	\item \textbf{Calculate Eigenvectors and Eigenvalues} of the covariance
-	matrix. The Eigenvectors are then normalized to make further calculations
-	easier,
-	\item \textbf{Generate feature vector} which indicated which features are
-	characteristic or meant to keep in de data set, by comparing the
-	Eigenvalues.
-\end{enumerate}
-
-The workings of the PCA \cite{smith2002tutorial} relies on the concepts of
-standard deviation, variance,
-covariance, eigenvectors and eigenvalues of matrices. These concepts will be
-discussed very briefly \footnote{For a more in-depth discussion we would like
-to refer the reader to \cite{jolliffe2005principal}}. The standard deviation
-and variance
-of a set are measures for the
-spread around the mean for a single dimension, or feature. The covariance
-between two features (dimensions of the data points) indicates how they are
-related; when positive the two
-features will increase together; when negative one will decrease when the
-other increases and when zero they are unrelated. The covariance matrix gives
-all the covariances for all pairs of features. This matrix is symmetrical
-about the diagonal and the values on the diagonal are the variances for each
-feature. For this matrix the eigenvectors can be computed. Eigenvectors have
-the characteristic that when they multiply a matrix, the resulting vector is a
-multiple of the Eigenvector. The amount by which it is the multiple is the
-Eigenvalue. This is illustrated in formulae \ref{eq:no-eigenvector} and
-\ref{eq:eigenvector}. The second formula shows an Eigenvector
-$ \left( \begin{smallmatrix}
-3 \\ 2 \end{smallmatrix} \right)$ and its
-associated Eigenvalue, 4. The vector in the first formula is not an
-Eigenvector.
-
-\begin{equation} \label{eq:no-eigenvector}
-	\begin{pmatrix} 2 & 3 \\ 2 & 1 \end{pmatrix}
-\times
-	\begin{pmatrix} 1 \\ 3 \end{pmatrix}
-=
-	\begin{pmatrix} 2 \cdot 1 + 3 \cdot 3 \\ 2 \cdot 1 + 1 \cdot 3
-	\end{pmatrix}
-=
-	\begin{pmatrix} 11 \\ 5 \end{pmatrix}
-\end{equation}
-
-\begin{equation} \label{eq:eigenvector}
-	\begin{pmatrix} 2 & 3 \\ 2 & 1 \end{pmatrix}
-\times
-	\begin{pmatrix} 3 \\ 2 \end{pmatrix}
-=
-	\begin{pmatrix} 2 \cdot 3 + 3 \cdot 2 \\ 2 \cdot 3 + 1 \cdot 2
-	\end{pmatrix}
-=
-	\begin{pmatrix} 12 \\ 8 \end{pmatrix}
-=
-	4 \times \begin{pmatrix} 3 \\ 2 \end{pmatrix}
-\end{equation}
-
-It are these Eigenvectors and Eigenvalues that makes PCA possible. Given a
-matrix of size $n \times n$ (only square matrices have Eigenvectors), $n$
-Eigenvectors can be found. Each Eigenvector is perpendicular, or orthogonal,
-to the others. The Eigenvectors combined describes the lines over which the
-data is plotted. The Eigenvector with the largest Eigenvalue is the
-\emph{principal} component of the data set. It is said that it is the most
-significant feature. Thus the Eigenvectors, and thereby the features or
-dimensions of the data set, can be sorted on significance by the Eigenvalues.
-
-With this ordering there are a few applications possible. The first is to just
-make a (comprehensible) representation of the data. The Eigenvectors describe
-the cloud of data points and thus are a compressed representation of the data.
-When the original data points are to be compressed, it is possible to remove
-the least significant features and reconstruct the data points from the
-resulting Eigenvectors. This will yield in a lossy compression. *** [figure to
-illustrate]. An extension of this application is to determine the number of
-features needed to keep the compression within a certain error criterion, as
-used in \cite{barbivc2004segmenting}. Sets of data can then be distinguished
-by the number of features needed to describe the points.
-
-*** [graph, a bit of formulae with sets and dimensions]
-
-%-------------------
-
-\subsection{Hidden Markov Models}
-Many dynamic systems can be viewed as some sequence of consecutive states
-which are observable by the world. These kind of systems are known as Markov
-Chains. When the state in which the system is can not be observed with
-complete certainty, a system known as a Hidden Markov Model (HMM) can be
-constructed. It has been shown that HMMs are effective on in the task of
-speech recognition \cite{rabiner1989tutorial}. Because of some resemblance
-between natural speech construction and human activities, modifications have
-been able to classify human activities \cite{guenterberg2009distributed} and
-extensions are implemented which create layers of HMMs to handle different
-granularity over time and reduce retraining
-\cite{oliver2002layered,perdikis2008recognition}. A HMM is a state-space model
-representing a single model for each label which needs to be classified.
-
-Formally, a Hidden Markov Model is characterized by the following elements
-\cite{rabiner1989tutorial}:
-\begin{enumerate}
-	\item A number of $N$ states in the model. These states are a
-	representation of the underlying mechanism of the system and can not be
-	directly observed from outside (hence the \emph{hidden} model). Individual
-	states are referenced by $S = \{S_1, S_2, \dots, S_N \}$ and a single
-	state in time $t$ is denoted by $q_t$.
-	\item An alphabet of $M$ discrete observable symbols for each state. An
-	observed symbol is the only indication of the systems current situation.
-	Each individual symbol is referenced by $V = \{ v_1, v_2, \dots, v_m \}$.
-	\item A state transition probability distribution $A = \{ a_{ij} \}$
-	indication the probability of being in state $j$ at time $t+1$ when the
-	state on time $t$ was $i$:
-	\begin{eqnarray}
-		a_{ij} = P [ q_{t+1} = S_j | q_t = S_i ], & 1 \le i, j \le N
-	\end{eqnarray}
-	Note that by setting the probability $a_{ij}=0$ for some $i$ and $j$
-	indicates that $j$ is not reachable from state $i$.
-	\item A symbol observation probability distribution in state $j$, $B = \{
-	b_j(k) \}$, where
-	\begin{eqnarray}
-	b_j(k) = P[ v_k\ \mathrm{at} \ t | q_t = S_j], & 1 \le j \le N \nonumber \\
-	& 1 \le k \le M
-	\end{eqnarray}
-	Note that every symbol can be observed in every state.
-	\item The initial state distribution in which the system starts $\pi= \{
-	\pi_i \}$ where
-	\begin{eqnarray}
-	\pi_i = P[q_1 = S_i], & 1 \le i \le N
-	\end{eqnarray}
-\end{enumerate}
-This system is often abbreviated to
-\begin{equation}
-	\lambda = \{ \pi_i, a_{ij}, b_j(k)\}
-\end{equation}
-The HMM will act as a generator giving a sequence of $T$ observations, each
-being a symbol from $V$:
-\begin{equation}
-	O = O_1, O_2 \dots O_T
-\end{equation}
-
-Given the characteristics of a specific HMM, there are three major application
-which serve the classification of an activity. The first application is, given
-a sequence of observations $O = O_1,O_2,\dots,O_T$, to label this sequence
-with the activity with
-the highest associated probability for that label, e.g. compute $P(O|\lambda)$
-for each model $\lambda$. The second task a HMM can perform is to
-\emph{explain} a observation sequence $O$, by choosing an optimal
-corresponding state sequence $Q = q_1,q_2,\dots,q_T$. The last application is
-more of a problem statement, being the task to train the HMM to adjust the
-parameters $\lambda = \{ \pi_i, a_{ij}, b_j(k)\}$ to maximize $P(O|\lambda)$.
-In the overall task of activity recognition, one would first use the last
-application to train a set of HMMs with the provided and labeled training
-data. The second application can be used to further optimize the HMM (set the
-number of states, adjust vector codebook etc.) and improve the capability of
-modeling the activities. Finally, when the system is provided with unlabeled
-data the first application is used to identify the observed sequence with
-trained examples, by comparing the probability result of multiple HMMs.
-
-Generally a HMM can be fully connected, or ergodic model. Other configurations
-are also possible, of which the left-right model, or Bakis model, is popular
-for temporal applications \cite{rabiner1989tutorial}. This because the
-structure of the state probability distribution resembles the construction of
-possible states over time. This left-right model can be generalized by setting
-\begin{eqnarray}
-	a_{ij} = 0, & j > i + \Delta
-\end{eqnarray}
-for some $\Delta$, indicating the possible \emph{distance} of the jumps over
-states.
-
-*** [INSERT GRAPHICS SHOWING STRUCTURE OF HMM]
-
-*** [discrete and continuous HHMs]
-
-*** Applications of HHM: \cite{shi2009towards} uses 10 coefficients extracted
-with FFT from 6 axis (3 times accelerometer and 3 times gyroscope). Example of
-continuous HMM (?). Recognition rates fo 50\%-100\%.
-
-In \cite{oliver2002layered} it is stated that HMMs are robust mechanisms with
-respect to changes in temporal segmentation of observations, although they
-suffer from a few problems when applied to reason about longer and complexer
-sequences over time. They state that, especially with limited training data,
-the HMM tend to be overfitted and tend to lack structure and have an excess of
-parameters. The overcome this limitation, they introduce a Layered Hidden
-Markov Model (LHHM) which they show is more robust to temporal changes in
-observations. They make use of the hypothesis from the field of psychology,
-that human behavior is hierarchically structured \cite{zacks2001event}. In
-\cite{perdikis2008recognition} a layered construction of HMMs is used which
-feeds the low-level data from a sliding window to the lower layer. This layer
-will recognize simple motions, such as moving the arm in a pre-defined
-segment. The results from these lower HMMs are then fed to a higher-level
-layer of HMMs which are able to recognize abstract motions, such as adjusting
-a monitor and picking up a pen. The results showed that the layered system was
-better at recognition then a similar single-layered setup. It was also robust
-in the sense that when the environment was changed, only the relatively simple
-low-level HMMs needed to be retrained.
-
-*** [DIAGRAM WITH EXAMPLE OF LOW- AND HIGH-LEVEL HMMs, SUCH AS MOVING ARM,
-PICKING UP PEN]
-
-*** Viterbi algorithm / path; find the most likely sequence of cause-states
-for
-the observed measures.
-
-%-------------------
-
-\subsection{Segmentation as clustering}
-In the previous section the discussed methods all relied on the stream of data
-points and tried to find cuts the discriminate between successive different
-type of activities. An other approach is to consider the tasks of segmentation
-as a type of clustering \cite{zhou2008aligned}. In clustering the objective is
-to assign labels, or classes, to all the data points indication a similar type
-of activity. A clustering $\mathcal{L}$ is thereby more informative then a
-segmentation but is also harder to produce.
-
-A clustering $\mathcal{L}$ is generated from a sequence of elements
-$\mathbf{X}$ which is decomposed in $m$ disjoint segments, each belonging to
-one of the $k$ classes. A segment $\mathbf{Y}_i \hat{=}
-\mathbf{X}_{[s_i,s_{i+1})}$ is composed of frames from position $s_i$ to
-$s_{i+1}$. A vector $g_{ci} = 1$ indicates class membership if $\mathbf{Y}_i$
-belongs to class $c$, otherwise $g_{ci} = 0$.
-
-When regarding segmentation of human motion as a task of clustering the
-difficulty is to model the temporal variability of actions
-and defining a robust metric between temporal actions. To overcome this,
-\cite{zhou2008aligned} introduces Aligned Cluster Analysis (ACA), by minimizing
-\begin{equation} \label{eq:ACA}
-J_{\mathit{ACA}}(\mathbf{G},\mathbf{s}) =
-\sum_{c=i}^{K} \sum_{i=1}^{m} g_{ci} \mathit{dist}_c
-(\mathbf{X}_{[s_i,s_{i+1})})
-\end{equation}
-
-The characteristic of ACA is that is enables segments to span over different
-number of data points, whereas the standard kernel $k$-means algorithm results
-in equally sized segments. [!!! TRUE?!] The second difference is that the
-kernel used in $\mathit{dist}_c$ to measure the distance from a segment to the
-class which it is assigned to uses the Dynamic Time Alignment Kernel
-[REFERENCE?] to measure between time series.
-
-%----------------------------------------------------------------------------------------
-
-\section{Clustering}
-
-This section will give an introduction and overview to clustering of data.
-When processing time series of data the line of distinction between
-segmentation and clustering is very fine. This section will introduce
-clustering for the purpose as used in this project.
-
-Data clustering can be used with two different applications, exploratory or
-confirmatory \cite{jain1999data}. In the former, the purpose is to discover
-clusters in a point cloud of data points. Here the cluster is defined as a
-group of homogeneous data points measured by some coherence measure, often a
-distance function with the data points being defined as vectors. The produced
-clusters, which together form a representation of the data examined, can be
-used in the
-latter application to assign new provided data points to a cluster and thereby
-classify them.
-
-Depended on the precise setup, clustering can be used as a successive step
-after or an implementation of temporal segmentation. When temporal
-segmentation is used to create successive coherent data points, clustering can
-be used to recognize the same activities in different points of time. The data
-points provided to the clustering will be some representation of each segment
-and each segment will be labeled with a activity. Another mechanism could be
-to extract features from the raw data points and use these to create clusters
-directly *** [temporal modifications]. Both mechanisms requires the
-exploratory and confirmatory phases.
-
-The exploratory phase roughly consists of the following five steps
-\cite{jain1999data}
-\begin{enumerate}
-	\item data point representation,
-	\item definition of coherence,
-	\item grouping data points,
-	\item cluster abstraction,
-	\item output qualification
-\end{enumerate}
-
-The first step is to find a \emph{representation} of the data points and
-clusters, considering the number, type and scale of the features and the
-number of desired classes to cluster in. When dealing with high-dimensional
-data points a \emph{feature selection} can be performed to use only the most
-discriminating features. When the raw features are not useful enough,
-\emph{extraction} can be used to create new synthetic features. ***
-[qualitative versus quantitative/conceptual]
-
-When the data points are represented in a meaningful way, the measure of
-\emph{coherence} between pairs of points must be defined. Often this is
-implemented as the Euclidean
-or Mahalanobis distance when the data points are represented as vectors or
-other similarity measures for conceptual patterns.
-
-In the most critical step, the \emph{grouping}, the data points with the
-highest coherence are linked together to form a cluster. The result can be a
-hard partition over the data or a fuzzy membership degree to each cluster for
-each data point. When applying rules for merging and splitting cluster, a
-hierarchical partition is constructed. Partitional clustering algorithms
-assign data points to clusters by optimizing some criterion, e.g. the mean
-squared distance from each data point to the clusters' centroid.
-
-Especially in case of large data sets, the resulting clusters will be defined
-by many data points. To form a compacter and simpler representation,
-\emph{abstraction} can be used. This will simply analysis by humans or
-automated systems. A common abstraction is to use the clusters' centroid
-\cite{diday1976clustering} or parameters of Gaussian patterns.
-
-The final step which can be applied to the resulting partition is some
-\emph{qualification}. The quality of multiple partitions can be compared and
-the validity of the partition can be determined. A partition is valid if
-reasonably it could not been constructed by change or as some random process.
-External references of models can be used to compare, or the internal data
-points can be examined. *** [better wordings]
-
-%-------------------
-
-\subsection{Formal definition}
-
-%-------------------
-
-\subsection{Application is research fields}
-
-%----------------------------------------------------------------------------------------
-
-\section{Temporal pattern recognition}
-
-\subsection{Dynamic Time Warping}
-Used to measure similarity between time sequences. Exact matching is
-high-cost, so approximations such as Minimum Bounding Rectangles are used.
-
-\subsection{k-means clustering}
-[EXPLAIN] divide data set $n$ into $k$ clusters.
-
-Among many unsupervised clustering techniques, $k$-means is successfully
-applied to large data sets. It is simple to implement and linear in time
-complexity so computationally attractive \cite{jain1999data}. A drawback of
-the method is that the results of the algorithm greatly depends on the initial
-configuration (the data points which will act as centroids) and the number of
-cluster $k$ must be determined beforehand.
-
-Generally, the $k$-means methods will minimize the squared error for a
-clustering $\mathcal{L}$ criterion which is defined as the distance from the
-data points the centroid for each cluster in $\mathcal{K}$. This is expressed
-as optimizing to a local optimum the energy function
-\begin{equation} \label{eq:k-means energy}
-e^2(\mathcal{K},\mathcal{L}) =
-\sum_{j=i}^{K}\sum_{i=1}^{n_j}\|\mathbf{x}_i^{(j)} -
-\mathbf{c}_j\|^2
-\end{equation}
-
-There are several limitation on the $k$-means method. One of these is that
-only spherical shapes of cluster can be generated. One of the extensions is
-kernel $k$-means \cite{scholkopf1998nonlinear}, which implicitly projects the
-data points to a higher dimension and thereby is able to form irregular shaped
-cluster.
-
-\subsection{Self-organizing Map}
-
-\subsection{Support Vector Machine}
-
-\subsection{Na\"{i}ve Bayes}
-
-\section{Unsupervised clustering of temporal patterns}
+\section{Outline}
+\begin{itemize}
+  \item Literature review about Temporal Segmentation (previous draft was more about classification)
+  \item Consider methods for the context of filter-methods for classification
+  \item Take a loot at 3-4 different kind of methods for change detection:
+    \begin{itemize}
+      \item Dimensionality reduction
+      \item Density-ratio estimation
+      \item Support Vector Machines (?) - if there are more sources about this
+      \item CUSUM - or other more traditional methods
+    \end{itemize}
+  \item With each method, shortly look at characteristics, strengths and weaknesses and consider applicability to accelerometer sensor data
+\end{itemize}

### 2013-05-08 Chapters/Chapter3.tex
Create outlines for new focus: temporal segmentation

@@
+% !TEX root = ../main.tex
 % Chapter 3
 
-\chapter{Experiments} % Main chapter title
+\chapter{Change detection by Density-Estimation}
 
-\label{Chapter3} % For referencing the chapter elsewhere, use \ref{Chapter1}
+\label{Chapter3} % For referencing the chapter elsewhere, use \ref{Chapter3}
 
-\lhead{Chapter 3. \emph{Experiments}} % This is for the header on each 
-%page - 
-%perhaps a shortened title
+\lhead{Chapter 3. \emph{Density-Ratio estimation}} % This is for the header on each page - perhaps a shortened title
 
 %----------------------------------------------------------------------------------------
-
-\section{}
+\section{Outline}
+\begin{itemize}
+  \item In-depth analysis on one of the methods of Chapter \ref{Chapter2}
+  \item This method (e.g. Density-Ratio estimation) will be the basis for the real research
+  \item Explain why this methods seems worthy and interesting
+  \item Look at problems when applied to accelerometer sensor data
+  \item \emph{The problems discovered here will give rise to the problem statement at the Introduction / beginning of research}
+  \item Opens the possibility for own method
+\end{itemize}

### 2013-05-08 Chapters/Chapter4.tex
New template, merged old text
Create outlines for new focus: temporal segmentation

--- 2013-01-30 Chapters/Chapter3.tex
@@
-% Chapter 3
+% !TEX root = ../main.tex
+% Chapter 4
 
-\chapter{Experiments} % Main chapter title
+\chapter{Proposed method}
 
-\label{Chapter3} % For referencing the chapter elsewhere, use \ref{Chapter1}
+\label{Chapter4} % For referencing the chapter elsewhere, use \ref{Chapter4}
 
-\lhead{Chapter 3. \emph{Experiments}} % This is for the header on each 
-%page - 
-%perhaps a shortened title
+\lhead{Chapter 4. \emph{Proposed method}} % This is for the header on each page - perhaps a shortened title
 
 %----------------------------------------------------------------------------------------
-
-\section{}
+\section{Outline}
+\begin{itemize}
+  \item Based on the problem statement with current research as stated in Chapter \ref{Chapter3}
+  \item Adjust method to needs
+  \item Explain using graphs, pseudo-algorithms. Make clear distinction in origin of ideas and why to apply
+\end{itemize}

### 2013-05-08 Chapters/Chapter5.tex
Create outlines for new focus: temporal segmentation

% !TEX root = ../main.tex
% Chapter 5

\chapter{Result}

\label{Chapter5} % For referencing the chapter elsewhere, use \ref{Chapter5}

\lhead{Chapter 5. \emph{Results}} % This is for the header on each page - perhaps a shortened title

%----------------------------------------------------------------------------------------
\section{Outline}
\begin{itemize}
  \item Compare proposed method with methods of Chapter \ref{Chapter2}
  \item Provide plots, tables, graphs, error rates, precision, etc.
  \item Apply to a multiple of data, to compare to previous research - use that data
  \item Give theoretical analysis about performance. Big-O, memory, run-time, precision.
  \item This sections needs programmed implementations of own method and the ones compared
\end{itemize}

### 2013-05-08 Chapters/Chapter6.tex
Create outlines for new focus: temporal segmentation

% !TEX root = ../main.tex
% Chapter 6

\chapter{Real-world applications}

\label{Chapter6} % For referencing the chapter elsewhere, use \ref{Chapter6}

\lhead{Chapter 6. \emph{Real-world applications}} % This is for the header on each page - perhaps a shortened title

%----------------------------------------------------------------------------------------
\section{Outline}
\begin{itemize}
  \item Apply proposed method to real-world applications, such as
    \begin{itemize}
      \item Daily life activity recognition (as the original context of this thesis is)
      \item PowerHouse sensor data
      \item Stock data?
    \end{itemize}
  \item Relate back to filter vs. wrapper methods - give results with different methods?
\end{itemize}

### 2013-05-08 Chapters/Chapter7.tex
Create outlines for new focus: temporal segmentation

% !TEX root = ../main.tex
% Chapter 7

\chapter{Conclusion}

\label{Chapter7} % For referencing the chapter elsewhere, use \ref{Chapter7}

\lhead{Chapter 7. \emph{Conclusion}} % This is for the header on each page - perhaps a shortened title

%----------------------------------------------------------------------------------------
\section{Outline}
\begin{itemize}
  \item Conclude research
  \item Future research
\end{itemize}

### 2013-05-08 Chapters/old_classification/Chapter1/literature_review_nonparametric_clustering.tex
Create outlines for new focus: temporal segmentation

%
% !TEX root = ../../main.tex

\subsection{Non-parametric clustering methods}\label{sec:lit_review_nonparametric}
Many clustering methods rely on either a known number of expected clusters on implement an iterative approach to discover the ideal number of clusters.
In non-parametric approaches the number of clusters is a priori unknown and this discovery is intrinsic in the method, besides the labeling of the data.
This subsection will review the applied non-parametric methods in the field of clustering, where possible of time series.

### 2013-05-10 Chapters/Chapter2/svm.tex
Write on SVM segmentation

% !TEX root = ../../main.tex
\section{Change-detection by Support Vector Machines}

Introduced by Vapnik \cite{vapnik1998statistical, vapnik1999nature}, Support Vector Machines offer a way to segment, and classify, linear separable data.
When combined with a mapping function, which maps the data from the input space $I$ to a higher dimension feature space $F$, the input data can be non-linear separable.
The linear hyperplane, which segments the data in the feature space $F$, yields to a non-linear segmentation in the lower-dimensional input space $I$.
Instead of explicitly mapping the input data to the higher dimensional space, a kernel function can be used.
This kernel function can calculate values of the feature space directly, without the need to first map the input values to this space.
This process is referred to as the kernel trick.

*** Let sigma be a mapping from I to F such that the dot product in F can be computed using some simple kernel ***

\subsection{One-class Support Vector Machine}
The proposed method of Camci \cite{camci2010change} uses a one-class support vector machine to segment time series data.
One-class SVMs are used to describe the current data under consideration, by assuming all data points are from the same class \cite{tax2001one}.
The class is described by a spherical boundary around the data with center $c$ and radius $r$, such that the volume is minimized.
Following the definition of Camci \cite{camci2010change}, the class description is obtained by minimizing $r^2$:
\begin{equation}
  \mathrm{Min}\ r^2
\end{equation}
\begin{equation}
  \mathrm{Subject\ to} : \|\mathbf{x}_i - \mathbf{c}\|^2 \le r^2\ \forall i,\ \mathbf{x}_i : i \mathrm{th\ data\ point}
\end{equation}

To be able to handle outliers in the input data, a penalty cost function $\varepsilon_i$ for each outlier can be added.

*** Add new function and constraints? ***

Using this one-class SVM formulation, differences between two (consecutive) windows of data points with size $w$ can be obtained.
The first window is used as the input set, $h_1$ and the second as the test set $h_t$.
For the first window a one-class SVM is constructed, yielding in a representation by $c_1$ and $r_1$.
When the data points of the second window belong to the same class, the representation of that one-class SVM would equal the first:
\begin{equation}
  c_1 = c_2, r_1 = r_2
\end{equation}

*** First tell more about (underlying) probability density functions, to relate to other methods ***

In case the second window of data points does not belong to the same class, i.e. the probability density function that describes the data differs from the first, the describing values of the second window will differ from the first.
The amount of difference can be expressed by a dissimilarity measure over the representations.
When the dissimilarity between the two windows exceeds some predefined threshold $th$, there exists a change point between the windows.

This process can be visualized as done in *** insert figure of four circles ***. The second window, $h_2$ can be constructed from the first by e.g. a shift of one data point. *** explain data point positions by circle ***.

Note that a difference in the SVM center $c$ or radius $r$ represent a change in the mean and variance, respectively.

### 2013-05-13 Chapters/Chapter2.tex
Add HTML version of PDF to repo.
Execute `pdf2htmlEX main.pdf` to regenerate it.
Write on SVM segmentation
Write about statistical models

@@
 \end{itemize}
+
+\section{Statistical framework}
+Many applications require the detection of time points at which underlying properties of a system change.
+This problem thus has received a lot of attention in the fields of data mining, etc... *** list and refs ***.
+
+Some methods, such as the CUSUM (cumulative sum) \cite{basseville1993detection} and GLR (generalized likelihood ratio) \cite{gustafsson1996marginalized, gustafsson2000adaptive} formulate this problem in the statistical framework in the following sense.
+A time series is represented by an underlying probability distribution generating the data.
+The PDFs (probability distribution function) of two consecutive segments around a target point are compared.
+If they differ significantly, the target point is considered as a change point \cite{kawahara2009change}.
+
+%----------------------------------------------------------------------------------------
+\section{Change-detection by Dimensionality Reduction / Covariance structure}
+
+%----------------------------------------------------------------------------------------
+\input{./Chapters/Chapter2/density_ratio}
+
+%----------------------------------------------------------------------------------------
+\input{./Chapters/Chapter2/svm}
+
+%----------------------------------------------------------------------------------------
+\section{Change-detection by Cumulative Sum}

### 2013-05-13 Chapters/Chapter2/density_ratio.tex
Write about statistical models

% !TEX root = ../../main.tex
\section{Change-detection by Density-Ratio Estimation}

Formulate the problem of detecting change in the statistical framework.
Consider the probability distributions from which two consecutive segments of time series around a target time point are generated.
When the disitrubtions differ significantly the target time point is regarded as a change point.

CUSUM (cumulative sum) \cite{basseville1993detection} and GLR (generalized likelihood ratio)

The distribution over the values of time series data can be represented with a probability density function (pdf).
Two sections of a time series data can be generated with the same underlying pdf or each with a different.

### 2013-05-14 Chapters/Chapter2.tex
Add introduction Chapter 2

--- 2013-05-08 Chapters/Chapter2.tex
@@
 \end{itemize}
+
+\section{Statistical framework}
+Many applications require the detection of time points at which the underlying properties of a system change.
+This problem thus has received a lot of attention in the fields of data mining, etc... *** list and refs ***.
+Often this problem is formulated in a statistical framework, by inspecting the data generating PDF (Probability Density Function) of the time series data.
+A change point is then defined as a significant change in the properties of the PDF, such as the mean and variance.
+
+The widely used CUSUM (cumulative sum) method by Basseville \etal \cite{basseville1993detection}, and the GLR (Generalized Likelihood Ratio) by Gustafsson \cite{gustafsson1996marginalized,gustafsson2000adaptive} take this approach.
+The former originates from control methods for detection from bench marks.
+The latter compares the logarithm of the likelihood ratio over two consecutive intervals.
+These two methods is discussed and analyzed in section \ref{cusum-glr}.
+
+The GLR method, as with others, relies on pre-specified parametric model assumptions and considers the data be independent over time, which makes it less flexible to real-world applications.
+The proposed methods by Kawahara \etal \cite{kawahara2009change} and Lui \etal \cite{liu2013change} try to overcome these problems by estimating the \emph{ratio} between the PDF, instead of estimating each PDF.
+This approach is discussed and analyzed in section \ref{density-ratio}.
+
+The density-estimation methods, as with the GLR, rely on the log-likelihood ratio between PDFs.
+The method of Camci \cite{camci2010change} takes an other approach within the statistical framework, by using a SVM (Support Vector Machine).
+One problem it tries to overcome is the (claimed) weakness of many methods to detect a decrease in variance.
+The method represents the distribution over the data points as a hyper-sphere in a higher dimension using kernel trick.
+A change in the PDF is represented by a change in the radius of this sphere.
+Section \ref{svm} discusses the SVM-method.
+
+The final method under consideration, change detection via (intrinsic) dimensionality reduction, takes a different point of view.
+Opposed to the other discussed methods, dimensionality reduction is framed in the MDL (Minimum Description Length) framework.
+It uses the estimated underlying number of parameters of the time series as a model for change detection.
+Section \ref{dim-reduction} discuses this method.
+
+%----------------------------------------------------------------------------------------
+% CUSUM and GLR
+\input{./Chapters/Chapter2/cusum_glr}
+
+%----------------------------------------------------------------------------------------
+% Density Ratio Estimation
+\input{./Chapters/Chapter2/density_ratio}
+
+%----------------------------------------------------------------------------------------
+% Support Vector Machines
+\input{./Chapters/Chapter2/svm}
+
+%----------------------------------------------------------------------------------------
+% Dimensionality reduciton
+\input{./Chapters/Chapter2/dim_reduction}

### 2013-05-14 Chapters/Chapter2/cusum_glr.tex
Add introduction Chapter 2

% !TEX root = ../../main.tex
\section{CUSUM and GLR}\label{cusum-glr}

Detection of variance: \cite{inclan1994use}.

### 2013-05-14 Chapters/Chapter2/cusum_glr.tex
Notes on CUSUM and GLR

@@
 
-Detection of variance: \cite{inclan1994use}.
+*** read book Basseville \cite{basseville1993detection} ***
+
+\subsection{CUSUM}
+Non-Bayesian change detection algorithm (thus: no prior distribution believe available for the change time).
+
+The CUSUM (cumulative sum) method is developed by Page \cite{page1954continuous} for the application of statistical quality control (it is also known as a control chart).
+
+Primary for detection of mean shift.
+
+\subsection{GLR}
+Also: maximum-likelihood estimation. "When applied to a dataset and a given statistical model, it provides estimates for the model's parameters."

### 2013-05-17 Appendices/AppendixA-summaries.tex
Starting summary Peter Flach

% !TEX root = ../main.tex
% Appendix A

\chapter{Summaries} % Main appendix title

\label{AppendixA} % For referencing this appendix elsewhere, use \ref{AppendixA}

\lhead{Appendix A. \emph{Summaries}} % This is for the header on each page - perhaps a shortened title

Please ignore this Appendix.
This appendix is for my own personal use.
It contains summaries of articles I have read.

%----------------------------------------------------------------------------------------
\section{Support Vector Machines}

\subsection{Machine learning: the art and science of algorithms that make sense of data}
Book by Peter Flach: \cite{flach2012machine}.
Mainly about chapter 7, ``Linear Models''.
Most important: section 7.3 - 7.5, about support vector machines and non-linearity.
\textbf{Some parts are direct text; do not use this text directly!}

\subsection{Linear models}
Models can be represented by their geometry of $d$ real-values features.
Data points are represented in the $d$-dimensional cartesian coordinate system/space $\mathcal{X} = \mathbb{R}^d$.
Geometric concepts such as lines and planes can be used for \emph{classification} and \emph{regression}.
An alternative approach is to use the distance between datapoints as a similarity measure, resulting from the geometrical representation.
Linear methods do not use that property, but rely on understanding of models in terms of lines and planes.

Linear models are of great interest in machine learning because of their simplicity.
A few manifestations of this simplicity are:
\begin{itemize}
  \item Linear models are \emph{parametric}, thus fixed small number of parameters that need to be learned from the data.
  \item Linear models are \emph{stable}, thus small variations in training data have small impact on the learned model. In logical models they can have large impact, because ``splitting rules'' in root have great impact.
  \item Due to relative few parameters, less likely to \emph{overfit} the training data.
\end{itemize}

The last two are summarized by saying that \emph{linear models have low variance but high bias}.
This is preferred with limited data and overfitting is to be avoided.

Linear models are well studied, in particular for the problem of linear regression.
This can be solved by the \emph{least-squares} method and classification as discussed in section \ref{least-squares}, the \emph{perceptron} as explained in section \ref{perceptron}.
Linear regression with the \emph{support vector machine} is handled in section \ref{svm-explained} and used for probability density estimation in section \ref{svm-pdf}.
The kernel trick used for learning non-linear models is explained in section \ref{non-linear}.

\subsection{Least-squares method}\label{least-squares}
The regression problem is to learn a function estimator $\hat{f}:\mathcal{X} \to \mathbb{R}$ from the examples $(x_i, f(x_i))$ where we assume $\mathcal{X} = \mathbb{R}^d$.
The difference between the actual and estimated function values are called \emph{residuals} $\epsilon_i = f(x_i) - \hat{f}(x_i)$.
The \emph{least-squares method} finds the estimation $\hat{f}$ by minimizing $\sum_{i=1}^{n} \epsilon_i^2$.
Univariate regressesion assumes a linear equation $y = a + b x$, with parameters $a$ and $b$ chosen such that the sum of squared residuals $\sum_{i=1}^{n} (y_i - (a + b x_i))^2$ is minimized.
Here the estimated parameter $\hat{a}$ is called the \emph{intercept} such that it goes through the (estimated) pooint $(\hat{x},\hat{y})$ and $\hat{b}$is the \emph{slope} which can be expressed by the (co)variances: $\hat{b} = \frac{\sigma_{xy}}{\sigma_{xx}}$.
In order to find the parameters, take the partial derivatives, set them to $0$ and solve for $a$ and $b$.

Although least-squares is sensitive to outliers, it works very well for such a simple method.
This can be explained as folows.
We can assume the underlying function is indeed linear but contanimated with random noise.
That means that our examples are actually $(x_i, f(x_i) + \epsilon_i)$ and $f(x) = ax + b$.
If we know $a$ and $b$ we can calculate what the residuals are, and by knowing $\sigma^2$ we can estimate \emph{the probability of observering the residuals}.
But since we don't know $a$ and $b$ we have to estimate them, by estimating the values for $a$ and $b$ that maximizes the probability of the residuals.
This is the \emph{maximum-likelihood estimate} (chapter 9 in the book).

The least-squares method can be used for a (binary) classifier, by encoding the target variable $y$ as classes by real numbers $-1$ (negative) and $1$ (positive).
It follows that $\matrixsym{X}^T\vectorsym(y) = P \vectorsym{\mu^+} - N \vectorsym{\mu^-}$, where $P$, $N$, $\vectorsym{\mu^+}$ and $\vectorsym{\mu^-}$ are the number of positive and negative examples, and the $d$-vectors containing each feature's mean values, resp.
The regression equation $y = \bar{y} + \hat{b}(x - \bar{x})$ can be used to obtain a decision boundary.
We need to determine the point $(x_0, y_0)$ such that $y_0$ is half-way between $y^+$ and $y^-$ (the positive and negative examples, i.e. $y_0 = 0$).

\subsection{Perceptron}\label{perceptron}
\subsection{Support Vector Machine}\label{svm-explained}
\subsection{Support Vector Machine Density Functions}\label{svm-pdf}
\subsection{Non-linear models}\label{non-linear}

%----------------------------------------------------------------------------------------

### 2013-05-18 Appendices/AppendixA-summaries.tex
Summary SVM

@@
 \subsection{Perceptron}\label{perceptron}
+Labelled data is \emph{linearly separable} if the exists a linear boundary separating the classes.
+The least-squares may find one, but it is not guaranteed.
+Image a perfect linearly separatable data set.
+Move all the positive points away from the negative, but one.
+At one point the new boundary will exclude (misqualify) the one original positive outlier.
+The \emph{perceptron} will guaranteed perform perfect separation when the data allows it to be.
+It was originally proposed as a \emph{simple neural network}.
+
 \subsection{Support Vector Machine}\label{svm-explained}

### 2013-05-21 Appendices/AppendixA-summaries.tex
Finished summary on perceptrons

@@
 Move all the positive points away from the negative, but one.
-At one point the new boundary will exclude (misqualify) the one original positive outlier.
+At one point the new boundary will exclude (misqualify) the one original positive outlier, due to the mean-statistics it relies on.
 The \emph{perceptron} will guaranteed perform perfect separation when the data allows it to be.
 It was originally proposed as a \emph{simple neural network}.
+It works by iterating over the training set and modifying the weight vector for every misclassified example ($\vectorsym{w} \cdot \vectorsym{x}_i < t$ for positive examples $\vectorsym{x}_i$).
+It uses a learning rate $\eta$, for a misclassified $y_i = \left\{-1,+1\right\}$: $\vectorsym{w}' = \vectorsym{w} + \eta y_i \vectorsym{x}_i$.
+The algorithm can be made \emph{online} by processing a stream of data points and and updating the weight vector only when a new data point is misclassified.
+
+When the algorithm is completed, every $y_i\vectorsym{x}_i$ is added $\alpha_i$ times to the weight vector (every time it was misclassified).
+Thus, the weight vector can be expressed as: $\vectorsym{w} = \sum_{i=i}^n \alpha_i y_i \vectorsym{x}_i$.
+In other words: the weight vector is a linear combination of the training instances.
+The dual form of the algorithm learns the instance weights $\alpha_i$ rather than the features weights $\vectorsym{w}_i$.
+An instance $\vectorsym{x}$ is then classified as $\hat{y} = sign(\sum_{i=1}^n \alpha_i y_i \vectorsym{x}_i \cdot \vectorsym{x})$.
+This means that during the training only the pairwise dot-products of the data is needed; this results in the \emph{n}-by-\emph{n} Gram matrix $\matrixsym{G} = \matrixsym{X}\matrixsym{X}^T$.
+This instance-based perspective will be further discussed in section \ref{svm-explained} about the support vector machine.
 

### 2013-05-21 Appendices/AppendixA-summaries.tex
Finish summary basics SVM

@@
 \subsection{Support Vector Machine}\label{svm-explained}
+A training example can be expressed by its \emph{margin}: $c(x)\hat{s}(x)$, where $c(x)$ is $+1$ for positive and $-1$ for negative examples and $\hat{s}(x)$ is the score.
+The score can be expressed as $\hat{s}(\vectorsym{x}) = \vectorsym{w} \cdot \vectorsym{x} - t$.
+A true positive example $\vectorsym{x}_i$ has a margin $\vectorsym{w} \cdot \vectorsym{x}_i > 0$ and a true negative $\vectorsym{x}_j$ has $\vectorsym{w} \cdot \vectorsym{x}_j < 0$.
+If $m^+$ and $m^-$ are the smallest positive and negative examples, then we want the sum of these to be as large as possible.
+\emph{The training examples with these minimal values are closest to the decision boundary $t$ and are called the support vectors.}
+The decision boundary is defined as a linear combination of the support vectors.
+The margin is thus defined as $\frac{m}{\norm{\vectorsym{w}}}$.
+Minimizing the margin (which is often set to $1$ and rescaling is allowed) yields to minimizing $\norm{\vectorsym{w}}$, or: $\frac{1}{2}\norm{\vectorsym{w}}^2$, restricted that none of the training points fall inside the margin.
+This gives the following quadratic, constrained optimisation problem:
+
+\begin{equation}
+\begin{aligned}
+  & \vectorsym{w}^*, t^* = \argmin\limits_{\vectorsym{w},t} \frac{1}{2} \norm{w}^2 & & \text{subject to } y_i(\vectorsym{w} \cdot \vectorsym{x}_i - t) \geq 1, 1 \leq i \leq n
+\end{aligned}
+\end{equation}
+
+This equation can be transformed with the Lagrange multipliers by adding the constraints to the minimization part with multipliers $\alpha_i$.
+Taking the partial derivative with respect to $t$ and setting it to $0$, we find that for the optimal solution (threshold) $t$ we have $\sum_{i=1}^{n} \alpha_i y_i = 0$.
+When we take the partial derivative with respect to $w\vectorsym{w}$ we see that the Langrange multipliers define the weight vector as a linear combination of the training examples.
+This partial derivative is $0$ for an optimal weight we get that $\vectorsym{w} = \sum_{i=1}^{n} \alpha_i y_i \vectorsym{x}_i$, \emph{which is the same expression as for the perceptrond derived in section \ref{perceptron}}.
+By plugging $\vectorsym{w}$ and $t$ back into the Langrange equation, we can eliminate these and get the dual optimization problem entirely formulated in terms of the Lagrange multipliers:
+
+\begin{equation}
+  \Lambda(\alpha_1, \dots, \alpha_n) = - \frac{1}{2} \sum\limits_{i=i}^n \sum\limits_{j=1}^n \alpha_i \alpha_j y_i y_j \vectorsym{x}_i \cdot \vectorsym{x}_j + \sum\limits_{i=1}^n \alpha_i
+\end{equation}
+
+The dual problem maximizes this function under positivity constaints and one equality constraint:
+
+\begin{equation}
+\begin{aligned}
+  \alpha_1^*, \dots, \alpha_n^* = \argmax\limits_{\alpha_1, \dots, \alpha_n} - \frac{1}{2} \sum\limits_{i=i}^n \sum\limits_{j=1}^n \alpha_i \alpha_j y_i y_j \vectorsym{x}_i \cdot \vectorsym{x}_j + \sum\limits_{i=1}^n \alpha_i \\
+  \text{subject to } \alpha_i \geq 0, 1 \leq i \leq n \text{ and } \sum\limits_{i=1}^n \alpha_i y_i = 0
+\end{aligned}
+\end{equation}
+
+This shows to important properties:
+\begin{enumerate}
+  \item Searching for the maximum-margin decision boundary is equivalent to searching for the support vectors; they are the training examples with non-zero Lagrange multipliers.
+  \item The optimization problem is entirely defined by pairwise dot products between training instances: the entries of the Gram matrix.
+\end{enumerate}
+
+The second property enables powerful adaption for support vector machines to learn non-linear decision boundaries, as discussed in section \ref{non-linear}.
+
+An other solution to non-linear separable data, that is when the constraints $\vectorsym{w} \cdot \vectorsym{x}_i - t \geq 1$ are not jointly satisfiable, is to add \emph{slack variables} $\xi_i$, one for each example.
+This allows them to be in the margin, of even at the wrong side of the boundary -- known as boundary errors.
+Thus, the constaints become $\vectorsym{w} \cdot \vectorsym{x}_i - t \geq 1 - \xi_i$.
+
+``In summary, \emph{support vector machines are linear classifiers that construct the unique decision boundary that maximises the distance to the nearest training examples (the support vectors)}.
+Training an SVM involves solving a large quadratic optimisation problem and is usually best left to a dedicated numerical solver.''
+
 \subsection{Support Vector Machine Density Functions}\label{svm-pdf}

### 2013-05-21 Appendices/AppendixA-summaries.tex
Finish summery probabilities from linear models

@@
 
+%--------------------------------------------
 \subsection{Linear models}
@@
 This can be solved by the \emph{least-squares} method and classification as discussed in section \ref{least-squares}, the \emph{perceptron} as explained in section \ref{perceptron}.
-Linear regression with the \emph{support vector machine} is handled in section \ref{svm-explained} and used for probability density estimation in section \ref{svm-pdf}.
+Linear regression with the \emph{support vector machine} is handled in section \ref{svm-explained} and used for probability density estimation in section \ref{linear-classifier-pdf}.
 The kernel trick used for learning non-linear models is explained in section \ref{non-linear}.
 
+%--------------------------------------------
 \subsection{Least-squares method}\label{least-squares}
@@
 
+%--------------------------------------------
 \subsection{Perceptron}\label{perceptron}
@@
 
+%--------------------------------------------
 \subsection{Support Vector Machine}\label{svm-explained}
@@
 
-\subsection{Support Vector Machine Density Functions}\label{svm-pdf}
+%--------------------------------------------
+\subsection{Density Functions from linear classifiers}\label{linear-classifier-pdf}
+The score of an data point can be used to obtain the signed distance of $\vectorsym{x}_i$ to the dicision boundary:
+\begin{equation}
+  d(\vectorsym{x}_i) = \frac{\hat{s}(\vectorsym{x}_i)}{\norm{w}} = \frac{\vectorsym{w} \cdot \vectorsym{x}_i - t}{\norm{w}} = \vectorsym{w}' \cdot \vectorsym{x}_i - t'
+\end{equation}
+where $\vectorsym{w}' = \vectorsym{w} / \norm{\vectorsym{w}}$ rescaled to unit length and $t' = t / \norm{w}$ corresponds to the rescaled intercept.
+this geometric interpretation of the scores enables them to turn into probabilities.
+Let $\bar{d}^+ = \vectorsym{w} \cdot \vectorsym{\mu}^+ - t$ denote the mean distance of the positive examples to the boundary, where $\vectorsym{\mu}^+$ is the mean of positive examples (in the grid) and $\vectorsym{w}$ is unit length.
+We can assume that the distance of the examples is normally distributed around the mean (which give a bell curve when plotted).
+
+If we obtain a new point $\vectorsym{x}$ we can get the class by $sign(d(\vectorsym{x}))$.
+We would like, instead, to get the probability (using Bayes' rule)
+\begin{equation}
+  \hat{p}(\vectorsym{x}) = P(+|d(\vectorsym{x}) = \frac{P(d(\vectorsym{x})|+)P(+)}{P(d(\vectorsym{x})|+)P(+) + P(d(\vectorsym{x})|-)P(-)} = \frac{LR}{LR + 1/clr}
+\end{equation}
+where $LR$ is the likelihood ratio obtained from the normal score distributions, als $clr$ is the class ratio.
+With some rewriting we can convert $d$ into a probability by means of the mapping $d \mapsto \frac{exp(d)}{exp(d) + 1}$, which is the \emph{logistic function}.
+The logarithm of the likelihood ratio is linear in $\vectorsym{x}$ and such models are called \emph{log-linear models}.
+This logistic calibration procedure can change the location of the decision boundary but not ts direction.
+There may be an alternative weight vector with a different direction that assign a higher likelihood to the data.
+Finding that maximum-likelihood linear classifier using the logistic model is called \emph{logistic regression}.
+
+%--------------------------------------------
 \subsection{Non-linear models}\label{non-linear}

### 2013-05-21 Appendices/AppendixA-summaries.tex
Finish summery book Flach

@@
 \end{equation}
-where $LR$ is the likelihood ratio obtained from the normal score distributions, als $clr$ is the class ratio.
+where $LR$ is the likelihood ratio obtained from the normal score distributions, and $clr$ is the class ratio.
 With some rewriting we can convert $d$ into a probability by means of the mapping $d \mapsto \frac{exp(d)}{exp(d) + 1}$, which is the \emph{logistic function}.
@@
 \subsection{Non-linear models}\label{non-linear}
+Linear methods such as least-squares for regression can be used for binary classification, yielding in the basic linear classifier.
+The (heuristic) perceptron guaranteers to classify correctly linear separable data points.
+Support vector machines find the unique decision boundary with maximum margin and can be adapted to non-linear separable data.
+These methods can be adjusted to learn non-linear boundaries.
+The main idea is to transform the data from the \emph{input space} non-linearly to a \emph{feature space} (which can, but does need to be in a higher dimension) in which linear classification can be applied.
+The mapping back from the feature space to the input space is often non-trivial (e.g.
+mapping $(x,y)$ to feature space by $(x^2, y^2)$, yields in four coordinates when transformed back to the input space).
+
+\emph{The remarkable thing is that often the feature space does not have to be explicitly constructed, as we can perform all necessary operations in input space.}
+For instance; the perceptron algorithm mainly depends on the dot product of $\vectorsym{x}_i \cdot \vectorsym{x}_j$.
+Assuming $\vectorsym{x}_i = (x_i, y_i)$ and $\vectorsym{x}_j = (x_j, y_j)$, the dot product can be written as $\vectorsym{x}_i \cdot \vectorsym{x}_j = x_i x_j + y_i y_j$.
+The instances in quadratic feature space are $(x_i^2, y_i^2)$ and $(x_j^2, y_j^2)$ and their dot product is $(x_i^2, y_i^2) \cdot (x_j^2, y_j^2) = x_i^2 x_j^2 + y_i^2 y_j^2$.
+This is almost equal to $(\vectorsym{x}_i \cdot \vectorsym{x}_j)^2 = (x_i x_j)^2 + (y_i y_j^2 + 2 x_i x_j y_i y_j$, but not quite because of the third term.
+We can make the equations equal by \emph{extending the feature space} (to a higher dimension) with a third feature $\sqrt{2xy}$, so the feature space is $\phi(\vectorsym{x}_i) = (x_i^2, y_i^2, \sqrt{2x_i y_i})$.
+
+If we define $\kappa(\vectorsym{x}_i, \vectorsym{x}_j) = (\vectorsym{x}_i, \vectorsym{x}_j)^2$ and replace $\vectorsym{x}_i \cdot \vectorsym{x}_j$ with $\kappa(\vectorsym{x}_i, \vectorsym{x}_j)$ in the (perceptron) algorithm, we obtain the \emph{kernel perceptron} with the degree $p = 2$.
+We are not restricted to polynomial kernels; an often used kernel is the \emph{Gaussian kernel}, defined as:
+\begin{equation}
+  \kappa(\vectorsym{x}_i, \vectorsym{x}_j) = exp(\dfrac{-\norm{\vectorsym{x}_i - \vectorsym{x}_j}^2}{2\sigma^2})
+\end{equation}
+where $\sigma$ is known as the \emph{bandwidth} parameter.
+We can think of the Gaussian kernel as imposing a Gaussian (i.e., multivariate normal) surface on each support vector in instance space, so that the boundary is defined in therms of those Gaussian surfaces.
+Kernel methods are best known in combination with support vector machines.
+Notice that the soft margin optimization problem is defined in terms of dot product between training examples, and thus the `kernel trick' can be applied.
+Note that the decision boundary learn with a non-linear kernel cannot be represented by a simple weight vector in input space.
+Thus, to classify a new example $\vectorsym{x}$ we need to evaluate $y_i \sum_{j=1}^n \alpha_j y_j \kappa(\vectorsym(x), \vectorsym{x}_j)$ (the Gram matrix?) involving all training examples, or at least all with non-zero multipliers $a_j$ (the support vectors).
 

### 2013-05-22 Appendices/AppendixA-summaries.tex
Working on SVM

@@
 %--------------------------------------------
-\subsection{Linear models}
+\subsubsection{Linear models}
 Models can be represented by their geometry of $d$ real-values features.
@@
 %--------------------------------------------
-\subsection{Least-squares method}\label{least-squares}
+\subsubsection{Least-squares method}\label{least-squares}
 The regression problem is to learn a function estimator $\hat{f}:\mathcal{X} \to \mathbb{R}$ from the examples $(x_i, f(x_i))$ where we assume $\mathcal{X} = \mathbb{R}^d$.
@@
 %--------------------------------------------
-\subsection{Perceptron}\label{perceptron}
+\subsubsection{Perceptron}\label{perceptron}
 Labelled data is \emph{linearly separable} if the exists a linear boundary separating the classes.
@@
 %--------------------------------------------
-\subsection{Support Vector Machine}\label{svm-explained}
+\subsubsection{Support Vector Machine}\label{svm-explained}
 A training example can be expressed by its \emph{margin}: $c(x)\hat{s}(x)$, where $c(x)$ is $+1$ for positive and $-1$ for negative examples and $\hat{s}(x)$ is the score.
@@
 %--------------------------------------------
-\subsection{Density Functions from linear classifiers}\label{linear-classifier-pdf}
+\subsubsection{Density Functions from linear classifiers}\label{linear-classifier-pdf}
 The score of an data point can be used to obtain the signed distance of $\vectorsym{x}_i$ to the dicision boundary:
@@
 %--------------------------------------------
-\subsection{Non-linear models}\label{non-linear}
+\subsubsection{Non-linear models}\label{non-linear}
 Linear methods such as least-squares for regression can be used for binary classification, yielding in the basic linear classifier.
@@
 %----------------------------------------------------------------------------------------
+
+\subsection{Change Point Detection In Time Series Data Using Support Vectors}
+Paper by Fatih Camci \cite{camci2010change}
+About segmentation with SVMs.
+Will be main material for section \ref{svm} about SVMs.
+
+\subsubsection{Introduction}
+Interprets change detection as finding the transition points from one underlying time series generation model to another.
+The change point is mostly represented in a sudden change in mean or variance.
+Existing models detect changes in mean and increase in variance, \emph{but fail to recognize decrease in variance}.
+Many methods require some model (like Auto-Regressive [AR]) to fit the time series in order to eliminate the noise.
+Thus, the effectiveness of the method is tied to the fitness degree of the model to the time series data.
+These two problems (lack of variance decrease detection and model-bound fitness degree) leads to this work; Support Vector based Change Point Detection targeting changes in variance and/or mean without any assumption of model fitting of data distribution.
+This method \emph{does not use a time series model} for fitting and targets \emph{both increase and decrease} in mean and variance.
+
+\subsubsection{Related work}
+Change Point Detection (CPD) can be categorized in posterior (off-line) and sequential (on-line).
+Sequential receive data sequentially and analyze previously obtained data to detect the possible change in current time.
+This method is based on sequential analysis and focuses on change on mean and variance in time domain.
+Other methods generally suffer from:
+\begin{itemize}
+  \item Inability / inefficiency in detecting variance decrease.
+  \item Assumptions about the statistical distribution of the data, obtained as error of fitting the (AR) model.
+  \item Necessity of training the model with possible changes.
+\end{itemize}
+
+\subsubsection{Support vector based one-class classification}

### 2013-05-22 Chapters/Chapter2/cusum_glr.tex
Working on SVM

@@
 
+The MOSUM of squares test for monitoring variance changes \cite{hsu2007mosum}.
+
+Use of Cumulative Sums of Squares for Retrospective Detection of Changes of Variance \cite{inclan1994use}
+
 \subsection{GLR}
-Also: maximum-likelihood estimation. "When applied to a dataset and a given statistical model, it provides estimates for the model's parameters."
+Also: maximum-likelihood estimation. ``When applied to a dataset and a given statistical model, it provides estimates for the model's parameters.''

### 2013-05-23 Appendices/AppendixA-summaries.tex
Add one-class summary. Fix some typs

@@
 Models can be represented by their geometry of $d$ real-values features.
-Data points are represented in the $d$-dimensional cartesian coordinate system/space $\mathcal{X} = \mathbb{R}^d$.
+Data points are represented in the $d$-dimensional Cartesian coordinate system/space $\mathcal{X} = \mathbb{R}^d$.
 Geometric concepts such as lines and planes can be used for \emph{classification} and \emph{regression}.
-An alternative approach is to use the distance between datapoints as a similarity measure, resulting from the geometrical representation.
+An alternative approach is to use the distance between data points as a similarity measure, resulting from the geometrical representation.
 Linear methods do not use that property, but rely on understanding of models in terms of lines and planes.
@@
 Although least-squares is sensitive to outliers, it works very well for such a simple method.
-This can be explained as folows.
-We can assume the underlying function is indeed linear but contanimated with random noise.
+This can be explained as follows.
+We can assume the underlying function is indeed linear but contaminated with random noise.
 That means that our examples are actually $(x_i, f(x_i) + \epsilon_i)$ and $f(x) = ax + b$.
-If we know $a$ and $b$ we can calculate what the residuals are, and by knowing $\sigma^2$ we can estimate \emph{the probability of observering the residuals}.
+If we know $a$ and $b$ we can calculate what the residuals are, and by knowing $\sigma^2$ we can estimate \emph{the probability of observing the residuals}.
 But since we don't know $a$ and $b$ we have to estimate them, by estimating the values for $a$ and $b$ that maximizes the probability of the residuals.
@@
 \subsubsection{Perceptron}\label{perceptron}
-Labelled data is \emph{linearly separable} if the exists a linear boundary separating the classes.
+Labeled data is \emph{linearly separable} if the exists a linear boundary separating the classes.
 The least-squares may find one, but it is not guaranteed.
-Image a perfect linearly separatable data set.
+Image a perfect linearly separable data set.
 Move all the positive points away from the negative, but one.
-At one point the new boundary will exclude (misqualify) the one original positive outlier, due to the mean-statistics it relies on.
+At one point the new boundary will exclude (mis qualify) the one original positive outlier, due to the mean-statistics it relies on.
 The \emph{perceptron} will guaranteed perform perfect separation when the data allows it to be.
@@
 Minimizing the margin (which is often set to $1$ and rescaling is allowed) yields to minimizing $\norm{\vectorsym{w}}$, or: $\frac{1}{2}\norm{\vectorsym{w}}^2$, restricted that none of the training points fall inside the margin.
-This gives the following quadratic, constrained optimisation problem:
+This gives the following quadratic, constrained optimization problem:
 
@@
 Taking the partial derivative with respect to $t$ and setting it to $0$, we find that for the optimal solution (threshold) $t$ we have $\sum_{i=1}^{n} \alpha_i y_i = 0$.
-When we take the partial derivative with respect to $w\vectorsym{w}$ we see that the Langrange multipliers define the weight vector as a linear combination of the training examples.
-This partial derivative is $0$ for an optimal weight we get that $\vectorsym{w} = \sum_{i=1}^{n} \alpha_i y_i \vectorsym{x}_i$, \emph{which is the same expression as for the perceptrond derived in section \ref{perceptron}}.
-By plugging $\vectorsym{w}$ and $t$ back into the Langrange equation, we can eliminate these and get the dual optimization problem entirely formulated in terms of the Lagrange multipliers:
+When we take the partial derivative with respect to $w\vectorsym{w}$ we see that the Lagrange multipliers define the weight vector as a linear combination of the training examples.
+This partial derivative is $0$ for an optimal weight we get that $\vectorsym{w} = \sum_{i=1}^{n} \alpha_i y_i \vectorsym{x}_i$, \emph{which is the same expression as for the perceptron derived in section \ref{perceptron}}.
+By plugging $\vectorsym{w}$ and $t$ back into the Lagrange equation, we can eliminate these and get the dual optimization problem entirely formulated in terms of the Lagrange multipliers:
 
@@
 
-The dual problem maximizes this function under positivity constaints and one equality constraint:
+The dual problem maximizes this function under positivity constraints and one equality constraint:
 
@@
 This allows them to be in the margin, of even at the wrong side of the boundary -- known as boundary errors.
-Thus, the constaints become $\vectorsym{w} \cdot \vectorsym{x}_i - t \geq 1 - \xi_i$.
+Thus, the constraints become $\vectorsym{w} \cdot \vectorsym{x}_i - t \geq 1 - \xi_i$.
 
-``In summary, \emph{support vector machines are linear classifiers that construct the unique decision boundary that maximises the distance to the nearest training examples (the support vectors)}.
-Training an SVM involves solving a large quadratic optimisation problem and is usually best left to a dedicated numerical solver.''
+``In summary, \emph{support vector machines are linear classifiers that construct the unique decision boundary that maximizes the distance to the nearest training examples (the support vectors)}.
+Training an SVM involves solving a large quadratic optimization problem and is usually best left to a dedicated numerical solver.''
 
@@
 \subsubsection{Density Functions from linear classifiers}\label{linear-classifier-pdf}
-The score of an data point can be used to obtain the signed distance of $\vectorsym{x}_i$ to the dicision boundary:
+The score of an data point can be used to obtain the signed distance of $\vectorsym{x}_i$ to the decision boundary:
 \begin{equation}
@@
 \subsubsection{Support vector based one-class classification}
+Although SVM was originally designed for two-class classification, it has been successfully applied to multi-class and one-class classification.
+SVM-based one-class classification gives the minimum volume closed spherical boundary around the data, represented by center $c$ and radius $r$.
+It minimizes $r^2$ (representing structural error), and uses a penalty coefficient $C$ for each outlier with distance $\xi_i$ from the hyper-sphere boundary:
+\begin{equation}
+\begin{aligned}
+  \text{Min } r^2 + C \sum_{i} \xi_i \\
+  \text{Subject to : } \norm{\vectorsym{x}_i - c}^2 \leq r^2 + \xi_i & & \xi_i \geq 0, & & \forall i, \vectorsym{x}_i: i\text{th data point}
+\end{aligned}
+\end{equation}
+This quadratic optimization problem can be transformed to its dual form by introduction Lagrange multipliers $\alpha_i$.
+If, for a data point, the multiplier $\alpha_i = 0$, then that point is inside the sphere.
+When it is $0 < \alpha_i < C$, then it is on the boundary.
+Data points for which the multiplier is $\alpha_i = C$ are located outside the sphere (and are penalized).
+The dual form is:
+\begin{equation}
+\begin{aligned}
+  \text{Max } \sum_i \alpha_i (\vectorsym{x}_i \cdot \vectorsym{x}_i) - \sum_{i,j} \alpha_i \alpha_j (\vectorsym{x}_i \cdot \vectorsym{x}_j) \\
+  \text{Subject to : } 0 \leq \alpha_i \leq C & & \forall i, & & \sum_i \alpha_i = 1
+\end{aligned}
+\end{equation}
+Note that only dot-products of the data points $\vectorsym{x}$ appear.
+In order to transform the data points to a higher dimension, to create a good representational hyper-sphere, kernels replace the dot products without compromising computational complexity.
+The problem then becomes:
+\begin{equation}
+  \text{Max } \sum_i \alpha_i K(\vectorsym{x}_i, \vectorsym{x}_i) - \sum_{i,j} \alpha_i \alpha_j K(\vectorsym{x}_i, \vectorsym{x}_j)
+\end{equation}
+It has been shown that Gaussian kernels offer better performance for one-class classification the others.
+The optimization of the \emph{scale parameter} has led to several implementations.
+As can been seen, there are no assumptions about the data distribution or independency made.

### 2013-05-23 Appendices/AppendixA-summaries.tex
working on algorithm section

@@
 As can been seen, there are no assumptions about the data distribution or independency made.
+
+\subsubsection{Problem formulation}
+[Not summarized here, useful for e.g. section \ref{statistical-framework}]
+
+\subsubsection{SVCPD: The algorithm}
+Instead of using statistical properties of the data, for each window of size $w$ a hyper-sphere is constructed without increasing computational complexity due to the \emph{kernel trick}.
+The window size is related to the sensitivity of the method to change; small windows are sensitive with high false alarm rate whilst large windows are slow to detect change and have low alarm rates.
+The algorithm is:
+
+\begin{center}\begin{table}
+\begin{tabular}{ l p{12cm} }
+  \hline
+  Step & Action \\
+  \hline
+  1 & Start with $n$ observations and construct hyper-sphere \\
+  2 & Add next observation $x_t$ and drop first one \\
+  3 & Identify new hyper-sphere and its \emph{approximate radius} \\
+  4 & if $x_t$ is outside hyper-sphere, mark $t$ as change points and continue from step 2 \\
+  5 & calculate radius average of last $w$ hyper-planes \\
+  6 & calculate radius ratio $\hbar$. If lower than ${th}_{low}$ or greater than ${th}_{high}$ then mark $t$ as change point \\
+  7 & continue from step 2 \\
+  \hline
+\end{tabular}
+\caption{Support Vector machine based Change Point Detection algorithm}
+\end{table}\end{center}

### 2013-05-23 Chapters/Chapter2.tex
Starting summary Peter Flach
working on algorithm section

@@
 
-\section{Statistical framework}
+\section{Statistical framework}\label{statistical-framework}
 Many applications require the detection of time points at which the underlying properties of a system change.

### 2013-05-23 Chapters/Chapter2/svm.tex
Add introduction Chapter 2
Finish SVM summary

@@
 % !TEX root = ../../main.tex
-\section{Change-detection by Support Vector Machines}
+\section{Change-detection by Support Vector Machines}\label{svm}
 
@@
 \subsection{One-class Support Vector Machine}
+
+*** Replace with / add information from summary in \ref{svm-explained}. ***
+
 The proposed method of Camci \cite{camci2010change} uses a one-class support vector machine to segment time series data.

### 2013-05-24 Appendices/AppendixA-summaries.tex
Finish SVM summary
Summary CUSUM method

@@
 The window size is related to the sensitivity of the method to change; small windows are sensitive with high false alarm rate whilst large windows are slow to detect change and have low alarm rates.
-The algorithm is:
+The algorithm is listed in table \ref{tab:alg-svcdp}.
+Note that SVCPD can be applied directly to multidimensional data, whilst many other methods can only be applied to one-dimensional data.
 
@@
 \caption{Support Vector machine based Change Point Detection algorithm}
+\label{tab:alg-svcdp}
 \end{table}\end{center}
+
+%----------------------------------------------------------------------------------------
+\section{CUSUM for variance}
+
+\subsection{Use of Cumulative Sums of Squares for Retrospective Detection of Changes of Variance}
+Carla Inclan and George C. Tiao \cite{inclan1994use}, 1944, 162 refs.
+
+%--------------------------------------------
+\subsubsection{Introduction}
+This paper is about reflective detection of multiple changes of variance in a sequence of independent observations.
+This is a statistical method, which differs from others (in that field) such as Bayesian method (Bayes ratio, posterior odds), maximum likelihood methods and (autoregressive) models.
+This approach uses the centered version of cumulative sums of squares to search for change points systematically and iteratively (and reflective).
+
+%--------------------------------------------
+\subsubsection{Centered Cumulative Sum of Squares}
+The cumulative sum of squares is often used for change detection in the mean.
+It is defined as $C_k = \sum_{i=1}^k \alpha_t^2$ for a series of uncorrelated random variables $\{\alpha_t\}$ with mean $0$ and variance $\sigma_t^2, t = 1, 2, \dots, T$.
+The centered (and normalized) cumulative sum of squares is:
+\begin{equation}
+  \begin{aligned}
+  D_k = \frac{C_k}{C_T} - \frac{k}{T}, & & k = 1, \dots, T, & & \text{with } D_0 = D_T = 0
+  \end{aligned}
+\end{equation}
+For homogeneous variance the plot of $D_k$ against $k$ (the first $k$ elements of the series) will oscillate around $0$.
+When a sudden change in variance occurs, the pattern of the plot of $D_k$ will break out some specified boundaries with high probability.
+For $C_k$ it holds that, under homogeneous variance, the plot will be a straight line with slope $\sigma^2$.
+With one of more change points the plot appears as a line of several straight pieces.
+The plot of $D_k$ creates a peak for a smaller and a trough for a larger variance, is visually more clear and breaks out a predefined value.
+The search for a change point is variance is than to find $k^* = \text{max}_k \abs{D_k}$.
+If the value of $D_k$ at $k^*$ exceeds a predefined value (e.g.
+$D_{0.5}^* = 1.358$, for $\sqrt{T / 2D_k}$ because of the Brownian bridge property), that value of $k^*$ will be an estimate for a change point.
+
+There is a relation between $D_k$ and the $F$ statistic, which is used for testing equality of variances between two independent samples.
+For a fixed $k$, $D_k(F)$ is a monotone function of $F$ (it depends only on $k$ through $k/T$).
+An important distinction: the $F$ statistic is \emph{used with known $k$}, whereas we are looking for $\text{max}_k \abs{D_k}$ to determine the location of the change point.
+
+When we assume that $\{\alpha_t\}$ is normally distributed with mean $0$ and variances $\sigma_t^2$, then we can obtain the \emph{likelihood ratio} for testing the hypothesis of one change against the hypothesis of no change in the variance.
+When maximizing the likelihood estimator for a location $\kappa$, we can find the log-likelihood ratio $LR_{0,1}$.
+Although $LR_{0,1}$ and $\text{max}_k \abs{D_k}$ are related, they are not the same.
+The latter puts more weight near the middle of the series is thus biased toward $T/2$.
+
+The (expected) value of $D_k$ given a change in variance differs in the context.
+If a smaller variance corresponds to the smaller portion of the series, then it will be harden to find the change point using $D_k$.
+There is a masking effect when there are multiple change points in the series; the order of small, medium and large variances result in the value of $D_k$.
+The iterative algorithm presented in this paper in section \ref{subsec:icss} is designed to lessen the masking effect.
+
+%--------------------------------------------
+\subsubsection{Multiple changes: Iterated cumulative sums of squares}\label{subsec:icss}
+In case of a single change point the $D_k$ method would succeed.
+But we are interested in multiple change points of variance, and thus the usefulness of the $D_k$ reduces due to the masking effect.
+A solution is to iteratively applying the method and dividing the series at each possible change point.
+The algorithm is presented in table \ref{tab:icss}.
+It is the third steps which reduces the masking effect and helps to ``fine tune'' the algorithm by (re)moving the potential change points by checking each location given the adjacent ones.
+
+\begin{center}\begin{table}
+\begin{tabular}{ l p{12cm} }
+  \hline
+  Step & Action \\
+  \hline
+  0 & Let $t_1 = 1$ \\
+  1 & Calculate $D_k(\alpha[t_1 : T])$.
+Let $k^*(\alpha[t_1 : T])$ be the point at which $\text{max}_k \abs{D_k(\alpha[t_1 : T])}$.
+Let $D^*$ be the asymptotically critical value and $M$ the max value in the series segment (?).
+If $M > D^*$ then consider $k^*$ to be a change point.
+Else, there is no change point and the algorithm stops.
+\\
+  2a & Repeat for the first part (up to the change point), until no more change points are found. \\
+  2b & Repeat for the second part (from the change point forward), until no more change points are found. \\
+  3 & When two or more change points are found; check for each $\alpha[j-1 : j+1]$ if there is indeed a change point ($j$).
+Repeat until the number of change points does not change and each new found change point is ``close'' enough to previous.
+\\
+  \hline
+\end{tabular}
+\caption{Iterated Cumulative Sums of Squares (ICSS) Algorithm}
+\label{tab:alg-svcdp}
+\end{table}\end{center}
+
+\subsubsection{Results}
+When the ICSS algorithm was applied to stock data, it resulted in comparable results as the maximum likelihood estimates and Bayesian analysis.
+The performance (CPU-time and correct observations with artificial data) of ICSS outperforms the other two.
+The heavy computational burden of posterior odds can be partially alleviated by the maximum log-likelihood method.
+The ICSS algorithm avoids calculating a function at all possible locations of change points due the iterative manner.

### 2013-05-28 Appendices/AppendixA-summaries.tex
Summary Direct Density Ratio Estimation

@@
 
+%----------------------------------------------------------------------------------------
+%----------------------------------------------------------------------------------------
+%----------------------------------------------------------------------------------------
 %----------------------------------------------------------------------------------------
@@
 %----------------------------------------------------------------------------------------
+%----------------------------------------------------------------------------------------
+%----------------------------------------------------------------------------------------
+%----------------------------------------------------------------------------------------
 \section{CUSUM for variance}
@@
 The ICSS algorithm avoids calculating a function at all possible locations of change points due the iterative manner.
+
+%----------------------------------------------------------------------------------------
+%----------------------------------------------------------------------------------------
+%----------------------------------------------------------------------------------------
+%----------------------------------------------------------------------------------------
+\section{Density Ratio Estimation}
+
+\subsection{Change-Point Detection in Time-Series Data by Direct Density-Ratio Estimation}
+Kawahara and Sugiyama \cite{kawahara2009change}, 2009, 45 refs.
+
+``This paper provides a change-point detection algorithm based on direct density-ratio estimation that can be computed very efficiently in an online manner''.
+
+%--------------------------------------------
+\subsubsection{Introduction}
+The problem of change-point detection is well studied over the last decades in the field of statistics.
+A common statistical formulation of change-point detection is to consider the probability distributions over past and present data intervals, and regard the target time as a change-point if the two distributions are significantly different.
+Some approaches (such as CUSUM and GLR) make use of the \emph{log-likelihood ratio}, and are extensively explored in the data mining community.
+Many approaches (novelty detection, maximum-likelihood ratio, learning of autoregressive models, subspace identification) rely on pre-specified parametric models (probability density models, autoregressive models, state-space models).
+That makes it less applicable to real-life problems.
+There have been some non-parametric density estimation approaches proposed, but that is known to be a hard problem.
+The key idea of this paper to directly estimate the \emph{ratio} of the probability densities (also known as \emph{importance}).
+The Kullback-Leibler Importance Estimation Procedure (KLIEP) is an example, but it is a batch algorithm.
+This paper introduces an online version of the KLIEP algorithm and develops a flexible and computationally efficient change-point detection method.
+This method is equipped with a natural cross validation procedure and thus the value of tuning parameters can be objectively determined.
+
+\subsubsection{Problem formulation and Basic Approach}
+Let $\vectorsym{y}(t) (\in \mathbb{R}^d)$ be a $d$-dimensional time series sample at time $t$.
+The task is to detect whether there exists a change point between two consecutive time intervals, called the \emph{reference} and \emph{test} intervals.
+The conventional algorithms consider the likelihood ratio over samples from the two intervals.
+Since time-series samples generally are not independent over time it is hard to deal with them directly.
+To overcome this difficulty, we consider \emph{sequences} of samples in the intervals: $\matrixsym{Y}(t)(\in \mathbb{R}^{dk})$ is the forward subsequence of length $k$ at time $t$.
+This is a common practice in subspace identification since it takes implicitly time correlation into consideration.
+The algorithm in the paper is based on the logarithm of the likelihood ration of the \emph{sequence sample} $\matrixsym{Y}$:
+\begin{equation}
+  s(\matrixsym{Y}) = \text{ln}\frac{p_\text{te}(\matrixsym{Y})}{p_\text{rf}(\matrixsym{Y})}
+\end{equation}
+where $p_{\text{te}}(\matrixsym{Y})$ and $p_{\text{rf}}(\matrixsym{Y})$ are the probability density functions of the reference and test sequence samples.
+Let $t_{\text{rf}}$ and $t_{\text{te}}$ be the starting points of the reference and test intervals.
+Decide if there is a change-point between the reference and test interval by monitoring the logarithm of the likelihood ratio:
+\begin{equation}
+  S = \sum_{i=1}^{n_\text{te}} \text{ln} \frac{p_{\text{te}}(\matrixsym{Y}_\text{te}(i))}{p_\text{rf}(\matrixsym{Y}_\text{te}(i))}
+\end{equation}
+If, for a predefined $\mu > 0$, it holds that $S > \mu$ then a change occurs.
+The remaining question is how to calculate the density ratio, because it is unknown and we need to estimate it from examples.
+The naive approach would be to first estimate the densities for the reference and test interval separately and then take the ratio.
+This approach via non-parametric density estimation may not be effective --- directly estimating the density ratio without estimating the densities would be more promising.
+
+The direct estimation of the density ratio is based on the Kullback-Leibler Importance Estimation Procedure.
+Let us model the density ratio $w(\matrixsym{Y})$ by a non-parametric Gaussian kernel model:
+\begin{equation}
+  \hat{w}(\matrixsym{Y}) = \sum_{l=1}^{n_\text{te}} \alpha_l K_\sigma(\matrixsym{Y}, \matrixsym{Y}_\text{te}(l)),
+\end{equation}
+where $\{\alpha_l\}_{l=1}^{n_\text{te}}$ are parameters to be learned from the data samples and $K_\sigma(\matrixsym{Y}, \matrixsym{Y}')$ is the Gaussian kernel function with mean $\matrixsym{Y}'$ and standard deviation $\sigma$.
+
+\subsubsection{Online Algorithm}
+...

### 2013-05-30 Chapters/Chapter2.tex
Add literature list to sections

--- 2013-05-13 Chapters/Chapter2.tex
@@
     \begin{itemize}
-      \item Dimensionality reduction
+      \item Introduction with a lot of techniques
+      \item Explain why look at a few
+      \item CUSUM - or other more traditional methods
       \item Density-ratio estimation
       \item Support Vector Machines (?) - if there are more sources about this
-      \item CUSUM - or other more traditional methods
+      \item (Dimensionality reduction)
     \end{itemize}
@@
 
-\section{Statistical framework}
-Many applications require the detection of time points at which underlying properties of a system change.
-This problem thus has received a lot of attention in the fields of data mining, etc... *** list and refs ***.
+% \section{Statistical framework}\label{statistical-framework}
+% Many applications require the detection of time points at which the underlying properties of a system change.
+% This problem thus has received a lot of attention in the fields of data mining, etc... *** list and refs ***.
+% Often this problem is formulated in a statistical framework, by inspecting the data generating PDF (Probability Density Function) of the time series data.
+% A change point is then defined as a significant change in the properties of the PDF, such as the mean and variance.
 
-Some methods, such as the CUSUM (cumulative sum) \cite{basseville1993detection} and GLR (generalized likelihood ratio) \cite{gustafsson1996marginalized, gustafsson2000adaptive} formulate this problem in the statistical framework in the following sense.
-A time series is represented by an underlying probability distribution generating the data.
-The PDFs (probability distribution function) of two consecutive segments around a target point are compared.
-If they differ significantly, the target point is considered as a change point \cite{kawahara2009change}.
+% The widely used CUSUM (cumulative sum) method by Basseville \etal \cite{basseville1993detection}, and the GLR (Generalized Likelihood Ratio) by Gustafsson \cite{gustafsson1996marginalized,gustafsson2000adaptive} take this approach.
+% The former originates from control methods for detection from bench marks.
+% The latter compares the logarithm of the likelihood ratio over two consecutive intervals.
+% These two methods is discussed and analyzed in section \ref{cusum-glr}.
+
+% The GLR method, as with others, relies on pre-specified parametric model assumptions and considers the data be independent over time, which makes it less flexible to real-world applications.
+% The proposed methods by Kawahara \etal \cite{kawahara2009change} and Lui \etal \cite{liu2013change} try to overcome these problems by estimating the \emph{ratio} between the PDF, instead of estimating each PDF.
+% This approach is discussed and analyzed in section \ref{density-ratio}.
+
+% The density-estimation methods, as with the GLR, rely on the log-likelihood ratio between PDFs.
+% The method of Camci \cite{camci2010change} takes an other approach within the statistical framework, by using a SVM (Support Vector Machine).
+% One problem it tries to overcome is the (claimed) weakness of many methods to detect a decrease in variance.
+% The method represents the distribution over the data points as a hyper-sphere in a higher dimension using kernel trick.
+% A change in the PDF is represented by a change in the radius of this sphere.
+% Section \ref{svm} discusses the SVM-method.
+
+% The final method under consideration, change detection via (intrinsic) dimensionality reduction, takes a different point of view.
+% Opposed to the other discussed methods, dimensionality reduction is framed in the MDL (Minimum Description Length) framework.
+% It uses the estimated underlying number of parameters of the time series as a model for change detection.
+% Section \ref{dim-reduction} discuses this method.
 
 %----------------------------------------------------------------------------------------
-\section{Change-detection by Dimensionality Reduction / Covariance structure}
+% CUSUM and GLR
+\input{./Chapters/Chapter2/cusum_glr}
 
 %----------------------------------------------------------------------------------------
+% Density Ratio Estimation
 \input{./Chapters/Chapter2/density_ratio}
@@
 %----------------------------------------------------------------------------------------
+% Support Vector Machines
 \input{./Chapters/Chapter2/svm}
@@
 %----------------------------------------------------------------------------------------
-\section{Change-detection by Cumulative Sum}
+% Dimensionality reduciton
+\input{./Chapters/Chapter2/dim_reduction}

### 2013-05-30 Chapters/Chapter2/cusum_glr.tex
Add literature list to sections

@@
 *** read book Basseville \cite{basseville1993detection} ***
+
+Papers:
+\begin{itemize}
+  \item Use of Cumulative Sums of Squares for Retrospective Detection of Changes of Variance, 1994, 162 refs \cite{inclan1994use} \\
+  Implemented this algorithm.
+  \item An adaptive CUSUM-based test for signal change detection, 2006, 15 refs. \cite{alippi2006adaptive}
+  \item The MOSUM of squares test for monitoring variance changes \cite{hsu2007mosum}
+\end{itemize}
 

### 2013-05-30 Chapters/Chapter2/density_ratio.tex
Add introduction Chapter 2
Add literature list to sections

@@
 % !TEX root = ../../main.tex
-\section{Change-detection by Density-Ratio Estimation}
+\section{Change-detection by Density-Ratio Estimation}\label{density-ratio}
+
+Papers:
+\begin{itemize}
+  \item Change-Point Detection in Time-Series Data by Direct Density-Ratio Estimation, 2009, 45 refs \cite{kawahara2009change}
+  \item Change-Point Detection in Time-Series Data by Relative Density-Ratio Estimation, 2013, 3 refs \cite{liu2013change}
+  \item Density ratio estimation in machine learning (Book), 2009, 24 refs \cite{sugiyama2012density}
+\end{itemize}
 

### 2013-05-31 Chapters/Chapter2/cusum.tex
Ch2. Sec. CUSUM

% !TEX root = ../../main.tex
\section{CUSUM}\label{cusum}

*** read book Basseville \cite{basseville1993detection} ***

Papers:
\begin{itemize}
  \item Use of Cumulative Sums of Squares for Retrospective Detection of Changes of Variance, 1994, 162 refs \cite{inclan1994use} \\
  Implemented this algorithm.
  \item An adaptive CUSUM-based test for signal change detection, 2006, 15 refs. \cite{alippi2006adaptive}
  \item The MOSUM of squares test for monitoring variance changes \cite{hsu2007mosum}
\end{itemize}

\subsection{CUSUM}
Non-Bayesian change detection algorithm (thus: no prior distribution believe available for the change time).
The CUSUM (cumulative sum) method is developed by Page \cite{page1954continuous} for the application of statistical quality control (it is also known as a control chart).
Primary for detection of mean shift.
The MOSUM of squares test for monitoring variance changes \cite{hsu2007mosum}.
Use of Cumulative Sums of Squares for Retrospective Detection of Changes of Variance \cite{inclan1994use}

An other used approach in the statistical framework of change detection is the Cumulative Sum of Squares (CUSUM) as introduced by Page \cite{page1954continuous}.
Originally used for quality control in production environments, its main function is to detect change in the mean of measurements.
Many extensions to this method have been proposed.
Some focus on the change in mean, such the method of Alippi and Roveri \cite{alippi2006adaptive}.
Others apply the method the problems in which the change of variance is under consideration.
Among others are there the centered version of the cumulative sums, introduced by Brown, Durbin and Evans \cite{brown1975techniques} and the Moving Sum (MOSUM) of squares by \cite{hsu2007mosum}.

The method of Incl\'{a}n and Tiao \cite{inclan1994use} builds on the centered version of CUSUM to detect changes in variance.
Using an iterative method algorithm --- Iterative Cumulative Sum of Squares (ICSS), they are able to find multiple change points in a reflective manner.
Let $C_k = \sum_{i=1}^k \alpha_t^2$ be the cumulative sum of squares for a series of uncorrelated random variables $\{\alpha_t\}$ of length $T$.
The centered (and normalized) sum is squares is defined as
\begin{equation}
  \begin{aligned}
  D_k = \frac{C_k}{C_T} - \frac{k}{T}, & & k = 1, \dots, T, & & \text{with } D_0 = D_T = 0.
  \end{aligned}
\end{equation}
For a series with homogeneous variance, the value of $D_k$ will oscillate around $0$.
When there is a sudden change, the value will increase and exceed some predefined boundary with high probability.
The behavior of $D_k$ is related a Brownian bridge.
By using an iterative algorithm, the method is able to minimize the masking effect of successive change points.

One of the motivations for the ICSS algorithm was the heavy computational burden involved with Bayesian methods, which need to calculate the posterior odds for the log-likelihood ratio testing.
The ICSS algorithm avoids applying a function at all possible locations, due to the iterative search.
The authors recommend the algorithm for analysis of long sequences.

### 2013-05-31 Chapters/Chapter2.tex
Ch2. Sec. CUSUM
Work on Ch2

--- 2013-05-23 Chapters/Chapter2.tex
@@
 
+*** Remove notion of GLR ***
+
 The widely used CUSUM (cumulative sum) method by Basseville \etal \cite{basseville1993detection}, and the GLR (Generalized Likelihood Ratio) by Gustafsson \cite{gustafsson1996marginalized,gustafsson2000adaptive} take this approach.
@@
 The latter compares the logarithm of the likelihood ratio over two consecutive intervals.
-These two methods is discussed and analyzed in section \ref{cusum-glr}.
+These two methods is discussed and analyzed in section \ref{cusum}.
 
@@
 
-The final method under consideration, change detection via (intrinsic) dimensionality reduction, takes a different point of view.
-Opposed to the other discussed methods, dimensionality reduction is framed in the MDL (Minimum Description Length) framework.
-It uses the estimated underlying number of parameters of the time series as a model for change detection.
-Section \ref{dim-reduction} discuses this method.
+% The final method under consideration, change detection via (intrinsic) dimensionality reduction, takes a different point of view.
+% Opposed to the other discussed methods, dimensionality reduction is framed in the MDL (Minimum Description Length) framework.
+% It uses the estimated underlying number of parameters of the time series as a model for change detection.
+% Section \ref{dim-reduction} discuses this method.
 
@@
 % CUSUM and GLR
-\input{./Chapters/Chapter2/cusum_glr}
+\input{./Chapters/Chapter2/cusum}
 
@@
 % Dimensionality reduciton
-\input{./Chapters/Chapter2/dim_reduction}
+% \input{./Chapters/Chapter2/dim_reduction}

### 2013-05-31 Chapters/Chapter2/cusum.tex
Work on Ch2

@@
 An other used approach in the statistical framework of change detection is the Cumulative Sum of Squares (CUSUM) as introduced by Page \cite{page1954continuous}.
-Originally used for quality control in production environments, its main function is to detect change in the mean of measurements.
+Originally used for quality control in production environments, its main function is to detect change in the mean of measurements and has been applied to this problem \cite{basseville1993detection}.
 Many extensions to this method have been proposed.

### 2013-05-31 Chapters/Chapter2/density_ratio.tex
Work on Ch2

@@
   \item Density ratio estimation in machine learning (Book), 2009, 24 refs \cite{sugiyama2012density}
+  \item Direct importance estimation for covariate shift adaptation, 2008, 84 refs \cite{sugiyama2008direct}
 \end{itemize}
 
-Formulate the problem of detecting change in the statistical framework.
-Consider the probability distributions from which two consecutive segments of time series around a target time point are generated.
-When the disitrubtions differ significantly the target time point is regarded as a change point.
+% Formulate the problem of detecting change in the statistical framework.
+% Consider the probability distributions from which two consecutive segments of time series around a target time point are generated.
+% When the disitrubtions differ significantly the target time point is regarded as a change point.
 
-CUSUM (cumulative sum) \cite{basseville1993detection} and GLR (generalized likelihood ratio)
+% CUSUM (cumulative sum) \cite{basseville1993detection} and GLR (generalized likelihood ratio)
 
-The distribution over the values of time series data can be represented with a probability density function (pdf).
-Two sections of a time series data can be generated with the same underlying pdf or each with a different.
+% The distribution over the values of time series data can be represented with a probability density function (pdf).
+% Two sections of a time series data can be generated with the same underlying pdf or each with a different.
+
+Many approaches to detect change points, regarded as a change in the underlying probabilistic generation, monitor the logarithm of the likelihood ratio between two consecutive intervals.
+Some methods which rely on this are novelty detection, maximum-likelihood estimation and online learning of autoregressive models \cite{kawahara2009change}.
+A limitation of these methods is that they rely on pre-specified parametric models.
+Non-parametric models for density estimation have been proposed, but it is said to be a hard problem.
+A solutions to this is to estimate the \emph{ratio} of probabilities instead of the probabilities themselves.
+One of the recent methods to achieve this is the Kullback-Leibler Importance Estimation Procedure (KLIEP) by Sugiyama \etal \cite{sugiyama2008direct}.
+
+The method proposed by Kawahara and Sugiyama \cite{kawahara2009change} is composed of an online version of the KLIEP algorithm.
+The method also considers \emph{sequences} of samples (rather than samples directly) because the time series samples are generally not independent over time.
+An advantage over other non-parametric approaches, such as sequential one-class support vector machines is that the model has an natural cross-validation procedure.
+This makes that the value of tuning parameters, such as the kernel bandwidth, can be objectively obtained.
+
+In their formulation change is detected by monitoring the logarithm of the likelihood ratio between the reference (past) and test (current) time intervals
+\begin{equation}
+  S = \sum_{i=1}^{n_\text{te}} \text{ln} \frac{p_{\text{te}}(\matrixsym{Y}_\text{te}(i))}{p_\text{rf}(\matrixsym{Y}_\text{te}(i))}
+\end{equation}
+Where $\matrixsym{Y}_\text{te}(i)$ is a sequence of samples from the test interval.
+A change is detected when $S > \mu$, for some predetermined threshold $\mu$.
+The question is then to how calculate the density ratio
+\begin{equation}
+  w(\matrixsym{Y}) := \frac{p_\text{te}(\matrixsym{Y)})}{p_\text{rf}(\matrixsym{Y})}
+\end{equation}
+because this ratio is unknown and should be estimated.
+The naive approach is to estimate the ratio by taking the ratio of the estimated densities.
+Since this is known to be a hard problem and sensitive for errors, the solution would be to estimate the ratio directly.
+
+The procedure of the method proposed by Kawahara and Sugiyama \cite{kawahara2009change} is to first apply the batch KLIEP algorithm with model selection for initial parameter $\vectorsym{\alpha}$ and kernel width calculation.
+Then for every new sample the reference and test intervals are shifted and the calculated parameters $\vectorsym{\alpha}$ are updated.
+Finally the logarithm of the likelihood ratio is evaluated.
+If it is beyond the threshold $\mu$ the current time is reported as a change point.
+
+Improvements in this line of research, by Liu \etal \cite{liu2013change} has lead the application of improved density-ratio estimation methods to the problem of change detection.
+Such an improvement is the Unconstrained Least-Squares Importance Fitting (uLSIF) method \cite{kanamori2009least} and an extension which possess a superior non-parametric convergence property: Relative uLSIF (RuLSIF) \cite{yamada2013relative}.

### 2013-05-31 Chapters/Chapter2/svm.tex
Work on Ch2

@@
 Introduced by Vapnik \cite{vapnik1998statistical, vapnik1999nature}, Support Vector Machines offer a way to segment, and classify, linear separable data.
+This technique can also be applied to estimate density functions of given time series \cite{weston1999support}.
 When combined with a mapping function, which maps the data from the input space $I$ to a higher dimension feature space $F$, the input data can be non-linear separable.
@@
 
-*** Let sigma be a mapping from I to F such that the dot product in F can be computed using some simple kernel ***
+The method of Camci \cite{camci2010change} uses one-class support vector machines to segment time series data.
+For the current interval a hyper-sphere is constructed, using the kernel trick, which circumscribes all the data.
+To cope with possible errors or outliers a soft-margin is applied.
+The data can thus be represented using a center $c_1$ and radius $r_1$ of the hyper-sphere.
+New data points are consecutively mapped from the input space to the feature space.
+The retrieved (high dimension) data point can be in- or outside the earlier constructed hyper-sphere, thereby giving information about a possible change point.
 
 \subsection{One-class Support Vector Machine}
+*** This section is too detailed for the Literature Review, but keep it for now ***
 
@@
 \begin{equation}
-  c_1 = c_2, r_1 = r_2
+\begin{aligned}
+  c_1 = c_2, & &  r_1 = r_2
+\end{aligned}
 \end{equation}

### 2013-06-04 Chapters/Chapter2.tex
Rough finish of chapter 2

@@
 \section{Outline}
+\emph{Not intented for the reader.}
 \begin{itemize}
@@
 
-*** Remove notion of GLR ***
+The widely used CUSUM (cumulative sum) method by Basseville \etal \cite{basseville1993detection} takes this approach.
+It originates from control methods for detection from bench marks.
+This method and some derivatiges are discussed and analyzed in section \ref{cusum}.
 
-The widely used CUSUM (cumulative sum) method by Basseville \etal \cite{basseville1993detection}, and the GLR (Generalized Likelihood Ratio) by Gustafsson \cite{gustafsson1996marginalized,gustafsson2000adaptive} take this approach.
-The former originates from control methods for detection from bench marks.
-The latter compares the logarithm of the likelihood ratio over two consecutive intervals.
-These two methods is discussed and analyzed in section \ref{cusum}.
-
-The GLR method, as with others, relies on pre-specified parametric model assumptions and considers the data be independent over time, which makes it less flexible to real-world applications.
+Many methods rely on pre-specified parametric model assumptions and considers the data be independent over time, which makes it less flexible to real-world applications.
 The proposed methods by Kawahara \etal \cite{kawahara2009change} and Lui \etal \cite{liu2013change} try to overcome these problems by estimating the \emph{ratio} between the PDF, instead of estimating each PDF.
@@
 
-The density-estimation methods, as with the GLR, rely on the log-likelihood ratio between PDFs.
+The density-estimation methods rely on the log-likelihood ratio between PDFs.
 The method of Camci \cite{camci2010change} takes an other approach within the statistical framework, by using a SVM (Support Vector Machine).
@@
 % Section \ref{dim-reduction} discuses this method.
+
+*** Maybe add dimensionality reduction, but for now leave out ***
 

### 2013-06-04 Chapters/Chapter2/cusum.tex
Rough finish of chapter 2

@@
 
-*** read book Basseville \cite{basseville1993detection} ***
+% *** read book Basseville \cite{basseville1993detection} ***
 
-Papers:
-\begin{itemize}
-  \item Use of Cumulative Sums of Squares for Retrospective Detection of Changes of Variance, 1994, 162 refs \cite{inclan1994use} \\
-  Implemented this algorithm.
-  \item An adaptive CUSUM-based test for signal change detection, 2006, 15 refs. \cite{alippi2006adaptive}
-  \item The MOSUM of squares test for monitoring variance changes \cite{hsu2007mosum}
-\end{itemize}
+% Papers:
+% \begin{itemize}
+%   \item Use of Cumulative Sums of Squares for Retrospective Detection of Changes of Variance, 1994, 162 refs \cite{inclan1994use} \\
+%   Implemented this algorithm.
+%   \item An adaptive CUSUM-based test for signal change detection, 2006, 15 refs. \cite{alippi2006adaptive}
+%   \item The MOSUM of squares test for monitoring variance changes \cite{hsu2007mosum}
+% \end{itemize}
 
-\subsection{CUSUM}
 Non-Bayesian change detection algorithm (thus: no prior distribution believe available for the change time).
@@
 
-An other used approach in the statistical framework of change detection is the Cumulative Sum of Squares (CUSUM) as introduced by Page \cite{page1954continuous}.
+An often used approach in the statistical framework of change detection is the Cumulative Sum of Squares (CUSUM) as introduced by Page \cite{page1954continuous}.
 Originally used for quality control in production environments, its main function is to detect change in the mean of measurements and has been applied to this problem \cite{basseville1993detection}.
+It is an non-Bayesian method and thus makes no assumptions (or: prior belief distributions) for the change points.
 Many extensions to this method have been proposed.
@@
 
-The method of Incl\'{a}n and Tiao \cite{inclan1994use} builds on the centered version of CUSUM to detect changes in variance.
-Using an iterative method algorithm --- Iterative Cumulative Sum of Squares (ICSS), they are able to find multiple change points in a reflective manner.
+The method of Incl\'{a}n and Tiao \cite{inclan1994use} builds on the centered version of CUSUM \cite{brown1975techniques} to detect changes in variance.
+Using the Iterative Cumulative Sum of Squares (ICSS) algorithm they are able to find multiple change points in a reflective manner.
 Let $C_k = \sum_{i=1}^k \alpha_t^2$ be the cumulative sum of squares for a series of uncorrelated random variables $\{\alpha_t\}$ of length $T$.

### 2013-06-04 Chapters/Chapter2/density_ratio.tex
Rough finish of chapter 2

@@
 
-Papers:
-\begin{itemize}
-  \item Change-Point Detection in Time-Series Data by Direct Density-Ratio Estimation, 2009, 45 refs \cite{kawahara2009change}
-  \item Change-Point Detection in Time-Series Data by Relative Density-Ratio Estimation, 2013, 3 refs \cite{liu2013change}
-  \item Density ratio estimation in machine learning (Book), 2009, 24 refs \cite{sugiyama2012density}
-  \item Direct importance estimation for covariate shift adaptation, 2008, 84 refs \cite{sugiyama2008direct}
-\end{itemize}
+% Papers:
+% \begin{itemize}
+%   \item Change-Point Detection in Time-Series Data by Direct Density-Ratio Estimation, 2009, 45 refs \cite{kawahara2009change}
+%   \item Change-Point Detection in Time-Series Data by Relative Density-Ratio Estimation, 2013, 3 refs \cite{liu2013change}
+%   \item Density ratio estimation in machine learning (Book), 2009, 24 refs \cite{sugiyama2012density}
+%   \item Direct importance estimation for covariate shift adaptation, 2008, 84 refs \cite{sugiyama2008direct}
+% \end{itemize}
 
@@
 A limitation of these methods is that they rely on pre-specified parametric models.
-Non-parametric models for density estimation have been proposed, but it is said to be a hard problem.
+Non-parametric models for density estimation have been proposed, but it is said to be a hard problem \cite{hardle2004nonparametric, sugiyama2012density}.
 A solutions to this is to estimate the \emph{ratio} of probabilities instead of the probabilities themselves.
@@
 The method also considers \emph{sequences} of samples (rather than samples directly) because the time series samples are generally not independent over time.
-An advantage over other non-parametric approaches, such as sequential one-class support vector machines is that the model has an natural cross-validation procedure.
+An advantage over other non-parametric approaches, such as sequential one-class support vector machines, is that the model has an natural cross-validation procedure.
 This makes that the value of tuning parameters, such as the kernel bandwidth, can be objectively obtained.

### 2013-06-04 Chapters/Chapter2/svm.tex
Rough finish of chapter 2

@@
 
-Introduced by Vapnik \cite{vapnik1998statistical, vapnik1999nature}, Support Vector Machines offer a way to segment, and classify, linear separable data.
+Introduced by Cortes and Vapnik \cite{vapnik1998statistical, vapnik1999nature}, Support Vector Machines offer a way to segment, and classify, linear separable data.
 This technique can also be applied to estimate density functions of given time series \cite{weston1999support}.
@@
 
-The method of Camci \cite{camci2010change} uses one-class support vector machines to segment time series data.
-For the current interval a hyper-sphere is constructed, using the kernel trick, which circumscribes all the data.
-To cope with possible errors or outliers a soft-margin is applied.
-The data can thus be represented using a center $c_1$ and radius $r_1$ of the hyper-sphere.
+The proposed method of Camci \cite{camci2010change} uses a one-class support vector machine to segment time series data.
+One-class SVMs are used to describe the current data under consideration, by assuming all data points are from the same class \cite{tax2001one}.
+To cope with possible errors or outliers a soft-margin is applied \cite{cortes1995support}.
+The class is described by a spherical boundary around the data with center $c$ and radius $r$, such that the volume is minimized.
 New data points are consecutively mapped from the input space to the feature space.
 The retrieved (high dimension) data point can be in- or outside the earlier constructed hyper-sphere, thereby giving information about a possible change point.
-
-\subsection{One-class Support Vector Machine}
-*** This section is too detailed for the Literature Review, but keep it for now ***
-
-*** Replace with / add information from summary in \ref{svm-explained}. ***
-
-The proposed method of Camci \cite{camci2010change} uses a one-class support vector machine to segment time series data.
-One-class SVMs are used to describe the current data under consideration, by assuming all data points are from the same class \cite{tax2001one}.
-The class is described by a spherical boundary around the data with center $c$ and radius $r$, such that the volume is minimized.
 Following the definition of Camci \cite{camci2010change}, the class description is obtained by minimizing $r^2$:
@@
 To be able to handle outliers in the input data, a penalty cost function $\varepsilon_i$ for each outlier can be added.
-
-*** Add new function and constraints? ***
 
@@
 
-*** First tell more about (underlying) probability density functions, to relate to other methods ***
-
 In case the second window of data points does not belong to the same class, i.e.
 the probability density function that describes the data differs from the first, the describing values of the second window will differ from the first.
+A difference in the SVM center $c$ or radius $r$ represent a change in the mean and variance, respectively.
 The amount of difference can be expressed by a dissimilarity measure over the representations.
@@
 
-This process can be visualized as done in *** insert figure of four circles ***.
-The second window, $h_2$ can be constructed from the first by e.g.
-a shift of one data point.
-*** explain data point positions by circle ***.
-
-Note that a difference in the SVM center $c$ or radius $r$ represent a change in the mean and variance, respectively.
+*** Give visual explanation with circles and in- and out-side new data points ***

### 2013-06-27 Appendices/AppendixB-camci.tex
Putting Camci-flaws in text

% !TEX root = ../main.tex
% Appendix B

\chapter{Session with Anne 24-06-2013 - Paper Camci analysis} % Main appendix title

\label{AppendixB} % For referencing this appendix elsewhere, use \ref{AppendixA}

\lhead{Appendix B. \emph{Paper Camci}} % This is for the header on each page - perhaps a shortened title

\emph{Please ignore this Appendix.
This appendix is for my own personal use.
This chapter will look at the paper of Camci \cite{camci2010change} (``Change point detection in time series using support vectors'') and will answer many question that the paper leaves open.
The goal is to make a better justification for the used techniques and made assumptions.
}
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------

\section{Density estimation / Data description Vapnik's principle}
Short look in density estimation. Other techniques to perform this.

Vapnik: ``do not solve a more complicated intermediate problem''.

%---------
Following Vapnik's principle, one should ``When solving a problem of interest, do not solve a more general problem as an intermediate step.'' \cite{vapnik1998statistical}.
In \cite{liu2013change} it is stated that the \emph{Support Vector Machine} by Cortes and Vapnik \cite{cortes1995support} is representative example of this principle.
Instead of estimating the more general data generating probability distributions, it only learns a decision boundary to differentiate between the two distributions.

The proposed method SVDD of Tax and Duin \cite{tax1999support, tax2004support} models the boundary of data under consideration.
Thereby it characterizes a data set and can be used to detect novel data or outliers.
The performance is compared to methods which model the distribution's density instead, using Receiver-Operating Characteristic (ROC) curves and false negative rates.
The compared methods are:
\begin{inparaenum}[\itshape (1)]
  \item normal density which estimates the mean and covariance matrix;
  \item the Parzen density where the width of the Parzen kernel is estimated;
  \item a Gaussian Mixture Model optimized using EM; and
  \item the Nearest-Neighbor Method which compares the local density of an object with the density of the nearest neighbor in the target set.
\end{inparaenum}
The results show that when the problem formulation is to characterize an area in a feature space (and not the complete density distribution) SVDD gives a good data description.

The method of Camci \cite{camci2010change} uses a Support Vector based method to find change points in the data.
It does not explicitly create a density estimation, but instead relies on the spherical boundary and uses its ability to detect novel data or outliers to detect change points in time series data.

*** Thus methods of data descriptions should be compared (which are more general) and not only density estimation ***

\section{Change point definition}
Periodicity/consecutive data vs unique/irregular data?

Change in model parameters (mean/variance, of linear/non-linear)? --> Model selection?

Definition of continuity --> windows the domain and problem.
Will result in definition of dis-continuity --> this is the goal to find
Relation with double differentiation.

\section{Data and model}
Why assume (model of) data is Gaussian/normal distribution? Thus, piecewise linear with mean and variance as changing properties. Why not set of polynomial models, as for example in \cite{chamroukhi2013joint}?

What is the best model for accelerometer data of human activities? --> Look to result of model-selecting papers.

Should we build a model of the data? (Gaussian distribution is also the model). And be able to reconstruct?

Why is it better (as Camci states) that a method makes no assumptions about the form of data/distribution of data? Is it that there are less parameters to estimate?

\section{Segmentation (SVM) method}
Why use SVM for density estimation? Look for justification, perhaps a review paper which compares/mentions SVM for temporal segmentation of (human activity type of) data?

Why use RBF/Gaussian kernel? Is it because of one-class SVM of because of form the data? Why not polynomial/linear?
\begin{itemize}
  \item ``Use RBF when relation between class and data is non-linear''
  \item ``RBF uses less paramters (C for penalty/soft margin and gamme for kernel width) than non-linear polynomial kernels''
  \item (arguments from \cite{zhang2011optimal} ``Optimal model selection for posture recognition in home-based healthcare'')
\end{itemize}

\subsection{Higher dimension mapping (including kernel)}
What does the mapping from the data-space to higher dimension looks like?

What is a RBF kernel?

What form has the higher dimensional space?

How do relations over data, such as distance, volume and noise, act in the higher dimensional space?

What is the kernel trick to not explicit do the mapping to the higher space?

\section{Relation to other methods}

\subsection{Novelty/outlier detection}
\subsection{Scale parameter}
\subsection{Robust statistics / "M-Estimators"}
Is the method robust, in the sense that outliers have restricted impact on the quality.

What is the relation to M-Estimators (Wikipedia: ``M-estimators are a broad class of estimators, which are obtained as the minima of sums of functions of the data'', http://en.wikipedia.org/wiki/M-estimator)

\section{Quality metrics}
As used in Camci: benefit, false alarm rate

Asymmetric test: feed data from front-to-back and back-to-front; how far are matched datapoint apart.

Model reconstruction error: try to reconstruct the simulated models, test similarity (BIC?). Is that a good measure? (If we are only interested in finding change points...)

### 2013-06-27 Appendices/AppendixB-camci.tex
Adding justifications to Camci

@@
 }
+
 %----------------------------------------------------------------------------------------
 %----------------------------------------------------------------------------------------
+%----------------------------------------------------------------------------------------
+%----------------------------------------------------------------------------------------
+\section{Density estimation / Data description / Vapnik's principle}
+Following Vapnik's principle, one should ``When solving a problem of interest, do not solve a more general problem as an intermediate step.'' \cite{vapnik1998statistical} when a limited amount of data is available.
+For the problem of change detection we are only interested in some characteristics of the data.
+Solving the complete density estimation might require more data than actually needed when the requested characteristic is a closed boundary around the data.
 
-\section{Density estimation / Data description Vapnik's principle}
-Short look in density estimation. Other techniques to perform this.
-
-Vapnik: ``do not solve a more complicated intermediate problem''.
-
-%---------
-Following Vapnik's principle, one should ``When solving a problem of interest, do not solve a more general problem as an intermediate step.'' \cite{vapnik1998statistical}.
 In \cite{liu2013change} it is stated that the \emph{Support Vector Machine} by Cortes and Vapnik \cite{cortes1995support} is representative example of this principle.
@@
   \item a Gaussian Mixture Model optimized using EM; and
-  \item the Nearest-Neighbor Method which compares the local density of an object with the density of the nearest neighbor in the target set.
+  \item the Nearest-Neighbor Method which compares the local density of an object with the density of the nearest neighbor in the target set (and is thus, just as SVDD, a boundary-based method).
 \end{inparaenum}
 The results show that when the problem formulation is to characterize an area in a feature space (and not the complete density distribution) SVDD gives a good data description.
+
+The study of Tax on one-class classifiers \cite{tax2001one} further compares density methods, boundary methods (amongst which SVDD) and reconstruction methods.
+One promesing result for SVDD is that it performs well on read-world data, for which generalization is needed.
 
@@
 
-*** Thus methods of data descriptions should be compared (which are more general) and not only density estimation ***
+*** Thus methods of data descriptions should be compared (which are more general) and not only density estimation? ***
 
+%----------------------------------------------------------------------------------------
+%----------------------------------------------------------------------------------------
+%----------------------------------------------------------------------------------------
+%----------------------------------------------------------------------------------------
 \section{Change point definition}
+The method of Camci \cite{camci2010change} regards a change point as the moment in time that the underlying stochastic process has changed, say from $p^1$ to $p^2$.
+It assumes that each of these stochastic processes is modeled following a Gaussian distribution, such that a change can occur in the value of the mean and/or the variance; $p^1 \sim N(\mu_1, \sigma_1^2)$.
+The CUSUM-based method of \cite{inclan1994use} also regards each semgent as a Gaussian distribution.
+
+The method if Kawahara \etal \cite{kawahara2009change} is based on the log likelihood ratio of test samples, and the method by Liu \etal \cite{liu2013change} uses a comparable dissimilarity measure using the KLIEP algorithm.
+
+The method of Chamroukhi \etal \cite{chamroukhi2013joint} is based on a Hidden Markov Model and logistic regression.
+It assumes a $K$-state hidden process with a (hidden) state sequence, each state providing the parameters (amongst which the order) for a polynomial.
+The order of the model segment is determined by model selecting, often using the Bayesian Information Criterion (BIC) or the similair Akaike Information Criterion (AIC) \cite{akaike1974new}, as in \cite{he2008activity}.
+
+*** REVIEW zoeken: change points in time series ***
+
 Periodicity/consecutive data vs unique/irregular data?
@@
 
+%----------------------------------------------------------------------------------------
+%----------------------------------------------------------------------------------------
+%----------------------------------------------------------------------------------------
+%----------------------------------------------------------------------------------------
 \section{Data and model}
@@
 Why is it better (as Camci states) that a method makes no assumptions about the form of data/distribution of data? Is it that there are less parameters to estimate?
+%----------------
 
+Many methods describe and compare methods to construct classifier models for the classification of accelerometer data, such as \cite{kwapisz2011activity} and \cite{zhang2011optimal} (often using extracted features from the raw data signal).
+In contrast, we could not find a clear characterization of accelerometer data obtained from human activities.
+When the problem of temporal segmentation is regarded in this context, a formalization of the data under consideration is needed.
+Some assume the data follows a piecewise linear set of segments with a mean and noise/variance modeled by a normal (Gaussian) distribution (such as \cite{camci2010change}).
+Other approaches regard the data as a set of polynomials, which can be estimated by regression (such as \cite{chamroukhi2013joint}) and apply a form of model selection to each segment.
+
+%----------------------------------------------------------------------------------------
+%----------------------------------------------------------------------------------------
+%----------------------------------------------------------------------------------------
+%----------------------------------------------------------------------------------------
 \section{Segmentation (SVM) method}
+Overview of segmentation methods.
+Collection of papers use relative and direct density-ratio estimation \cite{kawahara2009change, liu2013change}.
+
 Why use SVM for density estimation? Look for justification, perhaps a review paper which compares/mentions SVM for temporal segmentation of (human activity type of) data?
@@
 
+%----------------------------------------------------------------------------------------
+%----------------------------------------------------------------------------------------
+%----------------------------------------------------------------------------------------
+%----------------------------------------------------------------------------------------
 \section{Relation to other methods}
@@
 
+%----------------------------------------------------------------------------------------
+%----------------------------------------------------------------------------------------
+%----------------------------------------------------------------------------------------
+%----------------------------------------------------------------------------------------
 \section{Quality metrics}

### 2013-06-28 Appendices/AppendixB-camci.tex
Added refs, comparing stuff

@@
 
-*** Thus methods of data descriptions should be compared (which are more general) and not only density estimation? ***
+*** \emph{Thus methods of data descriptions should be compared (which are more general) and not only density estimation?} ***
+
+The paper of Yin \etal \cite{yin2008sensor} makes a distinction between similarity based (using a defined distance measure) and model based (which characterize the data using predictive models) approach.
+The one-class SVM is used in the model-approach to filter out normal activities in order to detect abnormality behavior.
 
@@
 
+*** TODO: make a clear distinction between model types; similarity (distance based) or model based, as in \cite{yin2008sensor} ***
+
 %----------------------------------------------------------------------------------------
@@
 \end{itemize}
+
+%-------------------------
+
+Re-evaluatie \cite{guo2012adaptive}, uses (amongst others) a pca-based dimension-detection method (?).
+Also uses model-selection as an intermediate step.
 
@@
 \subsection{Novelty/outlier detection}
+*** Read \cite{hodge2004survey} ``A survey of outlier detection methodologies'' to compare with other methodologies. ***
+
 \subsection{Scale parameter}

### 2013-07-03 Appendices/AppendixA-summaries.tex
Work on Ch2
Rough finish of chapter 2
Adding acronym-glosarries

@@
 
-Please ignore this Appendix.
+\emph{Please ignore this Appendix.
 This appendix is for my own personal use.
 It contains summaries of articles I have read.
-
+}
 %----------------------------------------------------------------------------------------
@@
 A solution is to iteratively applying the method and dividing the series at each possible change point.
-The algorithm is presented in table \ref{tab:icss}.
+The algorithm is presented in table \ref{tab:alg-icss}.
 It is the third steps which reduces the masking effect and helps to ``fine tune'' the algorithm by (re)moving the potential change points by checking each location given the adjacent ones.
@@
 \end{tabular}
-\caption{Iterated Cumulative Sums of Squares (ICSS) Algorithm}
-\label{tab:alg-svcdp}
+\caption{Iterated Cumulative Sums of Squares Algorithm}
+\label{tab:alg-icss}
 \end{table}\end{center}
@@
 \subsubsection{Results}
-When the ICSS algorithm was applied to stock data, it resulted in comparable results as the maximum likelihood estimates and Bayesian analysis.
-The performance (CPU-time and correct observations with artificial data) of ICSS outperforms the other two.
+When the \gls{icss} algorithm was applied to stock data, it resulted in comparable results as the maximum likelihood estimates and Bayesian analysis.
+The performance (CPU-time and correct observations with artificial data) of \gls{icss} outperforms the other two.
 The heavy computational burden of posterior odds can be partially alleviated by the maximum log-likelihood method.
-The ICSS algorithm avoids calculating a function at all possible locations of change points due the iterative manner.
+The \gls{icss} algorithm avoids calculating a function at all possible locations of change points due the iterative manner.
 
@@
 A common statistical formulation of change-point detection is to consider the probability distributions over past and present data intervals, and regard the target time as a change-point if the two distributions are significantly different.
-Some approaches (such as CUSUM and GLR) make use of the \emph{log-likelihood ratio}, and are extensively explored in the data mining community.
+Some approaches (such as \gls{cusum} and GLR) make use of the \emph{log-likelihood ratio}, and are extensively explored in the data mining community.
 Many approaches (novelty detection, maximum-likelihood ratio, learning of autoregressive models, subspace identification) rely on pre-specified parametric models (probability density models, autoregressive models, state-space models).

### 2013-07-03 Appendices/AppendixB-camci.tex
wip
Adding acronym-glosarries

@@
 
-The proposed method SVDD of Tax and Duin \cite{tax1999support, tax2004support} models the boundary of data under consideration.
+The proposed method \gls{svdd} of Tax and Duin \cite{tax1999support, tax2004support} models the boundary of data under consideration.
 Thereby it characterizes a data set and can be used to detect novel data or outliers.
@@
 
-Why use SVM for density estimation? Look for justification, perhaps a review paper which compares/mentions SVM for temporal segmentation of (human activity type of) data?
+Why use \gls{svm} for density estimation? Look for justification, perhaps a review paper which compares/mentions SVM for temporal segmentation of (human activity type of) data?
 
@@
   \item (arguments from \cite{zhang2011optimal} ``Optimal model selection for posture recognition in home-based healthcare'')
+  \item Survey \cite{hodge2004survey} states RBF is similair to Gaussian Mixture model (page 25).
 \end{itemize}
+
+*** TODO: read \cite{mavroforakis2006geometric} and \cite{bennett2000duality} on geometry of SVMs ***
 
@@
 %----------------------------------------------------------------------------------------
+
+*** Note: in \cite{mavroforakis2006geometric} it is explained why the inner-product between two vectors is a logical choice for the distance/similarity measure. ***
+
 \section{Relation to other methods}
@@
 *** Read \cite{hodge2004survey} ``A survey of outlier detection methodologies'' to compare with other methodologies. ***
+\begin{itemize}
+  \item One-class classification is referred to as ``Type 3'', with \emph{semi-supervised recognition or detection}.
+  \item It explains why one-class can be beneficial over type-2 where negative examples needs to be provided.
+  \item We are interested in type-3, so compare SVM with the others stated in this survey regarding type-3.
+  \item Often refers to the convex hull of the data set. Link with geometric approach as in \cite{bennett2000duality,mavroforakis2006geometric}?
+  \item It states that PCA and regression techniques are linear models and thus often are too simple for practical applications.
+SVMs try to find a hyperplane in higher dimensional space; linear models to implement complex class boundaries.
+It refers to \cite{tax1999support}.
+\end{itemize}
 

### 2013-07-03 Chapters/Chapter2.tex
Adding acronym-glosarries

@@
 This problem thus has received a lot of attention in the fields of data mining, etc... *** list and refs ***.
-Often this problem is formulated in a statistical framework, by inspecting the data generating PDF (Probability Density Function) of the time series data.
-A change point is then defined as a significant change in the properties of the PDF, such as the mean and variance.
+Often this problem is formulated in a statistical framework, by inspecting the data generating \gls{pdf} of the time series data.
+A change point is then defined as a significant change in the properties of the \gls{pdf}, such as the mean and variance.
 
-The widely used CUSUM (cumulative sum) method by Basseville \etal \cite{basseville1993detection} takes this approach.
+The widely used \gls{cusum} method by Basseville \etal \cite{basseville1993detection} takes this approach.
 It originates from control methods for detection from bench marks.
@@
 Many methods rely on pre-specified parametric model assumptions and considers the data be independent over time, which makes it less flexible to real-world applications.
-The proposed methods by Kawahara \etal \cite{kawahara2009change} and Lui \etal \cite{liu2013change} try to overcome these problems by estimating the \emph{ratio} between the PDF, instead of estimating each PDF.
+The proposed methods by Kawahara \etal \cite{kawahara2009change} and Lui \etal \cite{liu2013change} try to overcome these problems by estimating the \emph{ratio} between the \gls{pdf}, instead of estimating each \gls{pdf}.
 This approach is discussed and analyzed in section \ref{density-ratio}.
 
-The density-estimation methods rely on the log-likelihood ratio between PDFs.
-The method of Camci \cite{camci2010change} takes an other approach within the statistical framework, by using a SVM (Support Vector Machine).
+The density-estimation methods rely on the log-likelihood ratio between \glspl{pdf}.
+The method of Camci \cite{camci2010change} takes an other approach within the statistical framework, by using a \gls{svm}.
 One problem it tries to overcome is the (claimed) weakness of many methods to detect a decrease in variance.
 The method represents the distribution over the data points as a hyper-sphere in a higher dimension using kernel trick.
-A change in the PDF is represented by a change in the radius of this sphere.
-Section \ref{svm} discusses the SVM-method.
+A change in the \gls{pdf} is represented by a change in the radius of this sphere.
+Section \ref{svm} discusses the \gls{svm}-method.
 

### 2013-07-03 Chapters/Chapter2/cusum.tex
Adding acronym-glosarries

@@
 Non-Bayesian change detection algorithm (thus: no prior distribution believe available for the change time).
-The CUSUM (cumulative sum) method is developed by Page \cite{page1954continuous} for the application of statistical quality control (it is also known as a control chart).
+The \gls{cusum} method is developed by Page \cite{page1954continuous} for the application of statistical quality control (it is also known as a control chart).
 Primary for detection of mean shift.
-The MOSUM of squares test for monitoring variance changes \cite{hsu2007mosum}.
+The \gls{mosum} of squares test for monitoring variance changes \cite{hsu2007mosum}.
 Use of Cumulative Sums of Squares for Retrospective Detection of Changes of Variance \cite{inclan1994use}
 
-An often used approach in the statistical framework of change detection is the Cumulative Sum of Squares (CUSUM) as introduced by Page \cite{page1954continuous}.
+An often used approach in the statistical framework of change detection is the \gls{cusum} as introduced by Page \cite{page1954continuous}.
 Originally used for quality control in production environments, its main function is to detect change in the mean of measurements and has been applied to this problem \cite{basseville1993detection}.
@@
 Others apply the method the problems in which the change of variance is under consideration.
-Among others are there the centered version of the cumulative sums, introduced by Brown, Durbin and Evans \cite{brown1975techniques} and the Moving Sum (MOSUM) of squares by \cite{hsu2007mosum}.
+Among others are there the centered version of the cumulative sums, introduced by Brown, Durbin and Evans \cite{brown1975techniques} and the \gls{mosum} of squares by \cite{hsu2007mosum}.
 
-The method of Incl\'{a}n and Tiao \cite{inclan1994use} builds on the centered version of CUSUM \cite{brown1975techniques} to detect changes in variance.
-Using the Iterative Cumulative Sum of Squares (ICSS) algorithm they are able to find multiple change points in a reflective manner.
+The method of Incl\'{a}n and Tiao \cite{inclan1994use} builds on the centered version of \gls{cusum} \cite{brown1975techniques} to detect changes in variance.
+Using the \gls{icss} algorithm they are able to find multiple change points in a reflective manner.
 Let $C_k = \sum_{i=1}^k \alpha_t^2$ be the cumulative sum of squares for a series of uncorrelated random variables $\{\alpha_t\}$ of length $T$.
@@
 
-One of the motivations for the ICSS algorithm was the heavy computational burden involved with Bayesian methods, which need to calculate the posterior odds for the log-likelihood ratio testing.
-The ICSS algorithm avoids applying a function at all possible locations, due to the iterative search.
+One of the motivations for the \gls{icss} algorithm was the heavy computational burden involved with Bayesian methods, which need to calculate the posterior odds for the log-likelihood ratio testing.
+The \gls{icss} algorithm avoids applying a function at all possible locations, due to the iterative search.
 The authors recommend the algorithm for analysis of long sequences.

### 2013-07-03 glossary-acronyms.tex
Adding acronym-glosarries

% !TEX root = main.tex

% A C R O N Y M S

\newacronym{cusum}{CUSUM}{Cumulative Sum}

\newacronym{icss}{ICSS}{Iterated Cumulative Sums of Squares}

\newacronym{mosum}{MOSUM}{Moving Sum}

\newacronym{pdf}{PDF}{Probability Density Function}

\newacronym{svdd}{SVDD}{Support Vector Data Description}

\newacronym{svm}{SVM}{Support Vector Machine}

### 2013-07-03 main.tex
More on HMM
Add marco for et al. abbrevation
Starting summary Peter Flach
Add double bar notations
Use `\norm` for double bar notions, following
http://tex.stackexchange.com/questions/43008/absolute-value-symbols/43009#43009
Add `argmin` and `argmax` defs.
Putting Camci-flaws in text
Adding acronym-glosarries

@@
 %numbers), remove 'numbers'
+\usepackage{isomath}
+\usepackage{paralist}
 \hypersetup{urlcolor=blue, colorlinks=true} % Colors hyperlinks in blue -
@@
 
+% Use double bar notations: http://tex.stackexchange.com/questions/43008/absolute-value-symbols/43009#43009
+\usepackage{mathtools}
+\DeclarePairedDelimiter\abs{\lvert}{\rvert}%
+\DeclarePairedDelimiter\norm{\lVert}{\rVert}%
+
+% Swap the definition of \abs* and \norm*, so that \abs
+% and \norm resizes the size of the brackets, and the
+% starred version does not.
+\makeatletter
+\let\oldabs\abs
+\def\abs{\@ifstar{\oldabs}{\oldabs*}}
+%
+\let\oldnorm\norm
+\def\norm{\@ifstar{\oldnorm}{\oldnorm*}}
+\makeatother
+% End of double bar notations
+
+\DeclareMathOperator*{\argmin}{\arg\!\min}
+\DeclareMathOperator*{\argmax}{\arg\!\max}
+
+% \usepackage{hyperref}
+\usepackage[nomain,acronym,xindy,toc]{glossaries}
+\renewcommand*{\glstextformat}[1]{\textcolor{black}{#1}}
+\makeglossaries
+\usepackage[xindy]{imakeidx}
+\makeindex
+
 \begin{document}
+\include{glossary-acronyms}
+\include{glossaries}
 
@@
 \newcommand{\HRule}{\rule{\linewidth}{0.5mm}} % New command to make the lines in the title page
+
+\ExplSyntaxOn
+\newcommand\latinabbrev[1]{
+  \peek_meaning:NTF . {% Same as \@ifnextchar
+    #1\@}%
+  { \peek_catcode:NTF a {% Check whether next char has same catcode as \'a, i.e., is a letter
+      #1.\@\ }%
+    {#1.\@\ }}}
+\ExplSyntaxOff
+
+%Omit final dot from each def.
+\def\eg{\latinabbrev{e.g}}
+\def\etal{\latinabbrev{\textit{et al}}}
+\def\etc{\latinabbrev{etc}}
+\def\ie{\latinabbrev{i.e}}
 
@@
 
+% %----------------------------------------------------------------------------------------
+% %	ABBREVIATIONS
+% %----------------------------------------------------------------------------------------
+
+% \clearpage % Start a new page
+
+% \setstretch{1.5} % Set the line spacing to 1.5, this makes the following tables easier to read
+
+% \lhead{\emph{Abbreviations}} % Set the left side page header to "Abbreviations"
+% \listofsymbols{ll} % Include a list of Abbreviations (a table of two columns)
+% {
+% \textbf{LAH} & \textbf{L}ist \textbf{A}bbreviations \textbf{H}ere \\
+% %\textbf{Acronym} & \textbf{W}hat (it) \textbf{S}tands \textbf{F}or \\
+% \textbf{PCA} & \textbf{P}rinciple \textbf{C}omponent \textbf{A}nalysis
+% }
+
 %----------------------------------------------------------------------------------------
-%	ABBREVIATIONS
+% GLOSSARIES
 %----------------------------------------------------------------------------------------
 
-\clearpage % Start a new page
+\clearpage
 
-\setstretch{1.5} % Set the line spacing to 1.5, this makes the following tables easier to read
+\lhead{\emph{Acronyms}}
 
-\lhead{\emph{Abbreviations}} % Set the left side page header to "Abbreviations"
-\listofsymbols{ll} % Include a list of Abbreviations (a table of two columns)
-{
-\textbf{LAH} & \textbf{L}ist \textbf{A}bbreviations \textbf{H}ere \\
-%\textbf{Acronym} & \textbf{W}hat (it) \textbf{S}tands \textbf{F}or \\
-}
+\printglossaries
 
-%----------------------------------------------------------------------------------------
-%	PHYSICAL CONSTANTS/OTHER DEFINITIONS
-%----------------------------------------------------------------------------------------
+% %----------------------------------------------------------------------------------------
+% %	PHYSICAL CONSTANTS/OTHER DEFINITIONS
+% %----------------------------------------------------------------------------------------
 
-\clearpage % Start a new page
+% \clearpage % Start a new page
 
-\lhead{\emph{Physical Constants}} % Set the left side page header to "Physical Constants"
+% \lhead{\emph{Physical Constants}} % Set the left side page header to "Physical Constants"
 
-\listofconstants{lrcl} % Include a list of Physical Constants (a four column table)
-{
-Speed of Light & $c$ & $=$ & $2.997\ 924\ 58\times10^{8}\ \mbox{ms}^{-\mbox{s}}$ (exact)\\
-% Constant Name & Symbol & = & Constant Value (with units) \\
-}
+% \listofconstants{lrcl} % Include a list of Physical Constants (a four column table)
+% {
+% Speed of Light & $c$ & $=$ & $2.997\ 924\ 58\times10^{8}\ \mbox{ms}^{-\mbox{s}}$ (exact)\\
+% % Constant Name & Symbol & = & Constant Value (with units) \\
+% }
 
@@
 
+\clearpage
 \setstretch{1.3} % Return the line spacing back to 1.3
@@
 \input{./Chapters/Chapter6}
-%\input{./Chapters/Chapter7}
+\input{./Chapters/Chapter7}
 
@@
 
-\input{./Appendices/AppendixA}
+\input{./Appendices/AppendixA-Summaries}
+\input{./Appendices/AppendixB-Camci}
 %\input{./Appendices/AppendixB}

### 2013-07-03 Appendices/AppendixB-camci.tex
Add README. Change title

@@
 \section{Density estimation / Data description / Vapnik's principle}
-Following Vapnik's principle, one should ``When solving a problem of interest, do not solve a more general problem as an intermediate step.'' \cite{vapnik1998statistical} when a limited amount of data is available.
+Following Vapnik's principle, one should ``When solving a problem of interest, do not solve a more general problem as an intermediate step'' \cite{vapnik1998statistical} when a limited amount of data is available.
 For the problem of change detection we are only interested in some characteristics of the data.

### 2013-07-03 Appendices/AppendixA-summaries.tex
wip

@@
 \subsection{Machine learning: the art and science of algorithms that make sense of data}
-Book by Peter Flach: \cite{flach2012machine}.
+Book by Peter Flach:~\cite{flach2012machine}.
 Mainly about chapter 7, ``Linear Models''.
@@
 \subsection{Change Point Detection In Time Series Data Using Support Vectors}
-Paper by Fatih Camci \cite{camci2010change}
+Paper by Fatih Camci~\cite{camci2010change}
 About segmentation with SVMs.
@@
 \subsection{Use of Cumulative Sums of Squares for Retrospective Detection of Changes of Variance}
-Carla Inclan and George C. Tiao \cite{inclan1994use}, 1944, 162 refs.
+Carla Inclan and George C. Tiao~\cite{inclan1994use}, 1944, 162 refs.
 
@@
 \subsection{Change-Point Detection in Time-Series Data by Direct Density-Ratio Estimation}
-Kawahara and Sugiyama \cite{kawahara2009change}, 2009, 45 refs.
+Kawahara and Sugiyama~\cite{kawahara2009change}, 2009, 45 refs.
 
@@
 The key idea of this paper to directly estimate the \emph{ratio} of the probability densities (also known as \emph{importance}).
-The Kullback-Leibler Importance Estimation Procedure (KLIEP) is an example, but it is a batch algorithm.
-This paper introduces an online version of the KLIEP algorithm and develops a flexible and computationally efficient change-point detection method.
+The \gls{kliep} is an example, but it is a batch algorithm.
+This paper introduces an online version of the \gls{kliep} algorithm and develops a flexible and computationally efficient change-point detection method.
 This method is equipped with a natural cross validation procedure and thus the value of tuning parameters can be objectively determined.
@@
 ...
+
+%----------------------------------------------------------------------------------------
+%----------------------------------------------------------------------------------------
+%----------------------------------------------------------------------------------------
+%----------------------------------------------------------------------------------------
+\section{Outlier Detection methods}
+\subsection{A Survey of Outlier Detection Methodologies}
+Hodge and Austin,~\cite{hodge2004survey}, 2004, 734 refs.
+
+\subsection{Summary}
+This survey tooks a look at different methodologies that perform outlier detection.
+It gives two definitions of outliers, one of which relates to the problem statement in this thesis (taken from~\cite{barnett1994outliers}):
+\begin{quote}An observation (or subset of observations) which appears to be inconsistent with the remainder of that set of data.\end{quote}
+
+The survey introduces three fundamental approaches to the problem of outlier detection:
+\begin{description}
+  \item[Type 1] - no prior knowledge about the data; analogous to \emph{unsupervised clustering}. Flags the remote points as outliers; mainly batch-processing systems.
+  \item[Type 2] - model (and requires) both normal and abnormal data; analogous to \emph{supervised classification}. Is able to process new data online.
+  \item[Type 3] - Model only normal data (and in very few cases abnormal).
+Generally refered to as \emph{novelty detection}, analogous to \emph{semi-supervised recognition or detection}.
+It only requires pre-classified normal data.
+It aims to define a boundary of normality.
+\end{description}
+
+The type of approaches of interest in this research is of \textbf{type 3}.

### 2013-07-03 Appendices/AppendixB-camci.tex
wip

@@
 
-In \cite{liu2013change} it is stated that the \emph{Support Vector Machine} by Cortes and Vapnik \cite{cortes1995support} is representative example of this principle.
+In \cite{liu2013change} it is stated that the \gls{svm} by Cortes and Vapnik \cite{cortes1995support} is a representative example of this principle.
 Instead of estimating the more general data generating probability distributions, it only learns a decision boundary to differentiate between the two distributions.
@@
 Thereby it characterizes a data set and can be used to detect novel data or outliers.
-The performance is compared to methods which model the distribution's density instead, using Receiver-Operating Characteristic (ROC) curves and false negative rates.
+The performance is compared to methods which model the distribution's density instead, using \gls{roc} curves and false negative rates.
 The compared methods are:
@@
   \item a Gaussian Mixture Model optimized using EM; and
-  \item the Nearest-Neighbor Method which compares the local density of an object with the density of the nearest neighbor in the target set (and is thus, just as SVDD, a boundary-based method).
+  \item the Nearest-Neighbor Method which compares the local density of an object with the density of the nearest neighbor in the target set (and is thus, just as \gls{svdd}, a boundary-based method).
 \end{inparaenum}
-The results show that when the problem formulation is to characterize an area in a feature space (and not the complete density distribution) SVDD gives a good data description.
+The results show that when the problem formulation is to characterize an area in a feature space (and not the complete density distribution) \gls{svdd} gives a good data description.
 
-The study of Tax on one-class classifiers \cite{tax2001one} further compares density methods, boundary methods (amongst which SVDD) and reconstruction methods.
-One promesing result for SVDD is that it performs well on read-world data, for which generalization is needed.
+The study of Tax on \glspl{occ} \cite{tax2001one} further compares density methods, boundary methods (amongst which \gls{svdd}) and reconstruction methods.
+One promesing result for \gls{svdd} is that it performs well on read-world data, for which generalization is needed.
 
@@
 The paper of Yin \etal \cite{yin2008sensor} makes a distinction between similarity based (using a defined distance measure) and model based (which characterize the data using predictive models) approach.
-The one-class SVM is used in the model-approach to filter out normal activities in order to detect abnormality behavior.
+The \gls{oc-svm} is used in the model-approach to filter out normal activities in order to detect abnormality behavior.
 
@@
 It assumes that each of these stochastic processes is modeled following a Gaussian distribution, such that a change can occur in the value of the mean and/or the variance; $p^1 \sim N(\mu_1, \sigma_1^2)$.
-The CUSUM-based method of \cite{inclan1994use} also regards each semgent as a Gaussian distribution.
+The \gls{cusum}-based method of \cite{inclan1994use} also regards each semgent as a Gaussian distribution.
 
-The method if Kawahara \etal \cite{kawahara2009change} is based on the log likelihood ratio of test samples, and the method by Liu \etal \cite{liu2013change} uses a comparable dissimilarity measure using the KLIEP algorithm.
+The method if Kawahara \etal \cite{kawahara2009change} is based on the log likelihood ratio of test samples, and the method by Liu \etal \cite{liu2013change} uses a comparable dissimilarity measure using the \gls{kliep} algorithm.
 
@@
 It assumes a $K$-state hidden process with a (hidden) state sequence, each state providing the parameters (amongst which the order) for a polynomial.
-The order of the model segment is determined by model selecting, often using the Bayesian Information Criterion (BIC) or the similair Akaike Information Criterion (AIC) \cite{akaike1974new}, as in \cite{he2008activity}.
+The order of the model segment is determined by model selecting, often using the \gls{bic} or the similair \gls{aic} \cite{akaike1974new}, as in \cite{he2008activity}.
 
@@
 
-Why use RBF/Gaussian kernel? Is it because of one-class SVM of because of form the data? Why not polynomial/linear?
+Why use RBF/Gaussian kernel? Is it because of \gls{oc-svm} of because of form the data? Why not polynomial/linear?
 \begin{itemize}

### 2013-07-03 Chapters/Chapter2/density_ratio.tex
wip

@@
 A solutions to this is to estimate the \emph{ratio} of probabilities instead of the probabilities themselves.
-One of the recent methods to achieve this is the Kullback-Leibler Importance Estimation Procedure (KLIEP) by Sugiyama \etal \cite{sugiyama2008direct}.
+One of the recent methods to achieve this is the \gls{kliep} by Sugiyama \etal \cite{sugiyama2008direct}.
 
-The method proposed by Kawahara and Sugiyama \cite{kawahara2009change} is composed of an online version of the KLIEP algorithm.
+The method proposed by Kawahara and Sugiyama \cite{kawahara2009change} is composed of an online version of the \gls{kliep} algorithm.
 The method also considers \emph{sequences} of samples (rather than samples directly) because the time series samples are generally not independent over time.
@@
 
-The procedure of the method proposed by Kawahara and Sugiyama \cite{kawahara2009change} is to first apply the batch KLIEP algorithm with model selection for initial parameter $\vectorsym{\alpha}$ and kernel width calculation.
+The procedure of the method proposed by Kawahara and Sugiyama \cite{kawahara2009change} is to first apply the batch \gls{kliep} algorithm with model selection for initial parameter $\vectorsym{\alpha}$ and kernel width calculation.
 Then for every new sample the reference and test intervals are shifted and the calculated parameters $\vectorsym{\alpha}$ are updated.

### 2013-07-03 Appendices/AppendixA-summaries.tex
Added stuff of survey outlier methodologies

--- 2013-07-03 Appendices/AppendixA-summaries.tex
@@
 ...
+
+%----------------------------------------------------------------------------------------
+%----------------------------------------------------------------------------------------
+%----------------------------------------------------------------------------------------
+%----------------------------------------------------------------------------------------
+\section{Outlier Detection methods}
+\subsection{A Survey of Outlier Detection Methodologies}
+Hodge and Austin, \cite{hodge2004survey}, 2004, 734 refs.
+
+\subsection{Summary}
+This survey takes a look at different methodologies that perform outlier detection.
+It gives two definitions of outliers, one of which relates to the problem statement in this thesis (taken from \cite{barnett1994outliers}):
+\begin{quote}An observation (or subset of observations) which appears to be inconsistent with the remainder of that set of data.\end{quote}
+
+The survey introduces three fundamental approaches to the problem of outlier detection:
+\begin{description}
+  \item[Type 1] - no prior knowledge about the data; analogous to \emph{unsupervised clustering}. Flags the remote points as outliers; mainly batch-processing systems.
+  \item[Type 2] - model (and requires) both normal and abnormal data; analogous to \emph{supervised classification}. Is able to process new data online.
+  \item[Type 3] - Model only normal data (and in very few cases abnormal).
+Generally referred to as \emph{novelty detection}, analogous to \emph{semi-supervised recognition or detection}.
+It only requires pre-classified normal data.
+It aims to define a boundary of normality.
+\end{description}
+
+The type of approaches of interest in this research is of \textbf{type 3}.
+It is characterized by the ability to recognize new data as normal when it lies within the constructed boundary and as a novelty otherwise.
+This ability removes the need of sample-abnormality data, which may be hard (or costly) to produce.
+The method of Tax \etal \cite{tax1999supportdata} is stated to be of \textbf{type 2}, and one can argue that other methods of Tax \etal \cite{tax2001one,tax2004support} are by their one-class classification instances of \textbf{type 3} methods.
+
+The survey states that density-based methods estimate the density distribution of the training data.
+Outliers are identified as data points lying in a low-density region.

### 2013-07-03 Chapters/old_classification/Chapter2.tex
Added stuff of survey outlier methodologies

% Chapter 2
% !TEX root = ../main.tex

\chapter{Techniques} % Main chapter title

\label{Chapter2} % For referencing the chapter elsewhere, use~\ref{Chapter1}

\lhead{Chapter 2. \emph{Techniques}} % This is for the header on each
%page -
%perhaps a shortened title

%----------------------------------------------------------------------------------------

\input{./Chapters/Chapter2/Section_signals}

\section{Temporal Segmentation}
This section will give an introduction and in-depth analysis of temporal segmentation.

\subsection{Aims of segmentation}
When processing and analyzing time series of data, e.g. motion measurements, stock market fluctuations or natural language, first a low-level division between the discriminative parts of the stream must be made.
One can view this as splitting the series into the \emph{atoms}, which are the building blocks of the total stream.
These building blocks will be the aggregation of non-overlapping, internally homogeneous segments \cite{himberg2001time}.
This means that the data points inside a segment should have some resemblance relation to each other and their difference lies between some boundary.
The process of segmenting can be viewed as a subproblem to context analysis of time series.
Temporal segmentation is closely related to temporal clustering, although it is a stricter, and simpler, process.
Whereby clustering only restricts the data points on their distance relation (as used in a Voronoi diagram), within a segment the data points must also be contiguous.

The task of segmentation can be performed in a manual matter, by cutting and labeling parts of the stream into coherent parts.
This would require human (expert) knowledge and does not yield a clear cut because of ambiguity.
With increasing storage abilities and easier motion capture systems, there is a desire for automated systems which perform the segmentation task unsupervised.
Some algorithms used have the (often desired) side-effect of also clustering the segments, such that classes of segments can be discovered in the time series.
These algorithms would not only be able to make a distinction between walking, sitting and walking, but would also recognize the reappearance of the walking activity.

\subsection{Formal definition}
Formally, temporal segmentation is dividing a time series $s$, which consists of $N$ samples $\mathbf{x}(1),\mathbf{x}(2),\dots,\mathbf{x}(N)$ from $\mathbf{R}^d$.
Individual \emph{segments} are referenced by $s(a,b)$, consisting of the consecutive samples $\mathbf{x}(a),\mathbf{x}(a+1),\dots,\mathbf{x}(b)$, $a \le b$.
Let $s_1 = s(a,b)$ and $s_2 = s(b+1,c)$ be two segments, then their concatenation is $s_1s_2 = s(a,c)$.
A segmentation $S$ of $s$ consists of a sequence of $k$ non-empty segments $s_1s_2 \dots s_k = s$.
This notation is adopted from \cite{himberg2001time}.

As stated, informally each segment should be internally homogeneous.
This can formally be measured with an cost function $F$, indication the heterogeneity of a segment.
The overall aim is to minimize the cost $F$. The cost of a segment is a function from the data points and the number of data points
$n = b - a + 1$ and is expressed as
\begin{equation}
	\label{eq:segment_cost}
	\mathrm{cost}_F (s(a,b)) = F(\mathbf{x};n|\mathbf{x} \in s(a,b))
\end{equation}
The cost of a \emph{k-segmentation} $S$ is the summation of the costs of the $k$ segments:
\begin{equation}
	\label{eq:segmentation_cost}
	\mathrm{Cost}_F (s_1 s_2 \dots s_k) = \sum_{i=1}^{k} \mathrm{cost}_F (s_k)
\end{equation}
With the objective of minimizing the cost function, the optimal $k$-segmentation $S_F^\mathit{opt}(s;k)$ is the segmentation with minimal $\mathrm{Cost}_F(s_1 s_2 \dots s_k)$ over all possible $k$-segmentations.

The cost function, to calculate the heterogeneity of a (set of) segment(s), can be any function.
A simple and natural function would be the sum of variances of the segments.
The overall cost function would then be
\begin{equation}
	\label{eq:cost_variances}
	\mathrm{Cost}_V = \frac{1}{N} \sum_{i=i}^{k} \sum_{j=c_{i-1}+1}^{c_i} \|
	\mathbf{x}(j) - \mu_i \|^2
\end{equation}
where $\mu_i$ is the mean vector of data points in segment $s_i$.

\subsection{Application in research fields}

[CHARACTERISTICS OF HUMAN MOTION]
temporal variability, invariance over time, metrics over actions.

[COMPUTER VISION]

[GRAPHICS/VIDEO]

[DATA-MINING]

[MODEL BASED]

To analyze time series it is often preferred to divide the stream in segments of correlated data.
After dividing, each segment represent a period in time in which the same activity is performed.
Or, stated otherwise, it results in transitions moments between activities.

\subsection{PCA Based Methods}

Many fields of research have been active in the unsupervised segmentation of data.
Many authors rely on a form of Principal Component Analysis (PCA), as used a.o. in \cite{barbivc2004segmenting}.
Often PCA is used to reduce the dimensionality of the data being processed [REFERENCE] by only using the top $r$ dimensions to describe the data set.
It is observed that data series of simple motions have a lower dimensionality then complexer motions.
When a simple (repetitive) motion is about to end and fluently transforms in a new motion, there will be a window of time in which a high dimensionality will be present, due to the new motion.
After this period of transition, the dimensionality will decrease, since only the new simple motion is present in the window of time.
The first algorithm of \cite{barbivc2004segmenting} is based on this principle.

Given a set of data points, a lower dimensional hyperplane can by constructed to which the data points can be projected.
This projection on a lower dimension introduces a error to the original position.
When the error is fixed, less dimensions are needed for simple motions in which movements of body parts are highly correlated.
For segments in which the data points are lesser correlated, e.g. because of transition state, a higher degree of dimensions of the hyperplane is needed to represent the data with equal error degree.
When the dimensionality is reduced from $d$ to $r \le d$, the ratio of error $E_r$ can be calculated as
\begin{equation}
	E_r = \frac{\sum_{j=1}^{r} \sigma_j^2}{\sum_{j=1}^{d} \sigma_j^2}
\end{equation}

where $\sigma_j$ are the singular values as a result from Singular Value Decomposition (SVD), which is closely related to PCA \cite{shlens2005tutorial}.

In \cite{barbivc2004segmenting} the stream of frames is analyzed on cut-frames to find transitions between action.
First, for a number of $k$ frames (e.g. the equivalent of 2.5 seconds) the required dimensionality $r$ to keep the error $E_r$ below some threshold $\tau$ is calculated.
This will yield in an error $e_i$ for the first $i$ frames.
When more frames are added to the window, the error will increase with a low constant when it is still in the same activity, due to noise in the activity.
When a new activity starts, e.g. at frame $j$, the error at frame $j$ will start increasing faster.
This can be expressed by the derivative of the error rate $d_i = e_i - e_{i-l}$, where $l$ is a constant to remove noise.
From this derivative the mean and standard deviation can be calculated, for each point.
When a derivative $d_j$ rises more than a factor $k_\sigma = 3$ standard deviations from the mean, a transition point is encountered.
The previous frames are then cut from the sequence (as a segment) and the algorithm starts over.

*** [figure to illustrate derivative and standard deviation]

A second approach in \cite{barbivc2004segmenting} uses the probabilistic variant of PCA (PPCA) to model the data set as a Gaussian distribution instead of ignoring the frames which do not fit in the subspace.
Over windows of frames the mean and variance are calculated.
In a forward manner the Mahalanobis distance of a new window of frames is calculated, which represents the likelihood of the new window belonging to the same segment as the original widow.
When the distance decreases, the likelihood increases which happens when the motions in the becomes more homogeneous.
When a peak in the distance is reached, the new window of frames indicates a heterogeneous collection of motions in the window and thus a low likelihood of membership and a indication of a transition.
In order to distinct activities and sub-activities (which require a subset of motions is a distinct activity) the algorithm is also processed backward over the data series.

*** [figure to illustrate mahalanobis distance measure and peaks]

The third algorithm in \cite{barbivc2004segmenting} is based on the observation that data points (frames) tend to form clusters in the space.
These clusters are represented by $k$ Gaussian distributions for which each the Expectation-Maximization (EM) algorithm estimates the mean $m_j$, covariance matrix $\sum_{j}$ and prior $\pi_j$.
With all the Gaussian distributions estimated, the data points are assigned to the cluster with the highest membership likelihood.
When two consecutive frames $x_i$ and $x_{i+1}$ belong to different clusters, a transition of activities is recognized.
Note that this algorithm succeeds in segmenting the data and also labels the similar simple activities.

A drawback in this system, and many others which implement a variant of the $k$-means algorithm, is that the number of clusters $k$ need to be predetermined.
To cope with this, often the algorithm is performed multiple times for different values of $k$.
Using some criteria, e.g. the Bayesian Information Criterion \cite{pelleg2000x} or the Davies-Bouldin Index which guides $k$-means clustering as used in \cite{krause2003unsupervised}.

\subsection{Statistical methods}
An method to process data, or create a basis for further processing, is by analyzing the statistical properties of a data set.
When there is a continuous stream of data, it is possible to extract events by comparing the statistical properties of sliding time windows.
By definition, events have a different statistical profile then their background.
When two consecutive actions are regarded as the background, then a transition is an event between them.

In \cite{guenterberg2009automatic} a automatic system is used to segment a continuous stream of data.
By calculating features as activity level from the standard deviation of a sliding window and applying Adaptive Thresholds, the data can be segmented in active and rest parts.
There is no model representation.

\subsection{Hidden Markov Models Based Methods}

\subsection{Bayesian Methods}

\subsection{Principal Component Analysis}
When working with simple 2- or 3-dimensional data it is often, for humans, easy to discover patterns in the set.
The data can be plotted and the lines or planes among which the data points lie gives an indication of the pattern.
With Principal Component Analysis (PCA) this process is also possible for automated systems.
With this method the overall form of a set of data points can be represented, the most discriminative dimensions can be found, the dimensionality of a set can be reduced (which yield in data compression) or the result can be used to characterize and differentiate sets of data points.

The following steps are performed, which are explained below and illustrated in figure *** [add figure with plot of sets]:

\begin{enumerate}
	\item \textbf{Gather data} and represent them in a chosen number of features,
	\item \textbf{Subtract mean} to center the data around the axis. The new data set will have mean zero,
	\item \textbf{Generate covariance matrix} by calculating all pairwise feature variances,
	\item \textbf{Calculate Eigenvectors and Eigenvalues} of the covariance matrix.
	The Eigenvectors are then normalized to make further calculations easier,
	\item \textbf{Generate feature vector} which indicated which features are characteristic or meant to keep in de data set, by comparing the Eigenvalues.
\end{enumerate}

The workings of the PCA \cite{smith2002tutorial} relies on the concepts of standard deviation, variance, covariance, eigenvectors and eigenvalues of matrices.
These concepts will be discussed very briefly \footnote{For a more in-depth discussion we would like to refer the reader to \cite{jolliffe2005principal}}.
The standard deviation and variance of a set are measures for the spread around the mean for a single dimension, or feature.
The covariance between two features (dimensions of the data points) indicates how they are related; when positive the two features will increase together; when negative one will decrease when the other increases and when zero they are unrelated.
The covariance matrix gives all the covariances for all pairs of features.
This matrix is symmetrical about the diagonal and the values on the diagonal are the variances for each feature.
For this matrix the eigenvectors can be computed. Eigenvectors have the characteristic that when they multiply a matrix, the resulting vector is a multiple of the Eigenvector.
The amount by which it is the multiple is the Eigenvalue.
This is illustrated in formulae~\ref{eq:no-eigenvector} and \ref{eq:eigenvector}.
The second formula shows an Eigenvector $ \left( \begin{smallmatrix} 3 \\ 2 \end{smallmatrix} \right)$ and its associated Eigenvalue, 4.
The vector in the first formula is not an Eigenvector.

\begin{equation}
	\label{eq:no-eigenvector}
	\begin{pmatrix} 2 & 3 \\ 2 & 1 \end{pmatrix}
	\times
	\begin{pmatrix} 1 \\ 3 \end{pmatrix}
	=
	\begin{pmatrix} 2 \cdot 1 + 3 \cdot 3 \\ 2 \cdot 1 + 1 \cdot 3
	\end{pmatrix}
	=
	\begin{pmatrix} 11 \\ 5 \end{pmatrix}
\end{equation}

\begin{equation}
	\label{eq:eigenvector}
	\begin{pmatrix} 2 & 3 \\ 2 & 1 \end{pmatrix}
	\times
	\begin{pmatrix} 3 \\ 2 \end{pmatrix}
	=
	\begin{pmatrix} 2 \cdot 3 + 3 \cdot 2 \\ 2 \cdot 3 + 1 \cdot 2
	\end{pmatrix}
	=
	\begin{pmatrix} 12 \\ 8 \end{pmatrix}
	=
	4 \times \begin{pmatrix} 3 \\ 2 \end{pmatrix}
\end{equation}

 It are these Eigenvectors and Eigenvalues that makes PCA possible.
 Given a matrix of size $n \times n$ (only square matrices have Eigenvectors), $n$ Eigenvectors can be found.
 Each Eigenvector is perpendicular, or orthogonal, to the others.
 The Eigenvectors combined describes the lines over which the data is plotted.
 The Eigenvector with the largest Eigenvalue is the \emph{principal} component of the data set.
 It is said that it is the most significant feature.
 Thus the Eigenvectors, and thereby the features or dimensions of the data set, can be sorted on significance by the Eigenvalues.

With this ordering there are a few applications possible.
The first is to just make a (comprehensible) representation of the data.
The Eigenvectors describe the cloud of data points and thus are a compressed representation of the data.
When the original data points are to be compressed, it is possible to remove the least significant features and reconstruct the data points from the resulting Eigenvectors.
This will yield in a lossy compression. *** [figure to illustrate].
An extension of this application is to determine the number of features needed to keep the compression within a certain error criterion, as used in \cite{barbivc2004segmenting}.
Sets of data can then be distinguished by the number of features needed to describe the points.

*** [graph, a bit of formulae with sets and dimensions]

%-------------------

\subsection{Hidden Markov Models}
Many dynamic systems can be viewed as some sequence of consecutive states which are observable by the world.
These kind of systems are known as Markov Chains.
When the state in which the system is can not be observed with complete certainty, a system known as a Hidden Markov Model (HMM) can be constructed.
It has been shown that HMMs are effective on in the task of speech recognition \cite{rabiner1989tutorial}.
Because of some resemblance between natural speech construction and human activities, modifications have been able to classify human activities \cite{guenterberg2009distributed} and extensions are implemented which create layers of HMMs to handle different granularity over time and reduce retraining \cite{oliver2002layered,perdikis2008recognition}.
A HMM is a state-space model representing a single model for each label which needs to be classified.

Formally, a Hidden Markov Model is characterized by the following elements
\cite{rabiner1989tutorial}:
\begin{enumerate}
	\item
		A number of $N$ states in the model.
		These states are a representation of the underlying mechanism of the system and can not be directly observed from outside (hence the \emph{hidden} model).
		Individual states are referenced by $S = \{S_1, S_2, \dots, S_N \}$ and a single state in time $t$ is denoted by $q_t$.
	\item
		An alphabet of $M$ discrete observable symbols for each state.
		An observed symbol is the only indication of the systems current situation.
		Each individual symbol is referenced by $V = \{ v_1, v_2, \dots, v_m \}$.
	\item
		A state transition probability distribution $A = \{ a_{ij} \}$ indication the probability of being in state $j$ at time $t+1$ when the state on time $t$ was $i$:
		\begin{eqnarray}
			a_{ij} = P [ q_{t+1} = S_j | q_t = S_i ], & 1 \le i, j \le N
		\end{eqnarray}
		Note that by setting the probability $a_{ij}=0$ for some $i$ and $j$ indicates that $j$ is not reachable from state $i$.
	\item
		A symbol observation probability distribution in state $j$, $B = \{b_j(k) \}$, where
		\begin{eqnarray}
			b_j(k) = P[ v_k\ \mathrm{at} \ t | q_t = S_j], & 1 \le j \le N \nonumber \\
			& 1 \le k \le M
		\end{eqnarray}
		Note that every symbol can be observed in every state.
	\item
	The initial state distribution in which the system starts $\pi= \{ \pi_i \}$ where
	\begin{eqnarray}
		\pi_i = P[q_1 = S_i], & 1 \le i \le N
	\end{eqnarray}
\end{enumerate}

This system is often abbreviated to
\begin{equation}
	\lambda = \{ \pi_i, a_{ij}, b_j(k)\}
\end{equation}
The HMM will act as a generator giving a sequence of $T$ observations, each being a symbol from $V$:
\begin{equation}
	O = O_1, O_2 \dots O_T
\end{equation}

Given the characteristics of a specific HMM, there are three major application which serve the classification of an activity.
The first application is, given a sequence of observations $O = O_1,O_2,\dots,O_T$, to label this sequence with the activity with the highest associated probability for that label, e.g. compute $P(O|\lambda)$ for each model $\lambda$.
The second task a HMM can perform is to \emph{explain} a observation sequence $O$, by choosing an optimal corresponding state sequence $Q = q_1,q_2,\dots,q_T$.
The last application is more of a problem statement, being the task to train the HMM to adjust the parameters $\lambda = \{ \pi_i, a_{ij}, b_j(k)\}$ to maximize $P(O|\lambda)$.
In the overall task of activity recognition, one would first use the last application to train a set of HMMs with the provided and labeled training data.
The second application can be used to further optimize the HMM (set the number of states, adjust vector codebook etc.) and improve the capability of modeling the activities.
Finally, when the system is provided with unlabeled data the first application is used to identify the observed sequence with trained examples, by comparing the probability result of multiple HMMs.

Generally a HMM can be fully connected, or ergodic model.
Other configurations are also possible, of which the left-right model, or Bakis model, is popular for temporal applications \cite{rabiner1989tutorial}.
This because the structure of the state probability distribution resembles the construction of possible states over time.
This left-right model can be generalized by setting
\begin{eqnarray}
	a_{ij} = 0, & j > i + \Delta
\end{eqnarray}
for some $\Delta$, indicating the possible \emph{distance} of the jumps over states.

*** [INSERT GRAPHICS SHOWING STRUCTURE OF HMM]

*** [discrete and continuous HHMs]

*** Applications of HHM: \cite{shi2009towards} uses 10 coefficients extracted with FFT from 6 axis (3 times accelerometer and 3 times gyroscope).
Example of continuous HMM (?).
Recognition rates fo 50\%-100\%.

In \cite{oliver2002layered} it is stated that HMMs are robust mechanisms with respect to changes in temporal segmentation of observations, although they suffer from a few problems when applied to reason about longer and complexer sequences over time.
They state that, especially with limited training data, the HMM tend to be overfitted and tend to lack structure and have an excess of parameters.
The overcome this limitation, they introduce a Layered Hidden Markov Model (LHHM) which they show is more robust to temporal changes in observations.
They make use of the hypothesis from the field of psychology, that human behavior is hierarchically structured \cite{zacks2001event}.
In \cite{perdikis2008recognition} a layered construction of HMMs is used which feeds the low-level data from a sliding window to the lower layer.
This layer will recognize simple motions, such as moving the arm in a pre-defined segment.
The results from these lower HMMs are then fed to a higher-level layer of HMMs which are able to recognize abstract motions, such as adjusting a monitor and picking up a pen.
The results showed that the layered system was better at recognition then a similar single-layered setup.
It was also robust in the sense that when the environment was changed, only the relatively simple low-level HMMs needed to be retrained.

*** [DIAGRAM WITH EXAMPLE OF LOW- AND HIGH-LEVEL HMMs, SUCH AS MOVING ARM, PICKING UP PEN]

*** Viterbi algorithm / path; find the most likely sequence of cause-states for the observed measures.

%-------------------

\subsection{Segmentation as clustering}
In the previous section the discussed methods all relied on the stream of data points and tried to find cuts the discriminate between successive different type of activities.
An other approach is to consider the tasks of segmentation as a type of clustering \cite{zhou2008aligned}.
In clustering the objective is to assign labels, or classes, to all the data points indication a similar type of activity.
A clustering $\mathcal{L}$ is thereby more informative then a segmentation but is also harder to produce.

A clustering $\mathcal{L}$ is generated from a sequence of elements $\mathbf{X}$ which is decomposed in $m$ disjoint segments, each belonging to one of the $k$ classes.
A segment $\mathbf{Y}_i \hat{=} \mathbf{X}_{[s_i,s_{i+1})}$ is composed of frames from position $s_i$ to $s_{i+1}$.
A vector $g_{ci} = 1$ indicates class membership if $\mathbf{Y}_i$ belongs to class $c$, otherwise $g_{ci} = 0$.

When regarding segmentation of human motion as a task of clustering the difficulty is to model the temporal variability of actions and defining a robust metric between temporal actions.
To overcome this, \cite{zhou2008aligned} introduces Aligned Cluster Analysis (ACA), by minimizing
\begin{equation}
	\label{eq:ACA}
	J_{\mathit{ACA}}(\mathbf{G},\mathbf{s}) = \sum_{c=i}^{K} \sum_{i=1}^{m} g_{ci} \mathit{dist}_c (\mathbf{X}_{[s_i,s_{i+1})})
\end{equation}

The characteristic of ACA is that is enables segments to span over different number of data points, whereas the standard kernel $k$-means algorithm results in equally sized segments.
[!!! TRUE?!] The second difference is that the kernel used in $\mathit{dist}_c$ to measure the distance from a segment to the class which it is assigned to uses the Dynamic Time Alignment Kernel [REFERENCE?] to measure between time series.

%----------------------------------------------------------------------------------------

\section{Clustering}

This section will give an introduction and overview to clustering of data.
When processing time series of data the line of distinction between segmentation and clustering is very fine.
This section will introduce clustering for the purpose as used in this project.

Data clustering can be used with two different applications, exploratory or confirmatory \cite{jain1999data}.
In the former, the purpose is to discover clusters in a point cloud of data points.
Here the cluster is defined as a group of homogeneous data points measured by some coherence measure, often a distance function with the data points being defined as vectors.
The produced clusters, which together form a representation of the data examined, can be used in the latter application to assign new provided data points to a cluster and thereby classify them.

Depended on the precise setup, clustering can be used as a successive step after or an implementation of temporal segmentation.
When temporal segmentation is used to create successive coherent data points, clustering can be used to recognize the same activities in different points of time.
The data points provided to the clustering will be some representation of each segment and each segment will be labeled with a activity.
Another mechanism could be to extract features from the raw data points and use these to create clusters directly *** [temporal modifications].
Both mechanisms requires the exploratory and confirmatory phases.

The exploratory phase roughly consists of the following five steps
\cite{jain1999data}
\begin{enumerate}
	\item data point representation,
	\item definition of coherence,
	\item grouping data points,
	\item cluster abstraction,
	\item output qualification
\end{enumerate}

The first step is to find a \emph{representation} of the data points and clusters, considering the number, type and scale of the features and the number of desired classes to cluster in.
When dealing with high-dimensional data points a \emph{feature selection} can be performed to use only the most discriminating features.
When the raw features are not useful enough, \emph{extraction} can be used to create new synthetic features. *** [qualitative versus quantitative/conceptual]

When the data points are represented in a meaningful way, the measure of \emph{coherence} between pairs of points must be defined.
Often this is implemented as the Euclidean or Mahalanobis distance when the data points are represented as vectors or other similarity measures for conceptual patterns.

In the most critical step, the \emph{grouping}, the data points with the highest coherence are linked together to form a cluster.
The result can be a hard partition over the data or a fuzzy membership degree to each cluster for each data point.
When applying rules for merging and splitting cluster, a hierarchical partition is constructed.
Partitional clustering algorithms assign data points to clusters by optimizing some criterion, e.g. the mean squared distance from each data point to the clusters' centroid.

Especially in case of large data sets, the resulting clusters will be defined by many data points.
To form a compacter and simpler representation, \emph{abstraction} can be used.
This will simply analysis by humans or automated systems.
A common abstraction is to use the clusters' centroid \cite{diday1976clustering} or parameters of Gaussian patterns.

The final step which can be applied to the resulting partition is some \emph{qualification}.
The quality of multiple partitions can be compared and the validity of the partition can be determined.
A partition is valid if reasonably it could not been constructed by change or as some random process.
External references of models can be used to compare, or the internal data points can be examined. *** [better wordings]

%-------------------

\subsection{Formal definition}

%-------------------

\subsection{Application is research fields}

%----------------------------------------------------------------------------------------

\section{Temporal pattern recognition}

\subsection{Dynamic Time Warping}
Used to measure similarity between time sequences.
Exact matching is high-cost, so approximations such as Minimum Bounding Rectangles are used.

\subsection{k-means clustering}
[EXPLAIN] divide data set $n$ into $k$ clusters. Follows the outline in figure~\ref{fig:k_means}

Among many unsupervised clustering techniques, $k$-means is successfully applied to large data sets.
It is simple to implement and linear in time complexity so computationally attractive \cite{jain1999data}.
A drawback of the method is that the results of the algorithm greatly depends on the initial configuration (the data points which will act as centroids) and the number of cluster $k$ must be determined beforehand.

\begin{figure}[htbp]
	\centering
		\includegraphics{./Figures/graphs/k_means.png}
		\rule{35em}{0.5pt}
	\caption[K-means]{Outline of the $k$-means algorithm.}
	\label{fig:k_means}
\end{figure}

Generally, the $k$-means methods will minimize the squared error for a clustering $\mathcal{L}$ criterion which is defined as the distance from thedata points the centroid for each cluster in $\mathcal{K}$.
This is expressed as optimizing to a local optimum the energy function
\begin{equation}
	\label{eq:k-means energy}
	e^2(\mathcal{K},\mathcal{L}) =
	\sum_{j=i}^{K}\sum_{i=1}^{n_j}\|\mathbf{x}_i^{(j)} -
	\mathbf{c}_j\|^2
\end{equation}

There are several limitation on the $k$-means method.
One of these is that only spherical shapes of cluster can be generated.
One of the extensions is kernel $k$-means \cite{scholkopf1998nonlinear}, which implicitly projects the data points to a higher dimension and thereby is able to form irregular shaped cluster.

\subsection{Self-organizing Map}

\subsection{Support Vector Machine}

\subsection{Na\"{i}ve Bayes}

\section{Unsupervised clustering of temporal patterns}

### 2013-07-03 Chapters/Chapter3.tex
Rough finish of chapter 2
chapter 3: svm outlier

--- 2013-01-30 Chapters/Chapter3.tex
@@
+% !TEX root = ../main.tex
 % Chapter 3
 
-\chapter{Experiments} % Main chapter title
+\chapter{Change detection by Support Vector Machines}
 
-\label{Chapter3} % For referencing the chapter elsewhere, use \ref{Chapter1}
+\label{Chapter3} % For referencing the chapter elsewhere, use \ref{Chapter3}
 
-\lhead{Chapter 3. \emph{Experiments}} % This is for the header on each 
-%page - 
-%perhaps a shortened title
+\lhead{Chapter 3. \emph{Change detection by Support Vector Machines}} % This is for the header on each page - perhaps a shortened title
 
 %----------------------------------------------------------------------------------------
+\section{Outline}
+\emph{Not intented for the reader.}
+\begin{itemize}
+  \item Explain the connection between change detection and outlier detection
+  \item Explain why and how SVM can be used for outlier detection, and thus for change detection in time series
+  \item Explain options when using SVM, such as RBF kernel
+  \item Relate characteristics of (accelerometer) data to the used (RBF?) kernel
+  \item Define some quality metrics --> this should be in a different chapter (About experiments)
+  \item Note: try to avoid the perspective of \emph{density estimation}. Instead: \emph{data description} or \emph{boundary description}.
+\end{itemize}
 
-\section{}
+\input{./Chapters/Chapter3/outlier_detection}
+\input{./Chapters/Chapter3/one_class_svm}
+\input{./Chapters/Chapter3/svm_parameters}
+\input{./Chapters/Chapter3/data_characteristics}
+\input{./Chapters/Chapter3/quality_metrics}

### 2013-07-03 Chapters/Chapter3/one_class_svm.tex
chapter 3: svm outlier

% !TEX root = ../../main.tex
\section{One-Class Support Vector Machine}\label{sec:one_class_svm}

### 2013-07-03 Chapters/Chapter3/outlier_detection.tex
chapter 3: svm outlier

% !TEX root = ../../main.tex
\section{Outlier and change detection}\label{sec:outlier_and_change_detection}

### 2013-07-04 Appendices/AppendixC.tex
wip

% !TEX root = ../main.tex
% Appendix C

\chapter{Summary of papers and principle formulas}\label{AppendixC}
\lhead{Appendix C. \emph{Paper summaries and formulas}}

\section{Change point detection in time series data using support vectors, by Camci}
Paper: \cite{camci2010change}.
The main concept of this paper is to construct a hypersphere around the data and thereby generating a boundary.
A change point is detected when the radius of the hypersphere grows or shrinks significantly, or when a data point falls outside the boundary.

The main cost function being minimized:
%
\begin{equation}
  \minimize{r}
    {r^2 + C \sum_{i} \xi_i \norm{x}}
    {\norm{\vectorsym{x}_i - c}^2 \leq r^2 + \xi_i, \quad \xi_i \geq 0 \quad \forall i, \vectorsym{x}_i: i\text{th data point}}
\end{equation}
%
Where $r$ is the radius of a (hyper)circle with center $\vectorsym{c}$, $C$ is the penalty coefficient for every outlier and $\xi_i$ is the distance from the $i$th data point to hypersphere (also known as the slack variable).

The dual form by introducing the Lagrange multipliers ($\alpha_i, \alpha_i \geq 0$) and eliminating the slack variables $\xi$ is:
%
\begin{equation}
  \maximize{\vectorsym{\alpha}}
    {\sum_i \alpha_i (\vectorsym{x}_i \cdot \vectorsym{x}_i) - \sum_{i,j} \alpha_i \alpha_j (\vectorsym{x}_i \cdot \vectorsym{x}_j)}
    {0 \leq \alpha_i \leq C \quad\forall i, \quad \sum_i \alpha_i = 1}
\end{equation}
%
To allow for a non-linear relation between the data points and the data boundary, the inner product can be replaced by a (\eg Gaussian) kernel function: $K(\vectorsym{x}_i, \vectorsym{x}_j)$.

%---------------------------------------------
%---------------------------------------------
\section{Change-Point detection in time-series data by Direct Density-Ratio Estimation}
Paper: \cite{kawahara2009change}.

The density ratio $w(\matrixsym{Y})$ is modeled by a Gaussian kernel over sequences $\matrixsym{Y}$ of samples (sequence $\matrixsym{Y}_{te}(l)$ is the test sequence from the $l$th position on):
%
\begin{equation}
  \hat{w}(\matrixsym{Y}) = \sum_{l=1}^{n_\text{te}} \alpha_l K_\sigma(\matrixsym{Y}, \matrixsym{Y}_\text{te}(l)),
\end{equation}
%
where $\{\alpha_l\}_{l=1}^{n_\text{te}}$ are parameters to be learned from the data samples and $K_\sigma(\matrixsym{Y}, \matrixsym{Y}')$ is the Gaussian kernel function with mean $\matrixsym{Y}'$ and standard deviation $\sigma$.
The learned parameters minimize the Kullback-Leibler divergence from the sequence to the test sequence.
The maximization problem then becomes:
%
\begin{equation}
  \maximize{\{\alpha_l\}_{l=1}^{n_\text{te}}}
    {\sum_{i=1}^{n_\text{te}} \text{log} \left(\sum_{l=1}^{n_\text{te}} \alpha_l K_\sigma(\matrixsym{Y}_\text{te}(i), \matrixsym{Y}_\text{te}(l) ) \right)}
    {\frac{1}{n_\text{rf}} \sum_{i=1}^{n_\text{rf}} \sum_{l=1}^{n_\text{te}} \alpha_l K_\sigma(\matrixsym{Y}_\text{rf}(i), \matrixsym{Y}_\text{te}(l)) = 1, ~ \text{and} ~ \alpha_1, \dots ,\alpha_{n_\text{te}} \geq 1}
\end{equation}
%
With the estimated parameters the logarithm of the likelihood ratio between the test and reference interval can be calculated, which signals a change point if it is beyond a certain threshold $\mu$:
%
\begin{equation}
  S = \sum_{i=1}^{n_\text{te}} \text{ln}
    \frac
    {p_\text{te}(\matrixsym{Y}_\text{te}(i))}
    {p_\text{rf}(\matrixsym{Y}_\text{te}(i))}
\end{equation}

%---------------------------------------------
%---------------------------------------------
\section{Joint segmentation of multivariate time series with hidden process regression for human activity recognition, by Chamroukhi}
Paper: \cite{chamroukhi2013joint}.
This approach models the time series data by a parameterized regression, filtered with a \gls{hmm} to smooth out high frequency activity transitions.
With each observation $i$, generated by a $K$-state hidden process, an activity label $z_i$ (and thus sequence) is associated.
Observations follow a regression model:
%
\begin{equation}
  y_i = \beta_{z_i}^T \vectorsym{t}_i + \sigma_{z_i} \epsilon_i; \quad \epsilon_i \sim \mathcal{N}(0,1), \quad (i=1, \dots ,n)
\end{equation}
%
The observations get a label assigned by maximizing the logistic probability ($\pi_k$):
%
\begin{equation}
  \hat{z}_i = \argmaxold \limits_{1 \leq k \leq K} \pi_k (t_i; \hat{\vectorsym{w}}), \quad (i=1, \dots, n)
\end{equation}
%

### 2013-07-08 Appendices/AppendixC.tex
Work on mathematical formulas

@@
 
-\section{Change point detection in time series data using support vectors, by Camci}
+\section{Change point detection in time series data using support vectors, by Camci}\label{sec:camci}
 Paper: \cite{camci2010change}.
@@
 %---------------------------------------------
+\clearpage
 \section{Change-Point detection in time-series data by Direct Density-Ratio Estimation}
@@
 %---------------------------------------------
+\clearpage
 \section{Joint segmentation of multivariate time series with hidden process regression for human activity recognition, by Chamroukhi}
@@
 %
+
+%---------------------------------------------
+%---------------------------------------------
+\clearpage
+\section{Support Vector Data Description, by Tax and Duin}
+Paper: \cite{tax2004support}.
+The paper proposes the \gls{svdd} method, analogous to the \gls{svc} of Vapnik \cite{vapnik1998statistical}, based on the separating hyperplane of Sch{\"o}lkopf \etal \cite{scholkopf1999sv}.
+Where \gls{svc} is ablo to distinguish data between two classes, \gls{svdd} obtains a closed boundary around the target class and can detect outliers.
+
+Method and formulas very similair to description in Section \ref{sec:camci}.
+
+%---------------------------------------------
+%---------------------------------------------
+\clearpage
+\section{Support Vector Density Estimation, by Weston et al.}
+Paper: \cite{weston1999support}.
+Using the notation of \cite{weston1999support}, the distribution function of a density function $p(x)$ is represented as:
+%
+\begin{equation}
+  F(x) = P(X \leq x) = \int_{- \infty}^x p(t)\,\mathrm{d}t
+\end{equation}
+%
+To find the density the following linear equation need to be solved:
+%
+\begin{equation}
+  \int_{-\infty}^\infty \theta(x-t)p(t)\,\mathrm{d}t = F(x)
+\end{equation}
+%
+where
+%
+\[
+  \theta(x) =
+  \begin{cases}
+    1,& x > 0 \\
+    0,& otherwise
+  \end{cases}
+\]
+%
+In this problem the distribution function $F(x)$ is unknown and instead we are given the \gls{idd} data $x_1,\dots,x_l$ generated by $F$.
+
+The empirical distribution function can now be constructed as:
+%
+\begin{equation}
+  F_l(x) = \frac{1}{l} \sum_{i=1}^l \theta(x-x_i)
+\end{equation}
+
+%---------------------------------------------
+%---------------------------------------------
+\clearpage
+\section{An online algorithm for segmenting time series, by Keogh et al.}
+Paper: \cite{keogh2001online}.
+This method approximates the signal with piecewise linear representation.
+Change points are encountered at the time at which a new segment is used to represent the signal.
+The method uses linear regression by taking the best fitting line in the least squares sense, since that minimizes the Euclidian distance which is used as a quality metric.
+Linear interpolation is considered but since that always has a greater sum of squares error it is disregarded.
+
+Linear regression assumes a relation from $n$ observations $\matrixsym{X}$ to the dependend variable $\vectorsym{y}$ using the parameter vector $\vectorsym{\beta}$:
+%
+\begin{equation}
+  \vectorsym{y} = \matrixsym{X}\vectorsym{\beta} + \epsilon
+\end{equation}
+%
+%
+The error function, the sum of squared residuals, being minimized by searching for the best estimation of $\vectorsym{\beta}$ is:
+%
+\begin{equation}
+  \argmin{b} = \sum_{i=1}^n (y_i - x_i' b)^2 = (\vectorsym{y}-\matrixsym{X}b)^T (\vectorsym{y} - \matrixsym{X}b)
+\end{equation}
+%

### 2013-07-08 Appendices/AppendixC.tex
WIP

@@
 %
+
+%---------------------------------------------
+%---------------------------------------------
+\clearpage
+\section{Online novelty detection on temporal sequences, by Ma and Perkins}
+Paper: \cite{ma2003online}.
+This method uses support vectors for regression (in contrast with of classification).
+The regression function, using a kernel function $K(x_i, x_j)$ can be written as:
+%
+\begin{equation}
+  f(\vectorsym{x}) = \sum_{i=1}^l \theta_i K(\vectorsym{x}_i, \vectorsym{x}) + b,
+\end{equation}
+%
+where $\theta_i$ is a coefficient resulting from the Lagrange multipliers of the original minimization problem.
+A small fraction these of coefficients are non-zero, and the corresponding samples $\vectorsym{x}_i$ are the \emph{support vectors}.
+The regression function $f(\vectorsym{x})$ is non-linear when a non-linear kernel is chosen.
+
+The regression function is used to created a model of past observations.
+A matching function is constructed which determines the matching value $V(t_0)$ of a new observations with the constructed model.
+This matching value is the residual of the regression function at $t_0$.
+
+The algorithm determines \emph{(novel) events}, \emph{occurrences} and \emph{surprises}.
+\emph{Novel events} are defined as a series of observations for which the confidence value over the number of suprises (out-of-model observations) is high enough.
+Events thus have a length; they are constructed of a sub-series of observations.
+
+The papers presents an alternative implementation in order to handle fixed-resource environments and thus induce an online algorithm.
+After $W$ observations have been observed and used for the trained model, the oldest observation is disregarded before the newly obtained observation is incorporated.
+
+\emph{Note:} the Support Vector approach in this paper is used to select the observations to use in the regression model.
+This differs from one-class applications of support vector machines.
+The same authors have also presented a paper which does use one-class construction using support vector machines: \cite{ma2003time}.

### 2013-07-09 Appendices/AppendixC.tex
Add paper to Appendix C

@@
 Paper: \cite{tax2004support}.
-The paper proposes the \gls{svdd} method, analogous to the \gls{svc} of Vapnik \cite{vapnik1998statistical}, based on the separating hyperplane of Sch{\"o}lkopf \etal \cite{scholkopf1999sv}.
+The paper proposes the \gls{svdd} method, analogous to the \gls{svc} of Vapnik \cite{vapnik1998statistical}, based on the separating hyper-plane of Sch{\"o}lkopf \etal \cite{scholkopf1999sv}.
 Where \gls{svc} is ablo to distinguish data between two classes, \gls{svdd} obtains a closed boundary around the target class and can detect outliers.
@@
 The same authors have also presented a paper which does use one-class construction using support vector machines: \cite{ma2003time}.
+
+%---------------------------------------------
+%---------------------------------------------
+\clearpage
+\section{Time-series novelty detection using one-class support vector machines, by Ma and Perkins}
+Paper: \cite{ma2003time}
+This approach is very similar to the other paper of this author discussed in the preview section \cite{ma2003online}.
+The difference is that this method does create a \gls{svm}-classifier to detect in and out of class examples, whilst the other uses the support vectors to construct a regression function.
+
+The method constructs a hyper-plane which separates as many as possible data points in the feature space with the largest margin from the origin.
+This is a different from (and more like the original \gls{svm}-proposal by Sch\"olkopf \cite{scholkopf1999support}) the one-class methodology by Tax which creates a boundary around the data \cite{tax2004support}.
+
+The hyper-plane in feature space is represented as:
+%
+\begin{equation}\label{eq:hyper-plane-ma}
+  f(\vectorsym{x}) = \matrixsym{W}^\transpose \Phi(\vectorsym{x}) - \rho,
+\end{equation}
+%
+where $\Phi(\vectorsym{x})$ maps a vector $\vectorsym{x}$ from the input space $I$ to the (potentially infinite dimensional) feature space $F$.
+$\matrixsym{W}$ and $\rho$ are determined by solving a quadratic optimization problem.
+The dual formulation (using Lagrange multipliers $\alpha_i$) is:
+%
+\begin{equation}
+  \matrixsym{W} = \sum_{i=1}^l \alpha_i \Phi(\vectorsym{x}_i),
+\end{equation}
+%
+where $0 \leq \alpha_i \leq \frac{1}{\nu l}$.
+The parameter $\nu \in (0,1)$ is set to trade-off the smoothness of $f(\vectorsym{x})$ and acts as a upper bound on the fraction of outliers over all the data examples in $\vectorsym{x}$ \cite{scholkopf1999support}.
+
+Using the \emph{kernel trick} the inner product of two vectors in feature space $F$ can be replaced by a kernel function $K$, which is often the \gls{rbf}.
+The equation of the hyper-plane (\ref{eq:hyper-plane-ma}) then becomes the following (non-linear) function:
+%
+\begin{equation}
+  f(\vectorsym{x}) = \sum_{i=1}^l \alpha_i K(\vectorsym{x}_i, \vectorsym{x}) - \rho
+\end{equation}
+%
+
+The form of the input vector $\vectorsym{x}$ is considered to be the \emph{(projected) phase space} representation of the original time series.
+Just like \cite{kawahara2012sequential} which constructs \emph{sequences} of samples and in contrast with \cite{camci2010change}, each element of $\vectorsym{x}$ is a vector with the size of the \emph{embedding dimension} $E$ of the time series.
+Thus, a time series $x(t)$ is converted to a set of vectors $T_E(N)$:
+%
+\begin{equation}
+  T_E(N) = \{ \vectorsym{x}_E(t),\ t = E \cdots N\},
+\end{equation}
+%
+where
+%
+\begin{equation}
+  \vectorsym{x}_E(t) = [x(t - E + 1)\ x(t - E + 2)\ \cdots\ x(t)]
+\end{equation}
+%
+If a point of this vector $\vectorsym{x}_E(t)$ is regarded (in the feature space $F$) as an outlier, all corresponding values in the original time series are also regarded as such.

### 2013-07-10 Appendices/AppendixC.tex
WIP

@@
 If a point of this vector $\vectorsym{x}_E(t)$ is regarded (in the feature space $F$) as an outlier, all corresponding values in the original time series are also regarded as such.
+
+%---------------------------------------------
+%---------------------------------------------
+\section{Least squares one-class support vector machine, by Choi}
+Paper: \cite{choi2009least}.
+The proposed method uses a \gls{svm} for a similarity/distance comparison for testing examples to training examples.
+Instead of other methods, such as the standard one-class \gls{svm} by Sch\"olkopf \cite{scholkopf1999support} or the \gls{svdd} of Tax and Duin \cite{tax2004support}, it does not create a boundary for the training data.
+Instead it uses a least-squared approach to construct a hyperplane to which most of the training examples lie close to.
+
+The objective function to be minimized of the standard one-class method by Sch\"olkopf is formulated as:
+%
+\begin{equation}
+  \minimize{\vectorsym{w}}
+    {\frac{1}{2} \norm{\vectorsym{w}}^2 - \rho + C \sum_j \xi_j}
+    {\vectorsym{w} \cdot \phi(\vectorsym{x}_j) \geq \rho - \xi_j \quad \text{and} \quad \xi_j \geq 0}
+\end{equation}
+%
+where $\phi$ is a mapping to the feature space.
+
+The least-squares one-class support vector machine has a small variation on the above formula, which results in the following minimization problem:
+%
+\begin{equation}\label{eq:minimize_least_squares_one_class_svm}
+  \minimize{\vectorsym{w}}
+    {\frac{1}{2} \norm{\vectorsym{w}}^2 - \rho + \frac{1}{2} C \sum_j \xi_j^2}
+    {\vectorsym{w} \cdot \phi(\vectorsym{x}_j) = \rho - \xi_j}
+\end{equation}
+%
+The slack variable (for which in the original formulation $\xi_j \geq 0$ should hold) now represents an error caused by a training example $\vectorsym{x}_j$ with relation to the hyperplane, \ie $\xi_j = \rho - \vectorsym{w} \cdot \phi(\vectorsym{x}_j)$.
+
+In other words, the minimization problem (\ref{eq:minimize_least_squares_one_class_svm})) results in a hyperplane with maximal distance from the origin and for which the sum of the squares of errors $\xi_j^2$ are minimized.

### 2013-08-26 Appendices/AppendixD-planning.tex
WIP

% !TEX root = ../main.tex
% Appendix D

\chapter{Planning} % Main appendix title

\label{AppendixD}

\lhead{Appendix D. \emph{Planning}}

\begin{center}
  \begin{tabulary}{\textwidth}{|c|L|L|c|}
    \hline
    Chapter & Title & Extra & Deadline \\
    \hline \hline
    1 & Introduction, abstract & Write at last & 15/10/2013 \\
    \hline
    2 & Literature review & Partially writting. Maybe focus more on SVM and quick look at other segmentation techniques & 15/9/2013 \\
    \hline
    3 & (Change detection By) One-Class Support Vector Machines & Mix of current chapter and blog post & 1/10/2013 \\
    \hline
    4 & Proposed method & Propose new method, or just setup of testing the OC-SVMs with accelerometer data & Date \\
    \hline
    5 & \textbf{Results} & Test with known and public databases of accelerometer data. For this, two implementations (in Matlab?) are needed & 15/8/2013 \\
    \hline
    6 & Real-world applications & Test with own labeled accelerometer data & 1/9/2013 \\
    \hline
    7 & Conclusion & Write last, together with introduction & 15/10/2013 \\
    \hline
  \end{tabulary}
\end{center}

### 2013-08-26 Chapters/Chapter3/outlier_detection.tex
WIP

@@
 % !TEX root = ../../main.tex
-\section{Outlier and change detection}\label{sec:outlier_and_change_detection}
+\section{Outlier and change detection}\label{sec:outlier_detection}
+
+\subsection{General framework for outlier detection}
+In \cite{schubert2012local} a general framework for outlier detection is proposed.
+The authors compare existing methods and identified the common building block of the algorithms.
+The focus is on unsupervised methods, using information from a local selection of data objects for the detection of outliers.
+The following common algorithmic building blocks are identified:
+
+\begin{enumerate}
+  \item Context: a ``local'' context of an object $o$ for model building
+  \item Model: the method used for building the model
+  \item Reference: a ``reference'' context of object $o$ for model comparison
+  \item Comparison: the method used for model comparison
+  \item Normalization: a (global) normalization procedure
+\end{enumerate}
+
+Here the \emph{context} and \emph{reference} are sets of objects used for model building and model comparison, respectively.
+
+\subsection{Regression, Classification, etc}
+
+General framework of outlier detection, change detection in context of (simple?) regression, classification.
+
+\subsection{M-estimators}
+*** Move this section to somewhere in the end; is specification/extension, not base material ***
+To make a method less sensitive to outliers (in the training data) techniques from robust statistics are applied.
+The term \emph{robustness} has many interpretations, one of them that it ``signifies insensi­tivity to small deviations from the [prior] assumptions [about the underlying situation]'', according to Huber \cite{huber2009robust}.
+Methods have been proposed to apply robust statistics in the form of M-estimators to \gls{svm}s, such as \cite{choi2009least,chen2004m,suykens1999least}

### 2013-09-02 Appendices/AppendixE-Data-structure.tex
Add quotes on performance of SVMs

% !TEX root = ../main.tex
% Appendix E

\chapter{Data structure quotes}

\cite{muller1997predicting}, ``Predicting Time Series with Support Vector Machines'', p.6. \\
``Our experiments show that SVR methods work particularly well if the data is sparse (i.e. we have little data in a high dimensional space). This is due to their good inherent regularization properties.''

\cite{mountrakis2011support}, ``Support vector machines in remote sensing: A review'', p.10. \\
``Most of the findings show that there is empirical evidence to support the theoretical formulation and motivation behind SVMs.
The most important characteristic is SVM’s ability to generalize well from a limited amount and/or quality of training data.
Compared to alternative methods such as backpropagation neural networks, SVMs can yield comparable accuracy using a much smaller training sample size.
This is in line with the ``support vector'' concept that relies only on a few data points to define the classifier's hyperplane.
This property has been exploited and has proved to be very useful in many of the applications we have seen thus far, mainly because acquisition of ground truth for remote sensing data is generally an expensive process.''\\
(And more)

\cite{camps2008kernel}, ``Kernel-based framework for multitemporal and multisource remote sensing data classification and change detection'', p.30. \\
``As core learners, the binary SVC and the one-class SVDD classiﬁer were used, and they were also benchmarked with neural networks in real scenarios.
In general, neural networks show inferior results compared to non-linear kernel classiﬁers, which is a direct consequence of their difﬁculties when working with very high dimensional input samples particularly important when stacking together other information sources such as contextual or multi-temporal''

### 2013-09-23 Chapters/Chapter4/change_indication.tex
Writing on method

% !TEX root = ../../main.tex
\section{Change Indication}\label{sec:method_change_indication}

### 2013-09-23 Chapters/Chapter4/overview.tex
Writing on method

% !TEX root = ../../main.tex
\section{Overview}\label{sec:method_overview}

\begin{figure}
  \centering
    \includegraphics{./Figures/graphs/chapter4/method_setup.pdf}
  \caption[Method setup]{Schematic representation of the change detection method.}
  \label{fig:method_overview}
\end{figure}

This section gives a description of the method used for the experiments and change detection mechanism.
First described is the method to process the gathered sensor data.
A schematic overview is given in figure \ref{fig:method_overview} and shows the steps of the method.
A more detailed explanation of the ``Update model'' step follows.
This section is finalized with the experiments setup, annotation of data streams and quality measure.

\subsection{Change detection method}
As graphically represented in figure \ref{fig:method_overview}, the change detection method starts by processing the data from sensor, such as the accelerometer, magnetic orientation and rotation meaures \footnote{Something about the origin of the streams; all from the same sensor or different sensors?}.

The first step is to process the raw streams of data originating from a multiple of sensors.
The two processes applied are alignment and normalization.
Due to noisy sampling, not all the timestamps in the data streams are sensored at the same timestamp.
Since the \gls{svdd} method requires all the data stream at every timestamp and can not handle missing data on one of the timestamps, all the unique timestamps are filtered out.
Whilst this results in an overall filtering effect, in practice between $1\%$ and $5\%$ of each data stream is disregarded.
The effect of this filtering is not significant and the data is not modified.

Due to the nature of the sensor signals, a normalization step is required in order to set the weight for all the data streams equal.
The range of the accelerometer signal typically spans $-20$ to $20$, the magnetic field from $-45$ to $45$ and the rotations range is from $-1$ to $1$.
This means that a relative small change in the accelerometer stream could have a much larger impact on the model than the same (absolute) change in the rotation stream, whilst the latter has a larger relative impact.
The normalization step ensures that all data is weighted equally and changes in the data are all proportional.

In step $3$ the \gls{svdd} model is initialized.
The first full window over the data stream is used to construct an initial model.
During the initialization the parameters for the \gls{svdd} are provided, begin the kernel type (radial), \gls{rbf} with $\sigma$ and the outlier-fraction $C$.

Step $4$ is executed for every step-size $s$ data points in the stream.
Every update the oldest $s$ data points are removed from and $s$ new data points are added to the \gls{svdd} model.

\subsection{Model updating}

\subsection{Experiments}

### 2013-10-01 Appendices/AppendixA-summaries.tex
wip
Fix typos, add comments

@@
 The regression equation $y = \bar{y} + \hat{b}(x - \bar{x})$ can be used to obtain a decision boundary.
-We need to determine the point $(x_0, y_0)$ such that $y_0$ is half-way between $y^+$ and $y^-$ (the positive and negative examples, i.e. $y_0 = 0$).
+We need to determine the point $(x_0, y_0)$ such that $y_0$ is half-way between $y^+$ and $y^-$ (the positive and negative examples, \ie $y_0 = 0$).
 
@@
 \begin{aligned}
-  & \vectorsym{w}^*, t^* = \argmin\limits_{\vectorsym{w},t} \frac{1}{2} \norm{w}^2 & & \text{subject to } y_i(\vectorsym{w} \cdot \vectorsym{x}_i - t) \geq 1, 1 \leq i \leq n
+  \argmin{\vectorsym{w},t} = \frac{1}{2} \norm{w}^2 & & \text{subject to } y_i(\vectorsym{w} \cdot \vectorsym{x}_i - t) \geq 1, 1 \leq i \leq n
 \end{aligned}
@@
 The dual problem maximizes this function under positivity constraints and one equality constraint:
-
+*** TODO: fix equation (now commented) ***
 \begin{equation}
 \begin{aligned}
-  \alpha_1^*, \dots, \alpha_n^* = \argmax\limits_{\alpha_1, \dots, \alpha_n} - \frac{1}{2} \sum\limits_{i=i}^n \sum\limits_{j=1}^n \alpha_i \alpha_j y_i y_j \vectorsym{x}_i \cdot \vectorsym{x}_j + \sum\limits_{i=1}^n \alpha_i \\
+  % \alpha_1^*, \dots, \alpha_n^* = \argmax\limits_{\alpha_1, \dots, \alpha_n} - \frac{1}{2} \sum\limits_{i=i}^n \sum\limits_{j=1}^n \alpha_i \alpha_j y_i y_j \vectorsym{x}_i \cdot \vectorsym{x}_j + \sum\limits_{i=1}^n \alpha_i \\
   \text{subject to } \alpha_i \geq 0, 1 \leq i \leq n \text{ and } \sum\limits_{i=1}^n \alpha_i y_i = 0
@@
 where $\sigma$ is known as the \emph{bandwidth} parameter.
-We can think of the Gaussian kernel as imposing a Gaussian (i.e., multivariate normal) surface on each support vector in instance space, so that the boundary is defined in therms of those Gaussian surfaces.
+We can think of the Gaussian kernel as imposing a Gaussian (\ie, multivariate normal) surface on each support vector in instance space, so that the boundary is defined in therms of those Gaussian surfaces.
 Kernel methods are best known in combination with support vector machines.

### 2013-10-01 Chapters/Chapter2.tex
Fix typos, add comments

@@
       \item Support Vector Machines (?) - if there are more sources about this
-      \item (Dimensionality reduction)
+      \item (Dimensionality reduction) --> probably not
+      \item \emph{Not to much attention to all techniques, focus is on SVM}
     \end{itemize}
@@
 Many applications require the detection of time points at which the underlying properties of a system change.
-This problem thus has received a lot of attention in the fields of data mining, etc... *** list and refs ***.
-Often this problem is formulated in a statistical framework, by inspecting the data generating \gls{pdf} of the time series data.
+This problem has received a lot of attention in the fields of data mining, etc... *** list and refs ***.
+Often this problem is formulated in a statistical framework, by inspecting the underlying data generating \gls{pdf} of the time series data.
 A change point is then defined as a significant change in the properties of the \gls{pdf}, such as the mean and variance.
@@
 Many methods rely on pre-specified parametric model assumptions and considers the data be independent over time, which makes it less flexible to real-world applications.
-The proposed methods by Kawahara \etal \cite{kawahara2009change} and Lui \etal \cite{liu2013change} try to overcome these problems by estimating the \emph{ratio} between the \gls{pdf}, instead of estimating each \gls{pdf}.
+The methods proposed by Kawahara \etal \cite{kawahara2009change} and Lui \etal \cite{liu2013change} try to overcome these problems by estimating the \emph{ratio} between the \gls{pdf}, instead of estimating each \gls{pdf}.
 This approach is discussed and analyzed in section \ref{density-ratio}.
@@
 One problem it tries to overcome is the (claimed) weakness of many methods to detect a decrease in variance.
-The method represents the distribution over the data points as a hyper-sphere in a higher dimension using kernel trick.
+The method represents the distribution over the data points as a hyper-sphere in a higher dimension using the kernel trick.
 A change in the \gls{pdf} is represented by a change in the radius of this sphere.
 Section \ref{svm} discusses the \gls{svm}-method.
+*** Put emphasis that this is the source of inspiration for the chosen method? ***
 

### 2013-10-01 Chapters/Chapter2/cusum.tex
Fix typos, add comments

@@
 
-Non-Bayesian change detection algorithm (thus: no prior distribution believe available for the change time).
+\emph{Notes:}
+Non-Bayesian change detection algorithm (\ie no prior distribution believe available for the change time).
 The \gls{cusum} method is developed by Page \cite{page1954continuous} for the application of statistical quality control (it is also known as a control chart).
@@
 For a series with homogeneous variance, the value of $D_k$ will oscillate around $0$.
-When there is a sudden change, the value will increase and exceed some predefined boundary with high probability.
+In case of a sudden change, the value will increase and exceed some predefined boundary with high probability.
 The behavior of $D_k$ is related a Brownian bridge.

### 2013-10-01 Chapters/Chapter2/density_ratio.tex
Fix typos, add comments

@@
 A change is detected when $S > \mu$, for some predetermined threshold $\mu$.
-The question is then to how calculate the density ratio
+The question is then how to calculate the density ratio
 \begin{equation}
@@
 Finally the logarithm of the likelihood ratio is evaluated.
-If it is beyond the threshold $\mu$ the current time is reported as a change point.
+If it is above the threshold $\mu$ the current time is reported as a change point.
 
-Improvements in this line of research, by Liu \etal \cite{liu2013change} has lead the application of improved density-ratio estimation methods to the problem of change detection.
-Such an improvement is the Unconstrained Least-Squares Importance Fitting (uLSIF) method \cite{kanamori2009least} and an extension which possess a superior non-parametric convergence property: Relative uLSIF (RuLSIF) \cite{yamada2013relative}.
+Improvements in this line of research, by Liu \etal \cite{liu2013change} has led the application of improved density-ratio estimation methods to the problem of change detection.
+Such an improvement is the Unconstrained Least-Squares Importance Fitting (uLSIF) method \cite{kanamori2009least} and an extension which possesses a superior non-parametric convergence property: Relative uLSIF (RuLSIF) \cite{yamada2013relative}.

### 2013-10-01 Chapters/Chapter2/svm.tex
Fix typos, add comments

@@
 When combined with a mapping function, which maps the data from the input space $I$ to a higher dimension feature space $F$, the input data can be non-linear separable.
-The linear hyperplane, which segments the data in the feature space $F$, yields to a non-linear segmentation in the lower-dimensional input space $I$.
+A separating linear hyperplane that segments the data in the feature space $F$, yields a non-linear segmentation in the lower-dimensional input space $I$.
 Instead of explicitly mapping the input data to the higher dimensional space, a kernel function can be used.
@@
 The retrieved (high dimension) data point can be in- or outside the earlier constructed hyper-sphere, thereby giving information about a possible change point.
+*** NOTE: using the in/out position of a new datapoint is different from the approach in this thesis, by only using the model properties. ***
 Following the definition of Camci \cite{camci2010change}, the class description is obtained by minimizing $r^2$:
@@
 
-In case the second window of data points does not belong to the same class, i.e.
-the probability density function that describes the data differs from the first, the describing values of the second window will differ from the first.
+In case the second window of data points does not belong to the same class, \ie the probability density function that describes the data differs from the first, the describing values of the second window will differ from the first.
 A difference in the SVM center $c$ or radius $r$ represent a change in the mean and variance, respectively.
@@
 *** Give visual explanation with circles and in- and out-side new data points ***
+
+*** Make clear what approach we use (model properties) and create link/bridge to next Chapter (3), which discusses SVM in detail. ***

### 2013-10-01 Chapters/Chapter3/one_class_svm.tex
WIP
Fix typos, add comments

@@
 \section{One-Class Support Vector Machine}\label{sec:one_class_svm}
+*** Use material from blog post ***
+\subsection{Support Vector Machine}
+\emph{Explain classical two-class problem and workings of \gls{svm}; minimization function etc.}
+
+\subsection{Standard One-Class SVM}
+\emph{Explain workings of standard one-class svm by the method of Sch\"olkopf, \cite{scholkopf1999support}}
+
+\subsection{Support Vector Data Description}
+\emph{Explain working of \gls{svdd} by Tax and Duin, \cite{tax2004support}}
+
+*** Maybe add section about ``Change indication'' rather than ``detection''? ***

### 2013-10-01 Chapters/Chapter3/outlier_detection.tex
Fix typos, add comments

@@
 The term \emph{robustness} has many interpretations, one of them that it ``signifies insensi­tivity to small deviations from the [prior] assumptions [about the underlying situation]'', according to Huber \cite{huber2009robust}.
-Methods have been proposed to apply robust statistics in the form of M-estimators to \gls{svm}s, such as \cite{choi2009least,chen2004m,suykens1999least}
+Methods have been proposed to apply robust statistics in the form of M-estimators to \gls{svm}s, such as \cite{choi2009least,chen2004m,suykens1999least}.
+
+*** Question: do slack-variables (allowance of outliers) make SVM robust? ***

### 2013-10-01 Chapters/Chapter3/one_class_svm.tex
Text on two-class SVM

@@
 *** Use material from blog post ***
-\subsection{Support Vector Machine}
+
+\subsection{Support Vector Machine}\label{subsec:svm}
+We will first discuss the traditional two-class \gls{svm} before we consider the one-class variant, as introduced by Cortes and Vapnik in \cite{cortes1995support}.
+Consider a data set $\Omega = \{ (x_1, y_1),\allowbreak (x_2, y_2), \dots , (x_n, y_n) \}$; points $x_i \in \mathbb{R}^d$ in a (for instance two-dimensional) space where $x_i$ is the $i$-th input data point and $y_i \in \{-1, 1\}$ is the $i$-th output pattern, indicating the class membership.
+
+A \gls{svm} can create a boundary between linear-separable data points, making it a non-probabilistic binary linear classifier.
+More flexible non-linear boundaries can be obtained by the use of a non-linear function $\phi(x)$, as illustrated in Figure \ref{fig:kernel_mapping}.
+This function maps the input data from space $I$ to a higher dimensional space $F$.
+The \gls{svm} can create a linear separating hyperplane in the space $F$ that separates the data points from the two classes.
+When the hyperplane is projected to the (lower) original input space $I$ it creates a non-linear separating curve.
+The mapping and projection of data points can be efficient (and implicit) performed by using the kernel trick, which is discussed in section \ref{subsec:kernel_trick}.
+
+The separating hyperplane is represented by
+\begin{equation}
+w^T x + b = 0,
+\end{equation}
+with $w \in F$ and $b \in R$.
+The hyperplane that is created determines the \emph{margin} between the classes; the minimal distance from one of the data points to the hyperplane.
+In geometric sense, $w$ is the normal vector to the hyperplane and $\frac{b}{\lVert{w}\rVert}$ determines the offset of the hyperplane to the origin.
+Since the distance between the two margins is equal to $\frac{2}{\lVert{w}\rVert}$, the maximum-margin hyperplane is found by minimizing $\lVert{w}\rVert$.
+The data points which lie on the margin are the \emph{support vectors}.
+This geometrical interpretation is illustrated in Figure \ref{fig:svm_hyperplane}.
+All data points for which $y_i = -1$ are on one side of the hyperplane and all other data points (for which $y_i = 1$) are on the other side.
+The minimal distance from a data point to the hyperplane is for both classes equal.
+This results in a \emph{maximal margin} between the two classes.
+Thus, the \gls{svm} searches for a maximal separating hyperplane.
+
+\begin{figure}
+\centering
+  \includegraphics[width=0.5\textwidth]{./Figures/chapter3/svm_separating_plane_with_margin.png}
+  \caption[\gls{svm} and the separating hyperplane]{Illustration of the separating hyperplane of a \gls{svm}.
+Here $w$ is the normal vector for the separating hyperplane and the distance between the two margins is $\frac{2}{\lVert{w}\rVert}$.}
+  \label{fig:svm_hyperplane}
+\end{figure}
+
+With every classification method there is a risk of overfitting.
+In that case the random error or noise of the data set is described instead of the underlying data.
+The \gls{svm} classifier can use a \emph{soft margin} by allowing some data points to lie within the margin, instead of on the margin or farther away from the hyperplane.
+For this it introduces \emph{slack variables} $\xi_i$ for each data point and the constant $C > 0$ the determines the trade-off between maximizing the margin and the number of data points within that margin (and thus the training errors).
+The objective function for a \gls{svm} is the following minimization function:
+
+\begin{equation}\label{frm:svm_objective}
+  \operatorname*{min}_{w,\ b,\ \xi_i} \frac{ \lVert{w}\rVert^2 }{2} + C \sum_{i=1}^n \xi_i
+\end{equation}
+\begin{equation}
+  \begin{multlined}
+  \mbox{ subject to: } \\*
+  \begin{aligned}
+  y_i( w^T \phi(x_i) + b) \geq & 1 - \xi_i & \mbox{ for all } i = 1, \dots, n \\*
+   & \xi_i \geq 0 & \mbox{ for all } i = 1, \dots, n\\*
+  \end{aligned}
+  \end{multlined}
+\end{equation}
+
+*** TODO: better format of above formula ***
+
+This minimization problem can be solved (using quadratic programming) and transformed to its Lagrange dual formulation.
+To do so, the Lagrange multipliers $a_i >= 0$ are introduced and the decision function becomes:
+\begin{equation}\label{frm:svm_lagrange}
+  f(x) = \operatorname{sgn}( \sum_{i=1}^n \alpha_i y_i K(x, x_i) + b),
+\end{equation}
+where $K(x, x_i) = \phi(x)^T\phi(x_i)$ (which is further discussed in section \ref{subsec:kernel_trick}).
+Here every data point $i$ for which $a_i > 0$ is weighted in the decision function and thus ``supports'' the classification machine: hence the name ``\acrlong{svm}''.
+Since it is shown that under certain circumstances \gls{svm}s show an equality to sparse representations \cite{girosi1998equivalence}, there will often be relatively few Lagrange multipliers with a non-zero value.
+
+\subsection{Kernel trick}\label{subsec:kernel_trick}
+In the previous section, \ref{subsec:svm}, the mapping function $\phi(x)$ and the kernel function $K$ were briefly mentioned.
+The decision function in equation \ref{frm:svm_lagrange} only relies on the dot-products of mapped data points in the feature space $F$ (\ie all pairwise distances between the data points in that space).
+It shows \cite{flach2012machine} that as long as any function has the same result, without an explicit mapping to the higher dimension $F$, the dot-products can be substituted by the kernel function $K$.
+This is known as the \emph{kernel trick} and gives the \gls{svm} the ability to create non-linear decision function without high computational complexity.
+This mapping is illustrated in Figure \ref{fig:kernel_mapping}.
+Here the non-linear separating boundary in the input space $I$ is mapped, via $\phi$, to a linear boundary in the feature space $F$.
+
+The kernel function $K$ can have different forms, such as linear, polynomial and sigmoidal but the most used (and flexible) form is the Gaussian \gls{rbf}, in which the feature space $F$ is a Hilbert Space of infinite dimensions *** ref needed ***:
+\begin{equation}
+  K(x, x') = \operatorname{exp} \left( - \frac{ \lVert x - x' \rVert ^2}{2 \sigma^2 } \right),
+\end{equation}
+where $\sigma \in R$ is a kernel parameter and $\lVert x - x' \rVert$ is the dissimilarity measure expressed in Euclidean distance.
+
+\begin{figure}
+\centering
+  \includegraphics[width=0.8\textwidth]{./Figures/chapter3/svm_kernel_mapping.png}
+  \caption[Kernel mapping]{The non-linear boundary in the input space $I$ (left) is transformed to a linear boundary in the feature space $F$ (right) by mapping the data points with the function $\phi$.
+The kernel trick uses a function $K$ which performs an implicit mapping.}
+  \label{fig:kernel_mapping}
+\end{figure}
+
 \emph{Explain classical two-class problem and workings of \gls{svm}; minimization function etc.}

### 2013-10-02 Chapters/Chapter3/one_class_svm.tex
Add stuff on Scholkopf one-svm

@@
 \section{One-Class Support Vector Machine}\label{sec:one_class_svm}
-*** Use material from blog post ***
 
+%--------------------------------------------
 \subsection{Support Vector Machine}\label{subsec:svm}
@@
 
-\begin{equation}\label{frm:svm_objective}
+\begin{equation}\label{eq:svm_objective}
   \operatorname*{min}_{w,\ b,\ \xi_i} \frac{ \lVert{w}\rVert^2 }{2} + C \sum_{i=1}^n \xi_i
@@
 
-This minimization problem can be solved (using quadratic programming) and transformed to its Lagrange dual formulation.
+This minimization problem can be solved (using \gls{qp}) and transformed to its Lagrange dual formulation.
 To do so, the Lagrange multipliers $a_i >= 0$ are introduced and the decision function becomes:
-\begin{equation}\label{frm:svm_lagrange}
+\begin{equation}\label{eq:svm_lagrange}
   f(x) = \operatorname{sgn}( \sum_{i=1}^n \alpha_i y_i K(x, x_i) + b),
@@
 
+%--------------------------------------------
+
 \subsection{Kernel trick}\label{subsec:kernel_trick}
 In the previous section, \ref{subsec:svm}, the mapping function $\phi(x)$ and the kernel function $K$ were briefly mentioned.
-The decision function in equation \ref{frm:svm_lagrange} only relies on the dot-products of mapped data points in the feature space $F$ (\ie all pairwise distances between the data points in that space).
+The decision function in equation \ref{eq:svm_lagrange} only relies on the dot-products of mapped data points in the feature space $F$ (\ie all pairwise distances between the data points in that space).
 It shows \cite{flach2012machine} that as long as any function has the same result, without an explicit mapping to the higher dimension $F$, the dot-products can be substituted by the kernel function $K$.
@@
 
-\emph{Explain classical two-class problem and workings of \gls{svm}; minimization function etc.}
+%--------------------------------------------
 
-\subsection{Standard One-Class SVM}
-\emph{Explain workings of standard one-class svm by the method of Sch\"olkopf, \cite{scholkopf1999support}}
+\subsection{Reproducing kernel Hilbert space}
+\emph{Write something on this topic? Or leave it implicit?}
 
-\subsection{Support Vector Data Description}
+%--------------------------------------------
+
+\subsection{One-Class Support Vector Machines}\label{subsec:one-class-svm}
+The classical implementation of a \gls{svm} is to classify a dataset in two distinct classes.
+This is a common usecase, although sometimes there is no training data for both classes available.
+Still, one would like to classify new data points as regular, or in-class, or out-of-class, \eg in the case of a novelty detection.
+With that problem only data from one class is available and the objective is to recognize new data points that are not part of that class.
+This problem is closely related to density estimation.
+In that context, the problem can be the following.
+Given an underlying probabilty distribution $P$ the goal is to find a subset $S$ for which the probability that a data point from $P$ lies outside $S$ equals some predetermined value $\nu$ between $0$ and $1$ \cite{scholkopf1999support}.
+
+We will discuss two implementations of \gls{oc-svm}s.
+The first is the \gls{nu-svm} by Sch\"olkopf \etal \cite{scholkopf1999support} and closely follows the above problem statement regarding density estimation.
+The second, on which we will base our change detection method, is the \gls{svdd} by Tax and Duin \cite{tax2004support}.
+
+%--------------------------------------------
+
+\subsection{\acrlong{nu-svm}s}\label{subsec:nu-svm}
+The first of the \gls{oc-svm} methods we will discuss is often referred to as \gls{nu-svm} and introduced by Sch\"olkopf \etal \cite{scholkopf1999support}.
+Instead of estimating the density function of an distribution $P$, it focuses on an easier problem: the algorithm find regions in the input where the ``probability density lives''.
+This results in a function such that most of the data is in the region where the function is nonzero.
+
+The constructed decision function $f$ resembles the function discussed in Section \ref{subsec:svm}.
+It returns the value $+1$ in a (possibly small) region capturing most of the data points, and $-1$ elsewhere.
+The method maps the data points from input space $I$ to a feature space $F$ (following classical \gls{svm}s).
+In that space $F$ it separates the data points with maximal margin from the origin, with a hyperplane.
+For a new data points $x$, the function value $f(x)$ determines wheter the data point is part of the distribution (\ie the value is $+1$) or a novelty (\ie the value is $-1$).
+The hyperplane is represented by $g(x) = w \cdot \phi(x) + \rho = 0$ and the decision function is $f(x) = \operatorname{sgn}(g(x))$.
+This hyperplane and the separation from the origin is illustrated in Figure \ref{fig:nu-svm}.
+
+The objective function to find the separating hyperplane is the following minimization function, which can be solved using \gls{qp}:
+
+\begin{equation}\label{eq:nu-svm_objective}
+  \operatorname*{min}_{w,\ \xi_i,\ \rho } \frac{\lVert w \rVert ^2}{2} + \frac{1}{\nu n} \sum_{i=1}^n \xi_i - \rho
+\end{equation}
+\begin{equation}
+  \begin{multlined}
+    \mbox{ subject to: } \\
+    \begin{aligned}
+      (w \cdot \phi(x_i)) \geq & \rho - \xi_i & \mbox{ for all } i = 1, \dots, n \\
+      & \xi_i \geq 0 & \mbox{ for all } i = 1, \dots, n \\
+    \end{aligned}
+  \end{multlined}
+\end{equation}
+
+In the classical \gls{svm} objective function, as denoted in Equation \ref{eq:svm_objective}, the parameter $C$ decided the smoothness of the boundary, with respect to the slack variables $\xi_i$.
+In the formulation of \gls{nu-svm} the equivalent parameter is $\nu \in (0,1)$ (hence the name).
+It characterizes the solution in two ways:
+\begin{enumerate}
+  \item
+  \item
+\end{enumerate}
+
+\begin{figure}
+  \centering
+    \includegraphics[width=0.5\textwidth,keepaspectratio]{./Figures/chapter3/nu-svm.pdf}
+  \caption[\gls{nu-svm}]{Graphical representation of \gls{nu-svm}.
+The separating hyperplane $w \cdot \phi(x_i) + \rho = 0$ create a maximal margin, in the feature space, between the data points and the origin.
+Slack variables $\xi_i$ are used to create a soft margin.}
+  \label{fig:nu-svm}
+\end{figure}
+
+%--------------------------------------------
+
+\subsection{\acrlong{svdd}}\label{subsec:svdd}
 \emph{Explain working of \gls{svdd} by Tax and Duin, \cite{tax2004support}}

### 2013-10-03 Chapters/Chapter3/one_class_svm.tex
Complete part on v-SVM

@@
 \end{figure}
+*** Add wikipedia reference. How to do? http://en.wikipedia.org/wiki/File:Svm_max_sep_hyperplane_with_margin.png ***
 
@@
 \end{figure}
+*** Add wikipedia reference. How to do? http://en.wikipedia.org/wiki/File:Kernel_Machine.png ***
 
@@
 
+\begin{figure}
+  \centering
+    \includegraphics[width=0.5\textwidth,keepaspectratio]{./Figures/chapter3/nu-svm.pdf}
+  \caption[\gls{nu-svm}]{Graphical representation of \gls{nu-svm}.
+The separating hyperplane $w \cdot \phi(x_i) + \rho = 0$ create a maximal margin, in the feature space, between the data points and the origin.
+Slack variables $\xi_i$ are used to create a soft margin.}
+  \label{fig:nu-svm}
+\end{figure}
+
 The objective function to find the separating hyperplane is the following minimization function, which can be solved using \gls{qp}:
@@
 
+The decision function in the dual formulation with Lagrange multipliers is denoted as:
+\begin{equation}\label{eq:nu-svm_lagrange}
+f(x) = \operatorname{sgn}((w \cdot \phi(x_i)) - \rho) = \operatorname{sgn}( \sum_{i=1}^n \alpha_i K(x, x_i) - \rho)
+\end{equation}
+
 In the classical \gls{svm} objective function, as denoted in Equation \ref{eq:svm_objective}, the parameter $C$ decided the smoothness of the boundary, with respect to the slack variables $\xi_i$.
@@
 \begin{enumerate}
-  \item
-  \item
+  \item $\nu$ is an upper bound on the fraction of outliers, \ie training examples regarded as out-of-class.
+  \item $\nu$ is a lower bound on the fraction of \gls{sv}s, \ie training examples with a nonzero Lagrange multiplier $\alpha_i$.
 \end{enumerate}
+When $\nu$ approaches $0$, the penalty factor for nonzero Lagrange multipliers ($\frac{1}{\nu n}$) becomes infinite, and thus the solution resembles a \emph{hard margin} boundary.
 
-\begin{figure}
-  \centering
-    \includegraphics[width=0.5\textwidth,keepaspectratio]{./Figures/chapter3/nu-svm.pdf}
-  \caption[\gls{nu-svm}]{Graphical representation of \gls{nu-svm}.
-The separating hyperplane $w \cdot \phi(x_i) + \rho = 0$ create a maximal margin, in the feature space, between the data points and the origin.
-Slack variables $\xi_i$ are used to create a soft margin.}
-  \label{fig:nu-svm}
-\end{figure}
+This method creates a \emph{hyperplane}, characterized by $w$ and $\rho$, that separates the data with maximal margin from the origin in the feature space $F$.
+In the following section we will discuss an alternative method, which uses an circumscribing \emph{hypersphere} to characterize the data (and the region of distribution $P$ where the density (or: support) lives).
 

### 2013-10-03 Chapters/Chapter3/one_class_svm.tex
SVDD

@@
   \caption[\gls{svm} and the separating hyperplane]{Illustration of the separating hyperplane of a \gls{svm}.
-Here $w$ is the normal vector for the separating hyperplane and the distance between the two margins is $\frac{2}{\lVert{w}\rVert}$.}
+Here $w$ is the normal vector for the separating hyperplane and the distance between the two margins is $\frac{2}{\lVert{w}\rVert}$.
+*** Add wikipedia reference.
+How to do?
+***}
   \label{fig:svm_hyperplane}
 \end{figure}
-*** Add wikipedia reference. How to do? http://en.wikipedia.org/wiki/File:Svm_max_sep_hyperplane_with_margin.png ***
 
@@
   \caption[Kernel mapping]{The non-linear boundary in the input space $I$ (left) is transformed to a linear boundary in the feature space $F$ (right) by mapping the data points with the function $\phi$.
-The kernel trick uses a function $K$ which performs an implicit mapping.}
+The kernel trick uses a function $K$ which performs an implicit mapping.
+*** Add wikipedia reference.
+How to do?
+***}
   \label{fig:kernel_mapping}
 \end{figure}
-*** Add wikipedia reference. How to do? http://en.wikipedia.org/wiki/File:Kernel_Machine.png ***
 
@@
 \subsection{\acrlong{svdd}}\label{subsec:svdd}
-\emph{Explain working of \gls{svdd} by Tax and Duin, \cite{tax2004support}}
+The method introduced by Tax and Duin \cite{tax2004support}, known as \acrlong{svdd}, takes as spherical instead of planar approach.
+The boundary, created in feature space $F$, forms a hypersphere around the (high density region of the) data.
+The volumne of this hypersphere is minimized to get the smallest enclosing boundary.
+By allowing outliers using slacks variables, in the same manner as classical \gls{svm} and \gls{nu-svm}, a soft margin is constructed.
+
+The constructed hypersphere is characterized by a center $\mathbf{a}$ and a radius $R > 0$ as distance from the center to (any data point that is a \gls{sv} on) the boundary, for which the volume, and thus the radius $R$, will be minimized.
+
+\begin{equation}\label{eq:svdd_objective}
+  \operatorname*{min}_{R,\ \mathbf{a} } R^2 + C \sum_{i=1}^n \xi_i
+\end{equation}
+\begin{equation}
+  \begin{multlined}
+    \mbox{ subject to: } \\
+    \begin{aligned}
+      \lVert x_i - \mathbf{a} \rVert ^ 2 \leq & R^2 + \xi_i & \mbox{ for all } i = 1, \dots, n \\
+      & \xi_i \geq 0 & \mbox{ for all } i = 1, \dots, n \\
+    \end{aligned}
+  \end{multlined}
+\end{equation}
+
+\begin{figure}
+  \centering
+    \includegraphics[width=0.5\textwidth,keepaspectratio]{./Figures/chapter3/nu-vs-svdd.pdf}
+  \caption[Difference \gls{nu-svm} and \gls{svdd}]{Graphical representation of the difference between \gls{nu-svm} (left) and \gls{svdd} (right).
+Note that for the sake of simplicity the kernel functions are not applied.}
+  \label{fig:nu-vs-svdd}
+\end{figure}
 

### 2013-10-07 Chapters/Chapter3/one_class_svm.tex
Wip

@@
 We will first discuss the traditional two-class \gls{svm} before we consider the one-class variant, as introduced by Cortes and Vapnik in \cite{cortes1995support}.
-Consider a data set $\Omega = \{ (x_1, y_1),\allowbreak (x_2, y_2), \dots , (x_n, y_n) \}$; points $x_i \in \mathbb{R}^d$ in a (for instance two-dimensional) space where $x_i$ is the $i$-th input data point and $y_i \in \{-1, 1\}$ is the $i$-th output pattern, indicating the class membership.
+Consider a data set $\Omega = \{ (x_1, y_1),\allowbreak (x_2, y_2), \dots , (x_n, y_n) \}$; points $x_i \in \mathbb{R}^d$ in a (for instance two-dimensional) space where $x_i$ is the $\mathcal{I}$-th input data point and $y_i \in \{-1, 1\}$ is the $\mathcal{I}$-th output pattern, indicating the class membership.
 
@@
 More flexible non-linear boundaries can be obtained by the use of a non-linear function $\phi(x)$, as illustrated in Figure \ref{fig:kernel_mapping}.
-This function maps the input data from space $I$ to a higher dimensional space $F$.
-The \gls{svm} can create a linear separating hyperplane in the space $F$ that separates the data points from the two classes.
-When the hyperplane is projected to the (lower) original input space $I$ it creates a non-linear separating curve.
+This function maps the input data from space $\mathcal{I}$ to a higher dimensional space $\mathcal{F}$.
+The \gls{svm} can create a linear separating hyperplane in the space $\mathcal{F}$ that separates the data points from the two classes.
+When the hyperplane is projected to the (lower) original input space $\mathcal{I}$ it creates a non-linear separating curve.
 The mapping and projection of data points can be efficient (and implicit) performed by using the kernel trick, which is discussed in section \ref{subsec:kernel_trick}.
@@
 Here $w$ is the normal vector for the separating hyperplane and the distance between the two margins is $\frac{2}{\lVert{w}\rVert}$.
-*** Add wikipedia reference.
-How to do?
-***}
+Image from Wikipedia.org\protect\footnote}
   \label{fig:svm_hyperplane}
 \end{figure}
+\footnotetext{\url{http://en.wikipedia.org/wiki/File:Svm_max_sep_hyperplane_with_margin.png}}
 
@@
 where $K(x, x_i) = \phi(x)^T\phi(x_i)$ (which is further discussed in section \ref{subsec:kernel_trick}).
-Here every data point $i$ for which $a_i > 0$ is weighted in the decision function and thus ``supports'' the classification machine: hence the name ``\acrlong{svm}''.
+Here every data point $\mathcal{I}$ for which $a_i > 0$ is weighted in the decision function and thus ``supports'' the classification machine: hence the name ``\acrlong{svm}''.
 Since it is shown that under certain circumstances \gls{svm}s show an equality to sparse representations \cite{girosi1998equivalence}, there will often be relatively few Lagrange multipliers with a non-zero value.
@@
 In the previous section, \ref{subsec:svm}, the mapping function $\phi(x)$ and the kernel function $K$ were briefly mentioned.
-The decision function in equation \ref{eq:svm_lagrange} only relies on the dot-products of mapped data points in the feature space $F$ (\ie all pairwise distances between the data points in that space).
-It shows \cite{flach2012machine} that as long as any function has the same result, without an explicit mapping to the higher dimension $F$, the dot-products can be substituted by the kernel function $K$.
+The decision function in equation \ref{eq:svm_lagrange} only relies on the dot-products of mapped data points in the feature space $\mathcal{F}$ (\ie all pairwise distances between the data points in that space).
+It shows \cite{flach2012machine} that as long as any function has the same result, without an explicit mapping to the higher dimension $\mathcal{F}$, the dot-products can be substituted by the kernel function $K$.
 This is known as the \emph{kernel trick} and gives the \gls{svm} the ability to create non-linear decision function without high computational complexity.
 This mapping is illustrated in Figure \ref{fig:kernel_mapping}.
-Here the non-linear separating boundary in the input space $I$ is mapped, via $\phi$, to a linear boundary in the feature space $F$.
+Here the non-linear separating boundary in the input space $\mathcal{I}$ is mapped, via $\phi$, to a linear boundary in the feature space $\mathcal{F}$.
 
-The kernel function $K$ can have different forms, such as linear, polynomial and sigmoidal but the most used (and flexible) form is the Gaussian \gls{rbf}, in which the feature space $F$ is a Hilbert Space of infinite dimensions *** ref needed ***:
+The kernel function $K$ can have different forms, such as linear, polynomial and sigmoidal but the most used (and flexible) form is the Gaussian \gls{rbf}, in which the feature space $\mathcal{F}$ is a Hilbert Space of infinite dimensions *** ref needed ***:
 \begin{equation}
@@
   \includegraphics[width=0.8\textwidth]{./Figures/chapter3/svm_kernel_mapping.png}
-  \caption[Kernel mapping]{The non-linear boundary in the input space $I$ (left) is transformed to a linear boundary in the feature space $F$ (right) by mapping the data points with the function $\phi$.
+  \caption[Kernel mapping]{The non-linear boundary in the input space $\mathcal{I}$ (left) is transformed to a linear boundary in the feature space $\mathcal{F}$ (right) by mapping the data points with the function $\phi$.
 The kernel trick uses a function $K$ which performs an implicit mapping.
-*** Add wikipedia reference.
-How to do?
-***}
+Image from Wikipedia.org\footnotemark}
   \label{fig:kernel_mapping}
 \end{figure}
+\footnotetext{\url{http://en.wikipedia.org/wiki/File:Kernel_Machine.png}}
 
@@
 
-The constructed decision function $f$ resembles the function discussed in Section \ref{subsec:svm}.
+The constructed decision function $\mathcal{F}$ resembles the function discussed in Section \ref{subsec:svm}.
 It returns the value $+1$ in a (possibly small) region capturing most of the data points, and $-1$ elsewhere.
-The method maps the data points from input space $I$ to a feature space $F$ (following classical \gls{svm}s).
-In that space $F$ it separates the data points with maximal margin from the origin, with a hyperplane.
+The method maps the data points from input space $\mathcal{I}$ to a feature space $\mathcal{F}$ (following classical \gls{svm}s).
+In that space $\mathcal{F}$ it separates the data points with maximal margin from the origin, with a hyperplane.
 For a new data points $x$, the function value $f(x)$ determines wheter the data point is part of the distribution (\ie the value is $+1$) or a novelty (\ie the value is $-1$).
@@
 
-This method creates a \emph{hyperplane}, characterized by $w$ and $\rho$, that separates the data with maximal margin from the origin in the feature space $F$.
+This method creates a \emph{hyperplane}, characterized by $w$ and $\rho$, that separates the data with maximal margin from the origin in the feature space $\mathcal{F}$.
 In the following section we will discuss an alternative method, which uses an circumscribing \emph{hypersphere} to characterize the data (and the region of distribution $P$ where the density (or: support) lives).
@@
 The method introduced by Tax and Duin \cite{tax2004support}, known as \acrlong{svdd}, takes as spherical instead of planar approach.
-The boundary, created in feature space $F$, forms a hypersphere around the (high density region of the) data.
-The volumne of this hypersphere is minimized to get the smallest enclosing boundary.
+The boundary, created in feature space $\mathcal{F}$, forms a hypersphere around the (high density region of the) data.
+The volume of this hypersphere is minimized to get the smallest enclosing boundary.
 By allowing outliers using slacks variables, in the same manner as classical \gls{svm} and \gls{nu-svm}, a soft margin is constructed.
 
+\begin{figure}
+  \centering
+    \includegraphics[width=0.5\textwidth,keepaspectratio]{./Figures/chapter3/nu-vs-svdd.pdf}
+  \caption[Difference \gls{nu-svm} and \gls{svdd}]{Graphical representation of the difference between \gls{nu-svm} (left) and \gls{svdd} (right).
+Note that for the sake of simplicity the kernel functions are not applied.}
+  \label{fig:nu-vs-svdd}
+\end{figure}
+
 The constructed hypersphere is characterized by a center $\mathbf{a}$ and a radius $R > 0$ as distance from the center to (any data point that is a \gls{sv} on) the boundary, for which the volume, and thus the radius $R$, will be minimized.
+The center $\mathbf{a}$ is a linear combination of of the support vectors.
+Like the classical \gls{svm} and \gls{svdd} it can be required that all the distances from the data points $x_i$ to the center $\mathbf{a}$ are strict less then $R$ (or equivalent measure, to create a hard margin) or soft margins are allowed by using slack variables $\xi_i$.
+In the case of a soft margin, the penalty is determined by $C$ and the minimization is expressed as Equation \ref{eq:svdd_objective}.
+This principle is illustrated in the right image or Figure \ref{fig:nu-vs-svdd}.
+Instead of a separating hyperplane, constructed by \gls{nu-svm} and illustrated on the left of the Figure, the \gls{svdd} creates a hypersphere (in the illustration a cricle) around the data points.
+By using kernel functions (\eg the \gls{rbf}) the hyperspheres in the high dimensional feature space $\mathcal{F}$ corresponds to a flexible and tight enclosing boundary in input space $\mathcal{I}$, illustrated in Figure \ref{fig:svdd-boundary}.
 
@@
 
+The dual formulation, by using Lagrange multipliers $\alpha_i$, can be used to test a new data point $z$ and is the following:
+
+\begin{equation}
+  \lVert z - \mathbf{x} \rVert ^2 = \sum_{i=1}^n \alpha_i \operatorname{exp} \left( \frac{- \lVert z - x_i \rVert ^ 2}{\sigma^2} \right) \geq -R^2/2 + C_R
+\end{equation}
+
+%--------------------------------------------
+
+\subsection{Kernel Function}\label{subsec:kernel_function}
+
+\begin{itemize}
+  \item Explain RBF kernel
+  \item Explain RKHS? (Hilbert Space)
+\end{itemize}
+
 \begin{figure}
   \centering
-    \includegraphics[width=0.5\textwidth,keepaspectratio]{./Figures/chapter3/nu-vs-svdd.pdf}
-  \caption[Difference \gls{nu-svm} and \gls{svdd}]{Graphical representation of the difference between \gls{nu-svm} (left) and \gls{svdd} (right).
-Note that for the sake of simplicity the kernel functions are not applied.}
-  \label{fig:nu-vs-svdd}
+    \includegraphics[width=0.5\textwidth,keepaspectratio]{./Figures/chapter3/svdd-boundary.pdf}
+  \caption[\gls{svdd} boundary]{The \gls{svdd} method trained on a banana-shaped data set with different sigma-values for the \gls{rbf} kernel.
+Solid circles are support vectors, the dashed line is the boundary.
+Image by Tax \cite{tax2001one}.}
+  \label{fig:svdd-boundary}
 \end{figure}

### 2013-10-07 Chapters/Chapter3/one_class_svm.tex

@@
 We will first discuss the traditional two-class \gls{svm} before we consider the one-class variant, as introduced by Cortes and Vapnik in \cite{cortes1995support}.
-Consider a data set $\Omega = \{ (x_1, y_1),\allowbreak (x_2, y_2), \dots , (x_n, y_n) \}$; points $x_i \in \mathbb{R}^d$ in a (for instance two-dimensional) space where $x_i$ is the $\mathcal{I}$-th input data point and $y_i \in \{-1, 1\}$ is the $\mathcal{I}$-th output pattern, indicating the class membership.
+Consider a data set $\Omega = \{ (x_1, y_1),\allowbreak (x_2, y_2), \dots , (x_n, y_n) \}$; points $x_i \in \mathbb{R}^d$ in a (for instance two-dimensional) space where $x_i$ is the $i$-th input data point and $y_i \in \{-1, 1\}$ is the $i$-th output pattern, indicating the class membership.
 
@@
 Here every data point $\mathcal{I}$ for which $a_i > 0$ is weighted in the decision function and thus ``supports'' the classification machine: hence the name ``\acrlong{svm}''.
-Since it is shown that under certain circumstances \gls{svm}s show an equality to sparse representations \cite{girosi1998equivalence}, there will often be relatively few Lagrange multipliers with a non-zero value.
+Since it is shown that under certain circumstances \gls{svm}s show an equality to sparse representations \cite{girosi1998equivalence,smola1998connection}, there will often be relatively few Lagrange multipliers with a non-zero value.
 
@@
 
-The kernel function $K$ can have different forms, such as linear, polynomial and sigmoidal but the most used (and flexible) form is the Gaussian \gls{rbf}, in which the feature space $\mathcal{F}$ is a Hilbert Space of infinite dimensions *** ref needed ***:
+The kernel function $K$ can have different forms, such as linear, polynomial and sigmoidal but the most used (and flexible) form is the Gaussian \gls{rbf}.
+As Smola \etal \cite{smola1998connection} state, this Gaussian kernel yields good performance, especially when no assumptions can be made about the data.
+The kernel maps input space $\mathcal{I}$ to the feature space $\mathcal{F}$ which is a Hilbert Space of infinite dimensions *** ref needed ***:
 \begin{equation}
@@
 
+$k(\mathbf{x}_i, \mathbf{x})$ is a kernel function computing a dot product in feature space, introduced by Aizerman \etal \cite{aizerman1964theoretical}.
+``Training a \gls{svm} with Gaussian \gls{rbf} kernels corresponds to minimizing the specific cost function with a regularization operator'' \cite{smola1998connection}. \\
+``Gaussian kernels tend to yield good performance under general smoothness assumptions and should be considered especially of no additional knowledge of the data is available.'' \cite{smola1998connection}.
+\\
+``Choosing a small width of the kernels leads to high generalization error as it effectively decouples the separate basis functions of the kernel expansion into very localized functions which is equivalent to memorizing the data, whereas a wide kernel tends to oversmooth.'' \cite{smola1998connection}.
+\\
+
+sparsity, RKHS: \cite{girosi1998equivalence}
+
 \begin{figure}

### 2013-10-07 Appendices/AppendixC.tex
Overview of papers for literature review 'Segmentation'

@@
 \clearpage
-\section{Joint segmentation of multivariate time series with hidden process regression for human activity recognition, by Chamroukhi}
+\section{Joint segmentation of multivariate time series with hidden process regression for human activity recognition, by Chamroukhi}\label{sec:appendix-C-joint-segmentation}
 Paper: \cite{chamroukhi2013joint}.
@@
 \clearpage
-\section{An online algorithm for segmenting time series, by Keogh et al.}
+\section{An online algorithm for segmenting time series, by Keogh et al.}\label{sec:appendix-C-online-keogh}
 Paper: \cite{keogh2001online}.

### 2013-10-07 Chapters/Chapter2/temporal_segmentation.tex
Overview of papers for literature review 'Segmentation'

% !TEX root = ../../main.tex
\section{Temporal Segmentation}\label{sec:temporal_segmentation}

\begin{itemize}
  \item Given overview of segmentation techniques, for times series data
  \item Use different ``point-of-views'', or terms
  \item ``Segmentation''
  \item ``Change detection''
  \item ``Novelty detection''
  \item Specific view on \gls{svm}s
\end{itemize}

--- Segmentation ---

``Time series segmentation for context recognition in mobile devices'' \cite{himberg2001time}. 158, 2001 \\

``An Adaptive Approach for Online Segmentation of Multi-Dimensional Mobile Data'' \cite{guo2012adaptive}. 4, 2012 \\

``Joint segmentation of multivariate time series with hidden process regression for human activity recognition'' \cite{chamroukhi2013joint}. 2013. See \ref{sec:appendix-C-joint-segmentation}. \\

``Segmentation and Recognition of Motion Streams by Similarity Search'' \cite{li2007segmentation}. 29, 2007 \\

``Novel Online Methods for Time Series Segmentation'' \cite{liu2008novel}. 22, 2008 \\

``An online algorithm for segmenting time series'' \cite{keogh2001online}. See \ref{sec:appendix-C-online-keogh}. 538, 2001 \\
``Segmenting time series: A survey and novel approach'' \cite{keogh2004segmenting}. 242, 2004 \\

``Distributed Segmentation and Classification of Human Actions Using a Wearable Motion Sensor Network'' \cite{yang2008distributed}. 44, 2008 \\

``An Automatic Segmentation Technique in Body Sensor Networks based on Signal Energy'' \cite{guenterberg2009automatic}. 14, 2009 \\

``Online Segmentation of Time Series Based on Polynomial Least-Squares Approximations'' \cite{fuchs2010online}. 24, 2010 \\

``Aligned Cluster Analysis for Temporal Segmentation of Human Motion'' \cite{zhou2008aligned}. 63, 2008 \\

--- Change detection ---

``A unifying framework for detecting outliers and change points from time series'' \cite{takeuchi2006unifying}. 87, 2006 \\

``Bayesian online changepoint detection'' \cite{adams2007bayesian}. 85, 2007 \\

``An adaptive cusum-based test for signal change detection'' \cite{alippi2006adaptive}. 18, 2006 \\

``An efficient algorithm for estimating a change-point'' \cite{cheng2009efficient}. 5, 2009 \\

``The marginalized likelihood ratio test for detecting abrupt changes'' \cite{gustafsson1996marginalized}. 80, 1996 \\

``The MOSUM of squares test for monitoring variance changes'' \cite{hsu2007mosum}. 4, 2007 \\

``Use of cumulative sums of squares for retrospective detection of changes of variance'' \cite{inclan1994use}. 643, 1994 \\

``Change-point detection in time-series data by direct density-ratio estimation'' \cite{kawahara2009change}. 52, 2009 \\
``Sequential change-point detection based on direct density-ratio estimation'' \cite{kawahara2012sequential}. 22, 2012 \\
``Change-point detection in time-series data by relative density-ratio estimation'' \cite{liu2013change}. 11, 2013 \\

``Change point detection in time series data using support vectors'' \cite{camci2010change}. 3, 2010 \\

--- Novelty detection ---

``Online novelty detection on temporal sequences'' \cite{ma2003online}. 146, 2003 \\
``Time-series novelty detection using one-class support vector machines'' \cite{ma2003time}. 78, 2003 \\

``Novelty detection: a review—part 1: statistical approaches'' \cite{markou2003novelty}. 697, 2003 \\

``Support Vector Method for Novelty Detection'' \cite{scholkopf1999support}. 337, 1999 \\

--- SVMs ---

``Change point detection in time series data using support vectors'' \cite{camci2010change}. 3, 2010 \\

``Time-series novelty detection using one-class support vector machines'' \cite{ma2003time}. 78, 2003 \\

``Support Vector Method for Novelty Detection'' \cite{scholkopf1999support}. 337, 1999 \\

``Support vector domain description'' \cite{tax1999support}. 907, 1999
``Support vector data description applied to machine vibration analysis'' \cite{tax1999supportdata}. 76, 1999 \\

``Human Activity Recognition on Smartphones using a Multiclass Hardware-Friendly Support Vector Machine'' \cite{anguitahuman}. 13, 2012 \\

``A geometric approach to support vector machine (SVM) classification'' \cite{mavroforakis2006geometric}. 136, 2006 \\

``Support vector machines in remote sensing: A review'' \cite{mountrakis2011support}. 150, 2011 \\

``Predicting time series with support vector machines'' \cite{muller1997predicting}. 693, 1997 \\

### 2013-10-08 Chapters/Chapter2/temporal_segmentation.tex
Literature review, temporal segmentation

@@
 
+This section gives an overview of the literature on temporal segmentation in the context of human activity recognition.
+It takes a look on different implementations and methodologies.
+A wide range of terms and subtle differences are used in the field, such as `segmentation', `change detection', `novelty detection' and `outlier detection'.
+These will be the categorical terms for which we discuss the literature.
+Finally we will discuss other applications of \gls{svm}s in the context of these terms.
+
+\subsection{Segmentation}\label{subsec:segmentation}
+*** This subsection is mainly from previous draft version ***
+
+*** TODO: Create compact notation, one sentence per paper max ***
+
+Segmentation methods can roughly be categorized in three methods in the way the data is processed, as discussed by Avci \etal \cite{avci2010activity}:
+\begin{itemize}
+  \item \textbf{Top-down} methods iteratively divide the signal in segments by splitting at the best location.
+  The algorithm starts with two segments and completes when a certain condition is met, such as when an error value or number of segments $k$ is reached.
+  These methods process the data points recursively, which results in a complexity of $O(n^2k)$.
+  \item \textbf{Bottom-up} methods are the natural complement to top-down methods.
+  They start with creating $n/2$ segments and iteratively join adjacent segments while the value of a cost function for that operation is below a certain value.
+  Given the average segment length $L$, the complexity of this method is $O(nL)$.
+  \item \textbf{Sliding-window} methods are simple and intuitive for online segmenting purposes.
+  It starts with a small initial subsequence of the time series.
+  New data points are joined in the segment until the fit-error is above a threshold.
+  Since the data is only processed very locally, these methods can yield in poor results \cite{keogh2001online}.
+  The complexity is equal to the bottom-up approach, $O(nL)$, where $L$ is the average segment length.
+  \item \textbf{\acrlong{swab}}, as introduced by Keogh \etal \cite{keogh2001online}, joins the ability of the sliding window mechanism to process time series online and the bottom-up approach the create superior segments in terms of fit-error.
+  The algorithm processes the data in two stages.
+  The first stage is to join new data points in the current segment created by a sliding window, and pass this to a buffer with space for a few segments.
+  The buffer then processes the data Bottom-up and returns the first (left-most) segment as final segment.
+  Because this second stage retains some (semi-)global view of the data, the results are comparative with normal Bottom-up.
+  It is stated by Keogh \etal that the complexity of \gls{swab} is a small constant factor worse than that of regular Bottom-up.
+\end{itemize}
+It is clear that for the application of this research sliding-window and preferably \gls{swab}-based algorithms should be considered.
+
+The \gls{swab} method proposed by Keogh \etal \cite{keogh2001online} is dependent on an user setting, providing the maximum error when performing both stages.
+Each segment is approximated by using piecewise linear representation (PLR), an often used method.
+The user provided error threshold controls the granularity and number of segments.
+Other methods have been proposed, such as an adaptive threshold based on the signal energy by Guenterberg \etal \cite{guenterberg2009automatic}, the adaptive \gls{cusum}-based test by Alippi \etal \cite{alippi2006adaptive} and the \gls{mosum} by Hsu \cite{hsu2007mosum} in order to eliminate this user-dependency.
+The latter of these methods is able to process the accelerometer values directly, although better results are obtained when features of the signal are processed, as done in the former method.
+Here the signal energy, mean and standard deviation are used to segment activities and by adding all the axial time series together, the Signal-To-Noise ration is increased, resulting in a robuster method.
+
+The method of Guenterberg \etal extracts features from the raw sensor signal to base the segmentation on other properties than the pure values.
+The method of Bernecker \etal \cite{bernecker2012activity} uses other statistical properties, namely autocorrelation, to distinguish periodic from non-periodic segments.
+Using the \gls{swab} method the self-similarity of a one-dimensional signal is obtained.
+The authors claim that only a slight modification is needed to perform the method on multi-dimensional data.
+After the segmentation phase, the method of Bernecker \etal extracts other statistical features which are used in the classification phase.
+
+The proposal of Guo \etal \cite{guo2012adaptive} dynamically determines which features should be used for the segmentation and simultaneously determines the best model to fit the segment.
+For each of the three dimensions features such as the mean, variance, covariance, correlation, energy and entropy are calculated.
+By extending the \gls{swab} method, for every frame a feature set is selected, using an enhanced version of \gls{pca}.
+The research also considered the (Stepwise) Feasable Space Window as introduced by \cite{liu2008novel}, but since it results in a higher error rate than \gls{swab} the latter was chosen to extend.
+Whereas the before mentioned algorithms use a linear representation, this methods considers linear, quadratic and cubical representations for each segment.
+This differs from other methods where the model is fixed for the whole time series, such as \cite{fuchs2010online}, which is stated to perform inferior on non-stationary time series such as daily life.
+
+The time series data from a sensor can be considered as a stochastic process *** is this a correct term? ***.
+Probabilistic models can be constructed on that signal, yielding in probabilistic and Bayesian based segmentation methods.
+The \gls{cusum}-methods takes a statistical approach and relies on the log-likelihood ratio \cite{gustafsson1996marginalized} to measure the difference between two distributions.
+To calculate the ratio, the probability density functions need to be calculated.
+The method of Kawahara \etal \cite{kawahara2009change} proposes to estimate the ratio of probability densities (known as the \emph{importance}), based on the log likelihood of test samples, directly, without explicit estimation of the densities.
+The method by Liu \etal \cite{liu2013change} uses a comparable dissimilarity measure using the \gls{kliep} algorithm.
+They claim this results in a robuster approach for real-world scenarios.
+Although this is a model-based method, no strong assumptions (parameter settings) are made on the models.
+
+The method of Adams and MacKay \cite{adams2007bayesian} builds a probabilistic model on the segment run length, given the observed data so far.
+Instead of modeling the values of the data points as a probabilistic distribution, the length of segments as a function if time is modeled by calculating the posterior probability.
+It uses a prior estimate for the run length and a predictive distribution for newly-observed data, given the data since the last change point.
+This method contrasts with the approach of Guralnik and Srivastava \cite{guralnik1999event} in which change points are detected by a change in the (parameters of an) underlying, observed, model.
+For each new data point, the likelihoods of being a change point and part of the current segment are calculated, without a prior model (and thus is a non-Bayesian approach).
+It is observed that when no change point is detected for a long period of time, the computational complexity increases significantly.
+
+Another application of \gls{pca} is to characterize the data by determining the dimensionality of a sequence of data points.
+The proposed method of Berbi\v{c} \etal \cite{barbivc2004segmenting} determines the number of dimensions (features) needed to approximate a sequence within a specified error.
+With the observation that more dimensions are needed to keep the error below the threshold when transitions between actions occur, cut-points can be located and segments will be created.
+The superior extension of their approach uses a Probabilistic \gls{pca} algorithm to incorporate the dimensions outside the selected set as noise.
+
+In the method by \cite{himberg2001time} a cost function is defined over segments of data which is to be minimized.
+The cost functions thereby searches for internally homogeneous segments of data, reflecting states in which the devices and the user are.
+The cost function can be any arbitrary function and in the implementation the sum of variances over the segments is used.
+Both in a local and global iterative replacement procedure (as an alternative for the computationally hard dynamic programming algorithm) the best breakpoint locations $c_i$ for a pre-defined number of segments $1 \leq i \leq k$ are optimized.
+
+Many methods obtain an implicit segmentation as a result of classification over a sliding window *** add refs ***.
+The method of \cite{yang2008distributed} explicitly performs segmentation and classification simultaneously.
+It argues that the classification of a pre-segmented test-sequences becomes straightforward with many classical algorithms to choose from.
+The algorithm matches test examples with the \emph{sparsest} linear representation of mixture subspace models of training examples, searching over different temporal resolutions.
+
+The method of Chamroukhi \etal \cite{chamroukhi2013joint} is based on a \gls{hmm} and logistic regression.
+It assumes a $K$-state hidden process with a (hidden) state sequence, each state providing the parameters (amongst which the order) for a polynomial.
+The order of the model segment is determined by model selecting, often using the \gls{bic} or the similar \gls{aic} \cite{akaike1974new}, as in \cite{he2008activity}.
+
+Field of computer vision: \cite{zhou2008aligned}, \cite{li2007segmentation}.
+
+\subsection{Change detection}\label{subsec:change_detection}
+*** change order of this and temporal segmentation sections? So first change, then segmentation? *** \\
+Whereas the above mentioned researches focus on \emph{segmentation}, many have focused on \emph{change detection}.
+Although these techniques are closely related, there is a subtle difference.
+In the case of \emph{change detection} to goal is to find, possibly unrelated, sudden change points in a signal \cite{takeuchi2006unifying}.
+In contrast, the goal of \emph{temporal segmentation} is to find homogeneous segments of data, which can be the result of multiple detected changes.
+
+The \gls{icss} by Incl\'{a}n and Tiao \cite{inclan1994use} is a statistical method which obtains comparable results (when applied to stock data) compared to \gls{mle} and Bayesian.
+Whereas \gls{cusum} can be applied to search for a change in mean, the \gls{icss} is adapted to find changes in variance.
+It obtains a \emph{likelihood ratio} for testing the hypothesis of one change against no change in the variance.
+Using an iterative approach, all possible change points are considered.
+The proposal of \cite{cheng2009efficient} extends on the \gls{cusum}-based methods to find change points in mean and variance, by creating a more efficient and accurate algorithm.
+
+*** TODO: move CUSUM based techniques to this subsection ***
+
 --- Segmentation ---
 
-``Time series segmentation for context recognition in mobile devices'' \cite{himberg2001time}. 158, 2001 \\
+% ``An online algorithm for segmenting time series'' \cite{keogh2001online}. See \ref{sec:appendix-C-online-keogh}. 538, 2001 \\
+% ``Segmenting time series: A survey and novel approach'' \cite{keogh2004segmenting}. 242, 2004 \\
 
-``An Adaptive Approach for Online Segmentation of Multi-Dimensional Mobile Data'' \cite{guo2012adaptive}. 4, 2012 \\
+% ``Time series segmentation for context recognition in mobile devices'' \cite{himberg2001time}. 158, 2001 \\
 
-``Joint segmentation of multivariate time series with hidden process regression for human activity recognition'' \cite{chamroukhi2013joint}. 2013. See \ref{sec:appendix-C-joint-segmentation}. \\
+% ``An Adaptive Approach for Online Segmentation of Multi-Dimensional Mobile Data'' \cite{guo2012adaptive}. 4, 2012 \\
+
+% ``Joint segmentation of multivariate time series with hidden process regression for human activity recognition'' \cite{chamroukhi2013joint}. 0, 2013. See \ref{sec:appendix-C-joint-segmentation}. \\
 
@@
 
-``Novel Online Methods for Time Series Segmentation'' \cite{liu2008novel}. 22, 2008 \\
+% ``Novel Online Methods for Time Series Segmentation'' \cite{liu2008novel}. 22, 2008 \\
 
-``An online algorithm for segmenting time series'' \cite{keogh2001online}. See \ref{sec:appendix-C-online-keogh}. 538, 2001 \\
-``Segmenting time series: A survey and novel approach'' \cite{keogh2004segmenting}. 242, 2004 \\
+% ``Distributed Segmentation and Classification of Human Actions Using a Wearable Motion Sensor Network'' \cite{yang2008distributed}. 44, 2008 \\
 
-``Distributed Segmentation and Classification of Human Actions Using a Wearable Motion Sensor Network'' \cite{yang2008distributed}. 44, 2008 \\
+% ``An Automatic Segmentation Technique in Body Sensor Networks based on Signal Energy'' \cite{guenterberg2009automatic}. 14, 2009 \\
 
-``An Automatic Segmentation Technique in Body Sensor Networks based on Signal Energy'' \cite{guenterberg2009automatic}. 14, 2009 \\
-
-``Online Segmentation of Time Series Based on Polynomial Least-Squares Approximations'' \cite{fuchs2010online}. 24, 2010 \\
+% ``Online Segmentation of Time Series Based on Polynomial Least-Squares Approximations'' \cite{fuchs2010online}. 24, 2010 \\
 
 ``Aligned Cluster Analysis for Temporal Segmentation of Human Motion'' \cite{zhou2008aligned}. 63, 2008 \\
+
+% ``Segmenting motion capture data into distinct behaviors'' \cite{barbivc2004segmenting}. 270, 2004 \\
 
@@
 
-``Bayesian online changepoint detection'' \cite{adams2007bayesian}. 85, 2007 \\
+% ``Bayesian online changepoint detection'' \cite{adams2007bayesian}. 85, 2007 \\
 
-``An adaptive cusum-based test for signal change detection'' \cite{alippi2006adaptive}. 18, 2006 \\
+% ``An adaptive cusum-based test for signal change detection'' \cite{alippi2006adaptive}. 18, 2006 \\
 
-``An efficient algorithm for estimating a change-point'' \cite{cheng2009efficient}. 5, 2009 \\
+% ``An efficient algorithm for estimating a change-point'' \cite{cheng2009efficient}. 5, 2009 \\
 
-``The marginalized likelihood ratio test for detecting abrupt changes'' \cite{gustafsson1996marginalized}. 80, 1996 \\
+% ``The marginalized likelihood ratio test for detecting abrupt changes'' \cite{gustafsson1996marginalized}. 80, 1996 \\
 
-``The MOSUM of squares test for monitoring variance changes'' \cite{hsu2007mosum}. 4, 2007 \\
+% ``The MOSUM of squares test for monitoring variance changes'' \cite{hsu2007mosum}. 4, 2007 \\
 
-``Use of cumulative sums of squares for retrospective detection of changes of variance'' \cite{inclan1994use}. 643, 1994 \\
+% ``Use of cumulative sums of squares for retrospective detection of changes of variance'' \cite{inclan1994use}. 643, 1994 \\
 
-``Change-point detection in time-series data by direct density-ratio estimation'' \cite{kawahara2009change}. 52, 2009 \\
+% ``Change-point detection in time-series data by direct density-ratio estimation'' \cite{kawahara2009change}. 52, 2009 \\
 ``Sequential change-point detection based on direct density-ratio estimation'' \cite{kawahara2012sequential}. 22, 2012 \\
-``Change-point detection in time-series data by relative density-ratio estimation'' \cite{liu2013change}. 11, 2013 \\
+% ``Change-point detection in time-series data by relative density-ratio estimation'' \cite{liu2013change}. 11, 2013 \\
 
@@
 ``Support Vector Method for Novelty Detection'' \cite{scholkopf1999support}. 337, 1999 \\
+
+--- Outlier detection ---
+``A unifying framework for detecting outliers and change points from time series'' \cite{takeuchi2006unifying}. 87, 2006 \\
+
+``Outliers in statistical data'' \cite{barnett1994outliers} (book). 3745, 1994 \\
+
+``A survey of outlier detection methodologies'' \cite{hodge2004survey}. 791, 2004 \\
+
+``Local outlier detection reconsidered: a generalized view on locality with applications to spatial, video, and network outlier detection'' \cite{schubert2012local}. 2, 2012 \\
 

### 2013-10-08 Chapters/Chapter3.tex
wip
Fix typos, add comments
Add note on paper

@@
 \begin{itemize}
-  \item Explain the connection between change detection and outlier detection
+  \item Content follows blogpost, but more formal
+  \item Explain the connection between change detection and \{outlier detection, density estimation, \etc\}
+  \item Explain change detection in relation with ``M-Estimators''.
   \item Explain why and how SVM can be used for outlier detection, and thus for change detection in time series
@@
 
+*** TODO: ***
+Follow methodology of ``A unifying framework for detecting outliers and change points from time series'' \cite{takeuchi2006unifying}.
+It creates a two-stage process of first searching for outliers, and then using the ``outlier-score'' to find (sudden) change points, by a weighted average of a moving window.
+That eventual score can be thresholded (as in the paper) or processed with something like CUSUM (proposal).
+Looks like my proposed methods, in that is combines outliers and gives a score to change points.
+
 \input{./Chapters/Chapter3/outlier_detection}

### 2013-10-09 Chapters/Chapter4/overview.tex

@@
   \centering
-    \includegraphics{./Figures/graphs/chapter4/method_setup.pdf}
+    \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{./Figures/graphs/chapter4/method_setup.pdf}
   \caption[Method setup]{Schematic representation of the change detection method.}
@@
 \subsection{Change detection method}
-As graphically represented in figure \ref{fig:method_overview}, the change detection method starts by processing the data from sensor, such as the accelerometer, magnetic orientation and rotation meaures \footnote{Something about the origin of the streams; all from the same sensor or different sensors?}.
+As graphically represented in figure \ref{fig:method_overview}, the change detection method starts by processing the data from sensor, such as the accelerometer, magnetic orientation and rotation measures \footnote{Something about the origin of the streams; all from the same sensor or different sensors?}.
 
@@
 The two processes applied are alignment and normalization.
-Due to noisy sampling, not all the timestamps in the data streams are sensored at the same timestamp.
+Due to noisy sampling, not all the timestamps in the data streams are sensed at the same timestamp.
 Since the \gls{svdd} method requires all the data stream at every timestamp and can not handle missing data on one of the timestamps, all the unique timestamps are filtered out.
@@
 Due to the nature of the sensor signals, a normalization step is required in order to set the weight for all the data streams equal.
-The range of the accelerometer signal typically spans $-20$ to $20$, the magnetic field from $-45$ to $45$ and the rotations range is from $-1$ to $1$.
+The range of the accelerometer signal typically spans $-20$ to $20$, the magnetic field from $-60$ to $60$ and the rotations range is from $-1$ to $1$.
 This means that a relative small change in the accelerometer stream could have a much larger impact on the model than the same (absolute) change in the rotation stream, whilst the latter has a larger relative impact.
@@
 Every update the oldest $s$ data points are removed from and $s$ new data points are added to the \gls{svdd} model.
+The model is (partially) reconstructed and new model properties, such as the radius of the hypersphere and the number of outliers, are the result of this step \footnote{Other measures are also possible, for instance the distance from all the outliers to the boundary of the hypersphere}.
+*** MORE ON THE UPDATING STEP ***
+
+This final step of this method, step $5$, is the interpretation of the model properties.
+Many algorithms can be used for this process, all which take a one-dimensional time series as input and determine where change has occured.
+In our setup we used the \gls{rt} and \gls{cusum} methods, to show the modularity of this step.
 
@@
 \subsection{Experiments}
+For the experiments we used a HTC Sensation XE smartphone as recording device.
+The activities were recorded using a free Android application \cite{sensorlogger}.
+This application was chosen for its convenient data format of the sensor recording and its regularity of the sampling interval.
+Table \ref{tab:recorded_metrics} lists the recorded metrics.
+For our expiments we used the data for the accelerometer, magnatief field and rotation.
+
+\begin{center}\begin{table}
+  \begin{tabulary}{\textwidth}{|l|L|c|c|}
+    \hline
+    Metric & Description & Units of measure & Typical range \\
+    \hline \hline
+    Accelerometer & Acceleration force along each axis. & $m/s^2$ & $-20$ -- $20$ \\
+    \hline
+    Gravity & Force of gravity along each axi.s & $m/s^2$ & $-10$ -- $10$\\
+    \hline
+    Gyroscope & & $rad/s$ & $-15$ -- $15$\\
+    \hline
+    Light & Light sensitive sensor at the front of the phone. & & $0$ -- $10000$ \\
+    \hline
+    Linear acceleration & & $m/s^2$ & $-20$ -- $20$ \\
+    \hline
+    Magnetic field & Geomagnetic field strength along each axis. & $\mu T$ & $-60$ -- $60$ \\
+    \hline
+    Orientation & & & $-100$ -- $360$ \\
+    \hline
+    Rotation & & & $-1$ -- $1$\\
+    \hline
+  \end{tabulary}
+  \caption{Measured metrics. The set of axis is always the triple (x, y, z) direction.}
+  \label{tab:recorded_metrics}
+\end{table}\end{center}

### 2013-10-10 Chapters/Chapter2/temporal_segmentation.tex
Remove note
working on outline of proposed method

@@
   \item ``Novelty detection''
-  \item Specific view on \gls{svm}s
+  \item Specific view on \glspl{svm}
 \end{itemize}
@@
 These will be the categorical terms for which we discuss the literature.
-Finally we will discuss other applications of \gls{svm}s in the context of these terms.
+Finally we will discuss other applications of \glspl{svm} in the context of these terms.
 
@@
 
-The time series data from a sensor can be considered as a stochastic process *** is this a correct term? ***.
+The time series data from a sensor can be considered as being drawn from a certain stochastic process.
 Probabilistic models can be constructed on that signal, yielding in probabilistic and Bayesian based segmentation methods.

### 2013-10-10 Chapters/Chapter3/one_class_svm.tex
working on outline of proposed method

@@
 Here every data point $\mathcal{I}$ for which $a_i > 0$ is weighted in the decision function and thus ``supports'' the classification machine: hence the name ``\acrlong{svm}''.
-Since it is shown that under certain circumstances \gls{svm}s show an equality to sparse representations \cite{girosi1998equivalence,smola1998connection}, there will often be relatively few Lagrange multipliers with a non-zero value.
+Since it is shown that under certain circumstances \glspl{svm} show an equality to sparse representations \cite{girosi1998equivalence,smola1998connection}, there will often be relatively few Lagrange multipliers with a non-zero value.
 
@@
 
-We will discuss two implementations of \gls{oc-svm}s.
+We will discuss two implementations of \glspl{oc-svm}.
 The first is the \gls{nu-svm} by Sch\"olkopf \etal \cite{scholkopf1999support} and closely follows the above problem statement regarding density estimation.
@@
 
-\subsection{\acrlong{nu-svm}s}\label{subsec:nu-svm}
+\subsection{\acrlong{nu-svm}}\label{subsec:nu-svm}
 The first of the \gls{oc-svm} methods we will discuss is often referred to as \gls{nu-svm} and introduced by Sch\"olkopf \etal \cite{scholkopf1999support}.
@@
 It returns the value $+1$ in a (possibly small) region capturing most of the data points, and $-1$ elsewhere.
-The method maps the data points from input space $\mathcal{I}$ to a feature space $\mathcal{F}$ (following classical \gls{svm}s).
+The method maps the data points from input space $\mathcal{I}$ to a feature space $\mathcal{F}$ (following classical \glspl{svm}).
 In that space $\mathcal{F}$ it separates the data points with maximal margin from the origin, with a hyperplane.
@@
   \item $\nu$ is an upper bound on the fraction of outliers, \ie training examples regarded as out-of-class.
-  \item $\nu$ is a lower bound on the fraction of \gls{sv}s, \ie training examples with a nonzero Lagrange multiplier $\alpha_i$.
+  \item $\nu$ is a lower bound on the fraction of \glspl{sv}, \ie training examples with a nonzero Lagrange multiplier $\alpha_i$.
 \end{enumerate}

### 2013-10-10 Chapters/Chapter3/outlier_detection.tex
working on outline of proposed method

@@
 The term \emph{robustness} has many interpretations, one of them that it ``signifies insensi­tivity to small deviations from the [prior] assumptions [about the underlying situation]'', according to Huber \cite{huber2009robust}.
-Methods have been proposed to apply robust statistics in the form of M-estimators to \gls{svm}s, such as \cite{choi2009least,chen2004m,suykens1999least}.
+Methods have been proposed to apply robust statistics in the form of M-estimators to \glspl{svm}, such as \cite{choi2009least,chen2004m,suykens1999least}.
 

### 2013-10-10 Chapters/Chapter4/change_indication.tex
working on outline of proposed method

@@
 \section{Change Indication}\label{sec:method_change_indication}
+*** TODO: change section title ***
+*** TODO: ***
+\emph{Follow methodology of ``A unifying framework for detecting outliers and change points from time series'' \cite{takeuchi2006unifying}.
+It creates a two-stage process of first searching for outliers, and then using the ``outlier-score'' to find (sudden) change points, by a weighted average of a moving window.
+That eventual score can be thresholded (as in the paper) or processed with something like CUSUM (proposal).
+Looks like my proposed methods, in that is combines outliers and gives a score to change points.}
+
+The proposed method of this thesis follows the unifying framework as introduced by Takeuchi and Yamanishi \cite{takeuchi2006unifying} and an similar implementation by Camci \cite{camci2010change} with \glspl{svm}.
+The unifying framework combines the detection of outliers with change points and divides it in two stages.
+The first stage determines the outliers in a time series by giving a score based on the deviation from a learned model, and thereby creates a new time series.
+The second stage runs on that new created time series and calculates a average over a window of the outlier scores.
+The problem of change detection is then reduced to outlier detection over that average-scored time series.
+This method is named \gls{changeFinder} by the authors.
+The implementation by Camci, which uses \glspl{svm} to detect changes is named \acrlong{svcpd}.
+
+The problem statement and formal definition, following Tackeuchi and Yamanishi \cite{takeuchi2006unifying} and Camci \cite{camci2010change} is the following.
+The algorithm needs to find \emph{sudden} changes in the time series data.
+In other words, slowly changing properties in the data are not considered to be changes.
+This is in line with the search of changes in activities, since we are only interested in different activities (which are represented by sudden changes) instead of changes within an activity.
+Considered a time series $x_1 x_1 \dots$, which is drawn from a stochastic process $p$.
+Each $x_t$ (t = 1, 2, \dots) is a $d$-dimensional real valued vector and $p$ a probability density function of the sequence $x_1 x_2 \dots$.
+Assume $p$ can be decomposed in two different \gls{iid} stationary stochastic processes $p^1$ and $p^2$ and are one-dimensional Gaussian density functions.
+For a time point $a$ data points for which $t < a$ are drawn from $p^1 = N(\mu_1, \sigma_1^2)$ and for $t \geq a$ from $p^2 = N(\mu_2, \sigma_a^2)$.
+If $p^1$ and $p^2$ are different, then the time point $t = a$ is a \emph{change point}.
+In \cite{takeuchi2006unifying} the similarity between the stochastic processes are expressed by the \gls{kliep} divergence $D(p^2||p^1)$.
+The problem with this measure is that, as the authors conclude and Camci discusses, it is not able to detect a change by decrease in variance.
+
+Whereas \gls{changeFinder} uses double probability estimation algorithm, our approach follows \gls{svcpd} by constructing a \gls{svm} over a sliding window.
+The \gls{svcpd} algorithm uses the location of new data points in the feature space $\mathcal{F}$ with respect to the hypersphere and the hypersphere's radius $R$ to determine whether the new data point represents a change point.
+
+*** TODO: What is my contribution compared to Camci? Distance of outliers to hypersphere? ***

### 2013-10-10 glossary-acronyms.tex
wip
Work on mathematical formulas
Add paper to Appendix C
WIP: chapter 4, description of experiment methods
Add stuff on Scholkopf one-svm
Complete part on v-SVM
Literature review, temporal segmentation
working on outline of proposed method

@@
 
+\newacronym{aic}{AIC}{Akaike Information Criterion}
+
+\newacronym{bic}{BIC}{Bayesian Information Criterion}
+
 \newacronym{cusum}{CUSUM}{Cumulative Sum}
+
+\newacronym{hmm}{HMM}{Hidden Markov Model}
 
@@
 
+\newacronym{iid}{i.i.d.}{independently and identically distributed}
+
+\newacronym{kliep}{KLIEP}{Kullback-Leibler Importance Estimation Procedure}
+
+\newacronym{mle}{MLE}{Maximum Likelihood Estimates}
+
 \newacronym{mosum}{MOSUM}{Moving Sum}
 
+\newacronym{nu-svm}{$\nu$-SVM}{$\nu$-Support Vector Machine}
+
+\newacronym{occ}{OCC}{One-Class Classifier}
+
+\newacronym{oc-svm}{OC-SVM}{One-Class Support Vector Machine}
+
+\newacronym{rbf}{RBF}{Radial Base Function}
+
+\newacronym{roc}{ROC}{Receiver-Operating Characteristic}
+
+\newacronym{pca}{PCA}{Principal Component Analysis}
+
+\newacronym{rt}{RT}{Ratio-Thresholding}
+
 \newacronym{pdf}{PDF}{Probability Density Function}
+
+\newacronym{qp}{QP}{Quadratic Programming}
+
+\newacronym{sv}{SV}{Support Vector}
+
+\newacronym{svc}{SVC}{Support Vector Classifier}
+
+\newacronym{svcpd}{SVCPD}{Support Vector based Change Detection}
 
@@
 \newacronym{svm}{SVM}{Support Vector Machine}
+
+\newacronym{swab}{SWAB}{Sliding Window And Buttom-up}

### 2013-10-21 Chapters/Chapter2/temporal_segmentation.tex
WIP: literature review SVM

@@
 
-In the method by \cite{himberg2001time} a cost function is defined over segments of data which is to be minimized.
+In the method by Himberg \etal \cite{himberg2001time} a cost function is defined over segments of data which is to be minimized.
 The cost functions thereby searches for internally homogeneous segments of data, reflecting states in which the devices and the user are.
@@
 Many methods obtain an implicit segmentation as a result of classification over a sliding window *** add refs ***.
-The method of \cite{yang2008distributed} explicitly performs segmentation and classification simultaneously.
+The method of Yang \etal \cite{yang2008distributed} explicitly performs segmentation and classification simultaneously.
 It argues that the classification of a pre-segmented test-sequences becomes straightforward with many classical algorithms to choose from.
@@
 
-The \gls{icss} by Incl\'{a}n and Tiao \cite{inclan1994use} is a statistical method which obtains comparable results (when applied to stock data) compared to \gls{mle} and Bayesian.
+The \gls{icss} by Incl\'{a}n and Tiao \cite{inclan1994use} is a statistical method which obtains results (when applied to stock data) comparable to \gls{mle} and Bayesian *** Bayesian What? ***.
 Whereas \gls{cusum} can be applied to search for a change in mean, the \gls{icss} is adapted to find changes in variance.
@@
 
+\subsection{Support Vector Machines}\label{subsec:svm}
+In this section we review methods of change detection based on the applications of \glspl{svm} as model construction method.
+A number of proposals have been made using this method.
+Sch{\"o}lkopf \etal \cite{scholkopf1999support} applies Vapnik's principle, to never solve a problem which is more general than the one that one is actually interested in, to novelty detection.
+In the case of novelty detection, they argue there is no need for density estimation of the distribution.
+Simple algorithms estimate the density by considering how many data points fall in a region of interest.
+The \gls{nu-svm} method instead starts with the number of data points that should be whithin the region and estimates a region with that desired property.
+It builds on the method of Vapnik and Vladimir \cite{vapnik1963pattern}, which characterizes a set of data by separating it from the origin.
+The \gls{nu-svm} method adds the kernel method, allowing non-linear decision functions, and incorporates `softness' by the $\nu$-parameter.
+Whereas the method in \cite{vapnik1963pattern} focusses on two-class problems, the \gls{nu-svm} method introduces the method the one-class problems.
+
+The method introduced by Ma and Perkins \cite{ma2003time} creates a \emph{projected phase} of a time series data, which is intuitively equal to applying a high-pass filter to the time series.
+The projected phase of the time series construct vectors, which are then classified by the \gls{nu-svm} method.
+The method is applied to a simple synthetic sinusoidal signal with small additional noise and a small segment with large additional noise.
+The algorithm is able to detect that segment, without false alarms.
+
+With the same concepts as \gls{nu-svm}, the \gls{svdd} method by Tax and Duin \cite{tax1999support,tax2004support} uses a separating hypersphere (in contrast with a hyperplane) to characterize the data.
+The data points that lie outside of the created hypersphere are considered to be outliers, which number is a pre-determined fraction of the total number of data points.
+The \gls{svcpd} method of Camci \cite{camci2010change} uses this one-class classification method and applies it to time-series data.
+By using an \gls{oc-svm} the method is able to detect changes in mean variance.
+Especially the detection of variance \emph{decrease} is an improvement over other methods that rely in the Kullback-Leibler divergence, since these are inable to detect decreases in variance \cite{takeuchi2006unifying}.
+
+\textbf{The main goal of this these is to apply the \gls{svcpd} method to accelerometer data.}
+*** Beter verwoorden ***
+
 --- SVMs ---
@@
 ``Time-series novelty detection using one-class support vector machines'' \cite{ma2003time}. 78, 2003 \\
-
-``Support Vector Method for Novelty Detection'' \cite{scholkopf1999support}. 337, 1999 \\
 

### 2013-10-21 Chapters/Chapter4.tex
Rough finish of chapter 2
Writing on method
WIP: literature review SVM

@@
 \end{itemize}
+
+\subsection{Experiment notes}
+\begin{itemize}
+  \item ICSS/CUSUM is goed in het vinden van variance changes. Niet goed in mean veranderingen.
+  \item Problemen met SVM methode: ook na de verandering blijft de threshold nog stijgen;
+  \item geprobeerd direct model-reset te doen na change point, maar dan kan je achteraf niet meer verder analyseren.
+  \item Veranderingen in mean geven ``extra'' change point wanneer alle data weer in het window zit. Zie \ref{fig:jumping_mean_ratios} op de 50-tallen.
+
+\end{itemize}
+
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/notes/jumping_mean_camci.eps}
+  \caption[Jumping mean camci]{Jumping mean camci}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/notes/jumping_mean_camci_thresholds.eps}
+  \caption[Jumping mean camci thresholds]{Jumping mean camci, thresholds}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/notes/jumping_mean_camci_ratios_10.eps}
+  \caption[Jumping mean camci ratios]{Jumping mean camci ratios, last 10}
+  \label{fig:jumping_mean_ratios}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/notes/jumping_variance.eps}
+  \caption[Jumping variance]{Jumping variance}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/notes/jumping_variance_thresholds.eps}
+  \caption[Jumping variance thresholds]{Jumping variance, thresholds}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/notes/jumping_variance_ratios_10.eps}
+  \caption[Jumping variance ratios]{Jumping variance ratios, last 10}
+\end{figure}
+
+\input{./Chapters/Chapter4/change_indication}
+\input{./Chapters/Chapter4/overview}
+\input{./Chapters/Chapter4/algorithms}

### 2013-10-21 Chapters/Chapter6/data_gathering.tex
WIP: literature review SVM

% !TEX root = ../../main.tex
\section{Data Gathering}\label{sec:data_gathering}

%---- Run 1: walk-run-roemer
\begin{figure}
\centering
  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/run-1-walk-run-roemer/data_plot_rot_annotated.eps}
  \caption[R1: rotation]{Run 1: Walk-run-roemer, rotation}
\end{figure}

\begin{figure}
\centering
  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/run-1-walk-run-roemer/data_plot_mag_annotated.eps}
  \caption[R1: mag]{Run 1: Walk-run-roemer, Mag}
\end{figure}

\begin{figure}
\centering
  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/run-1-walk-run-roemer/data_plot_acc_annotated.eps}
  \caption[R1: accelerometer]{Run 1: Walk-run-roemer, accelerometer}
\end{figure}

%---- Run 2: walk-run-jos
\begin{figure}
\centering
  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/run-2-walk-run-jos/data_plot_rot_annotated.eps}
  \caption[R2: rotation]{Run 2: Walk-run-jos, rotation}
\end{figure}

\begin{figure}
\centering
  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/run-2-walk-run-jos/data_plot_mag_annotated.eps}
  \caption[R2: mag]{Run 2: Walk-run-jos, Mag}
\end{figure}

\begin{figure}
\centering
  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/run-2-walk-run-jos/data_plot_acc_annotated.eps}
  \caption[R2: accelerometer]{Run 2: Walk-run-jos, accelerometer}
\end{figure}

### 2013-10-22 Chapters/Chapter2/svm.tex
WIP

--- 2013-06-04 Chapters/Chapter2/svm.tex
@@
 % !TEX root = ../../main.tex
-\section{Change-detection by Support Vector Machines}\label{svm}
+\section{Support Vector Machines in Change Detection}\label{sec:svm_in_change_detection}
 
-Introduced by Cortes and Vapnik \cite{vapnik1998statistical, vapnik1999nature}, Support Vector Machines offer a way to segment, and classify, linear separable data.
-This technique can also be applied to estimate density functions of given time series \cite{weston1999support}.
-When combined with a mapping function, which maps the data from the input space $I$ to a higher dimension feature space $F$, the input data can be non-linear separable.
-The linear hyperplane, which segments the data in the feature space $F$, yields to a non-linear segmentation in the lower-dimensional input space $I$.
-Instead of explicitly mapping the input data to the higher dimensional space, a kernel function can be used.
-This kernel function can calculate values of the feature space directly, without the need to first map the input values to this space.
-This process is referred to as the kernel trick.
+Many proposals in the field of remote sensing and \gls{har} make use of \glspl{svm} as a model construction method.
+An elaborate overview of applications is given in \cite{mountrakis2011support}.
+In this section we review methods of change detection based on the applications of \glspl{svm} as model construction method.
+A number of proposals have been made using this method.
+Sch{\"o}lkopf \etal \cite{scholkopf1999support} applies Vapnik's principle, to never solve a problem which is more general than the one that one is actually interested in, to novelty detection.
+In the case of novelty detection, they argue there is no need for density estimation of the distribution.
+Simple algorithms estimate the density by considering how many data points fall in a region of interest.
+The \gls{nu-svm} method instead starts with the number of data points that should be within the region and estimates a region with that desired property.
+It builds on the method of Vapnik and Vladimir \cite{vapnik1963pattern}, which characterizes a set of data by separating it from the origin.
+The \gls{nu-svm} method adds the kernel method, allowing non-linear decision functions, and incorporates `softness' by the $\nu$-parameter.
+Whereas the method in \cite{vapnik1963pattern} focuses on two-class problems, the \gls{nu-svm} method introduces the method the one-class problems.
 
-The proposed method of Camci \cite{camci2010change} uses a one-class support vector machine to segment time series data.
-One-class SVMs are used to describe the current data under consideration, by assuming all data points are from the same class \cite{tax2001one}.
-To cope with possible errors or outliers a soft-margin is applied \cite{cortes1995support}.
-The class is described by a spherical boundary around the data with center $c$ and radius $r$, such that the volume is minimized.
-New data points are consecutively mapped from the input space to the feature space.
-The retrieved (high dimension) data point can be in- or outside the earlier constructed hyper-sphere, thereby giving information about a possible change point.
-Following the definition of Camci \cite{camci2010change}, the class description is obtained by minimizing $r^2$:
-\begin{equation}
-  \mathrm{Min}\ r^2
-\end{equation}
-\begin{equation}
-  \mathrm{Subject\ to} : \|\mathbf{x}_i - \mathbf{c}\|^2 \le r^2\ \forall i,\ \mathbf{x}_i : i \mathrm{th\ data\ point}
-\end{equation}
+The method introduced by Ma and Perkins \cite{ma2003time} creates a \emph{projected phase} of a time series data, which is intuitively equal to applying a high-pass filter to the time series.
+The projected phase of the time series combines a history or data points to a vector, which are then classified by the \gls{nu-svm} method.
+The method is applied to a simple synthetic sinusoidal signal with small additional noise and a small segment with large additional noise.
+The algorithm is able to detect that segment, without false alarms.
 
-To be able to handle outliers in the input data, a penalty cost function $\varepsilon_i$ for each outlier can be added.
+The algorithms of \glspl{svm} has been applied to the problem of \gls{har}, as by Anguita \etal \cite{anguitahuman}.
+In that research a multiclass classification problem is solved using \glspl{svm} and the One-Vs-All method.
+The method exploits fixed-point arithmetic to create a hardware-friendly algorithm, which can be executed on a smartphone.\footnote{Smartphones have limited resources and thus require energy-efficient algorithms.
+In \cite{anguita2007hardware} a Hardware-Friendly \gls{svm} is introduced.
+It uses less memory, processor time and power consumption, with a loss of precision.}
 
-Using this one-class SVM formulation, differences between two (consecutive) windows of data points with size $w$ can be obtained.
-The first window is used as the input set, $h_1$ and the second as the test set $h_t$.
-For the first window a one-class SVM is constructed, yielding in a representation by $c_1$ and $r_1$.
-When the data points of the second window belong to the same class, the representation of that one-class SVM would equal the first:
-\begin{equation}
-\begin{aligned}
-  c_1 = c_2, & &  r_1 = r_2
-\end{aligned}
-\end{equation}
+With the same concepts as \gls{nu-svm}, the \gls{svdd} method by Tax and Duin \cite{tax1999support,tax2004support} uses a separating hypersphere (in contrast with a hyperplane) to characterize the data.
+The data points that lie outside of the created hypersphere are considered to be outliers, which number is a pre-determined fraction of the total number of data points.
 
-In case the second window of data points does not belong to the same class, i.e.
-the probability density function that describes the data differs from the first, the describing values of the second window will differ from the first.
-A difference in the SVM center $c$ or radius $r$ represent a change in the mean and variance, respectively.
-The amount of difference can be expressed by a dissimilarity measure over the representations.
-When the dissimilarity between the two windows exceeds some predefined threshold $th$, there exists a change point between the windows.
+The method by Yin and Yang \cite{yin2008sensor} uses the \gls{svdd} method in the first phase of a two-phase algorithm to filter commonly available normal activities.
+The \gls{svcpd} method of Camci \cite{camci2010change} uses \gls{svdd} applies it to time-series data.
+By using an \gls{oc-svm} the method is able to detect changes in mean variance.
+Especially the detection of variance \emph{decrease} is an improvement over other methods that rely in the Kullback-Leibler divergence, since these are unable to detect decreases in variance \cite{takeuchi2006unifying}.
+This main research subject of this thesis is to apply the method of Camci \cite{camci2010change} to sensor data (such as accelerometer time series) obtained by on-body smartphones.
 
-*** Give visual explanation with circles and in- and out-side new data points ***
+The inner workings of \glspl{oc-svm} are discussed in detail in Section \ref{sec:one_class_svm}.
+
+% \section{Change-detection by Support Vector Machines}\label{svm}
+% *** REMOVE THIS SECTION? ***
+
+% Introduced by Cortes and Vapnik \cite{vapnik1998statistical, vapnik1999nature}, Support Vector Machines offer a way to segment, and classify, linear separable data.
+% This technique can also be applied to estimate density functions of given time series \cite{weston1999support}.
+% When combined with a mapping function, which maps the data from the input space $I$ to a higher dimension feature space $F$, the input data can be non-linear separable.
+% A separating linear hyperplane that segments the data in the feature space $F$, yields a non-linear segmentation in the lower-dimensional input space $I$.
+% Instead of explicitly mapping the input data to the higher dimensional space, a kernel function can be used.
+% This kernel function can calculate values of the feature space directly, without the need to first map the input values to this space.
+% This process is referred to as the kernel trick.
+
+% The proposed method of Camci \cite{camci2010change} uses a one-class support vector machine to segment time series data.
+% One-class SVMs are used to describe the current data under consideration, by assuming all data points are from the same class \cite{tax2001one}.
+% To cope with possible errors or outliers a soft-margin is applied \cite{cortes1995support}.
+% The class is described by a spherical boundary around the data with center $c$ and radius $r$, such that the volume is minimized.
+% New data points are consecutively mapped from the input space to the feature space.
+% The retrieved (high dimension) data point can be in- or outside the earlier constructed hyper-sphere, thereby giving information about a possible change point.
+% *** NOTE: using the in/out position of a new datapoint is different from the approach in this thesis, by only using the model properties. ***
+% Following the definition of Camci \cite{camci2010change}, the class description is obtained by minimizing $r^2$:
+% \begin{equation}
+%   \mathrm{Min}\ r^2
+% \end{equation}
+% \begin{equation}
+%   \mathrm{Subject\ to} : \|\mathbf{x}_i - \mathbf{c}\|^2 \le r^2\ \forall i,\ \mathbf{x}_i : i \mathrm{th\ data\ point}
+% \end{equation}
+
+% To be able to handle outliers in the input data, a penalty cost function $\varepsilon_i$ for each outlier can be added.
+
+% Using this one-class SVM formulation, differences between two (consecutive) windows of data points with size $w$ can be obtained.
+% The first window is used as the input set, $h_1$ and the second as the test set $h_t$.
+% For the first window a one-class SVM is constructed, yielding in a representation by $c_1$ and $r_1$.
+% When the data points of the second window belong to the same class, the representation of that one-class SVM would equal the first:
+% \begin{equation}
+% \begin{aligned}
+%   c_1 = c_2, & &  r_1 = r_2
+% \end{aligned}
+% \end{equation}
+
+% In case the second window of data points does not belong to the same class, \ie the probability density function that describes the data differs from the first, the describing values of the second window will differ from the first.
+% A difference in the SVM center $c$ or radius $r$ represent a change in the mean and variance, respectively.
+% The amount of difference can be expressed by a dissimilarity measure over the representations.
+% When the dissimilarity between the two windows exceeds some predefined threshold $th$, there exists a change point between the windows.
+
+% *** Give visual explanation with circles and in- and out-side new data points ***
+
+% *** Make clear what approach we use (model properties) and create link/bridge to next Chapter (3), which discusses SVM in detail. ***

### 2013-10-22 Chapters/Chapter2/temporal_segmentation.tex
WIP

--- 2013-10-10 Chapters/Chapter2/temporal_segmentation.tex
@@
 
-This section gives an overview of the literature on temporal segmentation in the context of human activity recognition.
+This section gives an overview of the literature on temporal segmentation in the context of \gls{har}.
 It takes a look on different implementations and methodologies.
@@
 Field of computer vision: \cite{zhou2008aligned}, \cite{li2007segmentation}.
-
-\subsection{Change detection}\label{subsec:change_detection}
-*** change order of this and temporal segmentation sections? So first change, then segmentation? *** \\
-Whereas the above mentioned researches focus on \emph{segmentation}, many have focused on \emph{change detection}.
-Although these techniques are closely related, there is a subtle difference.
-In the case of \emph{change detection} to goal is to find, possibly unrelated, sudden change points in a signal \cite{takeuchi2006unifying}.
-In contrast, the goal of \emph{temporal segmentation} is to find homogeneous segments of data, which can be the result of multiple detected changes.
-
-The \gls{icss} by Incl\'{a}n and Tiao \cite{inclan1994use} is a statistical method which obtains comparable results (when applied to stock data) compared to \gls{mle} and Bayesian.
-Whereas \gls{cusum} can be applied to search for a change in mean, the \gls{icss} is adapted to find changes in variance.
-It obtains a \emph{likelihood ratio} for testing the hypothesis of one change against no change in the variance.
-Using an iterative approach, all possible change points are considered.
-The proposal of \cite{cheng2009efficient} extends on the \gls{cusum}-based methods to find change points in mean and variance, by creating a more efficient and accurate algorithm.
-
-*** TODO: move CUSUM based techniques to this subsection ***
 
@@
 
+\subsection{Change detection}\label{subsec:change_detection}
+*** change order of this and temporal segmentation sections? So first change, then segmentation? *** \\
+Whereas the above mentioned researches focus on \emph{segmentation}, many have focused on \emph{change detection}.
+Although these techniques are closely related, there is a subtle difference.
+In the case of \emph{change detection} to goal is to find, possibly unrelated, sudden change points in a signal \cite{takeuchi2006unifying}.
+In contrast, the goal of \emph{temporal segmentation} is to find homogeneous segments of data, which can be the result of multiple detected changes.
+
+The \gls{icss} by Incl\'{a}n and Tiao \cite{inclan1994use} is a statistical method which obtains results (when applied to stock data) comparable to \gls{mle} and Bayesian *** Bayesian What? ***.
+Whereas \gls{cusum} can be applied to search for a change in mean, the \gls{icss} is adapted to find changes in variance.
+It obtains a \emph{likelihood ratio} for testing the hypothesis of one change against no change in the variance.
+Using an iterative approach, all possible change points are considered.
+The proposal of \cite{cheng2009efficient} extends on the \gls{cusum}-based methods to find change points in mean and variance, by creating a more efficient and accurate algorithm.
+
+*** TODO: move CUSUM based techniques to this subsection ***
+
 --- Change detection ---
@@
 
+\subsection{Novelty detection}\label{subsec:novelty_detection}
+
 --- Novelty detection ---
@@
 ``Support Vector Method for Novelty Detection'' \cite{scholkopf1999support}. 337, 1999 \\
+
+\subsection{Outlier detection}\label{subsec:outlier_detection}
 
@@
 
---- SVMs ---
+% --- SVMs ---
 
-``Change point detection in time series data using support vectors'' \cite{camci2010change}. 3, 2010 \\
+% ``Change point detection in time series data using support vectors'' \cite{camci2010change}. 3, 2010 \\
 
-``Time-series novelty detection using one-class support vector machines'' \cite{ma2003time}. 78, 2003 \\
+% ``Time-series novelty detection using one-class support vector machines'' \cite{ma2003time}. 78, 2003 \\
 
-``Support Vector Method for Novelty Detection'' \cite{scholkopf1999support}. 337, 1999 \\
+% ``Support vector domain description'' \cite{tax1999support}. 907, 1999
+% ``Support vector data description applied to machine vibration analysis'' \cite{tax1999supportdata}. 76, 1999 \\
 
-``Support vector domain description'' \cite{tax1999support}. 907, 1999
-``Support vector data description applied to machine vibration analysis'' \cite{tax1999supportdata}. 76, 1999 \\
+% ``Human Activity Recognition on Smartphones using a Multiclass Hardware-Friendly Support Vector Machine'' \cite{anguitahuman}. 13, 2012 \\
 
-``Human Activity Recognition on Smartphones using a Multiclass Hardware-Friendly Support Vector Machine'' \cite{anguitahuman}. 13, 2012 \\
+% ``Support vector machines in remote sensing: A review'' \cite{mountrakis2011support}. 150, 2011 \\
 
-``A geometric approach to support vector machine (SVM) classification'' \cite{mavroforakis2006geometric}. 136, 2006 \\
-
-``Support vector machines in remote sensing: A review'' \cite{mountrakis2011support}. 150, 2011 \\
-
-``Predicting time series with support vector machines'' \cite{muller1997predicting}. 693, 1997 \\
+% ``Sensor-based abnormal human-activity detection'' \cite{yin2008sensor}. 74, 2008 (builds one-class SVM of all normal traces) \\

### 2013-10-22 Chapters/Chapter3/one_class_classification.tex
WIP

% !TEX root = ../../main.tex
\section{One-Class Classification}\label{sec:one_class_classification}

*** TODO: bridge that explains why \gls{occ} is interesting ***

\begin{figure}
  \centering
    \includegraphics[width=0.5\textwidth,keepaspectratio]{./Figures/chapter3/occ_methods.pdf}
  \caption[\gls{occ} methods]{Overview of \gls{occ} methods categorized in Density Estimators, Reconstruction methods and Boundary methods. This categorization follows the definition of Tax in \cite{tax2001one}.}
  \label{fig:occ-methods}
\end{figure}

\subsection{Problem formulation}\label{subsec:occ-problem-formulation}
*** Introduce problem of \gls{occ}. Relate to two-class classification. ***

\subsection{One-Class Classification methods}\label{subsec:occ-methods}
In his thesis \cite{tax2001one}, Tax orders a collection of \gls{occ}-methods into three categories, visually represented in Figure \ref{fig:occ-methods}.
The first category consists of methods that estimate the density of the training data and set a threshold on this density.
Among those are Gaussian models, \glspl{gmm} and Parzen density estimators.
In order to get good generalization results with these methods, the dimensionality of the data and the complexity of the density need to be restricted.
This can cause a large bias on the data.
When a good probablity model is assumed, these methods work very well, since when one threshold is optimized, a minimum volume is automatically found for the given probability density model \cite{tax2001one}.

\subsection{Support Vector Data Description}\label{subsec:occ-svdd}
*** Short introduction of \gls{svdd}. More details in \ref{subsec:oc-svm-svdd}. ***

*** TODO: end with bridge to next chapter, \gls{oc-svm} ***

-- Literature --

``A survey of recent trends in one class classification'' \cite{khan2010survey}. 25, 2010 \\

``One-class classification by combining density and class probability estimation'' \cite{hempstalk2008one}. 51, 2008 \\

``On simple one-class classification methods'' \cite{noumir2012simple}. 2012 \\

``One-class classification'' \cite{tax2001one}. 693, 2001 \\

``The One-Class Classification Approach to Data Description and to Models Applicability Domain'' \cite{baskin2010one}. 19, 2009 \\

``One-class classifier networks for target recognition applications'' \cite{moya1993one}. 70, 1993. Origin of term ``One-Class Classification'' \\

### 2013-10-22 Chapters/Chapter3/one_class_svm.tex
WIP

@@
 
-\subsection{\acrlong{svdd}}\label{subsec:svdd}
+\subsection{\acrlong{svdd}}\label{subsec:oc-svm-svdd}
 The method introduced by Tax and Duin \cite{tax2004support}, known as \acrlong{svdd}, takes as spherical instead of planar approach.
@@
 *** Maybe add section about ``Change indication'' rather than ``detection''? ***
+
+\subsection{More literature}
+*** These papers need to be referenced and included ***
+
+``A geometric approach to support vector machine (SVM) classification'' \cite{mavroforakis2006geometric}. 136, 2006 \\
+
+``Least squares one-class support vector machine'' \cite{choi2009least}. 27, 2009 \\

### 2013-10-22 Chapters/Chapter3/one_class_classification.tex
Working on One-Class Classification texts

@@
 *** TODO: bridge that explains why \gls{occ} is interesting ***
+
+\subsection{Problem formulation}\label{subsec:occ-problem-formulation}
+*** Introduce problem of \gls{occ}. Relate to two-class classification. ***
+
+The problem of \gls{occ} is closely related to the (traditional) two-class classification situation\footnote{Two-class problems are considered as to be the basic problem, since multi-class classification problems can be decomposed into multiple two-class problems \cite{fukunaga1990introduction}.}.
+In the case of traditional classification algorithms, the problem is to classify an unknown object to one of the pre-defined categories.
+This problem is solved by determining a decision boundary in the feature space of the data objects and label the new data object based on the location relative to this boundary.
+
+In case of the \gls{occ} problem, only one class (often refered to as the target class) of data is available in the training phase, or the aquisition of outliers is very expensive.
+
+Figure \ref{fig:two-vs-one-classification} graphically shows the difference between two and one-class classification and the consequences for outlier objects.
+
+\begin{figure}
+  \centering
+    \includegraphics[width=0.5\textwidth,keepaspectratio]{./Figures/chapter3/two-vs-one-classification.pdf}
+  \caption[Difference between two and one-class classification]{This plot shows the difference between two and one-class classification.
+The solid line indicates a possible decision boundary between the $+$ and $*$ example objects.
+The dashed circle indicates the closed boundary around all the data objects.
+In the first type the object $o$ is considered to be member of the $*$-class, whilst in the latter (\gls{occ}) formulation it is an outlier.}
+  \label{fig:two-vs-one-classification}
+\end{figure}
+
+\subsection{One-Class Classification methods}\label{subsec:occ-methods}
+In his thesis \cite{tax2001one}, Tax orders a collection of \gls{occ}-methods into three categories, visually represented in Figure \ref{fig:occ-methods}.
+The first category consists of methods that estimate the density of the training data and set a threshold on this density.
+Among those are Gaussian models, \glspl{gmm} and Parzen density estimators.
+In order to get good generalization results with these methods, the dimensionality of the data and the complexity of the density need to be restricted.
+This can cause a large bias on the data.
+When a good probability model is assumed, these methods work very well, since when one threshold is optimized, a minimum volume is automatically found for the given probability density model \cite{tax2001one}.
+
+Boundary methods are based on Vapnik's principle\footnote{With a limited amount of data available, on should avoid solving a more general problem as an intermediate step to solve the original problem \cite{vapnik1998statistical}.} which imply in this case that estimating the complete data density for a \gls{occ} may be too complex, in case one is only interested in the closed boundary.
+Examples of methods that focus on the boundary (a direct threshold) of the training data distribution are K-centers, Nearest-neighborhood and \gls{svdd}.
+Especially the latter has a strong bias towards minimal volume solutions.
+These type of methods are sensitive to scaling of features, since they rely on a well-defined distance measure.
+The number of objects that is required is smaller then in the case of density methods.
+The boundary method \gls{svdd}, constructed by Tax, will be further discussed in Section \ref{subsec:oc-svm-svdd}.
+
+Reconstruction methods take a different approach.
+Instead of focusing on classification of the data and thus on the discriminating features of data objects, they model the data.
+This results in a compact representation of the target data and for any object a reconstruction error can be defined.
+Outliers are data objects with a high reconstruction error, since they are worse represented by the constructed model.
+Examples of reconstruction methods are K-means, \gls{pca} and some neural network implementations.
 
@@
 \end{figure}
-
-\subsection{Problem formulation}\label{subsec:occ-problem-formulation}
-*** Introduce problem of \gls{occ}. Relate to two-class classification. ***
-
-\subsection{One-Class Classification methods}\label{subsec:occ-methods}
-In his thesis \cite{tax2001one}, Tax orders a collection of \gls{occ}-methods into three categories, visually represented in Figure \ref{fig:occ-methods}.
-The first category consists of methods that estimate the density of the training data and set a threshold on this density.
-Among those are Gaussian models, \glspl{gmm} and Parzen density estimators.
-In order to get good generalization results with these methods, the dimensionality of the data and the complexity of the density need to be restricted.
-This can cause a large bias on the data.
-When a good probablity model is assumed, these methods work very well, since when one threshold is optimized, a minimum volume is automatically found for the given probability density model \cite{tax2001one}.
-
-\subsection{Support Vector Data Description}\label{subsec:occ-svdd}
-*** Short introduction of \gls{svdd}. More details in \ref{subsec:oc-svm-svdd}. ***
 

### 2013-10-23 Chapters/Chapter3/one_class_classification.tex
More on OCC

@@
 In the case of traditional classification algorithms, the problem is to classify an unknown object to one of the pre-defined categories.
+Every object $i$ is represented as a vector $\vectorsym{x}_i = (x_{i,1},\dots,x_{i,d}), x_{i,j} \in \mathbb{R}$ with $d$ real-valued measurements.
+Using this notation, an object $\vectorsym{x}_i$ thus represents one point in a feature space: $\vectorsym{x}_i \in \mathcal{X} = \mathbb{R}^d$.
+The two classes of objects, $\omega_1$ and $\omega_2$, are labeled $-1$ and $+1$ respectively.
+The objects with a label $y_i \in \{-1, +1\}$ attached are in the training set $\mathcal{X}^{tr}$ (note that it can be both positive and negative example objects).
 This problem is solved by determining a decision boundary in the feature space of the data objects and label the new data object based on the location relative to this boundary.
-
-In case of the \gls{occ} problem, only one class (often refered to as the target class) of data is available in the training phase, or the aquisition of outliers is very expensive.
-
-Figure \ref{fig:two-vs-one-classification} graphically shows the difference between two and one-class classification and the consequences for outlier objects.
+This is expressed by the decision function $y = f(\vectorsym{x})$:
+\begin{equation}\label{eq:decision_function_classification}
+  f: \mathbb{R}^d \rightarrow \{-1, +1\}
+\end{equation}
+In case of the \gls{occ} problem, only one class (often referred as the target class, or positive examples) of training data is used to create a decision boundary.
+The goal is to determine whether a new data object belongs to the target class or to the unknown class and thus is an outlier.
+One could argue that this problem is equal to the traditional two-class problem by considering all other classes as negative examples, although there are important differences.
+In pure \gls{occ} problems there are no negative example objects available.
+This could be because the acquisition of these examples is very expensive, or because there are only examples of the `normal' state of the system and the goal is to detect `abnormal' states.
+Since the algorithm's goal is to differentiate between normal and abnormal objects (relative to the training objects), \gls{occ} is often called outlier, novelty or anomaly detection, depending on the origin of the problem to which the algorithm is applied\footnote{The term \acrlong{occ} originates from Moya \etal \cite{moya1993one}.}.
+The difference between two and one-class classification and the consequence for outlier objects is illustrated in Figure \ref{fig:two-vs-one-classification}.
+In the two-class classification problem the object \textbf{o} will be member of the \textbf{-1} class whilst the \gls{occ} problem will label it as an outlier.
+In \cite{tax2001one} a more detailed analysis of the \gls{occ} is given.
 
@@
   \caption[Difference between two and one-class classification]{This plot shows the difference between two and one-class classification.
-The solid line indicates a possible decision boundary between the $+$ and $*$ example objects.
+The solid line indicates a possible decision boundary between the \textbf{+1} and \textbf{-1} example objects.
 The dashed circle indicates the closed boundary around all the data objects.
-In the first type the object $o$ is considered to be member of the $*$-class, whilst in the latter (\gls{occ}) formulation it is an outlier.}
+In the first type the object \textbf{o} is considered to be member of the \textbf{-1}-class, whilst in the latter (\gls{occ}) formulation it is an outlier.}
   \label{fig:two-vs-one-classification}
 \end{figure}
+
+The \gls{occ} algorithms have been applied to a wide range of applications.
+The first is, obviously, outlier detection of objects which do not resemble the bulk of the training data.
+It can be a goal by itself and can be used as a filtering mechanism for other data processing methods.
+Often methods that rely on data-characteristics are sensitive for remote regions in the data set.
+Using \gls{occ} these remote regions can be removed from the data set.
+A second application is for the problem as described above, in which only data from a single target class is available.
+When this problems originates from \eg a monitoring process, the \gls{occ} is able to recognize abnormal system states, without the need to create (or simulate) faulty states beforehand.
+The final possible application given by Tax \cite{tax2001one} is the comparison of two data sets.
+By constructing a \gls{occ}-classifier using a training data set, one can compare that set to a new data set.
+This will result in a similarity-measure.
+It is easy to see here a relation with similiary measures over \glspl{pdf} and the Kullback-Leibler divergence.
+This application will be further discussed in Section \ref{subsec:change-detection-with-occ}.
 
@@
 These type of methods are sensitive to scaling of features, since they rely on a well-defined distance measure.
-The number of objects that is required is smaller then in the case of density methods.
-The boundary method \gls{svdd}, constructed by Tax, will be further discussed in Section \ref{subsec:oc-svm-svdd}.
+The number of objects that is required is smaller than in the case of density methods.
+The boundary method \gls{svdd}, constructed by Tax and which has shown good performance \cite{khan2010survey}, will be further discussed in Section \ref{subsec:oc-svm-svdd}.
 
@@
 
+In \cite{khan2010survey,noumir2012simple} an overview of applications for \gls{occ}-algorithms, and explictly for \gls{svm}-based methods (such as \gls{svdd} and \gls{nu-svm}), is given.
+It shows succesful applications for, amonst others, problems in the field of Handwritten Digit Recognition, Face Recognition Applications, Spam Detection and Anomaly Detection \cite{li2003improving,perdisci2006using}.
+As discussed in Section \ref{sec:outlier_detection}, this can be used for change detection.
+
+\subsection{Change detection with \gls{occ}-methods}\label{subsec:change-detection-with-occ}
+*** Write -short- how \gls{occ}-methods can be used for change detection. Relate to previous section on change detection, \ref{sec:outlier_detection} ***
+
 *** TODO: end with bridge to next chapter, \gls{oc-svm} ***
@@
 
-``A survey of recent trends in one class classification'' \cite{khan2010survey}. 25, 2010 \\
+% ``A survey of recent trends in one class classification'' \cite{khan2010survey}. 25, 2010 \\
 
@@
 
-``On simple one-class classification methods'' \cite{noumir2012simple}. 2012 \\
-
-``One-class classification'' \cite{tax2001one}. 693, 2001 \\
+% ``One-class classification'' \cite{tax2001one}. 693, 2001 \\
 

### 2013-10-23 Chapters/Chapter2/svm.tex
Working on One-Class Classification texts
WIP

@@
 The algorithms of \glspl{svm} has been applied to the problem of \gls{har}, as by Anguita \etal \cite{anguitahuman}.
-In that research a multiclass classification problem is solved using \glspl{svm} and the One-Vs-All method.
+In that research a multi-class classification problem is solved using \glspl{svm} and the One-Vs-All method.
 The method exploits fixed-point arithmetic to create a hardware-friendly algorithm, which can be executed on a smartphone.\footnote{Smartphones have limited resources and thus require energy-efficient algorithms.
@@
 By using an \gls{oc-svm} the method is able to detect changes in mean variance.
-Especially the detection of variance \emph{decrease} is an improvement over other methods that rely in the Kullback-Leibler divergence, since these are unable to detect decreases in variance \cite{takeuchi2006unifying}.
+Especially the detection of variance \emph{decrease} is an improvement over other methods that rely in the \gls{kl} divergence, since these are unable to detect decreases in variance \cite{takeuchi2006unifying}.
 This main research subject of this thesis is to apply the method of Camci \cite{camci2010change} to sensor data (such as accelerometer time series) obtained by on-body smartphones.

### 2013-10-23 Chapters/Chapter3.tex
working on outline of proposed method
WIP

--- 2013-07-03 Chapters/Chapter3.tex
@@
 
-\input{./Chapters/Chapter3/outlier_detection}
+This chapter discusses the concepts and algorithms that will be used as a basis for the proposed method, as discussed in Chapter \ref{Chapter4}.
+The first sections formulates the problem of change detection and relates it to outlier and novelty detection.
+It transforms the traditional problem to change detection for times series data.
+In Section \ref{sec:one_class_classification} the problem of \acrlong{occ} is discussed and shows a number of implementations.
+Two \acrlong{svm}-based methods, \gls{nu-svm} and \gls{svdd}, are further disussed in section \ref{sec:one_class_svm}.
+*** TODO: add \ref{sec:svm_parameters} ***
+For the application to the problem of \acrlong{har}, the characteristics of the collected sensor data is discussed in \ref{sec:sensor_data_characteristics}.
+Finally a theoretic set of quality metrics is discussed in Section \ref{sec:quality_metrics}.
+
+\input{./Chapters/Chapter3/change_detection_time_series}
+\input{./Chapters/Chapter3/one_class_classification}
 \input{./Chapters/Chapter3/one_class_svm}
 \input{./Chapters/Chapter3/svm_parameters}
-\input{./Chapters/Chapter3/data_characteristics}
+\input{./Chapters/Chapter3/sensor_data_characteristics}
 \input{./Chapters/Chapter3/quality_metrics}

### 2013-10-23 Chapters/Chapter3/change_detection_time_series.tex
WIP

% !TEX root = ../../main.tex
\section{Change Detection in Time Series}\label{sec:change_detection_time_series}
*** The section should relate the problem of outlier detection to change detection.\\
It should transform the traditional problem formulation to that of time series data.
Build to one-class classification, which can be used for outlier detection. ***

\subsection{Relation to outlier detection}
The problems of outlier and novelty detection, segmentation, and change detection are closely related.
The terminology depends on the field of application, but there are subtle differences.
The problem of outlier detection is concerned with finding data objects in a data set which have small resemblance with the majority of the data objects.
These objects can be regarded as erroneous measurements.
In the case of novelty detection these objects are considered to be member of a new class of objects.
The unifying framework of Takeuchi and Yamanishi \cite{takeuchi2006unifying}, \gls{changeFinder}, creates a two stage process expressing change detection in terms of outlier detection.
The first stage determines the outliers in a time series by giving a score based on the deviation from a learned model, and thereby creates a new time series.
The second stage runs on that new created time series and calculates a average over a window of the outlier scores.
The problem of change detection is then reduced to outlier detection over that average-scored time series.
The implementation by Camci \cite{camci2010change}, \gls{svcpd}, implements outlier detection with the \gls{svdd} algorithm to detect changes in mean and variance.

\subsection{Problem formulation}\label{subsec:change_detection_problem_formulation}
*** Give the problem of change definition. Follow definition of Camci \cite{camci2010change} ***

*** Note: the following block is copied from \ref{sec:method_change_indication} *** \\
The problem of change point detection can be formulated using different type of models, as discussed in \ref{subsec:segmentation}.
The methods by Takeuchi and Yamanishi \cite{takeuchi2006unifying} and Camci \cite{camci2010change} use the following formulation for change detection, which we will also use for our problem formulation.
The algorithm searches for \emph{sudden} changes in the time series data.
In other words, slowly changing properties in the data are not considered to be changes.
Considered a time series $x_1 x_1 \dots$, which is drawn from a stochastic process $p$.
Each $x_t$ (t = 1, 2, \dots) is a $d$-dimensional real valued vector and $p$ a probability density function of the sequence $x_1 x_2 \dots$.
Assume $p$ can be decomposed in two different \gls{iid} stationary stochastic processes $p^1$ and $p^2$ and are one-dimensional Gaussian density functions.
For a time point $a$ data points for which $t < a$ are drawn from $p^1 = N(\mu_1, \sigma_1^2)$ and for $t \geq a$ from $p^2 = N(\mu_2, \sigma_a^2)$.
If $p^1$ and $p^2$ are different, then the time point $t = a$ is a \emph{change point}.
In \cite{takeuchi2006unifying} the similarity between the stochastic processes are expressed by the \gls{kl} divergence $D(p^2||p^1)$.
The problem with this measure is that, as the authors conclude and Camci discusses, it is not able to detect a change by decrease in variance.

The definition of change point being \emph{sudden} changes in the time series data is in line with the search of changes in activities.
Since we are only interested in different activities (which are represented by sudden changes), slight changes within an activity are not of interest.
Section \ref{sec:sensor_data_characteristics} discusses the representation of activities and changes between them in the data in more detail.

*** TODO: add more formal definition or analysis of change detection ***

-- Literature --

``Online segmentation of time series based on polynomial least-squares approximations'', \cite{fuchs2010online}. 25, 2010 \\

% \subsection{General framework for outlier detection}
% In \cite{schubert2012local} a general framework for outlier detection is proposed.
% The authors compare existing methods and identified the common building block of the algorithms.
% The focus is on unsupervised methods, using information from a local selection of data objects for the detection of outliers.
% The following common algorithmic building blocks are identified:

% \begin{enumerate}
%   \item Context: a ``local'' context of an object $o$ for model building
%   \item Model: the method used for building the model
%   \item Reference: a ``reference'' context of object $o$ for model comparison
%   \item Comparison: the method used for model comparison
%   \item Normalization: a (global) normalization procedure
% \end{enumerate}

% Here the \emph{context} and \emph{reference} are sets of objects used for model building and model comparison, respectively.

% \subsection{Regression, Classification, etc}

% General framework of outlier detection, change detection in context of (simple?) regression, classification.

% \subsection{M-estimators}
% *** Move this section to somewhere in the end; is specification/extension, not base material ***
% To make a method less sensitive to outliers (in the training data) techniques from robust statistics are applied.
% The term \emph{robustness} has many interpretations, one of them that it ``signifies insensi­tivity to small deviations from the [prior] assumptions [about the underlying situation]'', according to Huber \cite{huber2009robust}.
% Methods have been proposed to apply robust statistics in the form of M-estimators to \glspl{svm}, such as \cite{choi2009least,chen2004m,suykens1999least}.

% *** Question: do slack-variables (allowance of outliers) make SVM robust? ***

### 2013-10-23 Chapters/Chapter3/one_class_classification.tex
WIP

@@
 
-*** TODO: bridge that explains why \gls{occ} is interesting ***
+*** TODO: bridge that explains why \gls{occ} is interesting, relating it to outlier/change detection ***
 
@@
 This will result in a similarity-measure.
-It is easy to see here a relation with similiary measures over \glspl{pdf} and the Kullback-Leibler divergence.
-This application will be further discussed in Section \ref{subsec:change-detection-with-occ}.
+It is easy to see here a relation with similiary measures over \glspl{pdf} and the \gls{kl} divergence. \\
+*** Add reference to \ref{subsec:change_detection_problem_formulation} which also mentions \gls{kl} divergence? ***
 
 \subsection{One-Class Classification methods}\label{subsec:occ-methods}
-In his thesis \cite{tax2001one}, Tax orders a collection of \gls{occ}-methods into three categories, visually represented in Figure \ref{fig:occ-methods}.
+The methods and algorithms used for the \gls{occ}-problem can be organized into three categories \cite{tax2001one,noumir2012simple}, visually represented in Figure \ref{fig:occ-methods}.
 The first category consists of methods that estimate the density of the training data and set a threshold on this density.
@@
 Boundary methods are based on Vapnik's principle\footnote{With a limited amount of data available, on should avoid solving a more general problem as an intermediate step to solve the original problem \cite{vapnik1998statistical}.} which imply in this case that estimating the complete data density for a \gls{occ} may be too complex, in case one is only interested in the closed boundary.
-Examples of methods that focus on the boundary (a direct threshold) of the training data distribution are K-centers, Nearest-neighborhood and \gls{svdd}.
+Examples of methods that focus on the boundary (a direct threshold) of the training data distribution are K-centers, Nearest-neighborhood and \gls{svdd} or a combination of those methods \cite{hempstalk2008one}.
 Especially the latter has a strong bias towards minimal volume solutions.
@@
 Outliers are data objects with a high reconstruction error, since they are worse represented by the constructed model.
-Examples of reconstruction methods are K-means, \gls{pca} and some neural network implementations.
+Examples of reconstruction methods are K-means, \gls{pca} and different kind of neural network implementations.
 
@@
 It shows succesful applications for, amonst others, problems in the field of Handwritten Digit Recognition, Face Recognition Applications, Spam Detection and Anomaly Detection \cite{li2003improving,perdisci2006using}.
-As discussed in Section \ref{sec:outlier_detection}, this can be used for change detection.
+As discussed in Section \ref{sec:change_detection_time_series}, this can be used for change detection in time series data.
 
-\subsection{Change detection with \gls{occ}-methods}\label{subsec:change-detection-with-occ}
-*** Write -short- how \gls{occ}-methods can be used for change detection. Relate to previous section on change detection, \ref{sec:outlier_detection} ***
+In this Section we have discussed the problem of \gls{occ} and different kind of implementations.
+An often used implementation is the \gls{svm}-based method \cite{noumir2012simple}, since is shows good performance in comparitive researches.
+*** TODO: add refs for good results of OC-SVM***
+In the following section (\ref{sec:one_class_svm}) two implementations of \gls{oc-svm} will be discussed, the \gls{svdd} method of Tax and Duin \cite{tax2004support} and the \gls{nu-svm}-algoritm by Sch\"olkopf.
 
-*** TODO: end with bridge to next chapter, \gls{oc-svm} ***
-
--- Literature --
+% -- Literature --
 
@@
 
-``One-class classification by combining density and class probability estimation'' \cite{hempstalk2008one}. 51, 2008 \\
+% ``One-class classification by combining density and class probability estimation'' \cite{hempstalk2008one}. 51, 2008 \\
 
@@
 
-``The One-Class Classification Approach to Data Description and to Models Applicability Domain'' \cite{baskin2010one}. 19, 2009 \\
+% ``The One-Class Classification Approach to Data Description and to Models Applicability Domain'' \cite{baskin2010one}. 19, 2009 \\
 
-``One-class classifier networks for target recognition applications'' \cite{moya1993one}. 70, 1993. Origin of term ``One-Class Classification'' \\
+% ``One-class classifier networks for target recognition applications'' \cite{moya1993one}. 70, 1993. Origin of term ``One-Class Classification'' \\

### 2013-10-23 Chapters/Chapter4/change_indication.tex
WIP

@@
 
-The problem statement and formal definition, following Tackeuchi and Yamanishi \cite{takeuchi2006unifying} and Camci \cite{camci2010change} is the following.
+*** TODO: the following paragraph is copied into \ref{subsec:change_detection_problem_formulation}. Remove here? Or shorten? ***\\
+The problem statement and formal definition, following Takeuchi and Yamanishi \cite{takeuchi2006unifying} and Camci \cite{camci2010change} is the following.
 The algorithm needs to find \emph{sudden} changes in the time series data.
@@
 The problem with this measure is that, as the authors conclude and Camci discusses, it is not able to detect a change by decrease in variance.
+\\*** End copied paragraph ***
 

### 2013-10-24 Chapters/Chapter3/one_class_svm.tex
More on OCC
Adding notes from Learning from Data

@@
 The hyperplane that is created determines the \emph{margin} between the classes; the minimal distance from one of the data points to the hyperplane.
-In geometric sense, $w$ is the normal vector to the hyperplane and $\frac{b}{\lVert{w}\rVert}$ determines the offset of the hyperplane to the origin.
+In geometric sense, $w$ is the normal vector indication the direction of the hyperplane and $\frac{b}{\lVert{w}\rVert}$ determines the offset of the hyperplane to the origin.
 Since the distance between the two margins is equal to $\frac{2}{\lVert{w}\rVert}$, the maximum-margin hyperplane is found by minimizing $\lVert{w}\rVert$.
@@
 
-*** TODO: better format of above formula ***
+*** TODO: better format of above formulas. Check with notation and formulas of chapter 9 of \cite{cherkassky2007learning}. ***
 
 This minimization problem can be solved (using \gls{qp}) and transformed to its Lagrange dual formulation.
+\\ *** TODO: Add explanation why to use dual, which is: because the problem formulation scales with the number of samples $n$ and not the dimesionality $d$ of the samples.
+More in Section 9.3 of \cite{cherkassky2007learning}. *** \\
 To do so, the Lagrange multipliers $a_i >= 0$ are introduced and the decision function becomes:
@@
 Since it is shown that under certain circumstances \glspl{svm} show an equality to sparse representations \cite{girosi1998equivalence,smola1998connection}, there will often be relatively few Lagrange multipliers with a non-zero value.
+
+*** TODO: Add more on sparsity and good results, because of sparse representations, even in case of data of high dimensionality. e.g. \cite{cherkassky2007learning}. ***
 
@@
 The second, on which we will base our change detection method, is the \gls{svdd} by Tax and Duin \cite{tax2004support}.
+
+*** Add something about robustness, discusses in section 9.2 from \cite{cherkassky2007learning} ***
 
@@
 
-\subsection{More literature}
+-- Literature --
+
 *** These papers need to be referenced and included ***
@@
 ``Least squares one-class support vector machine'' \cite{choi2009least}. 27, 2009 \\
+
+``On simple one-class classification methods'' \cite{noumir2012simple}. 2012  --> decouples radius and center optimization, gives fast approximations instead of precise results. \\
+
+``Learning from Data'' \cite{cherkassky2007learning}. 2007, 1547 (book).

### 2013-10-24 Appendices/AppendixA-summaries.tex
WIP
Added the TODO command

@@
 The dual problem maximizes this function under positivity constraints and one equality constraint:
-*** TODO: fix equation (now commented) ***
+\TODO{fix equation (now commented)}
 \begin{equation}
@@
 About segmentation with SVMs.
-Will be main material for section \ref{svm} about SVMs.
+Will be main material for section \ref{sec:svm_in_change_detection} about SVMs.
 

### 2013-10-24 Appendices/AppendixB-camci.tex
Added the TODO command

@@
 
-*** \emph{Thus methods of data descriptions should be compared (which are more general) and not only density estimation?} ***
+\TODO{\emph{Thus methods of data descriptions should be compared (which are more general) and not only density estimation?}}
 
@@
 
-*** REVIEW zoeken: change points in time series ***
+\TODO{REVIEW zoeken: change points in time series}
 
@@
 
-*** TODO: make a clear distinction between model types; similarity (distance based) or model based, as in \cite{yin2008sensor} ***
+\TODO{make a clear distinction between model types; similarity (distance based) or model based, as in \cite{yin2008sensor}}
 
@@
 
-*** TODO: read \cite{mavroforakis2006geometric} and \cite{bennett2000duality} on geometry of SVMs ***
+\TODO{read \cite{mavroforakis2006geometric} and \cite{bennett2000duality} on geometry of SVMs}
 
@@
 
-*** Note: in \cite{mavroforakis2006geometric} it is explained why the inner-product between two vectors is a logical choice for the distance/similarity measure. ***
+\TODO{Note: in \cite{mavroforakis2006geometric} it is explained why the inner-product between two vectors is a logical choice for the distance/similarity measure.}
 
@@
 \subsection{Novelty/outlier detection}
-*** Read \cite{hodge2004survey} ``A survey of outlier detection methodologies'' to compare with other methodologies. ***
+\TODO{Read \cite{hodge2004survey} ``A survey of outlier detection methodologies'' to compare with other methodologies.}
 \begin{itemize}

### 2013-10-24 Chapters/Chapter2.tex
Overview of papers for literature review 'Segmentation'
WIP
Added the TODO command

@@
 Many applications require the detection of time points at which the underlying properties of a system change.
-This problem has received a lot of attention in the fields of data mining, etc... *** list and refs ***.
+This problem has received a lot of attention in the fields of data mining, etc... \TODO{list and refs}.
 Often this problem is formulated in a statistical framework, by inspecting the underlying data generating \gls{pdf} of the time series data.
@@
 A change in the \gls{pdf} is represented by a change in the radius of this sphere.
-Section \ref{svm} discusses the \gls{svm}-method.
-*** Put emphasis that this is the source of inspiration for the chosen method? ***
+Section \ref{sec:svm_in_change_detection} discusses the \gls{svm}-methods.
+\TODO{Put emphasis that this is the source of inspiration for the chosen method?}
 
@@
 
-*** Maybe add dimensionality reduction, but for now leave out ***
+\TODO{Maybe add dimensionality reduction, but for now leave out}
 
@@
 %----------------------------------------------------------------------------------------
+% Temporal Segmentation
+\input{./Chapters/Chapter2/temporal_segmentation.tex}
+
+%----------------------------------------------------------------------------------------
 % Support Vector Machines

### 2013-10-24 Chapters/Chapter2/temporal_segmentation.tex
WIP
Added the TODO command

@@
 \subsection{Segmentation}\label{subsec:segmentation}
-*** This subsection is mainly from previous draft version ***
+\TODO{This subsection is mainly from previous draft version}
 
-*** TODO: Create compact notation, one sentence per paper max ***
+\TODO{Create compact notation, one sentence per paper max}
 
@@
 
-Many methods obtain an implicit segmentation as a result of classification over a sliding window *** add refs ***.
+Many methods obtain an implicit segmentation as a result of classification over a sliding window \TODO{add refs}.
 The method of Yang \etal \cite{yang2008distributed} explicitly performs segmentation and classification simultaneously.
@@
 \subsection{Change detection}\label{subsec:change_detection}
-*** change order of this and temporal segmentation sections? So first change, then segmentation? *** \\
+\TODO{change order of this and temporal segmentation sections? So first change, then segmentation?} \\
 Whereas the above mentioned researches focus on \emph{segmentation}, many have focused on \emph{change detection}.
@@
 
-The \gls{icss} by Incl\'{a}n and Tiao \cite{inclan1994use} is a statistical method which obtains results (when applied to stock data) comparable to \gls{mle} and Bayesian *** Bayesian What? ***.
+The \gls{icss} by Incl\'{a}n and Tiao \cite{inclan1994use} is a statistical method which obtains results (when applied to stock data) comparable to \gls{mle} and Bayesian \TODO{Bayesian What?}.
 Whereas \gls{cusum} can be applied to search for a change in mean, the \gls{icss} is adapted to find changes in variance.
@@
 
-*** TODO: move CUSUM based techniques to this subsection ***
+Section \ref{sec:change_detection_time_series} discusses the problem of change detection in time series further, and gives a formal problem definition.
+
+\TODO{move CUSUM based techniques to this subsection}
 

### 2013-10-24 Chapters/Chapter3.tex
Added the TODO command

@@
 Two \acrlong{svm}-based methods, \gls{nu-svm} and \gls{svdd}, are further disussed in section \ref{sec:one_class_svm}.
-*** TODO: add \ref{sec:svm_parameters} ***
+\TODO{add \ref{sec:svm_parameters}}
 For the application to the problem of \acrlong{har}, the characteristics of the collected sensor data is discussed in \ref{sec:sensor_data_characteristics}.

### 2013-10-24 Chapters/Chapter3/change_detection_time_series.tex
Added the TODO command

@@
 \subsection{Problem formulation}\label{subsec:change_detection_problem_formulation}
-*** Give the problem of change definition. Follow definition of Camci \cite{camci2010change} ***
+\TODO{Give the problem of change definition. Follow definition of Camci \cite{camci2010change}}
 
-*** Note: the following block is copied from \ref{sec:method_change_indication} *** \\
+\TODO{Note: the following block is copied from \ref{sec:method_change_indication}} \\
 The problem of change point detection can be formulated using different type of models, as discussed in \ref{subsec:segmentation}.
@@
 
-*** TODO: add more formal definition or analysis of change detection ***
+\TODO{add more formal definition or analysis of change detection}
 
@@
 % \subsection{M-estimators}
-% *** Move this section to somewhere in the end; is specification/extension, not base material ***
+% \TODO{Move this section to somewhere in the end; is specification/extension, not base material}
 % To make a method less sensitive to outliers (in the training data) techniques from robust statistics are applied.
@@
 
-% *** Question: do slack-variables (allowance of outliers) make SVM robust? ***
+% \TODO{Question: do slack-variables (allowance of outliers) make SVM robust?}

### 2013-10-24 Chapters/Chapter3/one_class_classification.tex
Added the TODO command

@@
 
-*** TODO: bridge that explains why \gls{occ} is interesting, relating it to outlier/change detection ***
+\TODO{bridge that explains why \gls{occ} is interesting, relating it to outlier/change detection}
 
 \subsection{Problem formulation}\label{subsec:occ-problem-formulation}
-*** Introduce problem of \gls{occ}. Relate to two-class classification. ***
+\TODO{Introduce problem of \gls{occ}. Relate to two-class classification.}
 
@@
 It is easy to see here a relation with similiary measures over \glspl{pdf} and the \gls{kl} divergence. \\
-*** Add reference to \ref{subsec:change_detection_problem_formulation} which also mentions \gls{kl} divergence? ***
+\TODO{Add reference to \ref{subsec:change_detection_problem_formulation} which also mentions \gls{kl} divergence?}
 
@@
 An often used implementation is the \gls{svm}-based method \cite{noumir2012simple}, since is shows good performance in comparitive researches.
-*** TODO: add refs for good results of OC-SVM***
+\TODO{Add refs for good results of OC-SVM}
 In the following section (\ref{sec:one_class_svm}) two implementations of \gls{oc-svm} will be discussed, the \gls{svdd} method of Tax and Duin \cite{tax2004support} and the \gls{nu-svm}-algoritm by Sch\"olkopf.

### 2013-10-24 Chapters/Chapter3/one_class_svm.tex
Added the TODO command

@@
 
-*** TODO: better format of above formulas. Check with notation and formulas of chapter 9 of \cite{cherkassky2007learning}. ***
+\TODO{better format of above formulas. Check with notation and formulas of chapter 9 of \cite{cherkassky2007learning}.}
 
 This minimization problem can be solved (using \gls{qp}) and transformed to its Lagrange dual formulation.
-\\ *** TODO: Add explanation why to use dual, which is: because the problem formulation scales with the number of samples $n$ and not the dimesionality $d$ of the samples.
-More in Section 9.3 of \cite{cherkassky2007learning}. *** \\
+\\ \TODO{Add explanation why to use dual, which is: because the problem formulation scales with the number of samples $n$ and not the dimesionality $d$ of the samples.
+More in Section 9.3 of \cite{cherkassky2007learning}.} \\
 To do so, the Lagrange multipliers $a_i >= 0$ are introduced and the decision function becomes:
@@
 
-*** TODO: Add more on sparsity and good results, because of sparse representations, even in case of data of high dimensionality. e.g. \cite{cherkassky2007learning}. ***
+\TODO{Add more on sparsity and good results, because of sparse representations, even in case of data of high dimensionality. e.g. \cite{cherkassky2007learning}.}
 
@@
 As Smola \etal \cite{smola1998connection} state, this Gaussian kernel yields good performance, especially when no assumptions can be made about the data.
-The kernel maps input space $\mathcal{I}$ to the feature space $\mathcal{F}$ which is a Hilbert Space of infinite dimensions *** ref needed ***:
+The kernel maps input space $\mathcal{I}$ to the feature space $\mathcal{F}$ which is a Hilbert Space of infinite dimensions \TODO{ref needed}:
 \begin{equation}
@@
 
-*** Add something about robustness, discusses in section 9.2 from \cite{cherkassky2007learning} ***
+\TODO{Add something about robustness, discusses in section 9.2 from \cite{cherkassky2007learning}}
 
@@
 
-*** Maybe add section about ``Change indication'' rather than ``detection''? ***
+\TODO{Maybe add section about ``Change indication'' rather than ``detection''?}
 
@@
 
-*** These papers need to be referenced and included ***
+\TODO{These papers need to be referenced and included}
 

### 2013-10-24 Chapters/Chapter4/change_indication.tex
Added the TODO command

--- 2013-10-10 Chapters/Chapter4/change_indication.tex
@@
 \section{Change Indication}\label{sec:method_change_indication}
-*** TODO: change section title ***
-*** TODO: ***
+\TODO{change section title}
+\TODO{***}
 \emph{Follow methodology of ``A unifying framework for detecting outliers and change points from time series'' \cite{takeuchi2006unifying}.
@@
 
-The problem statement and formal definition, following Tackeuchi and Yamanishi \cite{takeuchi2006unifying} and Camci \cite{camci2010change} is the following.
+\TODO{the following paragraph is copied into \ref{subsec:change_detection_problem_formulation}. Remove here? Or shorten?}\\
+The problem statement and formal definition, following Takeuchi and Yamanishi \cite{takeuchi2006unifying} and Camci \cite{camci2010change} is the following.
 The algorithm needs to find \emph{sudden} changes in the time series data.
@@
 The problem with this measure is that, as the authors conclude and Camci discusses, it is not able to detect a change by decrease in variance.
+\\\TODO{End copied paragraph}
 
@@
 
-*** TODO: What is my contribution compared to Camci? Distance of outliers to hypersphere? ***
+\TODO{What is my contribution compared to Camci? Distance of outliers to hypersphere?}

### 2013-10-24 Chapters/old_classification/Chapter1/lit_rev_segm_and_class.tex
Added the TODO command

--- 2013-03-11 Chapters/Chapter1/literature_review.tex
@@
 % !TEX root = ../../main.tex
-
-% \section{Literature review}
-% Look into earlier application mentioned in the literature to this kind of  problems. Look for similarities in the problem and address where the  techniques used fail or are not applicable.
-
-% *** Most of the activity recognition techniques can be classified in two  different types.
-% Some rely on a state-space model, in which the activities to be recognized are represented in a statistical manner.
-% Bayesian networks, Finite State Machines and Hidden Markov Models can be considered among those.
-% Other techniques process the directly as a pattern recognition task, such as  Support Vector Machines, Neural Networks, Dynamic Time Warping, and Bayes and K-means clustering.
-
-% \subsection{Segmentation}
-% Describe the different approaches to segmentation.
-% Group different viewpoints, e.g. cut-points, change detection, segmentation, etc.
-
-% \section{Literature review}
-% This section will provide a review of the existing literature and research of the different subjects discussed in this thesis.
-% First the different approaches to mere segmentation of signals is discussed, followed by *** other techniques. ***
-
-% \subsection{Segmentation}
-% One of the first stages in analyzing the sensor data time series is to segment the signal.
-% The goal is to set cut-points in the signal such that all the data points in between consecutive cut-points have some homogeneous property among them.
-*** "organizing data into homogeneous groups where the within-group-object similarity is minimized and the between-group-object dissimilarity is maximized." ***
-% This results in the more abstract notion of change (point) detection.
-% The general concern is to detect the occurrence and time identification of changes in the signal, often expressed as changes in the probability distribution of the time series.
-% This literature review will discuss the different types of change detection as applied in the research.
-% To each applications question about the methods, complexity and usability etc. are asked.
-% On account of usability, it is desired to have a successful segmentation algorithm with minimal user interaction, e.g. defining thresholds and error rates.
-% ** something about what is considered, e.g. complexity, measurements, e.g. ***
-
-% Following the organization of \cite{warren2005clustering}, research can be structured in three groups.
-% The data being processed determines the classification, which can be the raw data, indirectly via features extracted from the raw data or indirectly via models extracted from the raw data.
-
-% One of the more simple and directs methods is to Piecewise Linear Represent the signal, while keeping the approximation error below a certain value, as discussed in \cite{keogh2001online}.
-% As with many algorithms, PLR can be applied to a batch of or real time stream input data (also known as off-line and on-line analyzes, respectively).
-% Keogh et al. \cite{keogh2001online} propose a method which combines these different techniques to form a Sliding Window and Bottom-up (SWAB) analyzes.
-% The algorithm works by first, per window of data points, creating for every two data points a segment.
-% Neighboring segments are joined together, as long as the approximation error is below the provided value.
-% This analyzes can be used as a starting point to further analyze the data segments.
-% Although Keogh et al. apply the SWAB algorithm only direct on the raw data, it is easily imagined how to apply it to extracted features of the signal, e.g. the energy or mean.
-% In its default application, the method yields little information about each segment; it is merely a linear representation.
-% The SWAB algorithm works by segmenting parts of the data series by linear approximation while keeping the error below a certain rate.
-% This rate needs to be user defined (e.g. by an expert), which makes it not possible to apply directly on any signal.
-% This method only divides the signal into segments, which representations can be processed further to form clusters or related segments.
-% The complexity of this method is relatively low, $O(Ln)$, where $L$ is the expected segment length and $n$ the number of data points.
-
-% The task of segmenting any form of data is made easier when the number of segments is beforehand known.
-% Many authors apply the well-known \emph{k-means} clustering *** refs needed *** on this type of problems.
-% An incremental error-minimizing version can determine the optimal number of segments (or actually, clusters in this case).
-% The proposed method of \cite{himberg2001time}, Global Iterative Replacement, resembles the method of Keogh et al., in the sense that it merges segments.
-% The algorithm is applied on two different data sets.
-% The first data sets can be considered as extracted features from sensor data, whilst the second data set works on model properties and thus can give a richer contextual segmentation.
-% For both types of data sets a cost function is defined, which represents the internal heterogeneity of a segment (e.g. the variance of the data within a segment).
-% The difference with the method of Keogh et al. is that with GIR the number of segments is known beforehand and during the algorithm this number is, through splits and merges, kept constant.
-% By running multiple instances of the GIR algorithm with each an increasing number of $k$ segments (bounded by user provided values), the optimal number and location of cut points is determined.
-
-% Other approaches form a higher level analyses of the raw data.
-% As used in \cite{barbivc2004segmenting}, from the field of computer graphics, the dimensionality of the data can be used to distinguish between actions.
-% Using Principal Component Analyses, Barbi{\v{c}} et al.
-use the dimensionality required to project the data points on a hyperplane while keeping the error below a certain rate as a measure of similarity.
-% The error threshold is fixed as a range of three times the standard deviation from the mean of the error.
-% In an extension of their method, using Probabilistic PCA \cite{tipping1999probabilistic}, the Mahanalobis distance \cite{duda1995pattern} for each frame represents similarity of concurrent motions.
-% By traversing the Mahanalobis distance as a function of the frames, cut-points are placed on specific points of interest.
-% *** NOTE to self: thus this method can segment (and group) unknown and new activities? ***
-% In a third method the authors use a Gaussian Mixture Model *** elaborate on this one [does it segment of direct classify?] ***.
-% Using this segmenting, and comparing the segments with earlier gathered representations of motions, the PPCA method gets the best results, with precision of 92\% and recall of 95\%.
-% The test subjects wore 14 sensors on joints of the body and each frame is represented as a 56-dimensional vector.
-% The test considered seven simple human motions: walking, running, sitting down, forward jumping, climbing, arm stretching and punching.
-
-% It is very common the use a higher level of features extracted from the raw data.
-% The approach of Zhou et al. \cite{zhou2008aligned} maps the data points to a higher dimension by using a kernel function.
-% The proposed method, Aligned Cluster Analysis, extends kernel $k$-means clustering on this higher dimensional data set.
-% Using the Dynamic Time Aligned Kernel (DTAK) metric, similarity between segments is measured.
-% Due to the properties of $k$-means clustering, discovering an unknown number of segments in the data set requires and iterative approach.
-
-% *** motifs ***
-
-% *** HMM ***
-
-% *** Finite mixture models ***
-
-% *** Infinite mixture models ***
-
-\section{Literature review}
-This section will provide a literature review per coherent subject of this thesis.
-In subsection \ref{sec:lit_review_segmentation} the research in segmentation and classification of human activities will be discussed.
 
@@
 
-During the last ten years much research is performed, as a result of smaller and cheaper measuring devices. *** cite needed ***.
-With the introduction of smartphones with sensory hardware the research has shifted to these devices, as they create a more natural and ubiquitous setting for the test subject. *** cite needed ***.
+During the last ten years much research is performed, as a result of smaller and cheaper measuring devices. \TODO{cite needed}.
+With the introduction of smartphones with sensory hardware the research has shifted to these devices, as they create a more natural and ubiquitous setting for the test subject. \TODO{cite needed}.
 
@@
 
-Very much used in machine learning problems are neural networks. *** cite needed ***
-Although these kind of networks typically perform well on spatial-typed problems *** cite needed ***, adjustments have been made to apply them for temporal classification.
+Very much used in machine learning problems are neural networks. \TODO{cite needed}
+Although these kind of networks typically perform well on spatial-typed problems \TODO{cite needed}, adjustments have been made to apply them for temporal classification.
 The methods of \cite{yang2008using, kwapisz2011activity} create a multi-level neural network to generate complex discriminators as activity classifiers.
@@
 Some researchers have made their used data set explicitly public available and other data sets are just available, providing valuable labeled activity patterns.
-*** This should be in experiments setup?
-***
-*** List here the datasets used ***
+\TODO{This should be in experiments setup?}
+\TODO{List here the datasets used}
 
@@
   \hline
+  many more... & & & & & & \\
+  \hline
 \end{tabular}
@@
 \end{table}
-
-\input{./Chapters/Chapter1/paper_list}

### 2013-10-24 Chapters/old_classification/Chapter1/lit_rev_temp_segm.tex
Create outlines for new focus: temporal segmentation
Added the TODO command

--- 2013-03-22 Chapters/Chapter1/lit_rev_temp_segm.tex
@@
 
-*** Piecewise linear approximation ***
+\TODO{Piecewise linear approximation}
 
@@
 
-The time series data from a sensor can be considered as a stochastic process *** is this a correct term? ***.
+The time series data from a sensor can be considered as a stochastic process \TODO{is this a correct term?}.
 Probabilistic models can be constructed on that signal, yielding in probabilistic and Bayesian based segmentation methods.
@@
 It is observed that when no change point is detected for a long period of time, the computational complexity increases significantly.
+
+\TODO{TODO} aan het einde van de lopende tekst richting-zinnen maken, die het einde aanduiden of een erg beknopte samenvatting geven / conclusie van literatuuronderzoek.

### 2013-10-24 Chapters/Chapter3/one_class_svm.tex
Add info on kernels

@@
 For this it introduces \emph{slack variables} $\xi_i$ for each data point and the constant $C > 0$ the determines the trade-off between maximizing the margin and the number of data points within that margin (and thus the training errors).
+The slack variables minimize the \emph{sum of deviations} rather than the \emph{number} of incorrect data points \cite{cherkassky2007learning}.
 The objective function for a \gls{svm} is the following minimization function:
@@
 
+A common class for kernel function is the \acrlong{rbf}.
+These function are of the form
+\begin{equation}
+  g(x) = \operatorname*{sign} \left(  \sum_{i=1}^{n} \alpha_i \operatorname*{exp} \left\{ \frac{|\vectorsym{x} - \vectorsym{x}_i|^2}{\sigma^2} \right\} \right),
+\end{equation}
+where $\sigma$ defines the width.
+The inner product kernel is of the form
+\begin{equation}
+  H(\vectorsym{x}, \vectorsym{x}') = \operatorname*{exp} \left\{ - \frac{| \vectorsym{x} - \vectorsym{x}'|^2}{\sigma^2} \right\}.
+\end{equation}
+
+Note that with the \gls{rbf} class of kernel functions the number of basis functions, the center parameters that correspond to the support vectors, and the weights in the output layer are all automatically determined via the optimal hyperplane \cite{cherkassky2007learning}.
+The width parameter is equal for all basis functions and determined a priori.
+Other examples of kernel function classes are \emph{pylonomial functions} of degree $q$, \emph{neural networks}, \emph{splines} of order $m$ with $b$ nodes, and \emph{Fourier expansions}.
+\TODO{More on basis functions. \cite{cherkassky2007learning}, chapter 9.4}
+
 $k(\mathbf{x}_i, \mathbf{x})$ is a kernel function computing a dot product in feature space, introduced by Aizerman \etal \cite{aizerman1964theoretical}.

### 2013-10-25 Chapters/Chapter3.tex
Draft, sent to Anne

@@
 Two \acrlong{svm}-based methods, \gls{nu-svm} and \gls{svdd}, are further disussed in section \ref{sec:one_class_svm}.
-\TODO{add \ref{sec:svm_parameters}}
 For the application to the problem of \acrlong{har}, the characteristics of the collected sensor data is discussed in \ref{sec:sensor_data_characteristics}.
@@
 \input{./Chapters/Chapter3/one_class_svm}
-\input{./Chapters/Chapter3/svm_parameters}
 \input{./Chapters/Chapter3/sensor_data_characteristics}

### 2013-10-25 Chapters/Chapter3/one_class_svm.tex
Draft, sent to Anne

@@
 When the hyperplane is projected to the (lower) original input space $\mathcal{I}$ it creates a non-linear separating curve.
+This is illustrated in Figure \ref{fig:svm_mapping_spaces}.
 The mapping and projection of data points can be efficient (and implicit) performed by using the kernel trick, which is discussed in section \ref{subsec:kernel_trick}.
@@
 Thus, the \gls{svm} searches for a maximal separating hyperplane.
+
+\begin{figure}
+\centering
+  \includegraphics[width=0.5\textwidth]{./Figures/chapter3/svm_mapping_spaces.pdf}
+  \caption[Mapping spaces in \gls{svm}]{The input vector $\vectorsym{x}$ from input space $\mathcal{I}$ is mapped by a non-linear function $\phi(\vectorsym{x})$ to a feature vector $\vectorsym{z}$ in the high dimensional feature space $\mathcal{F}$.
+The weights of $\vectorsym{w}$ create a linear separating hypeplace, which maps the high dimensional vector to the predicted outcome $\vectorsym{y}$.
+\TODO{Is this figure obsolete?
+Figure \ref{fig:kernel_mapping} shows almost the same}}
+  \label{fig:svm_mapping_spaces}
+\end{figure}
 
@@
 
-This minimization problem can be solved (using \gls{qp}) and transformed to its Lagrange dual formulation.
-\\ \TODO{Add explanation why to use dual, which is: because the problem formulation scales with the number of samples $n$ and not the dimesionality $d$ of the samples.
-More in Section 9.3 of \cite{cherkassky2007learning}.} \\
-To do so, the Lagrange multipliers $a_i >= 0$ are introduced and the decision function becomes:
+This minimization problem can be solved (using \gls{qp}) and transformed to its (Lagrange) dual formulation.
+In the dual formulation the problem scales with the number of training examples $n$ instead of the dimensionality $d$ of the samples.
+Solving this problem directly in the high dimensional feature space $\mathcal{F}$ makes it untractable.
+The linear approximation function corresponds to the kernel function in the dual formulation.
+Solving this dual formulation is equivalent to solving the primal formulation \cite{cherkassky2007learning}.
+In the dual formulation the Lagrange multipliers $a_i >= 0$ are introduced and the decision function becomes:
 \begin{equation}\label{eq:svm_lagrange}
@@
 With that problem only data from one class is available and the objective is to recognize new data points that are not part of that class.
-This problem is closely related to density estimation.
+This unsupervised learning problem is closely related to density estimation.
 In that context, the problem can be the following.
@@
 
-\TODO{Add something about robustness, discusses in section 9.2 from \cite{cherkassky2007learning}}
+\TODO{Add something about robustness, discusses in section 9.2 from \cite{cherkassky2007learning,chen2004m}}
 
@@
 Instead of a separating hyperplane, constructed by \gls{nu-svm} and illustrated on the left of the Figure, the \gls{svdd} creates a hypersphere (in the illustration a cricle) around the data points.
-By using kernel functions (\eg the \gls{rbf}) the hyperspheres in the high dimensional feature space $\mathcal{F}$ corresponds to a flexible and tight enclosing boundary in input space $\mathcal{I}$, illustrated in Figure \ref{fig:svdd-boundary}.
+By using kernel functions (\eg the \gls{rbf}) the hyperspheres in the high dimensional feature space $\mathcal{F}$ corresponds to a flexible and tight enclosing boundary in input space $\mathcal{I}$.
+Possible resulting closed boundaries are illustrated in Figure \ref{fig:svdd-boundary}.
+This enclosing boundary is obtained by minimizing:
 
@@
 
+\begin{figure}
+  \centering
+    \includegraphics[width=0.5\textwidth,keepaspectratio]{./Figures/chapter3/svdd-boundary.pdf}
+  \caption[\gls{svdd} boundary]{The \gls{svdd} method trained on a banana-shaped data set with different sigma-values for the \gls{rbf} kernel.
+Solid circles are support vectors, the dashed line is the boundary.
+Image by Tax \cite{tax2001one}.}
+  \label{fig:svdd-boundary}
+\end{figure}
+
 %--------------------------------------------
@@
 \begin{equation}
-  H(\vectorsym{x}, \vectorsym{x}') = \operatorname*{exp} \left\{ - \frac{| \vectorsym{x} - \vectorsym{x}'|^2}{\sigma^2} \right\}.
+  K(\vectorsym{x}, \vectorsym{x}') = \operatorname*{exp} \left\{ - \frac{| \vectorsym{x} - \vectorsym{x}'|^2}{\sigma^2} \right\}.
 \end{equation}
@@
 Note that with the \gls{rbf} class of kernel functions the number of basis functions, the center parameters that correspond to the support vectors, and the weights in the output layer are all automatically determined via the optimal hyperplane \cite{cherkassky2007learning}.
-The width parameter is equal for all basis functions and determined a priori.
+The width parameter $\sigma$ is equal for all basis functions and is set a priori and determines the flexibility and complexity of the boundary.
+In Section \ref{subsec:svm_model_parameters} this (hyper)parameter for a \gls{svm} is further discussed.
 Other examples of kernel function classes are \emph{pylonomial functions} of degree $q$, \emph{neural networks}, \emph{splines} of order $m$ with $b$ nodes, and \emph{Fourier expansions}.
@@
 
+-- Quotes and refs to use --
 $k(\mathbf{x}_i, \mathbf{x})$ is a kernel function computing a dot product in feature space, introduced by Aizerman \etal \cite{aizerman1964theoretical}.
@@
 
-\begin{figure}
-  \centering
-    \includegraphics[width=0.5\textwidth,keepaspectratio]{./Figures/chapter3/svdd-boundary.pdf}
-  \caption[\gls{svdd} boundary]{The \gls{svdd} method trained on a banana-shaped data set with different sigma-values for the \gls{rbf} kernel.
-Solid circles are support vectors, the dashed line is the boundary.
-Image by Tax \cite{tax2001one}.}
-  \label{fig:svdd-boundary}
-\end{figure}
+%--------------------------------------------
+
+\subsection{SVM model parameters}\label{subsec:svm_model_parameters}
+\gls{svm}-model selectiong and tuning depends on two type of parameters \cite{cherkassky2007learning}:
+\begin{enumerate}
+  \item Parameters controlling the `margin' size,
+  \item Model parameterization, \eg the kernel type and complexity parameters.
+  For the \gls{rbf} kernel the width parameter determines the model complexity.
+\end{enumerate}
+
+In case of a \gls{rbf} kernel, the width parameter $\sigma$ determines the flexibility and complexity of the boundary.
+The value of this parameter greatly determines the outcomes of the algorithm (\eg \gls{svdd}) as illustrated in Figure \ref{fig:svdd-boundary}.
+With a small value for the kernel width $\sigma$, each data point will tend to be used as a support vector and will represent a small Gaussian.
+The solution then corresponds with standard non-parametric density estimation.
+With a large value for the width $\sigma$, the boudary approximates the spherical boundary.
+
+\TODO{More on SVM parameters, \cite{cherkassky2007learning}, section 9.8, page 446}
+
+\TODO{Analysis of C/$\nu$-parameter, which sets a lower bound for the fraction of support vectors and an uppor bound for samples that will be outliers.
+\cite{flach2012machine} section 9.8 and \cite{scholkopf2002learning} and other Sch\"olkopf material.}
 
@@
 ``On simple one-class classification methods'' \cite{noumir2012simple}. 2012  --> decouples radius and center optimization, gives fast approximations instead of precise results. \\
-
-``Learning from Data'' \cite{cherkassky2007learning}. 2007, 1547 (book).

### 2013-10-25 Chapters/Chapter5/artificial_data.tex
Draft, sent to Anne

% !TEX root = ../../main.tex
\section{Artificial Data}\label{sec:artificial_data}
\TODO{Use data from \cite{camci2010change,takeuchi2006unifying}}
In this section we present three data sets as used by \cite{camci2010change,takeuchi2006unifying} to provide for a objetive performance comparison.

### 2013-10-28 Chapters/Chapter4.tex
WIP: Chapter 4

--- 2013-05-08 Chapters/Chapter4.tex
@@
 \end{itemize}
+
+\subsection{Experiment notes}
+\begin{itemize}
+  \item ICSS/CUSUM is goed in het vinden van variance changes. Niet goed in mean veranderingen.
+  \item Problemen met SVM methode: ook na de verandering blijft de threshold nog stijgen;
+  \item geprobeerd direct model-reset te doen na change point, maar dan kan je achteraf niet meer verder analyseren.
+  \item Veranderingen in mean geven ``extra'' change point wanneer alle data weer in het window zit. Zie \ref{fig:jumping_mean_ratios} op de 50-tallen.
+\end{itemize}
+
+\begin{description}
+  \item[Data gathering] Explain data gathering methods. Refer to chapter 5 and 6 for articifial and real-world details.
+  \item[Model construction] Explain SVDD model construction and updating.
+  \item[Model properties] Explain the SVDD properties that are used to calculate the change indication.
+  \item[Change detection] Explain the interpretation of the properties, the possible (modulair) methods to give change indication
+\end{description}
+
+\TODO{Write chapter introduction/outline. Give overview of method approach and refer to each section for details of that stage.}
+
+\input{./Chapters/Chapter4/data_gathering}
+\input{./Chapters/Chapter4/model_construction}
+\input{./Chapters/Chapter4/model_properties}
+\input{./Chapters/Chapter4/change_detection}
+
+ --- OLD ---
+\input{./Chapters/Chapter4/change_indication}
+\input{./Chapters/Chapter4/overview}
+\input{./Chapters/Chapter4/algorithms}

### 2013-10-28 Chapters/Chapter4/change_detection.tex
WIP: Chapter 4

% !TEX root = ../../main.tex
\section{Change Detection}\label{sec:method_change_detection}

### 2013-10-28 Chapters/Chapter4/data_gathering.tex
WIP: Chapter 4

% !TEX root = ../../main.tex
\section{Data Gathering}\label{sec:method_data_gathering}
In this section we briefly discuss the different data gathering methods used for the change detection algorithms and experiments.
Section \ref{subsec:data_gathering_artificial} reviews the artificial data sets we will use.
In Section \ref{subsec:data_gathering_real_world} an overview of the real-world data sets used is provided.
Both sections refer to Chapters \ref{Chapter5} and \ref{Chapter6} for more details, respectively.

% ------
\subsection{Artificial data}\label{subsec:data_gathering_artificial}
In order to provide an objective comparison to other methods, we will use artificial data sets which are also used in the researches on which our method is based.
The data sets are from Takeuchi and Yamanishi \cite{takeuchi2006unifying} and Camci \cite{camci2010change}.
Both construct a collection of one-dimensional time series data according to the second order \gls{ar} model:
\begin{equation}
  x_t = a_1 x_{t-1} + a_2 x_{t-2} + \epsilon_t.
\end{equation}
In the different data series the mean and variance of the Gaussian random variable $\epsilon_t$ differs and changes at pre-determined change points.
Using this data set an objective quality measure over the change detection methods can be obtained and compared.
All the used data sets are listed and analyzed in Chapter \ref{Chapter5}.

% ------
\subsection{Real-world data}\label{subsec:data_gathering_real_world}
In the second type of data sets we apply our method of change detection and temporal segmentation to real-world data sets.
Our setup records the activities of humans performed both in and out door in an uncontrolled environment.
Activities performed include walking, running in both a straight line and a curve, standing, walking up and downstairs and sitting.
Using the time series data sensor output in the form of acceleration, gravity and rotation metrics, our method gives change points based on those signals.
By comparing the discovered change points with annotated change points (based on video recordings of the performed activities) we are able to give subjective results.
In Chapter \ref{Chapter6} we give a detailed analysis of the performed activities and the recorded data sets.

### 2013-10-28 Chapters/Chapter4/model_construction.tex
WIP: Chapter 4

% !TEX root = ../../main.tex
\section{Model Construction: Incremental SVDD}\label{sec:method_model_construction}
After the data is collected and pre-processed (or in the case of the artificial data sets: generated), we construct an online incremental sliding window model construction algorithm.
We follow the method and implementation introduced by Tax and Laskov \cite{tax2003online}, the \acrlong{isvdd} method.
This method combines the techniques of online, unsupervised and incremental learning methods with the earlier introduced \gls{oc-svm} algorithm \gls{svdd}.
The method is first initialized with a window length and then in every step a new data object is added to and the last data object is removed from the working set.

Using the following abstract form of the \gls{svm} optimization problem, the extension of the incremental \gls{svm} to the \gls{svdd} can be carried out:
\begin{equation}
  \operatorname*{max}_\mu \operatorname*{min}_{\substack{
    0 \le x \le C \\
    \vectorsym{a}^T \vectorsym{x} + b = 0}
  } : W = -\vectorsym{c}^T\vectorsym{x} + \frac{1}{2}\vectorsym{x}^T K\vectorsym{x} + \mu(\vectorsym{a}^T\vectorsym{x} + b),
\end{equation}
where $\vectorsym{c}$ and $\vectorsym{a}$ are $n \times 1$ vectors, $K$ is a $n \times n$ matrix and $b$ is a scalar.
The \gls{svdd} implementation of this abstract form is set by the parameters $\vectorsym{c}=\operatorname*{diag}(K)$, $\vectorsym{a} = \vectorsym{y}$ and $b=1$.
The procudure for the incremental version has two operations: adding and removing a data object $k$.
When a data object $k$ added, its weight $x_k$ is initially set to $0$.
In case of an object removal, the weight is forced to be $x_k=0$.
Both the operations conclude with the recalculation $\mu$ and the weights $\vectorsym{x}$ for all the objects, in order to obtain the optimal solution for the enlarged or reduced data set.
The incremental learning algorithm follows from these two operations: new data objects are added to and old data objects are removed from the working set.

The size of the initial window of data objects has a lower bound determined by the hyperparameter $C$ (Equation \ref{eq:svdd_objective}).
Because of the equality constraint $\sum_{i=1}^n a_i x_i = 1$ and the box constraint $0 \le x_i \le C$, the number of objects in the working set must be at least $\ceil{\frac{1}{C}}$.
\TODO{How to ensure large enough window/working set}

### 2013-10-28 Chapters/Chapter4/model_properties.tex
WIP: Chapter 4

% !TEX root = ../../main.tex
\section{Model Properties}\label{sec:method_model_properties}

### 2013-10-28 Chapters/Chapter5/artificial_data.tex
WIP: Chapter 4

@@
 In this section we present three data sets as used by \cite{camci2010change,takeuchi2006unifying} to provide for a objetive performance comparison.
+
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/notes/jumping_mean_camci.eps}
+  \caption[Jumping mean camci]{Jumping mean camci}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/notes/jumping_mean_camci_thresholds.eps}
+  \caption[Jumping mean camci thresholds]{Jumping mean camci, thresholds}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/notes/jumping_mean_camci_ratios_10.eps}
+  \caption[Jumping mean camci ratios]{Jumping mean camci ratios, last 10}
+  \label{fig:jumping_mean_ratios}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/notes/jumping_variance.eps}
+  \caption[Jumping variance]{Jumping variance}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/notes/jumping_variance_thresholds.eps}
+  \caption[Jumping variance thresholds]{Jumping variance, thresholds}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/notes/jumping_variance_ratios_10.eps}
+  \caption[Jumping variance ratios]{Jumping variance ratios, last 10}
+\end{figure}

### 2013-10-29 Chapters/Chapter3/one_class_classification.tex
Better formulas on SVDD

@@
 \TODO{Add refs for good results of OC-SVM}
-In the following section (\ref{sec:one_class_svm}) two implementations of \gls{oc-svm} will be discussed, the \gls{svdd} method of Tax and Duin \cite{tax2004support} and the \gls{nu-svm}-algoritm by Sch\"olkopf.
+In the following section (\ref{sec:one_class_svm}) two implementations of \gls{oc-svm} will be discussed, the \gls{svdd} method of Tax and Duin \cite{tax1999support} and the \gls{nu-svm}-algoritm by Sch\"olkopf.
 

### 2013-10-29 Chapters/Chapter3/one_class_svm.tex
Better formulas on SVDD

@@
 The first is the \gls{nu-svm} by Sch\"olkopf \etal \cite{scholkopf1999support} and closely follows the above problem statement regarding density estimation.
-The second, on which we will base our change detection method, is the \gls{svdd} by Tax and Duin \cite{tax2004support}.
+The second, on which we will base our change detection method, is the \gls{svdd} by Tax and Duin \cite{tax1999support}.
 
@@
 \subsection{\acrlong{svdd}}\label{subsec:oc-svm-svdd}
-The method introduced by Tax and Duin \cite{tax2004support}, known as \acrlong{svdd}, takes as spherical instead of planar approach.
+The method introduced by Tax and Duin \cite{tax1999support}, known as \acrlong{svdd}, takes a spherical instead of planar approach.
 The boundary, created in feature space $\mathcal{F}$, forms a hypersphere around the (high density region of the) data.
 The volume of this hypersphere is minimized to get the smallest enclosing boundary.
+The chance of accepting outlier objects is thereby also minimized \cite{tax2003online}.
 By allowing outliers using slacks variables, in the same manner as classical \gls{svm} and \gls{nu-svm}, a soft margin is constructed.
@@
 Possible resulting closed boundaries are illustrated in Figure \ref{fig:svdd-boundary}.
-This enclosing boundary is obtained by minimizing:
-
+This enclosing boundary is obtained by minimizing the following error function $L$ which contains the volume of the hypersphere and the distance from the boundary to the outlier objects:
 \begin{equation}\label{eq:svdd_objective}
-  \operatorname*{min}_{R,\ \mathbf{a} } R^2 + C \sum_{i=1}^n \xi_i
+  L(R, \vectorsym{a}, \vectorsym{\xi}) = R^2 + C \sum_{i=1}^n \xi_i
 \end{equation}
@@
 \end{equation}
+In the dual Lagrangian formulation of this error function $L$ the multipliers $\vectorsym{\alpha}$ are maximized:
+\begin{equation}\label{eq:svdd_lagrange}
+  L = \sum_{i} \alpha_i(x_i \cdot x_i) - \sum_{i,j} \alpha_i \alpha_j(x_i \cdot x_j)
+\end{equation}
+\begin{equation}
+  \begin{multlined}
+    \mbox{ subject to: } \\
+    \begin{aligned}
+    0 \le \alpha_i \le C, \sum_{i} \alpha_i = 1
+    \end{aligned}
+  \end{multlined}
+\end{equation}
 
-The dual formulation, by using Lagrange multipliers $\alpha_i$, can be used to test a new data point $z$ and is the following:
+In the maximization of Equation \ref{eq:svdd_lagrange} a large fraction of the multipliers $\alpha_i$ become zero and for a small fraction $\alpha_i > 0$.
+This small fraction, for which $\alpha_i$ is non-zero, are called the \glspl{sv} and these objects lie on the boundary of the description.
+The center of the hypersphere only depends on this small number of \glspl{sv} and the objects for which $\alpha_i = 0$ can be disregarded from the solution.
+Testing the membership of a (new) object $\vectorsym{z}$ is done by determining if the distance to the center $\vectorsym{a}$ of the sphere is equal or smaller to the radius $R$:
+\begin{equation}
+  \lvert \vectorsym{z} - \vectorsym{a} \rVert ^2 = (\vectorsym{z} \cdot \vectorsym{z}) - 2 \sum_{i} \alpha_i(\vectorsym{z} \cdot \vectorsym{x}_i) + \sum_{i,j}(\vectorsym{x}_i \cdot \vectorsym{x}_j) \le R^2
+\end{equation}
 
-\begin{equation}
-  \lVert z - \mathbf{x} \rVert ^2 = \sum_{i=1}^n \alpha_i \operatorname{exp} \left( \frac{- \lVert z - x_i \rVert ^ 2}{\sigma^2} \right) \geq -R^2/2 + C_R
-\end{equation}
+\TODO{Note of inner products and that it can be replaced by kernel functions. Good ref: page 4/158 from \cite{tax2002uniform} and \cite{tax1999support}.}
 
@@
 
+\TODO{Note that \gls{svdd} and \gls{nu-svm} have identical solutions when data is preprocessed to have unit form (in the case of \gls{nu-svm}) \cite{tax2002uniform}.}
+
 %--------------------------------------------
-
 \subsection{Kernel Function}\label{subsec:kernel_function}

### 2013-10-29 Chapters/Chapter4/model_construction.tex
Better formulas on SVDD

@@
 The \gls{svdd} implementation of this abstract form is set by the parameters $\vectorsym{c}=\operatorname*{diag}(K)$, $\vectorsym{a} = \vectorsym{y}$ and $b=1$.
-The procudure for the incremental version has two operations: adding and removing a data object $k$.
+The procedure for the incremental version has two operations: adding and removing a data object $k$.
 When a data object $k$ added, its weight $x_k$ is initially set to $0$.
@@
 Because of the equality constraint $\sum_{i=1}^n a_i x_i = 1$ and the box constraint $0 \le x_i \le C$, the number of objects in the working set must be at least $\ceil{\frac{1}{C}}$.
-\TODO{How to ensure large enough window/working set}
+Thus the algorithm is initialized by selection the first $\ceil{\frac{1}{C}}$ objects for the working set.
+In every step of the loop of the algorithm at least the same number of objects must be added as there are removed.
+By analyzing the \gls{kkt} conditions, \cite{tax2003online} shows the optimality of the algorithm.
+
+From experiments it shows that the online, \gls{isvdd}, method results in less false alarms than the static \gls{svdd}.
+An explanation for this is that \gls{isvdd} follows the changing data distribution, such that small changes over time, like a drift in mean or increase in frequency, continuously re-model the \gls{svm} representation.

### 2013-10-29 Chapters/Chapter3/one_class_svm.tex
Fix missing vertical bar
Wip: chapter 3 formulas, chapter 4 model construction and properties

@@
 Testing the membership of a (new) object $\vectorsym{z}$ is done by determining if the distance to the center $\vectorsym{a}$ of the sphere is equal or smaller to the radius $R$:
-\begin{equation}
-  \lvert \vectorsym{z} - \vectorsym{a} \rVert ^2 = (\vectorsym{z} \cdot \vectorsym{z}) - 2 \sum_{i} \alpha_i(\vectorsym{z} \cdot \vectorsym{x}_i) + \sum_{i,j}(\vectorsym{x}_i \cdot \vectorsym{x}_j) \le R^2
+\begin{equation}\label{eq:svdd_test_object}
+  \lVert \vectorsym{z} - \vectorsym{a} \rVert ^2 = (\vectorsym{z} \cdot \vectorsym{z}) - 2 \sum_{i} \alpha_i(\vectorsym{z} \cdot \vectorsym{x}_i) + \sum_{i,j}(\vectorsym{x}_i \cdot \vectorsym{x}_j) \le R^2
 \end{equation}
@@
 \TODO{Note of inner products and that it can be replaced by kernel functions. Good ref: page 4/158 from \cite{tax2002uniform} and \cite{tax1999support}.}
+
+Because the Gaussian \gls{rbf} often yields good (\ie tight) boundaries, this set of kernels fuctions is commonly used:
+\begin{equation}
+  (\vectorsym{x} \cdot \vectorsym{y}) \rightarrow K(\vectorsym{x},\vectorsym{y}) = \operatorname{exp}\left(- \frac{\lVert \vectorsym{x} - \vectorsym{y} \rVert ^2}{\sigma^2}\right)
+\end{equation}
+Using this kernel function, the Lagrangian error function $L$ of Equation \ref{eq:svdd_lagrange} changes to:
+\begin{equation}\label{eq:svdd_lagrange_kernel}
+  L = 1 - \sum_{i} \alpha_i^2 - \sum_{i \ne j} \alpha_i \alpha_j K(x_i, x_j)
+\end{equation}
+Using Equation \ref{eq:svdd_test_object}, the following kernel formulation needs to hold for a new object $\vectorsym{z}$ to lie within the hypersphere:
+\begin{equation}\label{eq:svdd_inequality}
+  \sum_{i} \alpha_i K(\vectorsym{z}, x_i) \le \frac{1}{2} \left( 1 - R + \sum_{i,j} \alpha_i \alpha_j K(x_i, x_j) \right)
+\end{equation}
 
@@
 \end{figure}
+
+\begin{figure}
+  \centering
+    \includegraphics[width=0.8\textwidth,keepaspectratio]{./Figures/chapter3/svdd-parameter-sigma.pdf}
+  \caption[\gls{svdd} boundary]{The \gls{svdd} method trained on a banana-shaped data set with different $\sigma$-values for the \gls{rbf} kernel.
+Solid circles are support vectors, the dashed line is the boundary.}
+  \label{fig:svdd-boundary-sigma}
+\end{figure}
+
+\TODO{Use Figure \ref{fig:svdd-boundary} or \ref{fig:svdd-boundary-sigma}?}
 
@@
 ``Training a \gls{svm} with Gaussian \gls{rbf} kernels corresponds to minimizing the specific cost function with a regularization operator'' \cite{smola1998connection}. \\
+
 ``Gaussian kernels tend to yield good performance under general smoothness assumptions and should be considered especially of no additional knowledge of the data is available.'' \cite{smola1998connection}.
 \\
+
 ``Choosing a small width of the kernels leads to high generalization error as it effectively decouples the separate basis functions of the kernel expansion into very localized functions which is equivalent to memorizing the data, whereas a wide kernel tends to oversmooth.'' \cite{smola1998connection}.
@@
 
-sparsity, RKHS: \cite{girosi1998equivalence}
+``For small values of $\sigma$ almost all $\alpha_i >0$ and the \gls{svdd} resembles a Parzen density estimation.'' \cite{tax2002uniform} page 4, \cite{tax1999support} page 4. \\
+
+sparsity, RKHS: \cite{girosi1998equivalence} \\
+
+Kernels for Support Vector Classifiers where first (?) used by Vapnik \cite{vapnik1998statistical}.
 
@@
 The value of this parameter greatly determines the outcomes of the algorithm (\eg \gls{svdd}) as illustrated in Figure \ref{fig:svdd-boundary}.
-With a small value for the kernel width $\sigma$, each data point will tend to be used as a support vector and will represent a small Gaussian.
-The solution then corresponds with standard non-parametric density estimation.
+With a small value for the kernel width $\sigma$, each data point will tend to be used as a support vector (for almost all $\alpha_i > 0$) and the \gls{svdd} solution resembles a Parzen density estimation.
+For large values of $\sigma$, the solution will resemble the original hypersphere solution is obtained (in contrast with a tight boundary around the data).
 With a large value for the width $\sigma$, the boudary approximates the spherical boundary.
@@
 \TODO{Maybe add section about ``Change indication'' rather than ``detection''?}
+
+-- Notes --
+\begin{itemize}
+  \item Low target rejection rate $f_{T-}$ and low outlier acceptance rate $f_{O+}$.
+When only target examples are present, the first one can be estimated by the number of support vectors that we obtain as a solution of Lagrangian \ref{eq:svdd_lagrange_kernel}.
+\end{itemize}
 

### 2013-10-29 Chapters/Chapter4/model_construction.tex
Wip: chapter 3 formulas, chapter 4 model construction and properties

@@
 \section{Model Construction: Incremental SVDD}\label{sec:method_model_construction}
-After the data is collected and pre-processed (or in the case of the artificial data sets: generated), we construct an online incremental sliding window model construction algorithm.
+After the data is collected and pre-processed (or in the case of the artificial data sets: generated), we construct an online incremental sliding window algorithm.
 We follow the method and implementation introduced by Tax and Laskov \cite{tax2003online}, the \acrlong{isvdd} method.
@@
 An explanation for this is that \gls{isvdd} follows the changing data distribution, such that small changes over time, like a drift in mean or increase in frequency, continuously re-model the \gls{svm} representation.
+For our method we use the re-modeling capability of the \gls{isvdd} algorithm to obtain change indication measure.
+The following section will discuss the properties used from the constructed \gls{oc-svm} model at every loop-step of the algorithm.

### 2013-10-29 Chapters/Chapter4/model_properties.tex
Wip: chapter 3 formulas, chapter 4 model construction and properties

@@
 \section{Model Properties}\label{sec:method_model_properties}
+
+-- Notes --
+\begin{itemize}
+  \item Error measure of \gls{svdd} is based on both volume of the sphere and number of outliers.
+  \item When the outliers are uniformly distributed around the target data, we should minimize the volume of the description.
+  \item To do this (minimize the volume) we need to estimate the volume.
+  \item \cite{tax2002uniform} states that when the outliers are uniformly distributed, the fraction of outliers that is accepted gives an estimate of the volume of the target data description (with respect to the volume of the outlier distribution).
+  \item When the fraction of outlier is kept constant (by the parameters $C$ and $\sigma$), and when it is known that more outliers are present in current working set, the volume must increase.
+  \item Equation \ref{eq:svdd_inequality} gives that the right term must increase for a larger volume, \ie the value of $R$ will increase.
+  \item Thus, an increase of value $R$ indicates an increase of outliers in the working set.
+  \item This indicates a possible new activity in the time series data.
+\end{itemize}
+
+In the previous section we have discussed the \gls{isvdd} method, which creates a \gls{oc-svm} representation of a working set of data objects at every step of the algorithms loop.
+This section shows how we interpret the constructed model, to obtain a measure which can be used for a indication of change points.
+How this obtained measure is used to indicate a change is discussed in the following section.
+
+The \gls{isvdd} algorithm creates a spherical \gls{oc-svm} representation of the working set at every step of the algorithm.
+This model is obtained by the minimization of Equation \ref{eq:svdd_objective}, which incorporates the radius $R$ of the sphere and the distances $\vectorsym{\xi}$ from the outliers to the boundary.
+We will use the the radius $R$ of the hypersphere as an indication of change.
+
+In \cite{tax2002uniform} Tax and Duin provide an analysis of the error of the \gls{svdd} algorithm.
+This error is based on
+\begin{inparaenum}[\itshape 1\upshape)]
+\item the fraction $f_{T-}$ of target objects that is rejected, and
+\item the fraction $f_{O+}$ of outliers that is accepted.
+\end{inparaenum}
+Since in \gls{occ} situation typically there are (almost) no examples of outlier objects, Tax and Duin construct a method to generate outliers, based on the assumption that the outliers are uniformly distributed around the target set.
+To minimize the error, calculated by the fractions $f_{T-}$ and $f_{O-}$, we should minize the volume of the target data description (\ie the boundary of \gls{svdd}).
+This is since the the fraction of accepted outliers $f_{O-}$ is an estimate of the volume of the target data description, with respect to the volume of the outlier distribution.
+Tax and Duin provide a method to optimize the parameters of the \gls{svdd} method, being the trade-off parameter $C$ and the \gls{rbf} kernel width $\sigma$.
+This optimization will result in the modification of the radius $R$ of Equation \ref{eq:svdd_objective} and affects the Lagrangian inequality \ref{eq:svdd_inequality}.
+
+\TODO{outliers is kept constant, thus Radius must increase}

### 2013-11-01 Chapters/Chapter2.tex
Replace `takes` by `follows`

@@
 
-The widely used \gls{cusum} method by Basseville \etal \cite{basseville1993detection} takes this approach.
+The widely used \gls{cusum} method by Basseville \etal \cite{basseville1993detection} follows this approach.
 It originates from control methods for detection from bench marks.
@@
 The density-estimation methods rely on the log-likelihood ratio between \glspl{pdf}.
-The method of Camci \cite{camci2010change} takes an other approach within the statistical framework, by using a \gls{svm}.
+The method of Camci \cite{camci2010change} follows an other approach within the statistical framework, by using a \gls{svm}.
 One problem it tries to overcome is the (claimed) weakness of many methods to detect a decrease in variance.
@@
 
-% The final method under consideration, change detection via (intrinsic) dimensionality reduction, takes a different point of view.
+% The final method under consideration, change detection via (intrinsic) dimensionality reduction, follows a different point of view.
 % Opposed to the other discussed methods, dimensionality reduction is framed in the MDL (Minimum Description Length) framework.

### 2013-11-01 Chapters/Chapter2/temporal_segmentation.tex
Replace `takes` by `follows`

@@
 Probabilistic models can be constructed on that signal, yielding in probabilistic and Bayesian based segmentation methods.
-The \gls{cusum}-methods takes a statistical approach and relies on the log-likelihood ratio \cite{gustafsson1996marginalized} to measure the difference between two distributions.
+The \gls{cusum}-methods follows a statistical approach and relies on the log-likelihood ratio \cite{gustafsson1996marginalized} to measure the difference between two distributions.
 To calculate the ratio, the probability density functions need to be calculated.

### 2013-11-01 Chapters/Chapter3/one_class_svm.tex
Replace `takes` by `follows`

@@
 \subsection{\acrlong{svdd}}\label{subsec:oc-svm-svdd}
-The method introduced by Tax and Duin \cite{tax1999support}, known as \acrlong{svdd}, takes a spherical instead of planar approach.
+The method introduced by Tax and Duin \cite{tax1999support}, known as \acrlong{svdd}, follows a spherical instead of planar approach.
 The boundary, created in feature space $\mathcal{F}$, forms a hypersphere around the (high density region of the) data.

### 2013-11-01 Appendices/AppendixA-summaries.tex
Prefix refs with non-breaking space and capitalize refs

@@
 
-\label{AppendixA} % For referencing this appendix elsewhere, use \ref{AppendixA}
+\label{AppendixA} % For referencing this appendix elsewhere, use~\ref{AppendixA}
 
@@
 Linear models are well studied, in particular for the problem of linear regression.
-This can be solved by the \emph{least-squares} method and classification as discussed in section \ref{least-squares}, the \emph{perceptron} as explained in section \ref{perceptron}.
-Linear regression with the \emph{support vector machine} is handled in section \ref{svm-explained} and used for probability density estimation in section \ref{linear-classifier-pdf}.
-The kernel trick used for learning non-linear models is explained in section \ref{non-linear}.
+This can be solved by the \emph{least-squares} method and classification as discussed in Section~\ref{least-squares}, the \emph{perceptron} as explained in Section~\ref{perceptron}.
+Linear regression with the \emph{support vector machine} is handled in Section~\ref{svm-explained} and used for probability density estimation in Section~\ref{linear-classifier-pdf}.
+The kernel trick used for learning non-linear models is explained in Section~\ref{non-linear}.
 
@@
 This means that during the training only the pairwise dot-products of the data is needed; this results in the \emph{n}-by-\emph{n} Gram matrix $\matrixsym{G} = \matrixsym{X}\matrixsym{X}^T$.
-This instance-based perspective will be further discussed in section \ref{svm-explained} about the support vector machine.
+This instance-based perspective will be further discussed in Section~\ref{svm-explained} about the support vector machine.
 
@@
 When we take the partial derivative with respect to $w\vectorsym{w}$ we see that the Lagrange multipliers define the weight vector as a linear combination of the training examples.
-This partial derivative is $0$ for an optimal weight we get that $\vectorsym{w} = \sum_{i=1}^{n} \alpha_i y_i \vectorsym{x}_i$, \emph{which is the same expression as for the perceptron derived in section \ref{perceptron}}.
+This partial derivative is $0$ for an optimal weight we get that $\vectorsym{w} = \sum_{i=1}^{n} \alpha_i y_i \vectorsym{x}_i$, \emph{which is the same expression as for the perceptron derived in Section~\ref{perceptron}}.
 By plugging $\vectorsym{w}$ and $t$ back into the Lagrange equation, we can eliminate these and get the dual optimization problem entirely formulated in terms of the Lagrange multipliers:
@@
 
-The second property enables powerful adaption for support vector machines to learn non-linear decision boundaries, as discussed in section \ref{non-linear}.
+The second property enables powerful adaption for support vector machines to learn non-linear decision boundaries, as discussed in Section~\ref{non-linear}.
 
@@
 About segmentation with SVMs.
-Will be main material for section \ref{sec:svm_in_change_detection} about SVMs.
+Will be main material for Section~\ref{sec:svm_in_change_detection} about SVMs.
 
@@
 \subsubsection{Problem formulation}
-[Not summarized here, useful for e.g. section \ref{statistical-framework}]
+[Not summarized here, useful for e.g. Section~\ref{statistical-framework}]
 
@@
 The window size is related to the sensitivity of the method to change; small windows are sensitive with high false alarm rate whilst large windows are slow to detect change and have low alarm rates.
-The algorithm is listed in table \ref{tab:alg-svcdp}.
+The algorithm is listed in table~\ref{tab:alg-svcdp}.
 Note that SVCPD can be applied directly to multidimensional data, whilst many other methods can only be applied to one-dimensional data.
@@
 There is a masking effect when there are multiple change points in the series; the order of small, medium and large variances result in the value of $D_k$.
-The iterative algorithm presented in this paper in section \ref{subsec:icss} is designed to lessen the masking effect.
+The iterative algorithm presented in this paper in Section~\ref{subsec:icss} is designed to lessen the masking effect.
 
@@
 A solution is to iteratively applying the method and dividing the series at each possible change point.
-The algorithm is presented in table \ref{tab:alg-icss}.
+The algorithm is presented in table~\ref{tab:alg-icss}.
 It is the third steps which reduces the masking effect and helps to ``fine tune'' the algorithm by (re)moving the potential change points by checking each location given the adjacent ones.

### 2013-11-01 Appendices/AppendixC.tex
working on outline of proposed method
WIP
Prefix refs with non-breaking space and capitalize refs

@@
 \clearpage
+
 \section{Joint segmentation of multivariate time series with hidden process regression for human activity recognition, by Chamroukhi}\label{sec:appendix-C-joint-segmentation}
@@
 
-Method and formulas very similair to description in Section \ref{sec:camci}.
+Method and formulas very similair to description in Section~\ref{sec:camci}.
 
@@
 %
-In this problem the distribution function $F(x)$ is unknown and instead we are given the \gls{idd} data $x_1,\dots,x_l$ generated by $F$.
+In this problem the distribution function $F(x)$ is unknown and instead we are given the \gls{iid} data $x_1,\dots,x_l$ generated by $F$.
 

### 2013-11-01 Chapters/Chapter1.tex
Rough finish of chapter 2
Fix typos, add comments
WIP
Prefix refs with non-breaking space and capitalize refs

@@
 
-\label{Chapter1} % For referencing the chapter elsewhere, use \ref{Chapter1}
+\label{Chapter1} % For referencing the chapter elsewhere, use~\ref{Chapter1}
 
@@
 \section{Outline}
+\emph{Not intented for the reader.}
 \begin{itemize}
-  \item Context of research (human activity recognition), real-world applications
+  \item Context of research (\gls{har}), real-world applications
   \item Current methods, wrapper vs. filter methods
-  \item Problem statement with current filter methods (which follows from Chapter \ref{Chapter3} which goes in-depth with methods).
+  \item Problem statement with current filter methods (which follows from Chapter~\ref{Chapter3} which goes in-depth with methods).
   \item Purpose of this research. E.g. "Find a better algorithm for short-activity segmentation"
+  \item Provide the \emph{motivation} for this approach
   \item Relate to real-world applications
+  \item Outline for rest of the thesis
 \end{itemize}
+
+-- Overall todo's --
+\TODO{The term \emph{emperical data/distruction} is nowhere used in the thesis. Is missing.}

### 2013-11-01 Chapters/Chapter2.tex
Prefix refs with non-breaking space and capitalize refs

@@
 
-\label{Chapter2} % For referencing the chapter elsewhere, use \ref{Chapter2}
+\label{Chapter2} % For referencing the chapter elsewhere, use~\ref{Chapter2}
 
@@
 The widely used \gls{cusum} method by Basseville \etal \cite{basseville1993detection} follows this approach.
-It originates from control methods for detection from bench marks.
-This method and some derivatiges are discussed and analyzed in section \ref{cusum}.
+It originates from control methods for detection from benchmarks.
+This method and some derivatiges are discussed and analyzed in Section~\ref{cusum}.
 
 Many methods rely on pre-specified parametric model assumptions and considers the data be independent over time, which makes it less flexible to real-world applications.
-The methods proposed by Kawahara \etal \cite{kawahara2009change} and Lui \etal \cite{liu2013change} try to overcome these problems by estimating the \emph{ratio} between the \gls{pdf}, instead of estimating each \gls{pdf}.
-This approach is discussed and analyzed in section \ref{density-ratio}.
+The methods proposed by Kawahara \etal \cite{kawahara2009change} and Lui \etal \cite{liu2013change} try to overcome these problems by estimating the \emph{ratio} between the \glspl{pdf}, instead of estimating each \gls{pdf}.
+This approach is discussed and analyzed in Section~\ref{density-ratio}.
 
@@
 A change in the \gls{pdf} is represented by a change in the radius of this sphere.
-Section \ref{sec:svm_in_change_detection} discusses the \gls{svm}-methods.
+Section~\ref{sec:svm_in_change_detection} discusses the \gls{svm}-methods.
 \TODO{Put emphasis that this is the source of inspiration for the chosen method?}
@@
 % It uses the estimated underlying number of parameters of the time series as a model for change detection.
-% Section \ref{dim-reduction} discuses this method.
+% Section~\ref{dim-reduction} discuses this method.
 

### 2013-11-01 Chapters/Chapter2/svm.tex
Added the TODO command
Draft, sent to Anne
Prefix refs with non-breaking space and capitalize refs

@@
 
-Many proposals in the field of remote sensing and \gls{har} make use of \glspl{svm} as a model construction method.
+Many proposals in the field of remote sensing and \gls{har} make use of \glspl{svm} as a (supervised) model learning method.
 An elaborate overview of applications is given in \cite{mountrakis2011support}.
@@
 
-The inner workings of \glspl{oc-svm} are discussed in detail in Section \ref{sec:one_class_svm}.
+The inner workings of \glspl{oc-svm} are discussed in detail in Section~\ref{sec:one_class_svm}.
+
+\TODO{Add notion of `supervised' vs `unsupervised' learning. \gls{oc-svm} is unsupervised.}
 
 % \section{Change-detection by Support Vector Machines}\label{svm}
-% *** REMOVE THIS SECTION? ***
+% \TODO{REMOVE THIS SECTION?}
 
@@
 % The retrieved (high dimension) data point can be in- or outside the earlier constructed hyper-sphere, thereby giving information about a possible change point.
-% *** NOTE: using the in/out position of a new datapoint is different from the approach in this thesis, by only using the model properties. ***
+% \TODO{NOTE: using the in/out position of a new datapoint is different from the approach in this thesis, by only using the model properties.}
 % Following the definition of Camci \cite{camci2010change}, the class description is obtained by minimizing $r^2$:
@@
 
-% *** Give visual explanation with circles and in- and out-side new data points ***
+% \TODO{Give visual explanation with circles and in- and out-side new data points}
 
-% *** Make clear what approach we use (model properties) and create link/bridge to next Chapter (3), which discusses SVM in detail. ***
+% \TODO{Make clear what approach we use (model properties) and create link/bridge to next Chapter (3), which discusses SVM in detail.}

### 2013-11-01 Chapters/Chapter2/temporal_segmentation.tex
Prefix refs with non-breaking space and capitalize refs

@@
 
-% ``An online algorithm for segmenting time series'' \cite{keogh2001online}. See \ref{sec:appendix-C-online-keogh}. 538, 2001 \\
+% ``An online algorithm for segmenting time series'' \cite{keogh2001online}. See~\ref{sec:appendix-C-online-keogh}. 538, 2001 \\
 % ``Segmenting time series: A survey and novel approach'' \cite{keogh2004segmenting}. 242, 2004 \\
@@
 
-% ``Joint segmentation of multivariate time series with hidden process regression for human activity recognition'' \cite{chamroukhi2013joint}. 0, 2013. See \ref{sec:appendix-C-joint-segmentation}. \\
+% ``Joint segmentation of multivariate time series with hidden process regression for human activity recognition'' \cite{chamroukhi2013joint}. 0, 2013. See~\ref{sec:appendix-C-joint-segmentation}. \\
 
@@
 
-Section \ref{sec:change_detection_time_series} discusses the problem of change detection in time series further, and gives a formal problem definition.
+Section~\ref{sec:change_detection_time_series} discusses the problem of change detection in time series further, and gives a formal problem definition.
 

### 2013-11-01 Chapters/Chapter3.tex
Prefix refs with non-breaking space and capitalize refs

@@
 
-\label{Chapter3} % For referencing the chapter elsewhere, use \ref{Chapter3}
+\label{Chapter3} % For referencing the chapter elsewhere, use~\ref{Chapter3}
 
@@
 
-This chapter discusses the concepts and algorithms that will be used as a basis for the proposed method, as discussed in Chapter \ref{Chapter4}.
+This chapter discusses the concepts and algorithms that will be used as a basis for the proposed method, as discussed in Chapter~\ref{Chapter4}.
 The first sections formulates the problem of change detection and relates it to outlier and novelty detection.
 It transforms the traditional problem to change detection for times series data.
-In Section \ref{sec:one_class_classification} the problem of \acrlong{occ} is discussed and shows a number of implementations.
-Two \acrlong{svm}-based methods, \gls{nu-svm} and \gls{svdd}, are further disussed in section \ref{sec:one_class_svm}.
-For the application to the problem of \acrlong{har}, the characteristics of the collected sensor data is discussed in \ref{sec:sensor_data_characteristics}.
-Finally a theoretic set of quality metrics is discussed in Section \ref{sec:quality_metrics}.
+In Section~\ref{sec:one_class_classification} the problem of \acrlong{occ} is discussed and shows a number of implementations.
+Two \acrlong{svm}-based methods, \gls{nu-svm} and \gls{svdd}, are further disussed in Section~\ref{sec:one_class_svm}.
+For the application to the problem of \acrlong{har}, the characteristics of the collected sensor data is discussed in~\ref{sec:sensor_data_characteristics}.
+Finally a theoretic set of quality metrics is discussed in Section~\ref{sec:quality_metrics}.
 

### 2013-11-01 Chapters/Chapter3/change_detection_time_series.tex
Prefix refs with non-breaking space and capitalize refs

@@
 
-\TODO{Note: the following block is copied from \ref{sec:method_change_indication}} \\
-The problem of change point detection can be formulated using different type of models, as discussed in \ref{subsec:segmentation}.
+\TODO{Note: the following block is copied from~\ref{sec:method_change_indication}} \\
+The problem of change point detection can be formulated using different type of models, as discussed in~\ref{subsec:segmentation}.
 The methods by Takeuchi and Yamanishi \cite{takeuchi2006unifying} and Camci \cite{camci2010change} use the following formulation for change detection, which we will also use for our problem formulation.
@@
 Since we are only interested in different activities (which are represented by sudden changes), slight changes within an activity are not of interest.
-Section \ref{sec:sensor_data_characteristics} discusses the representation of activities and changes between them in the data in more detail.
+Section~\ref{sec:sensor_data_characteristics} discusses the representation of activities and changes between them in the data in more detail.
 

### 2013-11-01 Chapters/Chapter3/one_class_classification.tex
Prefix refs with non-breaking space and capitalize refs

@@
 Since the algorithm's goal is to differentiate between normal and abnormal objects (relative to the training objects), \gls{occ} is often called outlier, novelty or anomaly detection, depending on the origin of the problem to which the algorithm is applied\footnote{The term \acrlong{occ} originates from Moya \etal \cite{moya1993one}.}.
-The difference between two and one-class classification and the consequence for outlier objects is illustrated in Figure \ref{fig:two-vs-one-classification}.
+The difference between two and one-class classification and the consequence for outlier objects is illustrated in Figure~\ref{fig:two-vs-one-classification}.
 In the two-class classification problem the object \textbf{o} will be member of the \textbf{-1} class whilst the \gls{occ} problem will label it as an outlier.
@@
 It is easy to see here a relation with similiary measures over \glspl{pdf} and the \gls{kl} divergence. \\
-\TODO{Add reference to \ref{subsec:change_detection_problem_formulation} which also mentions \gls{kl} divergence?}
+\TODO{Add reference to~\ref{subsec:change_detection_problem_formulation} which also mentions \gls{kl} divergence?}
 
 \subsection{One-Class Classification methods}\label{subsec:occ-methods}
-The methods and algorithms used for the \gls{occ}-problem can be organized into three categories \cite{tax2001one,noumir2012simple}, visually represented in Figure \ref{fig:occ-methods}.
+The methods and algorithms used for the \gls{occ}-problem can be organized into three categories \cite{tax2001one,noumir2012simple}, visually represented in Figure~\ref{fig:occ-methods}.
 The first category consists of methods that estimate the density of the training data and set a threshold on this density.
@@
 The number of objects that is required is smaller than in the case of density methods.
-The boundary method \gls{svdd}, constructed by Tax and which has shown good performance \cite{khan2010survey}, will be further discussed in Section \ref{subsec:oc-svm-svdd}.
+The boundary method \gls{svdd}, constructed by Tax and which has shown good performance \cite{khan2010survey}, will be further discussed in Section~\ref{subsec:oc-svm-svdd}.
 
@@
 It shows succesful applications for, amonst others, problems in the field of Handwritten Digit Recognition, Face Recognition Applications, Spam Detection and Anomaly Detection \cite{li2003improving,perdisci2006using}.
-As discussed in Section \ref{sec:change_detection_time_series}, this can be used for change detection in time series data.
+As discussed in Section~\ref{sec:change_detection_time_series}, this can be used for change detection in time series data.
 

### 2013-11-01 Chapters/Chapter3/one_class_svm.tex
Prefix refs with non-breaking space and capitalize refs

@@
 A \gls{svm} can create a boundary between linear-separable data points, making it a non-probabilistic binary linear classifier.
-More flexible non-linear boundaries can be obtained by the use of a non-linear function $\phi(x)$, as illustrated in Figure \ref{fig:kernel_mapping}.
+More flexible non-linear boundaries can be obtained by the use of a non-linear function $\phi(x)$, as illustrated in Figure~\ref{fig:kernel_mapping}.
 This function maps the input data from space $\mathcal{I}$ to a higher dimensional space $\mathcal{F}$.
@@
 When the hyperplane is projected to the (lower) original input space $\mathcal{I}$ it creates a non-linear separating curve.
-This is illustrated in Figure \ref{fig:svm_mapping_spaces}.
-The mapping and projection of data points can be efficient (and implicit) performed by using the kernel trick, which is discussed in section \ref{subsec:kernel_trick}.
+This is illustrated in Figure~\ref{fig:svm_mapping_spaces}.
+The mapping and projection of data points can be efficient (and implicit) performed by using the kernel trick, which is discussed in Section~\ref{subsec:kernel_trick}.
 
@@
 The data points which lie on the margin are the \emph{support vectors}.
-This geometrical interpretation is illustrated in Figure \ref{fig:svm_hyperplane}.
+This geometrical interpretation is illustrated in Figure~\ref{fig:svm_hyperplane}.
 All data points for which $y_i = -1$ are on one side of the hyperplane and all other data points (for which $y_i = 1$) are on the other side.
@@
 \TODO{Is this figure obsolete?
-Figure \ref{fig:kernel_mapping} shows almost the same}}
+Figure~\ref{fig:kernel_mapping} shows almost the same}}
   \label{fig:svm_mapping_spaces}
@@
 \end{equation}
-where $K(x, x_i) = \phi(x)^T\phi(x_i)$ (which is further discussed in section \ref{subsec:kernel_trick}).
+where $K(x, x_i) = \phi(x)^T\phi(x_i)$ (which is further discussed in Section~\ref{subsec:kernel_trick}).
 Here every data point $\mathcal{I}$ for which $a_i > 0$ is weighted in the decision function and thus ``supports'' the classification machine: hence the name ``\acrlong{svm}''.
@@
 \subsection{Kernel trick}\label{subsec:kernel_trick}
-In the previous section, \ref{subsec:svm}, the mapping function $\phi(x)$ and the kernel function $K$ were briefly mentioned.
-The decision function in equation \ref{eq:svm_lagrange} only relies on the dot-products of mapped data points in the feature space $\mathcal{F}$ (\ie all pairwise distances between the data points in that space).
+In the previous section,~\ref{subsec:svm}, the mapping function $\phi(x)$ and the kernel function $K$ were briefly mentioned.
+The decision function in equation~\ref{eq:svm_lagrange} only relies on the dot-products of mapped data points in the feature space $\mathcal{F}$ (\ie all pairwise distances between the data points in that space).
 It shows \cite{flach2012machine} that as long as any function has the same result, without an explicit mapping to the higher dimension $\mathcal{F}$, the dot-products can be substituted by the kernel function $K$.
 This is known as the \emph{kernel trick} and gives the \gls{svm} the ability to create non-linear decision function without high computational complexity.
-This mapping is illustrated in Figure \ref{fig:kernel_mapping}.
+This mapping is illustrated in Figure~\ref{fig:kernel_mapping}.
 Here the non-linear separating boundary in the input space $\mathcal{I}$ is mapped, via $\phi$, to a linear boundary in the feature space $\mathcal{F}$.
@@
 
-The constructed decision function $\mathcal{F}$ resembles the function discussed in Section \ref{subsec:svm}.
+The constructed decision function $\mathcal{F}$ resembles the function discussed in Section~\ref{subsec:svm}.
 It returns the value $+1$ in a (possibly small) region capturing most of the data points, and $-1$ elsewhere.
@@
 The hyperplane is represented by $g(x) = w \cdot \phi(x) + \rho = 0$ and the decision function is $f(x) = \operatorname{sgn}(g(x))$.
-This hyperplane and the separation from the origin is illustrated in Figure \ref{fig:nu-svm}.
+This hyperplane and the separation from the origin is illustrated in Figure~\ref{fig:nu-svm}.
 
@@
 
-In the classical \gls{svm} objective function, as denoted in Equation \ref{eq:svm_objective}, the parameter $C$ decided the smoothness of the boundary, with respect to the slack variables $\xi_i$.
+In the classical \gls{svm} objective function, as denoted in Equation~\ref{eq:svm_objective}, the parameter $C$ decided the smoothness of the boundary, with respect to the slack variables $\xi_i$.
 In the formulation of \gls{nu-svm} the equivalent parameter is $\nu \in (0,1)$ (hence the name).
@@
 Like the classical \gls{svm} and \gls{svdd} it can be required that all the distances from the data points $x_i$ to the center $\mathbf{a}$ are strict less then $R$ (or equivalent measure, to create a hard margin) or soft margins are allowed by using slack variables $\xi_i$.
-In the case of a soft margin, the penalty is determined by $C$ and the minimization is expressed as Equation \ref{eq:svdd_objective}.
-This principle is illustrated in the right image or Figure \ref{fig:nu-vs-svdd}.
+In the case of a soft margin, the penalty is determined by $C$ and the minimization is expressed as Equation~\ref{eq:svdd_objective}.
+This principle is illustrated in the right image or Figure~\ref{fig:nu-vs-svdd}.
 Instead of a separating hyperplane, constructed by \gls{nu-svm} and illustrated on the left of the Figure, the \gls{svdd} creates a hypersphere (in the illustration a cricle) around the data points.
 By using kernel functions (\eg the \gls{rbf}) the hyperspheres in the high dimensional feature space $\mathcal{F}$ corresponds to a flexible and tight enclosing boundary in input space $\mathcal{I}$.
-Possible resulting closed boundaries are illustrated in Figure \ref{fig:svdd-boundary}.
+Possible resulting closed boundaries are illustrated in Figure~\ref{fig:svdd-boundary}.
 This enclosing boundary is obtained by minimizing the following error function $L$ which contains the volume of the hypersphere and the distance from the boundary to the outlier objects:
@@
 
-In the maximization of Equation \ref{eq:svdd_lagrange} a large fraction of the multipliers $\alpha_i$ become zero and for a small fraction $\alpha_i > 0$.
+In the maximization of Equation~\ref{eq:svdd_lagrange} a large fraction of the multipliers $\alpha_i$ become zero and for a small fraction $\alpha_i > 0$.
 This small fraction, for which $\alpha_i$ is non-zero, are called the \glspl{sv} and these objects lie on the boundary of the description.
@@
 \end{equation}
-Using this kernel function, the Lagrangian error function $L$ of Equation \ref{eq:svdd_lagrange} changes to:
+Using this kernel function, the Lagrangian error function $L$ of Equation~\ref{eq:svdd_lagrange} changes to:
 \begin{equation}\label{eq:svdd_lagrange_kernel}
@@
 \end{equation}
-Using Equation \ref{eq:svdd_test_object}, the following kernel formulation needs to hold for a new object $\vectorsym{z}$ to lie within the hypersphere:
+Using Equation~\ref{eq:svdd_test_object}, the following kernel formulation needs to hold for a new object $\vectorsym{z}$ to lie within the hypersphere:
 \begin{equation}\label{eq:svdd_inequality}
@@
 
-\TODO{Use Figure \ref{fig:svdd-boundary} or \ref{fig:svdd-boundary-sigma}?}
+\TODO{Use Figure~\ref{fig:svdd-boundary} or~\ref{fig:svdd-boundary-sigma}?}
 
@@
 The width parameter $\sigma$ is equal for all basis functions and is set a priori and determines the flexibility and complexity of the boundary.
-In Section \ref{subsec:svm_model_parameters} this (hyper)parameter for a \gls{svm} is further discussed.
+In Section~\ref{subsec:svm_model_parameters} this (hyper)parameter for a \gls{svm} is further discussed.
 Other examples of kernel function classes are \emph{pylonomial functions} of degree $q$, \emph{neural networks}, \emph{splines} of order $m$ with $b$ nodes, and \emph{Fourier expansions}.
@@
 In case of a \gls{rbf} kernel, the width parameter $\sigma$ determines the flexibility and complexity of the boundary.
-The value of this parameter greatly determines the outcomes of the algorithm (\eg \gls{svdd}) as illustrated in Figure \ref{fig:svdd-boundary}.
+The value of this parameter greatly determines the outcomes of the algorithm (\eg \gls{svdd}) as illustrated in Figure~\ref{fig:svdd-boundary}.
 With a small value for the kernel width $\sigma$, each data point will tend to be used as a support vector (for almost all $\alpha_i > 0$) and the \gls{svdd} solution resembles a Parzen density estimation.
@@
   \item Low target rejection rate $f_{T-}$ and low outlier acceptance rate $f_{O+}$.
-When only target examples are present, the first one can be estimated by the number of support vectors that we obtain as a solution of Lagrangian \ref{eq:svdd_lagrange_kernel}.
+When only target examples are present, the first one can be estimated by the number of support vectors that we obtain as a solution of Lagrangian~\ref{eq:svdd_lagrange_kernel}.
 \end{itemize}

### 2013-11-01 Chapters/Chapter4.tex
Prefix refs with non-breaking space and capitalize refs

@@
 
-\label{Chapter4} % For referencing the chapter elsewhere, use \ref{Chapter4}
+\label{Chapter4} % For referencing the chapter elsewhere, use~\ref{Chapter4}
 
@@
 \begin{itemize}
-  \item Based on the problem statement with current research as stated in Chapter \ref{Chapter3}
+  \item Based on the problem statement with current research as stated in Chapter~\ref{Chapter3}
   \item Adjust method to needs
@@
   \item geprobeerd direct model-reset te doen na change point, maar dan kan je achteraf niet meer verder analyseren.
-  \item Veranderingen in mean geven ``extra'' change point wanneer alle data weer in het window zit. Zie \ref{fig:jumping_mean_ratios} op de 50-tallen.
+  \item Veranderingen in mean geven ``extra'' change point wanneer alle data weer in het window zit. Zie~\ref{fig:jumping_mean_ratios} op de 50-tallen.
 \end{itemize}

### 2013-11-01 Chapters/Chapter4/data_gathering.tex
Prefix refs with non-breaking space and capitalize refs

@@
 In this section we briefly discuss the different data gathering methods used for the change detection algorithms and experiments.
-Section \ref{subsec:data_gathering_artificial} reviews the artificial data sets we will use.
-In Section \ref{subsec:data_gathering_real_world} an overview of the real-world data sets used is provided.
-Both sections refer to Chapters \ref{Chapter5} and \ref{Chapter6} for more details, respectively.
+Section~\ref{subsec:data_gathering_artificial} reviews the artificial data sets we will use.
+In Section~\ref{subsec:data_gathering_real_world} an overview of the real-world data sets used is provided.
+Both sections refer to Chapters~\ref{Chapter5} and~\ref{Chapter6} for more details, respectively.
 
@@
 Using this data set an objective quality measure over the change detection methods can be obtained and compared.
-All the used data sets are listed and analyzed in Chapter \ref{Chapter5}.
+All the used data sets are listed and analyzed in Chapter~\ref{Chapter5}.
 
@@
 By comparing the discovered change points with annotated change points (based on video recordings of the performed activities) we are able to give subjective results.
-In Chapter \ref{Chapter6} we give a detailed analysis of the performed activities and the recorded data sets.
+In Chapter~\ref{Chapter6} we give a detailed analysis of the performed activities and the recorded data sets.

### 2013-11-01 Chapters/Chapter4/model_properties.tex
Prefix refs with non-breaking space and capitalize refs

@@
   \item When the fraction of outlier is kept constant (by the parameters $C$ and $\sigma$), and when it is known that more outliers are present in current working set, the volume must increase.
-  \item Equation \ref{eq:svdd_inequality} gives that the right term must increase for a larger volume, \ie the value of $R$ will increase.
+  \item Equation~\ref{eq:svdd_inequality} gives that the right term must increase for a larger volume, \ie the value of $R$ will increase.
   \item Thus, an increase of value $R$ indicates an increase of outliers in the working set.
@@
 The \gls{isvdd} algorithm creates a spherical \gls{oc-svm} representation of the working set at every step of the algorithm.
-This model is obtained by the minimization of Equation \ref{eq:svdd_objective}, which incorporates the radius $R$ of the sphere and the distances $\vectorsym{\xi}$ from the outliers to the boundary.
+This model is obtained by the minimization of Equation~\ref{eq:svdd_objective}, which incorporates the radius $R$ of the sphere and the distances $\vectorsym{\xi}$ from the outliers to the boundary.
 We will use the the radius $R$ of the hypersphere as an indication of change.
@@
 Tax and Duin provide a method to optimize the parameters of the \gls{svdd} method, being the trade-off parameter $C$ and the \gls{rbf} kernel width $\sigma$.
-This optimization will result in the modification of the radius $R$ of Equation \ref{eq:svdd_objective} and affects the Lagrangian inequality \ref{eq:svdd_inequality}.
+This optimization will result in the modification of the radius $R$ of Equation~\ref{eq:svdd_objective} and affects the Lagrangian inequality~\ref{eq:svdd_inequality}.
 

### 2013-11-01 Chapters/Chapter4/overview.tex
Added the TODO command
Draft, sent to Anne
Prefix refs with non-breaking space and capitalize refs

@@
   \centering
-    \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{./Figures/graphs/chapter4/method_setup.pdf}
+    \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{./Figures/chapter4/method_setup.pdf}
   \caption[Method setup]{Schematic representation of the change detection method.}
@@
 First described is the method to process the gathered sensor data.
-A schematic overview is given in figure \ref{fig:method_overview} and shows the steps of the method.
+A schematic overview is given in figure~\ref{fig:method_overview} and shows the steps of the method.
 A more detailed explanation of the ``Update model'' step follows.
@@
 \subsection{Change detection method}
-As graphically represented in figure \ref{fig:method_overview}, the change detection method starts by processing the data from sensor, such as the accelerometer, magnetic orientation and rotation measures \footnote{Something about the origin of the streams; all from the same sensor or different sensors?}.
+As graphically represented in figure~\ref{fig:method_overview}, the change detection method starts by processing the data from sensor, such as the accelerometer, magnetic orientation and rotation measures \footnote{Something about the origin of the streams; all from the same sensor or different sensors?}.
 
@@
 The model is (partially) reconstructed and new model properties, such as the radius of the hypersphere and the number of outliers, are the result of this step \footnote{Other measures are also possible, for instance the distance from all the outliers to the boundary of the hypersphere}.
-*** MORE ON THE UPDATING STEP ***
+\TODO{MORE ON THE UPDATING STEP}
 
@@
 This application was chosen for its convenient data format of the sensor recording and its regularity of the sampling interval.
-Table \ref{tab:recorded_metrics} lists the recorded metrics.
+Table~\ref{tab:recorded_metrics} lists the recorded metrics.
 For our expiments we used the data for the accelerometer, magnatief field and rotation.

### 2013-11-01 Chapters/Chapter5.tex
Rough finish of chapter 2
Draft, sent to Anne
WIP: Chapter 4
Prefix refs with non-breaking space and capitalize refs

@@
 
-\chapter{Result}
+\chapter{Artificial data results}
 
-\label{Chapter5} % For referencing the chapter elsewhere, use \ref{Chapter5}
+\label{Chapter5} % For referencing the chapter elsewhere, use~\ref{Chapter5}
 
-\lhead{Chapter 5. \emph{Results}} % This is for the header on each page - perhaps a shortened title
+\lhead{Chapter 5. \emph{Artificial data results}} % This is for the header on each page - perhaps a shortened title
 
@@
 \section{Outline}
+\emph{Not intented for the reader.}
 \begin{itemize}
-  \item Compare proposed method with methods of Chapter \ref{Chapter2}
+  \item Compare proposed method with methods of Chapter~\ref{Chapter2}
+  \item Provide KL-analysis (as Takeuchi does)
   \item Provide plots, tables, graphs, error rates, precision, etc.
@@
 \end{itemize}
+
+% \input{./Chapters/Chapter5/artificial_data}
+% \input{./Chapters/Chapter5/real_world_data}

### 2013-11-01 Chapters/Chapter6.tex
Rough finish of chapter 2
WIP: literature review SVM
WIP: Chapter 4
Prefix refs with non-breaking space and capitalize refs

@@
 
-\chapter{Real-world applications}
+\chapter{Real-world results}
 
-\label{Chapter6} % For referencing the chapter elsewhere, use \ref{Chapter6}
+\label{Chapter6} % For referencing the chapter elsewhere, use~\ref{Chapter6}
 
-\lhead{Chapter 6. \emph{Real-world applications}} % This is for the header on each page - perhaps a shortened title
+\lhead{Chapter 6. \emph{Real-world results}} % This is for the header on each page - perhaps a shortened title
 
@@
 \end{itemize}
+
+\input{./Chapters/Chapter6/data_gathering}

### 2013-11-01 Chapters/ChapterTemplate.tex
Prefix refs with non-breaking space and capitalize refs

% Chapter Template

\chapter{Chapter Title Here} % Main chapter title

\label{ChapterX} % Change X to a consecutive number; for referencing this chapter elsewhere, use~\ref{ChapterX}

\lhead{Chapter X. \emph{Chapter Title Here}} % Change X to a consecutive number; this is for the header on each page - perhaps a shortened title

%----------------------------------------------------------------------------------------
%	SECTION 1
%----------------------------------------------------------------------------------------

\section{Main Section 1}

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Aliquam ultricies lacinia euismod. Nam tempus risus in dolor rhoncus in interdum enim tincidunt. Donec vel nunc neque. In condimentum ullamcorper quam non consequat. Fusce sagittis tempor feugiat. Fusce magna erat, molestie eu convallis ut, tempus sed arcu. Quisque molestie, ante a tincidunt ullamcorper, sapien enim dignissim lacus, in semper nibh erat lobortis purus. Integer dapibus ligula ac risus convallis pellentesque.

%-----------------------------------
%	SUBSECTION 1
%-----------------------------------
\subsection{Subsection 1}

Nunc posuere quam at lectus tristique eu ultrices augue venenatis. Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Aliquam erat volutpat. Vivamus sodales tortor eget quam adipiscing in vulputate ante ullamcorper. Sed eros ante, lacinia et sollicitudin et, aliquam sit amet augue. In hac habitasse platea dictumst.

%-----------------------------------
%	SUBSECTION 2
%-----------------------------------

\subsection{Subsection 2}
Morbi rutrum odio eget arcu adipiscing sodales. Aenean et purus a est pulvinar pellentesque. Cras in elit neque, quis varius elit. Phasellus fringilla, nibh eu tempus venenatis, dolor elit posuere quam, quis adipiscing urna leo nec orci. Sed nec nulla auctor odio aliquet consequat. Ut nec nulla in ante ullamcorper aliquam at sed dolor. Phasellus fermentum magna in augue gravida cursus. Cras sed pretium lorem. Pellentesque eget ornare odio. Proin accumsan, massa viverra cursus pharetra, ipsum nisi lobortis velit, a malesuada dolor lorem eu neque.

%----------------------------------------------------------------------------------------
%	SECTION 2
%----------------------------------------------------------------------------------------

\section{Main Section 2}

Sed ullamcorper quam eu nisl interdum at interdum enim egestas. Aliquam placerat justo sed lectus lobortis ut porta nisl porttitor. Vestibulum mi dolor, lacinia molestie gravida at, tempus vitae ligula. Donec eget quam sapien, in viverra eros. Donec pellentesque justo a massa fringilla non vestibulum metus vestibulum. Vestibulum in orci quis felis tempor lacinia. Vivamus ornare ultrices facilisis. Ut hendrerit volutpat vulputate. Morbi condimentum venenatis augue, id porta ipsum vulputate in. Curabitur luctus tempus justo. Vestibulum risus lectus, adipiscing nec condimentum quis, condimentum nec nisl. Aliquam dictum sagittis velit sed iaculis. Morbi tristique augue sit amet nulla pulvinar id facilisis ligula mollis. Nam elit libero, tincidunt ut aliquam at, molestie in quam. Aenean rhoncus vehicula hendrerit.

### 2013-11-01 Chapters/old_classification/Chapter1.tex
Prefix refs with non-breaking space and capitalize refs

% !TEX root = ../main.tex
% Chapter 1
%----------------------------------------------------------------------------------------

%perhaps a shortened title
*** [GRAPHICS diagram: sensors --> signal pre-processing --> feature extraction --> analysis [* segmentation, * labeling] --> conclusions]
*** Give context problems, means to reach (e.g. monitoring of patients,  getting conclusions over health and activity).
\chapter{Introduction} % Main chapter title
\label{Chapter1} % For referencing the chapter elsewhere, use~\ref{Chapter1}
\lhead{Chapter 1. \emph{Introduction}} % This is for the header on each page -

\section{Literature review}
This section will provide a literature review per coherent subject of this thesis.

% \input{./Chapters/Chapter1/literature_review_nonparametric_clustering}
\input{./Chapters/Chapter1/lit_rev_segm_and_class}
\input{./Chapters/Chapter1/lit_rev_temp_segm}
\input{./Chapters/Chapter1/paper_list}

\section{Problem statement}

% \section{System architecture}
% In subSection~\ref{sec:lit_review_segmentation} the research in segmentation and classification of human activities will be discussed.
% SubSection~\ref{sec:lit_review_nonparametric} will take a look on applied non-parametric methods in clustering (time series) of data.

### 2013-11-01 Chapters/old_classification/Chapter2/Section_signals.tex
Prefix refs with non-breaking space and capitalize refs

--- 2013-02-08 Chapters/Chapter2/Section_signals.tex
@@
     This differs from mere multisensor integration, in which the controlling system is responsible for processing and interpreting the data from different sources.
-    Figure \ref{fig:sensor_fusion} shows the difference.
+    Figure~\ref{fig:sensor_fusion} shows the difference.
 
@@
       \centering
-        \includegraphics{./Figures/graphs/sensor_fusion.png}
+        \includegraphics{./Figures/sensor_fusion.png}
         % \rule{35em}{0.5pt}
@@
       \item \textbf{Intermediate-level:} Also known as \emph{feature fusion} combines the features which can be extracted from the data.
-      See also section \ref{sec:feature_extraction}.
-      This data would be applicable to segmentation **[???].
+      See also Section~\ref{sec:feature_extraction}.
+      This data would be applicable to segmentation ***[???].
       \item \textbf{High-level:} Fusion of this level will support decision making, optionally followed by actions taken by the system to the environment or conclusion being drawn from the environment.

### 2013-11-01 Chapters/Chapter1.tex
Add notes from gerard

@@
 -- Overall todo's --
-\TODO{The term \emph{emperical data/distruction} is nowhere used in the thesis. Is missing.}
+\begin{itemize}
+  \item \TODO{The term \emph{emperical data/distruction} is nowhere used in the thesis. or: Emperical cumulative distribution. Especially Chapter 2.}
+  \item \TODO{Check for ``tang-constructions'': saying to much in a single sentence.}
+  \item \TODO{Overall: meer focus op de \emph{waarom} vraag/antwoord bij keuzes. Bouwt aan motivatie en context van het geheel.}
+  \item \TODO{Nu staat nog veel (met name Chapter 2) in ``aantekening'' style; ombouwen naar uitleggen, lezer begeleiden, duiden, verlaren etc.}
+  \item \TODO{Test moet condenser: bepalen wat er uit gelaten mag worden.}
+\end{itemize}

### 2013-11-01 Chapters/Chapter2/cusum.tex
Added the TODO command
Add notes from gerard

@@
 
-% *** read book Basseville \cite{basseville1993detection} ***
+% \TODO{read book Basseville \cite{basseville1993detection}}
 
@@
 Originally used for quality control in production environments, its main function is to detect change in the mean of measurements and has been applied to this problem \cite{basseville1993detection}.
-It is an non-Bayesian method and thus makes no assumptions (or: prior belief distributions) for the change points.
+It is a non-Bayesian method and thus makes no assumptions, in the form of prior belief distributions, for the change points.
 Many extensions to this method have been proposed.
@@
 Using the \gls{icss} algorithm they are able to find multiple change points in a reflective manner.
+\TODO{Explain ``reflective'', or use term as batch, post-processing etc.}
 Let $C_k = \sum_{i=1}^k \alpha_t^2$ be the cumulative sum of squares for a series of uncorrelated random variables $\{\alpha_t\}$ of length $T$.
-The centered (and normalized) sum is squares is defined as
+The centered (and normalized) sum of squares is defined as
 \begin{equation}
@@
 In case of a sudden change, the value will increase and exceed some predefined boundary with high probability.
-The behavior of $D_k$ is related a Brownian bridge.
 By using an iterative algorithm, the method is able to minimize the masking effect of successive change points.

### 2013-11-01 Chapters/Chapter2/density_ratio.tex
Add notes from gerard

@@
 
-Many approaches to detect change points, regarded as a change in the underlying probabilistic generation, monitor the logarithm of the likelihood ratio between two consecutive intervals.
+Many approaches to detect change points monitor the logarithm of the likelihood ratio between two consecutive intervals.
+A change is point is regarded to be the moment in time when the underlying probabisitic generation function changes.
 Some methods which rely on this are novelty detection, maximum-likelihood estimation and online learning of autoregressive models \cite{kawahara2009change}.
@@
 
-Improvements in this line of research, by Liu \etal \cite{liu2013change} has led the application of improved density-ratio estimation methods to the problem of change detection.
+Improvements in this line of research by Liu \etal \cite{liu2013change} has led the application of improved density-ratio estimation methods to the problem of change detection.
 Such an improvement is the Unconstrained Least-Squares Importance Fitting (uLSIF) method \cite{kanamori2009least} and an extension which possesses a superior non-parametric convergence property: Relative uLSIF (RuLSIF) \cite{yamada2013relative}.

### 2013-11-01 Chapters/Chapter3/change_detection_time_series.tex
Add notes from gerard

@@
 The methods by Takeuchi and Yamanishi \cite{takeuchi2006unifying} and Camci \cite{camci2010change} use the following formulation for change detection, which we will also use for our problem formulation.
-The algorithm searches for \emph{sudden} changes in the time series data.
+The algorithm searches for sudden changes in the time series data.
 In other words, slowly changing properties in the data are not considered to be changes.
@@
 In \cite{takeuchi2006unifying} the similarity between the stochastic processes are expressed by the \gls{kl} divergence $D(p^2||p^1)$.
-The problem with this measure is that, as the authors conclude and Camci discusses, it is not able to detect a change by decrease in variance.
+It is observed that this measure is not able to detect a change by decrease in variance \cite{takeuchi2006unifying,camci2010change}.
+This problem is the motivation for Camci \cite{camci2010change} to create the \gls{svcpd} algorithm.
 
-The definition of change point being \emph{sudden} changes in the time series data is in line with the search of changes in activities.
+The definition of change point being sudden changes in the time series data is in line with the search of changes in activities.
 Since we are only interested in different activities (which are represented by sudden changes), slight changes within an activity are not of interest.

### 2013-11-01 Chapters/Chapter3/one_class_svm.tex
Fixing typos

@@
 -- Quotes and refs to use --
+``Continuity assumptions: This is a general assumption in pattern recognition: two objects near in feature space should also resemble each other in real life'' \cite{tax2001one}.
+
 $k(\mathbf{x}_i, \mathbf{x})$ is a kernel function computing a dot product in feature space, introduced by Aizerman \etal \cite{aizerman1964theoretical}.
+
 ``Training a \gls{svm} with Gaussian \gls{rbf} kernels corresponds to minimizing the specific cost function with a regularization operator'' \cite{smola1998connection}. \\
@@
 
-Kernels for Support Vector Classifiers where first (?) used by Vapnik \cite{vapnik1998statistical}.
+Kernels for Support Vector Classifiers were first (?) used by Vapnik \cite{vapnik1998statistical}.
 

### 2013-11-01 Chapters/Chapter4/data_gathering.tex
Fixing typos

@@
 In order to provide an objective comparison to other methods, we will use artificial data sets which are also used in the researches on which our method is based.
-The data sets are from Takeuchi and Yamanishi \cite{takeuchi2006unifying} and Camci \cite{camci2010change}.
+These are data sets used by Takeuchi and Yamanishi \cite{takeuchi2006unifying} and Camci \cite{camci2010change}.
 Both construct a collection of one-dimensional time series data according to the second order \gls{ar} model:
@@
 In the second type of data sets we apply our method of change detection and temporal segmentation to real-world data sets.
-Our setup records the activities of humans performed both in and out door in an uncontrolled environment.
-Activities performed include walking, running in both a straight line and a curve, standing, walking up and downstairs and sitting.
-Using the time series data sensor output in the form of acceleration, gravity and rotation metrics, our method gives change points based on those signals.
-By comparing the discovered change points with annotated change points (based on video recordings of the performed activities) we are able to give subjective results.
+Our setup records the activities of humans performed both in- and outdoor in an uncontrolled environment.
+Activities performed include sitting, standing, walking, running in a straight and curve line, and walking up- and downstairs.
+Our method uses the signals from the accelerometer, gravity, and rotation sensors.
+These time series data are used to detect change points.
+A video recording from the performed activity is used to annotate the time series with real change points.
+The discovered change point are compared with these annotated change points to give a subjective quality measure.
 In Chapter~\ref{Chapter6} we give a detailed analysis of the performed activities and the recorded data sets.

### 2013-11-01 Chapters/Chapter4/model_construction.tex
Prefix refs with non-breaking space and capitalize refs
Fixing typos

@@
 After the data is collected and pre-processed (or in the case of the artificial data sets: generated), we construct an online incremental sliding window algorithm.
-We follow the method and implementation introduced by Tax and Laskov \cite{tax2003online}, the \acrlong{isvdd} method.
-This method combines the techniques of online, unsupervised and incremental learning methods with the earlier introduced \gls{oc-svm} algorithm \gls{svdd}.
-The method is first initialized with a window length and then in every step a new data object is added to and the last data object is removed from the working set.
+We follow the \gls{svcpd} method by Camci \cite{camci2010change} and the \gls{isvdd} implementation by Tax and Laskov \cite{tax2003online}.
+That latter algorithm combines the techniques of online, unsupervised and incremental learning methods with the earlier introduced \gls{oc-svm} algorithm \gls{svdd}.
+The method is first initialized with a fixed window length and then in every step a new data object is added to and the oldest data object is removed from the working set.
 
@@
 The \gls{svdd} implementation of this abstract form is set by the parameters $\vectorsym{c}=\operatorname*{diag}(K)$, $\vectorsym{a} = \vectorsym{y}$ and $b=1$.
-The procedure for the incremental version has two operations: adding and removing a data object $k$.
+The procedure for the incremental version has two operations for a data object $k$: add to and remove from the working set.
 When a data object $k$ added, its weight $x_k$ is initially set to $0$.
@@
 Both the operations conclude with the recalculation $\mu$ and the weights $\vectorsym{x}$ for all the objects, in order to obtain the optimal solution for the enlarged or reduced data set.
-The incremental learning algorithm follows from these two operations: new data objects are added to and old data objects are removed from the working set.
+These two operations are the basis for \gls{isvdd}.
+With respect to the timed index, the working set has new data objects are added and old data objects are removed.
 
-The size of the initial window of data objects has a lower bound determined by the hyperparameter $C$ (Equation \ref{eq:svdd_objective}).
+The size of the initial window of data objects has a lower bound determined by the hyperparameter $C$ (Equation~\ref{eq:svdd_objective}).
 Because of the equality constraint $\sum_{i=1}^n a_i x_i = 1$ and the box constraint $0 \le x_i \le C$, the number of objects in the working set must be at least $\ceil{\frac{1}{C}}$.
 Thus the algorithm is initialized by selection the first $\ceil{\frac{1}{C}}$ objects for the working set.
-In every step of the loop of the algorithm at least the same number of objects must be added as there are removed.
-By analyzing the \gls{kkt} conditions, \cite{tax2003online} shows the optimality of the algorithm.
+In every loop of the algorithm the number of elements added are at least equal to the number of elements removed.
+The optimality of the algorithm is proved by analysis of the \gls{kkt} conditions \cite{tax2003online}.
 
-From experiments it shows that the online, \gls{isvdd}, method results in less false alarms than the static \gls{svdd}.
-An explanation for this is that \gls{isvdd} follows the changing data distribution, such that small changes over time, like a drift in mean or increase in frequency, continuously re-model the \gls{svm} representation.
+From experiments it shows that the (online) \gls{isvdd} method results in less false alarms than the static \gls{svdd}.
+An explanation for this is that \gls{isvdd} follows the changing data distribution, \ie small changes over time.
+Such changes are a drift in mean or increase in frequency.
+The \gls{isvdd} algorithm continuously re-models the \gls{svm} representation.
 For our method we use the re-modeling capability of the \gls{isvdd} algorithm to obtain change indication measure.
 The following section will discuss the properties used from the constructed \gls{oc-svm} model at every loop-step of the algorithm.
+
+\TODO{make more explicit that the re-modeling is a strong property for our change-detection method.}

### 2013-11-01 Chapters/Chapter4/model_properties.tex
Fixing typos

@@
 This section shows how we interpret the constructed model, to obtain a measure which can be used for a indication of change points.
-How this obtained measure is used to indicate a change is discussed in the following section.
+The following section discusses how this obtained measure is used to indicate change.
 
@@
 \end{inparaenum}
-Since in \gls{occ} situation typically there are (almost) no examples of outlier objects, Tax and Duin construct a method to generate outliers, based on the assumption that the outliers are uniformly distributed around the target set.
-To minimize the error, calculated by the fractions $f_{T-}$ and $f_{O-}$, we should minize the volume of the target data description (\ie the boundary of \gls{svdd}).
-This is since the the fraction of accepted outliers $f_{O-}$ is an estimate of the volume of the target data description, with respect to the volume of the outlier distribution.
-Tax and Duin provide a method to optimize the parameters of the \gls{svdd} method, being the trade-off parameter $C$ and the \gls{rbf} kernel width $\sigma$.
+Since in \gls{occ} situation typically there are (almost) no examples of outlier objects, Tax and Duin construct a method to generate outliers based on the assumption that the outliers are uniformly distributed around the target set.
+To minimize the error, calculated by the fractions $f_{T-}$ and $f_{O-}$, we should minimize the volume of the target data description (\ie the boundary of \gls{svdd}).
+This is because the fraction of accepted outliers $f_{O-}$ is an estimate of the volume of the target data description, with respect to the volume of the outlier distribution.
+Tax and Duin provide a method to optimize the parameters of the \gls{svdd} method, \ie the trade-off parameter $C$ and the \gls{rbf} kernel width $\sigma$.
 This optimization will result in the modification of the radius $R$ of Equation~\ref{eq:svdd_objective} and affects the Lagrangian inequality~\ref{eq:svdd_inequality}.
 
-\TODO{outliers is kept constant, thus Radius must increase}
+Since in our method the parameters $C$ and $\sigma$ are kept constant, and thereby also the fraction of target objects being rejected, the only free parameter is the radius $R$.
+During the \gls{isvdd} algorithm the volume of the hypersphere is being minimized.
+This means that only the radius of the sphere will change during the algorithm.
+We will interpret changes in the radius $R$ with the same continuity assumptions as discussed in Section~\ref{subsec:kernel_function}.
+This means that when two objects (in this case radius sizes) are close to each other, their original objects are also close.
+In other words, when the time series changes its data objects, the corresponding radius $R$ will also change.
+Since the data objects in the working set become more heterogeneous, the radius of the hypersphere will increase.
+When the data objects change from a heterogeneous set to a more homogeneous, we expect the radius to decrease in value, since the data objects are closer to each other in the feature space.

### 2013-11-04 Chapters/Chapter4.tex
WIP: Chapter 4

@@
 
+\TODO{Puts emphasis on two-stage framework by Takeuchi and Yamanishi \cite{takeuchi2006unifying}.}.
+
+\begin{figure}
+  \centering
+    \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{./Figures/chapter4/method_setup_short.pdf}
+  \caption[Method setup]{Schematic overview of the change detection method.
+The first step is the data gathering, described in Section \ref{sec:method_data_gathering}.
+Ater the pre-processing, the data is used to construct a \gls{svdd} model, as described in Section \ref{sec:method_model_construction}.
+Section \ref{sec:method_model_features} describes which features of the model are used for the final change indication algorithm, discussed in Section \ref{sec:method_change_detection}.}
+  \label{fig:method_overview_short}
+\end{figure}
+
 \input{./Chapters/Chapter4/data_gathering}
 \input{./Chapters/Chapter4/model_construction}
-\input{./Chapters/Chapter4/model_properties}
+\input{./Chapters/Chapter4/model_features}
 \input{./Chapters/Chapter4/change_detection}

### 2013-11-04 Chapters/Chapter4/change_detection.tex
WIP: Chapter 4

@@
 \section{Change Detection}\label{sec:method_change_detection}
+In the following sections we haved discussed what data is used for the model construction how the model is interpreted in the context of change detection.
+In this section we discuss the methods applicable to the extracted radius $R$ of the hypersphere, in order to indicate a change in the underlying distribution properties.
+This employs the second stage of our algorithm, analogue to the second stage of the unifying framework by Takeuchi and Yamanishi~\cite{takeuchi2006unifying}.
+
+In the previous sections we have discussed how a multi-dimensional signal can be reduced to a single dimension time series data based on the radius $R$.
+We have argued that in the case of a change in the underlying data, where it becomes more heterogeneous, we expect an increase in the radius $R$.
+In the same manner we expect that when the change is in the oldest data objects of the working set the radius $R$ will decrease, since the data becomes more homogeneous.
+An abstract representation of these expectations is illustrated in Figure~\ref{fig:radius_expectation}.
+
+\begin{figure}
+  \centering
+    \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{./Figures/chapter4/expected_behaviour.pdf}
+  \caption[Expected radius behaviour]{Expected behaviour of the radius $R$ of the hypersphere.
+The upper part shows a typical sinusoidal time series signal.
+The lower graph visualizes an abstract expectation of the values of $R$.
+Five possible windows are illustrated.
+The left, middle, and right windows cover an area of homogeneous signal.
+The expected value of $R$ is low.
+The other two windows cover a change entering and leaving the window, respectively.
+At these locations the radius $R$ is expected to increase temporarily.
+}
+  \label{fig:radius_expectation}
+\end{figure}
+
+The problem of change detection in complex multi-dimensional time series is now reduced to finding change in a one-dimensional time series.
+Besides the dimensionality reduction, the form of the signal is also simplified.
+Whereas the original signal was represented as a sinusoidal or second order \gls{ar} model, the new time series is of much simpler form.
+In our method we leave open the precise method which is used to find change points in the obtained change indication values.
+Below, we will gives two examples of methods which can be applied.
+The first example is a ratio-based thresholding mechanism used by Camci~\cite{camci2010change}.
+
+\subsection{Ratio-thresholding}
+
+\begin{equation}\label{eq:ratio_radius}
+  \hbar_t = \frac{r_t}{\operatorname*{mean}(r_{y:t-1})},
+\end{equation}
+where $r_t$ is the approximated radius of the hypersphere at time $t$, $y$ is the time of the last change point (or $1$ if there is non) and $\operatorname*{mean}(r_{y:t-1})$ is the average of the previous approximated radii.

### 2013-11-04 Chapters/Chapter4/data_gathering.tex
WIP: Chapter 4

@@
 In order to provide an objective comparison to other methods, we will use artificial data sets which are also used in the researches on which our method is based.
-These are data sets used by Takeuchi and Yamanishi \cite{takeuchi2006unifying} and Camci \cite{camci2010change}.
-Both construct a collection of one-dimensional time series data according to the second order \gls{ar} model:
+These are the data sets used by Takeuchi and Yamanishi \cite{takeuchi2006unifying} and Camci \cite{camci2010change}.
+Both construct a collection of one-dimensional time series data according to a second order \gls{ar} model:
 \begin{equation}
@@
 In the second type of data sets we apply our method of change detection and temporal segmentation to real-world data sets.
-Our setup records the activities of humans performed both in- and outdoor in an uncontrolled environment.
-Activities performed include sitting, standing, walking, running in a straight and curve line, and walking up- and downstairs.
+For our setup we record the activities of humans performed both in- and outdoor in an uncontrolled environment.
+Activities performed include sitting, standing, walking, running in a straight and curved line, and walking up- and downstairs.
 Our method uses the signals from the accelerometer, gravity, and rotation sensors.

### 2013-11-04 Chapters/Chapter4/model_construction.tex
WIP: Chapter 4

@@
 We follow the \gls{svcpd} method by Camci \cite{camci2010change} and the \gls{isvdd} implementation by Tax and Laskov \cite{tax2003online}.
-That latter algorithm combines the techniques of online, unsupervised and incremental learning methods with the earlier introduced \gls{oc-svm} algorithm \gls{svdd}.
-The method is first initialized with a fixed window length and then in every step a new data object is added to and the oldest data object is removed from the working set.
+That latter algorithm combines the techniques of online, unsupervised, and incremental learning methods with the earlier introduced \gls{oc-svm} algorithm \gls{svdd}.
+The method is initialized with a fixed window length and then in every step a new data object is added to and the oldest data object is removed from the working set.
 
@@
 In case of an object removal, the weight is forced to be $x_k=0$.
-Both the operations conclude with the recalculation $\mu$ and the weights $\vectorsym{x}$ for all the objects, in order to obtain the optimal solution for the enlarged or reduced data set.
+Both the operations conclude with the recalculation  of $\mu$ and the weights $\vectorsym{x}$ for all the objects, in order to obtain the optimal solution for the enlarged or reduced data set.
 These two operations are the basis for \gls{isvdd}.
-With respect to the timed index, the working set has new data objects are added and old data objects are removed.
+With respect to the timed index, new data objects are added to and old data objects are removed from the working set.
 

### 2013-11-04 Chapters/Chapter4/model_features.tex
WIP: Chapter 4

--- 2013-11-01 Chapters/Chapter4/model_properties.tex
@@
 % !TEX root = ../../main.tex
-\section{Model Properties}\label{sec:method_model_properties}
+\section{Model Features}\label{sec:method_model_features}
 
@@
 In the previous section we have discussed the \gls{isvdd} method, which creates a \gls{oc-svm} representation of a working set of data objects at every step of the algorithms loop.
-This section shows how we interpret the constructed model, to obtain a measure which can be used for a indication of change points.
+This section shows how we interpret the constructed model and extract features to obtain a measure which can be used for a indication of change points.
 The following section discusses how this obtained measure is used to indicate change.
@@
 Since in \gls{occ} situation typically there are (almost) no examples of outlier objects, Tax and Duin construct a method to generate outliers based on the assumption that the outliers are uniformly distributed around the target set.
-To minimize the error, calculated by the fractions $f_{T-}$ and $f_{O-}$, we should minimize the volume of the target data description (\ie the boundary of \gls{svdd}).
-This is because the fraction of accepted outliers $f_{O-}$ is an estimate of the volume of the target data description, with respect to the volume of the outlier distribution.
+To minimize the error, calculated by the fractions $f_{T-}$ and $f_{O+}$, we should minimize the volume of the target data description (\ie the boundary of \gls{svdd}).
+This is because the fraction of accepted outliers $f_{O+}$ is an estimate of the volume of the target data description, with respect to the volume of the outlier distribution.
 Tax and Duin provide a method to optimize the parameters of the \gls{svdd} method, \ie the trade-off parameter $C$ and the \gls{rbf} kernel width $\sigma$.
@@
 This means that when two objects (in this case radius sizes) are close to each other, their original objects are also close.
-In other words, when the time series changes its data objects, the corresponding radius $R$ will also change.
-Since the data objects in the working set become more heterogeneous, the radius of the hypersphere will increase.
-When the data objects change from a heterogeneous set to a more homogeneous, we expect the radius to decrease in value, since the data objects are closer to each other in the feature space.
+In other words, when the underlying distribution of time series is changed, and thus the characteristics of the data objects changes, the corresponding radius $R$ will also change.
+Since in that case the data objects in the working set become more heterogeneous, the radius of the hypersphere will increase.
+When, instead, the data objects change from a heterogeneous set to a more homogeneous, we expect the radius to decrease in value, since the data objects are closer to each other in the feature space.
+The following section discusses the algorithms which can be applied to the extracted radius $R$ as a volume estimate.
+
+\TODO{Put emphasis on reduction of multi-dimensions to single dimension. Also stage 1 and 2 from unifying framework \cite{takeuchi2006unifying}.}

### 2013-11-04 glossary-acronyms.tex
WIP
Working on One-Class Classification texts
More on OCC
WIP: Chapter 4
Better formulas on SVDD
Wip: chapter 3 formulas, chapter 4 model construction and properties

@@
 % A C R O N Y M S
+
+\newacronym{ar}{AR}{Autoregressive}
 
@@
 
+\newacronym{har}{HAR}{Human Activity Recognition}
+
+\newacronym{gmm}{GMM}{Gaussian Mixture Model}
+
 \newacronym{hmm}{HMM}{Hidden Markov Model}
@@
 
+\newacronym{isvdd}{I-SVDD}{Incremental Support Vector Data Description}
+
+\newacronym{kl}{KL}{Kullback-Leibler}
+
 \newacronym{kliep}{KLIEP}{Kullback-Leibler Importance Estimation Procedure}
+
+\newacronym{kkt}{KKT}{Karush–-Kuhn–-Tucker}
 
@@
 
-\newacronym{occ}{OCC}{One-Class Classifier}
+\newacronym{occ}{OCC}{One-Class Classification}
 
 \newacronym{oc-svm}{OC-SVM}{One-Class Support Vector Machine}
+
+\newacronym{pca}{PCA}{Principal Component Analysis}
+
+\newacronym{pdf}{PDF}{Probability Density Function}
+
+\newacronym{qp}{QP}{Quadratic Programming}
 
@@
 
-\newacronym{pca}{PCA}{Principal Component Analysis}
-
 \newacronym{rt}{RT}{Ratio-Thresholding}
-
-\newacronym{pdf}{PDF}{Probability Density Function}
-
-\newacronym{qp}{QP}{Quadratic Programming}
 
@@
 
-\newacronym{svcpd}{SVCPD}{Support Vector based Change Detection}
+\newacronym{svcpd}{SVCPD}{Support Vector based Change Point Detection}
 
@@
 
-\newacronym{swab}{SWAB}{Sliding Window And Buttom-up}
+\newacronym{swab}{SWAB}{Sliding Window And Bottom-up}

### 2013-11-05 Chapters/Chapter4/change_detection.tex
WIP: radius ratio thresholding

@@
     \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{./Figures/chapter4/expected_behaviour.pdf}
-  \caption[Expected radius behaviour]{Expected behaviour of the radius $R$ of the hypersphere.
+  \caption[Expected radius behavior]{Expected behavior of the radius $R$ of the hypersphere.
 The upper part shows a typical sinusoidal time series signal.
@@
 \subsection{Ratio-thresholding}
-
+The first example of an algorithm that interprets the change indication time series and transforms it to change detection is the radius ratio thresholding method used by Camci~\cite{camci2010change}.
+The method calculates the radius ratio $\hbar_t$ by taking the ratio of the current radius $r_t$ at time $t$ to the average radii since the last change point $y$:
 \begin{equation}\label{eq:ratio_radius}
@@
 \end{equation}
-where $r_t$ is the approximated radius of the hypersphere at time $t$, $y$ is the time of the last change point (or $1$ if there is non) and $\operatorname*{mean}(r_{y:t-1})$ is the average of the previous approximated radii.
+where $\operatorname*{mean}(r_{y:t-1})$ is the average of the previous approximated radii.
+The ratio $\hbar_t$ is compared with two thresholds: the low threshold $th_\text{low}$ and the high threshold $th_\text{high}$.
+When the data objects in the working set becomes more homogeneous at time $t$, the radius $R$ of the hypersphere and thereby the ratio $\hbar_t$ will decrease.
+When the new value of $\hbar_r$ is lower than the low threshold $th_\text{low}$, a change detection is observed for time $t$.
+The same holds for a more heterogeneous set of data and an increase of $\hbar_r$ higher than $th_\text{high}$.
+
+The appropriate values for the thresholds are highly depended on the characteristics of the data and the trade-off parameter $C$ (as discussed in Section~\ref{subsec:oc-svm-svdd}).
+This trade-off parameters regulates the fraction of data objects that will be rejected from the constructed model.
+Since a high value of $C$ assigns a high penalty to outliers, data objects are more likely to be incorporated into the hypersphere by increasing the radius.
+For low values of $C$ the cost of rejecting data objects is, compared to the benefits of a smaller hypersphere, relatively low.
+This gives a relation between the parameters $C$, $th_\text{low}$, and $th_\text{high}$ and the sensitivity of the radius ratio based thresholding method.
+A high value of $C$, or values close to $1$ for $th_\text{low}$ and $th_\text{high}$ result in a sensitive change detection procedure.
+The reverse results in less sensitive methods.
+The correct values for the intended sensitivity need to be empirically determined.

### 2013-11-05 Chapters/Chapter4/change_detection.tex
Add note on historic data
WIP: chapter 4

@@
 The other two windows cover a change entering and leaving the window, respectively.
-At these locations the radius $R$ is expected to increase temporarily.
-}
+At these locations the radius $R$ is expected to increase temporarily.}
   \label{fig:radius_expectation}
@@
 Below, we will gives two examples of methods which can be applied.
-The first example is a ratio-based thresholding mechanism used by Camci~\cite{camci2010change}.
+The first is a family of methods based on the \gls{cusum} method.
+The second example is a ratio-based thresholding mechanism used by Camci~\cite{camci2010change}.
+We conclude this section by discussing some post-processing techniques we have used, to decrease the number of false positives.
+
+\subsection{CUSUM based methods}
+Many of the simple change detection methods are based on \gls{cusum}, originally introduced by Page~\cite{page1954continuous}.
+Many variations and extensions of the method are proposed and used (\cite{inclan1994use,alippi2006adaptive,hsu2007mosum}).
+Here we will discuss the simple form, which only detects changes for an increase of the examples, in this case the radius $R$.
+The cumulative sum for all the values is calculated:
+\begin{equation}
+  S_n = \sum_{k=1}^n r_k
+\end{equation}
+and a change is detected when the value of the cumulative sums minus the minimum encountered exceeds a predetermined threshold $h$:
+\begin{equation}
+  S_n - \operatorname*{min}_{0 \le i < n} S_i \ge h.
+\end{equation}
+
+Extensions of the \gls{cusum} method have been proposed.
+Amongst others, extensions which extend its use to two-threshold methods, and methods appropiate for changes in mean or variance \cite{inclan1994use}.
+Whilst adaptive methods have been proposed \cite{alippi2006adaptive}, most methods require manual determination of the threshold parameter $h$.
 
 \subsection{Ratio-thresholding}
-The first example of an algorithm that interprets the change indication time series and transforms it to change detection is the radius ratio thresholding method used by Camci~\cite{camci2010change}.
+The second example of an algorithm that interprets the change indication time series and transforms it to change detection is the radius ratio thresholding method used by Camci~\cite{camci2010change}.
+The method shares characteristics with \gls{cusum}, since it relies on summation of historic data and thresholds for change detection.
 The method calculates the radius ratio $\hbar_t$ by taking the ratio of the current radius $r_t$ at time $t$ to the average radii since the last change point $y$:
@@
 where $\operatorname*{mean}(r_{y:t-1})$ is the average of the previous approximated radii.
+By using only historic data, this method can be incorporated in an online change detection method.
 The ratio $\hbar_t$ is compared with two thresholds: the low threshold $th_\text{low}$ and the high threshold $th_\text{high}$.
@@
 The correct values for the intended sensitivity need to be empirically determined.
+
+\subsection{Post-processing}
+In our method we use the ratio based thresholding, since it is also used by Camci~\cite{camci2010change}.
+From preliminary experiments we observed that a single change in the data can cause many detected change points, using the ratio based method.
+This phenomenon can be explained by the nature of the thresholding method.
+Since a single (high or low) threshold is set for ratio of the radii, high values are considered to be a change point before the highest point of the peak.
+As illustrated in Figure~\ref{fig:tresholding}, the peak at $t_1$ represents a single change point.
+The region from $t_2$ to $t_3$ is strictly increasing.
+At $t_2$ a change point is detected, since $\hbar_t$ exceeds the threshold.
+But since it is increasing up to $t_3$, all values of $\hbar_t$ in that period will trigger the ratio based method to indicate a change point.
+
+To overcome this problem, we apply a post-processing method on the generated change points.
+All the change points that are within a time period $\delta$ of each other are merged together.
+\TODO{Merged to the front, or to the back, depending on high/low threshold exceeding?}
+
+\begin{figure}
+  \centering
+    \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{./Figures/chapter4/signal_threshold.pdf}
+  \caption[Thresholding]{Abstract illustration of the radius ratio thresholding.
+The peak at $t_1$ generates a single change point.
+The region between $t_2$ and $t_3$ generates multiple change points, since the value of $\hbar_t$ keeps increasing.}
+  \label{fig:tresholding}
+\end{figure}

### 2013-11-05 Chapters/Chapter4/model_features.tex
WIP: chapter 4

@@
 \section{Model Features}\label{sec:method_model_features}
-
--- Notes --
-\begin{itemize}
-  \item Error measure of \gls{svdd} is based on both volume of the sphere and number of outliers.
-  \item When the outliers are uniformly distributed around the target data, we should minimize the volume of the description.
-  \item To do this (minimize the volume) we need to estimate the volume.
-  \item \cite{tax2002uniform} states that when the outliers are uniformly distributed, the fraction of outliers that is accepted gives an estimate of the volume of the target data description (with respect to the volume of the outlier distribution).
-  \item When the fraction of outlier is kept constant (by the parameters $C$ and $\sigma$), and when it is known that more outliers are present in current working set, the volume must increase.
-  \item Equation~\ref{eq:svdd_inequality} gives that the right term must increase for a larger volume, \ie the value of $R$ will increase.
-  \item Thus, an increase of value $R$ indicates an increase of outliers in the working set.
-  \item This indicates a possible new activity in the time series data.
-\end{itemize}
-
 In the previous section we have discussed the \gls{isvdd} method, which creates a \gls{oc-svm} representation of a working set of data objects at every step of the algorithms loop.
@@
 Tax and Duin provide a method to optimize the parameters of the \gls{svdd} method, \ie the trade-off parameter $C$ and the \gls{rbf} kernel width $\sigma$.
-This optimization will result in the modification of the radius $R$ of Equation~\ref{eq:svdd_objective} and affects the Lagrangian inequality~\ref{eq:svdd_inequality}.
+This optimization will result in the modification of the radius $R$ of Equation~\ref{eq:svdd_objective} and affects the Lagrangian inequality~(\ref{eq:svdd_inequality}).
 
@@
 When, instead, the data objects change from a heterogeneous set to a more homogeneous, we expect the radius to decrease in value, since the data objects are closer to each other in the feature space.
+
+With the \gls{isvdd} algorithm we have effectively implemented a form of dimensionality reduction by feature extraction, using the radius $R$.
 The following section discusses the algorithms which can be applied to the extracted radius $R$ as a volume estimate.
+We thereby follow the setup of the unifying framework by Takeuchi and Yamanishi~\cite{takeuchi2006unifying}, of which this section described the first stage.
 
-\TODO{Put emphasis on reduction of multi-dimensions to single dimension. Also stage 1 and 2 from unifying framework \cite{takeuchi2006unifying}.}
+\TODO{Notion of dimensionality reduction is a bit vague here, because it applies to a window of objects.}

### 2013-11-06 Chapters/Chapter1.tex
WIP

@@
 \end{itemize}
+
+\begin{figure}
+  \centering
+    \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{./Figures/chapter1/thesis_goal.pdf}
+  \caption[Thesis goal]{The goal of this thesis.
+The scope of this thesis is Temporal Segmentation, which can be useful in the context of Temporal Classification.
+Given a data set, we apply the construction of \gls{oc-svm} models as a form of dimensionality reduction.
+Using the reduced representation, being the radius $R$ of the constructed hypersphere, we apply direct change detection methods.
+The detected change points can support the classification of homogeneous segments of data.}
+  \label{fig:thesis_goal}
+\end{figure}

### 2013-11-06 Chapters/Chapter3/one_class_svm.tex
WIP

@@
 \subsection{\acrlong{nu-svm}}\label{subsec:nu-svm}
+\TODO{Check notations with book of Scholkopf \cite{scholkopf2002learning}, chapter 8, page 227.} \\
 The first of the \gls{oc-svm} methods we will discuss is often referred to as \gls{nu-svm} and introduced by Sch\"olkopf \etal \cite{scholkopf1999support}.
@@
 
-\TODO{Note that \gls{svdd} and \gls{nu-svm} have identical solutions when data is preprocessed to have unit form (in the case of \gls{nu-svm}) \cite{tax2002uniform}.}
+\TODO{Note that \gls{svdd} and \gls{nu-svm} have identical solutions when data is preprocessed to have unit form (in the case of \gls{nu-svm}) \cite{tax2002uniform}.
+and the \gls{rbf} kernel is used.
+Page 234 of \cite{scholkopf2002learning}.}
 
@@
 
-Kernels for Support Vector Classifiers were first (?) used by Vapnik \cite{vapnik1998statistical}.
+Kernels for Support Vector Classifiers were first (?) used by Vapnik \cite{vapnik1998statistical}. \\
+
+Use Chapter 2 from ``Learning with kernels'', \cite{scholkopf2002learning}.
 

### 2013-11-06 Chapters/Chapter4/model_construction.tex
WIP

@@
 The \gls{isvdd} algorithm continuously re-models the \gls{svm} representation.
-For our method we use the re-modeling capability of the \gls{isvdd} algorithm to obtain change indication measure.
-The following section will discuss the properties used from the constructed \gls{oc-svm} model at every loop-step of the algorithm.
-
-\TODO{make more explicit that the re-modeling is a strong property for our change-detection method.}
+This re-modeling of the \gls{svm} representation is beneficial for our purpose, since we are only interested in sudden changes between activities.
+The following section will discuss the features extracted from the constructed \gls{oc-svm} model at every loop-step of the algorithm, before Section~\ref{sec:method_change_detection} describes how these features are used to detect change.

### 2013-11-07 Chapters/Chapter2.tex
WIP: Working on organization of Chapter 2

@@
 
-\TODO{Maybe add dimensionality reduction, but for now leave out}
+-- Nieuwe indeling Chapter 2 --
+
+\begin{description}
+  \item[Intro/outline] Introduction and overview of this chapter. Put in the context of temporal classification.
+  \item[Temporal segmentation] Give overview of used temporal segmentation methods, move towards a split in model building/dimensionality reduction and change detection methods
+  \item[Dimensionality reduction] Explain that \gls{dr} can be useful in data processing. Name a few techniques, make bridge to \gls{oc-svm}.
+  \item[Change detection] Describe methods for \emph{direct} change detection. End with notion
+    \begin{description}
+      \item[Model fitting]
+      \item[Cusum]
+      \item[Density ratio]
+    \end{description}
+  \item[Support Vector Machine based methods]
+
+\end{description}
+
+\section{Overview}\label{sec:literature_review_overview}
+As described in Chapter~\ref{Chapter1}, the focus of this research is the application of temporal segmentation to accelerometer data.
+This is with the assumption that explicit temporal segmentation can aid the process of temporal classification.
+
+\TODO{Give examples of temporal classification techniques}
+
+In recent years a lot is research is performed in the classification of human activities, recorded by on-body sensors.
+Especially since smartphones with accelerometer sensors became more available, research has focussed on recordings from these devices.
+
+\TODO{Include highlights from old-classification text}
+
+In the following section of this is chapter we begin with a review of current literature on temporal segmentation.
+Many techniques of temporal segmentation rely in change detection algorithms.
+In Section~\ref{sec:change_detection_methods} we will give an overview of those kind of algorithms.
+An often encountered problem is the \emph{curse of dimensionality}.
+To overcome this problem, we look at \gls{dr} methods in Section~\ref{sec:dimensionality_reduction}.
+We end this chapter with a review on \gls{svm} based methods and algorithms.
+This latter section relates the \gls{oc-svm} methods via \gls{dr} to temporal segmentation.
+
+\input{./Chapters/Chapter2/temporal_segmentation.tex}
+\input{./Chapters/Chapter2/change_detection_methods.tex}
+\input{./Chapters/Chapter2/dimensionality_reduction.tex}
+\input{./Chapters/Chapter2/svm_based.tex}
+
+-- OLD --
 
@@
 % Temporal Segmentation
-\input{./Chapters/Chapter2/temporal_segmentation.tex}
+\input{./Chapters/Chapter2/temporal_segmentation_old.tex}
 
@@
 \input{./Chapters/Chapter2/svm}
-
-%----------------------------------------------------------------------------------------
-% Dimensionality reduciton
-% \input{./Chapters/Chapter2/dim_reduction}

### 2013-11-07 Chapters/Chapter2/temporal_segmentation.tex
WIP: Working on organization of Chapter 2

--- 2013-10-24 Chapters/Chapter2/temporal_segmentation.tex
@@
 
-\begin{itemize}
-  \item Given overview of segmentation techniques, for times series data
-  \item Use different ``point-of-views'', or terms
-  \item ``Segmentation''
-  \item ``Change detection''
-  \item ``Novelty detection''
-  \item Specific view on \glspl{svm}
-\end{itemize}
-
-This section gives an overview of the literature on temporal segmentation in the context of \gls{har}.
-It takes a look on different implementations and methodologies.
-A wide range of terms and subtle differences are used in the field, such as `segmentation', `change detection', `novelty detection' and `outlier detection'.
-These will be the categorical terms for which we discuss the literature.
-Finally we will discuss other applications of \glspl{svm} in the context of these terms.
-
-\subsection{Segmentation}\label{subsec:segmentation}
-\TODO{This subsection is mainly from previous draft version}
-
-\TODO{Create compact notation, one sentence per paper max}
-
-Segmentation methods can roughly be categorized in three methods in the way the data is processed, as discussed by Avci \etal \cite{avci2010activity}:
-\begin{itemize}
-  \item \textbf{Top-down} methods iteratively divide the signal in segments by splitting at the best location.
-  The algorithm starts with two segments and completes when a certain condition is met, such as when an error value or number of segments $k$ is reached.
-  These methods process the data points recursively, which results in a complexity of $O(n^2k)$.
-  \item \textbf{Bottom-up} methods are the natural complement to top-down methods.
-  They start with creating $n/2$ segments and iteratively join adjacent segments while the value of a cost function for that operation is below a certain value.
-  Given the average segment length $L$, the complexity of this method is $O(nL)$.
-  \item \textbf{Sliding-window} methods are simple and intuitive for online segmenting purposes.
-  It starts with a small initial subsequence of the time series.
-  New data points are joined in the segment until the fit-error is above a threshold.
-  Since the data is only processed very locally, these methods can yield in poor results \cite{keogh2001online}.
-  The complexity is equal to the bottom-up approach, $O(nL)$, where $L$ is the average segment length.
-  \item \textbf{\acrlong{swab}}, as introduced by Keogh \etal \cite{keogh2001online}, joins the ability of the sliding window mechanism to process time series online and the bottom-up approach the create superior segments in terms of fit-error.
-  The algorithm processes the data in two stages.
-  The first stage is to join new data points in the current segment created by a sliding window, and pass this to a buffer with space for a few segments.
-  The buffer then processes the data Bottom-up and returns the first (left-most) segment as final segment.
-  Because this second stage retains some (semi-)global view of the data, the results are comparative with normal Bottom-up.
-  It is stated by Keogh \etal that the complexity of \gls{swab} is a small constant factor worse than that of regular Bottom-up.
-\end{itemize}
-It is clear that for the application of this research sliding-window and preferably \gls{swab}-based algorithms should be considered.
-
-The \gls{swab} method proposed by Keogh \etal \cite{keogh2001online} is dependent on an user setting, providing the maximum error when performing both stages.
-Each segment is approximated by using piecewise linear representation (PLR), an often used method.
-The user provided error threshold controls the granularity and number of segments.
-Other methods have been proposed, such as an adaptive threshold based on the signal energy by Guenterberg \etal \cite{guenterberg2009automatic}, the adaptive \gls{cusum}-based test by Alippi \etal \cite{alippi2006adaptive} and the \gls{mosum} by Hsu \cite{hsu2007mosum} in order to eliminate this user-dependency.
-The latter of these methods is able to process the accelerometer values directly, although better results are obtained when features of the signal are processed, as done in the former method.
-Here the signal energy, mean and standard deviation are used to segment activities and by adding all the axial time series together, the Signal-To-Noise ration is increased, resulting in a robuster method.
-
-The method of Guenterberg \etal extracts features from the raw sensor signal to base the segmentation on other properties than the pure values.
-The method of Bernecker \etal \cite{bernecker2012activity} uses other statistical properties, namely autocorrelation, to distinguish periodic from non-periodic segments.
-Using the \gls{swab} method the self-similarity of a one-dimensional signal is obtained.
-The authors claim that only a slight modification is needed to perform the method on multi-dimensional data.
-After the segmentation phase, the method of Bernecker \etal extracts other statistical features which are used in the classification phase.
-
-The proposal of Guo \etal \cite{guo2012adaptive} dynamically determines which features should be used for the segmentation and simultaneously determines the best model to fit the segment.
-For each of the three dimensions features such as the mean, variance, covariance, correlation, energy and entropy are calculated.
-By extending the \gls{swab} method, for every frame a feature set is selected, using an enhanced version of \gls{pca}.
-The research also considered the (Stepwise) Feasable Space Window as introduced by \cite{liu2008novel}, but since it results in a higher error rate than \gls{swab} the latter was chosen to extend.
-Whereas the before mentioned algorithms use a linear representation, this methods considers linear, quadratic and cubical representations for each segment.
-This differs from other methods where the model is fixed for the whole time series, such as \cite{fuchs2010online}, which is stated to perform inferior on non-stationary time series such as daily life.
-
-The time series data from a sensor can be considered as being drawn from a certain stochastic process.
-Probabilistic models can be constructed on that signal, yielding in probabilistic and Bayesian based segmentation methods.
-The \gls{cusum}-methods takes a statistical approach and relies on the log-likelihood ratio \cite{gustafsson1996marginalized} to measure the difference between two distributions.
-To calculate the ratio, the probability density functions need to be calculated.
-The method of Kawahara \etal \cite{kawahara2009change} proposes to estimate the ratio of probability densities (known as the \emph{importance}), based on the log likelihood of test samples, directly, without explicit estimation of the densities.
-The method by Liu \etal \cite{liu2013change} uses a comparable dissimilarity measure using the \gls{kliep} algorithm.
-They claim this results in a robuster approach for real-world scenarios.
-Although this is a model-based method, no strong assumptions (parameter settings) are made on the models.
-
-The method of Adams and MacKay \cite{adams2007bayesian} builds a probabilistic model on the segment run length, given the observed data so far.
-Instead of modeling the values of the data points as a probabilistic distribution, the length of segments as a function if time is modeled by calculating the posterior probability.
-It uses a prior estimate for the run length and a predictive distribution for newly-observed data, given the data since the last change point.
-This method contrasts with the approach of Guralnik and Srivastava \cite{guralnik1999event} in which change points are detected by a change in the (parameters of an) underlying, observed, model.
-For each new data point, the likelihoods of being a change point and part of the current segment are calculated, without a prior model (and thus is a non-Bayesian approach).
-It is observed that when no change point is detected for a long period of time, the computational complexity increases significantly.
-
-Another application of \gls{pca} is to characterize the data by determining the dimensionality of a sequence of data points.
-The proposed method of Berbi\v{c} \etal \cite{barbivc2004segmenting} determines the number of dimensions (features) needed to approximate a sequence within a specified error.
-With the observation that more dimensions are needed to keep the error below the threshold when transitions between actions occur, cut-points can be located and segments will be created.
-The superior extension of their approach uses a Probabilistic \gls{pca} algorithm to incorporate the dimensions outside the selected set as noise.
-
-In the method by Himberg \etal \cite{himberg2001time} a cost function is defined over segments of data which is to be minimized.
-The cost functions thereby searches for internally homogeneous segments of data, reflecting states in which the devices and the user are.
-The cost function can be any arbitrary function and in the implementation the sum of variances over the segments is used.
-Both in a local and global iterative replacement procedure (as an alternative for the computationally hard dynamic programming algorithm) the best breakpoint locations $c_i$ for a pre-defined number of segments $1 \leq i \leq k$ are optimized.
-
-Many methods obtain an implicit segmentation as a result of classification over a sliding window \TODO{add refs}.
-The method of Yang \etal \cite{yang2008distributed} explicitly performs segmentation and classification simultaneously.
-It argues that the classification of a pre-segmented test-sequences becomes straightforward with many classical algorithms to choose from.
-The algorithm matches test examples with the \emph{sparsest} linear representation of mixture subspace models of training examples, searching over different temporal resolutions.
-
-The method of Chamroukhi \etal \cite{chamroukhi2013joint} is based on a \gls{hmm} and logistic regression.
-It assumes a $K$-state hidden process with a (hidden) state sequence, each state providing the parameters (amongst which the order) for a polynomial.
-The order of the model segment is determined by model selecting, often using the \gls{bic} or the similar \gls{aic} \cite{akaike1974new}, as in \cite{he2008activity}.
-
-Field of computer vision: \cite{zhou2008aligned}, \cite{li2007segmentation}.
-
---- Segmentation ---
-
-% ``An online algorithm for segmenting time series'' \cite{keogh2001online}. See \ref{sec:appendix-C-online-keogh}. 538, 2001 \\
-% ``Segmenting time series: A survey and novel approach'' \cite{keogh2004segmenting}. 242, 2004 \\
-
-% ``Time series segmentation for context recognition in mobile devices'' \cite{himberg2001time}. 158, 2001 \\
-
-% ``An Adaptive Approach for Online Segmentation of Multi-Dimensional Mobile Data'' \cite{guo2012adaptive}. 4, 2012 \\
-
-% ``Joint segmentation of multivariate time series with hidden process regression for human activity recognition'' \cite{chamroukhi2013joint}. 0, 2013. See \ref{sec:appendix-C-joint-segmentation}. \\
-
-``Segmentation and Recognition of Motion Streams by Similarity Search'' \cite{li2007segmentation}. 29, 2007 \\
-
-% ``Novel Online Methods for Time Series Segmentation'' \cite{liu2008novel}. 22, 2008 \\
-
-% ``Distributed Segmentation and Classification of Human Actions Using a Wearable Motion Sensor Network'' \cite{yang2008distributed}. 44, 2008 \\
-
-% ``An Automatic Segmentation Technique in Body Sensor Networks based on Signal Energy'' \cite{guenterberg2009automatic}. 14, 2009 \\
-
-% ``Online Segmentation of Time Series Based on Polynomial Least-Squares Approximations'' \cite{fuchs2010online}. 24, 2010 \\
-
-``Aligned Cluster Analysis for Temporal Segmentation of Human Motion'' \cite{zhou2008aligned}. 63, 2008 \\
-
-% ``Segmenting motion capture data into distinct behaviors'' \cite{barbivc2004segmenting}. 270, 2004 \\
-
-\subsection{Change detection}\label{subsec:change_detection}
-\TODO{change order of this and temporal segmentation sections? So first change, then segmentation?} \\
-Whereas the above mentioned researches focus on \emph{segmentation}, many have focused on \emph{change detection}.
-Although these techniques are closely related, there is a subtle difference.
-In the case of \emph{change detection} to goal is to find, possibly unrelated, sudden change points in a signal \cite{takeuchi2006unifying}.
-In contrast, the goal of \emph{temporal segmentation} is to find homogeneous segments of data, which can be the result of multiple detected changes.
-
-The \gls{icss} by Incl\'{a}n and Tiao \cite{inclan1994use} is a statistical method which obtains results (when applied to stock data) comparable to \gls{mle} and Bayesian \TODO{Bayesian What?}.
-Whereas \gls{cusum} can be applied to search for a change in mean, the \gls{icss} is adapted to find changes in variance.
-It obtains a \emph{likelihood ratio} for testing the hypothesis of one change against no change in the variance.
-Using an iterative approach, all possible change points are considered.
-The proposal of \cite{cheng2009efficient} extends on the \gls{cusum}-based methods to find change points in mean and variance, by creating a more efficient and accurate algorithm.
-
-Section \ref{sec:change_detection_time_series} discusses the problem of change detection in time series further, and gives a formal problem definition.
-
-\TODO{move CUSUM based techniques to this subsection}
-
---- Change detection ---
-
-``A unifying framework for detecting outliers and change points from time series'' \cite{takeuchi2006unifying}. 87, 2006 \\
-
-% ``Bayesian online changepoint detection'' \cite{adams2007bayesian}. 85, 2007 \\
-
-% ``An adaptive cusum-based test for signal change detection'' \cite{alippi2006adaptive}. 18, 2006 \\
-
-% ``An efficient algorithm for estimating a change-point'' \cite{cheng2009efficient}. 5, 2009 \\
-
-% ``The marginalized likelihood ratio test for detecting abrupt changes'' \cite{gustafsson1996marginalized}. 80, 1996 \\
-
-% ``The MOSUM of squares test for monitoring variance changes'' \cite{hsu2007mosum}. 4, 2007 \\
-
-% ``Use of cumulative sums of squares for retrospective detection of changes of variance'' \cite{inclan1994use}. 643, 1994 \\
-
-% ``Change-point detection in time-series data by direct density-ratio estimation'' \cite{kawahara2009change}. 52, 2009 \\
-``Sequential change-point detection based on direct density-ratio estimation'' \cite{kawahara2012sequential}. 22, 2012 \\
-% ``Change-point detection in time-series data by relative density-ratio estimation'' \cite{liu2013change}. 11, 2013 \\
-
-``Change point detection in time series data using support vectors'' \cite{camci2010change}. 3, 2010 \\
-
-\subsection{Novelty detection}\label{subsec:novelty_detection}
-
---- Novelty detection ---
-
-``Online novelty detection on temporal sequences'' \cite{ma2003online}. 146, 2003 \\
-``Time-series novelty detection using one-class support vector machines'' \cite{ma2003time}. 78, 2003 \\
-
-``Novelty detection: a review—part 1: statistical approaches'' \cite{markou2003novelty}. 697, 2003 \\
-
-``Support Vector Method for Novelty Detection'' \cite{scholkopf1999support}. 337, 1999 \\
-
-\subsection{Outlier detection}\label{subsec:outlier_detection}
-
---- Outlier detection ---
-``A unifying framework for detecting outliers and change points from time series'' \cite{takeuchi2006unifying}. 87, 2006 \\
-
-``Outliers in statistical data'' \cite{barnett1994outliers} (book). 3745, 1994 \\
-
-``A survey of outlier detection methodologies'' \cite{hodge2004survey}. 791, 2004 \\
-
-``Local outlier detection reconsidered: a generalized view on locality with applications to spatial, video, and network outlier detection'' \cite{schubert2012local}. 2, 2012 \\
-
-% --- SVMs ---
-
-% ``Change point detection in time series data using support vectors'' \cite{camci2010change}. 3, 2010 \\
-
-% ``Time-series novelty detection using one-class support vector machines'' \cite{ma2003time}. 78, 2003 \\
-
-% ``Support vector domain description'' \cite{tax1999support}. 907, 1999
-% ``Support vector data description applied to machine vibration analysis'' \cite{tax1999supportdata}. 76, 1999 \\
-
-% ``Human Activity Recognition on Smartphones using a Multiclass Hardware-Friendly Support Vector Machine'' \cite{anguitahuman}. 13, 2012 \\
-
-% ``Support vector machines in remote sensing: A review'' \cite{mountrakis2011support}. 150, 2011 \\
-
-% ``Sensor-based abnormal human-activity detection'' \cite{yin2008sensor}. 74, 2008 (builds one-class SVM of all normal traces) \\
+Give overview of used temporal segmentation methods. Problem is that it is not really well-defined.

### 2013-11-07 Chapters/Chapter2/temporal_segmentation_old.tex
WIP: Working on organization of Chapter 2

% !TEX root = ../../main.tex
\section{Temporal Segmentation}\label{sec:temporal_segmentation_old}

\begin{itemize}
  \item Given overview of segmentation techniques, for times series data
  \item Use different ``point-of-views'', or terms
  \item ``Segmentation''
  \item ``Change detection''
  \item ``Novelty detection''
  \item Specific view on \glspl{svm}
\end{itemize}

This section gives an overview of the literature on temporal segmentation in the context of \gls{har}.
It takes a look on different implementations and methodologies.
A wide range of terms and subtle differences are used in the field, such as `segmentation', `change detection', `novelty detection' and `outlier detection'.
These will be the categorical terms for which we discuss the literature.
Finally we will discuss other applications of \glspl{svm} in the context of these terms.

\subsection{Segmentation}\label{subsec:segmentation}
\TODO{This subsection is mainly from previous draft version}

\TODO{Create compact notation, one sentence per paper max}

Segmentation methods can roughly be categorized in three methods in the way the data is processed, as discussed by Avci \etal \cite{avci2010activity}:
\begin{itemize}
  \item \textbf{Top-down} methods iteratively divide the signal in segments by splitting at the best location.
  The algorithm starts with two segments and completes when a certain condition is met, such as when an error value or number of segments $k$ is reached.
  These methods process the data points recursively, which results in a complexity of $O(n^2k)$.
  \item \textbf{Bottom-up} methods are the natural complement to top-down methods.
  They start with creating $n/2$ segments and iteratively join adjacent segments while the value of a cost function for that operation is below a certain value.
  Given the average segment length $L$, the complexity of this method is $O(nL)$.
  \item \textbf{Sliding-window} methods are simple and intuitive for online segmenting purposes.
  It starts with a small initial subsequence of the time series.
  New data points are joined in the segment until the fit-error is above a threshold.
  Since the data is only processed very locally, these methods can yield in poor results \cite{keogh2001online}.
  The complexity is equal to the bottom-up approach, $O(nL)$, where $L$ is the average segment length.
  \item \textbf{\acrlong{swab}}, as introduced by Keogh \etal \cite{keogh2001online}, joins the ability of the sliding window mechanism to process time series online and the bottom-up approach the create superior segments in terms of fit-error.
  The algorithm processes the data in two stages.
  The first stage is to join new data points in the current segment created by a sliding window, and pass this to a buffer with space for a few segments.
  The buffer then processes the data Bottom-up and returns the first (left-most) segment as final segment.
  Because this second stage retains some (semi-)global view of the data, the results are comparative with normal Bottom-up.
  It is stated by Keogh \etal that the complexity of \gls{swab} is a small constant factor worse than that of regular Bottom-up.
\end{itemize}
It is clear that for the application of this research sliding-window and preferably \gls{swab}-based algorithms should be considered.

The \gls{swab} method proposed by Keogh \etal \cite{keogh2001online} is dependent on an user setting, providing the maximum error when performing both stages.
Each segment is approximated by using piecewise linear representation (PLR), an often used method.
The user provided error threshold controls the granularity and number of segments.
Other methods have been proposed, such as an adaptive threshold based on the signal energy by Guenterberg \etal \cite{guenterberg2009automatic}, the adaptive \gls{cusum}-based test by Alippi \etal \cite{alippi2006adaptive} and the \gls{mosum} by Hsu \cite{hsu2007mosum} in order to eliminate this user-dependency.
The latter of these methods is able to process the accelerometer values directly, although better results are obtained when features of the signal are processed, as done in the former method.
Here the signal energy, mean and standard deviation are used to segment activities and by adding all the axial time series together, the Signal-To-Noise ration is increased, resulting in a robuster method.

The method of Guenterberg \etal extracts features from the raw sensor signal to base the segmentation on other properties than the pure values.
The method of Bernecker \etal \cite{bernecker2012activity} uses other statistical properties, namely autocorrelation, to distinguish periodic from non-periodic segments.
Using the \gls{swab} method the self-similarity of a one-dimensional signal is obtained.
The authors claim that only a slight modification is needed to perform the method on multi-dimensional data.
After the segmentation phase, the method of Bernecker \etal extracts other statistical features which are used in the classification phase.

The proposal of Guo \etal \cite{guo2012adaptive} dynamically determines which features should be used for the segmentation and simultaneously determines the best model to fit the segment.
For each of the three dimensions features such as the mean, variance, covariance, correlation, energy and entropy are calculated.
By extending the \gls{swab} method, for every frame a feature set is selected, using an enhanced version of \gls{pca}.
The research also considered the (Stepwise) Feasable Space Window as introduced by \cite{liu2008novel}, but since it results in a higher error rate than \gls{swab} the latter was chosen to extend.
Whereas the before mentioned algorithms use a linear representation, this methods considers linear, quadratic and cubical representations for each segment.
This differs from other methods where the model is fixed for the whole time series, such as \cite{fuchs2010online}, which is stated to perform inferior on non-stationary time series such as daily life.

The time series data from a sensor can be considered as being drawn from a certain stochastic process.
Probabilistic models can be constructed on that signal, yielding in probabilistic and Bayesian based segmentation methods.
The \gls{cusum}-methods follows a statistical approach and relies on the log-likelihood ratio \cite{gustafsson1996marginalized} to measure the difference between two distributions.
To calculate the ratio, the probability density functions need to be calculated.
The method of Kawahara \etal \cite{kawahara2009change} proposes to estimate the ratio of probability densities (known as the \emph{importance}), based on the log likelihood of test samples, directly, without explicit estimation of the densities.
The method by Liu \etal \cite{liu2013change} uses a comparable dissimilarity measure using the \gls{kliep} algorithm.
They claim this results in a robuster approach for real-world scenarios.
Although this is a model-based method, no strong assumptions (parameter settings) are made on the models.

The method of Adams and MacKay \cite{adams2007bayesian} builds a probabilistic model on the segment run length, given the observed data so far.
Instead of modeling the values of the data points as a probabilistic distribution, the length of segments as a function if time is modeled by calculating the posterior probability.
It uses a prior estimate for the run length and a predictive distribution for newly-observed data, given the data since the last change point.
This method contrasts with the approach of Guralnik and Srivastava \cite{guralnik1999event} in which change points are detected by a change in the (parameters of an) underlying, observed, model.
For each new data point, the likelihoods of being a change point and part of the current segment are calculated, without a prior model (and thus is a non-Bayesian approach).
It is observed that when no change point is detected for a long period of time, the computational complexity increases significantly.

Another application of \gls{pca} is to characterize the data by determining the dimensionality of a sequence of data points.
The proposed method of Berbi\v{c} \etal \cite{barbivc2004segmenting} determines the number of dimensions (features) needed to approximate a sequence within a specified error.
With the observation that more dimensions are needed to keep the error below the threshold when transitions between actions occur, cut-points can be located and segments will be created.
The superior extension of their approach uses a Probabilistic \gls{pca} algorithm to incorporate the dimensions outside the selected set as noise.

In the method by Himberg \etal \cite{himberg2001time} a cost function is defined over segments of data which is to be minimized.
The cost functions thereby searches for internally homogeneous segments of data, reflecting states in which the devices and the user are.
The cost function can be any arbitrary function and in the implementation the sum of variances over the segments is used.
Both in a local and global iterative replacement procedure (as an alternative for the computationally hard dynamic programming algorithm) the best breakpoint locations $c_i$ for a pre-defined number of segments $1 \leq i \leq k$ are optimized.

Many methods obtain an implicit segmentation as a result of classification over a sliding window \TODO{add refs}.
The method of Yang \etal \cite{yang2008distributed} explicitly performs segmentation and classification simultaneously.
It argues that the classification of a pre-segmented test-sequences becomes straightforward with many classical algorithms to choose from.
The algorithm matches test examples with the \emph{sparsest} linear representation of mixture subspace models of training examples, searching over different temporal resolutions.

The method of Chamroukhi \etal \cite{chamroukhi2013joint} is based on a \gls{hmm} and logistic regression.
It assumes a $K$-state hidden process with a (hidden) state sequence, each state providing the parameters (amongst which the order) for a polynomial.
The order of the model segment is determined by model selecting, often using the \gls{bic} or the similar \gls{aic} \cite{akaike1974new}, as in \cite{he2008activity}.

Field of computer vision: \cite{zhou2008aligned}, \cite{li2007segmentation}.

--- Segmentation ---

% ``An online algorithm for segmenting time series'' \cite{keogh2001online}. See~\ref{sec:appendix-C-online-keogh}. 538, 2001 \\
% ``Segmenting time series: A survey and novel approach'' \cite{keogh2004segmenting}. 242, 2004 \\

% ``Time series segmentation for context recognition in mobile devices'' \cite{himberg2001time}. 158, 2001 \\

% ``An Adaptive Approach for Online Segmentation of Multi-Dimensional Mobile Data'' \cite{guo2012adaptive}. 4, 2012 \\

% ``Joint segmentation of multivariate time series with hidden process regression for human activity recognition'' \cite{chamroukhi2013joint}. 0, 2013. See~\ref{sec:appendix-C-joint-segmentation}. \\

``Segmentation and Recognition of Motion Streams by Similarity Search'' \cite{li2007segmentation}. 29, 2007 \\

% ``Novel Online Methods for Time Series Segmentation'' \cite{liu2008novel}. 22, 2008 \\

% ``Distributed Segmentation and Classification of Human Actions Using a Wearable Motion Sensor Network'' \cite{yang2008distributed}. 44, 2008 \\

% ``An Automatic Segmentation Technique in Body Sensor Networks based on Signal Energy'' \cite{guenterberg2009automatic}. 14, 2009 \\

% ``Online Segmentation of Time Series Based on Polynomial Least-Squares Approximations'' \cite{fuchs2010online}. 24, 2010 \\

``Aligned Cluster Analysis for Temporal Segmentation of Human Motion'' \cite{zhou2008aligned}. 63, 2008 \\

% ``Segmenting motion capture data into distinct behaviors'' \cite{barbivc2004segmenting}. 270, 2004 \\

\subsection{Change detection}\label{subsec:change_detection}
\TODO{change order of this and temporal segmentation sections? So first change, then segmentation?} \\
Whereas the above mentioned researches focus on \emph{segmentation}, many have focused on \emph{change detection}.
Although these techniques are closely related, there is a subtle difference.
In the case of \emph{change detection} to goal is to find, possibly unrelated, sudden change points in a signal \cite{takeuchi2006unifying}.
In contrast, the goal of \emph{temporal segmentation} is to find homogeneous segments of data, which can be the result of multiple detected changes.

The \gls{icss} by Incl\'{a}n and Tiao \cite{inclan1994use} is a statistical method which obtains results (when applied to stock data) comparable to \gls{mle} and Bayesian \TODO{Bayesian What?}.
Whereas \gls{cusum} can be applied to search for a change in mean, the \gls{icss} is adapted to find changes in variance.
It obtains a \emph{likelihood ratio} for testing the hypothesis of one change against no change in the variance.
Using an iterative approach, all possible change points are considered.
The proposal of \cite{cheng2009efficient} extends on the \gls{cusum}-based methods to find change points in mean and variance, by creating a more efficient and accurate algorithm.

Section~\ref{sec:change_detection_time_series} discusses the problem of change detection in time series further, and gives a formal problem definition.

\TODO{move CUSUM based techniques to this subsection}

--- Change detection ---

``A unifying framework for detecting outliers and change points from time series'' \cite{takeuchi2006unifying}. 87, 2006 \\

% ``Bayesian online changepoint detection'' \cite{adams2007bayesian}. 85, 2007 \\

% ``An adaptive cusum-based test for signal change detection'' \cite{alippi2006adaptive}. 18, 2006 \\

% ``An efficient algorithm for estimating a change-point'' \cite{cheng2009efficient}. 5, 2009 \\

% ``The marginalized likelihood ratio test for detecting abrupt changes'' \cite{gustafsson1996marginalized}. 80, 1996 \\

% ``The MOSUM of squares test for monitoring variance changes'' \cite{hsu2007mosum}. 4, 2007 \\

% ``Use of cumulative sums of squares for retrospective detection of changes of variance'' \cite{inclan1994use}. 643, 1994 \\

% ``Change-point detection in time-series data by direct density-ratio estimation'' \cite{kawahara2009change}. 52, 2009 \\
``Sequential change-point detection based on direct density-ratio estimation'' \cite{kawahara2012sequential}. 22, 2012 \\
% ``Change-point detection in time-series data by relative density-ratio estimation'' \cite{liu2013change}. 11, 2013 \\

``Change point detection in time series data using support vectors'' \cite{camci2010change}. 3, 2010 \\

\subsection{Novelty detection}\label{subsec:novelty_detection}

--- Novelty detection ---

``Online novelty detection on temporal sequences'' \cite{ma2003online}. 146, 2003 \\
``Time-series novelty detection using one-class support vector machines'' \cite{ma2003time}. 78, 2003 \\

``Novelty detection: a review—part 1: statistical approaches'' \cite{markou2003novelty}. 697, 2003 \\

``Support Vector Method for Novelty Detection'' \cite{scholkopf1999support}. 337, 1999 \\

\subsection{Outlier detection}\label{subsec:outlier_detection}

--- Outlier detection ---
``A unifying framework for detecting outliers and change points from time series'' \cite{takeuchi2006unifying}. 87, 2006 \\

``Outliers in statistical data'' \cite{barnett1994outliers} (book). 3745, 1994 \\

``A survey of outlier detection methodologies'' \cite{hodge2004survey}. 791, 2004 \\

``Local outlier detection reconsidered: a generalized view on locality with applications to spatial, video, and network outlier detection'' \cite{schubert2012local}. 2, 2012 \\

% --- SVMs ---

% ``Change point detection in time series data using support vectors'' \cite{camci2010change}. 3, 2010 \\

% ``Time-series novelty detection using one-class support vector machines'' \cite{ma2003time}. 78, 2003 \\

% ``Support vector domain description'' \cite{tax1999support}. 907, 1999
% ``Support vector data description applied to machine vibration analysis'' \cite{tax1999supportdata}. 76, 1999 \\

% ``Human Activity Recognition on Smartphones using a Multiclass Hardware-Friendly Support Vector Machine'' \cite{anguita2012human}. 13, 2012 \\

% ``Support vector machines in remote sensing: A review'' \cite{mountrakis2011support}. 150, 2011 \\

% ``Sensor-based abnormal human-activity detection'' \cite{yin2008sensor}. 74, 2008 (builds one-class SVM of all normal traces) \\

### 2013-11-07 Chapters/Chapter4/data_gathering.tex
WIP

@@
 In Chapter~\ref{Chapter6} we give a detailed analysis of the performed activities and the recorded data sets.
+
+For the experiments we used a HTC Sensation XE smartphone as recording device.
+The activities were recorded using a free Android application \cite{sensorlogger}.
+This application was chosen for its convenient data format of the sensor recording and its regularity of the sampling interval.
+Table~\ref{tab:recorded_metrics} lists the recorded metrics.
+For our expiments we used the data for the accelerometer, magnatief field and rotation.
+
+\begin{center}\begin{table}
+  \begin{tabulary}{\textwidth}{|l|L|c|c|}
+    \hline
+    Metric & Description & Units of measure & Typical range \\
+    \hline \hline
+    Accelerometer & Acceleration force along each axis. & $m/s^2$ & $-20$ -- $20$ \\
+    \hline
+    Gravity & Force of gravity along each axi.s & $m/s^2$ & $-10$ -- $10$\\
+    \hline
+    Gyroscope & & $rad/s$ & $-15$ -- $15$\\
+    \hline
+    Light & Light sensitive sensor at the front of the phone. & & $0$ -- $10000$ \\
+    \hline
+    Linear acceleration & & $m/s^2$ & $-20$ -- $20$ \\
+    \hline
+    Magnetic field & Geomagnetic field strength along each axis. & $\mu T$ & $-60$ -- $60$ \\
+    \hline
+    Orientation & & & $-100$ -- $360$ \\
+    \hline
+    Rotation & & & $-1$ -- $1$\\
+    \hline
+  \end{tabulary}
+  \caption{Measured metrics. The set of axis is always the triple (x, y, z) direction.}
+  \label{tab:recorded_metrics}
+\end{table}\end{center}

### 2013-11-07 Chapters/Chapter4/model_construction.tex
WIP

@@
 We follow the \gls{svcpd} method by Camci \cite{camci2010change} and the \gls{isvdd} implementation by Tax and Laskov \cite{tax2003online}.
-That latter algorithm combines the techniques of online, unsupervised, and incremental learning methods with the earlier introduced \gls{oc-svm} algorithm \gls{svdd}.
+The latter algorithm combines the techniques of online, unsupervised, and incremental learning methods with the earlier introduced \gls{oc-svm} algorithm \gls{svdd}.
 The method is initialized with a fixed window length and then in every step a new data object is added to and the oldest data object is removed from the working set.

### 2013-11-11 Chapters/Chapter6/data_gathering.tex
Draft, sent to Anne
WIP
Cropped graphs

@@
 \section{Data Gathering}\label{sec:data_gathering}
+
+\TODO{Use common data sets. The WISDM dataset is not usefull because it has no transitions, only very clear blocks of the same activities. Include a plot to illustrate the problem?}
+
+\begin{figure}\label{fig:wisdm_excerpt}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/Chapter6/data_collection/wisdm_excerpt.eps}
+  \caption[Jumping variance ratios]{Excerpt from the WISDM data set \cite{kwapisz2011activity}.
+First five activities for subject $33$ are displayed.
+Due to the discontinuous nature of the recordings, this data set can not be used for change detection.}
+\end{figure}
 

### 2013-11-11 Chapters/Chapter3/one_class_svm.tex
Add UCI HAR plot and other notes

@@
 \end{equation}
+\TODO{Elaborate on imporant properties of page 56, in Appendix A: maximum-margin is searching for support vectors and pairwise dot products, Kernel matrix.}
 where $K(x, x_i) = \phi(x)^T\phi(x_i)$ (which is further discussed in Section~\ref{subsec:kernel_trick}).

### 2013-11-11 Chapters/Chapter4/change_detection.tex
Add UCI HAR plot and other notes

@@
 
-\subsection{Ratio-thresholding}
+\TODO{This subsection has double information. Cleanup. Mention iterated cumulative sum of squares of inclan \cite{inclan1994use}.}
+
+\subsection{Ratio-thresholding}\label{subsec:ratio_thresholding}
 The second example of an algorithm that interprets the change indication time series and transforms it to change detection is the radius ratio thresholding method used by Camci~\cite{camci2010change}.
@@
 \end{figure}
+
+\begin{center}\begin{table}
+\begin{tabular}{ l p{12cm} }
+  \hline
+  Step & Action \\
+  \hline
+  1 & Start with $n$ data objects as working set and construct hypersphere \\
+  2 & Add next data object $x_t$ and drop first one \\
+  3 & Identify new hyper-sphere and its \emph{approximate radius} \\
+  4 & calculate radius average of hyperspheres since last change point \\
+  5 & calculate radius ratio $\hbar$. If lower than ${th}_{low}$ or greater than ${th}_{high}$ then mark $t$ as change point \\
+  6 & continue from step 2, until no more data objects available \\
+  7 & Apply post-processing: merge close change points \\
+  \hline
+\end{tabular}
+\caption{Algorithm of the proposed change detection method, using the Ratio-Thresholding method of \ref{subsec:ratio_thresholding} for change detection.}
+\label{tab:algorithm_proposed_method}
+\end{table}\end{center}
+
+\TODO{Add reference to Table~\ref{tab:algorithm_proposed_method}.}

### 2013-11-11 Chapters/Chapter6/data_gathering.tex
Add UCI HAR plot and other notes

@@
 
-\TODO{Use common data sets. The WISDM dataset is not usefull because it has no transitions, only very clear blocks of the same activities. Include a plot to illustrate the problem?}
+\TODO{Use common data sets. The WISDM dataset is not useful because it has no transitions, only very clear blocks of the same activities. Include a plot to illustrate the problem?}
 
@@
   \includegraphics[width=1\textwidth]{./Figures/Chapter6/data_collection/wisdm_excerpt.eps}
-  \caption[Jumping variance ratios]{Excerpt from the WISDM data set \cite{kwapisz2011activity}.
+  \caption[WISDM Excerpt]{Excerpt from the WISDM data set \cite{kwapisz2011activity}.
 First five activities for subject $33$ are displayed.
 Due to the discontinuous nature of the recordings, this data set can not be used for change detection.}
+\end{figure}
+
+\begin{figure}\label{fig:uci_annotated}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/Chapter6/data_collection/uci_annotated.eps}
+  \caption[UCI HAR Excerpt]{Excerpt from the UCI HAR data set \cite{anguita2012human}.
+The recorded training activities for subject $25$ are displayed.
+Due to the imprecise (and possibly incorrect) labeling and the short duration of activities, the data set can not be used for the change detection method as described in this thesis.}
 \end{figure}

### 2013-11-13 Appendices/AppendixA-summaries.tex
WIP: kernels

@@
 An instance $\vectorsym{x}$ is then classified as $\hat{y} = sign(\sum_{i=1}^n \alpha_i y_i \vectorsym{x}_i \cdot \vectorsym{x})$.
-This means that during the training only the pairwise dot-products of the data is needed; this results in the \emph{n}-by-\emph{n} Gram matrix $\matrixsym{G} = \matrixsym{X}\matrixsym{X}^T$.
+This means that during the training only the pairwise dot products of the data is needed; this results in the \emph{n}-by-\emph{n} Gram matrix $\matrixsym{G} = \matrixsym{X}\matrixsym{X}^T$.
 This instance-based perspective will be further discussed in Section~\ref{svm-explained} about the support vector machine.
@@
 \end{equation}
-Note that only dot-products of the data points $\vectorsym{x}$ appear.
+Note that only dot products of the data points $\vectorsym{x}$ appear.
 In order to transform the data points to a higher dimension, to create a good representational hyper-sphere, kernels replace the dot products without compromising computational complexity.

### 2013-11-13 Chapters/Chapter2.tex
WIP: kernels

@@
 It originates from control methods for detection from benchmarks.
-This method and some derivatiges are discussed and analyzed in Section~\ref{cusum}.
+This method and some derivatiges are discussed and analyzed in Section~\ref{sec:cusum}.
 
@@
 The methods proposed by Kawahara \etal \cite{kawahara2009change} and Lui \etal \cite{liu2013change} try to overcome these problems by estimating the \emph{ratio} between the \glspl{pdf}, instead of estimating each \gls{pdf}.
-This approach is discussed and analyzed in Section~\ref{density-ratio}.
+This approach is discussed and analyzed in Section~\ref{sec:density-ratio}.
 

### 2013-11-13 Chapters/Chapter3/one_class_svm.tex
WIP: kernels

--- 2013-11-06 Chapters/Chapter3/one_class_svm.tex
@@
 This is illustrated in Figure~\ref{fig:svm_mapping_spaces}.
-The mapping and projection of data points can be efficient (and implicit) performed by using the kernel trick, which is discussed in Section~\ref{subsec:kernel_trick}.
+The mapping and projection of data points can be efficient (and implicit) performed by using the kernel trick, which is discussed in Section~\ref{subsec:kernels}.
+
+\begin{figure}
+\centering
+  \includegraphics[width=0.8\textwidth]{./Figures/chapter3/svm_kernel_mapping.png}
+  \caption[Kernel mapping]{The non-linear boundary in the input space $\mathcal{I}$ (left) is transformed to a linear boundary in the feature space $\mathcal{F}$ (right) by mapping the data points with the function $\phi$.
+The kernel trick uses a function $K$ which performs an implicit mapping.
+Image from Wikipedia.org\footnotemark}
+  \label{fig:kernel_mapping}
+\end{figure}
+\footnotetext{\url{http://en.wikipedia.org/wiki/File:Kernel_Machine.png}}
 
@@
 \end{equation}
-where $K(x, x_i) = \phi(x)^T\phi(x_i)$ (which is further discussed in Section~\ref{subsec:kernel_trick}).
+where $K(x, x_i) = \phi(x)^T\phi(x_i)$ is the dot product of data objects in feature space $\mathcal{F}$ (which is further discussed in Section~\ref{subsec:kernels}).
 Here every data point $\mathcal{I}$ for which $a_i > 0$ is weighted in the decision function and thus ``supports'' the classification machine: hence the name ``\acrlong{svm}''.
 Since it is shown that under certain circumstances \glspl{svm} show an equality to sparse representations \cite{girosi1998equivalence,smola1998connection}, there will often be relatively few Lagrange multipliers with a non-zero value.
+
+Using this formulation two imporant properties arise \cite{flach2012machine}:
+\begin{enumerate}
+  \item Searching for the maximum margin decision boundary is equivalent to searching for the support vectors; they are the training examples with non-zero Lagrange mulitpliers.
+  \item The optimization problem is entirely defined by pairwise dot products between training examples: the entries of the kernel matrix $K$.
+\end{enumerate}
+The second property is what enables powerful adaption for support vector machines to learn non-linear decision boundaries.
+The workings of the kernel matrix $K$ and the non-linear boundaries is discussed in the following section.
 
@@
 
-\subsection{Kernel trick}\label{subsec:kernel_trick}
-In the previous section,~\ref{subsec:svm}, the mapping function $\phi(x)$ and the kernel function $K$ were briefly mentioned.
-The decision function in equation~\ref{eq:svm_lagrange} only relies on the dot-products of mapped data points in the feature space $\mathcal{F}$ (\ie all pairwise distances between the data points in that space).
-It shows \cite{flach2012machine} that as long as any function has the same result, without an explicit mapping to the higher dimension $\mathcal{F}$, the dot-products can be substituted by the kernel function $K$.
+\subsection{Kernels}\label{subsec:kernels}
+In the previous section, the mapping function $\phi(x)$ and the kernel function $K$ were briefly mentioned.
+The decision function in Equation~\ref{eq:svm_lagrange} only relies on the dot products of mapped data points in the feature space $\mathcal{F}$ (\ie all pairwise distances between the data points in that space).
+It shows \cite{flach2012machine} that as long as any function has the same result, without an explicit mapping to the higher dimension $\mathcal{F}$, the dot products can be substituted by the kernel function $K$, as introduced by Aizerman~\etal \cite{aizerman1964theoretical} and applied to \glspl{svc} by Vapnik \cite{vapnik1998statistical}:
+\begin{equation}
+  K(\vectorsym{x}, \vectorsym{x}') = (\vectorsym{z} \cdot \vectorsym{z}') = \phi(\vectorsym{x}) \cdot \phi(\vectorsym{x}')
+\end{equation}
+where vectors $\vectorsym{z}$ and $\vectorsym{z}'$ are projections of data objects $\vectorsym{x}$ and $\vectorsym{x}'$ through $\phi(\vectorsym{x})$ on the features space $\mathcal{F}$ and the dot product kernel $K$ is determined by the sum
+\begin{equation}
+  K(\vectorsym{x}, \vectorsym{x}') = \sum_{j=1}^m g_j(\vectorsym{x}) g_j(\vectorsym{x}')
+\end{equation}
+of basis functions $g_j(\vectorsym{x})$.
+Note that the evaluation of dot products in the feature space $\mathcal{F}$ between vectors is performed indirectly via the evaluation of the kernel $K$ between (support) vectors in the input space $\mathcal{I}$.
 This is known as the \emph{kernel trick} and gives the \gls{svm} the ability to create non-linear decision function without high computational complexity.
-This mapping is illustrated in Figure~\ref{fig:kernel_mapping}.
-Here the non-linear separating boundary in the input space $\mathcal{I}$ is mapped, via $\phi$, to a linear boundary in the feature space $\mathcal{F}$.
 
 The kernel function $K$ can have different forms, such as linear, polynomial and sigmoidal but the most used (and flexible) form is the Gaussian \gls{rbf}.
+The basis functions have the form
+\begin{equation}
+  g(x) = \operatorname*{sign} \left(  \sum_{i=1}^{n} \alpha_i \operatorname*{exp} \left\{ \frac{\rVert \vectorsym{x} - \vectorsym{x}_i \rVert ^2}{\sigma^2} \right\} \right),
+\end{equation}
+where $\sigma$ defines the width and $\rVert \vectorsym{x} - \vectorsym{x}' \rVert ^2$ is the dissimilarity measure expressed as Euclidian distance.
+The inner product kernel $K$ then becomes
+\begin{equation}
+  K(\vectorsym{x}, \vectorsym{x}') = \operatorname*{exp} \left\{ - \frac{ \rVert \vectorsym{x} - \vectorsym{x}' \rVert ^2}{\sigma^2} \right\}.
+\end{equation}
+The kernel $K$ maps input space $\mathcal{I}$ to the feature space $\mathcal{F}$ which is a \gls{rkhs} of (theoretically) infinite dimensions.
 As Smola \etal \cite{smola1998connection} state, this Gaussian kernel yields good performance, especially when no assumptions can be made about the data.
-The kernel maps input space $\mathcal{I}$ to the feature space $\mathcal{F}$ which is a Hilbert Space of infinite dimensions \TODO{ref needed}:
-\begin{equation}
-  K(x, x') = \operatorname{exp} \left( - \frac{ \lVert x - x' \rVert ^2}{2 \sigma^2 } \right),
-\end{equation}
-where $\sigma \in R$ is a kernel parameter and $\lVert x - x' \rVert$ is the dissimilarity measure expressed in Euclidean distance.
+As an explanation, they show a correspondence between learning \glspl{svm} with \gls{rbf} kernels and good regularization operators.
+This may give insights in why \glspl{svm} have been found to exhibit high generalization ability (by learning with few training objects).
 
-\begin{figure}
-\centering
-  \includegraphics[width=0.8\textwidth]{./Figures/chapter3/svm_kernel_mapping.png}
-  \caption[Kernel mapping]{The non-linear boundary in the input space $\mathcal{I}$ (left) is transformed to a linear boundary in the feature space $\mathcal{F}$ (right) by mapping the data points with the function $\phi$.
-The kernel trick uses a function $K$ which performs an implicit mapping.
-Image from Wikipedia.org\footnotemark}
-  \label{fig:kernel_mapping}
-\end{figure}
-\footnotetext{\url{http://en.wikipedia.org/wiki/File:Kernel_Machine.png}}
+The number of basis functions, the center parameters that correspond with support vectors, and the weights in the ouput layer are all automatically determined via the optimal hyperplane in features space $\mathcal{F}$ \cite{cherkassky2007learning}.
+The width parameter $\sigma$ is equal for all basis functions and is set a priori and determines the flexibility and complexity of the boundary.
+In Section~\ref{subsec:svm_model_parameters} this (hyper)parameter for a \gls{svm} is further discussed.
 
-%--------------------------------------------
-
-\subsection{Reproducing kernel Hilbert space}
-\emph{Write something on this topic? Or leave it implicit?}
+The mapping from input space $\mathcal{I}$ to $\mathcal{F}$ via $\phi(x)$ is subject to the continuity assumptions.
+This general assumption in pattern recogition state that two near objects in feature space should also resemble each other in real life \cite{tax2001one}.
+Thus, objects which are close in feature space should be close in the original input space.
+When this assumption would not hold, the example objects would be scatter through the feature space and finding a decision function becomes very complex.
 
@@
 -- Quotes and refs to use --
-``Continuity assumptions: This is a general assumption in pattern recognition: two objects near in feature space should also resemble each other in real life'' \cite{tax2001one}.
-
-$k(\mathbf{x}_i, \mathbf{x})$ is a kernel function computing a dot product in feature space, introduced by Aizerman \etal \cite{aizerman1964theoretical}.
-
-``Training a \gls{svm} with Gaussian \gls{rbf} kernels corresponds to minimizing the specific cost function with a regularization operator'' \cite{smola1998connection}. \\
-
-``Gaussian kernels tend to yield good performance under general smoothness assumptions and should be considered especially of no additional knowledge of the data is available.'' \cite{smola1998connection}.
-\\
-
-``Choosing a small width of the kernels leads to high generalization error as it effectively decouples the separate basis functions of the kernel expansion into very localized functions which is equivalent to memorizing the data, whereas a wide kernel tends to oversmooth.'' \cite{smola1998connection}.
-\\
-
-``For small values of $\sigma$ almost all $\alpha_i >0$ and the \gls{svdd} resembles a Parzen density estimation.'' \cite{tax2002uniform} page 4, \cite{tax1999support} page 4. \\
-
-sparsity, RKHS: \cite{girosi1998equivalence} \\
-
-Kernels for Support Vector Classifiers were first (?) used by Vapnik \cite{vapnik1998statistical}. \\
+sparsity, \gls{rkhs}: \cite{girosi1998equivalence} \\
 
@@
 ``On simple one-class classification methods'' \cite{noumir2012simple}. 2012  --> decouples radius and center optimization, gives fast approximations instead of precise results. \\
+
+``Choosing a small width of the kernels leads to high generalization error as it effectively decouples the separate basis functions of the kernel expansion into very localized functions which is equivalent to memorizing the data, whereas a wide kernel tends to oversmooth.'' \cite{smola1998connection}.
+\\
+
+``For small values of $\sigma$ almost all $\alpha_i >0$ and the \gls{svdd} resembles a Parzen density estimation.'' \cite{tax2002uniform} page 4, \cite{tax1999support} page 4. \\

### 2013-11-13 Chapters/Chapter4/model_features.tex
WIP
WIP: kernels

@@
 This means that only the radius of the sphere will change during the algorithm.
-We will interpret changes in the radius $R$ with the same continuity assumptions as discussed in Section~\ref{subsec:kernel_function}.
+We will interpret changes in the radius $R$ with the same continuity assumptions as mentioned in Section~\ref{subsec:kernels}.
 This means that when two objects (in this case radius sizes) are close to each other, their original objects are also close.
@@
 
-\TODO{Notion of dimensionality reduction is a bit vague here, because it applies to a window of objects.}
+\TODO{Put the SVM method more in the light of dimensionality reduction. Applies to text in whole thesis...}

### 2013-11-14 Chapters/Chapter3/one_class_svm.tex
Remove old Kernels subsection

@@
 When this assumption would not hold, the example objects would be scatter through the feature space and finding a decision function becomes very complex.
+
+-- Quotes and refs to use --x
+
+sparsity, \gls{rkhs}: \cite{girosi1998equivalence} \\
+
+Use Chapter 2 from ``Learning with kernels'', \cite{scholkopf2002learning}.
 
@@
 %--------------------------------------------
-\subsection{Kernel Function}\label{subsec:kernel_function}
-
-\begin{itemize}
-  \item Explain RBF kernel
-  \item Explain RKHS? (Hilbert Space)
-\end{itemize}
-
-A common class for kernel function is the \acrlong{rbf}.
-These function are of the form
-\begin{equation}
-  g(x) = \operatorname*{sign} \left(  \sum_{i=1}^{n} \alpha_i \operatorname*{exp} \left\{ \frac{|\vectorsym{x} - \vectorsym{x}_i|^2}{\sigma^2} \right\} \right),
-\end{equation}
-where $\sigma$ defines the width.
-The inner product kernel is of the form
-\begin{equation}
-  K(\vectorsym{x}, \vectorsym{x}') = \operatorname*{exp} \left\{ - \frac{| \vectorsym{x} - \vectorsym{x}'|^2}{\sigma^2} \right\}.
-\end{equation}
-
-Note that with the \gls{rbf} class of kernel functions the number of basis functions, the center parameters that correspond to the support vectors, and the weights in the output layer are all automatically determined via the optimal hyperplane \cite{cherkassky2007learning}.
-The width parameter $\sigma$ is equal for all basis functions and is set a priori and determines the flexibility and complexity of the boundary.
-In Section~\ref{subsec:svm_model_parameters} this (hyper)parameter for a \gls{svm} is further discussed.
-Other examples of kernel function classes are \emph{pylonomial functions} of degree $q$, \emph{neural networks}, \emph{splines} of order $m$ with $b$ nodes, and \emph{Fourier expansions}.
-\TODO{More on basis functions. \cite{cherkassky2007learning}, chapter 9.4}
-
--- Quotes and refs to use --
-sparsity, \gls{rkhs}: \cite{girosi1998equivalence} \\
-
-Use Chapter 2 from ``Learning with kernels'', \cite{scholkopf2002learning}.
-
-%--------------------------------------------
 

### 2013-11-14 Chapters/Chapter4/change_detection.tex
Add caption titles to tables

@@
 \end{tabular}
-\caption{Algorithm of the proposed change detection method, using the Ratio-Thresholding method of \ref{subsec:ratio_thresholding} for change detection.}
+\caption[Proposed algorithm]{Algorithm of the proposed change detection method, using the Ratio-Thresholding method of \ref{subsec:ratio_thresholding} for change detection.}
 \label{tab:algorithm_proposed_method}

### 2013-11-14 Chapters/Chapter5/artificial_data_generation.tex
WIP: artificial data

--- 2013-10-28 Chapters/Chapter5/artificial_data.tex
@@
 % !TEX root = ../../main.tex
-\section{Artificial Data}\label{sec:artificial_data}
+\section{Artifical data generation}\label{sec:artificial_data_generation}
 \TODO{Use data from \cite{camci2010change,takeuchi2006unifying}}
 In this section we present three data sets as used by \cite{camci2010change,takeuchi2006unifying} to provide for a objetive performance comparison.
+
+Model:
+\begin{equation}
+  x_t = a_1 x_{t-1} + a_2 x_{t-2} + \epsilon_t,
+\end{equation}
+where $\epsilon_t$ is a Gaussian distribution, modeling the noise.
+The length of $x_t$ is $10000$ and change points are generated at each $y \times 10000^\text{th}$ data point, with $y = (1, 2, 3, \dots 9)$.
+
+First Takeuchi, reduced jumping mean:
+$\epsilon_t$ is Gaussian with mean $0$ and variance $\sigma^2 = 1$, $a_1 = 0.6$ and $a_2 = -0.5$.
+
+First Camci, fixed jumping mean.
+$\Delta y = 5$
+See Figure~\ref{fig:camci_mean_fixed}.
 
@@
   \caption[Jumping mean camci]{Jumping mean camci}
+  \label{fig:camci_mean_fixed}
 \end{figure}

### 2013-11-14 Chapters/Chapter5/results.tex
WIP: artificial data

% !TEX root = ../../main.tex
\section{Results}\label{sec:artificial_data_results}

### 2013-11-15 Chapters/Chapter5/artificial_data_generation.tex
Chapter 5 figures

@@
 
-First Takeuchi, reduced jumping mean:
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/notes/jumping_mean_takeuchi.eps}
+  \caption[Jumping mean Takeuchi, set 1]{Jumping mean takeuchi, set 1}
+  \label{fig:jumping_mean_takeuchi}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/notes/jumping_mean_takeuchi_thresholds.eps}
+  \caption[Jumping mean Takeuchi, set 1, thresholds]{Jumping mean takeuchi, set 1, thresholds}
+\end{figure}
+
+First Takeuchi, reduced jumping mean, Figure~\ref{fig:jumping_mean_takeuchi}:
 $\epsilon_t$ is Gaussian with mean $0$ and variance $\sigma^2 = 1$, $a_1 = 0.6$ and $a_2 = -0.5$.
 
-First Camci, fixed jumping mean.
-$\Delta y = 5$
-See Figure~\ref{fig:camci_mean_fixed}.
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/notes/jumping_variance_mean_takeuchi.eps}
+  \caption[Jumping variance and mean Takeuchi, set 2]{Jumping variance and mean takeuchi, set 2}
+  \label{fig:jumping_variance_and_mean_takeuchi}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/notes/jumping_variance_mean_takeuchi_thresholds_log_scale.eps}
+  \caption[Jumping variance mean Takeuchi, set 2, thresholds]{Jumping variance and mean takeuchi, set 2, thresholds (log scale)}
+\end{figure}
+
+Second Takeuchi, jumping mean with $\Delta y = 1$ and $\sigma^2 = 0.1 / (0.01 + (10000 - x)/1000)$.
+Figure~\ref{fig:jumping_variance_and_mean_takeuchi}.
 
@@
 \end{figure}
+
+First Camci, fixed jumping mean.
+$\Delta y = 5$
+See Figure~\ref{fig:camci_mean_fixed}.
 
@@
   \includegraphics[width=1\textwidth]{./Figures/notes/jumping_variance.eps}
-  \caption[Jumping variance]{Jumping variance}
+  \caption[Jumping variance]{Jumping variance, equal for Camci and Takeuchi, both set 3}
+  \label{fig:jumping_variance_camci_takeuchi}
 \end{figure}
+
+Third Camci and Takeuchi, jumping variance.
+When $y$ is odd, variance $\sigma^2$ changes from $1.0$ to $9.0$, and when $y$ is even variance $\sigma^2$ changes from $9.0$ to $1.0$.
+See Figure~\ref{fig:jumping_variance_camci_takeuchi}.
 

### 2013-11-18 Appendices/AppendixA-summaries.tex
Add caption titles to tables
Restructure Chapter 5 graphs of data sets

@@
 Linear models are well studied, in particular for the problem of linear regression.
-This can be solved by the \emph{least-squares} method and classification as discussed in Section~\ref{least-squares}, the \emph{perceptron} as explained in Section~\ref{perceptron}.
-Linear regression with the \emph{support vector machine} is handled in Section~\ref{svm-explained} and used for probability density estimation in Section~\ref{linear-classifier-pdf}.
-The kernel trick used for learning non-linear models is explained in Section~\ref{non-linear}.
+This can be solved by the \emph{least-squares} method and classification as discussed in \Cref{least-squares}, the \emph{perceptron} as explained in \Cref{perceptron}.
+Linear regression with the \emph{support vector machine} is handled in \Cref{svm-explained} and used for probability density estimation in \Cref{linear-classifier-pdf}.
+The kernel trick used for learning non-linear models is explained in \Cref{non-linear}.
 
@@
 This means that during the training only the pairwise dot products of the data is needed; this results in the \emph{n}-by-\emph{n} Gram matrix $\matrixsym{G} = \matrixsym{X}\matrixsym{X}^T$.
-This instance-based perspective will be further discussed in Section~\ref{svm-explained} about the support vector machine.
+This instance-based perspective will be further discussed in \Cref{svm-explained} about the support vector machine.
 
@@
 When we take the partial derivative with respect to $w\vectorsym{w}$ we see that the Lagrange multipliers define the weight vector as a linear combination of the training examples.
-This partial derivative is $0$ for an optimal weight we get that $\vectorsym{w} = \sum_{i=1}^{n} \alpha_i y_i \vectorsym{x}_i$, \emph{which is the same expression as for the perceptron derived in Section~\ref{perceptron}}.
+This partial derivative is $0$ for an optimal weight we get that $\vectorsym{w} = \sum_{i=1}^{n} \alpha_i y_i \vectorsym{x}_i$, \emph{which is the same expression as for the perceptron derived in \Cref{perceptron}}.
 By plugging $\vectorsym{w}$ and $t$ back into the Lagrange equation, we can eliminate these and get the dual optimization problem entirely formulated in terms of the Lagrange multipliers:
@@
 
-The second property enables powerful adaption for support vector machines to learn non-linear decision boundaries, as discussed in Section~\ref{non-linear}.
+The second property enables powerful adaption for support vector machines to learn non-linear decision boundaries, as discussed in \Cref{non-linear}.
 
@@
 About segmentation with SVMs.
-Will be main material for Section~\ref{sec:svm_in_change_detection} about SVMs.
+Will be main material for \Cref{sec:svm_in_change_detection} about SVMs.
 
@@
 \subsubsection{Problem formulation}
-[Not summarized here, useful for e.g. Section~\ref{statistical-framework}]
+[Not summarized here, useful for e.g. \Cref{statistical-framework}]
 
@@
 \end{tabular}
-\caption{Support Vector machine based Change Point Detection algorithm}
+\caption[\gls{svcpd} algorithm]{Support Vector machine based Change Point Detection algorithm}
 \label{tab:alg-svcdp}
@@
 There is a masking effect when there are multiple change points in the series; the order of small, medium and large variances result in the value of $D_k$.
-The iterative algorithm presented in this paper in Section~\ref{subsec:icss} is designed to lessen the masking effect.
+The iterative algorithm presented in this paper in \Cref{subsec:icss} is designed to lessen the masking effect.
 
@@
 \end{tabular}
-\caption{Iterated Cumulative Sums of Squares Algorithm}
+\caption[\gls{icss} algorithm]{Iterated Cumulative Sums of Squares Algorithm}
 \label{tab:alg-icss}

### 2013-11-18 Chapters/Chapter2.tex
Restructure Chapter 5 graphs of data sets

--- 2013-11-07 Chapters/Chapter2.tex
@@
 It originates from control methods for detection from benchmarks.
-This method and some derivatiges are discussed and analyzed in Section~\ref{cusum}.
+This method and some derivatiges are discussed and analyzed in \Cref{sec:cusum}.
 
@@
 The methods proposed by Kawahara \etal \cite{kawahara2009change} and Lui \etal \cite{liu2013change} try to overcome these problems by estimating the \emph{ratio} between the \glspl{pdf}, instead of estimating each \gls{pdf}.
-This approach is discussed and analyzed in Section~\ref{density-ratio}.
+This approach is discussed and analyzed in \Cref{sec:density-ratio}.
 
@@
 A change in the \gls{pdf} is represented by a change in the radius of this sphere.
-Section~\ref{sec:svm_in_change_detection} discusses the \gls{svm}-methods.
+\Cref{sec:svm_in_change_detection} discusses the \gls{svm}-methods.
 \TODO{Put emphasis that this is the source of inspiration for the chosen method?}
@@
 % It uses the estimated underlying number of parameters of the time series as a model for change detection.
-% Section~\ref{dim-reduction} discuses this method.
+% \Cref{dim-reduction} discuses this method.
 
@@
 \section{Overview}\label{sec:literature_review_overview}
-As described in Chapter~\ref{Chapter1}, the focus of this research is the application of temporal segmentation to accelerometer data.
+As described in \Cref{Chapter1}, the focus of this research is the application of temporal segmentation to accelerometer data.
 This is with the assumption that explicit temporal segmentation can aid the process of temporal classification.
@@
 Many techniques of temporal segmentation rely in change detection algorithms.
-In Section~\ref{sec:change_detection_methods} we will give an overview of those kind of algorithms.
+In \Cref{sec:change_detection_methods} we will give an overview of those kind of algorithms.
 An often encountered problem is the \emph{curse of dimensionality}.
-To overcome this problem, we look at \gls{dr} methods in Section~\ref{sec:dimensionality_reduction}.
+To overcome this problem, we look at \gls{dr} methods in \Cref{sec:dimensionality_reduction}.
 We end this chapter with a review on \gls{svm} based methods and algorithms.

### 2013-11-18 Chapters/Chapter2/svm.tex
WIP: Working on organization of Chapter 2
Restructure Chapter 5 graphs of data sets

@@
 
-The algorithms of \glspl{svm} has been applied to the problem of \gls{har}, as by Anguita \etal \cite{anguitahuman}.
+The algorithms of \glspl{svm} has been applied to the problem of \gls{har}, as by Anguita \etal \cite{anguita2012human}.
 In that research a multi-class classification problem is solved using \glspl{svm} and the One-Vs-All method.
@@
 
-The inner workings of \glspl{oc-svm} are discussed in detail in Section~\ref{sec:one_class_svm}.
+The inner workings of \glspl{oc-svm} are discussed in detail in \Cref{sec:one_class_svm}.
 

### 2013-11-18 Chapters/Chapter3.tex
Restructure Chapter 5 graphs of data sets

@@
 
-This chapter discusses the concepts and algorithms that will be used as a basis for the proposed method, as discussed in Chapter~\ref{Chapter4}.
+This chapter discusses the concepts and algorithms that will be used as a basis for the proposed method, as discussed in \Cref{Chapter4}.
 The first sections formulates the problem of change detection and relates it to outlier and novelty detection.
 It transforms the traditional problem to change detection for times series data.
-In Section~\ref{sec:one_class_classification} the problem of \acrlong{occ} is discussed and shows a number of implementations.
-Two \acrlong{svm}-based methods, \gls{nu-svm} and \gls{svdd}, are further disussed in Section~\ref{sec:one_class_svm}.
+In \Cref{sec:one_class_classification} the problem of \acrlong{occ} is discussed and shows a number of implementations.
+Two \acrlong{svm}-based methods, \gls{nu-svm} and \gls{svdd}, are further disussed in \Cref{sec:one_class_svm}.
 For the application to the problem of \acrlong{har}, the characteristics of the collected sensor data is discussed in~\ref{sec:sensor_data_characteristics}.
-Finally a theoretic set of quality metrics is discussed in Section~\ref{sec:quality_metrics}.
+Finally a theoretic set of quality metrics is discussed in \Cref{sec:quality_metrics}.
 

### 2013-11-18 Chapters/Chapter3/one_class_classification.tex
Restructure Chapter 5 graphs of data sets

@@
 Since the algorithm's goal is to differentiate between normal and abnormal objects (relative to the training objects), \gls{occ} is often called outlier, novelty or anomaly detection, depending on the origin of the problem to which the algorithm is applied\footnote{The term \acrlong{occ} originates from Moya \etal \cite{moya1993one}.}.
-The difference between two and one-class classification and the consequence for outlier objects is illustrated in Figure~\ref{fig:two-vs-one-classification}.
+The difference between two and one-class classification and the consequence for outlier objects is illustrated in \Cref{fig:two-vs-one-classification}.
 In the two-class classification problem the object \textbf{o} will be member of the \textbf{-1} class whilst the \gls{occ} problem will label it as an outlier.
@@
 \subsection{One-Class Classification methods}\label{subsec:occ-methods}
-The methods and algorithms used for the \gls{occ}-problem can be organized into three categories \cite{tax2001one,noumir2012simple}, visually represented in Figure~\ref{fig:occ-methods}.
+The methods and algorithms used for the \gls{occ}-problem can be organized into three categories \cite{tax2001one,noumir2012simple}, visually represented in \Cref{fig:occ-methods}.
 The first category consists of methods that estimate the density of the training data and set a threshold on this density.
@@
 The number of objects that is required is smaller than in the case of density methods.
-The boundary method \gls{svdd}, constructed by Tax and which has shown good performance \cite{khan2010survey}, will be further discussed in Section~\ref{subsec:oc-svm-svdd}.
+The boundary method \gls{svdd}, constructed by Tax and which has shown good performance \cite{khan2010survey}, will be further discussed in \Cref{subsec:oc-svm-svdd}.
 
@@
 It shows succesful applications for, amonst others, problems in the field of Handwritten Digit Recognition, Face Recognition Applications, Spam Detection and Anomaly Detection \cite{li2003improving,perdisci2006using}.
-As discussed in Section~\ref{sec:change_detection_time_series}, this can be used for change detection in time series data.
+As discussed in \Cref{sec:change_detection_time_series}, this can be used for change detection in time series data.
 

### 2013-11-18 Chapters/Chapter3/one_class_svm.tex
Restructure Chapter 5 graphs of data sets

@@
 A \gls{svm} can create a boundary between linear-separable data points, making it a non-probabilistic binary linear classifier.
-More flexible non-linear boundaries can be obtained by the use of a non-linear function $\phi(x)$, as illustrated in Figure~\ref{fig:kernel_mapping}.
+More flexible non-linear boundaries can be obtained by the use of a non-linear function $\phi(x)$, as illustrated in \Cref{fig:kernel_mapping}.
 This function maps the input data from space $\mathcal{I}$ to a higher dimensional space $\mathcal{F}$.
@@
 When the hyperplane is projected to the (lower) original input space $\mathcal{I}$ it creates a non-linear separating curve.
-This is illustrated in Figure~\ref{fig:svm_mapping_spaces}.
-The mapping and projection of data points can be efficient (and implicit) performed by using the kernel trick, which is discussed in Section~\ref{subsec:kernels}.
+This is illustrated in \Cref{fig:svm_mapping_spaces}.
+The mapping and projection of data points can be efficient (and implicit) performed by using the kernel trick, which is discussed in \Cref{subsec:kernels}.
 
@@
 The data points which lie on the margin are the \emph{support vectors}.
-This geometrical interpretation is illustrated in Figure~\ref{fig:svm_hyperplane}.
+This geometrical interpretation is illustrated in \Cref{fig:svm_hyperplane}.
 All data points for which $y_i = -1$ are on one side of the hyperplane and all other data points (for which $y_i = 1$) are on the other side.
@@
 \TODO{Is this figure obsolete?
-Figure~\ref{fig:kernel_mapping} shows almost the same}}
+\Cref{fig:kernel_mapping} shows almost the same}}
   \label{fig:svm_mapping_spaces}
@@
 \end{equation}
-where $K(x, x_i) = \phi(x)^T\phi(x_i)$ is the dot product of data objects in feature space $\mathcal{F}$ (which is further discussed in Section~\ref{subsec:kernels}).
+where $K(x, x_i) = \phi(x)^T\phi(x_i)$ is the dot product of data objects in feature space $\mathcal{F}$ (which is further discussed in \Cref{subsec:kernels}).
 Here every data point $\mathcal{I}$ for which $a_i > 0$ is weighted in the decision function and thus ``supports'' the classification machine: hence the name ``\acrlong{svm}''.
@@
 In the previous section, the mapping function $\phi(x)$ and the kernel function $K$ were briefly mentioned.
-The decision function in Equation~\ref{eq:svm_lagrange} only relies on the dot products of mapped data points in the feature space $\mathcal{F}$ (\ie all pairwise distances between the data points in that space).
+The decision function in \Cref{eq:svm_lagrange} only relies on the dot products of mapped data points in the feature space $\mathcal{F}$ (\ie all pairwise distances between the data points in that space).
 It shows \cite{flach2012machine} that as long as any function has the same result, without an explicit mapping to the higher dimension $\mathcal{F}$, the dot products can be substituted by the kernel function $K$, as introduced by Aizerman~\etal \cite{aizerman1964theoretical} and applied to \glspl{svc} by Vapnik \cite{vapnik1998statistical}:
@@
 The width parameter $\sigma$ is equal for all basis functions and is set a priori and determines the flexibility and complexity of the boundary.
-In Section~\ref{subsec:svm_model_parameters} this (hyper)parameter for a \gls{svm} is further discussed.
+In \Cref{subsec:svm_model_parameters} this (hyper)parameter for a \gls{svm} is further discussed.
 
@@
 
-The constructed decision function $\mathcal{F}$ resembles the function discussed in Section~\ref{subsec:svm}.
+The constructed decision function $\mathcal{F}$ resembles the function discussed in \Cref{subsec:svm}.
 It returns the value $+1$ in a (possibly small) region capturing most of the data points, and $-1$ elsewhere.
@@
 The hyperplane is represented by $g(x) = w \cdot \phi(x) + \rho = 0$ and the decision function is $f(x) = \operatorname{sgn}(g(x))$.
-This hyperplane and the separation from the origin is illustrated in Figure~\ref{fig:nu-svm}.
+This hyperplane and the separation from the origin is illustrated in \Cref{fig:nu-svm}.
 
@@
 
-In the classical \gls{svm} objective function, as denoted in Equation~\ref{eq:svm_objective}, the parameter $C$ decided the smoothness of the boundary, with respect to the slack variables $\xi_i$.
+In the classical \gls{svm} objective function, as denoted in \Cref{eq:svm_objective}, the parameter $C$ decided the smoothness of the boundary, with respect to the slack variables $\xi_i$.
 In the formulation of \gls{nu-svm} the equivalent parameter is $\nu \in (0,1)$ (hence the name).
@@
 Like the classical \gls{svm} and \gls{svdd} it can be required that all the distances from the data points $x_i$ to the center $\mathbf{a}$ are strict less then $R$ (or equivalent measure, to create a hard margin) or soft margins are allowed by using slack variables $\xi_i$.
-In the case of a soft margin, the penalty is determined by $C$ and the minimization is expressed as Equation~\ref{eq:svdd_objective}.
-This principle is illustrated in the right image or Figure~\ref{fig:nu-vs-svdd}.
+In the case of a soft margin, the penalty is determined by $C$ and the minimization is expressed as \Cref{eq:svdd_objective}.
+This principle is illustrated in the right image or \Cref{fig:nu-vs-svdd}.
 Instead of a separating hyperplane, constructed by \gls{nu-svm} and illustrated on the left of the Figure, the \gls{svdd} creates a hypersphere (in the illustration a cricle) around the data points.
 By using kernel functions (\eg the \gls{rbf}) the hyperspheres in the high dimensional feature space $\mathcal{F}$ corresponds to a flexible and tight enclosing boundary in input space $\mathcal{I}$.
-Possible resulting closed boundaries are illustrated in Figure~\ref{fig:svdd-boundary}.
+Possible resulting closed boundaries are illustrated in \Cref{fig:svdd-boundary}.
 This enclosing boundary is obtained by minimizing the following error function $L$ which contains the volume of the hypersphere and the distance from the boundary to the outlier objects:
@@
 
-In the maximization of Equation~\ref{eq:svdd_lagrange} a large fraction of the multipliers $\alpha_i$ become zero and for a small fraction $\alpha_i > 0$.
+In the maximization of \Cref{eq:svdd_lagrange} a large fraction of the multipliers $\alpha_i$ become zero and for a small fraction $\alpha_i > 0$.
 This small fraction, for which $\alpha_i$ is non-zero, are called the \glspl{sv} and these objects lie on the boundary of the description.
@@
 \end{equation}
-Using this kernel function, the Lagrangian error function $L$ of Equation~\ref{eq:svdd_lagrange} changes to:
+Using this kernel function, the Lagrangian error function $L$ of \Cref{eq:svdd_lagrange} changes to:
 \begin{equation}\label{eq:svdd_lagrange_kernel}
@@
 \end{equation}
-Using Equation~\ref{eq:svdd_test_object}, the following kernel formulation needs to hold for a new object $\vectorsym{z}$ to lie within the hypersphere:
+Using \Cref{eq:svdd_test_object}, the following kernel formulation needs to hold for a new object $\vectorsym{z}$ to lie within the hypersphere:
 \begin{equation}\label{eq:svdd_inequality}
@@
 
-\TODO{Use Figure~\ref{fig:svdd-boundary} or~\ref{fig:svdd-boundary-sigma}?}
+\TODO{Use \Cref{fig:svdd-boundary} or~\ref{fig:svdd-boundary-sigma}?}
 
@@
 In case of a \gls{rbf} kernel, the width parameter $\sigma$ determines the flexibility and complexity of the boundary.
-The value of this parameter greatly determines the outcomes of the algorithm (\eg \gls{svdd}) as illustrated in Figure~\ref{fig:svdd-boundary}.
+The value of this parameter greatly determines the outcomes of the algorithm (\eg \gls{svdd}) as illustrated in \Cref{fig:svdd-boundary}.
 With a small value for the kernel width $\sigma$, each data point will tend to be used as a support vector (for almost all $\alpha_i > 0$) and the \gls{svdd} solution resembles a Parzen density estimation.

### 2013-11-18 Chapters/Chapter4/change_detection.tex
Restructure Chapter 5 graphs of data sets

@@
 In the same manner we expect that when the change is in the oldest data objects of the working set the radius $R$ will decrease, since the data becomes more homogeneous.
-An abstract representation of these expectations is illustrated in Figure~\ref{fig:radius_expectation}.
+An abstract representation of these expectations is illustrated in \Cref{fig:radius_expectation}.
 
@@
 
-The appropriate values for the thresholds are highly depended on the characteristics of the data and the trade-off parameter $C$ (as discussed in Section~\ref{subsec:oc-svm-svdd}).
+The appropriate values for the thresholds are highly depended on the characteristics of the data and the trade-off parameter $C$ (as discussed in \Cref{subsec:oc-svm-svdd}).
 This trade-off parameters regulates the fraction of data objects that will be rejected from the constructed model.
@@
 Since a single (high or low) threshold is set for ratio of the radii, high values are considered to be a change point before the highest point of the peak.
-As illustrated in Figure~\ref{fig:tresholding}, the peak at $t_1$ represents a single change point.
+As illustrated in \Cref{fig:tresholding}, the peak at $t_1$ represents a single change point.
 The region from $t_2$ to $t_3$ is strictly increasing.

### 2013-11-18 Chapters/Chapter4/data_gathering.tex
Add caption titles to tables
Restructure Chapter 5 graphs of data sets

@@
 In this section we briefly discuss the different data gathering methods used for the change detection algorithms and experiments.
-Section~\ref{subsec:data_gathering_artificial} reviews the artificial data sets we will use.
-In Section~\ref{subsec:data_gathering_real_world} an overview of the real-world data sets used is provided.
+\Cref{subsec:data_gathering_artificial} reviews the artificial data sets we will use.
+In \Cref{subsec:data_gathering_real_world} an overview of the real-world data sets used is provided.
 Both sections refer to Chapters~\ref{Chapter5} and~\ref{Chapter6} for more details, respectively.
@@
 Using this data set an objective quality measure over the change detection methods can be obtained and compared.
-All the used data sets are listed and analyzed in Chapter~\ref{Chapter5}.
+All the used data sets are listed and analyzed in \Cref{Chapter5}.
 
@@
 The discovered change point are compared with these annotated change points to give a subjective quality measure.
-In Chapter~\ref{Chapter6} we give a detailed analysis of the performed activities and the recorded data sets.
+In \Cref{Chapter6} we give a detailed analysis of the performed activities and the recorded data sets.
 
@@
   \end{tabulary}
-  \caption{Measured metrics. The set of axis is always the triple (x, y, z) direction.}
+  \caption[Measured metrics]{Measured metrics. The set of axis is always the triple (x, y, z) direction.}
   \label{tab:recorded_metrics}

### 2013-11-18 Chapters/Chapter4/model_construction.tex
Restructure Chapter 5 graphs of data sets

@@
 
-The size of the initial window of data objects has a lower bound determined by the hyperparameter $C$ (Equation~\ref{eq:svdd_objective}).
+The size of the initial window of data objects has a lower bound determined by the hyperparameter $C$ (\Cref{eq:svdd_objective}).
 Because of the equality constraint $\sum_{i=1}^n a_i x_i = 1$ and the box constraint $0 \le x_i \le C$, the number of objects in the working set must be at least $\ceil{\frac{1}{C}}$.
@@
 This re-modeling of the \gls{svm} representation is beneficial for our purpose, since we are only interested in sudden changes between activities.
-The following section will discuss the features extracted from the constructed \gls{oc-svm} model at every loop-step of the algorithm, before Section~\ref{sec:method_change_detection} describes how these features are used to detect change.
+The following section will discuss the features extracted from the constructed \gls{oc-svm} model at every loop-step of the algorithm, before \Cref{sec:method_change_detection} describes how these features are used to detect change.

### 2013-11-18 Chapters/Chapter4/model_features.tex
Restructure Chapter 5 graphs of data sets

@@
 The \gls{isvdd} algorithm creates a spherical \gls{oc-svm} representation of the working set at every step of the algorithm.
-This model is obtained by the minimization of Equation~\ref{eq:svdd_objective}, which incorporates the radius $R$ of the sphere and the distances $\vectorsym{\xi}$ from the outliers to the boundary.
+This model is obtained by the minimization of \Cref{eq:svdd_objective}, which incorporates the radius $R$ of the sphere and the distances $\vectorsym{\xi}$ from the outliers to the boundary.
 We will use the the radius $R$ of the hypersphere as an indication of change.
@@
 Tax and Duin provide a method to optimize the parameters of the \gls{svdd} method, \ie the trade-off parameter $C$ and the \gls{rbf} kernel width $\sigma$.
-This optimization will result in the modification of the radius $R$ of Equation~\ref{eq:svdd_objective} and affects the Lagrangian inequality~(\ref{eq:svdd_inequality}).
+This optimization will result in the modification of the radius $R$ of \Cref{eq:svdd_objective} and affects the Lagrangian inequality~(\ref{eq:svdd_inequality}).
 
@@
 This means that only the radius of the sphere will change during the algorithm.
-We will interpret changes in the radius $R$ with the same continuity assumptions as mentioned in Section~\ref{subsec:kernels}.
+We will interpret changes in the radius $R$ with the same continuity assumptions as mentioned in \Cref{subsec:kernels}.
 This means that when two objects (in this case radius sizes) are close to each other, their original objects are also close.

### 2013-11-18 Chapters/Chapter4/overview.tex
Add notes from gerard
Add caption titles to tables
Restructure Chapter 5 graphs of data sets

@@
     \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{./Figures/chapter4/method_setup.pdf}
-  \caption[Method setup]{Schematic representation of the change detection method.}
+  \caption[Method setup]{Schematic representation of the change detection method. \TODO{Gerard noemt dit misschien ook goed voor bij de inleiding, als motivatie voor onderzoek. Overnemen?}}
   \label{fig:method_overview}
@@
 First described is the method to process the gathered sensor data.
-A schematic overview is given in figure~\ref{fig:method_overview} and shows the steps of the method.
+A schematic overview is given in \Cref{fig:method_overview} and shows the steps of the method.
 A more detailed explanation of the ``Update model'' step follows.
@@
 \subsection{Change detection method}
-As graphically represented in figure~\ref{fig:method_overview}, the change detection method starts by processing the data from sensor, such as the accelerometer, magnetic orientation and rotation measures \footnote{Something about the origin of the streams; all from the same sensor or different sensors?}.
+As graphically represented in \Cref{fig:method_overview}, the change detection method starts by processing the data from sensor, such as the accelerometer, magnetic orientation and rotation measures \footnote{Something about the origin of the streams; all from the same sensor or different sensors?}.
 

### 2013-11-18 Chapters/Chapter5/artificial_data_generation.tex
Restructure Chapter 5 graphs of data sets

--- 2013-10-28 Chapters/Chapter5/artificial_data.tex
@@
 % !TEX root = ../../main.tex
-\section{Artificial Data}\label{sec:artificial_data}
-\TODO{Use data from \cite{camci2010change,takeuchi2006unifying}}
-In this section we present three data sets as used by \cite{camci2010change,takeuchi2006unifying} to provide for a objetive performance comparison.
+\section{Artifical data generation}\label{sec:artificial_data_generation}
+In this section we present four data sets as used by \cite{camci2010change,takeuchi2006unifying} to provide for a objetive performance comparison.
+All the data sets are modeled by a second-order \gls{ar}-model as in \Cref{eq:artificial_data_sets_model}, where $\epsilon_t$ is a Gaussian distribution modeling the noise with mean $0$ and variance $\sigma^2 = 1$, $a_1 = 0.6$ and $a_2 = -0.5$.
+\begin{equation}\label{eq:artificial_data_sets_model}
+  x_t = a_1 x_{t-1} + a_2 x_{t-2} + \epsilon_t,
+\end{equation}
+The length of $x_t$ is $10000$ and change points are generated at each $y \times 10000^\text{th}$ data point, with $y = (1, 2, 3, \dots 9)$.
+
+Using this general \gls{ar}-model, we create four different data sets.
+Each data set is characterized by a change in mean, variance, or both of the Gaussian term $\epsilon_t$:
+\begin{enumerate}
+  \item \textbf{Fixed increasing mean:} at each change point $y$ the mean is increased with $\Delta(y) = 5$.
+  This set origins as the first data set of Camci~\cite{camci2010change}.
+  See \Cref{fig:camci_fixed_increasing_mean,fig:camci_fixed_increasing_mean_thresholds}.
+
+  \item \textbf{Reduced increasing mean:} at each change point $y$ the mean is increased with $\Delta(y) = 10 - y$.
+  This set origins as the first data set of Takeuchi \etal \cite{takeuchi2006unifying}.
+  See \Cref{fig:takeuchi_reduced_increasing_mean,fig:takeuchi_reduced_increasing_mean_thresholds}.
+
+  \item \textbf{Reduced increasing mean, increasing variance:} at each change point $y$ the mean is increased with $\Delta(y) = 10 - y$ and the standard deviation is of $\epsilon_t$ is $0.1 / (0.01 + (10000 - x)/1000)$.
+  This set origins as the second set of Camci~\cite{camci2010change}.
+  See \Cref{fig:camci_reduced_increasing_mean_increasing_variance,fig:camci_reduced_increasing_mean_increasing_variance_thresholds}
+
+  \item \textbf{Alternating variance:} at each change point $y$ the variance $\sigma^2$ is set to $9.0$ when $y$ is odd, and to $1.0$ otherwise.
+  This set origins as the thirds sets from Camci~\cite{camci2010change} and Takeuchi \etal \cite{takeuchi2006unifying}.
+  See \Cref{fig:camci_takeuchi_alternating_variance,fig:camci_takeuchi_alternating_variance_thresholds}.
+\end{enumerate}
+
+% =====
 
@@
 \centering
-  \includegraphics[width=1\textwidth]{./Figures/notes/jumping_mean_camci.eps}
-  \caption[Jumping mean camci]{Jumping mean camci}
+  \includegraphics[width=1\textwidth]{./Figures/notes/camci_fixed_increasing_mean.eps}
+  \caption[Fixed increasing mean]{Fixed increasing mean}
+  \label{fig:camci_fixed_increasing_mean}
 \end{figure}
@@
 \centering
-  \includegraphics[width=1\textwidth]{./Figures/notes/jumping_mean_camci_thresholds.eps}
-  \caption[Jumping mean camci thresholds]{Jumping mean camci, thresholds}
+  \includegraphics[width=1\textwidth]{./Figures/notes/camci_fixed_increasing_mean_thresholds.eps}
+  \caption[Fixed increasing mean, thresholds]{Fixed increasing mean, thresholds}
+  \label{fig:camci_fixed_increasing_mean_thresholds}
+\end{figure}
+
+% =====
+
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/notes/takeuchi_reduced_increasing_mean.eps}
+  \caption[Reduced increasing mean]{Reduced increasing mean}
+  \label{fig:takeuchi_reduced_increasing_mean}
 \end{figure}
@@
 \centering
-  \includegraphics[width=1\textwidth]{./Figures/notes/jumping_mean_camci_ratios_10.eps}
-  \caption[Jumping mean camci ratios]{Jumping mean camci ratios, last 10}
-  \label{fig:jumping_mean_ratios}
+  \includegraphics[width=1\textwidth]{./Figures/notes/takeuchi_reduced_increasing_mean_thresholds.eps}
+  \caption[Reduced increasing mean, thresholds]{Reduced increasing mean, thresholds}
+  \label{fig:takeuchi_reduced_increasing_mean_thresholds}
+\end{figure}
+
+% =====
+
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/notes/camci_reduced_increasing_mean_increasing_variance.eps}
+  \caption[Reduced increasing mean, increasing variance]{Reduced increasing mean, increasing variance.}
+  \label{fig:camci_reduced_increasing_mean_increasing_variance}
 \end{figure}
@@
 \centering
-  \includegraphics[width=1\textwidth]{./Figures/notes/jumping_variance.eps}
-  \caption[Jumping variance]{Jumping variance}
+  \includegraphics[width=1\textwidth]{./Figures/notes/camci_reduced_increasing_mean_increasing_variance_thresholds.eps}
+  \caption[Reduced increasing mean, increasing variance, thresholds]{Reduced increasing mean, increasing variance, thresholds (log scale)}
+  \label{fig:camci_reduced_increasing_mean_increasing_variance_thresholds}
+\end{figure}
+
+% =====
+
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/notes/camci_takeuchi_alternating_variance.eps}
+  \caption[Alternating variance]{Alternating variance}
+  \label{fig:camci_takeuchi_alternating_variance}
 \end{figure}
@@
 \centering
-  \includegraphics[width=1\textwidth]{./Figures/notes/jumping_variance_thresholds.eps}
-  \caption[Jumping variance thresholds]{Jumping variance, thresholds}
+  \includegraphics[width=1\textwidth]{./Figures/notes/camci_takeuchi_alternating_variance_thresholds.eps}
+  \caption[Alternating variance, thresholds]{Alternating variance, thresholds}
+  \label{fig:camci_takeuchi_alternating_variance_thresholds}
 \end{figure}
-
-\begin{figure}
-\centering
-  \includegraphics[width=1\textwidth]{./Figures/notes/jumping_variance_ratios_10.eps}
-  \caption[Jumping variance ratios]{Jumping variance ratios, last 10}
-\end{figure}

### 2013-11-18 intended_contents.tex
Restructure Chapter 5 graphs of data sets

@@
 Outline of the thesis is presented.
-Chapter~\ref{sec:Concepts} introduces the big parts of the puzzle.
-Chapter~\ref{sec:Setup} gives details about the algorithms used.
-Chapter~\ref{sec:Implementation} will give the implementation details.
-Chapter~\ref{sec:Experiments} will show how data is gathered and shows the data.
-Chapter~\ref{sec:Discussion} will discuss and relate the results
-Chapter~\ref{sec:Conclusion} will conclude and give final remarks.
+\Cref{sec:Concepts} introduces the big parts of the puzzle.
+\Cref{sec:Setup} gives details about the algorithms used.
+\Cref{sec:Implementation} will give the implementation details.
+\Cref{sec:Experiments} will show how data is gathered and shows the data.
+\Cref{sec:Discussion} will discuss and relate the results
+\Cref{sec:Conclusion} will conclude and give final remarks.
 
@@
 The focus is on the meaning of the terms and what they do and how they do it.
-Real implementations will be discussed in chapter~\ref{sec:Setup}.
+Real implementations will be discussed in \Cref{sec:Setup}.
 

### 2013-11-18 Chapters/Chapter5/artificial_data_generation.tex
Rename  dir to

@@
 \centering
-  \includegraphics[width=1\textwidth]{./Figures/notes/camci_fixed_increasing_mean.eps}
+  \includegraphics[width=1\textwidth]{./Figures/chapter5/camci_fixed_increasing_mean.eps}
   \caption[Fixed increasing mean]{Fixed increasing mean}
@@
 \centering
-  \includegraphics[width=1\textwidth]{./Figures/notes/camci_fixed_increasing_mean_thresholds.eps}
+  \includegraphics[width=1\textwidth]{./Figures/chapter5/camci_fixed_increasing_mean_thresholds.eps}
   \caption[Fixed increasing mean, thresholds]{Fixed increasing mean, thresholds}
@@
 \centering
-  \includegraphics[width=1\textwidth]{./Figures/notes/takeuchi_reduced_increasing_mean.eps}
+  \includegraphics[width=1\textwidth]{./Figures/chapter5/takeuchi_reduced_increasing_mean.eps}
   \caption[Reduced increasing mean]{Reduced increasing mean}
@@
 \centering
-  \includegraphics[width=1\textwidth]{./Figures/notes/takeuchi_reduced_increasing_mean_thresholds.eps}
+  \includegraphics[width=1\textwidth]{./Figures/chapter5/takeuchi_reduced_increasing_mean_thresholds.eps}
   \caption[Reduced increasing mean, thresholds]{Reduced increasing mean, thresholds}
@@
 \centering
-  \includegraphics[width=1\textwidth]{./Figures/notes/camci_reduced_increasing_mean_increasing_variance.eps}
+  \includegraphics[width=1\textwidth]{./Figures/chapter5/camci_reduced_increasing_mean_increasing_variance.eps}
   \caption[Reduced increasing mean, increasing variance]{Reduced increasing mean, increasing variance.}
@@
 \centering
-  \includegraphics[width=1\textwidth]{./Figures/notes/camci_reduced_increasing_mean_increasing_variance_thresholds.eps}
+  \includegraphics[width=1\textwidth]{./Figures/chapter5/camci_reduced_increasing_mean_increasing_variance_thresholds.eps}
   \caption[Reduced increasing mean, increasing variance, thresholds]{Reduced increasing mean, increasing variance, thresholds (log scale)}
@@
 \centering
-  \includegraphics[width=1\textwidth]{./Figures/notes/camci_takeuchi_alternating_variance.eps}
+  \includegraphics[width=1\textwidth]{./Figures/chapter5/camci_takeuchi_alternating_variance.eps}
   \caption[Alternating variance]{Alternating variance}
@@
 \centering
-  \includegraphics[width=1\textwidth]{./Figures/notes/camci_takeuchi_alternating_variance_thresholds.eps}
+  \includegraphics[width=1\textwidth]{./Figures/chapter5/camci_takeuchi_alternating_variance_thresholds.eps}
   \caption[Alternating variance, thresholds]{Alternating variance, thresholds}

### 2013-11-18 Chapters/Chapter5/results.tex
Add equations

@@
 \section{Results}\label{sec:artificial_data_results}
+For an objective measure of quality of the proposed method applied to the four aforementioned data sets, we employ two metrics, as defined by Takeuchi \etal \cite{takeuchi2006unifying}.
+The first metric is the $\operatorname*{average \_ benefit}$, for which we first need to define the notion of $\operatorname*{benefit}$ for each change:
+\begin{equation}\label{eq:benefit}
+  \operatorname*{benefit}(y) = 1 - |t_y - t_y^*| / 10  \mbox{~~~(when } |t_y - t_y^*| < 10 \mbox{)},
+\end{equation}
+where $t_y$ is the time of the discovered change point and $t_y^*$ is the true change time.
+We can then define the $\operatorname*{average \_ benefit}$ over all the discovered change points $Y$ as:
+\begin{equation}\label{eq:average_benefit}
+  \operatorname*{average \_ benefit}(Y) = \frac{\sum_{y=1}^Y \operatorname*{benefit}(y)}{Y}.
+\end{equation}
+Besides the benefit, which rewards discovered change points close to the actual change points, we need to penalize discovered change points which do not relate to a true change point.
+For this we use the \acrlong{far}, expressed as
+\begin{equation}\label{eq:false_alarm_rate}
+  \operatorname*{far}(Y) = \frac{Y_{\operatorname*{false}}}{Y}
+\end{equation}

### 2013-11-21 Chapters/Chapter3/one_class_classification.tex
Add intro in chapter on OCC

@@
 \section{One-Class Classification}\label{sec:one_class_classification}
+As discussed in the previous section, change detection in time series can be implemented using outlier, or novelty, detection.
+To regard a data point as an outlier, there must be a model of the normal time series data and a (dis)similiarity measure defined over the model and data objects.
+When a data point differs enough from the created model, it can be labeled as an outlier.
+The class of \gls{occ} algorithms is especially designed for that purpose.
+The algorithms build up a model of the data, assuming only normal data objects are available (or a very limited amount of example outlier data objects).
+This differs from classical classification algorithms, which commonly rely of both positive and negative examples.
 
-\TODO{bridge that explains why \gls{occ} is interesting, relating it to outlier/change detection}
+In \Cref{subsec:occ-problem-formulation} the problem formulation of \gls{occ} methods is explained.
+In \cref{subsec:occ-methods} an overview of \gls{occ} methods is given.
+The following section will discuss one specific set of methods, the \gls{oc-svm} which use \glspl{svm} to model the normal training data.
 

### 2013-11-21 Chapters/Chapter3/change_detection_time_series.tex
Restructure Chapter 5 graphs of data sets
Fix TODO note

@@
 \section{Change Detection in Time Series}\label{sec:change_detection_time_series}
-*** The section should relate the problem of outlier detection to change detection.\\
+\TODO{Write section introduction} \\
+\emph{The section should relate the problem of outlier detection to change detection.
 It should transform the traditional problem formulation to that of time series data.
-Build to one-class classification, which can be used for outlier detection. ***
+Build to one-class classification, which can be used for outlier detection}
 
@@
 \subsection{Problem formulation}\label{subsec:change_detection_problem_formulation}
-\TODO{Give the problem of change definition. Follow definition of Camci \cite{camci2010change}}
-
 \TODO{Note: the following block is copied from~\ref{sec:method_change_indication}} \\
@@
 Since we are only interested in different activities (which are represented by sudden changes), slight changes within an activity are not of interest.
-Section~\ref{sec:sensor_data_characteristics} discusses the representation of activities and changes between them in the data in more detail.
+\Cref{sec:sensor_data_characteristics} discusses the representation of activities and changes between them in the data in more detail.
 

### 2013-11-21 Chapters/Chapter3/one_class_classification.tex
WIP

@@
 \subsection{Problem formulation}\label{subsec:occ-problem-formulation}
-\TODO{Introduce problem of \gls{occ}. Relate to two-class classification.}
-
 The problem of \gls{occ} is closely related to the (traditional) two-class classification situation\footnote{Two-class problems are considered as to be the basic problem, since multi-class classification problems can be decomposed into multiple two-class problems \cite{fukunaga1990introduction}.}.
@@
 By constructing a \gls{occ}-classifier using a training data set, one can compare that set to a new data set.
-This will result in a similarity-measure.
-It is easy to see here a relation with similiary measures over \glspl{pdf} and the \gls{kl} divergence. \\
-\TODO{Add reference to~\ref{subsec:change_detection_problem_formulation} which also mentions \gls{kl} divergence?}
+This will result in a similarity-measure expressed in the properties of outliers.
+This measure is related to other similarity measures, such as \gls{pdf} estimation and the \gls{kl} divergence as discussed in \Cref{subsec:change_detection_problem_formulation}.
 

### 2013-11-22 Chapters/Chapter2.tex
'Refactoring'

@@
 
-\section{Overview}\label{sec:literature_review_overview}
-As described in \Cref{Chapter1}, the focus of this research is the application of temporal segmentation to accelerometer data.
-This is with the assumption that explicit temporal segmentation can aid the process of temporal classification.
-
-\TODO{Give examples of temporal classification techniques}
-
-In recent years a lot is research is performed in the classification of human activities, recorded by on-body sensors.
-Especially since smartphones with accelerometer sensors became more available, research has focussed on recordings from these devices.
-
-\TODO{Include highlights from old-classification text}
-
-In the following section of this is chapter we begin with a review of current literature on temporal segmentation.
-Many techniques of temporal segmentation rely in change detection algorithms.
-In \Cref{sec:change_detection_methods} we will give an overview of those kind of algorithms.
-An often encountered problem is the \emph{curse of dimensionality}.
-To overcome this problem, we look at \gls{dr} methods in \Cref{sec:dimensionality_reduction}.
-We end this chapter with a review on \gls{svm} based methods and algorithms.
-This latter section relates the \gls{oc-svm} methods via \gls{dr} to temporal segmentation.
-
+\input{./Chapters/Chapter2/overview.tex}
 \input{./Chapters/Chapter2/temporal_segmentation.tex}
 \input{./Chapters/Chapter2/change_detection_methods.tex}
-\input{./Chapters/Chapter2/dimensionality_reduction.tex}
+% \input{./Chapters/Chapter2/dimensionality_reduction.tex}
 \input{./Chapters/Chapter2/svm_based.tex}

### 2013-11-22 Chapters/Chapter2/cusum.tex
WIP: Working on organization of Chapter 2
'Refactoring'

@@
 % !TEX root = ../../main.tex
-\section{CUSUM}\label{cusum}
+\section{CUSUM}\label{sec:cusum}
 
@@
 The method of Incl\'{a}n and Tiao \cite{inclan1994use} builds on the centered version of \gls{cusum} \cite{brown1975techniques} to detect changes in variance.
-Using the \gls{icss} algorithm they are able to find multiple change points in a reflective manner.
-\TODO{Explain ``reflective'', or use term as batch, post-processing etc.}
+Using the batch \gls{icss} algorithm they are able to find multiple change points, offline while post-processing the data.
 Let $C_k = \sum_{i=1}^k \alpha_t^2$ be the cumulative sum of squares for a series of uncorrelated random variables $\{\alpha_t\}$ of length $T$.

### 2013-11-22 Chapters/Chapter2/overview.tex
'Refactoring'

% !TEX root = ../main.tex
\section{Overview}\label{sec:literature_review_overview}
As described in \Cref{Chapter1}, the focus of this research is the application of temporal segmentation to accelerometer data.
This is with the assumption that explicit temporal segmentation can aid the process of temporal classification.

\TODO{Give examples of temporal classification techniques}

In recent years a lot is research is performed in the classification of human activities, recorded by on-body sensors.
Especially since smartphones with accelerometer sensors became more available, research has focussed on recordings from these devices.

\TODO{Include highlights from old-classification text}

In the following sections of this is chapter we begin with a review of current literature on temporal segmentation of time series data.
Many techniques of temporal segmentation rely on change detection algorithms, implemented as outlier or novelty detection.
In \Cref{sec:change_detection_methods} we will give an overview of algorithms that detect changes in time series data, based on model assumptions over the data.
We end this chapter with a review on \gls{svm} based methods and algorithms, which will serve as a basis of this research.

\TODO{Note to self: methods based on model-fitting do not directly use change points}

\TODO{Plek bedenken waar SVM en Dimensionality Reduction aan elkaar gekoppeld kunnen worden. Ergens in chapter 3?}

### 2013-11-22 Chapters/Chapter2/temporal_segmentation.tex
'Refactoring'

--- 2013-11-01 Chapters/Chapter2/temporal_segmentation.tex
@@
 \section{Temporal Segmentation}\label{sec:temporal_segmentation}
+
+Give overview of used temporal segmentation methods. Problem is that it is not really well-defined.
 
@@
 These will be the categorical terms for which we discuss the literature.
-Finally we will discuss other applications of \glspl{svm} in the context of these terms.
+
+Aspects we will consider when discussing the literature are applicability to \gls{har}, parameter tuning and ... \TODO{What more aspects?}
+% Finally we will discuss other applications of \glspl{svm} in the context of these terms.
 
 \subsection{Segmentation}\label{subsec:segmentation}
-\TODO{This subsection is mainly from previous draft version}
 
-\TODO{Create compact notation, one sentence per paper max}
+Methods that apply temporal segmentation on time series data can be roughly categorized in three different methods, as discussed by Avci \etal \cite{avci2010activity}.
+In that survey the distinction between \emph{Top-Down}, \emph{Bottom-Up}, and \emph{Sliding-Window} approaches is made, based on the way data is processed by the algorithm.
+Since we are interested in on-line change detection, the literature discussed in this section will mainly be forms of sliding-window algorithms.
 
-Segmentation methods can roughly be categorized in three methods in the way the data is processed, as discussed by Avci \etal \cite{avci2010activity}:
-\begin{itemize}
-  \item \textbf{Top-down} methods iteratively divide the signal in segments by splitting at the best location.
-  The algorithm starts with two segments and completes when a certain condition is met, such as when an error value or number of segments $k$ is reached.
-  These methods process the data points recursively, which results in a complexity of $O(n^2k)$.
-  \item \textbf{Bottom-up} methods are the natural complement to top-down methods.
-  They start with creating $n/2$ segments and iteratively join adjacent segments while the value of a cost function for that operation is below a certain value.
-  Given the average segment length $L$, the complexity of this method is $O(nL)$.
-  \item \textbf{Sliding-window} methods are simple and intuitive for online segmenting purposes.
-  It starts with a small initial subsequence of the time series.
-  New data points are joined in the segment until the fit-error is above a threshold.
-  Since the data is only processed very locally, these methods can yield in poor results \cite{keogh2001online}.
-  The complexity is equal to the bottom-up approach, $O(nL)$, where $L$ is the average segment length.
-  \item \textbf{\acrlong{swab}}, as introduced by Keogh \etal \cite{keogh2001online}, joins the ability of the sliding window mechanism to process time series online and the bottom-up approach the create superior segments in terms of fit-error.
-  The algorithm processes the data in two stages.
-  The first stage is to join new data points in the current segment created by a sliding window, and pass this to a buffer with space for a few segments.
-  The buffer then processes the data Bottom-up and returns the first (left-most) segment as final segment.
-  Because this second stage retains some (semi-)global view of the data, the results are comparative with normal Bottom-up.
-  It is stated by Keogh \etal that the complexity of \gls{swab} is a small constant factor worse than that of regular Bottom-up.
-\end{itemize}
-It is clear that for the application of this research sliding-window and preferably \gls{swab}-based algorithms should be considered.
+In Keogh \etal \cite{keogh2001online} a comparison of the aforementioned approaches is made.
+The research also introduces the \gls{swab} method, which combines the simple and intuitive approach of the sliding-window approach with the (semi-)global view over the data from the bottom-up method.
+This proposed method is dependent on user provided parameters, setting the maximum error when performing the algorithm.
+Each segment is approximated by using a \gls{plr}.
+The provided error threshold controls the granularity and number of segments.
 
-The \gls{swab} method proposed by Keogh \etal \cite{keogh2001online} is dependent on an user setting, providing the maximum error when performing both stages.
-Each segment is approximated by using piecewise linear representation (PLR), an often used method.
-The user provided error threshold controls the granularity and number of segments.
 Other methods have been proposed, such as an adaptive threshold based on the signal energy by Guenterberg \etal \cite{guenterberg2009automatic}, the adaptive \gls{cusum}-based test by Alippi \etal \cite{alippi2006adaptive} and the \gls{mosum} by Hsu \cite{hsu2007mosum} in order to eliminate this user-dependency.
@@
 
---- Segmentation ---
+% --- Segmentation ---
 
@@
 
-Section~\ref{sec:change_detection_time_series} discusses the problem of change detection in time series further, and gives a formal problem definition.
+\Cref{sec:change_detection_time_series} discusses the problem of change detection in time series further, and gives a formal problem definition.
 
@@
 
-% ``Human Activity Recognition on Smartphones using a Multiclass Hardware-Friendly Support Vector Machine'' \cite{anguitahuman}. 13, 2012 \\
+% ``Human Activity Recognition on Smartphones using a Multiclass Hardware-Friendly Support Vector Machine'' \cite{anguita2012human}. 13, 2012 \\
 

### 2013-11-22 Chapters/Chapter3.tex
'Refactoring'

@@
 
+\TODO{Give this a nice place in this chapter somewhere, on model building}
+Segmentation methods can roughly be categorized in three methods in the way the data is processed, as discussed by Avci \etal \cite{avci2010activity}:
+\begin{itemize}
+  \item \textbf{Top-down} methods iteratively divide the signal in segments by splitting at the best location.
+  The algorithm starts with two segments and completes when a certain condition is met, such as when an error value or number of segments $k$ is reached.
+  These methods process the data points recursively, which results in a complexity of $O(n^2k)$.
+  \item \textbf{Bottom-up} methods are the natural complement to top-down methods.
+  They start with creating $n/2$ segments and iteratively join adjacent segments while the value of a cost function for that operation is below a certain value.
+  Given the average segment length $L$, the complexity of this method is $O(nL)$.
+  \item \textbf{Sliding-window} methods are simple and intuitive for online segmenting purposes.
+  It starts with a small initial subsequence of the time series.
+  New data points are joined in the segment until the fit-error is above a threshold.
+  Since the data is only processed very locally, these methods can yield in poor results \cite{keogh2001online}.
+  The complexity is equal to the bottom-up approach, $O(nL)$, where $L$ is the average segment length.
+  \item \textbf{\acrlong{swab}}, as introduced by Keogh \etal \cite{keogh2001online}, joins the ability of the sliding window mechanism to process time series online and the bottom-up approach the create superior segments in terms of fit-error.
+  The algorithm processes the data in two stages.
+  The first stage is to join new data points in the current segment created by a sliding window, and pass this to a buffer with space for a few segments.
+  The buffer then processes the data Bottom-up and returns the first (left-most) segment as final segment.
+  Because this second stage retains some (semi-)global view of the data, the results are comparative with normal Bottom-up.
+  It is stated by Keogh \etal that the complexity of \gls{swab} is a small constant factor worse than that of regular Bottom-up.
+\end{itemize}
+It is clear that for the application of this research sliding-window and preferably \gls{swab}-based algorithms should be considered. \\
+--- end ---
+
+\TODO{Put somewhere difference and emphasis of online vs. offline/batch/post-processing}
+
 \input{./Chapters/Chapter3/change_detection_time_series}

### 2013-11-29 Chapters/Chapter3.tex
Outline chapter 7 conclusion

@@
   These methods process the data points recursively, which results in a complexity of $O(n^2k)$.
-  \item \textbf{Bottom-up} methods are the natural complement to top-down methods.
+  \item \textbf{Bottom-up} methods are the natural complement to top-down methods.g
   They start with creating $n/2$ segments and iteratively join adjacent segments while the value of a cost function for that operation is below a certain value.

### 2013-11-29 Chapters/Chapter7.tex
Rough finish of chapter 2
Wip: chapter 3 formulas, chapter 4 model construction and properties
Prefix refs with non-breaking space and capitalize refs
Outline chapter 7 conclusion

@@
 
-\label{Chapter7} % For referencing the chapter elsewhere, use \ref{Chapter7}
+\label{Chapter7} % For referencing the chapter elsewhere, use~\ref{Chapter7}
 
@@
 \section{Outline}
+\emph{Not intented for the reader.}
 \begin{itemize}
-  \item Conclude research
-  \item Future research
+  \item Tie together, integrate, sythesize all the issues from the chapters, reflecting on introductory thesis statements.
+  \item Provide answers to research questions
+  \item Highlight study limitations
+  \item Directions for further research.
+  \item Content: Summarize
+  \begin{itemize}
+    \item What have I researched
+    \item Main arguments
+    \item How researched
+    \item What discovered
+    \item What pre-existing views were challenged (?)
   \end{itemize}
+  \item Overview of
+  \begin{itemize}
+    \item New knowledge
+    \item Significance of research
+    \item Limitations (data, concepts)
+    \item Speculations on implications of research
+    \item Further research and development, or: links with different fields, other methods applied to same data.
+  \end{itemize}
+  \item Must: make clear statement of my original contribution
+  \item Ideally:
+  \begin{itemize}
+    \item Shows links across key ideas spread in chapters
+    \item Show commitment and enthusiasm for research
+    \item Positive impression
+  \end{itemize}
+  \item Avoid:
+  \begin{itemize}
+    \item claim findings that have not been proven
+    \item introducing new data/concepts
+    \item Be self-critical, but dont put weaknesses or limitation in research
+    \item Too long/too short
+  \end{itemize}
+\end{itemize}
+
+Paragraph: into on conclusion; what have I researched and what is my contribution.
+Break into sections further down.
+
+===
+
+In this research we have applied \gls{occ} methods to accelerometer data, recorded during the performance of various human activities, in order to create a temporal segmentation of those recordings.
+The goal was to find the change points between activities, assuming that it will aid better understanding of the sensor values.
+To do so, we used the \gls{svdd} algorithm by Tax~\cite{tax1999support}, following the method of applying it to artificial data as done by Camci~\cite{camci2010change}.
+Instead of using both the information whether a new data point is an outlier and the radius of the hypersphere to indicate change, we only used on the latter model properties.
+The increase of the radius of the constructed hypersphere, which encloses the data objects through a sliding window, is extracted and used for indication of change.
+
+In the following sections we will...
+
+\section{What is done}
+What we have researched and how we did it.
+Repeat research question, justify the research.
+Provide context, background of the topic.
+Indicate the problem (which links to research question and key objectives).
+Explain what is further discussed in the conclusion (sequential map of subjects).
+
+\section{Main findings}
+What we have found, show links across chapters.
+
+For each finding a different paragraph.
+How did we arrive at this finding, and it challanges/improves on previous research.
+Not just a summary, but the effects and implications.
+
+\section{Other research}
+Relate the finding with earlier research.
+Provide similarities and differences.
+Better/worse results?
+
+\section{Limitations}
+Show what limitations were encountered or decided (data collection, scope).
+Note the focus of the limitations and how they worked together (to get positive note).
+This can create bridge to "future research".
+
+\section{Future research}
+What has not been covered? What is worthwhile.
+Show that I am thinking ahead and know the field of study.
+
+\section{Final section}
+Reminding of original contribution and significance.
+This section is the overall conclusion and thus should be very concise and precise.
+
+\TODO{Formulate the research questions in the introduction of thesis, and in the chapters.}

### 2013-12-02 Chapters/Chapter6/data_gathering.tex
WIP on Conclusion and stills of recordings

@@
 \end{figure}
+
+\TODO{Add stills from video recordings, showing the kind of activities performed.}
+
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/stills/jos cp walk-run.png}
+  \caption[]{}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/stills/jos run 2.png}
+  \caption[]{}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/stills/jos run.png}
+  \caption[]{}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/stills/jos sprint.png}
+  \caption[]{}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/stills/jos walking.png}
+  \caption[]{}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/stills/roemer run cw.png}
+  \caption[]{}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/stills/roemer walk.png}
+  \caption[]{}
+\end{figure}

### 2013-12-02 Chapters/Chapter7.tex
WIP on Conclusion and stills of recordings

@@
 \section{Outline}
-\emph{Not intented for the reader.}
+\emph{Not intended for the reader.}
 \begin{itemize}
-  \item Tie together, integrate, sythesize all the issues from the chapters, reflecting on introductory thesis statements.
+  \item Tie together, integrate, synthesize all the issues from the chapters, reflecting on introductory thesis statements.
   \item Provide answers to research questions
@@
 The goal was to find the change points between activities, assuming that it will aid better understanding of the sensor values.
-To do so, we used the \gls{svdd} algorithm by Tax~\cite{tax1999support}, following the method of applying it to artificial data as done by Camci~\cite{camci2010change}.
+To do so, we used the \gls{svdd} algorithm by Tax~\cite{tax1999support}, following the method of applying it to a sliding window of data, as done by Camci~\cite{camci2010change}.
 Instead of using both the information whether a new data point is an outlier and the radius of the hypersphere to indicate change, we only used on the latter model properties.
-The increase of the radius of the constructed hypersphere, which encloses the data objects through a sliding window, is extracted and used for indication of change.
+The increase of the radius of the constructed hypersphere, which encloses the data objects and increases with heterogeneous segments of data, is extracted and used for indication of change.
 
-In the following sections we will...
+\TODO{Write overview of this chapter. \emph{In the following section we will...}}
 
 \section{What is done}
-What we have researched and how we did it.
-Repeat research question, justify the research.
-Provide context, background of the topic.
-Indicate the problem (which links to research question and key objectives).
-Explain what is further discussed in the conclusion (sequential map of subjects).
+The scope of this research was the temporal segmentation of recordings, obtained from human activities.
+This was performed in the wider context of activity classification.
+Currently, many algorithms obtain an implicit segmentation as a side-effect of direct activity classification.
+In this research the primary goal was finding a temporal segmentation, which is assumed to be able to aid the classification step.
+
+We have considered a range of segmentation methods, from which the method by Camci~\cite{camci2010change} showed potential.
+It is based on the detection of outliers in a time series data, assuming that an increase of outliers indicates a change in the underlying generating model.
+For the detection of outliers it uses the \gls{svdd} method by Tax~\cite{tax1999support}.
+It tests every new data point with the constructed model; whether the data point is in or outside the model, and whether the model increases or decreases in radius size.
+The constructed sliding-window algorithm, \gls{svcpd}, is applied to artificial Gaussian noise data.
+
+In this research we have applied the method by Camci~\cite{camci2010change} to inertial signals, recorded by smartphones.
+We have used the accelerometer data for measurement of speed, the gyroscope data for measurement of rotation, and the magnetometer to measure the direction and orientation of the performed activities.
+The constructed method reduces this 9-dimensional signal to a single property, obtained from the constructed model, which can be interpreted to indicate change.
+
+The constructed model follows the implementation of Tax and Laskov~\cite{tax2003online} for an incremental version of \gls{svdd}, \gls{isvdd}.
+This algorithm constructs a \gls{oc-svm} model by processing the data over a sliding window.
+The model represents a hypersphere from which the radius size is abstracted.
+That property is of interest since the assumption is that a heterogeneous window of data will show an increase of radius, in relation to a homogeneous segment of data.
+The homogeneous segment of data will have a relatively small radius since the data points will be close together, resulting from the continuity assumptions.
+
+To test our approach, we have used the same data as used in the experiments from Camci~\cite{camci2010change} and Takeuchi and Yamanishi~\cite{takeuchi2006unifying}.
+Furthermore, since this research is focused around human activities, we have recorded and manually annotated in- and outdoor activities.
+The discovered change points were compared to the annotated change points, obtained from video recordings.
+In this research we found that other public available and wide used common data sets, such as the WISDM~\cite{kwapisz2011activity} and UCI HAR~\cite{anguita2012human}, were not useful for our purpose, since the activities are non-continuously recorded.
+
+\TODO{In the above, refer to the corresponding chapters/sections.}
 
@@
 For each finding a different paragraph.
-How did we arrive at this finding, and it challanges/improves on previous research.
+How did we arrive at this finding, and it challenges/improves on previous research.
 Not just a summary, but the effects and implications.
+
+While temporal segmentation of time series data is not uncommon, the application to inertial signals is currently not under much research attention.
+It is applied in other contexts, such as motion tracking and first person vision sensing, the segmentation obtained from inertial signals is often the implicit result of classification.
 
@@
 \section{Future research}
+The motivation for this research was, amonst others, the assumption that an explicit temporal segmentation of inertial signals can be used to obtain a better classification of the performed activities.
+There are two reasons for that.
+First, since the segmentation adds information about sequential data points (whether they belong to the same class), the a priori knowledge about the distribution changes.
+Second, the window of data used for model construction can be much larger (than the often used length of 1-3 seconds).
+This enables a better mathematical model to be constructed.
+A future research in this line of assumptions is to compare the performance of classification methods using this obtained temporal segmentation and those that do not.
+
+Since the temporal segmentation of inertial signal is not yet a very active field of study, more studies can be performed to obtain it.
+In our approach we have used the raw data directly.
+In the field of activity recognition many methods make use of pre-processing steps, from which one step often is feature extraction from the raw data.
+That, and other pre-processing steps, could be incorporated in the explicit temporal segmentation of inertial signals.
+Furture research could focus on extracted features that aid the detection of change points.
+
+\TODO{Add future research on \gls{oc-svm} applications.}
+
 What has not been covered? What is worthwhile.

### 2013-12-03 Chapters/Chapter2.tex
Cleanup of sections

@@
 \section{Statistical framework}\label{statistical-framework}
-Many applications require the detection of time points at which the underlying properties of a system change.
-This problem has received a lot of attention in the fields of data mining, etc... \TODO{list and refs}.
-Often this problem is formulated in a statistical framework, by inspecting the underlying data generating \gls{pdf} of the time series data.
-A change point is then defined as a significant change in the properties of the \gls{pdf}, such as the mean and variance.
-
-The widely used \gls{cusum} method by Basseville \etal \cite{basseville1993detection} follows this approach.
-It originates from control methods for detection from benchmarks.
-This method and some derivatiges are discussed and analyzed in \Cref{sec:cusum}.
-
-Many methods rely on pre-specified parametric model assumptions and considers the data be independent over time, which makes it less flexible to real-world applications.
-The methods proposed by Kawahara \etal \cite{kawahara2009change} and Lui \etal \cite{liu2013change} try to overcome these problems by estimating the \emph{ratio} between the \glspl{pdf}, instead of estimating each \gls{pdf}.
-This approach is discussed and analyzed in \Cref{sec:density-ratio}.
-
-The density-estimation methods rely on the log-likelihood ratio between \glspl{pdf}.
-The method of Camci \cite{camci2010change} follows an other approach within the statistical framework, by using a \gls{svm}.
-One problem it tries to overcome is the (claimed) weakness of many methods to detect a decrease in variance.
-The method represents the distribution over the data points as a hyper-sphere in a higher dimension using the kernel trick.
-A change in the \gls{pdf} is represented by a change in the radius of this sphere.
-\Cref{sec:svm_in_change_detection} discusses the \gls{svm}-methods.
-\TODO{Put emphasis that this is the source of inspiration for the chosen method?}
-
-% The final method under consideration, change detection via (intrinsic) dimensionality reduction, follows a different point of view.
-% Opposed to the other discussed methods, dimensionality reduction is framed in the MDL (Minimum Description Length) framework.
-% It uses the estimated underlying number of parameters of the time series as a model for change detection.
-% \Cref{dim-reduction} discuses this method.
 
@@
 \end{description}
-
 \input{./Chapters/Chapter2/overview.tex}
+\input{./Chapters/Chapter2/statistical_models.tex}
 \input{./Chapters/Chapter2/temporal_segmentation.tex}
 \input{./Chapters/Chapter2/change_detection_methods.tex}
-% \input{./Chapters/Chapter2/dimensionality_reduction.tex}
-\input{./Chapters/Chapter2/svm_based.tex}
 
@@
 %----------------------------------------------------------------------------------------
-% Temporal Segmentation
-\input{./Chapters/Chapter2/temporal_segmentation_old.tex}
-
-%----------------------------------------------------------------------------------------
 % Support Vector Machines

### 2013-12-03 Chapters/Chapter2/overview.tex
Cleanup of sections

@@
-% !TEX root = ../main.tex
+% !TEX root = ../../main.tex
 \section{Overview}\label{sec:literature_review_overview}
@@
 In recent years a lot is research is performed in the classification of human activities, recorded by on-body sensors.
-Especially since smartphones with accelerometer sensors became more available, research has focussed on recordings from these devices.
+Especially since smartphones with accelerometer sensors became more available, research has focused on recordings from these devices.
 
@@
 
-In the following sections of this is chapter we begin with a review of current literature on temporal segmentation of time series data.
-Many techniques of temporal segmentation rely on change detection algorithms, implemented as outlier or novelty detection.
-In \Cref{sec:change_detection_methods} we will give an overview of algorithms that detect changes in time series data, based on model assumptions over the data.
-We end this chapter with a review on \gls{svm} based methods and algorithms, which will serve as a basis of this research.
+In the following of this chapter we will start with an overview of statistical models used in this field of study.
+That section will act as a starting point to look at the different types, applications and methods of temporal segmentation.
+The following section, \Cref{sec:literature_review_temporal_segmentation} follows the distinction made and looks at a collection of temporal segmentation and change detection methods.
+It will also relate the concept of \gls{occ} to temporal segmentation and change detection.
+The final section of this chapter, \Cref{sec:literature_review_svm}, will discuss a specific class of algorithms for \gls{occ}, namely the methods that use \glspl{svm} for the classification task.
 
-\TODO{Note to self: methods based on model-fitting do not directly use change points}
+% In the following sections of this is chapter we begin with a review of current literature on temporal segmentation of time series data.
+% Many techniques of temporal segmentation rely on change detection algorithms, implemented as outlier or novelty detection.
+% In \Cref{sec:change_detection_methods} we will give an overview of algorithms that detect changes in time series data, based on model assumptions over the data.
+% We end this chapter with a review on \gls{svm} based methods and algorithms, which will serve as a basis of this research.
 
-\TODO{Plek bedenken waar SVM en Dimensionality Reduction aan elkaar gekoppeld kunnen worden. Ergens in chapter 3?}
+% \TODO{Note to self: methods based on model-fitting do not directly use change points}
+
+% \TODO{Plek bedenken waar SVM en Dimensionality Reduction aan elkaar gekoppeld kunnen worden. Ergens in chapter 3?}

### 2013-12-03 Chapters/Chapter2/statistical_models.tex
Cleanup of sections

% !TEX root = ../../main.tex
\section{Statistical models}\label{sec:literature_review_statistical_models}

\TODO{Introduce the diffent models and thus different approaches}

Many applications require the detection of time points at which the underlying properties of a system change.
This problem has received a lot of attention in the fields of data mining, etc... \TODO{list and refs}.
Often this problem is formulated in a statistical framework, by inspecting the underlying data generating \gls{pdf} of the time series data.
A change point is then defined as a significant change in the properties of the \gls{pdf}, such as the mean and variance.

The widely used \gls{cusum} method by Basseville \etal \cite{basseville1993detection} follows this approach.
It originates from control methods for detection from benchmarks.
This method and some derivatiges are discussed and analyzed in \Cref{sec:cusum}.

Many methods rely on pre-specified parametric model assumptions and considers the data be independent over time, which makes it less flexible to real-world applications.
The methods proposed by Kawahara \etal \cite{kawahara2009change} and Lui \etal \cite{liu2013change} try to overcome these problems by estimating the \emph{ratio} between the \glspl{pdf}, instead of estimating each \gls{pdf}.
This approach is discussed and analyzed in \Cref{sec:density-ratio}.

The density-estimation methods rely on the log-likelihood ratio between \glspl{pdf}.
The method of Camci \cite{camci2010change} follows an other approach within the statistical framework, by using a \gls{svm}.
One problem it tries to overcome is the (claimed) weakness of many methods to detect a decrease in variance.
The method represents the distribution over the data points as a hyper-sphere in a higher dimension using the kernel trick.
A change in the \gls{pdf} is represented by a change in the radius of this sphere.
\Cref{sec:literature_review_svm} discusses the \gls{svm}-methods.
\TODO{Put emphasis that this is the source of inspiration for the chosen method?}

-- NEW --

In the field of temporal segmentation of time series data we can distinguish different approaches and common frameworks in which they perform.
The problem is commonly formulated in the statistical framework and regarding the properties of the underlying data generating \gls{pdf} fo the time series data.
In that sense, a change point is then defined as a significant change in the properties of the \gls{pdf}, such as the mean or variance.

In the search for the change in properties, temporal segmentation and change point detection methods can roughly be categorized in three methods in the way the data is processed, as discussed by Avci \etal \cite{avci2010activity}:
\begin{itemize}
  \item \textbf{Top-Down} methods iteratively divide the signal in segments by splitting at the best location.
  The algorithm starts with two segments and completes when a certain condition is met, such as when an error value or number of segments $k$ is reached.
  % These methods process the data points recursively, which results in a complexity of $O(n^2k)$.
  \item \textbf{Bottom-Up} methods are the natural complement to top-down methods.
  They start with creating $n/2$ segments and iteratively join adjacent segments while the value of a cost function for that operation is below a certain value.
  \item \textbf{Sliding-window} methods are simple and intuitive for online segmenting purposes.
  It starts with a small initial subsequence of the time series.
  New data points are joined in the segment until the fit-error is above a threshold.
  Since the data is only processed very locally, these methods can yield in poor results \cite{keogh2001online}.
  \item \textbf{\acrlong{swab}}, as introduced by Keogh \etal \cite{keogh2001online}, combines the ability of the sliding window mechanism to process time series online and the bottom-up approach the create superior segments in terms of fit-error.
  The algorithm processes the data in two stages.
  The first stage is to join new data points in the current segment created by a sliding window.
  The second stage processes the data using Bottom-Up and return the first segment as the final result.
  Since this second stage retains some (semi-)global view of the data, the results are comparative with normal Bottom-Up.
\end{itemize}
It is clear that for the application of this research sliding-window and preferably \gls{swab}-based algorithms should be considered.
In the following we will dicuss classes of algorithms grouped by the type of decision function, assuming a \gls{swab}-based data processing order.

% CUSUM
A widely used class of methods is based on the \gls{cusum} approach by Basseville \etal \cite{basseville1993detection}.
This approach originates from control methods for detection and faults and benchmarking.
In \Cref{subsec:literature_review_temporal_segmentation_cusum} a further overview of this class of methods is discussed.

% Piecewise approximation

-- Density

-- Outlier based

### 2013-12-03 Chapters/Chapter2/temporal_segmentation.tex
WIP on Conclusion and stills of recordings
Cleanup of sections

@@
 % !TEX root = ../../main.tex
-\section{Temporal Segmentation}\label{sec:temporal_segmentation}
+\section{Temporal Segmentation}\label{sec:literature_review_temporal_segmentation}
+
+\TODO{Is there any research of temporal segmentation on accelerometer/inertia signals? Note if there is/isnt}
 
@@
 
-The method of Guenterberg \etal extracts features from the raw sensor signal to base the segmentation on other properties than the pure values.
+The method of Guenterberg \etal \cite{guenterberg2009automatic} extracts features from the raw sensor signal to base the segmentation on other properties than the pure values.
 The method of Bernecker \etal \cite{bernecker2012activity} uses other statistical properties, namely autocorrelation, to distinguish periodic from non-periodic segments.
@@
 
+\subsection{CUSUM}\label{subsec:literature_review_temporal_segmentation_cusum}
+
+\TODO{Link this subsection with previous and next}
+% Papers:
+% \begin{itemize}
+%   \item Use of Cumulative Sums of Squares for Retrospective Detection of Changes of Variance, 1994, 162 refs \cite{inclan1994use} \\
+%   Implemented this algorithm.
+%   \item An adaptive CUSUM-based test for signal change detection, 2006, 15 refs. \cite{alippi2006adaptive}
+%   \item The MOSUM of squares test for monitoring variance changes \cite{hsu2007mosum}
+% \end{itemize}
+
+\emph{Notes:}
+Non-Bayesian change detection algorithm (\ie no prior distribution believe available for the change time).
+The \gls{cusum} method is developed by Page \cite{page1954continuous} for the application of statistical quality control (it is also known as a control chart).
+Primary for detection of mean shift.
+The \gls{mosum} of squares test for monitoring variance changes \cite{hsu2007mosum}.
+Use of Cumulative Sums of Squares for Retrospective Detection of Changes of Variance \cite{inclan1994use}
+
+An often used approach in the statistical framework of change detection is the \gls{cusum} as introduced by Page \cite{page1954continuous}.
+Originally used for quality control in production environments, its main function is to detect change in the mean of measurements and has been applied to this problem \cite{basseville1993detection}.
+It is a non-Bayesian method and thus makes no assumptions, in the form of prior belief distributions, for the change points.
+Many extensions to this method have been proposed.
+Some focus on the change in mean, such the method of Alippi and Roveri \cite{alippi2006adaptive}.
+Others apply the method the problems in which the change of variance is under consideration.
+Among others are there the centered version of the cumulative sums, introduced by Brown, Durbin and Evans \cite{brown1975techniques} and the \gls{mosum} of squares by \cite{hsu2007mosum}.
+
+The method of Incl\'{a}n and Tiao \cite{inclan1994use} builds on the centered version of \gls{cusum} \cite{brown1975techniques} to detect changes in variance.
+Using the batch \gls{icss} algorithm they are able to find multiple change points, offline while post-processing the data.
+Let $C_k = \sum_{i=1}^k \alpha_t^2$ be the cumulative sum of squares for a series of uncorrelated random variables $\{\alpha_t\}$ of length $T$.
+The centered (and normalized) sum of squares is defined as
+\begin{equation}
+  \begin{aligned}
+  D_k = \frac{C_k}{C_T} - \frac{k}{T}, & & k = 1, \dots, T, & & \text{with } D_0 = D_T = 0.
+  \end{aligned}
+\end{equation}
+For a series with homogeneous variance, the value of $D_k$ will oscillate around $0$.
+In case of a sudden change, the value will increase and exceed some predefined boundary with high probability.
+By using an iterative algorithm, the method is able to minimize the masking effect of successive change points.
+
+One of the motivations for the \gls{icss} algorithm was the heavy computational burden involved with Bayesian methods, which need to calculate the posterior odds for the log-likelihood ratio testing.
+The \gls{icss} algorithm avoids applying a function at all possible locations, due to the iterative search.
+The authors recommend the algorithm for analysis of long sequences.
+
 \subsection{Change detection}\label{subsec:change_detection}
@@
 
---- Change detection ---
+\subsection{Novelty and Outlier detection}\label{subsec:novelty_and_outlier_detection}
+In the case of time series data, many approaches rely on the detection of outlier, or novel, data objects.
+An increase in outlier data objects indicates a higher probability of change, as discussed by Takeuchi and Yamanischi~\cite{takeuchi2006unifying}.
+The concepts of novelty detection and outlier detection are closely related and often used interchangeable.
 
-``A unifying framework for detecting outliers and change points from time series'' \cite{takeuchi2006unifying}. 87, 2006 \\
+The algorithm by Ma and Perkins~\cite{markou2003novelty} uses \gls{svr} to model a series of past observations.
+Using the regression, a matching function $V(t_0)$ is constructed which uses the residual value of $t_0$ to create an outlier (out-of-class) confidence value.
+The method defines novel \emph{events} as a series of observations for which the outlier confidence is high enough.
+In an alternative approach by the same authors~\cite{ma2003time} a classification method, in contrast with regression, is used to detect outlier observations.
+The latter approach uses the \gls{nu-svm} algorithm by Sch\"olkopf \etal \cite{scholkopf1999support}, which is further discussed in \Cref{sec:literature_review_svm}
 
-% ``Bayesian online changepoint detection'' \cite{adams2007bayesian}. 85, 2007 \\
-
-% ``An adaptive cusum-based test for signal change detection'' \cite{alippi2006adaptive}. 18, 2006 \\
-
-% ``An efficient algorithm for estimating a change-point'' \cite{cheng2009efficient}. 5, 2009 \\
-
-% ``The marginalized likelihood ratio test for detecting abrupt changes'' \cite{gustafsson1996marginalized}. 80, 1996 \\
-
-% ``The MOSUM of squares test for monitoring variance changes'' \cite{hsu2007mosum}. 4, 2007 \\
-
-% ``Use of cumulative sums of squares for retrospective detection of changes of variance'' \cite{inclan1994use}. 643, 1994 \\
-
-% ``Change-point detection in time-series data by direct density-ratio estimation'' \cite{kawahara2009change}. 52, 2009 \\
-``Sequential change-point detection based on direct density-ratio estimation'' \cite{kawahara2012sequential}. 22, 2012 \\
-% ``Change-point detection in time-series data by relative density-ratio estimation'' \cite{liu2013change}. 11, 2013 \\
-
-``Change point detection in time series data using support vectors'' \cite{camci2010change}. 3, 2010 \\
-
-\subsection{Novelty detection}\label{subsec:novelty_detection}
-
---- Novelty detection ---
-
-``Online novelty detection on temporal sequences'' \cite{ma2003online}. 146, 2003 \\
-``Time-series novelty detection using one-class support vector machines'' \cite{ma2003time}. 78, 2003 \\
-
-``Novelty detection: a review—part 1: statistical approaches'' \cite{markou2003novelty}. 697, 2003 \\
-
-``Support Vector Method for Novelty Detection'' \cite{scholkopf1999support}. 337, 1999 \\
-
-\subsection{Outlier detection}\label{subsec:outlier_detection}
-
---- Outlier detection ---
-``A unifying framework for detecting outliers and change points from time series'' \cite{takeuchi2006unifying}. 87, 2006 \\
-
-``Outliers in statistical data'' \cite{barnett1994outliers} (book). 3745, 1994 \\
-
-``A survey of outlier detection methodologies'' \cite{hodge2004survey}. 791, 2004 \\
-
-``Local outlier detection reconsidered: a generalized view on locality with applications to spatial, video, and network outlier detection'' \cite{schubert2012local}. 2, 2012 \\
-
-% --- SVMs ---
-
-% ``Change point detection in time series data using support vectors'' \cite{camci2010change}. 3, 2010 \\
-
-% ``Time-series novelty detection using one-class support vector machines'' \cite{ma2003time}. 78, 2003 \\
-
-% ``Support vector domain description'' \cite{tax1999support}. 907, 1999
-% ``Support vector data description applied to machine vibration analysis'' \cite{tax1999supportdata}. 76, 1999 \\
-
-% ``Human Activity Recognition on Smartphones using a Multiclass Hardware-Friendly Support Vector Machine'' \cite{anguita2012human}. 13, 2012 \\
-
-% ``Support vector machines in remote sensing: A review'' \cite{mountrakis2011support}. 150, 2011 \\
-
-% ``Sensor-based abnormal human-activity detection'' \cite{yin2008sensor}. 74, 2008 (builds one-class SVM of all normal traces) \\
+A extensive overview of novelty detection based on statistical models is given by Markou and Singh~\cite{markou2003novelty}.
+In the second part of their review~\cite{markou2003novelty-neural}, novelty detection by a variety of neural networks is discussed.
+The survey by Hodge and Austin~\cite{hodge2004survey} distinguishes between three types of outlier detection: \begin{inparaenum}[1)]
+    \item unsupervised,
+    \item supervised, and
+    \item semi-supervised learning.
+  \end{inparaenum}
+As we will see in \Cref{sec:one_class_classification}, in this research we are interested in methods from Type 3, from which \gls{occ} is a member.
+More theoretical background and foundation on outliers in statistical data can be found in \cite{barnett1994outliers}.

### 2013-12-03 Chapters/Chapter3/change_detection_time_series.tex
Cleanup of sections

@@
 \subsection{Problem formulation}\label{subsec:change_detection_problem_formulation}
-\TODO{Note: the following block is copied from~\ref{sec:method_change_indication}} \\
 The problem of change point detection can be formulated using different type of models, as discussed in~\ref{subsec:segmentation}.

### 2013-12-03 Chapters/Chapter3/one_class_classification.tex
Cleanup of sections

@@
 \section{One-Class Classification}\label{sec:one_class_classification}
+\TODO{Add ref to \cite{hodge2004survey}, which states 3 types of outlier detection. It puts \gls{occ} as of Type 3: \textbf{semi-supervised recognition or detection}}
+
 As discussed in the previous section, change detection in time series can be implemented using outlier, or novelty, detection.

### 2013-12-03 Chapters/Chapter6/data_gathering.tex
Crop still recording images and add description
Cleanup of sections

--- 2013-11-11 Chapters/Chapter6/data_gathering.tex
@@
 \end{figure}
+
+\TODO{Add stills from video recordings, showing the kind of activities performed.}
+
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/stills/jos_cp_walk-run.png}
+  \caption[Recording still 1]{Subject 1 on the transition from walking to running}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/stills/jos_run_2.png}
+  \caption[Recording still 2]{Subject 1 running}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/stills/jos_run.png}
+  \caption[Recording still 3]{Subject 1 running}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/stills/jos_sprint.png}
+  \caption[Recording still 4]{Subject 1 sprinting}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/stills/jos_walking.png}
+  \caption[Recording still 5]{Subject 1 walking}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/stills/roemer_run_cw.png}
+  \caption[Recording still 6]{Subject 2 running in clockwise corner}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/stills/roemer_walk.png}
+  \caption[Recording still 7]{Subject 2 walking}
+\end{figure}

### 2013-12-09 Chapters/Chapter6.tex
WIP
WIP: Working on organization of Chapter 2

@@
 
+This chapter discusses the real-world data sets used for this research.
+In \Cref{sec:data_sets} we discuss earlier used data sets and argue that these are inappropriate for our research.
+It is followed by an overview of the setup we have created to record continuous data sets from an uncontrolled environment, both in- and outdoor.
+The activities performed are discussed in \Cref{sec:data_gathering}.
+As discussed in \Cref{Chapter4}, the algorithm is subject to user parameters.
+In \Cref{sec:real_world_parameters} the effect and ranges of the parameters on the real world data sets are discussed.
+Finally, the performance of the algorithm is discussed in \Cref{sec:real_world_performance}.
+
+\input{./Chapters/Chapter6/data_sets}
 \input{./Chapters/Chapter6/data_gathering}
+\input{./Chapters/Chapter6/parameters}
+\input{./Chapters/Chapter6/performance}

### 2013-12-09 Chapters/Chapter6/data_gathering.tex
WIP

@@
 \section{Data Gathering}\label{sec:data_gathering}
+For our data sets of real world activities we have sampled both in- and outdoor activites.
+The subjects performing the activities are three males, in an age around 25.
+We have used two series of recordings.
+The first series are various outdoor recordings with segments of walking, running and standing.
+Some movement activities are performed in a straight line, whilst others are performed around corners of in a wide circular direction.
+The second series consist of mainly walking up and down the stairs, with small vertical parts between the segments.
+The stairs are all in a circular shape.
 
-\TODO{Use common data sets. The WISDM dataset is not useful because it has no transitions, only very clear blocks of the same activities. Include a plot to illustrate the problem?}
+\subsection{Outside}\label{subsec:outside}
 
-\begin{figure}\label{fig:wisdm_excerpt}
-\centering
-  \includegraphics[width=1\textwidth]{./Figures/Chapter6/data_collection/wisdm_excerpt.eps}
-  \caption[WISDM Excerpt]{Excerpt from the WISDM data set \cite{kwapisz2011activity}.
-First five activities for subject $33$ are displayed.
-Due to the discontinuous nature of the recordings, this data set can not be used for change detection.}
-\end{figure}
+\TODO{For each data set: discovered timepoint with acc, mag, rot, and all combined. Use closest change point}
 
-\begin{figure}\label{fig:uci_annotated}
-\centering
-  \includegraphics[width=1\textwidth]{./Figures/Chapter6/data_collection/uci_annotated.eps}
-  \caption[UCI HAR Excerpt]{Excerpt from the UCI HAR data set \cite{anguita2012human}.
-The recorded training activities for subject $25$ are displayed.
-Due to the imprecise (and possibly incorrect) labeling and the short duration of activities, the data set can not be used for the change detection method as described in this thesis.}
-\end{figure}
+\begin{center}\begin{table}
+  \begin{tabulary}{\textwidth}{|c|c|c|}
+    \hline
+    \multicolumn{3}{|c|}{Subj 1, run 1} \\
+    \hline \hline
+    Act. & Ann. & Disc. \\
+    \hline
+    Still & 0 & *0* \\
+    \hline
+    Shake & 3.5 & *3.5* \\
+    \hline
+    Still & 4.2 & *4.2* \\
+    \hline
+    In pocket & 5.5 & *5.5* \\
+    \hline
+    Walk & 7 & *7* \\
+    \hline
+    Run & 10 & *10* \\
+    \hline
+    Walk & 14 & *14* \\
+    \hline
+    Run & 17.2 & *17.2* \\
+    \hline
+    Walk & 22.1 & *22.1* \\
+    \hline
+    Turn CCW & 23.5 & *23.5* \\
+    \hline
+    Walk & 24.2 & *24.2* \\
+    \hline
+    Run (sprint) & 28 & *28* \\
+    \hline
+    Run & 31 & *31* \\
+    \hline
+    Walk & 33 & *33* \\
+    \hline
+    Run & 35.5 & *35.5* \\
+    \hline
+    Walk & 39 & *39* \\
+    \hline
+    Out pocket & 41 & *41* \\
+    \hline
+    Still & 43 & *43* \\
+    \hline
+  \end{tabulary}
+  \caption[Performed activities subject 1 run 1]{The outdoor series performed activities by subject 1, run 1.}
+  \label{tab:outdoor_series}
+\end{table}\end{center}
+
+\begin{center}\begin{table}
+  \begin{tabulary}{\textwidth}{|c|c|c|}
+    \hline
+    \multicolumn{3}{|c|}{Subj 2, run 1} \\
+    \hline \hline
+    Act. & Ann. & Disc. \\
+    \hline
+    Still & 0 & *0* \\
+    \hline
+    Shake & 4 & *4* \\
+    \hline
+    Still & 5 & *5* \\
+    \hline
+    In pocket & 8.8 & *8.8* \\
+    \hline
+    Walk & 12 & *12* \\
+    \hline
+    Run & 16 & *16* \\
+    \hline
+    Walk & 21.3 & *21.3* \\
+    \hline
+    Run & 24.5 & *24.5* \\
+    \hline
+    Walk & 28.3 & *28.3* \\
+    \hline
+    Turn CCW & 30 & *30* \\
+    \hline
+    Walk & 31.8 & *31.8* \\
+    \hline
+    Run & 32.8 & *32.8* \\
+    \hline
+    Walk & 37 & *37* \\
+    \hline
+    Run (sprint) & 39.5 & *39.5* \\
+    \hline
+    Run & 42 & *42* \\
+    \hline
+    Walk & 44.5 & *44.5* \\
+    \hline
+    Out pocket & 49.5 & *49.5* \\
+    \hline
+    Still & 52 & *52* \\
+    \hline
+    Shake & 55.7 & *55.7* \\
+    \hline
+    Still & 56.8 & *56.8* \\
+    \hline
+  \end{tabulary}
+  \caption[Performed activities subject 2 run 1]{The outdoor series performed activities by subject 2, run 1.}
+  \label{tab:outdoor_series}
+\end{table}\end{center}
+
+\TODO{Replace all times in the stars with the closest discovered change point}
 
@@
   \caption[R1: rotation]{Run 1: Walk-run-roemer, rotation}
+  \label{fig:data_gathering_run_1_rot}
 \end{figure}
@@
   \caption[R1: mag]{Run 1: Walk-run-roemer, Mag}
+  \label{fig:data_gathering_run_1_mag}
 \end{figure}
@@
   \caption[R1: accelerometer]{Run 1: Walk-run-roemer, accelerometer}
+  \label{fig:data_gathering_run_1_ann}
 \end{figure}
@@
   \caption[R2: rotation]{Run 2: Walk-run-jos, rotation}
+  \label{fig:data_gathering_run_2_rot}
 \end{figure}
@@
   \caption[R2: mag]{Run 2: Walk-run-jos, Mag}
+  \label{fig:data_gathering_run_2_mag}
 \end{figure}
@@
   \caption[R2: accelerometer]{Run 2: Walk-run-jos, accelerometer}
+  \label{fig:data_gathering_run_2_ann}
+\end{figure}
+
+%---- Run 3: Stairs remco
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/stairs-1-marc/data_plot_rot_annotated.eps}
+  \caption[R3: rotation]{Run 3: Indoor stairs Marc, rotation}
+  \label{fig:data_gathering_run_3_rot}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/stairs-1-marc/data_plot_mag_annotated.eps}
+  \caption[R3: mag]{Run 3: Indoor stairs Marc, Mag}
+  \label{fig:data_gathering_run_3_mag}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/stairs-1-marc/data_plot_acc_annotated.eps}
+  \caption[R3: accelerometer]{Run 3: Indoor stairs Marc, accelerometer}
+  \label{fig:data_gathering_run_3_ann}
 \end{figure}
@@
 \TODO{Add stills from video recordings, showing the kind of activities performed.}
+
+Stills from subject 3:
 
@@
   \caption[Recording still 1]{Subject 1 on the transition from walking to running}
+  \label{fig:data_gathering_still_1_still_walk_to_run}
 \end{figure}
@@
   \caption[Recording still 2]{Subject 1 running}
+  \label{fig:data_gathering_still_1_running}
 \end{figure}
@@
   \caption[Recording still 3]{Subject 1 running}
+  \label{fig:data_gathering_still_1_running_2}
 \end{figure}
@@
   \caption[Recording still 4]{Subject 1 sprinting}
+  \label{fig:data_gathering_still_1_sprint}
 \end{figure}
@@
   \caption[Recording still 5]{Subject 1 walking}
+  \label{fig:data_gathering_still_1_walk}
 \end{figure}
+
+Stills from subject 2:
 
@@
   \caption[Recording still 6]{Subject 2 running in clockwise corner}
+  \label{fig:data_gathering_still_2_run}
 \end{figure}
@@
   \caption[Recording still 7]{Subject 2 walking}
+  \label{fig:data_gathering_still_2_walk}
 \end{figure}
+
+Stills from subject 3:
+
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/stills/marc_downstairs.png}
+  \caption[Recording still 8]{Subject 3 descencing the stairs.}
+  \label{fig:data_gathering_still_3_descending}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/stills/marc_stairs_up_walk.png}
+  \caption[Recording still 9]{Subject 3 will start walking for a small fragment and will continue ascending the stairs.}
+  \label{fig:data_gathering_still_3_ascending}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/stills/marc_walk.png}
+  \caption[Recording still 10]{Subject 3 walking in the hallway. He will turn $180^{\circ}$ counter-clockwise.}
+  \label{fig:data_gathering_still_3_walk}
+\end{figure}

### 2013-12-09 Chapters/Chapter6/data_sets.tex
WIP

% !TEX root = ../../main.tex
\section{Data Sets}\label{sec:data_sets}
In the previous chapter we have applied our method to the artificial data sets from \cite{camci2010change,takeuchi2006unifying}.
In this chapter we apply the method to real-world data sets.
The data sets are inertial sensor data recorded during human activities, such as walking, running, ascending and descending stairs, and standing still.
To compare our results with other research, we have tried to apply our method to two commonly used data sets.
The first is the data from the \gls{wisdm} lab \cite{kwapisz2011activity}.
As an excerpt from the data in \Cref{fig:wisdm_excerpt} shows, the activities are recorded non-continuous.
Whilst this data set is convenient for activity classification, it does not allow for detection of change points.
The other commonly used data set is the \gls{uci-har} from the Machine Learning Repository and originates from \cite{anguita2012human}.
As shown in an except in \Cref{fig:uci_annotated}, the labeling to the data segments seems to be incorrect.
Furthermore, it is not clear how the data is recorded and whether all activities are performed continuously.

\begin{figure}
\centering
  \includegraphics[width=1\textwidth]{./Figures/Chapter6/data_collection/wisdm_excerpt.eps}
  \caption[WISDM Excerpt]{Excerpt from the WISDM data set \cite{kwapisz2011activity}. First five activities for subject $33$ are displayed. Due to the discontinuous nature of the recordings, this data set can not be used for change detection.}
  \label{fig:wisdm_excerpt}
\end{figure}

\begin{figure}
\centering
  \includegraphics[width=1\textwidth]{./Figures/Chapter6/data_collection/uci_annotated.eps}
  \caption[UCI HAR Excerpt]{Excerpt from the UCI HAR data set \cite{anguita2012human}. The recorded training activities for subject $25$ are displayed. Due to the imprecise (and possibly incorrect) labeling and the short duration of activities, the data set can not be used for the change detection method as described in this thesis.}
  \label{fig:uci_annotated}
\end{figure}

Because of the above described shortcomings of commonly used data sets, we have decided to record activities by ourselves.
The following characterizations were set for our data sets:
\begin{enumerate}
  \item The activities need to be performed and recorded in a continuous matter,
  \item The environment should be natural and uncontrolled, both in- and outdoor,
  \item There should be a clear annotation possible, as objective as possible,
  \item The hardware and software should be widely available.
\end{enumerate}

With these requirements in mind, we created the following setup.
All activities were recorded by one smartphone in the right front pants pocket of the subject.
The phones that were used are the \textsc{HTC Sensation XE} and \textsc{HTC Desire}, running the \textsc{Anroid} smartphone operating system.
During the activities the intertial sensor data is recorded using a free application, \textsc{Sensor Logger}~\cite{sensorlogger}.
This application records the sensor data to \textsc{CSV} files, which are then transformed to \textsc{MATLAB} compatible files.
The constructed algorithm processes the data in an online matter (using only current and historic data).
During the activities, the subject is recorded with a video recorder to enable manual labeling and determining the change points.

Using the video recordings, the data sets are manually annotated, as shown in \Cref{fig:data_gathering_run_1_rot}.
The graph displays from a single recording the rotational-axis values (see \Cref{tab:recorded_metrics} for explanation of the metrics), with manually annotated activities above the data.
The manually determines change points are used to draw quality conclusion over the algorithms performance, by comparing it to the discovered change points.
All used data recordings, including the video material and manual labeling, are made available \cite{vlasveld2013continuous}.
\TODO{Create web page with data sets}
The following section discusses the performed activities during the recordings.

### 2013-12-09 Chapters/Chapter6/data_gathering.tex
WIP: Chapter 6

@@
 \section{Data Gathering}\label{sec:data_gathering}
-For our data sets of real world activities we have sampled both in- and outdoor activites.
+For our data sets of real world activities we have sampled both in- and outdoor activities.
 The subjects performing the activities are three males, in an age around 25.
@@
 
-\subsection{Outside}\label{subsec:outside}
+\subsection{Outdoor straight lines}\label{subsec:outdoor_straight}
+The first series of outdoor performed activities consist of walking and running segments in a straight line.
+Halfway the recording is a $180^{\circ}$ turn.
+The activities consist of standing still, walking and running (both jogging and sprinting).
+The plots of \Cref{fig:data_gathering_run_1_rot,fig:data_gathering_run_1_acc,fig:data_gathering_run_1_mag} shows the recordings of Subject $1$.
+For Subject $2$ the plots are visible in \Cref{fig:data_gathering_run_2_rot,fig:data_gathering_run_2_acc,fig:data_gathering_run_2_mag}.
+The recorded metrics are from the accelerometer, rotational and magnetometer sensors.
+The solid black lines indicate manually annoted change points, the dashes lines indicate the closest discovered change point.
+A still from the video recording, showing Subject $2$ performing a (sprinting) run segment, is in \Cref{fig:data_gathering_still_1_sprint}.
 
+\TODO{Add found change points to plots}
+
+%---- Run 1: walk-run-roemer
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/run-1-walk-run-roemer/data_plot_rot_annotated.eps}
+  \caption[R1: rotation]{Run 1: Walk-run-roemer, rotation}
+  \label{fig:data_gathering_run_1_rot}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/run-1-walk-run-roemer/data_plot_mag_annotated.eps}
+  \caption[R1: mag]{Run 1: Walk-run-roemer, Mag}
+  \label{fig:data_gathering_run_1_mag}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/run-1-walk-run-roemer/data_plot_acc_annotated.eps}
+  \caption[R1: accelerometer]{Run 1: Walk-run-roemer, accelerometer}
+  \label{fig:data_gathering_run_1_acc}
+\end{figure}
+
+%---- Run 2: walk-run-jos
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/run-2-walk-run-jos/data_plot_rot_annotated.eps}
+  \caption[R2: rotation]{Run 2: Walk-run-jos, rotation}
+  \label{fig:data_gathering_run_2_rot}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/run-2-walk-run-jos/data_plot_mag_annotated.eps}
+  \caption[R2: mag]{Run 2: Walk-run-jos, Mag}
+  \label{fig:data_gathering_run_2_mag}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/run-2-walk-run-jos/data_plot_acc_annotated.eps}
+  \caption[R2: accelerometer]{Run 2: Walk-run-jos, accelerometer}
+  \label{fig:data_gathering_run_2_acc}
+\end{figure}
+
+%=== STILLS
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/stills/jos_sprint.png}
+  \caption[Recording still 4]{Subject 1 sprinting}
+  \label{fig:data_gathering_still_1_sprint}
+\end{figure}
+
+\subsection{Outdoor straight lines results}
 \TODO{For each data set: discovered timepoint with acc, mag, rot, and all combined. Use closest change point}
@@
 \begin{center}\begin{table}
-  \begin{tabulary}{\textwidth}{|c|c|c|}
+  \begin{tabulary}{\textwidth}{|l|c|c|}
     \hline
@@
     \hline
+    \hline
+    \multicolumn{2}{|l|}{Ratio Total CPs} & 3.2 \\
+    \hline
+    \multicolumn{2}{|l|}{Sum of differences} & 12.341 \\
+    \hline
   \end{tabulary}
@@
 
-\begin{center}\begin{table}
-  \begin{tabulary}{\textwidth}{|c|c|c|}
-    \hline
-    \multicolumn{3}{|c|}{Subj 2, run 1} \\
-    \hline \hline
-    Act. & Ann. & Disc. \\
-    \hline
-    Still & 0 & *0* \\
-    \hline
-    Shake & 4 & *4* \\
-    \hline
-    Still & 5 & *5* \\
-    \hline
-    In pocket & 8.8 & *8.8* \\
-    \hline
-    Walk & 12 & *12* \\
-    \hline
-    Run & 16 & *16* \\
-    \hline
-    Walk & 21.3 & *21.3* \\
-    \hline
-    Run & 24.5 & *24.5* \\
-    \hline
-    Walk & 28.3 & *28.3* \\
-    \hline
-    Turn CCW & 30 & *30* \\
-    \hline
-    Walk & 31.8 & *31.8* \\
-    \hline
-    Run & 32.8 & *32.8* \\
-    \hline
-    Walk & 37 & *37* \\
-    \hline
-    Run (sprint) & 39.5 & *39.5* \\
-    \hline
-    Run & 42 & *42* \\
-    \hline
-    Walk & 44.5 & *44.5* \\
-    \hline
-    Out pocket & 49.5 & *49.5* \\
-    \hline
-    Still & 52 & *52* \\
-    \hline
-    Shake & 55.7 & *55.7* \\
-    \hline
-    Still & 56.8 & *56.8* \\
-    \hline
-  \end{tabulary}
-  \caption[Performed activities subject 2 run 1]{The outdoor series performed activities by subject 2, run 1.}
-  \label{tab:outdoor_series}
-\end{table}\end{center}
-
 \TODO{Replace all times in the stars with the closest discovered change point}
 
-%---- Run 1: walk-run-roemer
-\begin{figure}
-\centering
-  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/run-1-walk-run-roemer/data_plot_rot_annotated.eps}
-  \caption[R1: rotation]{Run 1: Walk-run-roemer, rotation}
-  \label{fig:data_gathering_run_1_rot}
-\end{figure}
-
-\begin{figure}
-\centering
-  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/run-1-walk-run-roemer/data_plot_mag_annotated.eps}
-  \caption[R1: mag]{Run 1: Walk-run-roemer, Mag}
-  \label{fig:data_gathering_run_1_mag}
-\end{figure}
-
-\begin{figure}
-\centering
-  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/run-1-walk-run-roemer/data_plot_acc_annotated.eps}
-  \caption[R1: accelerometer]{Run 1: Walk-run-roemer, accelerometer}
-  \label{fig:data_gathering_run_1_ann}
-\end{figure}
-
-%---- Run 2: walk-run-jos
-\begin{figure}
-\centering
-  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/run-2-walk-run-jos/data_plot_rot_annotated.eps}
-  \caption[R2: rotation]{Run 2: Walk-run-jos, rotation}
-  \label{fig:data_gathering_run_2_rot}
-\end{figure}
-
-\begin{figure}
-\centering
-  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/run-2-walk-run-jos/data_plot_mag_annotated.eps}
-  \caption[R2: mag]{Run 2: Walk-run-jos, Mag}
-  \label{fig:data_gathering_run_2_mag}
-\end{figure}
-
-\begin{figure}
-\centering
-  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/run-2-walk-run-jos/data_plot_acc_annotated.eps}
-  \caption[R2: accelerometer]{Run 2: Walk-run-jos, accelerometer}
-  \label{fig:data_gathering_run_2_ann}
-\end{figure}
-
-%---- Run 3: Stairs remco
-\begin{figure}
-\centering
-  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/stairs-1-marc/data_plot_rot_annotated.eps}
-  \caption[R3: rotation]{Run 3: Indoor stairs Marc, rotation}
-  \label{fig:data_gathering_run_3_rot}
-\end{figure}
-
-\begin{figure}
-\centering
-  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/stairs-1-marc/data_plot_mag_annotated.eps}
-  \caption[R3: mag]{Run 3: Indoor stairs Marc, Mag}
-  \label{fig:data_gathering_run_3_mag}
-\end{figure}
-
-\begin{figure}
-\centering
-  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/stairs-1-marc/data_plot_acc_annotated.eps}
-  \caption[R3: accelerometer]{Run 3: Indoor stairs Marc, accelerometer}
-  \label{fig:data_gathering_run_3_ann}
-\end{figure}
-
-\TODO{Add stills from video recordings, showing the kind of activities performed.}
-
-Stills from subject 3:
+% ===== OUTDOOR FREE ACTIVITIES ========
+\subsection{Outdoor free}\label{subsec:outdoor_free}
+The second series of outdoor recordings consist of more free-form activities.
+The performed activities consist, as with the first set, of standing, walking, and running.
+Furthermore, the segments are performed in a circular manner.
+\eg the running activities are performed around a fountain with an estimated diameter of 10 meters.
+The plots of the recordings of Subject, for the accelerometer, rotation, and magnetometer are shown in...
+\TODO{Add plots}
+For Subject $2$ the still of the video recordings are in \Cref{fig:data_gathering_still_1_still_walk_to_run,fig:data_gathering_still_1_running,fig:data_gathering_still_1_running_2,fig:data_gathering_still_1_walk}.
+The stills for Subject $1$ are in \Cref{fig:data_gathering_still_2_run,fig:data_gathering_still_2_walk}.
 
@@
   \label{fig:data_gathering_still_1_running_2}
-\end{figure}
-
-\begin{figure}
-\centering
-  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/stills/jos_sprint.png}
-  \caption[Recording still 4]{Subject 1 sprinting}
-  \label{fig:data_gathering_still_1_sprint}
 \end{figure}
@@
 
-Stills from subject 3:
+\subsection{Indoor stairs}\label{subsec:indoor_stairs}
+The last series of activities recorded consist of walking up and down the stairs, indoor.
+The stairs are shaped in an semi-circular form, with a small vertical segment on each floor.
+The still of \Cref{fig:data_gathering_still_3_descending} illustrates the circular shape of the stairs.
+In \Cref{fig:data_gathering_still_3_ascending} the small flat segment is visible.
+Between the stairs the subject has walked in a hallway with a $180^{\circ}$ counter-clockwise turn, as shown in \Cref{fig:data_gathering_still_3_walk}.
+
+%---- Run 3: Stairs remco
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/stairs-1-marc/data_plot_rot_annotated.eps}
+  \caption[R3: rotation]{Run 3: Indoor stairs Marc, rotation}
+  \label{fig:data_gathering_run_3_rot}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/stairs-1-marc/data_plot_mag_annotated.eps}
+  \caption[R3: mag]{Run 3: Indoor stairs Marc, Mag}
+  \label{fig:data_gathering_run_3_mag}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/stairs-1-marc/data_plot_acc_annotated.eps}
+  \caption[R3: accelerometer]{Run 3: Indoor stairs Marc, accelerometer}
+  \label{fig:data_gathering_run_3_acc}
+\end{figure}
 

### 2013-12-11 Chapters/Chapter6/data_gathering.tex
WIP: Chapter 6

@@
 \subsection{Indoor stairs}\label{subsec:indoor_stairs}
-The last series of activities recorded consist of walking up and down the stairs, indoor.
+The final series of activities recorded consist of walking up and down the stairs, indoor.
 The stairs are shaped in an semi-circular form, with a small vertical segment on each floor.
@@
 In \Cref{fig:data_gathering_still_3_ascending} the small flat segment is visible.
-Between the stairs the subject has walked in a hallway with a $180^{\circ}$ counter-clockwise turn, as shown in \Cref{fig:data_gathering_still_3_walk}.
+Between the stairs the subject walked in a hallway with a $180^{\circ}$ counter-clockwise turn, as shown in \Cref{fig:data_gathering_still_3_walk}.
+The recorded sensor data of the performed activities are visible in the plots of \Cref{fig:data_gathering_run_3_rot,fig:data_gathering_run_3_mag,fig:data_gathering_run_3_acc}.
 

### 2013-12-11 Chapters/Chapter5/results.tex
Add todo
Add results of artificial data sets'

@@
 \begin{equation}\label{eq:false_alarm_rate}
-  \operatorname*{far}(Y) = \frac{Y_{\operatorname*{false}}}{Y}
+  \operatorname*{far}(Y) = \frac{Y_{\operatorname*{false}}}{Y}.
 \end{equation}
+
+For a high quality method the $\operatorname*{average\_benefit}$ should be high and the $\operatorname*{far}$ should be low.
+
+\TODO{Create plots of FAR and average benefit, as a kind of ROC curve.}
+
+\begin{table}
+  \centering
+  \begin{tabulary}{\textwidth}{|l|c|c|c|c|}
+    \cline{2-5}
+    \multicolumn{1}{l|}{} & Set 1 & Set 2 & Set 3 & Set 4 \\
+    \hline
+    High & 1.6 & 1.6 & 1.5 & 2.2 \\
+    \hline
+    Low & 0.1 & 0.1 & 0.5 & 0.1 \\
+    \hline
+    Closeness & 10 & 10 & 10 & 50 \\
+    \hline
+    Number of change points detected & 10 & 11 & 11 & 11 \\
+    \hline
+    Sum of offsets & 25 & 68 & 169 & 7 \\
+    \hline
+    Window & 50 & 100 & 50 & 50 \\
+    \hline
+    Sigma & 13 & 13 & 15 & 13 \\
+    \hline
+  \end{tabulary}
+  \caption[Results artificial data sets]{Results of the artificial data sets.}
+  \label{tab:results_artificial}
+\end{table}
+
+\TODO{Instead of offset, use mean and spread measure, in \Cref{tab:results_artificial}.}

### 2013-12-12 Chapters/Chapter5/artificial_data_generation.tex
Add final plots for artificial data

--- 2013-11-18 Chapters/Chapter5/artificial_data_generation.tex
@@
 % !TEX root = ../../main.tex
-\section{Artifical data generation}\label{sec:artificial_data_generation}
-In this section we present four data sets as used by \cite{camci2010change,takeuchi2006unifying} to provide for a objetive performance comparison.
+\section{Artificial data generation}\label{sec:artificial_data_generation}
+In this section we present four data sets as used by \cite{camci2010change,takeuchi2006unifying} to provide for a objective performance comparison.
 All the data sets are modeled by a second-order \gls{ar}-model as in \Cref{eq:artificial_data_sets_model}, where $\epsilon_t$ is a Gaussian distribution modeling the noise with mean $0$ and variance $\sigma^2 = 1$, $a_1 = 0.6$ and $a_2 = -0.5$.
@@
   This set origins as the first data set of Camci~\cite{camci2010change}.
-  See \Cref{fig:camci_fixed_increasing_mean,fig:camci_fixed_increasing_mean_thresholds}.
+  The data is plotted in \Cref{fig:camci_fixed_increasing_mean}.
 
@@
   This set origins as the first data set of Takeuchi \etal \cite{takeuchi2006unifying}.
-  See \Cref{fig:takeuchi_reduced_increasing_mean,fig:takeuchi_reduced_increasing_mean_thresholds}.
+  The data is plotted in \Cref{fig:takeuchi_reduced_increasing_mean}.
 
@@
   This set origins as the second set of Camci~\cite{camci2010change}.
-  See \Cref{fig:camci_reduced_increasing_mean_increasing_variance,fig:camci_reduced_increasing_mean_increasing_variance_thresholds}
+  The data is plotted in \Cref{fig:camci_reduced_increasing_mean_increasing_variance}.
 
@@
   This set origins as the thirds sets from Camci~\cite{camci2010change} and Takeuchi \etal \cite{takeuchi2006unifying}.
-  See \Cref{fig:camci_takeuchi_alternating_variance,fig:camci_takeuchi_alternating_variance_thresholds}.
+  The data is plotted in \Cref{fig:camci_takeuchi_alternating_variance}.
 \end{enumerate}
-
-% =====
 
@@
 \centering
-  \includegraphics[width=1\textwidth]{./Figures/notes/camci_fixed_increasing_mean.eps}
+  \includegraphics[width=1\textwidth]{./Figures/chapter5/set_1_data.eps}
   \caption[Fixed increasing mean]{Fixed increasing mean}
@@
 \centering
-  \includegraphics[width=1\textwidth]{./Figures/notes/camci_fixed_increasing_mean_thresholds.eps}
-  \caption[Fixed increasing mean, thresholds]{Fixed increasing mean, thresholds}
-  \label{fig:camci_fixed_increasing_mean_thresholds}
-\end{figure}
-
-% =====
-
-\begin{figure}
-\centering
-  \includegraphics[width=1\textwidth]{./Figures/notes/takeuchi_reduced_increasing_mean.eps}
+  \includegraphics[width=1\textwidth]{./Figures/chapter5/set_2_data.eps}
   \caption[Reduced increasing mean]{Reduced increasing mean}
@@
 \centering
-  \includegraphics[width=1\textwidth]{./Figures/notes/takeuchi_reduced_increasing_mean_thresholds.eps}
-  \caption[Reduced increasing mean, thresholds]{Reduced increasing mean, thresholds}
-  \label{fig:takeuchi_reduced_increasing_mean_thresholds}
-\end{figure}
-
-% =====
-
-\begin{figure}
-\centering
-  \includegraphics[width=1\textwidth]{./Figures/notes/camci_reduced_increasing_mean_increasing_variance.eps}
+  \includegraphics[width=1\textwidth]{./Figures/chapter5/set_3_data.eps}
   \caption[Reduced increasing mean, increasing variance]{Reduced increasing mean, increasing variance.}
@@
 \centering
-  \includegraphics[width=1\textwidth]{./Figures/notes/camci_reduced_increasing_mean_increasing_variance_thresholds.eps}
-  \caption[Reduced increasing mean, increasing variance, thresholds]{Reduced increasing mean, increasing variance, thresholds (log scale)}
-  \label{fig:camci_reduced_increasing_mean_increasing_variance_thresholds}
-\end{figure}
-
-% =====
-
-\begin{figure}
-\centering
-  \includegraphics[width=1\textwidth]{./Figures/notes/camci_takeuchi_alternating_variance.eps}
+  \includegraphics[width=1\textwidth]{./Figures/chapter5/set_4_data.eps}
   \caption[Alternating variance]{Alternating variance}
@@
 \end{figure}
-
-\begin{figure}
-\centering
-  \includegraphics[width=1\textwidth]{./Figures/notes/camci_takeuchi_alternating_variance_thresholds.eps}
-  \caption[Alternating variance, thresholds]{Alternating variance, thresholds}
-  \label{fig:camci_takeuchi_alternating_variance_thresholds}
-\end{figure}

### 2013-12-12 Chapters/Chapter5/results.tex
Add final plots for artificial data

@@
 
-\TODO{Create plots of FAR and average benefit, as a kind of ROC curve.}
+\TODO{Above metrics are not used, remove it and replace with own metric (offset, accuracy)}
+
+The results of all the data sets are provided in \Cref{tab:results_artificial}, together with the used parameters.
+For each data set it shows the high and low thresholds used in the \gls{rt} based change detection method.
+During the post-processing all detected change points within a certain \emph{closeness} range are merged, for which the value differs per data set.
+The difference between the actual and final detected change points are expressed as an offset.
+The data sets were processed with different window lengths and $\sigma$ value for the \gls{rbf} kernel.
+
+To give a better impression of the spread of the accuracy of the detected change points, a box plot for each data set is displayed in \Cref{fig:boxplot_artificial_sets}.
+With each data set embodying different characteristics, different observations can be drawn from the data sets.
+In the references figures, the blue lines represent the radius of the constructed hypersphere.
+The vertical red lines are the discovered change points.
+For each data set the following is observed:
+\begin{enumerate}
+  \item \textbf{Fixed increasing mean:} In \Cref{fig:camci_fixed_increasing_mean_thresholds} we see that each change point is almost equally fast and accurate detected.
+  This implies that the relative difference between the means of the Gaussian distribution has no effect on the detectability of the algorithm.
+  \item \textbf{Reduced increasing mean:} \Cref{fig:takeuchi_reduced_increasing_mean_thresholds} shows that the absolute difference between the means of the Gaussian distribution is a certrain influence to the detectability.
+  Although all change points are correctly detected (at the beginning of the series are two outliers), the change in radius becomes less significant at every change point.
+  \item \textbf{Reduced increasing mean, increasing variance:} In \Cref{fig:camci_takeuchi_alternating_variance_thresholds} we see the same effect, in regard to the mean, as with the previous data set.
+  Furthermore, the increasing variance makes change detection harder.
+  This is visible around the $9000$\textsuperscript{th} data point and after, where the radius keeps increasing.
+  \item \textbf{Alternating variance:} The ability to detect decreases in variance of the Gaussian distribution is displayed in \Cref{fig:camci_reduced_increasing_mean_increasing_variance_thresholds}.
+  Although some change points are discovered with a delay almost up to the length of the window processing the data, the decreases are all detected.
+  The box plot of this set, number 4 in \Cref{fig:boxplot_artificial_sets}, shows a wider spread for the results.
+\end{enumerate}
 
@@
     \hline
-    High & 1.6 & 1.6 & 1.5 & 2.2 \\
+    High threshold & 1.6 & 1.6 & 1.5 & 2.2 \\
     \hline
-    Low & 0.1 & 0.1 & 0.5 & 0.1 \\
+    Low threshold & 0.1 & 0.1 & 0.5 & 0.1 \\
     \hline
-    Closeness & 10 & 10 & 10 & 50 \\
+    Closeness merging & 10 & 10 & 10 & 50 \\
     \hline
@@
     \hline
-    Sum of offsets & 25 & 68 & 169 & 7 \\
+    Mean offset & 2.27 & 6.18 & 0.64 & 15.36 \\
     \hline
-    Window & 50 & 100 & 50 & 50 \\
+    Standard deviation offset & 1.19 & 7.94 & 0.92 & 17.70 \\
     \hline
-    Sigma & 13 & 13 & 15 & 13 \\
+    Window length & 50 & 100 & 50 & 50 \\
+    \hline
+    Sigma of \gls{rbf} & 13 & 13 & 15 & 13 \\
     \hline
@@
 
-\TODO{Instead of offset, use mean and spread measure, in \Cref{tab:results_artificial}.}
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/chapter5/boxplot_results_artificial_sets.eps}
+  \caption[Box plot results artificial data sets]{Box plot of the results for the artificial data sets, indicating the number of data points difference between the actual and detected change points.}
+  \label{fig:boxplot_artificial_sets}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/chapter5/set_1_results.eps}
+  \caption[Fixed increasing mean, thresholds]{Hypersphere radius sizes for the \emph{Fixed increasing mean} data set with detected change points as red vertical bars.}
+  \label{fig:camci_fixed_increasing_mean_thresholds}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/chapter5/set_2_results.eps}
+  \caption[Reduced increasing mean, thresholds]{Hypersphere radius sizes for the \emph{Reduced increasing mean} data set with detected change points as red vertical bars.}
+  \label{fig:takeuchi_reduced_increasing_mean_thresholds}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/chapter5/set_3_results.eps}
+  \caption[Reduced increasing mean, increasing variance, thresholds]{Hypersphere radius sizes for the \emph{Reduced increasing mean, increasing variance} data set with detected change points as red vertical bars.}
+  \label{fig:camci_reduced_increasing_mean_increasing_variance_thresholds}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/chapter5/set_4_results.eps}
+  \caption[Alternating variance, thresholds]{Hypersphere radius sizes for the \emph{Alternating variance} data set with detected change points as red vertical bars.}
+  \label{fig:camci_takeuchi_alternating_variance_thresholds}
+\end{figure}

### 2013-12-12 Chapters/Chapter5.tex
WIP
WIP: artificial data
Restructure Chapter 5 graphs of data sets
WIP: artificial data set results

@@
 \begin{itemize}
-  \item Compare proposed method with methods of Chapter~\ref{Chapter2}
+  \item Compare proposed method with methods of \Cref{Chapter2}
   \item Provide KL-analysis (as Takeuchi does)
@@
 
-% \input{./Chapters/Chapter5/artificial_data}
-% \input{./Chapters/Chapter5/real_world_data}
+In this chapter we will apply the proposed algorithm to artificial data sets.
+For comparison, we use some of the data sets that are used by Camci~\cite{camci2010change} and Takeuchi \etal \cite{takeuchi2006unifying}.
+These data sets are generated by Gaussian noise distributions.
+In \Cref{sec:artificial_data_generation} the precise generation method and characteristics of the data sets are discussed.
+For our final change detection mechanism we use the \gls{rt}-method, also applied by Camci \cite{camci2010change}.
+This is a simple and fast method and allows us to mainly focus on the \gls{oc-svm} model properties that should reflect changes in the data.
+The results of these two methods are discussed in \Cref{sec:artificial_data_results}.
+
+\input{./Chapters/Chapter5/artificial_data_generation}
+
+\input{./Chapters/Chapter5/results}

### 2013-12-12 Chapters/Chapter5/artificial_data_generation.tex
WIP: artificial data set results

@@
   This set origins as the first data set of Camci~\cite{camci2010change}.
+  The main characteristic is that the difference between the means of the Gaussian term $\epsilon_t$ is fixed and thus the relative distance becomes smaller.
   The data is plotted in \Cref{fig:camci_fixed_increasing_mean}.
@@
   This set origins as the first data set of Takeuchi \etal \cite{takeuchi2006unifying}.
+  In contrast with the previous set, in this set the absolute difference in mean is reduced over time.
   The data is plotted in \Cref{fig:takeuchi_reduced_increasing_mean}.
@@
   This set origins as the second set of Camci~\cite{camci2010change}.
+  This set combines the reduced increasing mean property of the previous set with an increasing variance, causing the change points to become more vague over time.
   The data is plotted in \Cref{fig:camci_reduced_increasing_mean_increasing_variance}.
@@
   This set origins as the thirds sets from Camci~\cite{camci2010change} and Takeuchi \etal \cite{takeuchi2006unifying}.
+  This data set is used to measure the ability of detecting changes in decrease of variance, as Camci~\cite{camci2010change} claims is not trivial.
   The data is plotted in \Cref{fig:camci_takeuchi_alternating_variance}.
@@
   \includegraphics[width=1\textwidth]{./Figures/chapter5/set_1_data.eps}
-  \caption[Fixed increasing mean]{Fixed increasing mean}
+  \caption[Fixed increasing mean]{Data set 1, \emph{Fixed increasing mean}.}
   \label{fig:camci_fixed_increasing_mean}
@@
   \includegraphics[width=1\textwidth]{./Figures/chapter5/set_2_data.eps}
-  \caption[Reduced increasing mean]{Reduced increasing mean}
+  \caption[Reduced increasing mean]{Data set 2, \emph{Reduced increasing mean}.}
   \label{fig:takeuchi_reduced_increasing_mean}
@@
   \includegraphics[width=1\textwidth]{./Figures/chapter5/set_3_data.eps}
-  \caption[Reduced increasing mean, increasing variance]{Reduced increasing mean, increasing variance.}
+  \caption[Reduced increasing mean, increasing variance]{Data set 3, \emph{Reduced increasing mean, increasing variance}.}
   \label{fig:camci_reduced_increasing_mean_increasing_variance}
@@
   \includegraphics[width=1\textwidth]{./Figures/chapter5/set_4_data.eps}
-  \caption[Alternating variance]{Alternating variance}
+  \caption[Alternating variance]{Data set 4, \emph{Alternating variance}.}
   \label{fig:camci_takeuchi_alternating_variance}

### 2013-12-13 Chapters/Chapter3/one_class_classification.tex
Fix typos

@@
 
-Boundary methods are based on Vapnik's principle\footnote{With a limited amount of data available, on should avoid solving a more general problem as an intermediate step to solve the original problem \cite{vapnik1998statistical}.} which imply in this case that estimating the complete data density for a \gls{occ} may be too complex, in case one is only interested in the closed boundary.
+Boundary methods are based on Vapnik's principle\footnote{With a limited amount of data available, one should avoid solving a more general problem as an intermediate step to solve the original problem \cite{vapnik1998statistical}.} which imply in this case that estimating the complete data density for a \gls{occ} may be too complex, in case one is only interested in the closed boundary.
 Examples of methods that focus on the boundary (a direct threshold) of the training data distribution are K-centers, Nearest-neighborhood and \gls{svdd} or a combination of those methods \cite{hempstalk2008one}.

### 2013-12-13 Chapters/Chapter3/one_class_svm.tex
Fix typos

@@
 The kernel trick uses a function $K$ which performs an implicit mapping.
-Image from Wikipedia.org\footnotemark}
+Image from Wikipedia.org}
   \label{fig:kernel_mapping}
 \end{figure}
-\footnotetext{\url{http://en.wikipedia.org/wiki/File:Kernel_Machine.png}}
 
@@
 The center $\mathbf{a}$ is a linear combination of of the support vectors.
-Like the classical \gls{svm} and \gls{svdd} it can be required that all the distances from the data points $x_i$ to the center $\mathbf{a}$ are strict less then $R$ (or equivalent measure, to create a hard margin) or soft margins are allowed by using slack variables $\xi_i$.
+Like the classical \gls{svm} and \gls{svdd} it can be required that all the distances from the data points $x_i$ to the center $\mathbf{a}$ are strict less than $R$ (or equivalent measure, to create a hard margin) or soft margins are allowed by using slack variables $\xi_i$.
 In the case of a soft margin, the penalty is determined by $C$ and the minimization is expressed as \Cref{eq:svdd_objective}.

### 2013-12-13 Appendices/AppendixF-final.tex
Add notes final planning

% !TEX root = ../main.tex
% Appendix F

\chapter{Afronden} % Main appendix title

\label{AppendixF} % For referencing this appendix elsewhere, use~\ref{AppendixA}

\lhead{Appendix F. \emph{Afronden}}

\begin{itemize}
  \item Typo's checken en fixen
  \item Chapter 2: Literature review
  \begin{itemize}
    \item Prima hoeveelheid, niet meer veel bij doen
    \item TODO's: alleen aan elkaar lijmen, geen nieuwe info of details
    \item Onderwerpen schrappen: "statistical model/background": mag korter en kleiner
    \item Indeling finalizen. Is niet zo belangrijk.
  \end{itemize}
  \item Chapter 3: inhoud is prima
  \begin{itemize}
    \item fig 3.4: laten staan. Relateren aan andere figuur, zelfde symbolen gebruiken
    \item TODO: alleen aan elkaar lijmen, geen nieuwe inhoud. Eventueel kort refereren naar bron voor meer duiding (parameters C etc)
    \item §3.5 en §3.6: overslaan, komt in §5 en §6.
  \end{itemize}
  \item Chapter 4:
  \begin{itemize}
    \item In begin aangeven wat mijn contributie is: * continue data met changes erin (gelabeled), en * toepassen van oc-svm op HAR data.
    \item Aangeven wat contributie in "proposed method" is: aangeven wat het zelfde en verschilt van Camci/Takeuchi. O.a.: toepassing op real-world data.
    \item Aangeven geen aparte "membership" check zoals Camci doet --> argumenteer robustheid (outliers e.d.).
    \item Naam van hoofdstuk: "Application to real-world HAR data" o.i.d. (past beter met pre-post processing sections). Meer focus op hele systeem, niet de algoritmes.
    \item Naam bedenken voor toepassing, om makkeljk te kunnen refereren.
  \end{itemize}
  \item Chapter 6:
  \begin{itemize}
    \item Foto's van recordings (veel) kleiner en minder. Alleen om setting te duiden.
    \item Een enkele plot van hele run. Verder kleine stukjes uit plots om interessante stukken te duiden (overgang lopen/rennen).
    \item Stukjes van plots gebruiken om verschillen te duiden.
    \item Echte resultaten in (box)plots en tabellen zetten. "Benefit" en FAR.
    \item Performance: vooral aangeven dat het nieuwe (subjectieve) data is. Voor echte vergelijkingen wijzen naar §5.
    \item Bijzondere mis-detecties (of extra detecties) aan het einde duiden ("remarks").
    \item Overall: het moet wat "cleaner"
  \end{itemize}
  \item Chapter 7: prima, aanvullen op de open plekken
\end{itemize}

Tussendoor:
\begin{itemize}
  \item Verantwoording SVM: generieke methode die op veel types data is gebruikt. "Hypothese" van onderzoek is: OC-SVM werkt op HAR data. Terug laten komen in §1 introduction en §7 conclusie.
\end{itemize}

\subsection{Prioriteiten}
\begin{enumerate}
  \item Hoofstuk 1, introduction
  \item Abstract
  \begin{itemize}
    \item Vraagstelling
    \item Antwoord
    \item Aanleiding tot onderzoek
    \item Samenvatting van onderzoek
    \item Wat gedaan is: Praktische toepassen en uitwerking OC-SVM
    \item Wat bijzonder is: applicatie in domein
  \end{itemize}
  \item Lijstje hierboven afwerken
  \item TODO's eruit filteren
  \item Resultaten chapter 6
  \item Conclusie afmaken
\end{enumerate}

\subsection{Planning}
\begin{itemize}
  \item Deadline: 10 januari, geprint op bureau Anne. Final versie geen todo's meer.
\end{itemize}

### 2013-12-13 Chapters/Chapter1.tex
Restructure Chapter 5 graphs of data sets
WIP: Introduction

@@
   \item Current methods, wrapper vs. filter methods
-  \item Problem statement with current filter methods (which follows from Chapter~\ref{Chapter3} which goes in-depth with methods).
+  \item Problem statement with current filter methods (which follows from \Cref{Chapter3} which goes in-depth with methods).
   \item Purpose of this research. E.g. "Find a better algorithm for short-activity segmentation"
@@
 \end{figure}
+
+\begin{itemize}
+  \item Begin with hook, statement, to draw attention
+  \item Cite previous work and explain why new research was necessary.
+  \item Statement of the goal of the thesis: why?
+  \item Context to understand background of goal and the significance
+  \item Acknowledgement of previous research and context
+  \item Explain scope: what will be, and wont be discussed
+  \item Verbal table of contents
+  \item
+\end{itemize}
+
+Introduction to the introduction \\
+Context \\
+Restatement of problem and significance (echo of opening, but now with the context) \\
+Restatement of the response \\
+Roadmap \\
+
+\section{Intro to intro}
+With the wide availability of smartphones and built-in inertial sensors, more and more applications of \gls{har} are introduced.
+Many approaches to recognizing the performed activity rely on classification of recorded data, which can be done online or in batches after the performance.
+Often an time-windowed based approach is used, which feeds short consecutive segments of data through a model construction phase.
+The constructed model is used to determine what (earlier learned) activity is performed.
+Besides the explicit classification of the data, an implicit result is a segmentation of performed activities over time.
+
+In this thesis the goal is to find the temporal segmentation of time series data explicitly, prior to the classification of the activities.
+This is done under the assumption that it could be beneficial for classification methods to possess this explicit segmentation, since the model construction phase can use more data than the time-windowed approach would allow.
+To find the temporal segmentation, we rely on change detection in time series data.
+In the context of this research, this time series data consist of recordings from inertial sensors found in smartphones, during the performance of human activities.
+We have recorded both in- and outdoor activities in an uncontrolled environment and in a continuous manner.
+
+For the change detection algorithm we have used a special form of \gls{svm}, the \gls{oc-svm} based \gls{svdd} as introduced by Tax~\cite{tax2001one}.
+This method models the data under consideration in the shape of a high-dimensional hypersphere.
+From this hypersphere the radius is extracted, since a change in data characteristics (resulting from a change in performed activity) should result in a change of radius.
+
+\section{Thesis structure}
+The structure of this thesis is as follows.
+In \Cref{Chapter2} a literature review is provided.
+We will look at the different interpretations and implementations for change, novelty, and outlier detection.
+Previous work in the field of \gls{har} is discussed and we end with existing applications of \glspl{svm} to change detection.
+The next chapter further analyses \glspl{svm} and two implementations of \glspl{oc-svm}.
+It relates the properties of the constructed \gls{oc-svm} models to change detection in time series data.
+That relation is applied to our problem formulation in \Cref{Chapter4}, where we construct our change detection method, based on the work of Camci~\cite{camci2010change}.
+The two following chapters apply the method to both artificial and real-world data.
+\Cref{Chapter5} show the objective performance of the method to artificial constructed data.
+In \Cref{Chapter6} we apply the method to our self-recorded real-world data.
+We show the ability of detecting change in \gls{har} data, recorded by inertial sensors.
+This thesis is conclused in \Cref{Chapter7}, in which we reflect on the performed research and state possibilities for feature research.

### 2013-12-16 Chapters/Chapter1.tex
WIP: Introduction chapter

--- 2013-11-06 Chapters/Chapter1.tex
@@
 % Chapter 1
+
+-- Overall todo's --
+\begin{itemize}
+  \item \TODO{The term \emph{emperical data/distruction} is nowhere used in the thesis. or: Emperical cumulative distribution. Especially Chapter 2.}
+  \item \TODO{Check for ``tang-constructions'': saying to much in a single sentence.}
+  \item \TODO{Overall: meer focus op de \emph{waarom} vraag/antwoord bij keuzes. Bouwt aan motivatie en context van het geheel.}
+\end{itemize}
 
@@
 
-%----------------------------------------------------------------------------------------
+With the wide availability of smartphones and built-in inertial sensors, more and more applications of \gls{har} are introduced \cite{avci2010activity,derawi2010accelerometer,guenterberg2009distributed}.
+Many approaches to recognize the performed activity rely on classification of recorded data \cite{devaul2001real,ward2006activity,yang2008using,anguita2012human,bao2004activity,bernecker2012activity,he2008activity}, which can be done online or in batches after the performance \cite{duque2012offline}.
+Often a time-windowed based approach is used, which feeds short consecutive segments of data through a model construction and recognition phase.
+The constructed model is used to determine which (earlier learned) activity is performed.
+Besides the explicit classification of the data, an implicit obtained result is a segmentation of performed activities over time.
 
-\section{Outline}
-\emph{Not intented for the reader.}
-\begin{itemize}
-  \item Context of research (\gls{har}), real-world applications
-  \item Current methods, wrapper vs. filter methods
-  \item Problem statement with current filter methods (which follows from Chapter~\ref{Chapter3} which goes in-depth with methods).
-  \item Purpose of this research. E.g. "Find a better algorithm for short-activity segmentation"
-  \item Provide the \emph{motivation} for this approach
-  \item Relate to real-world applications
-  \item Outline for rest of the thesis
-\end{itemize}
+In this thesis the goal is to find the temporal segmentation of time series data explicitly, prior to the classification of the activities.
+This is done under the assumption that it could be beneficial for classification methods to possess this explicit segmentation, since the model construction phase can use more data than the time-windowed approach would allow.
+To find a temporal segmentation, we rely on change detection in time series data.
+In the context of this research, this time series data consists of recordings from inertial sensors found in smartphones, such as the accelerometer signal, during the performance of human activities.
+We are interested in both in- and outdoor activities in an uncontrolled environment and in a continuous manner, which can be regarded as real-world data.
+Although the number of performed activities will be limited, we assume the temporal segmentation must be more robust then in the case of data from a controlled environment.
 
--- Overall todo's --
-\begin{itemize}
-  \item \TODO{The term \emph{emperical data/distruction} is nowhere used in the thesis. or: Emperical cumulative distribution. Especially Chapter 2.}
-  \item \TODO{Check for ``tang-constructions'': saying to much in a single sentence.}
-  \item \TODO{Overall: meer focus op de \emph{waarom} vraag/antwoord bij keuzes. Bouwt aan motivatie en context van het geheel.}
-  \item \TODO{Nu staat nog veel (met name Chapter 2) in ``aantekening'' style; ombouwen naar uitleggen, lezer begeleiden, duiden, verlaren etc.}
-  \item \TODO{Test moet condenser: bepalen wat er uit gelaten mag worden.}
-\end{itemize}
+For the change detection algorithm we have used a special form of \gls{svm}, the \gls{occ} based \gls{svdd} as introduced by Tax~\cite{tax2001one}.
+This method models the data under consideration in the shape of a high-dimensional hypersphere.
+From this hypersphere the radius is extracted, which can be used as an indication of change and be processed by one-dimensional change detection algorithms.
+Since a change in data characteristics, resulting from a change in performed activity, alters the data distribution, the model and thus the radius should reflect this change as well.
+
+\section{Scope}
+The scope of this thesis is to find a temporal segmentation of time series data from inertial sensors embedded in smartphones, obtained during the performance of human activities.
+As mentioned above, the segmentation is often obtained as a by-product from direct windowed classification.
+Other approaches have a very rigid form of segmentation, \eg by limiting the number of possible segment types \cite{himberg2001time,chamroukhi2013joint}, or have applied it to different type of time series data \cite{li2007segmentation,fuchs2010online,guenterberg2009automatic}.
+Others have extracted features from windows of data which are used for the segmentation \cite{guo2012adaptive}.
+These activities include sitting, standing, walking, running, and ascending and descending staircases.
+The problem of finding a temporal segmentation can be seen as a sub-problem to the problem of Activity Classification.
+This relation is illustrated in \Cref{fig:thesis_goal}.
+The figure shows that we can apply classification over windows of the time series data.
+In many approaches this is constructed with a (partially overlapping) sliding window.
+In the context of \gls{har}, the window length commonly consists of around $2$ seconds of data points.
+In our setup, we create a temporal segmentation instead of a fixed window length.
+Each segment should consist of homogeneous data, in the sense that the underlying data generating activity should not change within a segment.
+Likewise, the previous and next segments (if present) should be generated by a different performed activity.
+
+To find these segments of homogeneous data, we employ a model construction phase.
+The \gls{oc-svm} model transforms the data points to a high-dimensional hyperspherical shape from which we extract the radius as characterizing metric.
+In this phase the data is processed by a fixed window length and a change in data characteristics should reflect in a (significant) change of the hypersphere's radius.
+Such a sudden change in the radius thus indicates a change in the time series data and indicates the end of the current and beginning of a new segment.
+In \Cref{sec:literature_review_temporal_segmentation} a more detailed overview of temporal segmentation methods is discussed.
+The information about segments can be used in the classification phase of \gls{har}, which applies classification to windows (or in this case: segments) of data.
+The final result, which is outside the scope is this thesis, will be a full classification of activities over the time series data.
 
@@
 The scope of this thesis is Temporal Segmentation, which can be useful in the context of Temporal Classification.
-Given a data set, we apply the construction of \gls{oc-svm} models as a form of dimensionality reduction.
-Using the reduced representation, being the radius $R$ of the constructed hypersphere, we apply direct change detection methods.
+Given a data set, we construct a \gls{oc-svm} high-dimensional spherical model, from which we extract the radius.
+This metric is then applied to direct change detection algorithms.
 The detected change points can support the classification of homogeneous segments of data.}
@@
 \end{figure}
+
+\section{One-Class Support Vector Machines}
+In the setup, as described in the previous section, we employ a model construction phase.
+For this model we will use an \acrlong{oc-svm} implementation: the \acrlong{svdd} algorithm by Tax~\cite{tax2001one}.
+In earlier research multi-class \glspl{svm} are used for direct classification of inertial sensor data \cite{he2008activity,mountrakis2011support,anguita2012human}.
+Others have used \gls{oc-svm} classifiers for novelty, or outlier, detection \cite{scholkopf1999support,camci2010change,li2003improving,ma2003time,tax1999support}.
+However, to our knowledge, \gls{oc-svm} methods have not yet been applied to inertial sensor time series data directly.
+This algorithm transforms the time series data, which can be of any high dimension, to an higher dimensional feature space.
+When the data is assumed to be of the same class, it can be characterized by an enclosing boundary, in the shape of a hypersphere.
+Using this \gls{occ} method, we are able to test new data points of being a member of this class.
+When a new data point, after transformation to the feature space, lies within the boundary, it is a member of the class, and otherwise it is regarded as an outlier.
+
+In case the data distribution changes, the data points will have a larger variance which results in a larger hypersphere in the feature space.
+This means that with continuous model updating, a change in the underlying process of a time series data means the hypersphere will have a larger radius.
+Thus the \gls{oc-svm} method allows us to generate a one-dimensional time series data, reduced from a complex high-dimensional original signal.
+The final radius time series data can be inspected by simpler change detection methods, in a number of ways.
+\Cref{Chapter3} will discuss the \gls{oc-svm} methods in detail and \Cref{Chapter4} elaborates on the overall change detection method.
+
+\section{Data sets}
+In recent previous research two data sets for the context of \gls{har} have been made public.
+In order to benchmark our proposed method we tried to employ it to these data sets.
+However, it turned out the data sets were not appropriate for continuous temporal segmentation.
+The first set, the \gls{wisdm} from \cite{kwapisz2011activity}, consists of non-continuous recordings of activities.
+Obviously, this voids the usefulness of the data set of our research.
+The second set, the \gls{uci-har} from \cite{anguita2012human}, also lacks continuous recordings and furthermore exhibits patterns and labeling which seems to be incorrect.
+Besides these problems, since there are no visual images of the performed activities, there is no ground truth which can be consulted in case of ambiguity.
+To overcome these problems, we have created our own real-world data sets.
+
+Our own data sets are recorded by placing a smartphone with inertial sensors in the front right pants pocket of a subject.
+The subject was asked to perform common and simple activities in an uncontrolled environment, both in- and outdoor situated.
+During the performance the subjects were also recorded with a video camera from the third-person point of view.
+This allows us to create a ground truth for the temporal segmentation of performed activities.
+
+Besides the recorded data sets we have used artificial data sets to obtain an objective performance measure of the method.
+These artificial data sets are modeled to the sets used in \cite{camci2010change,takeuchi2006unifying}.
+In \Cref{Chapter5} we discuss the artificial data sets and results in detail.
+The real-world recordings and results are discussed in \Cref{Chapter6}.
+
+\section{Contributions}
+As a result from our proposed method and application to recorded real-world data sets, this thesis will provide the following contributions:
+\begin{enumerate}
+  \item \textbf{Continuous recorded human activity data:} the proposed method is applied to our own recorded human activity data.
+  This data set can be grounded by using the video recordings of the subject while performing the activities.
+  The data set, consisting of raw inertial sensor recordings, manual change point and activity labeling, and video recordings, are used for testing the method and are made publicly available \cite{vlasveld2013continuous}.
+  \item \textbf{Application of \gls{oc-svm} methods to inertial sensor data:} up to our knowledge, this thesis is the first application of \gls{oc-svm} methods, especially the \gls{svdd} method, to inertial sensor signals, such as accelerometer time series data.
+  Furthermore, this thesis contributes to the small amount of research that has a temporal segmentation of inertial sensor data as its primary objective.
+  \item \textbf{Adaptation of \acrshort{svcpd}:} we have adopted and slightly adapted the \acrshort{svcpd} method by Camci~\cite{camci2010change} in order to find change points in time series data.
+  In our simplification we have not encountered noticeably decrease in performance.
+\end{enumerate}
+
+\section{Thesis structure}
+The structure of this thesis is as follows.
+In \Cref{Chapter2} a literature review is provided.
+We will look at the different interpretations and implementations for change, novelty, and outlier detection.
+Previous work in the field of \gls{har} is discussed and we end with existing applications of \glspl{svm} to change detection.
+The next chapter further analyses \glspl{svm} and two implementations of \glspl{oc-svm}.
+It relates the properties of the constructed \gls{oc-svm} models to change detection in time series data.
+That relation is applied to our problem formulation in \Cref{Chapter4}, where we construct our change detection method \TODO{add name}, based on the work of Camci~\cite{camci2010change}.
+The two following chapters apply the method to both artificial and real-world data.
+\Cref{Chapter5} show the objective performance of the method to artificial constructed data.
+In \Cref{Chapter6} we apply the method to our self-recorded real-world data.
+In that chapter we show the ability of detecting change in \gls{har} data, recorded by inertial sensors.
+This thesis is concluded in \Cref{Chapter7}, in which we reflect on the performed research and state possibilities for feature research.

### 2013-12-16 Chapters/Chapter6/data_gathering.tex
Put stills in grid. Crop images

@@
   \label{fig:data_gathering_run_2_acc}
-\end{figure}
-
-%=== STILLS
-\begin{figure}
-\centering
-  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/stills/jos_sprint.png}
-  \caption[Recording still 4]{Subject 1 sprinting}
-  \label{fig:data_gathering_still_1_sprint}
 \end{figure}
@@
   \centering
-  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/stills/jos_cp_walk-run.png}
-  \caption[Recording still 1]{Subject 1 on the transition from walking to running}
-  \label{fig:data_gathering_still_1_still_walk_to_run}
-\end{figure}
-
-\begin{figure}
-\centering
-  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/stills/jos_run_2.png}
-  \caption[Recording still 2]{Subject 1 running}
-  \label{fig:data_gathering_still_1_running}
-\end{figure}
-
-\begin{figure}
-\centering
-  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/stills/jos_run.png}
-  \caption[Recording still 3]{Subject 1 running}
-  \label{fig:data_gathering_still_1_running_2}
-\end{figure}
-
-\begin{figure}
-\centering
-  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/stills/jos_walking.png}
-  \caption[Recording still 5]{Subject 1 walking}
-  \label{fig:data_gathering_still_1_walk}
-\end{figure}
-
-Stills from subject 2:
-
-\begin{figure}
-\centering
-  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/stills/roemer_run_cw.png}
-  \caption[Recording still 6]{Subject 2 running in clockwise corner}
-  \label{fig:data_gathering_still_2_run}
-\end{figure}
-
-\begin{figure}
-\centering
-  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/stills/roemer_walk.png}
-  \caption[Recording still 7]{Subject 2 walking}
-  \label{fig:data_gathering_still_2_walk}
+  \begin{subfigure}[b]{0.2\textwidth}
+    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/stills/jos_walking.png}
+    \caption{Walking}
+    \label{fig:stills_subject_2_walking}
+  \end{subfigure}%
+  \qquad \qquad \qquad
+  \begin{subfigure}[b]{0.2\textwidth}
+    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/stills/jos_cp_walk-run.png}
+    \caption{Change point}
+    \label{fig:stills_subject_2_change_point}
+  \end{subfigure}
+  \qquad \qquad \qquad
+  \begin{subfigure}[b]{0.2\textwidth}
+    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/stills/jos_run_2.png}
+    \caption{Running}
+    \label{fig:stills_subject_2_running}
+  \end{subfigure}
+  \par\bigskip
+  \begin{subfigure}[b]{0.2\textwidth}
+    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/stills/roemer_run_cw.png}
+    \caption{Running CW}
+    \label{fig:stills_subject_1_running_clockwise}
+  \end{subfigure}
+  \qquad \qquad \qquad
+  \begin{subfigure}[b]{0.2\textwidth}
+    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/stills/roemer_walk.png}
+    \caption{Walking}
+    \label{fig:stills_subject_1_walking}
+  \end{subfigure}
+  \qquad \qquad \qquad
+  \begin{subfigure}[b]{0.2\textwidth}
+    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/stills/jos_sprint.png}
+    \caption{Sprintng}
+    \label{fig:stills_subject_2_sprinting}
+  \end{subfigure}
+  \caption[Stills subject 2]{Stills of video recordings during performing of the outdoor activities of walking and running.
+\Cref{fig:stills_subject_2_change_point} shows the change point between the two activities.
+\Cref{fig:stills_subject_1_running_clockwise} shows the running in clockwise motion.}\label{fig:stills_subject_2}
 \end{figure}
@@
 The recorded sensor data of the performed activities are visible in the plots of \Cref{fig:data_gathering_run_3_rot,fig:data_gathering_run_3_mag,fig:data_gathering_run_3_acc}.
+
+\begin{figure}
+  \centering
+  \begin{subfigure}[b]{0.3\textwidth}
+    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/stills/marc_downstairs.png}
+    \caption{Downstairs}
+    \label{fig:stills_subject_3_downstairs}
+  \end{subfigure}%
+  ~
+  \begin{subfigure}[b]{0.3\textwidth}
+    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/stills/marc_walk.png}
+    \caption{Walking}
+    \label{fig:stills_subject_3_walking}
+  \end{subfigure}
+  ~
+  \begin{subfigure}[b]{0.3\textwidth}
+    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/stills/marc_stairs_up_walk.png}
+    \caption{Upstairs}
+    \label{fig:stills_subject_3_upstairs}
+  \end{subfigure}
+  \caption[Stills subject 3]{Stills of video recordings during performing of the indoor activities of walking, ascending and descending the stairs.}\label{fig:stills_subject_3}
+\end{figure}
 
@@
 \end{figure}
-
-\begin{figure}
-\centering
-  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/stills/marc_downstairs.png}
-  \caption[Recording still 8]{Subject 3 descencing the stairs.}
-  \label{fig:data_gathering_still_3_descending}
-\end{figure}
-
-\begin{figure}
-\centering
-  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/stills/marc_stairs_up_walk.png}
-  \caption[Recording still 9]{Subject 3 will start walking for a small fragment and will continue ascending the stairs.}
-  \label{fig:data_gathering_still_3_ascending}
-\end{figure}
-
-\begin{figure}
-\centering
-  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/stills/marc_walk.png}
-  \caption[Recording still 10]{Subject 3 walking in the hallway. He will turn $180^{\circ}$ counter-clockwise.}
-  \label{fig:data_gathering_still_3_walk}
-\end{figure}

### 2013-12-17 Chapters/Chapter1.tex
WIP: Finalizing chapter 3

@@
 Others have used \gls{oc-svm} classifiers for novelty, or outlier, detection \cite{scholkopf1999support,camci2010change,li2003improving,ma2003time,tax1999support}.
-However, to our knowledge, \gls{oc-svm} methods have not yet been applied to inertial sensor time series data directly.
+The research by Yin \etal  \cite{yin2008sensor} creates a \gls{oc-svm} classifier to filter normal activity traces, using training data.
+However, to our knowledge, \gls{oc-svm} methods have not yet been applied to inertial sensor time series data directly to find changes in behavior and create a temporal segmentation.
 This algorithm transforms the time series data, which can be of any high dimension, to an higher dimensional feature space.
@@
   The data set, consisting of raw inertial sensor recordings, manual change point and activity labeling, and video recordings, are used for testing the method and are made publicly available \cite{vlasveld2013continuous}.
-  \item \textbf{Application of \gls{oc-svm} methods to inertial sensor data:} up to our knowledge, this thesis is the first application of \gls{oc-svm} methods, especially the \gls{svdd} method, to inertial sensor signals, such as accelerometer time series data.
+  \item \textbf{Application of \gls{oc-svm} methods to inertial sensor data:} to our best knowledge, this thesis is the first application of \gls{oc-svm} methods, especially the \gls{svdd} method, to inertial sensor signals in the context of creating a temporal segmentation.
   Furthermore, this thesis contributes to the small amount of research that has a temporal segmentation of inertial sensor data as its primary objective.

### 2013-12-17 Chapters/Chapter2/svm.tex
Cleanup of sections
WIP: Finalizing chapter 3

@@
 % !TEX root = ../../main.tex
-\section{Support Vector Machines in Change Detection}\label{sec:svm_in_change_detection}
+\section{Support Vector Machines in Change Detection}\label{sec:literature_review_svm}
 
@@
 
-The method by Yin and Yang \cite{yin2008sensor} uses the \gls{svdd} method in the first phase of a two-phase algorithm to filter commonly available normal activities.
+The method by Yin \etal \cite{yin2008sensor} uses the \gls{svdd} method in the first phase of a two-phase algorithm to filter commonly available normal activities.
 The \gls{svcpd} method of Camci \cite{camci2010change} uses \gls{svdd} applies it to time-series data.

### 2013-12-17 Chapters/Chapter3.tex
Cleanup of sections
WIP: Finalizing chapter 3

--- 2013-11-18 Chapters/Chapter3.tex
@@
 
-%----------------------------------------------------------------------------------------
-\section{Outline}
-\emph{Not intented for the reader.}
-\begin{itemize}
-  \item Content follows blogpost, but more formal
-  \item Explain the connection between change detection and \{outlier detection, density estimation, \etc\}
-  \item Explain change detection in relation with ``M-Estimators''.
-  \item Explain why and how SVM can be used for outlier detection, and thus for change detection in time series
-  \item Explain options when using SVM, such as RBF kernel
-  \item Relate characteristics of (accelerometer) data to the used (RBF?) kernel
-  \item Define some quality metrics --> this should be in a different chapter (About experiments)
-  \item Note: try to avoid the perspective of \emph{density estimation}. Instead: \emph{data description} or \emph{boundary description}.
-\end{itemize}
-
-This chapter discusses the concepts and algorithms that will be used as a basis for the proposed method, as discussed in \Cref{Chapter4}.
-The first sections formulates the problem of change detection and relates it to outlier and novelty detection.
-It transforms the traditional problem to change detection for times series data.
-In \Cref{sec:one_class_classification} the problem of \acrlong{occ} is discussed and shows a number of implementations.
-Two \acrlong{svm}-based methods, \gls{nu-svm} and \gls{svdd}, are further disussed in \Cref{sec:one_class_svm}.
-For the application to the problem of \acrlong{har}, the characteristics of the collected sensor data is discussed in~\ref{sec:sensor_data_characteristics}.
-Finally a theoretic set of quality metrics is discussed in \Cref{sec:quality_metrics}.
+This chapter discusses the concepts and algorithms that will be used as a basis for the proposed method, as introduced in \Cref{Chapter4}.
+The first section formulates the problem of change detection and relates it to outlier and novelty detection.
+It transforms the traditional problem of outlier detection into change detection for times series data.
+In \Cref{sec:one_class_classification} the problem of \acrlong{occ} is explained and discusses a number of implementations.
+In the final sections of this chapter, \Cref{sec:one_class_svm}, two \gls{svm}-based \gls{occ} methods, \gls{nu-svm} and \gls{svdd}, and the influence of model parameters are further discussed in detail.
 
@@
 \input{./Chapters/Chapter3/one_class_svm}
-\input{./Chapters/Chapter3/sensor_data_characteristics}
-\input{./Chapters/Chapter3/quality_metrics}

### 2013-12-17 Chapters/Chapter3/change_detection_time_series.tex
WIP: Finalizing chapter 3

@@
 % !TEX root = ../../main.tex
-\section{Change Detection in Time Series}\label{sec:change_detection_time_series}
-\TODO{Write section introduction} \\
-\emph{The section should relate the problem of outlier detection to change detection.
-It should transform the traditional problem formulation to that of time series data.
-Build to one-class classification, which can be used for outlier detection}
+\section{Problem Formulation Change Detection}\label{sec:change_detection_time_series}
+% \TODO{Write section introduction} \\
+% \emph{The section should relate the problem of outlier detection to change detection.
+% It should transform the traditional problem formulation to that of time series data.
+% Build to one-class classification, which can be used for outlier detection}
 
-\subsection{Relation to outlier detection}
-The problems of outlier and novelty detection, segmentation, and change detection are closely related.
+% \subsection{Relation to outlier detection}
+The problems of outlier and novelty detection, segmentation, and change detection in time series data are closely related.
 The terminology depends on the field of application, but there are subtle differences.
@@
 
-\subsection{Problem formulation}\label{subsec:change_detection_problem_formulation}
+% \subsection{Problem formulation}\label{subsec:change_detection_problem_formulation}
 The problem of change point detection can be formulated using different type of models, as discussed in~\ref{subsec:segmentation}.
@@
 Since we are only interested in different activities (which are represented by sudden changes), slight changes within an activity are not of interest.
-\Cref{sec:sensor_data_characteristics} discusses the representation of activities and changes between them in the data in more detail.
 
-\TODO{add more formal definition or analysis of change detection}
+% -- Literature --
 
--- Literature --
-
-``Online segmentation of time series based on polynomial least-squares approximations'', \cite{fuchs2010online}. 25, 2010 \\
+% ``Online segmentation of time series based on polynomial least-squares approximations'', \cite{fuchs2010online}. 25, 2010 \\
 

### 2013-12-17 Chapters/Chapter3/one_class_classification.tex
WIP: Finalizing chapter 3

@@
 \section{One-Class Classification}\label{sec:one_class_classification}
-\TODO{Add ref to \cite{hodge2004survey}, which states 3 types of outlier detection. It puts \gls{occ} as of Type 3: \textbf{semi-supervised recognition or detection}}
-
 As discussed in the previous section, change detection in time series can be implemented using outlier, or novelty, detection.
@@
 The algorithms build up a model of the data, assuming only normal data objects are available (or a very limited amount of example outlier data objects).
+This is also known as novelty detection or semi-supervised detection and is of Type 3 in the system by Hodge and Austin~\cite{hodge2004survey}.
 This differs from classical classification algorithms, which commonly rely of both positive and negative examples.
@@
 This will result in a similarity-measure expressed in the properties of outliers.
-This measure is related to other similarity measures, such as \gls{pdf} estimation and the \gls{kl} divergence as discussed in \Cref{subsec:change_detection_problem_formulation}.
+This measure is related to other similarity measures, such as \gls{pdf} estimation and the \gls{kl} divergence as discussed in \Cref{subsec:change_detection_time_series}.
 
@@
 In this Section we have discussed the problem of \gls{occ} and different kind of implementations.
-An often used implementation is the \gls{svm}-based method \cite{noumir2012simple}, since is shows good performance in comparitive researches.
-\TODO{Add refs for good results of OC-SVM}
+An often used implementation is the \gls{svm}-based method \cite{noumir2012simple}, since is shows good performance in comparitive researches \cite{khan2010survey,smola1998connection}.
 In the following section (\ref{sec:one_class_svm}) two implementations of \gls{oc-svm} will be discussed, the \gls{svdd} method of Tax and Duin \cite{tax1999support} and the \gls{nu-svm}-algoritm by Sch\"olkopf.

### 2013-12-17 Chapters/Chapter3/one_class_svm.tex
WIP: Finalizing chapter 3

@@
 \section{One-Class Support Vector Machine}\label{sec:one_class_svm}
+In this section we will discuss the details of a \gls{svm} and \gls{oc-svm} implementations.
+The classical implementation of a \gls{svm} is to classify a dataset in two distinct classes.
+This is a common usecase, although sometimes there is no training data for both classes available.
+Still, one would like to classify new data points as regular, in-class, or out-of-class, \eg in the case of a novelty detection.
+With that problem only data examples from one class is available and the objective is to recognize new data points that are not part of that class.
+This unsupervised learning problem is closely related to density estimation.
+In that context, the problem can be the following.
+Given an underlying probabilty distribution $P$ the goal is to find a subset $S$ for which the probability that a data point from $P$ lies outside $S$ equals some predetermined value $\nu$ between $0$ and $1$ \cite{scholkopf1999support}.
+
+In the following of this section we will start with a discussion of traditional \glspl{svm}.
+In \Cref{subsec:kernels} we will show how kernels allow for non-linear decision functions.
+That is followed by two different implementations of \gls{oc-svm}: \gls{nu-svm} in \Cref{subsec:nu-svm} by Sch\"olkopf \etal \cite{scholkopf1999support}, which closely follows the above problem statement regarding density estimation, and \gls{svdd} by Tax and Duin \cite{tax1999support} in \Cref{subsec:oc-svm-svdd}.
+The final part of this section discusses the influende of model parameters on the performance of \glspl{svm}
 
@@
 When the hyperplane is projected to the (lower) original input space $\mathcal{I}$ it creates a non-linear separating curve.
-This is illustrated in \Cref{fig:svm_mapping_spaces}.
+A schematic overview of this process is illustrated in \Cref{fig:svm_mapping_spaces}.
+The original data points $x \in I$ are through $\phi(x)$ projected to the points $z \in F$.
+In that space the linear hyperplane is constructed by weights $w$.
+This results in a non-linear hyperplane $y$ in input space $I$.
 The mapping and projection of data points can be efficient (and implicit) performed by using the kernel trick, which is discussed in \Cref{subsec:kernels}.
@@
   \label{fig:kernel_mapping}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=0.8\textwidth]{./Figures/chapter3/svm_mapping_spaces.pdf}
+  \caption[Mapping spaces in \gls{svm}]{The input vector $\vectorsym{x}$ from input space $\mathcal{I}$ is mapped by a non-linear function $\phi(\vectorsym{x})$ to a feature vector $\vectorsym{z}$ in the high dimensional feature space $\mathcal{F}$.
+The weights of $\vectorsym{w}$ create a linear separating hypeplace, which maps the high dimensional vector to the predicted outcome $\vectorsym{y}$.}
+  \label{fig:svm_mapping_spaces}
 \end{figure}
@@
 \centering
-  \includegraphics[width=0.5\textwidth]{./Figures/chapter3/svm_mapping_spaces.pdf}
-  \caption[Mapping spaces in \gls{svm}]{The input vector $\vectorsym{x}$ from input space $\mathcal{I}$ is mapped by a non-linear function $\phi(\vectorsym{x})$ to a feature vector $\vectorsym{z}$ in the high dimensional feature space $\mathcal{F}$.
-The weights of $\vectorsym{w}$ create a linear separating hypeplace, which maps the high dimensional vector to the predicted outcome $\vectorsym{y}$.
-\TODO{Is this figure obsolete?
-\Cref{fig:kernel_mapping} shows almost the same}}
-  \label{fig:svm_mapping_spaces}
-\end{figure}
-
-\begin{figure}
-\centering
   \includegraphics[width=0.5\textwidth]{./Figures/chapter3/svm_separating_plane_with_margin.png}
@@
 Here $w$ is the normal vector for the separating hyperplane and the distance between the two margins is $\frac{2}{\lVert{w}\rVert}$.
-Image from Wikipedia.org\protect\footnote}
+Image from Wikipedia.org}
   \label{fig:svm_hyperplane}
 \end{figure}
-\footnotetext{\url{http://en.wikipedia.org/wiki/File:Svm_max_sep_hyperplane_with_margin.png}}
 
@@
   \begin{aligned}
-  y_i( w^T \phi(x_i) + b) \geq & 1 - \xi_i & \mbox{ for all } i = 1, \dots, n \\*
+  y_i( w^T \phi(x_i) + b) \geq \: & 1 - \xi_i & \mbox{ for all } i = 1, \dots, n \\*
    & \xi_i \geq 0 & \mbox{ for all } i = 1, \dots, n\\*
@@
 \end{equation}
-
-\TODO{better format of above formulas. Check with notation and formulas of chapter 9 of \cite{cherkassky2007learning}.}
 
@@
 where $K(x, x_i) = \phi(x)^T\phi(x_i)$ is the dot product of data objects in feature space $\mathcal{F}$ (which is further discussed in \Cref{subsec:kernels}).
-Here every data point $\mathcal{I}$ for which $a_i > 0$ is weighted in the decision function and thus ``supports'' the classification machine: hence the name ``\acrlong{svm}''.
+Here every data point in $\mathcal{I}$ for which $a_i > 0$ is weighted in the decision function and thus ``supports'' the classification machine: hence the name ``\acrlong{svm}''.
 Since it is shown that under certain circumstances \glspl{svm} show an equality to sparse representations \cite{girosi1998equivalence,smola1998connection}, there will often be relatively few Lagrange multipliers with a non-zero value.
@@
 \end{enumerate}
-The second property is what enables powerful adaption for support vector machines to learn non-linear decision boundaries.
+An effect of the first property, combined with the equality to sparse represenations, is that \glspl{svm} often have good results, even in the case of high dimensional data (also known as the curse of dimensionality) or limited training examples \cite{cherkassky2007learning}.
+The second property is what enables powerful adaption for \glspl{svm} to learn non-linear decision boundaries.
 The workings of the kernel matrix $K$ and the non-linear boundaries is discussed in the following section.
-
-\TODO{Add more on sparsity and good results, because of sparse representations, even in case of data of high dimensionality. e.g. \cite{cherkassky2007learning}.}
 
@@
 
--- Quotes and refs to use --x
+% -- Quotes and refs to use --x
 
-sparsity, \gls{rkhs}: \cite{girosi1998equivalence} \\
+% sparsity, \gls{rkhs}: \cite{girosi1998equivalence} \\
 
-Use Chapter 2 from ``Learning with kernels'', \cite{scholkopf2002learning}.
-
-%--------------------------------------------
-
-\subsection{One-Class Support Vector Machines}\label{subsec:one-class-svm}
-The classical implementation of a \gls{svm} is to classify a dataset in two distinct classes.
-This is a common usecase, although sometimes there is no training data for both classes available.
-Still, one would like to classify new data points as regular, or in-class, or out-of-class, \eg in the case of a novelty detection.
-With that problem only data from one class is available and the objective is to recognize new data points that are not part of that class.
-This unsupervised learning problem is closely related to density estimation.
-In that context, the problem can be the following.
-Given an underlying probabilty distribution $P$ the goal is to find a subset $S$ for which the probability that a data point from $P$ lies outside $S$ equals some predetermined value $\nu$ between $0$ and $1$ \cite{scholkopf1999support}.
-
-We will discuss two implementations of \glspl{oc-svm}.
-The first is the \gls{nu-svm} by Sch\"olkopf \etal \cite{scholkopf1999support} and closely follows the above problem statement regarding density estimation.
-The second, on which we will base our change detection method, is the \gls{svdd} by Tax and Duin \cite{tax1999support}.
-
-\TODO{Add something about robustness, discusses in section 9.2 from \cite{cherkassky2007learning,chen2004m}}
+% Use Chapter 2 from ``Learning with kernels'', \cite{scholkopf2002learning}.
 
@@
 \subsection{\acrlong{nu-svm}}\label{subsec:nu-svm}
-\TODO{Check notations with book of Scholkopf \cite{scholkopf2002learning}, chapter 8, page 227.} \\
 The first of the \gls{oc-svm} methods we will discuss is often referred to as \gls{nu-svm} and introduced by Sch\"olkopf \etal \cite{scholkopf1999support}.
@@
     \begin{aligned}
-      (w \cdot \phi(x_i)) \geq & \rho - \xi_i & \mbox{ for all } i = 1, \dots, n \\
+      (w \cdot \phi(x_i)) \geq \: & \rho - \xi_i & \mbox{ for all } i = 1, \dots, n \\
       & \xi_i \geq 0 & \mbox{ for all } i = 1, \dots, n \\
@@
   \centering
-    \includegraphics[width=0.5\textwidth,keepaspectratio]{./Figures/chapter3/nu-vs-svdd.pdf}
+    \includegraphics[width=0.8\textwidth,keepaspectratio]{./Figures/chapter3/nu-vs-svdd.pdf}
   \caption[Difference \gls{nu-svm} and \gls{svdd}]{Graphical representation of the difference between \gls{nu-svm} (left) and \gls{svdd} (right).
@@
 The center $\mathbf{a}$ is a linear combination of of the support vectors.
-Like the classical \gls{svm} and \gls{svdd} it can be required that all the distances from the data points $x_i$ to the center $\mathbf{a}$ are strict less than $R$ (or equivalent measure, to create a hard margin) or soft margins are allowed by using slack variables $\xi_i$.
-In the case of a soft margin, the penalty is determined by $C$ and the minimization is expressed as \Cref{eq:svdd_objective}.
+Like the classical \gls{svm} and \gls{svdd} it can be required that all the distances from the data points $x_i$ to the center $\mathbf{a}$ are strict less than $R$ (or equivalent measure).
+A soft margin can be allowed by using slack variables $\xi_i$.
+In that case, the penalty is determined by $C$ and the minimization is expressed as \Cref{eq:svdd_objective}.
 This principle is illustrated in the right image or \Cref{fig:nu-vs-svdd}.
@@
     \begin{aligned}
-      \lVert x_i - \mathbf{a} \rVert ^ 2 \leq & R^2 + \xi_i & \mbox{ for all } i = 1, \dots, n \\
+      \lVert x_i - \mathbf{a} \rVert ^ 2 \leq \: & R^2 + \xi_i & \mbox{ for all } i = 1, \dots, n \\
       & \xi_i \geq 0 & \mbox{ for all } i = 1, \dots, n \\
@@
     \begin{aligned}
-    0 \le \alpha_i \le C, \sum_{i} \alpha_i = 1
+    0 \le \alpha_i \le C, \: \sum_{i} \alpha_i = 1
     \end{aligned}
@@
 \end{equation}
-
-\TODO{Note of inner products and that it can be replaced by kernel functions. Good ref: page 4/158 from \cite{tax2002uniform} and \cite{tax1999support}.}
+As with \Cref{eq:svm_lagrange}, the solution of this equation only relies on dot products between the data points in $\vectorsym{x}$ and $\vectorsym{z}$.
+This means that the kernel projection and trick, as discussed in \Cref{subsec:kernels}, can be applied to \gls{svdd} as well \cite{tax1999support,tax2002uniform}.
 
@@
 \end{equation}
+When the Gaussian \gls{rbf} kernel is applied, and in case the data is preprocessed to have unit length (for the \gls{nu-svm} solution), the two different \gls{oc-svm} implementations \gls{nu-svm} and \gls{svdd} are shown to have identical solutions \cite{tax2002uniform,scholkopf2002learning}
 
@@
 \end{figure}
-
-\begin{figure}
-  \centering
-    \includegraphics[width=0.8\textwidth,keepaspectratio]{./Figures/chapter3/svdd-parameter-sigma.pdf}
-  \caption[\gls{svdd} boundary]{The \gls{svdd} method trained on a banana-shaped data set with different $\sigma$-values for the \gls{rbf} kernel.
-Solid circles are support vectors, the dashed line is the boundary.}
-  \label{fig:svdd-boundary-sigma}
-\end{figure}
-
-\TODO{Use \Cref{fig:svdd-boundary} or~\ref{fig:svdd-boundary-sigma}?}
-
-\TODO{Note that \gls{svdd} and \gls{nu-svm} have identical solutions when data is preprocessed to have unit form (in the case of \gls{nu-svm}) \cite{tax2002uniform}.
-and the \gls{rbf} kernel is used.
-Page 234 of \cite{scholkopf2002learning}.}
 
@@
 With a large value for the width $\sigma$, the boudary approximates the spherical boundary.
+The influence of the $\sigma$ parameter on the \gls{svdd} solution is illustrated in \Cref{fig:svdd-boundary-sigma}.
 
-\TODO{More on SVM parameters, \cite{cherkassky2007learning}, section 9.8, page 446}
+\begin{figure}
+  \centering
+    \includegraphics[width=0.8\textwidth,keepaspectratio]{./Figures/chapter3/svdd-parameter-sigma.pdf}
+  \caption[\gls{svdd} boundary]{The \gls{svdd} method trained on a banana-shaped data set with different $\sigma$-values for the \gls{rbf} kernel.
+Solid circles are support vectors, the dashed line is the boundary.}
+  \label{fig:svdd-boundary-sigma}
+\end{figure}
 
-\TODO{Analysis of C/$\nu$-parameter, which sets a lower bound for the fraction of support vectors and an uppor bound for samples that will be outliers.
-\cite{flach2012machine} section 9.8 and \cite{scholkopf2002learning} and other Sch\"olkopf material.}
+As discussed in \Cref{subsec:nu-svm}, the \gls{svm} parameter $C$ (or $\nu$ in case of \gls{nu-svm}) is of high influence on the ``smoothness'' of the decision function.
+It acts as an upper bound to the fraction of outliers and as a lower bound to the fraction of \glspl{sv}.
+A more detailed discussion of the influence of the \gls{svm} model parameters can be found in Section~9.8 of \cite{cherkassky2007learning} and Section~7.4 from \cite{flach2012machine}.
+A detailed discussion of the $\nu$ and kernel parameters can be found in \cite{scholkopf2002learning}.
 
-\TODO{Maybe add section about ``Change indication'' rather than ``detection''?}
+The following chapter will discuss our proposed method, which incorporates the \gls{svdd} algorithm.
+It relates the \gls{oc-svm} model constuction to outlier detection and eventually change detection, leading to finding a temporal segmentation of time series data.
 
--- Notes --
-\begin{itemize}
-  \item Low target rejection rate $f_{T-}$ and low outlier acceptance rate $f_{O+}$.
+% -- Notes --
+% \begin{itemize}
+  % \item Low target rejection rate $f_{T-}$ and low outlier acceptance rate $f_{O+}$.
 When only target examples are present, the first one can be estimated by the number of support vectors that we obtain as a solution of Lagrangian~\ref{eq:svdd_lagrange_kernel}.
-\end{itemize}
+% \end{itemize}
 
--- Literature --
+% -- Literature --
 
-\TODO{These papers need to be referenced and included}
+% ``A geometric approach to support vector machine (SVM) classification'' \cite{mavroforakis2006geometric}. 136, 2006 \\
 
-``A geometric approach to support vector machine (SVM) classification'' \cite{mavroforakis2006geometric}. 136, 2006 \\
+% ``Least squares one-class support vector machine'' \cite{choi2009least}. 27, 2009 \\
 
-``Least squares one-class support vector machine'' \cite{choi2009least}. 27, 2009 \\
+% ``On simple one-class classification methods'' \cite{noumir2012simple}. 2012  --> decouples radius and center optimization, gives fast approximations instead of precise results. \\
 
-``On simple one-class classification methods'' \cite{noumir2012simple}. 2012  --> decouples radius and center optimization, gives fast approximations instead of precise results. \\
-
-``Choosing a small width of the kernels leads to high generalization error as it effectively decouples the separate basis functions of the kernel expansion into very localized functions which is equivalent to memorizing the data, whereas a wide kernel tends to oversmooth.'' \cite{smola1998connection}.
+% ``Choosing a small width of the kernels leads to high generalization error as it effectively decouples the separate basis functions of the kernel expansion into very localized functions which is equivalent to memorizing the data, whereas a wide kernel tends to oversmooth.'' \cite{smola1998connection}.
 \\
 
-``For small values of $\sigma$ almost all $\alpha_i >0$ and the \gls{svdd} resembles a Parzen density estimation.'' \cite{tax2002uniform} page 4, \cite{tax1999support} page 4. \\
+% ``For small values of $\sigma$ almost all $\alpha_i >0$ and the \gls{svdd} resembles a Parzen density estimation.'' \cite{tax2002uniform} page 4, \cite{tax1999support} page 4. \\

### 2013-12-17 Chapters/Chapter4.tex
WIP
Restructure Chapter 5 graphs of data sets
Cleanup of sections
WIP: Finalizing chapter 3

@@
 \begin{itemize}
-  \item Based on the problem statement with current research as stated in Chapter~\ref{Chapter3}
+  \item Based on the problem statement with current research as stated in \Cref{Chapter3}
   \item Adjust method to needs
@@
   \item geprobeerd direct model-reset te doen na change point, maar dan kan je achteraf niet meer verder analyseren.
-  \item Veranderingen in mean geven ``extra'' change point wanneer alle data weer in het window zit. Zie~\ref{fig:jumping_mean_ratios} op de 50-tallen.
+  \item Veranderingen in mean geven ``extra'' change point wanneer alle data weer in het window zit. Zie \Cref{fig:camci_fixed_increasing_mean} op de 50-tallen.
 \end{itemize}
@@
 \TODO{Puts emphasis on two-stage framework by Takeuchi and Yamanishi \cite{takeuchi2006unifying}.}.
+
+\TODO{Put somewhere difference and emphasis of online vs. offline/batch/post-processing}
 
@@
 
- --- OLD ---
-\input{./Chapters/Chapter4/change_indication}
-\input{./Chapters/Chapter4/overview}
-\input{./Chapters/Chapter4/algorithms}
+ % --- OLD ---
+% \input{./Chapters/Chapter4/change_indication}
+% \input{./Chapters/Chapter4/overview}

### 2013-12-17 Chapters/Chapter3/one_class_classification.tex
WIP: Plots Chapter 6

@@
 This will result in a similarity-measure expressed in the properties of outliers.
-This measure is related to other similarity measures, such as \gls{pdf} estimation and the \gls{kl} divergence as discussed in \Cref{subsec:change_detection_time_series}.
+This measure is related to other similarity measures, such as \gls{pdf} estimation and the \gls{kl} divergence as discussed in \Cref{sec:change_detection_time_series}.
 

### 2013-12-17 Chapters/Chapter6/data_gathering.tex
WIP: Plots Chapter 6

@@
 The solid black lines indicate manually annoted change points, the dashes lines indicate the closest discovered change point.
-A still from the video recording, showing Subject $2$ performing a (sprinting) run segment, is in \Cref{fig:data_gathering_still_1_sprint}.
-
-\TODO{Add found change points to plots}
-
-%---- Run 1: walk-run-roemer
-\begin{figure}
-\centering
-  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/run-1-walk-run-roemer/data_plot_rot_annotated.eps}
-  \caption[R1: rotation]{Run 1: Walk-run-roemer, rotation}
-  \label{fig:data_gathering_run_1_rot}
-\end{figure}
-
-\begin{figure}
-\centering
-  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/run-1-walk-run-roemer/data_plot_mag_annotated.eps}
-  \caption[R1: mag]{Run 1: Walk-run-roemer, Mag}
-  \label{fig:data_gathering_run_1_mag}
-\end{figure}
-
-\begin{figure}
-\centering
-  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/run-1-walk-run-roemer/data_plot_acc_annotated.eps}
-  \caption[R1: accelerometer]{Run 1: Walk-run-roemer, accelerometer}
-  \label{fig:data_gathering_run_1_acc}
-\end{figure}
-
-%---- Run 2: walk-run-jos
-\begin{figure}
-\centering
-  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/run-2-walk-run-jos/data_plot_rot_annotated.eps}
-  \caption[R2: rotation]{Run 2: Walk-run-jos, rotation}
-  \label{fig:data_gathering_run_2_rot}
-\end{figure}
-
-\begin{figure}
-\centering
-  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/run-2-walk-run-jos/data_plot_mag_annotated.eps}
-  \caption[R2: mag]{Run 2: Walk-run-jos, Mag}
-  \label{fig:data_gathering_run_2_mag}
-\end{figure}
-
-\begin{figure}
-\centering
-  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/run-2-walk-run-jos/data_plot_acc_annotated.eps}
-  \caption[R2: accelerometer]{Run 2: Walk-run-jos, accelerometer}
-  \label{fig:data_gathering_run_2_acc}
-\end{figure}
+A still from the video recording, showing Subject $2$ performing a (sprinting) run segment, is in \Cref{fig:stills_subject_2_sprinting}.
 
@@
 \TODO{Add plots}
-For Subject $2$ the still of the video recordings are in \Cref{fig:data_gathering_still_1_still_walk_to_run,fig:data_gathering_still_1_running,fig:data_gathering_still_1_running_2,fig:data_gathering_still_1_walk}.
-The stills for Subject $1$ are in \Cref{fig:data_gathering_still_2_run,fig:data_gathering_still_2_walk}.
+For Subject~$2$ the still of the video recordings are in the upper row of \Cref{fig:stills_subject_1_and_2}, for Subject~$1$ in the lower row.
+
+\begin{figure}
+  \centering
+  \begin{subfigure}[b]{1\textwidth}
+    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/run-1-walk-run-roemer/data_plot_acc_annotated.eps}
+    \caption{Accelerometer}
+    \label{fig:recording_subject_1_run_1_accelerometer}
+  \end{subfigure}
+
+  \begin{subfigure}[b]{1\textwidth}
+    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/run-1-walk-run-roemer/data_plot_rot_annotated.eps}
+    \caption{Rotation}
+    \label{fig:recording_subject_1_run_1_rotation}
+  \end{subfigure} \\
+
+  \begin{subfigure}[b]{1\textwidth}
+    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/run-1-walk-run-roemer/data_plot_mag_annotated.eps}
+    \caption{Magnetic Field}
+    \label{fig:recording_subject_1_run_1_magnetic}
+  \end{subfigure} \\
+  \caption[Plots]{Annotated plots of a full recording performed by Subject~$2$.
+The vertical lines are manually determined change points.
+The label above each segment indicates the current performed activity.}\label{fig:plots_subject_1}
+\end{figure}
 
@@
 \Cref{fig:stills_subject_2_change_point} shows the change point between the two activities.
-\Cref{fig:stills_subject_1_running_clockwise} shows the running in clockwise motion.}\label{fig:stills_subject_2}
+\Cref{fig:stills_subject_1_running_clockwise} shows the running in clockwise motion.}\label{fig:stills_subject_1_and_2}
 \end{figure}
@@
 The stairs are shaped in an semi-circular form, with a small vertical segment on each floor.
-The still of \Cref{fig:data_gathering_still_3_descending} illustrates the circular shape of the stairs.
-In \Cref{fig:data_gathering_still_3_ascending} the small flat segment is visible.
-Between the stairs the subject walked in a hallway with a $180^{\circ}$ counter-clockwise turn, as shown in \Cref{fig:data_gathering_still_3_walk}.
+The still of \Cref{fig:stills_subject_3_downstairs} illustrates the circular shape of the stairs.
+In \Cref{fig:stills_subject_3_upstairs} the small flat segment is visible.
+Between the stairs the subject walked in a hallway with a $180^{\circ}$ counter-clockwise turn, as shown in \Cref{fig:stills_subject_3_walking}.
 The recorded sensor data of the performed activities are visible in the plots of \Cref{fig:data_gathering_run_3_rot,fig:data_gathering_run_3_mag,fig:data_gathering_run_3_acc}.

### 2013-12-18 Chapters/Chapter4.tex
WIP: chapter 4

@@
 
-\chapter{Proposed method}
+\chapter{Application to Human Activity Time Series Data}
 
@@
 
-\lhead{Chapter 4. \emph{Proposed method}} % This is for the header on each page - perhaps a shortened title
+\lhead{Chapter 4. \emph{Application to Human Activity Time Series Data}} % This is for the header on each page - perhaps a shortened title
 
 %----------------------------------------------------------------------------------------
-\section{Outline}
-\emph{Not intented for the reader.}
+
+Ideen naam methode:
 \begin{itemize}
-  \item Based on the problem statement with current research as stated in \Cref{Chapter3}
-  \item Adjust method to needs
-  \item Explain using graphs, pseudo-algorithms. Make clear distinction in origin of ideas and why to apply
+  \item Human Activity Support Vector Change Detection
+  \item Inertial Sensor Time Series Svm Temporal Segmentation
+  \item Support Vector bases Human Activity Temporal Segmentation
+  \item Human Activity Temporal Segmentation by Support Vector Machines: HATS-SVM
+  \item HATS with One-Class SVM: HATS-OCS, HATS-OS, HATSOS
+  \item OCS-HATS
+  \item HATS-SVDD
 \end{itemize}
 
-\subsection{Experiment notes}
+\section{Experiment notes}
 \begin{itemize}
   \item ICSS/CUSUM is goed in het vinden van variance changes. Niet goed in mean veranderingen.
-  \item Problemen met SVM methode: ook na de verandering blijft de threshold nog stijgen;
   \item geprobeerd direct model-reset te doen na change point, maar dan kan je achteraf niet meer verder analyseren.
-  \item Veranderingen in mean geven ``extra'' change point wanneer alle data weer in het window zit. Zie \Cref{fig:camci_fixed_increasing_mean} op de 50-tallen.
 \end{itemize}

### 2013-12-18 Chapters/Chapter4/data_gathering.tex
WIP
WIP: chapter 4

@@
 In this section we briefly discuss the different data gathering methods used for the change detection algorithms and experiments.
-\Cref{subsec:data_gathering_artificial} reviews the artificial data sets we will use.
+\Cref{subsec:data_gathering_artificial} discusses the artificial data sets we will use.
 In \Cref{subsec:data_gathering_real_world} an overview of the real-world data sets used is provided.
@@
 \end{equation}
-In the different data series the mean and variance of the Gaussian random variable $\epsilon_t$ differs and changes at pre-determined change points.
+Over the different data series the mean and variance of the Gaussian random variable $\epsilon_t$ differs and changes at pre-determined change points.
 Using this data set an objective quality measure over the change detection methods can be obtained and compared.
@@
 Activities performed include sitting, standing, walking, running in a straight and curved line, and walking up- and downstairs.
-Our method uses the signals from the accelerometer, gravity, and rotation sensors.
+Our method uses the signals from the accelerometer, magnetic field, and rotation sensors.
 These time series data are used to detect change points.
@@
 This application was chosen for its convenient data format of the sensor recording and its regularity of the sampling interval.
-Table~\ref{tab:recorded_metrics} lists the recorded metrics.
+Table~\ref{tab:recorded_metrics} lists all the recorded metrics.
 For our expiments we used the data for the accelerometer, magnatief field and rotation.
@@
     \hline \hline
-    Accelerometer & Acceleration force along each axis. & $m/s^2$ & $-20$ -- $20$ \\
+    Accelerometer & Acceleration force along each axis (including gravity). & $m/s^2$ & $-20$ -- $20$ \\
     \hline
-    Gravity & Force of gravity along each axi.s & $m/s^2$ & $-10$ -- $10$\\
+    Gravity & Force of gravity along each axis. & $m/s^2$ & $-10$ -- $10$\\
     \hline
-    Gyroscope & & $rad/s$ & $-15$ -- $15$\\
+    Gyroscope & Rate of rotationg around each axis. & $rad/s$ & $-15$ -- $15$\\
     \hline
@@
     \hline
-    Linear acceleration & & $m/s^2$ & $-20$ -- $20$ \\
+    Linear acceleration & Acceleration force along each axis (excluding gravity). & $m/s^2$ & $-20$ -- $20$ \\
     \hline
@@
     \hline
-    Orientation & & & $-100$ -- $360$ \\
+    Orientation & Degrees of rotation around the three physical axis. & Degrees & $-100$ -- $360$ \\
     \hline
-    Rotation & & & $-1$ -- $1$\\
+    Rotation & Measre of rotation around the device's rotation axis. & Unitless & $-1$ -- $1$\\
     \hline

### 2013-12-18 Chapters/Chapter4/model_construction.tex
WIP: chapter 4

@@
 Thus the algorithm is initialized by selection the first $\ceil{\frac{1}{C}}$ objects for the working set.
-In every loop of the algorithm the number of elements added are at least equal to the number of elements removed.
+In every loop of the algorithm the number of elements added are at least enough to keep $\ceil{\frac{1}{C}}$ objects in the window, after removing the oldest objects.
 The optimality of the algorithm is proved by analysis of the \gls{kkt} conditions \cite{tax2003online}.
@@
 This re-modeling of the \gls{svm} representation is beneficial for our purpose, since we are only interested in sudden changes between activities.
-The following section will discuss the features extracted from the constructed \gls{oc-svm} model at every loop-step of the algorithm, before \Cref{sec:method_change_detection} describes how these features are used to detect change.
+The following section will discuss the features extracted from the constructed \gls{oc-svm} model at every loop-step of the algorithm, and \Cref{sec:method_change_detection} describes how these features are used to detect change.

### 2013-12-18 Chapters/Chapter4/model_features.tex
WIP: chapter 4

@@
 In the previous section we have discussed the \gls{isvdd} method, which creates a \gls{oc-svm} representation of a working set of data objects at every step of the algorithms loop.
-This section shows how we interpret the constructed model and extract features to obtain a measure which can be used for a indication of change points.
-The following section discusses how this obtained measure is used to indicate change.
+This section shows how we interpret the constructed model and extract features to obtain a measure which can be used for an indication of change points.
+The next section discusses how this obtained measure is used to indicate change.
 
@@
 This model is obtained by the minimization of \Cref{eq:svdd_objective}, which incorporates the radius $R$ of the sphere and the distances $\vectorsym{\xi}$ from the outliers to the boundary.
-We will use the the radius $R$ of the hypersphere as an indication of change.
+We will use the radius $R$ of the hypersphere as an indication of change.
 
@@
 \end{inparaenum}
-Since in \gls{occ} situation typically there are (almost) no examples of outlier objects, Tax and Duin construct a method to generate outliers based on the assumption that the outliers are uniformly distributed around the target set.
+Since in \gls{occ} situations typically there are (almost) no examples of outlier objects, Tax and Duin construct a method to generate outliers based on the assumption that the outliers are uniformly distributed around the target set.
 To minimize the error, calculated by the fractions $f_{T-}$ and $f_{O+}$, we should minimize the volume of the target data description (\ie the boundary of \gls{svdd}).
-This is because the fraction of accepted outliers $f_{O+}$ is an estimate of the volume of the target data description, with respect to the volume of the outlier distribution.
+That is because the fraction of accepted outliers $f_{O+}$ is an estimate of the volume of the target data description, with respect to the volume of the outlier distribution.
 Tax and Duin provide a method to optimize the parameters of the \gls{svdd} method, \ie the trade-off parameter $C$ and the \gls{rbf} kernel width $\sigma$.
@@
 During the \gls{isvdd} algorithm the volume of the hypersphere is being minimized.
-This means that only the radius of the sphere will change during the algorithm.
+This means that only the radius of the sphere will change during the analysis over the time series data.
 We will interpret changes in the radius $R$ with the same continuity assumptions as mentioned in \Cref{subsec:kernels}.
@@
 We thereby follow the setup of the unifying framework by Takeuchi and Yamanishi~\cite{takeuchi2006unifying}, of which this section described the first stage.
-
-\TODO{Put the SVM method more in the light of dimensionality reduction. Applies to text in whole thesis...}

### 2013-12-19 Chapters/Chapter4.tex
WIP: Chapter 4

@@
 \lhead{Chapter 4. \emph{Application to Human Activity Time Series Data}} % This is for the header on each page - perhaps a shortened title
-
-%----------------------------------------------------------------------------------------
 
@@
 
-\section{Experiment notes}
-\begin{itemize}
-  \item ICSS/CUSUM is goed in het vinden van variance changes. Niet goed in mean veranderingen.
-  \item geprobeerd direct model-reset te doen na change point, maar dan kan je achteraf niet meer verder analyseren.
-\end{itemize}
-
-\begin{description}
-  \item[Data gathering] Explain data gathering methods. Refer to chapter 5 and 6 for articifial and real-world details.
-  \item[Model construction] Explain SVDD model construction and updating.
-  \item[Model properties] Explain the SVDD properties that are used to calculate the change indication.
-  \item[Change detection] Explain the interpretation of the properties, the possible (modulair) methods to give change indication
-\end{description}
-
-\TODO{Write chapter introduction/outline. Give overview of method approach and refer to each section for details of that stage.}
-
-\TODO{Puts emphasis on two-stage framework by Takeuchi and Yamanishi \cite{takeuchi2006unifying}.}.
-
-\TODO{Put somewhere difference and emphasis of online vs. offline/batch/post-processing}
-
 \begin{figure}
   \centering
-    \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{./Figures/chapter4/method_setup_short.pdf}
+    \includegraphics[width=\textwidth]{./Figures/chapter4/method_setup.pdf}
   \caption[Method setup]{Schematic overview of the change detection method.
@@
 Section \ref{sec:method_model_features} describes which features of the model are used for the final change indication algorithm, discussed in Section \ref{sec:method_change_detection}.}
-  \label{fig:method_overview_short}
+  \label{fig:method_overview}
 \end{figure}
 
+This chapter will introduce our method and setup for the temporal segmentation of human activities, using an \gls{oc-svm} approach.
+The setup follows the structure as illustrated in \Cref{fig:method_overview} and further discussed in \Cref{sec:method_overview}.
+It starts with the data collection phase, in which we both use artificial and real-world data.
+The data sets are discussed in \Cref{sec:data_gathering}.
+After (optional) pre-processing, the \gls{svdd} algorithm is used to create a model representation, which is iteratively updated.
+In \Cref{sec:method_model_construction} this construction phase is discussed and shows the windowed approach.
+From this constructed model we can extract some features.
+In our case, we use the approximated radius $R$ of the hypersphere to detect changes.
+\Cref{sec:method_model_features} goes into detail about the rationale of using that feature.
+Finally, in \Cref{sec:change_detection_methods} we discuss two different, but simple, methods to detect changes in the time series data, based on the extracted radius $R$.
+
+\input{./Chapters/Chapter4/method_overview}
 \input{./Chapters/Chapter4/data_gathering}
@@
 \input{./Chapters/Chapter4/change_detection}
-
- % --- OLD ---
-% \input{./Chapters/Chapter4/change_indication}
-% \input{./Chapters/Chapter4/overview}

### 2013-12-19 Chapters/Chapter4/change_detection.tex
WIP: Chapter 4

@@
 \section{Change Detection}\label{sec:method_change_detection}
-In the following sections we haved discussed what data is used for the model construction how the model is interpreted in the context of change detection.
-In this section we discuss the methods applicable to the extracted radius $R$ of the hypersphere, in order to indicate a change in the underlying distribution properties.
-This employs the second stage of our algorithm, analogue to the second stage of the unifying framework by Takeuchi and Yamanishi~\cite{takeuchi2006unifying}.
-
-In the previous sections we have discussed how a multi-dimensional signal can be reduced to a single dimension time series data based on the radius $R$.
-We have argued that in the case of a change in the underlying data, where it becomes more heterogeneous, we expect an increase in the radius $R$.
+In the previous sections we have discussed what data is used for the model construction, and how that model is interpreted in the context of change detection.
+We have shown how the multi-dimensional signal is reduced to a single dimension time series data, based on the extracted radius $R$ from the constructed hypersphere.
+This metric is used to discover changes in the underlying data generation process.
+We argued that in the case of the data objects set becomes more heterogeneous, we expect an increase in the radius $R$.
 In the same manner we expect that when the change is in the oldest data objects of the working set the radius $R$ will decrease, since the data becomes more homogeneous.
@@
 
-\begin{figure}
-  \centering
-    \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{./Figures/chapter4/expected_behaviour.pdf}
-  \caption[Expected radius behavior]{Expected behavior of the radius $R$ of the hypersphere.
-The upper part shows a typical sinusoidal time series signal.
-The lower graph visualizes an abstract expectation of the values of $R$.
-Five possible windows are illustrated.
-The left, middle, and right windows cover an area of homogeneous signal.
-The expected value of $R$ is low.
-The other two windows cover a change entering and leaving the window, respectively.
-At these locations the radius $R$ is expected to increase temporarily.}
-  \label{fig:radius_expectation}
-\end{figure}
+In this section we discuss the methods applicable to the extracted radius $R$ of the hypersphere, in order to discover a change in the underlying distribution properties.
+This employs the second stage of our algorithm, analogue to the second stage of the unifying framework by Takeuchi and Yamanishi~\cite{takeuchi2006unifying}.
+It is also based on the \gls{svcpd} method by Camci~\cite{camci2010change}, although we have slightly simplified the algorithm.
+
+In the \gls{svcpd} the first step for a new data point $\vectorsym{z}$ is to check whether the data point lies within the enclosing (hyperspherical) boundary of the constructed model.
+This will label the new data point as an in- or outlier, realtive to the current data set.
+Depending on that outcome, the model is updated to include the new data point.
+The second step is to inspect the (changed) model properties, \ie the radius $R$.
+In our method we only employ the second step: the inspection of the radius $R$.
+We regard the first step (determine if $\vectorsym{z}$ is an outlier) to be superfluent: in case it is an outlier, the model update step will change the radius $R$ in order to let the boudary include the new data point $\vectorsym{z}$.
+Would the new data object $\vectorsym{z}$ be already an inlier, then the updated model would not have a significant change in radius $R$.
+Our algorithm is outlined in \Cref{tab:algorithm_proposed_method}.
 
@@
 We conclude this section by discussing some post-processing techniques we have used, to decrease the number of false positives.
-
-\subsection{CUSUM based methods}
-Many of the simple change detection methods are based on \gls{cusum}, originally introduced by Page~\cite{page1954continuous}.
-Many variations and extensions of the method are proposed and used (\cite{inclan1994use,alippi2006adaptive,hsu2007mosum}).
-Here we will discuss the simple form, which only detects changes for an increase of the examples, in this case the radius $R$.
-The cumulative sum for all the values is calculated:
-\begin{equation}
-  S_n = \sum_{k=1}^n r_k
-\end{equation}
-and a change is detected when the value of the cumulative sums minus the minimum encountered exceeds a predetermined threshold $h$:
-\begin{equation}
-  S_n - \operatorname*{min}_{0 \le i < n} S_i \ge h.
-\end{equation}
-
-Extensions of the \gls{cusum} method have been proposed.
-Amongst others, extensions which extend its use to two-threshold methods, and methods appropiate for changes in mean or variance \cite{inclan1994use}.
-Whilst adaptive methods have been proposed \cite{alippi2006adaptive}, most methods require manual determination of the threshold parameter $h$.
-
-\TODO{This subsection has double information. Cleanup. Mention iterated cumulative sum of squares of inclan \cite{inclan1994use}.}
-
-\subsection{Ratio-thresholding}\label{subsec:ratio_thresholding}
-The second example of an algorithm that interprets the change indication time series and transforms it to change detection is the radius ratio thresholding method used by Camci~\cite{camci2010change}.
-The method shares characteristics with \gls{cusum}, since it relies on summation of historic data and thresholds for change detection.
-The method calculates the radius ratio $\hbar_t$ by taking the ratio of the current radius $r_t$ at time $t$ to the average radii since the last change point $y$:
-\begin{equation}\label{eq:ratio_radius}
-  \hbar_t = \frac{r_t}{\operatorname*{mean}(r_{y:t-1})},
-\end{equation}
-where $\operatorname*{mean}(r_{y:t-1})$ is the average of the previous approximated radii.
-By using only historic data, this method can be incorporated in an online change detection method.
-The ratio $\hbar_t$ is compared with two thresholds: the low threshold $th_\text{low}$ and the high threshold $th_\text{high}$.
-When the data objects in the working set becomes more homogeneous at time $t$, the radius $R$ of the hypersphere and thereby the ratio $\hbar_t$ will decrease.
-When the new value of $\hbar_r$ is lower than the low threshold $th_\text{low}$, a change detection is observed for time $t$.
-The same holds for a more heterogeneous set of data and an increase of $\hbar_r$ higher than $th_\text{high}$.
-
-The appropriate values for the thresholds are highly depended on the characteristics of the data and the trade-off parameter $C$ (as discussed in \Cref{subsec:oc-svm-svdd}).
-This trade-off parameters regulates the fraction of data objects that will be rejected from the constructed model.
-Since a high value of $C$ assigns a high penalty to outliers, data objects are more likely to be incorporated into the hypersphere by increasing the radius.
-For low values of $C$ the cost of rejecting data objects is, compared to the benefits of a smaller hypersphere, relatively low.
-This gives a relation between the parameters $C$, $th_\text{low}$, and $th_\text{high}$ and the sensitivity of the radius ratio based thresholding method.
-A high value of $C$, or values close to $1$ for $th_\text{low}$ and $th_\text{high}$ result in a sensitive change detection procedure.
-The reverse results in less sensitive methods.
-The correct values for the intended sensitivity need to be empirically determined.
-
-\subsection{Post-processing}
-In our method we use the ratio based thresholding, since it is also used by Camci~\cite{camci2010change}.
-From preliminary experiments we observed that a single change in the data can cause many detected change points, using the ratio based method.
-This phenomenon can be explained by the nature of the thresholding method.
-Since a single (high or low) threshold is set for ratio of the radii, high values are considered to be a change point before the highest point of the peak.
-As illustrated in \Cref{fig:tresholding}, the peak at $t_1$ represents a single change point.
-The region from $t_2$ to $t_3$ is strictly increasing.
-At $t_2$ a change point is detected, since $\hbar_t$ exceeds the threshold.
-But since it is increasing up to $t_3$, all values of $\hbar_t$ in that period will trigger the ratio based method to indicate a change point.
-
-To overcome this problem, we apply a post-processing method on the generated change points.
-All the change points that are within a time period $\delta$ of each other are merged together.
-\TODO{Merged to the front, or to the back, depending on high/low threshold exceeding?}
-
-\begin{figure}
-  \centering
-    \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{./Figures/chapter4/signal_threshold.pdf}
-  \caption[Thresholding]{Abstract illustration of the radius ratio thresholding.
-The peak at $t_1$ generates a single change point.
-The region between $t_2$ and $t_3$ generates multiple change points, since the value of $\hbar_t$ keeps increasing.}
-  \label{fig:tresholding}
-\end{figure}
 
@@
 
-\TODO{Add reference to Table~\ref{tab:algorithm_proposed_method}.}
+\subsection{CUSUM based methods}
+Many of the simple change detection methods are based on \gls{cusum}, originally introduced by Page~\cite{page1954continuous}.
+Other variations and extensions of the method are proposed and used (\cite{inclan1994use,alippi2006adaptive,hsu2007mosum}).
+Here we will discuss the simple form, which only detects changes as an increase in value of the examples, in this case the radius $R$.
+The cumulative sum for all the values is calculated:
+\begin{equation}
+  S_n = \sum_{k=1}^n r_k
+\end{equation}
+and a change is detected when the value of the cumulative sums minus the minimum encountered exceeds a predetermined threshold $h$:
+\begin{equation}
+  S_n - \operatorname*{min}_{0 \le i < n} S_i \ge h.
+\end{equation}
+
+Extensions of the \gls{cusum} method have been proposed.
+Amongst others, extensions which extend its use to two-threshold methods, and methods appropiate for changes in mean or variance \cite{inclan1994use}.
+Whilst adaptive methods have been proposed \cite{alippi2006adaptive}, most methods require manual determination of the threshold parameter $h$.
+
+\subsection{Ratio-thresholding}\label{subsec:ratio_thresholding}
+The second example of an algorithm that interprets the change indication time series and transforms it to change detection is the radius ratio thresholding method used by Camci~\cite{camci2010change}.
+The method shares characteristics with \gls{cusum}, since it relies on summation of historic data and thresholds for change detection.
+The method calculates the radius ratio $\hbar_t$ by taking the ratio of the current radius $r_t$ at time $t$ to the average radii since the last change point $y$:
+\begin{equation}\label{eq:ratio_radius}
+  \hbar_t = \frac{r_t}{\operatorname*{mean}(r_{y:t-1})},
+\end{equation}
+where $\operatorname*{mean}(r_{y:t-1})$ is the average of the previous approximated radii.
+By using only historic data, this method can be incorporated in an online change detection method.
+The ratio $\hbar_t$ is compared with two thresholds: the low threshold $th_\text{low}$ and the high threshold $th_\text{high}$.
+When the data objects in the working set become more homogeneous at time $t$, the radius $R$ of the hypersphere and thereby the ratio $\hbar_t$ will decrease.
+When the new value of $\hbar_r$ is lower than the low threshold $th_\text{low}$, a change detection is observed for time $t$.
+The same holds for a more heterogeneous set of data and an increase of $\hbar_r$ higher than $th_\text{high}$.
+This is illustrated in \Cref{fig:tresholding}.
+
+The appropriate values for the thresholds are highly depended on the characteristics of the data and the trade-off parameter $C$ (as discussed in \Cref{subsec:oc-svm-svdd}).
+This trade-off parameters regulates the fraction of data objects that will be rejected from the constructed model.
+Since a high value of $C$ assigns a high penalty to outliers, data objects are more likely to be incorporated into the hypersphere by increasing the radius.
+For low values of $C$ the cost of rejecting data objects is, compared to the benefits of a smaller hypersphere, relatively low.
+This gives a relation between the parameters $C$, $th_\text{low}$, $th_\text{high}$, and the sensitivity of the radius ratio based thresholding method.
+A high value of $C$, or values close to $1$ for $th_\text{low}$ and $th_\text{high}$ result in a sensitive change detection procedure.
+The reverse results in less sensitive methods.
+The correct values for the intended sensitivity need to be empirically determined.
+
+\begin{figure}
+  \centering
+    \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{./Figures/chapter4/signal_threshold.pdf}
+  \caption[Thresholding]{Abstract illustration of the radius ratio thresholding.
+The peak at $t_1$ generates a single change point.
+The region between $t_2$ and $t_3$ generates multiple change points, since the value of $\hbar_t$ keeps increasing.}
+  \label{fig:tresholding}
+\end{figure}
+
+\subsection{Post-processing}
+In our method we use the ratio based thresholding, since it is also used by Camci~\cite{camci2010change}.
+From preliminary experiments we observed that a single change in the data can cause many detected change points, using the ratio based method.
+This phenomenon can be explained by the nature of the thresholding method.
+Since a single (high or low) threshold is set for ratio of the radii, high values are considered to be a change point before the highest point of the peak.
+As illustrated in \Cref{fig:tresholding}, the peak at $t_1$ represents a single change point.
+The region from $t_2$ to $t_3$ is strictly increasing.
+At $t_2$ a change point is detected, since $\hbar_t$ exceeds the threshold.
+But since it is increasing up to $t_3$, all values of $\hbar_t$ in that period will trigger the ratio based method to indicate a change point.
+
+To overcome this problem, we apply a post-processing method on the generated change points.
+All the change points that are within a time period $\delta$ of each other are merged together.
+
+In the following two chapters we will take a detailed look at the used data sets and apply the method as described in this method to those sets.

### 2013-12-19 Chapters/Chapter4/method_overview.tex
WIP: Chapter 4

--- 2013-11-18 Chapters/Chapter4/overview.tex
@@
 % !TEX root = ../../main.tex
-\section{Overview}\label{sec:method_overview}
+\section{Algorithm overview}\label{sec:method_overview}
+The proposed method of this thesis follows the unifying framework as introduced by Takeuchi and Yamanishi \cite{takeuchi2006unifying} and an similar implementation by Camci \cite{camci2010change} with \glspl{svm}.
+The unifying framework relates the detection of outliers with change points and divides the whole process in two stages.
+The first stage determines the outliers in a time series by giving a score based on the deviation from a learned model, and thereby creates a new time series.
+The second stage runs on that new created time series and calculates a average over a window of the outlier scores.
+The problem of change detection is then reduced to outlier detection over that average-scored time series.
+This method is named \gls{changeFinder} by the authors.
+The implementation by Camci, which uses \glspl{svm} to detect changes is named \acrlong{svcpd}.
 
-\begin{figure}
-  \centering
-    \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{./Figures/chapter4/method_setup.pdf}
-  \caption[Method setup]{Schematic representation of the change detection method. \TODO{Gerard noemt dit misschien ook goed voor bij de inleiding, als motivatie voor onderzoek. Overnemen?}}
-  \label{fig:method_overview}
-\end{figure}
+Whereas \gls{changeFinder} uses double probability estimation algorithm, our approach follows \gls{svcpd} by constructing a \gls{svm} over a sliding window.
+The \gls{svcpd} algorithm uses the location of new data points in the feature space $\mathcal{F}$ with respect to the hypersphere and the hypersphere's radius $R$ to determine whether the new data point represents a change point.
+Our approach is a slight simplification of the algorithm: we only use the updated model properties (\ie the radius $R$) to detect a change.
+This difference is further discussed and justified in \Cref{sec:method_change_detection}.
 
@@
 A more detailed explanation of the ``Update model'' step follows.
-This section is finalized with the experiments setup, annotation of data streams and quality measure.
 
-\subsection{Change detection method}
-As graphically represented in \Cref{fig:method_overview}, the change detection method starts by processing the data from sensor, such as the accelerometer, magnetic orientation and rotation measures \footnote{Something about the origin of the streams; all from the same sensor or different sensors?}.
+As graphically represented in \Cref{fig:method_overview}, the change detection method starts by processing the data from sensor, such as the accelerometer, magnetic orientation and rotation metrics.
 
@@
 Every update the oldest $s$ data points are removed from and $s$ new data points are added to the \gls{svdd} model.
-The model is (partially) reconstructed and new model properties, such as the radius of the hypersphere and the number of outliers, are the result of this step \footnote{Other measures are also possible, for instance the distance from all the outliers to the boundary of the hypersphere}.
-\TODO{MORE ON THE UPDATING STEP}
+The \gls{isvdd} algorithm by Tax and Duin~\cite{tax2002uniform} is used for efficient incremental model updating.
+The model is (partially) reconstructed and new model properties, such as the radius of the hypersphere, is the result of this step \footnote{Other measures are also possible, for instance the distance from all the outliers to the boundary of the hypersphere or the number of outliers}.
 
@@
 In our setup we used the \gls{rt} and \gls{cusum} methods, to show the modularity of this step.
-
-\subsection{Model updating}
-
-\subsection{Experiments}
-For the experiments we used a HTC Sensation XE smartphone as recording device.
-The activities were recorded using a free Android application \cite{sensorlogger}.
-This application was chosen for its convenient data format of the sensor recording and its regularity of the sampling interval.
-Table~\ref{tab:recorded_metrics} lists the recorded metrics.
-For our expiments we used the data for the accelerometer, magnatief field and rotation.
-
-\begin{center}\begin{table}
-  \begin{tabulary}{\textwidth}{|l|L|c|c|}
-    \hline
-    Metric & Description & Units of measure & Typical range \\
-    \hline \hline
-    Accelerometer & Acceleration force along each axis. & $m/s^2$ & $-20$ -- $20$ \\
-    \hline
-    Gravity & Force of gravity along each axi.s & $m/s^2$ & $-10$ -- $10$\\
-    \hline
-    Gyroscope & & $rad/s$ & $-15$ -- $15$\\
-    \hline
-    Light & Light sensitive sensor at the front of the phone. & & $0$ -- $10000$ \\
-    \hline
-    Linear acceleration & & $m/s^2$ & $-20$ -- $20$ \\
-    \hline
-    Magnetic field & Geomagnetic field strength along each axis. & $\mu T$ & $-60$ -- $60$ \\
-    \hline
-    Orientation & & & $-100$ -- $360$ \\
-    \hline
-    Rotation & & & $-1$ -- $1$\\
-    \hline
-  \end{tabulary}
-  \caption[Measured metrics]{Measured metrics. The set of axis is always the triple (x, y, z) direction.}
-  \label{tab:recorded_metrics}
-\end{table}\end{center}
+This final step is further discussed in \Cref{sec:change_detection_methods}.

### 2013-12-19 Chapters/Chapter4/model_features.tex
WIP: Chapter 4

@@
 
+\begin{figure}
+  \centering
+    \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{./Figures/chapter4/expected_behaviour.pdf}
+  \caption[Expected radius behavior]{Expected behavior of the radius $R$ of the hypersphere.
+The upper part shows a typical sinusoidal time series signal.
+The lower graph visualizes an abstract expectation of the values of $R$.
+Five possible time windows are illustrated.
+The left, middle, and right windows cover an area of homogeneous signal.
+The expected value of $R$ is low.
+The other two windows cover a change entering and leaving the window, respectively.
+At these locations the radius $R$ is expected to increase temporarily.}
+  \label{fig:radius_expectation}
+\end{figure}
+
 Since in our method the parameters $C$ and $\sigma$ are kept constant, and thereby also the fraction of target objects being rejected, the only free parameter is the radius $R$.
@@
 When, instead, the data objects change from a heterogeneous set to a more homogeneous, we expect the radius to decrease in value, since the data objects are closer to each other in the feature space.
+This relation between the signal and expected radius $R$ is illustrated in \Cref{fig:radius_expectation}.
+It shows that in the two time windows that cover a heterogeneous set of data objects, the radius of the hypersphere is expected to be relatively high.
 

### 2013-12-19 Chapters/Chapter5.tex
WIP: Finalizing chapter 3
WIP: chapter 4
FrontmatteR

@@
 \lhead{Chapter 5. \emph{Artificial data results}} % This is for the header on each page - perhaps a shortened title
-
-%----------------------------------------------------------------------------------------
-\section{Outline}
-\emph{Not intented for the reader.}
 \begin{itemize}
-  \item Compare proposed method with methods of \Cref{Chapter2}
-  \item Provide KL-analysis (as Takeuchi does)
-  \item Provide plots, tables, graphs, error rates, precision, etc.
-  \item Apply to a multiple of data, to compare to previous research - use that data
-  \item Give theoretical analysis about performance. Big-O, memory, run-time, precision.
-  \item This sections needs programmed implementations of own method and the ones compared
+  \item Problemen met SVM methode: ook na de verandering blijft de threshold nog stijgen;
+  \item Veranderingen in mean geven ``extra'' change point wanneer alle data weer in het window zit. Zie \Cref{fig:camci_fixed_increasing_mean} op de 50-tallen.
 \end{itemize}
 
-In this chapter we will apply the proposed algorithm to artificial data sets.
-For comparison, we use some of the data sets that are used by Camci~\cite{camci2010change} and Takeuchi \etal \cite{takeuchi2006unifying}.
+In this chapter we will apply the proposed algorithm as stated in the previous chapter to artificial data sets.
+For comparison, we use some of the data sets that are used by Camci~\cite{camci2010change} and Takeuchi and Yamanishi~\cite{takeuchi2006unifying}.
 These data sets are generated by Gaussian noise distributions.
@@
 
+\TODO{Add short section on quality metrics}
+
 \input{./Chapters/Chapter5/results}

### 2013-12-19 Chapters/Chapter5/artificial_data_generation.tex
FrontmatteR

@@
   \item \textbf{Reduced increasing mean:} at each change point $y$ the mean is increased with $\Delta(y) = 10 - y$.
-  This set origins as the first data set of Takeuchi \etal \cite{takeuchi2006unifying}.
+  This set origins as the first data set of Takeuchi and Yamanishi~\cite{takeuchi2006unifying}.
   In contrast with the previous set, in this set the absolute difference in mean is reduced over time.
@@
   \item \textbf{Alternating variance:} at each change point $y$ the variance $\sigma^2$ is set to $9.0$ when $y$ is odd, and to $1.0$ otherwise.
-  This set origins as the thirds sets from Camci~\cite{camci2010change} and Takeuchi \etal \cite{takeuchi2006unifying}.
+  This set origins as the thirds sets from Camci~\cite{camci2010change} and Takeuchi and Yamanishi~\cite{takeuchi2006unifying}.
   This data set is used to measure the ability of detecting changes in decrease of variance, as Camci~\cite{camci2010change} claims is not trivial.

### 2013-12-19 Chapters/Chapter5/results.tex
FrontmatteR

@@
 \section{Results}\label{sec:artificial_data_results}
-For an objective measure of quality of the proposed method applied to the four aforementioned data sets, we employ two metrics, as defined by Takeuchi \etal \cite{takeuchi2006unifying}.
+For an objective measure of quality of the proposed method applied to the four aforementioned data sets, we employ two metrics, as defined by Takeuchi and Yamanishi~\cite{takeuchi2006unifying}.
 The first metric is the $\operatorname*{average \_ benefit}$, for which we first need to define the notion of $\operatorname*{benefit}$ for each change:

### 2013-12-19 main.tex
wip
Add paper to Appendix C
WIP
Add quotes on performance of SVMs
WIP: chapter 4, description of experiment methods
Added the TODO command
Prefix refs with non-breaking space and capitalize refs
Add caption titles to tables
Restructure Chapter 5 graphs of data sets
WIP: Introduction
Put stills in grid. Crop images
WIP: Chapter 4
FrontmatteR

@@
 \usepackage{paralist}
+\usepackage{tabulary}
+\usepackage{cleveref}
+\usepackage{subcaption}
 \hypersetup{urlcolor=blue, colorlinks=true} % Colors hyperlinks in blue -
@@
 
-\DeclareMathOperator*{\argmin}{\arg\!\min}
-\DeclareMathOperator*{\argmax}{\arg\!\max}
+\DeclareMathOperator*{\argminold}{\arg\!\min}
+\DeclareMathOperator*{\argmaxold}{\arg\!\max}
 
@@
 \makeindex
+
+\usepackage{templates/math_shorthands}
 
@@
 \newcommand{\HRule}{\rule{\linewidth}{0.5mm}} % New command to make the lines in the title page
+
+\newcommand{\TODO}[1]{\emph{\textbf{*** TODO: {#1} ***}}}
 
@@
 \textsc{\LARGE \univname}\\[1.5cm] % University name
-\textsc{\Large Doctoral Thesis}\\[0.5cm] % Thesis type
+\textsc{\Large Master's Degree Thesis}\\[0.5cm] % Thesis type
 
@@
 \emph{Author:}\\
-\href{http://www.johnsmith.com}{\authornames} % Author name - remove the \href bracket to remove the link
+\href{http://rvlasveld.github.io}{\authornames}, \href{http://www.uu.nl}\textsc{UU} % Author name - remove the \href bracket to remove the link
 \end{flushleft}
 \end{minipage}
-\begin{minipage}{0.4\textwidth}
+\begin{minipage}{0.5\textwidth}
 \begin{flushright} \large
-\emph{Supervisor:} \\
-\href{http://www.jamessmith.com}{\supname} % Supervisor name - remove the \href bracket to remove the link
+\emph{Supervisors:} \\
+\href{http://www.cs.uu.nl/staff/gv.html}{Dr. G.A.W. \textsc{Vreeswijk}, \href{http://www.uu.nl}\textsc{UU}}  \\ % Supervisor name - remove the \href bracket to remove the link
+\href{http://www.annevanrossum.com}{Anne \textsc{van Rossum}, \href{http://www.almende.com}\textsc{Almende B.V.}} % Supervisor name - remove the \href bracket to remove the link
 \end{flushright}
@@
 \textit{in the}\\[0.4cm]
-\groupname\\\deptname\\[2cm] % Research group name and department name
+\facname\\\deptname\\[2cm] % Research group name and department name
 
@@
 
-\textit{``Thanks to my solid academic training, today I can write hundreds of words on virtually any topic without possessing a shred of information, which is how I got a good job in journalism."}
+\textit{``All truly great thoughts are conceived by walking."}
 
 \begin{flushright}
-Dave Barry
+Friedrich Nietzsche
 \end{flushright}
@@
 }
+
+\TODO{Write abstract}
 
@@
 
+\glsresetall
+
 \lhead{\emph{List of Figures}} % Set the left side page header to "List of Figures"
@@
 
+\glsresetall
+
 \lhead{\emph{List of Tables}} % Set the left side page header to "List of Tables"
 \listoftables % Write out the List of Tables
+
+\glsresetall
 
@@
 
-\clearpage % Start a new page
+% \clearpage % Start a new page
 
-\lhead{\emph{Symbols}} % Set the left side page header to "Symbols"
+% \lhead{\emph{Symbols}} % Set the left side page header to "Symbols"
 
-\listofnomenclature{lll} % Include a list of Symbols (a three column table)
-{
-$a$ & distance & m \\
-$P$ & power & W (Js$^{-1}$) \\
-% Symbol & Name & Unit \\
+% \listofnomenclature{lll} % Include a list of Symbols (a three column table)
+% {
+% $a$ & distance & m \\
+% $P$ & power & W (Js$^{-1}$) \\
+% % Symbol & Name & Unit \\
 
-& & \\ % Gap to separate the Roman symbols from the Greek
+% & & \\ % Gap to separate the Roman symbols from the Greek
 
-$\omega$ & angular frequency & rads$^{-1}$ \\
-% Symbol & Name & Unit \\
-}
+% $\omega$ & angular frequency & rads$^{-1}$ \\
+% % Symbol & Name & Unit \\
+% }
 
@@
 
-\clearpage
-\setstretch{1.3} % Return the line spacing back to 1.3
+% \clearpage
+% \setstretch{1.3} % Return the line spacing back to 1.3
 
-\pagestyle{empty} % Page style needs to be empty for this page
+% \pagestyle{empty} % Page style needs to be empty for this page
 
-\dedicatory{For/Dedicated to/To my\ldots} % Dedication text
+% \dedicatory{For/Dedicated to/To my\ldots} % Dedication text
 
-\addtocontents{toc}{\vspace{2em}} % Add a gap in the Contents, for aesthetics
+% \addtocontents{toc}{\vspace{2em}} % Add a gap in the Contents, for aesthetics
 
@@
 
-\input{./Appendices/AppendixA-Summaries}
-\input{./Appendices/AppendixB-Camci}
-%\input{./Appendices/AppendixB}
-%\input{./Appendices/AppendixC}
+% \input{./Appendices/AppendixA-Summaries}
+% \input{./Appendices/AppendixB-Camci}
+% \input{./Appendices/AppendixC}
+% \input{./Appendices/AppendixD-Planning}
+% \input{./Appendices/AppendixE-Data-structure}
+% \input{./Appendices/AppendixF-final}
 
@@
 
-\bibliographystyle{unsrtnat} % Use the "unsrtnat" BibTeX style for formatting the Bibliography
+% \bibliographystyle{unsrtnat} % Use the "unsrtnat" BibTeX style for formatting the Bibliography
+\bibliographystyle{plainnat} % Use the "unsrtnat" BibTeX style for formatting the Bibliography
 

### 2013-12-20 Chapters/Chapter5.tex
WIP: Chapter 5

@@
 \lhead{Chapter 5. \emph{Artificial data results}} % This is for the header on each page - perhaps a shortened title
-\begin{itemize}
-  \item Problemen met SVM methode: ook na de verandering blijft de threshold nog stijgen;
-  \item Veranderingen in mean geven ``extra'' change point wanneer alle data weer in het window zit. Zie \Cref{fig:camci_fixed_increasing_mean} op de 50-tallen.
-\end{itemize}
-
 In this chapter we will apply the proposed algorithm as stated in the previous chapter to artificial data sets.
@@
 \input{./Chapters/Chapter5/artificial_data_generation}
-
-\TODO{Add short section on quality metrics}
-
+\input{./Chapters/Chapter5/quality_metrics}
 \input{./Chapters/Chapter5/results}

### 2013-12-20 Chapters/Chapter5/quality_metrics.tex
WIP: Chapter 5

% !TEX root = ../../main.tex
\section{Quality metrics}\label{sec:artificial_data_quality_metrics}
For an objective measure of quality of the proposed method applied to the four aforementioned data sets, we employ two metrics, inspired by and slighty altered from Takeuchi and Yamanishi~\cite{takeuchi2006unifying}.
The first metric is the $\operatorname*{average \_ benefit}$, for which we first need to define the notion of $\operatorname*{benefit}$ for each change:
\begin{equation}\label{eq:benefit}
  \operatorname*{benefit}(y) = 1 - |t_y - t_y^*|,
\end{equation}
where $t_y$ is the time of the closest discovered change point and $t_y^*$ is the true change time.
We can then define the $\operatorname*{average \_ benefit}$ over all the discovered change points $Y$ as:
\begin{equation}\label{eq:average_benefit}
  \operatorname*{average \_ benefit}(Y) = \frac{\sum_{y=1}^Y \operatorname*{benefit}(y)}{Y}.
\end{equation}
Besides the benefit, which rewards discovered change points close to the actual change points, we need to penalize discovered change points which do not relate to a true change point.
For this we use the \acrlong{far}, expressed as
\begin{equation}\label{eq:false_alarm_rate}
  \operatorname*{far}(Y) = \frac{Y_{\operatorname*{false}}}{Y}.
\end{equation}

For a high quality method the $\operatorname*{average\_benefit}$ should be high and the $\operatorname*{far}(Y)$ should be low.
In our results we will also list the parameters used, since these need to be user-defined and are part of the (manual) optimization process.
The parameters are:
\begin{itemize}
  \item \textbf{Window length}: The window length that is processed by the \gls{svdd} algorithm. This is the number of data points that are continuously modeled in the shape in a hypersphere.
  \item \textbf{Sigma of \gls{rbf}}: The value of $\sigma$ for the \gls{rbf} kernel. The effect of this parameter is discussed in \Cref{subsec:svm_model_parameters}.
  \item \textbf{High threshold}: The $th_\text{high}$ threshold value for the \gls{rt} method described in \Cref{subsec:ratio_thresholding}. A higher value indicates a more precise and robust setting.
  \item \textbf{Low threshold}: The $th_\text{low}$ threshold value for the \gls{rt} method described in \Cref{subsec:ratio_thresholding}. A lower value indicates a more precise and robust setting.
  \item \textbf{Closeness}: In the post-processing phase the detected change points which are within a certain \emph{closeness} range will be merged together. This parameter indicates the minimal number of data points that change points must differ in order to be considered separate detections.
\end{itemize}
For all the sets we have used a value of $C = 0.1$, which indicated that $10\%$ of the data points are used as outliers.
This is a normal setting for the \gls{svdd} method.

### 2013-12-20 Chapters/Chapter5/results.tex
WIP: Chapter 5

--- 2013-12-12 Chapters/Chapter5/results.tex
@@
 \section{Results}\label{sec:artificial_data_results}
-For an objective measure of quality of the proposed method applied to the four aforementioned data sets, we employ two metrics, as defined by Takeuchi \etal \cite{takeuchi2006unifying}.
-The first metric is the $\operatorname*{average \_ benefit}$, for which we first need to define the notion of $\operatorname*{benefit}$ for each change:
-\begin{equation}\label{eq:benefit}
-  \operatorname*{benefit}(y) = 1 - |t_y - t_y^*| / 10  \mbox{~~~(when } |t_y - t_y^*| < 10 \mbox{)},
-\end{equation}
-where $t_y$ is the time of the discovered change point and $t_y^*$ is the true change time.
-We can then define the $\operatorname*{average \_ benefit}$ over all the discovered change points $Y$ as:
-\begin{equation}\label{eq:average_benefit}
-  \operatorname*{average \_ benefit}(Y) = \frac{\sum_{y=1}^Y \operatorname*{benefit}(y)}{Y}.
-\end{equation}
-Besides the benefit, which rewards discovered change points close to the actual change points, we need to penalize discovered change points which do not relate to a true change point.
-For this we use the \acrlong{far}, expressed as
-\begin{equation}\label{eq:false_alarm_rate}
-  \operatorname*{far}(Y) = \frac{Y_{\operatorname*{false}}}{Y}.
-\end{equation}
-
-For a high quality method the $\operatorname*{average\_benefit}$ should be high and the $\operatorname*{far}$ should be low.
-
-\TODO{Above metrics are not used, remove it and replace with own metric (offset, accuracy)}
-
 The results of all the data sets are provided in \Cref{tab:results_artificial}, together with the used parameters.
@@
 The data sets were processed with different window lengths and $\sigma$ value for the \gls{rbf} kernel.
-
-To give a better impression of the spread of the accuracy of the detected change points, a box plot for each data set is displayed in \Cref{fig:boxplot_artificial_sets}.
-With each data set embodying different characteristics, different observations can be drawn from the data sets.
-In the references figures, the blue lines represent the radius of the constructed hypersphere.
-The vertical red lines are the discovered change points.
-For each data set the following is observed:
-\begin{enumerate}
-  \item \textbf{Fixed increasing mean:} In \Cref{fig:camci_fixed_increasing_mean_thresholds} we see that each change point is almost equally fast and accurate detected.
-  This implies that the relative difference between the means of the Gaussian distribution has no effect on the detectability of the algorithm.
-  \item \textbf{Reduced increasing mean:} \Cref{fig:takeuchi_reduced_increasing_mean_thresholds} shows that the absolute difference between the means of the Gaussian distribution is a certrain influence to the detectability.
-  Although all change points are correctly detected (at the beginning of the series are two outliers), the change in radius becomes less significant at every change point.
-  \item \textbf{Reduced increasing mean, increasing variance:} In \Cref{fig:camci_takeuchi_alternating_variance_thresholds} we see the same effect, in regard to the mean, as with the previous data set.
-  Furthermore, the increasing variance makes change detection harder.
-  This is visible around the $9000$\textsuperscript{th} data point and after, where the radius keeps increasing.
-  \item \textbf{Alternating variance:} The ability to detect decreases in variance of the Gaussian distribution is displayed in \Cref{fig:camci_reduced_increasing_mean_increasing_variance_thresholds}.
-  Although some change points are discovered with a delay almost up to the length of the window processing the data, the decreases are all detected.
-  The box plot of this set, number 4 in \Cref{fig:boxplot_artificial_sets}, shows a wider spread for the results.
-\end{enumerate}
 
@@
     \hline
+    Window length & 50 & 100 & 50 & 50 \\
+    \hline
+    Sigma of \gls{rbf} & 13 & 13 & 15 & 13 \\
+    \hline
     High threshold & 1.6 & 1.6 & 1.5 & 2.2 \\
@@
     \hline
-    Closeness merging & 10 & 10 & 10 & 50 \\
+    Closeness & 10 & 10 & 10 & 50 \\
     \hline
-    Number of change points detected & 10 & 11 & 11 & 11 \\
     \hline
-    Mean offset & 2.27 & 6.18 & 0.64 & 15.36 \\
+    $\operatorname*{far}(Y)$ & 0 & 0.1 & 0.1 & 0.1 \\
     \hline
-    Standard deviation offset & 1.19 & 7.94 & 0.92 & 17.70 \\
+    $\operatorname*{Average\_benefit}$ & 2.27 & 6.18 & 0.64 & 15.36 \\
     \hline
-    Window length & 50 & 100 & 50 & 50 \\
-    \hline
-    Sigma of \gls{rbf} & 13 & 13 & 15 & 13 \\
+    Standard deviation benefit & 1.19 & 7.94 & 0.92 & 17.70 \\
     \hline
   \end{tabulary}
-  \caption[Results artificial data sets]{Results of the artificial data sets.}
+  \caption[Results artificial data sets]{Parameter settings and results of the artificial data sets.}
   \label{tab:results_artificial}
@@
 
-\begin{figure}
-\centering
-  \includegraphics[width=1\textwidth]{./Figures/chapter5/set_1_results.eps}
-  \caption[Fixed increasing mean, thresholds]{Hypersphere radius sizes for the \emph{Fixed increasing mean} data set with detected change points as red vertical bars.}
-  \label{fig:camci_fixed_increasing_mean_thresholds}
-\end{figure}
+To give a better impression of the spread of the accuracy of the detected change points, a box plot for each data set is displayed in \Cref{fig:boxplot_artificial_sets} and are the results are visualized in \Cref{fig:plots_results_artificial_data_sets}.
+With each data set embodying different characteristics, different observations can be drawn from the data sets.
+In the references figures, the blue lines represent the radius of the constructed hypersphere.
+The vertical red lines are the discovered change points.
+For each data set the following is observed:
+\begin{enumerate}
+  \item \textbf{Fixed increasing mean:} In \Cref{fig:camci_fixed_increasing_mean_thresholds} we see that each change point is almost equally fast and accurate detected.
+  This implies that the relative difference between the means of the Gaussian distribution has no effect on the detectability of the algorithm.
+  \item \textbf{Reduced increasing mean:} \Cref{fig:takeuchi_reduced_increasing_mean_thresholds} shows that the absolute difference between the means of the Gaussian distribution is a certain influence to the detectability.
+  Although all change points are correctly detected (at the beginning of the series are two outliers), the change in radius becomes less significant at every change point.
+  \item \textbf{Reduced increasing mean, increasing variance:} In \Cref{fig:camci_takeuchi_alternating_variance_thresholds} we see the same effect, in regard to the mean, as with the previous data set.
+  Furthermore, the increasing variance makes change detection harder.
+  This is visible around the $9000$\textsuperscript{th} data point and after, where the radius keeps increasing.
+  \item \textbf{Alternating variance:} The ability to detect decreases in variance of the Gaussian distribution is displayed in \Cref{fig:camci_reduced_increasing_mean_increasing_variance_thresholds}.
+  Although some change points are discovered with a delay almost up to the length of the window processing the data, the decreases are all detected.
+  The box plot of this set, number 4 in \Cref{fig:boxplot_artificial_sets}, shows a wider spread for the results.
+\end{enumerate}
 
@@
   \centering
+  \begin{subfigure}[b]{1\textwidth}
+    \includegraphics[width=1\textwidth]{./Figures/chapter5/set_1_results.eps}
+    \caption[Fixed increasing mean, thresholds]{\emph{Fixed increasing mean} data set.}
+      \label{fig:camci_fixed_increasing_mean_thresholds}
+  \end{subfigure} \\
+
+  \begin{subfigure}[b]{1\textwidth}
    \includegraphics[width=1\textwidth]{./Figures/chapter5/set_2_results.eps}
-  \caption[Reduced increasing mean, thresholds]{Hypersphere radius sizes for the \emph{Reduced increasing mean} data set with detected change points as red vertical bars.}
+    \caption[Reduced increasing mean, thresholds]{\emph{Reduced increasing mean} data set.}
     \label{fig:takeuchi_reduced_increasing_mean_thresholds}
+  \end{subfigure} \\
+
+  \begin{subfigure}[b]{1\textwidth}
+    \includegraphics[width=1\textwidth]{./Figures/chapter5/set_3_results.eps}
+      \caption[Reduced increasing mean, increasing variance, thresholds]{\emph{Reduced increasing mean, increasing variance} data set.}
+      \label{fig:camci_reduced_increasing_mean_increasing_variance_thresholds}
+  \end{subfigure} \\
+
+  \begin{subfigure}[b]{1\textwidth}
+   \includegraphics[width=1\textwidth]{./Figures/chapter5/set_4_results.eps}
+    \caption[Alternating variance, thresholds]{\emph{Alternating variance} data set.}
+    \label{fig:camci_takeuchi_alternating_variance_thresholds}
+  \end{subfigure} \\
+
+  \caption[Artificial data set results]{Hypersphere radius sizes for the four artificial data sets. The detected change points are displayed as red vertical bars.}
+  \label{fig:plots_results_artificial_data_sets}
 \end{figure}
-
-\begin{figure}
-\centering
-  \includegraphics[width=1\textwidth]{./Figures/chapter5/set_3_results.eps}
-  \caption[Reduced increasing mean, increasing variance, thresholds]{Hypersphere radius sizes for the \emph{Reduced increasing mean, increasing variance} data set with detected change points as red vertical bars.}
-  \label{fig:camci_reduced_increasing_mean_increasing_variance_thresholds}
-\end{figure}
-
-\begin{figure}
-\centering
-  \includegraphics[width=1\textwidth]{./Figures/chapter5/set_4_results.eps}
-  \caption[Alternating variance, thresholds]{Hypersphere radius sizes for the \emph{Alternating variance} data set with detected change points as red vertical bars.}
-  \label{fig:camci_takeuchi_alternating_variance_thresholds}
-\end{figure}

### 2014-01-06 Chapters/Chapter5/quality_metrics.tex
Changed Benefit to Delay. WIP chapter 6

@@
 For an objective measure of quality of the proposed method applied to the four aforementioned data sets, we employ two metrics, inspired by and slighty altered from Takeuchi and Yamanishi~\cite{takeuchi2006unifying}.
-The first metric is the $\operatorname*{average \_ benefit}$, for which we first need to define the notion of $\operatorname*{benefit}$ for each change:
-\begin{equation}\label{eq:benefit}
-  \operatorname*{benefit}(y) = 1 - |t_y - t_y^*|,
+The first metric is the $\operatorname*{average \_ delay}$, for which we first need to define the notion of $\operatorname*{delay}$ for each change:
+\begin{equation}\label{eq:delay}
+  \operatorname*{delay}(y) = |t_y - t_y^*|,
 \end{equation}
 where $t_y$ is the time of the closest discovered change point and $t_y^*$ is the true change time.
-We can then define the $\operatorname*{average \_ benefit}$ over all the discovered change points $Y$ as:
-\begin{equation}\label{eq:average_benefit}
-  \operatorname*{average \_ benefit}(Y) = \frac{\sum_{y=1}^Y \operatorname*{benefit}(y)}{Y}.
+We can then define the $\operatorname*{average \_ delay}$ over all the discovered change points $Y$ as:
+\begin{equation}\label{eq:average_delay}
+  \operatorname*{average \_ delay}(Y) = \frac{\sum_{y=1}^Y \operatorname*{delay}(y)}{Y}.
 \end{equation}
-Besides the benefit, which rewards discovered change points close to the actual change points, we need to penalize discovered change points which do not relate to a true change point.
+Besides the delay, which should be low to indicate discovered change points near the actual change point, we need to penalize discovered change points which do not relate to a true change point.
 For this we use the \acrlong{far}, expressed as
@@
 
-For a high quality method the $\operatorname*{average\_benefit}$ should be high and the $\operatorname*{far}(Y)$ should be low.
+For a high quality method the $\operatorname*{average\_delay}$ and the $\operatorname*{far}(Y)$ should both be low.
 In our results we will also list the parameters used, since these need to be user-defined and are part of the (manual) optimization process.

### 2014-01-06 Chapters/Chapter5/results.tex
Changed Benefit to Delay. WIP chapter 6

@@
     \hline
-    $\operatorname*{Average\_benefit}$ & 2.27 & 6.18 & 0.64 & 15.36 \\
+    $\operatorname*{Average\_delay}$ & 2.27 & 6.18 & 0.64 & 15.36 \\
     \hline
-    Standard deviation benefit & 1.19 & 7.94 & 0.92 & 17.70 \\
+    Standard deviation Delay & 1.19 & 7.94 & 0.92 & 17.70 \\
     \hline
@@
   \includegraphics[width=1\textwidth]{./Figures/chapter5/boxplot_results_artificial_sets.eps}
-  \caption[Box plot results artificial data sets]{Box plot of the results for the artificial data sets, indicating the number of data points difference between the actual and detected change points.}
+  \caption[Box plot results artificial data sets]{Box plot of the results for the artificial data sets, indicating the number of data points between the actual and closest detected change points.
+A lower and more compact box plot is better.}
   \label{fig:boxplot_artificial_sets}
@@
 To give a better impression of the spread of the accuracy of the detected change points, a box plot for each data set is displayed in \Cref{fig:boxplot_artificial_sets} and are the results are visualized in \Cref{fig:plots_results_artificial_data_sets}.
+This last figure shows the number of data points between each actual and closest detected change point.
+Using the box plot notation, we can visualize not only the average but also the spread of the delay.
 With each data set embodying different characteristics, different observations can be drawn from the data sets.

### 2014-01-06 Chapters/Chapter6.tex
WIP: Finalizing chapter 3
Changed Benefit to Delay. WIP chapter 6

@@
 \section{Outline}
-\emph{Not intented for the reader.}
-\begin{itemize}
-  \item Apply proposed method to real-world applications, such as
-    \begin{itemize}
-      \item Daily life activity recognition (as the original context of this thesis is)
-      \item PowerHouse sensor data
-      \item Stock data?
-    \end{itemize}
-  \item Relate back to filter vs. wrapper methods - give results with different methods?
-\end{itemize}
 
@@
 As discussed in \Cref{Chapter4}, the algorithm is subject to user parameters.
-In \Cref{sec:real_world_parameters} the effect and ranges of the parameters on the real world data sets are discussed.
-Finally, the performance of the algorithm is discussed in \Cref{sec:real_world_performance}.
+% In \Cref{sec:real_world_parameters} the effect and ranges of the parameters on the real world data sets are discussed.
+Finally, the results of the algorithm is discussed in \Cref{sec:real_world_performance}.
+
+\TODO{Check for final contents and ordering.}
 
@@
 \input{./Chapters/Chapter6/data_gathering}
-\input{./Chapters/Chapter6/parameters}
-\input{./Chapters/Chapter6/performance}
+% \input{./Chapters/Chapter6/parameters}
+
+% \TODO{Add short section on quality metrics}
+
+\input{./Chapters/Chapter6/results}

### 2014-01-06 Chapters/Chapter6/data_gathering.tex
Changed Benefit to Delay. WIP chapter 6

@@
 \section{Data Gathering}\label{sec:data_gathering}
+\TODO{Determine subjects:}
+\begin{enumerate}
+  \item Roemer?
+  \item Jos?
+  \item Marc
+\end{enumerate}
 For our data sets of real world activities we have sampled both in- and outdoor activities.
-The subjects performing the activities are three males, in an age around 25.
+The subjects performing the activities are three males, of an age around 25.
 We have used two series of recordings.
-The first series are various outdoor recordings with segments of walking, running and standing.
-Some movement activities are performed in a straight line, whilst others are performed around corners of in a wide circular direction.
-The second series consist of mainly walking up and down the stairs, with small vertical parts between the segments.
+The first series consists of various outdoor recordings with segments of walking, running and standing.
+Some activities are performed in a straight line, whilst others are performed around corners or in a wide circular direction.
+The second series consists mainly of walking up and down the stairs, with small vertical parts between the segments.
 The stairs are all in a circular shape.
@@
 The activities consist of standing still, walking and running (both jogging and sprinting).
-The plots of \Cref{fig:data_gathering_run_1_rot,fig:data_gathering_run_1_acc,fig:data_gathering_run_1_mag} shows the recordings of Subject $1$.
-For Subject $2$ the plots are visible in \Cref{fig:data_gathering_run_2_rot,fig:data_gathering_run_2_acc,fig:data_gathering_run_2_mag}.
-The recorded metrics are from the accelerometer, rotational and magnetometer sensors.
+The plots of \Cref{fig:plots_subject_1} shows the recordings of Subject $1$.
+For Subject $2$ the plots are visible in \Cref{fig:plots_subject_2}.
+The recorded metrics originate from the accelerometer, rotational and magnetometer sensors.
 The solid black lines indicate manually annoted change points, the dashes lines indicate the closest discovered change point.
@@
 
-\subsection{Outdoor straight lines results}
-\TODO{For each data set: discovered timepoint with acc, mag, rot, and all combined. Use closest change point}
-
-\begin{center}\begin{table}
-  \begin{tabulary}{\textwidth}{|l|c|c|}
-    \hline
-    \multicolumn{3}{|c|}{Subj 1, run 1} \\
-    \hline \hline
-    Act. & Ann. & Disc. \\
-    \hline
-    Still & 0 & *0* \\
-    \hline
-    Shake & 3.5 & *3.5* \\
-    \hline
-    Still & 4.2 & *4.2* \\
-    \hline
-    In pocket & 5.5 & *5.5* \\
-    \hline
-    Walk & 7 & *7* \\
-    \hline
-    Run & 10 & *10* \\
-    \hline
-    Walk & 14 & *14* \\
-    \hline
-    Run & 17.2 & *17.2* \\
-    \hline
-    Walk & 22.1 & *22.1* \\
-    \hline
-    Turn CCW & 23.5 & *23.5* \\
-    \hline
-    Walk & 24.2 & *24.2* \\
-    \hline
-    Run (sprint) & 28 & *28* \\
-    \hline
-    Run & 31 & *31* \\
-    \hline
-    Walk & 33 & *33* \\
-    \hline
-    Run & 35.5 & *35.5* \\
-    \hline
-    Walk & 39 & *39* \\
-    \hline
-    Out pocket & 41 & *41* \\
-    \hline
-    Still & 43 & *43* \\
-    \hline
-    \hline
-    \multicolumn{2}{|l|}{Ratio Total CPs} & 3.2 \\
-    \hline
-    \multicolumn{2}{|l|}{Sum of differences} & 12.341 \\
-    \hline
-  \end{tabulary}
-  \caption[Performed activities subject 1 run 1]{The outdoor series performed activities by subject 1, run 1.}
-  \label{tab:outdoor_series}
-\end{table}\end{center}
-
-\TODO{Replace all times in the stars with the closest discovered change point}
+% \begin{center}\begin{table}
+%   \begin{tabulary}{\textwidth}{|l|c|c|}
+%     \hline
+%     \multicolumn{3}{|c|}{Subj 1, run 1} \\
+%     \hline \hline
+%     Act. & Ann. & Disc. \\
+%     \hline
+%     Still & 0 & *0* \\
+%     \hline
+%     Shake & 3.5 & *3.5* \\
+%     \hline
+%     Still & 4.2 & *4.2* \\
+%     \hline
+%     In pocket & 5.5 & *5.5* \\
+%     \hline
+%     Walk & 7 & *7* \\
+%     \hline
+%     Run & 10 & *10* \\
+%     \hline
+%     Walk & 14 & *14* \\
+%     \hline
+%     Run & 17.2 & *17.2* \\
+%     \hline
+%     Walk & 22.1 & *22.1* \\
+%     \hline
+%     Turn CCW & 23.5 & *23.5* \\
+%     \hline
+%     Walk & 24.2 & *24.2* \\
+%     \hline
+%     Run (sprint) & 28 & *28* \\
+%     \hline
+%     Run & 31 & *31* \\
+%     \hline
+%     Walk & 33 & *33* \\
+%     \hline
+%     Run & 35.5 & *35.5* \\
+%     \hline
+%     Walk & 39 & *39* \\
+%     \hline
+%     Out pocket & 41 & *41* \\
+%     \hline
+%     Still & 43 & *43* \\
+%     \hline
+%     \hline
+%     \multicolumn{2}{|l|}{Ratio Total CPs} & 3.2 \\
+%     \hline
+%     \multicolumn{2}{|l|}{Sum of differences} & 12.341 \\
+%     \hline
+%   \end{tabulary}
+%   \caption[Performed activities subject 1 run 1]{The outdoor series performed activities by subject 1, run 1.}
+%   \label{tab:outdoor_series}
+% \end{table}\end{center}
 
@@
   \centering
-  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/stairs-1-marc/data_plot_rot_annotated.eps}
-  \caption[R3: rotation]{Run 3: Indoor stairs Marc, rotation}
+  \begin{subfigure}[b]{1\textwidth}
+    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/stairs-1-marc/data_plot_rot_annotated.eps}
+    \caption{Run 3: Indoor stairs Marc, rotation}
     \label{fig:data_gathering_run_3_rot}
+  \end{subfigure} \\
+
+  \begin{subfigure}[b]{1\textwidth}
+    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/stairs-1-marc/data_plot_mag_annotated.eps}
+    \caption{Run 3: Indoor stairs Marc, Mag}
+    \label{fig:data_gathering_run_3_mag}
+  \end{subfigure} \\
+
+  \begin{subfigure}[b]{1\textwidth}
+    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/stairs-1-marc/data_plot_acc_annotated.eps}
+    \caption{Run 3: Indoor stairs Marc, accelerometer}
+    \label{fig:data_gathering_run_3_acc}
+  \end{subfigure}
+  \caption[Plots subject 3]{Annotated plots of a full recording performed by Subject~$3$ while walking the stairs.
+The vertical lines are manually determined change points.
+The label above each segment indicates the current performed activity.}\label{fig:plots_subject_3}
 \end{figure}
-
-\begin{figure}
-\centering
-  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/stairs-1-marc/data_plot_mag_annotated.eps}
-  \caption[R3: mag]{Run 3: Indoor stairs Marc, Mag}
-  \label{fig:data_gathering_run_3_mag}
-\end{figure}
-
-\begin{figure}
-\centering
-  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/stairs-1-marc/data_plot_acc_annotated.eps}
-  \caption[R3: accelerometer]{Run 3: Indoor stairs Marc, accelerometer}
-  \label{fig:data_gathering_run_3_acc}
-\end{figure}

### 2014-01-06 Chapters/Chapter6/data_sets.tex
Changed Benefit to Delay. WIP chapter 6

@@
 In this chapter we apply the method to real-world data sets.
-The data sets are inertial sensor data recorded during human activities, such as walking, running, ascending and descending stairs, and standing still.
+The data sets come from inertial sensor data recorded during human activities, such as walking, running, ascending and descending stairs, and standing still.
 To compare our results with other research, we have tried to apply our method to two commonly used data sets.
@@
 Whilst this data set is convenient for activity classification, it does not allow for detection of change points.
+Even if the gap between the activities would be removed, the transition segments between activities would still be missing and thus the data set would not reflect continuous recordings.
 The other commonly used data set is the \gls{uci-har} from the Machine Learning Repository and originates from \cite{anguita2012human}.
@@
 Because of the above described shortcomings of commonly used data sets, we have decided to record activities by ourselves.
-The following characterizations were set for our data sets:
+The following characterization requirements were set for our data sets:
 \begin{enumerate}
-  \item The activities need to be performed and recorded in a continuous matter,
+  \item The activities need to be performed and recorded in a continuous manner,
   \item The environment should be natural and uncontrolled, both in- and outdoor,
-  \item There should be a clear annotation possible, as objective as possible,
-  \item The hardware and software should be widely available.
+  \item There data should be clearly annotated, as objective as possible,
+  \item The required hardware (and software) should be widely available.
 \end{enumerate}
 
-With these requirements in mind, we created the following setup.
-All activities were recorded by one smartphone in the right front pants pocket of the subject.
+With these requirements in mind, we have created the following setup.
+All activities were recorded by a smartphone in the right front pants pocket of the subject.
 The phones that were used are the \textsc{HTC Sensation XE} and \textsc{HTC Desire}, running the \textsc{Anroid} smartphone operating system.
@@
 This application records the sensor data to \textsc{CSV} files, which are then transformed to \textsc{MATLAB} compatible files.
-The constructed algorithm processes the data in an online matter (using only current and historic data).
-During the activities, the subject is recorded with a video recorder to enable manual labeling and determining the change points.
+The constructed algorithm processes the data in an online manner (using only current and historic data).
+During the activities, the subject is recorded with a video camera to enable manual labeling and determination of the change points.
 
-Using the video recordings, the data sets are manually annotated, as shown in \Cref{fig:data_gathering_run_1_rot}.
+Using the video recordings, the data sets are manually annotated, as shown in the plots of \Cref{fig:plots_subject_1,fig:plots_subject_3}.
 The graph displays from a single recording the rotational-axis values (see \Cref{tab:recorded_metrics} for explanation of the metrics), with manually annotated activities above the data.
-The manually determines change points are used to draw quality conclusion over the algorithms performance, by comparing it to the discovered change points.
+The manually determined change points are used to draw quality conclusion over the algorithm's performance, by comparing it to the discovered change points.
 All used data recordings, including the video material and manual labeling, are made available \cite{vlasveld2013continuous}.

### 2014-01-06 Chapters/Chapter6/results.tex
Changed Benefit to Delay. WIP chapter 6

% !TEX root = ../../main.tex
\section{Results}\label{sec:real_world_results}

In the first paragraph we will explain that we use the same quality measures as used in \Cref{sec:artificial_data_quality_metrics} for objective tabular results.
We further argue that more important is the subjective, visual, inspection of the discovered change points regarding the sensor data.
To give both results, we will first give a table and box plot, like in \Cref{sec:artificial_data_results}.
After that, we will show characterizing parts of the plots with annotated and discovered change points.

\subsection{Objective measure}

\TODO{Table of search run (in the columns) and for each run the ratio of False Alarm Rate, Average benefit (closest CP) and STD benefit. Like chapter 5.}

\begin{table}
  \centering
  \begin{tabulary}{\textwidth}{|l|c|c|c|c|c|c|c|c|}
    \cline{2-9}
    \multicolumn{1}{l|}{} & Run 1 & Run 2 & Run 3 & Run 4 & Run 5 & Run 6 & Run 7 & Run 8 \\
    \hline
    Window length & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
    \hline
    Sigma of \gls{rbf} & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
    \hline
    High threshold & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
    \hline
    Low threshold & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
    \hline
    Closeness & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
    \hline
    \hline
    $\operatorname*{far}(Y)$ & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
    \hline
    $\operatorname*{Average\_delay}$ & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
    \hline
    STD Delay & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
    \hline
  \end{tabulary}
  \caption[Results real world runs]{Parameter settings and results of the real-world data sets.}
  \label{tab:results_real_world}
\end{table}

\TODO{Box plot of the runs, like in Chapter 5}

Here comes a list of observations and conclusions on the performance:
\begin{itemize}
  \item \textbf{Perhaps something in bold:} here we list a funny observation.
  \eg that when walking the circulair stairs it is harder to find the rotation on the flat surface.
  \item Or just something else.
\end{itemize}

\subsection{Subjective measures}
In this subsection we will provide a few plots with characterizing parts of the sensor data, annotated change points, and discovered change points.
We use it to illustrate some aspects of the method, on which it performs well and where not.
We do not provide all plots, because that would take up to much time.
We do try to give a subjective conclusion about the performance, backed by the provided examples.

### 2014-01-06 Chapters/Chapter6.tex
WIP chapter 6

--- 2013-11-01 Chapters/Chapter6.tex
@@
 
-%----------------------------------------------------------------------------------------
-\section{Outline}
-\emph{Not intented for the reader.}
-\begin{itemize}
-  \item Apply proposed method to real-world applications, such as
-    \begin{itemize}
-      \item Daily life activity recognition (as the original context of this thesis is)
-      \item PowerHouse sensor data
-      \item Stock data?
-    \end{itemize}
-  \item Relate back to filter vs. wrapper methods - give results with different methods?
-\end{itemize}
+In this chapter we discuss the real-world data sets used for this research and the performance of our proposed algorithm on these data sets.
+In \Cref{sec:data_sets} we discuss common available data sets and argue that these are inappropriate for our research.
+It is followed by an overview of the setup we have created to record our own continuous data sets in an uncontrolled environment, both for in- and outdoor activities.
+The series of activities performed are discussed in \Cref{sec:data_gathering}.
+In that section we show typical plots of the inertial sensor data and stills from the video recordings.
+That will illustrate the setting in which the activities are performed and the characteristics of continuous data.
+Finally, the results of the algorithm are discussed in \Cref{sec:real_world_results}.
+We will use the objective measures, introduced in \Cref{sec:artificial_data_quality_metrics}, and also try to give subjective claims on the performance of the algorithm applied to the real-world data.
 
+\input{./Chapters/Chapter6/data_sets}
 \input{./Chapters/Chapter6/data_gathering}
+\input{./Chapters/Chapter6/results}

### 2014-01-06 Chapters/Chapter6/data_gathering.tex
WIP chapter 6

@@
 The activities consist of standing still, walking and running (both jogging and sprinting).
-The plots of \Cref{fig:plots_subject_1} shows the recordings of Subject $1$.
-For Subject $2$ the plots are visible in \Cref{fig:plots_subject_2}.
+The plots of \Cref{fig:plots_subject_2} shows the recordings of Subject $2$.
 The recorded metrics originate from the accelerometer, rotational and magnetometer sensors.
@@
 \TODO{Add plots}
-For Subject~$2$ the still of the video recordings are in the upper row of \Cref{fig:stills_subject_1_and_2}, for Subject~$1$ in the lower row.
+For Subject~$1$ the still of the video recordings are in the upper row of \Cref{fig:stills_subject_1_and_2}, for Subject~$2$ in the lower row.
 
@@
 The vertical lines are manually determined change points.
-The label above each segment indicates the current performed activity.}\label{fig:plots_subject_1}
+The label above each segment indicates the current performed activity.}\label{fig:plots_subject_2}
 \end{figure}

### 2014-01-07 Chapters/Chapter6/results.tex
Add remarks on real-world data results

@@
 We do try to give a subjective conclusion about the performance, backed by the provided examples.
+
+\subsection{Notes: remarks}
+Mergen van change points:\\
+
+Aanpak 1: Na ieder gevonden change point een bepaalde periode $s$ alle anderen negeren. Probleem: je hebt dan bijna altijd na $s$ sowieso een `gevonden' change point.\\
+Aanpak 2: Alle change points die binnen $s$ van elkaar zitten mergen.
+Probleem: naar voren of naar achteren doen?
+Als je een heel lang blok hebt met CPs dicht bij elkaar, met mogelijk `echte' CPs, dan worden die ook genegeerd.
+Mogelijke oplossing: gebruik van `zekerheid' over de gevonden CPs om te mergen.
+
+Het ghosting treedt vooral op aan het begin van iedere run, met het `kaliberen': het begin van de echte activiteit zit er te dicht op en wordt niet goed herkend.
+
+Run 1 - Roemer\\
+
+\begin{itemize}
+  \item Doordat de thresholds niet adaptief zijn, wordt de gevoeligheid voor de hele methode globaal gezet.
+  Dit zorgt dat ruwe overgangen gevonden worden, maar subtielere worden overgeslagen.
+  Bijvoorbeeld: op 14s zit een overgang van rennen naar lopen, maar die is niet zo groot.
+Deze wordt niet gevonden, tenzij de gevoeligheid veel hoger gaat, maar dan krijg je weer teveel false positives.
+De overgang vlak erna, op 17s, gaat weer van lopen naar rennen en die wordt wel gevonden.
+  \item Op 33s van rennen naar lopen wordt juist weer wel goed gevonden.
+  \item Doordat in de data soms wel veranderingen zitten (die niet direct duidelijk zijn uit de video beelden) en change points dichtbij gemerged worden, kan er soms een groot verschil ontstaan.
+Bv op 8s en 9s.
+De changes op 7s en 10s worden gemist.
+  \item De overgang van rennen naar sprint(?) op 31s wordt gemist.
+\end{itemize}
+
+Run 2 - Jos \\
+\begin{itemize}
+  \item Op 37s is een annotate van rennen naar lopen.
+De methode detecteert het bijna een seconde eerder.
+Als je naar de ruwe data kijkt zie je daar ook een verandering, dus nog voordat je het uit de video-beelden zou opmerken --> heel subjectief dus.
+  \item Zelfde `ghosting' probleem als hierboven: bij overgang van Still naar Pocket wordt niets gevonden, doordat in de periode ervoor veel changes zijn.
+De data in `Still' is heel homogeen, dus kleine veranderingen worden als changes gezien, en die zijn er alsnog veel.
+--> \textbf{Hoe heterogener de data, hoe robuster}.
+\end{itemize}
+
+Run 3 - Roemer, around the corner \\
+\begin{itemize}
+  \item Met acc, mag, rot data is 90 graden CCW niet goed te vinden. Met mag,rot data wel --> te veel data is ook niet goed.
+  \item Het enige punt, de draai, wordt met mag,rot goed gevonden. (geen plot of data, want verder niet interessant). Ook Walk-in-hand, pocket en Out-pocket worden erg nauwkeurig gevonden!
+\end{itemize}
+
+Run 4 - Roemer, fountain \\
+\begin{itemize}
+  \item Na 24s: van Rennen naar Lopen, worden meerdere change points gevonden. In de data zie je ook dat iedere stap een rustigere stap is, dus de gevoeligheid pakt dat op.
+  \item De draaien in het Rennen rond 12s en 16s worden wel gevonden (bij hoge gevoeligheid, met delay), maar die rond 21s niet. (acc, mag, rot). Met alleen Rot ook niet.
+  \item Correct number of change points found, but still 11 total difference. Would be better if benchmarking periods (still, shake) would not be present. (overall geldt dat ook).
+  \item
+\end{itemize}
+
+Run 5 - Jos, fountain \\
+\begin{itemize}
+  \item Bij Acc,Mag,Rot worden de Turn CW punten niet goed gevonden. Bij alleen Rot wel (niet allemaal, maar wel beter).
+  \item Om bij Acc,Mag,Rot de Turn te vinden is hoge gevoeligheid nodig, maar dat geeft hoge FAR.
+  \item
+\end{itemize}
+
+Run 6 - Roemer, fountain \\
+\begin{itemize}
+  \item Om de overgang van rennen naar lopen op 18s te vinden is relatief hoge Low-threshold nodig (dus: hoge gevoeligheid).
+  \item Bij hoge Low-Threshold wordt draai op 24s ook gevonden. Anders niet.
+  \item Stilstaan op 38s geeft hele veel ruis in `thresholding' property. --> overgang naar lopen 2 sec later wordt niet goed gevonden. Komt doordat hypersphere heel klein wordt, en dus veel outliers?
+  \item
+\end{itemize}
+
+Run 7 - Jos, fountain \\
+Camera was leeg, dus mislukte run
+
+Run 8 - Marc, indoor \\
+\begin{itemize}
+  \item Op 24s, van Walk naar Downstairs: door het ghosting effect is het moeilijk te pakken.
+Relatief hoge merging-tijd nodig, maar dat verwijderd deze, omdat het begin van lopen dicht zit op het begin van downstairs.
+  \item Op 37s zit CCW Turn: is te herkennen (met acc, mag, rot) maar hoge gevoeligheid nodig.
+  \item In tweede downstairs block (24s) en upstairs blocken (42s, 54s) worden halverwege ook changes gevonden (met acc, mag, rot). Ook met alleen Acc: zit ook echt verschil in de data
+  \item Met alleen Mag data heel veel changes: cirkel-vorm van trap zorgt voor constante verandering.
+  \item Overall: hoge FAR, door cirkelvorm trap (?)
+  \item Verschil tussen down/upstairs en lopen is niet zo groot. Typisch: 33/34s: herkenning van overgang duurt even een stap.
+  \item Vooral op einde, vanaf 66s (lopen naar stoel), heel veel False Positives. Veel verschillende activiteiten kort op elkaar.
+\end{itemize}

### 2014-01-08 Chapters/Chapter6/data_sets.tex
WIP chapter 6
Add boxplot and remarks real-world results

@@
 All activities were recorded by a smartphone in the right front pants pocket of the subject.
-The phones that were used are the \textsc{HTC Sensation XE} and \textsc{HTC Desire}, running the \textsc{Anroid} smartphone operating system.
+The phones that were used are the \textsc{HTC Sensation XE} and \textsc{HTC Desire}, running the \textsc{Android} smartphone operating system.
 During the activities the intertial sensor data is recorded using a free application, \textsc{Sensor Logger}~\cite{sensorlogger}.
@@
 
-Using the video recordings, the data sets are manually annotated, as shown in the plots of \Cref{fig:plots_subject_1,fig:plots_subject_3}.
+\TODO{Mentiod and ref dd\_tools, libsvm}
+
+dd\_tools: \cite{Ddtools2013} and PR tools: \cite{van2005classification}.
+
+Using the video recordings, the data sets are manually annotated, as shown in the plots of \Cref{fig:plots_subject_2,fig:plots_subject_3}.
 The graph displays from a single recording the rotational-axis values (see \Cref{tab:recorded_metrics} for explanation of the metrics), with manually annotated activities above the data.

### 2014-01-08 Chapters/Chapter6/results.tex
Add boxplot and remarks real-world results

@@
 \section{Results}\label{sec:real_world_results}
+
+\TODO{Add notion of \textbf{empirical} results.}
 
@@
 
-\TODO{Box plot of the runs, like in Chapter 5}
+Here something to refer to \Cref{fig:boxplot_real_world_runs}, about what it shows.
+
+\begin{figure}
+\centering
+  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/boxplot_results_real_world_runs.eps}
+  \caption[Box plot results real-world runs]{Box plot of the results for the real-world runs, indicating the number of data points between the actual and closest detected change points.
+A lower and more compact box plot is better.}
+  \label{fig:boxplot_real_world_runs}
+\end{figure}
+
+We will state observations and remarks from the results which present and applicable to all the runs.
+After that, we will discuss special observations for each run individually.
+
+\TODO{Uitleggen over sensitivity: we willen die zo laag mogelijk vanwege stabiliteit.}
+
+About the overall performance and results, we have the following observations:
+\begin{itemize}
+  \item \textbf{Merging:} in our method we use a \emph{closeness} time period $t_c$ (usually up to $0.5$ - $1.0$ seconds) to merge discovered change points that have a very small distance between them.
+  Multiple merging-strategies can be applied, and in our method we use the naive implementation by simply ignoring all the discovered change points that occur less than $t_c$ seconds after the previous change point.
+  This works well when there is a block of noisy data with a high amount of (falsely) discovered change points.
+  On the flip side, this method will ignore real new change points which occur during the noisy period.
+  This problem on itself can be formulated as a change detection problem.
+  \item \textbf{Masking:} Since the parameters for the detection method are set globally, it is difficult to discover all the change points (and only the change points) without a high \gls{far}.
+  If a change point is proceeded by a noisy block, then the real change point will be merged with the previous change points.
+  In an other case, where a change point is very clear represented in the data but the next, close by, other change point is more subtle, the latter change point is \emph{masked} by the former.
+  This masking effect is also discussed in \cite{inclan1994use}, where an iterated approach is applied.
+  \item \textbf{False heterogeneity:} For our video-data synchronization, we started and ended each recording with a few seconds in which the recording smartphone was kept still in the air.
+  During this period, the data variance, and thus the constructed hypersphere, becomes very small.
+  As a results, even small movements are considered to be changes and in the case of the mid-air still smartphone a lot of false change points are detected.
+  Due to the merging effect described above, the final real change point (often shaking the smartphone) is not discovered.
+  \item \textbf{Incorrect weighting:} In our analyses we have used the data from the accelerometer, magnetic field, and rotation sensors.
+  In the segments which embodied movement in a circular manner (such as walking around a corner on the street, as in run 3 \TODO{add ref to image}), the turn was not discovered.
+  Alternatively, when we only used the magnetic field and rotation sensors, the turn was correctly discovered.
+  Other transitions, such as from walking to running, are harder to discover without the (linear) accelerometer sensor data.
+\end{itemize}
+
+\TODO{Add positive remarks}
+
+When looking at the individual runs, we have the following observations for each run number:
+\begin{enumerate}
+  \item \textbf{Subject 2, straight}
+    \begin{itemize}
+      \item The first transition from running to walking, around $14s$ is harder to discover than the transition for the same activities around $33s$.
+      This shows is that in real-world applications there is a diversity between the same transitions and activities.
+      \item Around $8s$, $9s$, and $16s$ a few steps (while walking) are regarded as change points.
+      During the running segments from $17s$ and $28s$ there is a lower probability of change for each step.
+    \end{itemize}
+  \item \textbf{Subject 1, straight}
+    \begin{itemize}
+      \item Following the video recordings, we have annotated a change point from running to walking around $37s$.
+      Our method discovers a change point almost a second before.
+      In retrospect, we can see that the data distribution indeed changes from the discovered change point on.
+      This shows us two important principles.
+      The first is that the annotations are very subjective.
+      The second is that between different activities the transition period is longer than we would think.
+      Looking at the data, we can see that the body slows down, even before we visually notice it on the video recordings.
+    \end{itemize}
+  \item \textbf{Subject 2, corner}
+    \begin{itemize}
+      \item The $90^{\circ}$ counter-clockwise turn during the walking activity is hard to discover when the accelerometer sensor data is included.
+      When only the magnetic field and rotation sensors are used, the turn requires a lower sensitivity.
+      With only these two sensors all the other change points in this run are also successfully discovered.
+    \end{itemize}
+  \item \textbf{Subject 2, fountain}
+    \begin{itemize}
+      \item Like in the first run, the walking segment from $24s$ results in a change point for each step.
+      Further inspection of the data reveals that each step is indeed different from the other.
+      Due to the global parameter settings, the sensitivity is too high for this segment to recognize it as one.
+      \item During the circular run, from $12s$ till $24$, there are two change points discovered.
+      The difference for the rotational vectors need to accumulate to a certain value before they have enough influence to let the rotation be regarded as a change point.
+      \TODO{Beter formuleren}.
+    \end{itemize}
+  \item \textbf{Subject 1, fountain}
+    \begin{itemize}
+      \item As with the other runs, the accelerometer data makes it harder to detect turns.
+      It requires a higher sensitivity, which results in a higher \gls{far}.
+    \end{itemize}
+  \item \textbf{Subject 2, fountain 2}
+    \begin{itemize}
+      \item During the standing segment around $38s$ there are a lot of false positives.
+      It seems to be the \emph{false heterogeneity} problem described above.
+    \end{itemize}
+  \setcounter{enumi}{7}
+  \item \textbf{Subject 3, indoor stairs}
+    \begin{itemize}
+      \item The walking segment around $22s$, between two segments of walking downstairs, shows little difference in the data.
+      To recognize it as a change point a high sensitivity and low closeness time period $t_c$ is required.
+      \item During some segments (downstairs from $24s$, upstairs from $42s$ and $54s$) the method discovers more change points than our annotation.
+      A closer inspecting of the raw data reveals indeed changes in behavior.
+      To exclude these (semi) false positives, a better tuning of parameters is required.
+      \item Because of the circular shape of the stairs, the magnetic field sensors constantly differs.
+      Although our method is build to exclude slow shifting changes (because we are only interested in sudden changes), with our used window width it still eventually results in change points.
+      \item The difference between taking the stairs and walking is smaller than, \eg, walking and running.
+      The delay between these segments seems to be larger, as illustrated around $33s$.
+    \end{itemize}
+\end{enumerate}
+
+\TODO{Add suggestion for adaptive parameter tuning, locally determined}
 

### 2014-01-09 Chapters/Chapter4/data_gathering.tex
Add boxplot and remarks real-world results
Finalizing.... still....and again...

@@
 
-For the experiments we used a HTC Sensation XE smartphone as recording device.
-The activities were recorded using a free Android application \cite{sensorlogger}.
+For the experiments we used a smartphone with inertial sensors as recording device.
+The activities were recorded using a free \textsc{Android} application \cite{sensorlogger}.
 This application was chosen for its convenient data format of the sensor recording and its regularity of the sampling interval.
@@
 For our expiments we used the data for the accelerometer, magnatief field and rotation.
+
+We implemented our algorithm in \textsc{MATLAB}.
+For the \gls{svdd} model construction phase we used the \textsc{dd\_tools} library by Tax \cite{Ddtools2013}, which is dependend on the \textsc{PR\_tools} package, originating from \cite{van2005classification}.
+In an earlier attempt we used the a \textsc{LibSVM} \cite{chang2011libsvm} add-on for the \gls{svdd} algorithm \cite{changrevisit}, but we had troubles making that implementation work with our data.
 

### 2014-01-09 Chapters/Chapter6.tex
Finalizing.... still....and again...

@@
 Finally, the results of the algorithm are discussed in \Cref{sec:real_world_results}.
-We will use the objective measures, introduced in \Cref{sec:artificial_data_quality_metrics}, and also try to give subjective claims on the performance of the algorithm applied to the real-world data.
+We will use the objective measures, introduced in \Cref{sec:artificial_data_quality_metrics}, and also give subjective claims on the performance of the algorithm applied to the real-world data.
+Since the data sets used are real-world recordings, we will put emphasis on the latter type of quality measures \ie the subjective claims.
 

### 2014-01-09 Chapters/Chapter6/data_gathering.tex
Finalizing.... still....and again...

--- 2014-01-06 Chapters/Chapter6/data_gathering.tex
@@
 \section{Data Gathering}\label{sec:data_gathering}
-\TODO{Determine subjects:}
-\begin{enumerate}
-  \item Roemer?
-  \item Jos?
-  \item Marc
-\end{enumerate}
-For our data sets of real world activities we have sampled both in- and outdoor activities.
+For our data sets of real-world activities we have sampled both in- and outdoor activities.
 The subjects performing the activities are three males, of an age around 25.
-We have used two series of recordings.
-The first series consists of various outdoor recordings with segments of walking, running and standing.
-Some activities are performed in a straight line, whilst others are performed around corners or in a wide circular direction.
-The second series consists mainly of walking up and down the stairs, with small vertical parts between the segments.
-The stairs are all in a circular shape.
+We have used four types of recordings.
+The first type (Run 1 and Run 2) consists of various outdoor recordings with segments of walking, running and standing.
+For the second type (Run 4, 5, and 6) some of the same activities are performed in a straight line, whilst others are performed around corners or in a wide circular direction.
+The third type (Run 8) consists mainly of walking up and down circular shaped stairs, with small horizontal parts between the segments.
+The last type is for Run 3, in which we walked in a straight line, took a $90^{\circ}$ counter-clockwise turn (around the corner) and continued walking.
+A more detailed description of the performed activities follows.
 
 \subsection{Outdoor straight lines}\label{subsec:outdoor_straight}
-The first series of outdoor performed activities consist of walking and running segments in a straight line.
+The first series of outdoor performed activities consist of standing still, walking, and running segments in a straight line.
 Halfway the recording is a $180^{\circ}$ turn.
-The activities consist of standing still, walking and running (both jogging and sprinting).
 The plots of \Cref{fig:plots_subject_1} shows the recordings of Subject $1$.
-For Subject $2$ the plots are visible in \Cref{fig:plots_subject_2}.
 The recorded metrics originate from the accelerometer, rotational and magnetometer sensors.
-The solid black lines indicate manually annoted change points, the dashes lines indicate the closest discovered change point.
+The solid black lines indicate manually annotated change points.
 A still from the video recording, showing Subject $2$ performing a (sprinting) run segment, is in \Cref{fig:stills_subject_2_sprinting}.
-
-% \begin{center}\begin{table}
-%   \begin{tabulary}{\textwidth}{|l|c|c|}
-%     \hline
-%     \multicolumn{3}{|c|}{Subj 1, run 1} \\
-%     \hline \hline
-%     Act. & Ann. & Disc. \\
-%     \hline
-%     Still & 0 & *0* \\
-%     \hline
-%     Shake & 3.5 & *3.5* \\
-%     \hline
-%     Still & 4.2 & *4.2* \\
-%     \hline
-%     In pocket & 5.5 & *5.5* \\
-%     \hline
-%     Walk & 7 & *7* \\
-%     \hline
-%     Run & 10 & *10* \\
-%     \hline
-%     Walk & 14 & *14* \\
-%     \hline
-%     Run & 17.2 & *17.2* \\
-%     \hline
-%     Walk & 22.1 & *22.1* \\
-%     \hline
-%     Turn CCW & 23.5 & *23.5* \\
-%     \hline
-%     Walk & 24.2 & *24.2* \\
-%     \hline
-%     Run (sprint) & 28 & *28* \\
-%     \hline
-%     Run & 31 & *31* \\
-%     \hline
-%     Walk & 33 & *33* \\
-%     \hline
-%     Run & 35.5 & *35.5* \\
-%     \hline
-%     Walk & 39 & *39* \\
-%     \hline
-%     Out pocket & 41 & *41* \\
-%     \hline
-%     Still & 43 & *43* \\
-%     \hline
-%     \hline
-%     \multicolumn{2}{|l|}{Ratio Total CPs} & 3.2 \\
-%     \hline
-%     \multicolumn{2}{|l|}{Sum of differences} & 12.341 \\
-%     \hline
-%   \end{tabulary}
-%   \caption[Performed activities subject 1 run 1]{The outdoor series performed activities by subject 1, run 1.}
-%   \label{tab:outdoor_series}
-% \end{table}\end{center}
 
@@
 \eg the running activities are performed around a fountain with an estimated diameter of 10 meters.
-The plots of the recordings of Subject, for the accelerometer, rotation, and magnetometer are shown in...
-\TODO{Add plots}
+The plots of the recordings of Subject 2, for the accelerometer, rotation, and magnetometer are shown in \Cref{fig:plots_subject_2}.
 For Subject~$2$ the still of the video recordings are in the upper row of \Cref{fig:stills_subject_1_and_2}, for Subject~$1$ in the lower row.
@@
   \end{subfigure} \\
-  \caption[Plots]{Annotated plots of a full recording performed by Subject~$2$.
+  \caption[Plots run 1]{Annotated plots of a full recording performed by Subject~$1$.
 The vertical lines are manually determined change points.
 The label above each segment indicates the current performed activity.}\label{fig:plots_subject_1}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \begin{subfigure}[b]{1\textwidth}
+    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/run-5-run-fountain-jos/data_plot_acc_annotated.eps}
+    \caption{Accelerometer}
+    \label{fig:recording_subject_2_run_5_accelerometer}
+  \end{subfigure}
+
+  \begin{subfigure}[b]{1\textwidth}
+    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/run-5-run-fountain-jos/data_plot_rot_annotated.eps}
+    \caption{Rotation}
+    \label{fig:recording_subject_2_run_5_rotation}
+  \end{subfigure} \\
+
+  \begin{subfigure}[b]{1\textwidth}
+    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/run-5-run-fountain-jos/data_plot_mag_annotated.eps}
+    \caption{Magnetic Field}
+    \label{fig:recording_subject_2_run_5_magnetic}
+  \end{subfigure} \\
+  \caption[Plots run 5]{Annotated plots of a full recording performed by Subject~$2$.
+The vertical lines are manually determined change points.
+The label above each segment indicates the current performed activity.}\label{fig:plots_subject_2}
 \end{figure}
@@
 The final series of activities recorded consist of walking up and down the stairs, indoor.
-The stairs are shaped in an semi-circular form, with a small vertical segment on each floor.
+The stairs are shaped in an semi-circular form, with a small horizontal segment on each floor.
 The still of \Cref{fig:stills_subject_3_downstairs} illustrates the circular shape of the stairs.

### 2014-01-09 Chapters/Chapter6/data_sets.tex
Finalizing.... still....and again...

--- 2014-01-06 Chapters/Chapter6/data_sets.tex
@@
 Because of the above described shortcomings of commonly used data sets, we have decided to record activities by ourselves.
-The following characterization requirements were set for our data sets:
+The following characterizing requirements were set for our data sets:
 \begin{enumerate}
@@
 With these requirements in mind, we have created the following setup.
-All activities were recorded by a smartphone in the right front pants pocket of the subject.
-The phones that were used are the \textsc{HTC Sensation XE} and \textsc{HTC Desire}, running the \textsc{Anroid} smartphone operating system.
+All activities were recorded by a smartphone worn on body in the right front pants pocket of the subject.
+The used phones are the \textsc{HTC Sensation XE} and \textsc{HTC Desire}, running the \textsc{Android} smartphone operating system.
 During the activities the intertial sensor data is recorded using a free application, \textsc{Sensor Logger}~\cite{sensorlogger}.
@@
 The constructed algorithm processes the data in an online manner (using only current and historic data).
+The \textsc{MATLAB} implementation uses the \gls{svdd} library \textsc{dd\_tools} by Tax \cite{Ddtools2013}.
 During the activities, the subject is recorded with a video camera to enable manual labeling and determination of the change points.
+As illustrated in the top row of \Cref{fig:stills_subject_1_and_2}, the manual labeling is an ambiguous task.
+In \Cref{fig:stills_subject_2_change_point} a transition between walking and running is shown.
+It is impossible to determine the exact location of the change point and thus objective quality measures (such as the delay) are of less importance.
 
-Using the video recordings, the data sets are manually annotated, as shown in the plots of \Cref{fig:plots_subject_1,fig:plots_subject_3}.
-The graph displays from a single recording the rotational-axis values (see \Cref{tab:recorded_metrics} for explanation of the metrics), with manually annotated activities above the data.
+Using the video recordings, the data sets are manually annotated, as shown in the plots of \Cref{fig:plots_subject_1,fig:plots_subject_2,fig:plots_subject_3}.
+The graphs display each from a single recording the inertial sensor values (see \Cref{tab:recorded_metrics} for explanation of the metrics), with manually annotated activities above the data.
 The manually determined change points are used to draw quality conclusion over the algorithm's performance, by comparing it to the discovered change points.
-All used data recordings, including the video material and manual labeling, are made available \cite{vlasveld2013continuous}.
-\TODO{Create web page with data sets}
+All of the used data sets, including the video material and manual labeling, are made publicly available \cite{vlasveld2013continuous} for further research.
+\TODO{Publish web page with data sets}
 The following section discusses the performed activities during the recordings.

### 2014-01-09 Chapters/Chapter6/results.tex
Fill table real-world results
Finalizing.... still....and again...

@@
 \section{Results}\label{sec:real_world_results}
+In this section we will provide the results of our methods applied to the real-world data sets and discuss the quality of the solution.
+As stated above, we will use the objective quality metrics as used in \Cref{sec:artificial_data_quality_metrics}.
+That will be in the form of a tabular overview of the (manually set) best meta-parameters and results, expressed in \gls{far} and $\operatorname*{Average\_delay}$ (following \Cref{sec:artificial_data_results}).
+Furthermore, we will discuss the results from an empirical point of view, since the problem of finding change points in human activities is ambiguous.
+By stating our observations and remarks and providing fragments of visual results, we will discuss the characterizing strengths and weaknesses of our method.
 
-\TODO{Add notion of \textbf{empirical} results.}
-
-In the first paragraph we will explain that we use the same quality measures as used in \Cref{sec:artificial_data_quality_metrics} for objective tabular results.
-We further argue that more important is the subjective, visual, inspection of the discovered change points regarding the sensor data.
-To give both results, we will first give a table and box plot, like in \Cref{sec:artificial_data_results}.
-After that, we will show characterizing parts of the plots with annotated and discovered change points.
-
-\subsection{Objective measure}
-
-\TODO{Table of search run (in the columns) and for each run the ratio of False Alarm Rate, Average benefit (closest CP) and STD benefit. Like chapter 5.}
+\subsection{Objective metrics}
+The employed two-stage algorithm requires the setting of user defined parameters.
+In \Cref{tab:results_real_world} for each run the results are listed, together with the used parameters.
+We have used a window length of 50 frames, which represents around $1.16$ seconds of data (the frame-rate of the recordings is not fixed, and we apply a slight filtering).
+The other parameter values are the result of an educated guess and manual optimization.
+In general, a high value for the upper threshold and low value for the lower threshold indicates a robust method, since it indicates that change points are significantly different from homogeneous segments.
+The closeness time frame should be low, since that indicates there are few false positives.
 
@@
   \centering
-  \begin{tabulary}{\textwidth}{|l|c|c|c|c|c|c|c|c|}
-    \cline{2-9}
-    \multicolumn{1}{l|}{} & Run 1 & Run 2 & Run 3 & Run 4 & Run 5 & Run 6 & Run 7 & Run 8 \\
+  \begin{tabulary}{\textwidth}{|l|c|c|c|c|c|c|c|}
+    \cline{2-8}
+    \multicolumn{1}{l|}{} & Run 1 & Run 2 & Run 3 & Run 4 & Run 5 & Run 6 & Run 7 \\
     \hline
-    Window length & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
+    Window length & 50 & 50 & 50 & 50 & 50 & 50 & 50 \\
     \hline
-    Sigma of \gls{rbf} & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
+    Sigma of \gls{rbf} & 13 & 13 & 13 & 13 & 13 & 13 & 4 \\
     \hline
-    High threshold & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
+    High threshold & 1.2 & 1.5 & 1.3 & 1.1 & 1.3 & 1.2 & 1.7 \\
     \hline
-    Low threshold & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
+    Low threshold & 0.8 & 0.8 & 0.6 & 0.7 & 0.7 & 0.8 & 0.6 \\
     \hline
-    Closeness & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
+    Closeness (s) & 0.7 & 0.85 & 1 & 0.6 & 0.7 & 1 & 0.8 \\
     \hline
     \hline
-    $\operatorname*{far}(Y)$ & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
+    $\operatorname*{far}(Y)$ & 0.1 & 0.05 & 0 & 0 & 0.2 & 0 & 0.48 \\
     \hline
-    $\operatorname*{Average\_delay}$ & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
+    $\operatorname*{Average\_delay}$ & 0.52 & 0.83 & 0.47 & 0.92 & 1.26 & 1.05 & 0.92 \\
     \hline
-    STD Delay & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
+    STD Delay & 0.29 & 0.55 & 0.52 & 0.75 & 1.67 & 0.78 & 0.92 \\
     \hline
@@
 
-Here something to refer to \Cref{fig:boxplot_real_world_runs}, about what it shows.
+To give a better impression on the quality of the $\operatorname*{Average\_delay}$, \Cref{fig:boxplot_real_world_runs} shows a box plot for each run.
+It shows the spread of the $\operatorname*{Average\_delay}$ in seconds over all the change points.
+A lower and more compact box plot is better.
 
@@
 
-We will state observations and remarks from the results which present and applicable to all the runs.
-After that, we will discuss special observations for each run individually.
+\subsection{Subjective observations and remarks}
+Since the manual annotation of change points is ambiguous, the objective quality metrics can give a distorted view of the performance.
+To overcome this, we will state observations and remarks found after close (visual) inspection of the annotated and discovered change points.
+We will start with general observations applicable to all the runs and will continue with individual remarks for each run.
+Where possible, a graphical illustration will support our observations.
 
-\TODO{Uitleggen over sensitivity: we willen die zo laag mogelijk vanwege stabiliteit.}
-
-About the overall performance and results, we have the following observations:
+Considering the overall performance and results, we have the following observations:
 \begin{itemize}
@@
   \item \textbf{Incorrect weighting:} In our analyses we have used the data from the accelerometer, magnetic field, and rotation sensors.
-  In the segments which embodied movement in a circular manner (such as walking around a corner on the street, as in run 3 \TODO{add ref to image}), the turn was not discovered.
+  In the segments which embodied movement in a circular manner (such as walking and running around a fountain), the turn was not (or at a different time point) discovered.
   Alternatively, when we only used the magnetic field and rotation sensors, the turn was correctly discovered.
@@
 \begin{enumerate}
-  \item \textbf{Subject 2, straight}
+  \item \textbf{Subject 2, straight walk and run}
     \begin{itemize}
@@
     \end{itemize}
-  \item \textbf{Subject 1, straight}
+  \item \textbf{Subject 1, straight walk and run}
     \begin{itemize}
@@
     \end{itemize}
-  \item \textbf{Subject 2, corner}
+  \item \textbf{Subject 2, walk around corner}
     \begin{itemize}
@@
     \end{itemize}
-  \item \textbf{Subject 2, fountain}
+  \item \textbf{Subject 2, walk and run around fountain}
     \begin{itemize}
@@
     \end{itemize}
-  \item \textbf{Subject 1, fountain}
+  \item \textbf{Subject 1, walk and run around fountain}
     \begin{itemize}
@@
     \end{itemize}
-  \item \textbf{Subject 2, fountain 2}
+  \item \textbf{Subject 2, walk and run fountain 2}
     \begin{itemize}
@@
     \end{itemize}
-  \setcounter{enumi}{7}
   \item \textbf{Subject 3, indoor stairs}
@@
 
-\TODO{Add suggestion for adaptive parameter tuning, locally determined}
+\TODO{Samenvatten observations en remarks, brugje naar possible improvements maken}
 
-Here comes a list of observations and conclusions on the performance:
-\begin{itemize}
-  \item \textbf{Perhaps something in bold:} here we list a funny observation.
-  \eg that when walking the circulair stairs it is harder to find the rotation on the flat surface.
-  \item Or just something else.
-\end{itemize}
+\section{Possible improvements}\label{sec:possible_improvements}
+From the above section we can extract a few possible improvements for the algorithm.
+Most of the observations and remarks are due to a different required sensitivity for local segments of the run.
+Currently, our method employs a global setting for the thresholds.
+A possible improvement would be to use locally optimized parameters, although that would require an reflective data processing approach or more a priori about the data.
+With our approach, the lack of a priori knowledge about the data distribution is one of the design decisions.
 
-\subsection{Subjective measures}
-In this subsection we will provide a few plots with characterizing parts of the sensor data, annotated change points, and discovered change points.
-We use it to illustrate some aspects of the method, on which it performs well and where not.
-We do not provide all plots, because that would take up to much time.
-We do try to give a subjective conclusion about the performance, backed by the provided examples.
-
-\subsection{Notes: remarks}
-Mergen van change points:\\
-
-Aanpak 1: Na ieder gevonden change point een bepaalde periode $s$ alle anderen negeren. Probleem: je hebt dan bijna altijd na $s$ sowieso een `gevonden' change point.\\
-Aanpak 2: Alle change points die binnen $s$ van elkaar zitten mergen.
-Probleem: naar voren of naar achteren doen?
-Als je een heel lang blok hebt met CPs dicht bij elkaar, met mogelijk `echte' CPs, dan worden die ook genegeerd.
-Mogelijke oplossing: gebruik van `zekerheid' over de gevonden CPs om te mergen.
-
-Het ghosting treedt vooral op aan het begin van iedere run, met het `kaliberen': het begin van de echte activiteit zit er te dicht op en wordt niet goed herkend.
-
-Run 1 - Roemer\\
-
-\begin{itemize}
-  \item Doordat de thresholds niet adaptief zijn, wordt de gevoeligheid voor de hele methode globaal gezet.
-  Dit zorgt dat ruwe overgangen gevonden worden, maar subtielere worden overgeslagen.
-  Bijvoorbeeld: op 14s zit een overgang van rennen naar lopen, maar die is niet zo groot.
-Deze wordt niet gevonden, tenzij de gevoeligheid veel hoger gaat, maar dan krijg je weer teveel false positives.
-De overgang vlak erna, op 17s, gaat weer van lopen naar rennen en die wordt wel gevonden.
-  \item Op 33s van rennen naar lopen wordt juist weer wel goed gevonden.
-  \item Doordat in de data soms wel veranderingen zitten (die niet direct duidelijk zijn uit de video beelden) en change points dichtbij gemerged worden, kan er soms een groot verschil ontstaan.
-Bv op 8s en 9s.
-De changes op 7s en 10s worden gemist.
-  \item De overgang van rennen naar sprint(?) op 31s wordt gemist.
-\end{itemize}
-
-Run 2 - Jos \\
-\begin{itemize}
-  \item Op 37s is een annotate van rennen naar lopen.
-De methode detecteert het bijna een seconde eerder.
-Als je naar de ruwe data kijkt zie je daar ook een verandering, dus nog voordat je het uit de video-beelden zou opmerken --> heel subjectief dus.
-  \item Zelfde `ghosting' probleem als hierboven: bij overgang van Still naar Pocket wordt niets gevonden, doordat in de periode ervoor veel changes zijn.
-De data in `Still' is heel homogeen, dus kleine veranderingen worden als changes gezien, en die zijn er alsnog veel.
---> \textbf{Hoe heterogener de data, hoe robuster}.
-\end{itemize}
-
-Run 3 - Roemer, around the corner \\
-\begin{itemize}
-  \item Met acc, mag, rot data is 90 graden CCW niet goed te vinden. Met mag,rot data wel --> te veel data is ook niet goed.
-  \item Het enige punt, de draai, wordt met mag,rot goed gevonden. (geen plot of data, want verder niet interessant). Ook Walk-in-hand, pocket en Out-pocket worden erg nauwkeurig gevonden!
-\end{itemize}
-
-Run 4 - Roemer, fountain \\
-\begin{itemize}
-  \item Na 24s: van Rennen naar Lopen, worden meerdere change points gevonden. In de data zie je ook dat iedere stap een rustigere stap is, dus de gevoeligheid pakt dat op.
-  \item De draaien in het Rennen rond 12s en 16s worden wel gevonden (bij hoge gevoeligheid, met delay), maar die rond 21s niet. (acc, mag, rot). Met alleen Rot ook niet.
-  \item Correct number of change points found, but still 11 total difference. Would be better if benchmarking periods (still, shake) would not be present. (overall geldt dat ook).
-  \item
-\end{itemize}
-
-Run 5 - Jos, fountain \\
-\begin{itemize}
-  \item Bij Acc,Mag,Rot worden de Turn CW punten niet goed gevonden. Bij alleen Rot wel (niet allemaal, maar wel beter).
-  \item Om bij Acc,Mag,Rot de Turn te vinden is hoge gevoeligheid nodig, maar dat geeft hoge FAR.
-  \item
-\end{itemize}
-
-Run 6 - Roemer, fountain \\
-\begin{itemize}
-  \item Om de overgang van rennen naar lopen op 18s te vinden is relatief hoge Low-threshold nodig (dus: hoge gevoeligheid).
-  \item Bij hoge Low-Threshold wordt draai op 24s ook gevonden. Anders niet.
-  \item Stilstaan op 38s geeft hele veel ruis in `thresholding' property. --> overgang naar lopen 2 sec later wordt niet goed gevonden. Komt doordat hypersphere heel klein wordt, en dus veel outliers?
-  \item
-\end{itemize}
-
-Run 7 - Jos, fountain \\
-Camera was leeg, dus mislukte run
-
-Run 8 - Marc, indoor \\
-\begin{itemize}
-  \item Op 24s, van Walk naar Downstairs: door het ghosting effect is het moeilijk te pakken.
-Relatief hoge merging-tijd nodig, maar dat verwijderd deze, omdat het begin van lopen dicht zit op het begin van downstairs.
-  \item Op 37s zit CCW Turn: is te herkennen (met acc, mag, rot) maar hoge gevoeligheid nodig.
-  \item In tweede downstairs block (24s) en upstairs blocken (42s, 54s) worden halverwege ook changes gevonden (met acc, mag, rot). Ook met alleen Acc: zit ook echt verschil in de data
-  \item Met alleen Mag data heel veel changes: cirkel-vorm van trap zorgt voor constante verandering.
-  \item Overall: hoge FAR, door cirkelvorm trap (?)
-  \item Verschil tussen down/upstairs en lopen is niet zo groot. Typisch: 33/34s: herkenning van overgang duurt even een stap.
-  \item Vooral op einde, vanaf 66s (lopen naar stoel), heel veel False Positives. Veel verschillende activiteiten kort op elkaar.
-\end{itemize}
+Furthermore, we discovered that sometimes, especially with movements incorporating circular motions, the addition of (linear) accelerometer data makes the discovery of change points harder.
+In such cases, each of the inertial sensor data streams should be relatively weighted (\eg using the covariance) in order to signal an overall change when it only occurs in one of the streams.

### 2014-01-09 Chapters/Chapter7.tex
WIP: Introduction chapter
Finalizing.... still....and again...

@@
 To do so, we used the \gls{svdd} algorithm by Tax~\cite{tax1999support}, following the method of applying it to a sliding window of data, as done by Camci~\cite{camci2010change}.
-Instead of using both the information whether a new data point is an outlier and the radius of the hypersphere to indicate change, we only used on the latter model properties.
+Instead of using both the information whether a new data point is an outlier and the radius of the hypersphere to indicate change, we only used the latter model properties.
 The increase of the radius of the constructed hypersphere, which encloses the data objects and increases with heterogeneous segments of data, is extracted and used for indication of change.
@@
 \section{What is done}
-The scope of this research was the temporal segmentation of recordings, obtained from human activities.
+The scope of this research was the temporal segmentation of recordings obtained from human activities.
 This was performed in the wider context of activity classification.
@@
 In this research the primary goal was finding a temporal segmentation, which is assumed to be able to aid the classification step.
+This problem is reduced to finding change points in temporal data sets originating from inertial sensors.
 
@@
 
-In this research we have applied the method by Camci~\cite{camci2010change} to inertial signals, recorded by smartphones.
+In this research we have applied the method by Camci~\cite{camci2010change} to data from inertial sensors, recorded by smartphones.
 We have used the accelerometer data for measurement of speed, the gyroscope data for measurement of rotation, and the magnetometer to measure the direction and orientation of the performed activities.
-The constructed method reduces this 9-dimensional signal to a single property, obtained from the constructed model, which can be interpreted to indicate change.
+The proposed method reduces this 9-dimensional signal to a single property, obtained from the constructed model, which can be interpreted to indicate change.
 
 The constructed model follows the implementation of Tax and Laskov~\cite{tax2003online} for an incremental version of \gls{svdd}, \gls{isvdd}.
-This algorithm constructs a \gls{oc-svm} model by processing the data over a sliding window.
-The model represents a hypersphere from which the radius size is abstracted.
+This algorithm creates a \gls{oc-svm} model by processing the data over a sliding window.
+The model represents a hypersphere from which the radius size is extracted.
 That property is of interest since the assumption is that a heterogeneous window of data will show an increase of radius, in relation to a homogeneous segment of data.
@@
 
-To test our approach, we have used the same data as used in the experiments from Camci~\cite{camci2010change} and Takeuchi and Yamanishi~\cite{takeuchi2006unifying}.
+To test our approach, we have used the same data as used in the experiments from Camci~\cite{camci2010change} and Takeuchi and Yamanishi~\cite{takeuchi2006unifying}, for which the results are comparable with the original research (see \Cref{sec:artificial_data_results}).
 Furthermore, since this research is focused around human activities, we have recorded and manually annotated in- and outdoor activities.
 The discovered change points were compared to the annotated change points, obtained from video recordings.
+The results where discussed in \Cref{sec:real_world_results}.
 In this research we found that other public available and wide used common data sets, such as the WISDM~\cite{kwapisz2011activity} and UCI HAR~\cite{anguita2012human}, were not useful for our purpose, since the activities are non-continuously recorded.
@@
 It is applied in other contexts, such as motion tracking and first person vision sensing, the segmentation obtained from inertial signals is often the implicit result of classification.
+
+\section{Contributions}
 
@@
 
+\TODO{Improvements of our method, \Cref{sec:possible_improvements}}
+
 What has not been covered? What is worthwhile.

### 2014-01-10 Chapters/Chapter1.tex
WIP: chapter 4
Finalizing chapter 7, conclusion

@@
   \item \TODO{Overall: meer focus op de \emph{waarom} vraag/antwoord bij keuzes. Bouwt aan motivatie en context van het geheel.}
+  \item \TODO{Consistent `data point' of `data object' gebruiken.}
 \end{itemize}
@@
 The research by Yin \etal  \cite{yin2008sensor} creates a \gls{oc-svm} classifier to filter normal activity traces, using training data.
-However, to our knowledge, \gls{oc-svm} methods have not yet been applied to inertial sensor time series data directly to find changes in behavior and create a temporal segmentation.
+However, to the best of our knowledge, \gls{oc-svm} methods have not yet been applied to inertial sensor time series data directly to find changes in behavior and create a temporal segmentation.
 This algorithm transforms the time series data, which can be of any high dimension, to an higher dimensional feature space.
@@
   The data set, consisting of raw inertial sensor recordings, manual change point and activity labeling, and video recordings, are used for testing the method and are made publicly available \cite{vlasveld2013continuous}.
-  \item \textbf{Application of \gls{oc-svm} methods to inertial sensor data:} to our best knowledge, this thesis is the first application of \gls{oc-svm} methods, especially the \gls{svdd} method, to inertial sensor signals in the context of creating a temporal segmentation.
+  \item \textbf{Application of \gls{oc-svm} methods to inertial sensor data:} to the best of our knowledge, this thesis is the first application of \gls{oc-svm} methods, especially the \gls{svdd} method, to inertial sensor signals in the context of creating a temporal segmentation.
   Furthermore, this thesis contributes to the small amount of research that has a temporal segmentation of inertial sensor data as its primary objective.

### 2014-01-10 Chapters/Chapter7.tex
Finalizing chapter 7, conclusion

@@
 
-%----------------------------------------------------------------------------------------
-\section{Outline}
-\emph{Not intended for the reader.}
-\begin{itemize}
-  \item Tie together, integrate, synthesize all the issues from the chapters, reflecting on introductory thesis statements.
-  \item Provide answers to research questions
-  \item Highlight study limitations
-  \item Directions for further research.
-  \item Content: Summarize
-  \begin{itemize}
-    \item What have I researched
-    \item Main arguments
-    \item How researched
-    \item What discovered
-    \item What pre-existing views were challenged (?)
-  \end{itemize}
-  \item Overview of
-  \begin{itemize}
-    \item New knowledge
-    \item Significance of research
-    \item Limitations (data, concepts)
-    \item Speculations on implications of research
-    \item Further research and development, or: links with different fields, other methods applied to same data.
-  \end{itemize}
-  \item Must: make clear statement of my original contribution
-  \item Ideally:
-  \begin{itemize}
-    \item Shows links across key ideas spread in chapters
-    \item Show commitment and enthusiasm for research
-    \item Positive impression
-  \end{itemize}
-  \item Avoid:
-  \begin{itemize}
-    \item claim findings that have not been proven
-    \item introducing new data/concepts
-    \item Be self-critical, but dont put weaknesses or limitation in research
-    \item Too long/too short
-  \end{itemize}
-\end{itemize}
-
-Paragraph: into on conclusion; what have I researched and what is my contribution.
-Break into sections further down.
-
-===
-
 In this research we have applied \gls{occ} methods to accelerometer data, recorded during the performance of various human activities, in order to create a temporal segmentation of those recordings.
@@
 
-\TODO{Write overview of this chapter. \emph{In the following section we will...}}
+In the following section we will further discuss the proposed method for temporal segmentation of inertial sensor signals from human activities.
+Our main findings and contributions are discussed in \Cref{sec:main_findings}.
+The limitations of our method and research is subject of \Cref{sec:limitations}.
+These two sections are followed by our suggestions for future research.
+\Cref{sec:final_remarks} is the final section of this thesis.
 
@@
 We have used the accelerometer data for measurement of speed, the gyroscope data for measurement of rotation, and the magnetometer to measure the direction and orientation of the performed activities.
-The proposed method reduces this 9-dimensional signal to a single property, obtained from the constructed model, which can be interpreted to indicate change.
+The proposed method, as discussed in \Cref{Chapter4}, reduces this 9-dimensional signal to a single property, obtained from the constructed model, which can be interpreted to indicate change.
 
@@
 
-\TODO{In the above, refer to the corresponding chapters/sections.}
+\section{Main findings / Contributions}\label{sec:main_findings}
+\TODO{Vraag aan Anne: wat is een goede titel voor deze section? Of moeten we Findings en Contributions splitsen?}
+While temporal segmentation of time series data is not uncommon, the application to inertial signals is currently not under much research attention.
+For other contexts it has been applied, such as motion tracking \cite{barbivc2004segmenting,li2007segmentation}.
+In the context of human activities the segmentation obtained from inertial sensor signals is often an implicit result of classification.
 
-\section{Main findings}
-What we have found, show links across chapters.
+In this research we have shown that it is possible to explicitly obtain a temporal segmentation, expressed as change points, of activities performed by humans.
+This was performed in a real-world setting, showing the robustness of the method for noisy data.
+Since manual labeling (the annotation of change points on the time series data) of human activities is ambiguous we have put emphasis on the subjective inspection of the results.
+The proposed method gives fairly good results, although parameter tuning is a subject of furter research.
 
-For each finding a different paragraph.
-How did we arrive at this finding, and it challenges/improves on previous research.
-Not just a summary, but the effects and implications.
+Our method is based on and a slightly simplification of the \gls{svcpd} by Camci~\cite{camci2010change}.
+We have tested our proposed method with the same artificial data and did not find a deterioration in the results.
 
-While temporal segmentation of time series data is not uncommon, the application to inertial signals is currently not under much research attention.
-It is applied in other contexts, such as motion tracking and first person vision sensing, the segmentation obtained from inertial signals is often the implicit result of classification.
+This research is, as far as we know, the first application of \gls{oc-svm} based methods to on-body worn inertial sensor signals, such as accelerometer data.
+Furthermore, it applies the algorithms, amongst others of Tax and Duin~\cite{tax1999support}, to real-world data.
+We have shown that the assumption of a change the performed activity results, due to a change in the data distribution, in an increase of the hypersphere's radius.
+This increase, and a similar decrease in the case of homogeneous activity distribution in the current window, can be used for an indication of change.
 
-\section{Contributions}
+Finally, due to encountered problems with commonly used data sets \cite{kwapisz2011activity,anguita2012human}, we have recorded our real-world data sets by ourselves.
+We have recorded all the activities with a smartphone worn in the right front pocket of the subjects.
+Furthermore, we have recorded the performing subjects from the third person point of view with a video camera.
+This allows for better interpretation of the results.
+All the recorded data is made publicity available~\cite{vlasveld2013continuous}.
 
-\section{Other research}
-Relate the finding with earlier research.
-Provide similarities and differences.
-Better/worse results?
+\TODO{Toevoegen: samenvatting van resultaten uit \Cref{Chapter6}: de subjectieve waarnemingen?}
 
-\section{Limitations}
-Show what limitations were encountered or decided (data collection, scope).
-Note the focus of the limitations and how they worked together (to get positive note).
-This can create bridge to "future research".
+\section{Limitations}\label{sec:limitations}
+Our proposed method was constructed under the assumption that an explicit temporal segmentation of inertial sensor data can aid the process of activity classification.
+We have expressed the problem of temporal segmentation as finding change points in the time series data.
+Since this is the first application of \gls{oc-svm} based algorithms to real-world accelerometer data, we have limited ourselves to creating a first method.
+The (results of the) proposed method is subject to parameters, which we have set manually based on empirical results and experience.
 
-\section{Future research}
+We have applied our method to real-world data, recorded in 7 runs.
+For a better and more robust analysis of the algorithm it can be applied to more real-world data.
+Although we have recorded and publicized our used data, application to other commonly used data sets would enable a fair comparison.
+Furthermore, since the manual annotation of change points is an ambiguous task, the results are very subjective.
+
+\section{Future research}\label{sec:further_research}
 The motivation for this research was, amonst others, the assumption that an explicit temporal segmentation of inertial signals can be used to obtain a better classification of the performed activities.
@@
 A future research in this line of assumptions is to compare the performance of classification methods using this obtained temporal segmentation and those that do not.
+This would confirm of reject the assumption that a temporal segmentation enables a better classification.
+
+As discussed in the previous section, the parameters for our proposed method (such as the \gls{svdd} settings and \gls{rbf} width $\sigma$), are set manually and globally.
+For further research we advise to create an automated optimization method for the parameters.
+Furthermore, to overcome the \emph{masking} effect discussed in \Cref{sec:real_world_results}, a method which determines the parameters locally should obtain better results.
 
 Since the temporal segmentation of inertial signal is not yet a very active field of study, more studies can be performed to obtain it.
-In our approach we have used the raw data directly.
+In our approach we have used (all) the raw data directly.
 In the field of activity recognition many methods make use of pre-processing steps, from which one step often is feature extraction from the raw data.
 That, and other pre-processing steps, could be incorporated in the explicit temporal segmentation of inertial signals.
-Furture research could focus on extracted features that aid the detection of change points.
+Future research could focus on extracted features that aid the detection of change points.
 
-\TODO{Add future research on \gls{oc-svm} applications.}
+In that same line, we have encountered situations in which the exclusion of some sensors, \eg the accelerometer, yields better results.
+This occurs mainly in situations where a rotation is part of the change (\eg walking around a corner).
+Thus depending on the desired granularity of the segmentation, it can be beneficial to disregard some of the time series data.
+Determining (dynamically) which sensors should be, and which should not be, used in the model construction phase (known as \emph{feature selection}) could be a subject of further research.
 
-\TODO{Improvements of our method, \Cref{sec:possible_improvements}}
+Finally, the application of \gls{oc-svm} based methods to new types of data is an interesting subject for further research in itself.
+It creates a powerful non-linear classification solution and combined with optimization techniques can reduce problems of relatively high dimensionality to a single (or a few) dimensions.
 
-What has not been covered? What is worthwhile.
-Show that I am thinking ahead and know the field of study.
-
-\section{Final section}
-Reminding of original contribution and significance.
-This section is the overall conclusion and thus should be very concise and precise.
-
-\TODO{Formulate the research questions in the introduction of thesis, and in the chapters.}
+\section{Final words}\label{sec:final_remarks}
+In this research we have constructed a method to create a temporal segmentation of inertial sensor time series data, originating from performed human activities.
+To this end we have applied \gls{occ} and \gls{oc-svm} based method to our own recorded data sets.
+We followed the approach by Camci~\cite{camci2010change} and used the theory and implementation of \gls{svdd} by Tax~\cite{tax2001one,Ddtools2013}.
+Although the performance of the method heavily depends on user set parameters, the results are promising.
+We believe that the temporal segmentation of time series data can aid the process of activity classification, which is a logical follow-up on this research.
+\TODO{Hier nog meer samenvatten?}

### 2014-01-10 Chapters/Chapter1.tex
Finalizing Chapter 2, literatur review...

@@
 We are interested in both in- and outdoor activities in an uncontrolled environment and in a continuous manner, which can be regarded as real-world data.
-Although the number of performed activities will be limited, we assume the temporal segmentation must be more robust then in the case of data from a controlled environment.
+Although the number of performed activities will be limited, we assume the temporal segmentation must be more robust than in the case of data from a controlled environment.
 
-For the change detection algorithm we have used a special form of \gls{svm}, the \gls{occ} based \gls{svdd} as introduced by Tax~\cite{tax2001one}.
+For the change detection algorithm we will use a special form of a \gls{svm}, the \gls{occ} based \gls{svdd}, as introduced by Tax~\cite{tax2001one}.
 This method models the data under consideration in the shape of a high-dimensional hypersphere.
@@
 Since a change in data characteristics, resulting from a change in performed activity, alters the data distribution, the model and thus the radius should reflect this change as well.
+In the reverse relation, a change in the radius indicates a change in the data distribution and thus a change in the performed activity.
 
@@
 In our setup, we create a temporal segmentation instead of a fixed window length.
-Each segment should consist of homogeneous data, in the sense that the underlying data generating activity should not change within a segment.
+Each segment will consist of homogeneous data, in the sense that the underlying data generating activity should not change within a segment.
 Likewise, the previous and next segments (if present) should be generated by a different performed activity.
@@
 \section{One-Class Support Vector Machines}
-In the setup, as described in the previous section, we employ a model construction phase.
+In the setup, as described in the previous section and visualized in \Cref{fig:thesis_goal}, we employ a model construction phase.
 For this model we will use an \acrlong{oc-svm} implementation: the \acrlong{svdd} algorithm by Tax~\cite{tax2001one}.
@@
 In recent previous research two data sets for the context of \gls{har} have been made public.
-In order to benchmark our proposed method we tried to employ it to these data sets.
+In order to benchmark our proposed method we tried to apply our method to these data sets.
 However, it turned out the data sets were not appropriate for continuous temporal segmentation.

### 2014-01-10 Chapters/Chapter2.tex
Finalizing Chapter 2, literatur review...

@@
 %----------------------------------------------------------------------------------------
-\section{Outline}
-\emph{Not intented for the reader.}
-\begin{itemize}
-  \item Literature review about Temporal Segmentation (previous draft was more about classification)
-  \item Consider methods for the context of filter-methods for classification
-  \item Take a loot at 3-4 different kind of methods for change detection:
-    \begin{itemize}
-      \item Introduction with a lot of techniques
-      \item Explain why look at a few
-      \item CUSUM - or other more traditional methods
-      \item Density-ratio estimation
-      \item Support Vector Machines (?) - if there are more sources about this
-      \item (Dimensionality reduction) --> probably not
-      \item \emph{Not to much attention to all techniques, focus is on SVM}
-    \end{itemize}
-  \item With each method, shortly look at characteristics, strengths and weaknesses and consider applicability to accelerometer sensor data
-\end{itemize}
+% \section{Outline}
+% \emph{Not intented for the reader.}
+% \begin{itemize}
+%   \item Literature review about Temporal Segmentation (previous draft was more about classification)
+%   \item Consider methods for the context of filter-methods for classification
+%   \item Take a loot at 3-4 different kind of methods for change detection:
+%     \begin{itemize}
+%       \item Introduction with a lot of techniques
+%       \item Explain why look at a few
+%       \item CUSUM - or other more traditional methods
+%       \item Density-ratio estimation
+%       \item Support Vector Machines (?) - if there are more sources about this
+%       \item (Dimensionality reduction) --> probably not
+%       \item \emph{Not to much attention to all techniques, focus is on SVM}
+%     \end{itemize}
+%   \item With each method, shortly look at characteristics, strengths and weaknesses and consider applicability to accelerometer sensor data
+% \end{itemize}
 
-\section{Statistical framework}\label{statistical-framework}
+% \section{Statistical framework}\label{statistical-framework}
 
--- Nieuwe indeling Chapter 2 --
+% -- Nieuwe indeling Chapter 2 --
 
-\begin{description}
-  \item[Intro/outline] Introduction and overview of this chapter. Put in the context of temporal classification.
-  \item[Temporal segmentation] Give overview of used temporal segmentation methods, move towards a split in model building/dimensionality reduction and change detection methods
-  \item[Dimensionality reduction] Explain that \gls{dr} can be useful in data processing. Name a few techniques, make bridge to \gls{oc-svm}.
-  \item[Change detection] Describe methods for \emph{direct} change detection. End with notion
-    \begin{description}
-      \item[Model fitting]
-      \item[Cusum]
-      \item[Density ratio]
-    \end{description}
-  \item[Support Vector Machine based methods]
+% \begin{description}
+%   \item[Intro/outline] Introduction and overview of this chapter. Put in the context of temporal classification.
+%   \item[Temporal segmentation] Give overview of used temporal segmentation methods, move towards a split in model building/dimensionality reduction and change detection methods
+%   \item[Dimensionality reduction] Explain that \gls{dr} can be useful in data processing. Name a few techniques, make bridge to \gls{oc-svm}.
+%   \item[Change detection] Describe methods for \emph{direct} change detection. End with notion
+%     \begin{description}
+%       \item[Model fitting]
+%       \item[Cusum]
+%       \item[Density ratio]
+%     \end{description}
+%   \item[Support Vector Machine based methods] test
 
-\end{description}
-\input{./Chapters/Chapter2/overview.tex}
+% \end{description}
+As described in \Cref{Chapter1}, the focus of this research is the application of temporal segmentation to accelerometer data.
+This is with the assumption that explicit temporal segmentation can aid the process of temporal classification.
+
+\TODO{Give examples of temporal classification techniques}
+
+In recent years a lot is research is performed in the classification of human activities, recorded by on-body sensors.
+Especially since smartphones with accelerometer sensors became more available, research has focused on recordings from these devices.
+
+\TODO{Include highlights from old-classification text}
+
+In the following of this chapter we will start with an overview of statistical models used in this field of study.
+That section will act as a starting point to look at the different types, applications and methods of temporal segmentation.
+The next section, \Cref{sec:literature_review_temporal_segmentation} follows the distinction made and looks at a collection of temporal segmentation and change detection methods.
+It will also relate the concept of \gls{occ} to temporal segmentation and change detection.
+The final section of this chapter, \Cref{sec:literature_review_svm}, will discuss a specific class of algorithms for \gls{occ}, namely the methods that use \glspl{svm} for the classification task.
+
 \input{./Chapters/Chapter2/statistical_models.tex}
 \input{./Chapters/Chapter2/temporal_segmentation.tex}
-\input{./Chapters/Chapter2/change_detection_methods.tex}
+% \input{./Chapters/Chapter2/change_detection_methods.tex}
 
--- OLD --
+% -- OLD --
 
@@
 % CUSUM and GLR
-\input{./Chapters/Chapter2/cusum}
+% \input{./Chapters/Chapter2/cusum}
 

### 2014-01-10 Chapters/Chapter2/statistical_models.tex
Finalizing Chapter 2, literatur review...

@@
 \section{Statistical models}\label{sec:literature_review_statistical_models}
-
-\TODO{Introduce the diffent models and thus different approaches}
-
 Many applications require the detection of time points at which the underlying properties of a system change.
-This problem has received a lot of attention in the fields of data mining, etc... \TODO{list and refs}.
 Often this problem is formulated in a statistical framework, by inspecting the underlying data generating \gls{pdf} of the time series data.
@@
 It originates from control methods for detection from benchmarks.
-This method and some derivatiges are discussed and analyzed in \Cref{sec:cusum}.
+This method, and some derivatives, are discussed and analyzed in \Cref{subsec:literature_review_temporal_segmentation_cusum}.
 
@@
 The density-estimation methods rely on the log-likelihood ratio between \glspl{pdf}.
-The method of Camci \cite{camci2010change} follows an other approach within the statistical framework, by using a \gls{svm}.
+The method of Camci~\cite{camci2010change} follows an other approach within the statistical framework, by using a \gls{svm}.
 One problem it tries to overcome is the (claimed) weakness of many methods to detect a decrease in variance.
-The method represents the distribution over the data points as a hyper-sphere in a higher dimension using the kernel trick.
+The method represents the distribution over the data points as a hypersphere in a higher dimension using the kernel trick.
 A change in the \gls{pdf} is represented by a change in the radius of this sphere.
-\Cref{sec:literature_review_svm} discusses the \gls{svm}-methods.
-\TODO{Put emphasis that this is the source of inspiration for the chosen method?}
-
--- NEW --
-
-In the field of temporal segmentation of time series data we can distinguish different approaches and common frameworks in which they perform.
-The problem is commonly formulated in the statistical framework and regarding the properties of the underlying data generating \gls{pdf} fo the time series data.
-In that sense, a change point is then defined as a significant change in the properties of the \gls{pdf}, such as the mean or variance.
+In \Cref{sec:literature_review_svm} more applications of the \gls{svm}-based methods are discussed.
+Since \gls{occ} and the \gls{oc-svm} method is central in our approach, \Cref{sec:one_class_classification,sec:one_class_svm} discuss these techniques in detail.
 
@@
 In \Cref{subsec:literature_review_temporal_segmentation_cusum} a further overview of this class of methods is discussed.
-
-% Piecewise approximation
-
--- Density
-
--- Outlier based

### 2014-01-10 Chapters/Chapter2/svm.tex
Finalizing Chapter 2, literatur review...

@@
 The \gls{nu-svm} method instead starts with the number of data points that should be within the region and estimates a region with that desired property.
-It builds on the method of Vapnik and Vladimir \cite{vapnik1963pattern}, which characterizes a set of data by separating it from the origin.
+It builds on the method of Vapnik and Vladimir \cite{vapnik1963pattern}, which characterizes a set of data points by separating it from the origin.
 The \gls{nu-svm} method adds the kernel method, allowing non-linear decision functions, and incorporates `softness' by the $\nu$-parameter.
@@
 
-The inner workings of \glspl{oc-svm} are discussed in detail in \Cref{sec:one_class_svm}.
-
-\TODO{Add notion of `supervised' vs `unsupervised' learning. \gls{oc-svm} is unsupervised.}
+The following chapter discusses \gls{occ}, \gls{oc-svm}, and change detection through \gls{oc-svm} in detail.
+It will show that the used implementation of \gls{oc-svm} is an example of the Type 3 semi-supervised learning methods, in the system of Hodge and Austin~\cite{hodge2004survey}.
 

### 2014-01-10 Chapters/Chapter2/temporal_segmentation.tex
Finalizing Chapter 2, literatur review...

@@
 
-\TODO{Is there any research of temporal segmentation on accelerometer/inertia signals? Note if there is/isnt}
+% \TODO{Is there any research of temporal segmentation on accelerometer/inertia signals? Note if there is/isnt}
 
-Give overview of used temporal segmentation methods. Problem is that it is not really well-defined.
+% Give overview of used temporal segmentation methods. Problem is that it is not really well-defined.
 
-\begin{itemize}
-  \item Given overview of segmentation techniques, for times series data
-  \item Use different ``point-of-views'', or terms
-  \item ``Segmentation''
-  \item ``Change detection''
-  \item ``Novelty detection''
-  \item Specific view on \glspl{svm}
-\end{itemize}
+% \begin{itemize}
+%   \item Given overview of segmentation techniques, for times series data
+%   \item Use different ``point-of-views'', or terms
+%   \item ``Segmentation''
+%   \item ``Change detection''
+%   \item ``Novelty detection''
+%   \item Specific view on \glspl{svm}
+% \end{itemize}
 
 This section gives an overview of the literature on temporal segmentation in the context of \gls{har}.
-It takes a look on different implementations and methodologies.
+It provides a look on different implementations and methodologies.
 A wide range of terms and subtle differences are used in the field, such as `segmentation', `change detection', `novelty detection' and `outlier detection'.
-These will be the categorical terms for which we discuss the literature.
+The following sections discuss these different methods.
 
-Aspects we will consider when discussing the literature are applicability to \gls{har}, parameter tuning and ... \TODO{What more aspects?}
-% Finally we will discuss other applications of \glspl{svm} in the context of these terms.
-
-\subsection{Segmentation}\label{subsec:segmentation}
+\subsection{Segmentation based methods}\label{subsec:segmentation}
 
@@
 Other methods have been proposed, such as an adaptive threshold based on the signal energy by Guenterberg \etal \cite{guenterberg2009automatic}, the adaptive \gls{cusum}-based test by Alippi \etal \cite{alippi2006adaptive} and the \gls{mosum} by Hsu \cite{hsu2007mosum} in order to eliminate this user-dependency.
-The latter of these methods is able to process the accelerometer values directly, although better results are obtained when features of the signal are processed, as done in the former method.
-Here the signal energy, mean and standard deviation are used to segment activities and by adding all the axial time series together, the Signal-To-Noise ration is increased, resulting in a robuster method.
+The latter of these methods is able to process the accelerometer values directly, although better results are obtained when features of the signal are processed, as done in the first method \cite{guenterberg2009automatic}.
+Here the signal energy, mean and standard deviation are used to segment activities and by adding all the axial time series together, the Signal-To-Noise ratio is increased, resulting in a robuster method.
 
@@
 The authors claim that only a slight modification is needed to perform the method on multi-dimensional data.
-After the segmentation phase, the method of Bernecker \etal extracts other statistical features which are used in the classification phase.
+After the segmentation phase, the method of Bernecker \etal \cite{bernecker2012activity} extracts other statistical features which are used in the classification phase.
 
@@
 The research also considered the (Stepwise) Feasable Space Window as introduced by \cite{liu2008novel}, but since it results in a higher error rate than \gls{swab} the latter was chosen to extend.
-Whereas the before mentioned algorithms use a linear representation, this methods considers linear, quadratic and cubical representations for each segment.
+Whereas the above mentioned algorithms use a linear representation, this methods considers linear, quadratic and cubical representations for each segment.
 This differs from other methods where the model is fixed for the whole time series, such as \cite{fuchs2010online}, which is stated to perform inferior on non-stationary time series such as daily life.
@@
 The method of Adams and MacKay \cite{adams2007bayesian} builds a probabilistic model on the segment run length, given the observed data so far.
-Instead of modeling the values of the data points as a probabilistic distribution, the length of segments as a function if time is modeled by calculating the posterior probability.
+Instead of modeling the values of the data points as a probabilistic distribution, the length of segments as a function of time is modeled by calculating the posterior probability.
 It uses a prior estimate for the run length and a predictive distribution for newly-observed data, given the data since the last change point.
@@
 
-Many methods obtain an implicit segmentation as a result of classification over a sliding window \TODO{add refs}.
+As stated before, oten methods obtain an implicit segmentation as a result of classification over a sliding window.
 The method of Yang \etal \cite{yang2008distributed} explicitly performs segmentation and classification simultaneously.
@@
 
-Field of computer vision: \cite{zhou2008aligned}, \cite{li2007segmentation}.
+% Field of computer vision: \cite{zhou2008aligned}, \cite{li2007segmentation}.
 
@@
 
-``Segmentation and Recognition of Motion Streams by Similarity Search'' \cite{li2007segmentation}. 29, 2007 \\
+% ``Segmentation and Recognition of Motion Streams by Similarity Search'' \cite{li2007segmentation}. 29, 2007 \\
 
@@
 
-``Aligned Cluster Analysis for Temporal Segmentation of Human Motion'' \cite{zhou2008aligned}. 63, 2008 \\
+% ``Aligned Cluster Analysis for Temporal Segmentation of Human Motion'' \cite{zhou2008aligned}. 63, 2008 \\
 
@@
 
-\TODO{Link this subsection with previous and next}
+% \TODO{Link this subsection with previous and next}
 % Papers:
@@
 
-\emph{Notes:}
-Non-Bayesian change detection algorithm (\ie no prior distribution believe available for the change time).
-The \gls{cusum} method is developed by Page \cite{page1954continuous} for the application of statistical quality control (it is also known as a control chart).
-Primary for detection of mean shift.
-The \gls{mosum} of squares test for monitoring variance changes \cite{hsu2007mosum}.
-Use of Cumulative Sums of Squares for Retrospective Detection of Changes of Variance \cite{inclan1994use}
+% \emph{Notes:}
+% Non-Bayesian change detection algorithm (\ie no prior distribution believe available for the change time).
+% The \gls{cusum} method is developed by Page \cite{page1954continuous} for the application of statistical quality control (it is also known as a control chart).
+% Primary for detection of mean shift.
+% The \gls{mosum} of squares test for monitoring variance changes \cite{hsu2007mosum}.
+% Use of Cumulative Sums of Squares for Retrospective Detection of Changes of Variance \cite{inclan1994use}
 
-An often used approach in the statistical framework of change detection is the \gls{cusum} as introduced by Page \cite{page1954continuous}.
-Originally used for quality control in production environments, its main function is to detect change in the mean of measurements and has been applied to this problem \cite{basseville1993detection}.
+An other often used approach in the statistical framework of change detection is the \gls{cusum} as introduced by Page \cite{page1954continuous}.
+Originally used for quality control in production environments, its main function is to detect change in the mean of measurements \cite{basseville1993detection}.
 It is a non-Bayesian method and thus makes no assumptions, in the form of prior belief distributions, for the change points.
@@
 The method of Incl\'{a}n and Tiao \cite{inclan1994use} builds on the centered version of \gls{cusum} \cite{brown1975techniques} to detect changes in variance.
+The obtained results (when applied to stock data) are comparable with the application of \gls{mle}.
 Using the batch \gls{icss} algorithm they are able to find multiple change points, offline while post-processing the data.
+Whereas \gls{cusum} can be applied to search for a change in mean, the \gls{icss} is adapted to find changes in variance.
 Let $C_k = \sum_{i=1}^k \alpha_t^2$ be the cumulative sum of squares for a series of uncorrelated random variables $\{\alpha_t\}$ of length $T$.
@@
 For a series with homogeneous variance, the value of $D_k$ will oscillate around $0$.
+This $D_k$ is used to obtain a \emph{likelihood ratio} for testing the hypothesis of one change against no change in the variance.
 In case of a sudden change, the value will increase and exceed some predefined boundary with high probability.
 By using an iterative algorithm, the method is able to minimize the masking effect of successive change points.
+The proposal of \cite{cheng2009efficient} extends on the \gls{cusum}-based methods to find change points in mean and variance, by creating a more efficient and accurate algorithm.
 
@@
 
-\subsection{Change detection}\label{subsec:change_detection}
-\TODO{change order of this and temporal segmentation sections? So first change, then segmentation?} \\
-Whereas the above mentioned researches focus on \emph{segmentation}, many have focused on \emph{change detection}.
-Although these techniques are closely related, there is a subtle difference.
-In the case of \emph{change detection} to goal is to find, possibly unrelated, sudden change points in a signal \cite{takeuchi2006unifying}.
-In contrast, the goal of \emph{temporal segmentation} is to find homogeneous segments of data, which can be the result of multiple detected changes.
+% \subsection{Change detection methods}\label{subsec:change_detection}
+% \TODO{change order of this and temporal segmentation sections? So first change, then segmentation?} \\
+% Whereas the above mentioned researches focus on \emph{segmentation}, many have focused on \emph{change detection}.
+% Although these techniques are closely related, there is a subtle difference.
+% In the case of \emph{change detection} to goal is to find, possibly unrelated, sudden change points in a signal \cite{takeuchi2006unifying}.
+% In contrast, the goal of \emph{temporal segmentation} is to find homogeneous segments of data, which can be the result of multiple detected changes.
 
-The \gls{icss} by Incl\'{a}n and Tiao \cite{inclan1994use} is a statistical method which obtains results (when applied to stock data) comparable to \gls{mle} and Bayesian \TODO{Bayesian What?}.
-Whereas \gls{cusum} can be applied to search for a change in mean, the \gls{icss} is adapted to find changes in variance.
-It obtains a \emph{likelihood ratio} for testing the hypothesis of one change against no change in the variance.
-Using an iterative approach, all possible change points are considered.
-The proposal of \cite{cheng2009efficient} extends on the \gls{cusum}-based methods to find change points in mean and variance, by creating a more efficient and accurate algorithm.
+% The \gls{icss} by Incl\'{a}n and Tiao \cite{inclan1994use} is a statistical method which obtains results (when applied to stock data) comparable to \gls{mle} and Bayesian \TODO{Bayesian What?}.
+% Whereas \gls{cusum} can be applied to search for a change in mean, the \gls{icss} is adapted to find changes in variance.
+% It obtains a \emph{likelihood ratio} for testing the hypothesis of one change against no change in the variance.
+% Using an iterative approach, all possible change points are considered.
+% The proposal of \cite{cheng2009efficient} extends on the \gls{cusum}-based methods to find change points in mean and variance, by creating a more efficient and accurate algorithm.
 
-\Cref{sec:change_detection_time_series} discusses the problem of change detection in time series further, and gives a formal problem definition.
+% \Cref{sec:change_detection_time_series} discusses the problem of change detection in time series further, and gives a formal problem definition.
 
-\TODO{move CUSUM based techniques to this subsection}
+% \TODO{move CUSUM based techniques to this subsection}
 

### 2014-01-10 Chapters/Chapter4.tex
WIP: Chapter 4
Finalizing chapter 7, conclusion
Finalizing Chapter 2, literatur review...

@@
 
-Ideen naam methode:
+\TODO{Decide on name}
 \begin{itemize}
@@
 This chapter will introduce our method and setup for the temporal segmentation of human activities, using an \gls{oc-svm} approach.
+It shows our method to apply a \gls{oc-svm} based temporal segmentation method to inertial sensor time series data, often used in the context of \acrlong{har}.
+As far as we know it is the first in its kind, especially on the application of \gls{svdd} and \gls{svcpd} to real-world \gls{har} data.
 The setup follows the structure as illustrated in \Cref{fig:method_overview} and further discussed in \Cref{sec:method_overview}.
@@
 \Cref{sec:method_model_features} goes into detail about the rationale of using that feature.
-Finally, in \Cref{sec:change_detection_methods} we discuss two different, but simple, methods to detect changes in the time series data, based on the extracted radius $R$.
+Finally, in \Cref{sec:method_change_detection} we discuss two different, but simple, methods to detect changes in the time series data, based on the extracted radius $R$.
 

### 2014-01-10 Chapters/Chapter1.tex
Added plots for chapter 6

@@
 % Chapter 1
-
--- Overall todo's --
-\begin{itemize}
-  \item \TODO{The term \emph{emperical data/distruction} is nowhere used in the thesis. or: Emperical cumulative distribution. Especially Chapter 2.}
-  \item \TODO{Check for ``tang-constructions'': saying to much in a single sentence.}
-  \item \TODO{Overall: meer focus op de \emph{waarom} vraag/antwoord bij keuzes. Bouwt aan motivatie en context van het geheel.}
-  \item \TODO{Consistent `data point' of `data object' gebruiken.}
-\end{itemize}
-
 \chapter{Introduction}

### 2014-01-10 Chapters/Chapter6/results.tex
Added plots for chapter 6

@@
 To overcome this, we will state observations and remarks found after close (visual) inspection of the annotated and discovered change points.
+In \Cref{fig:plots_all_runs_results} all the results are plotted.
+The black solid lines indicate our manual determined change points.
+The dashed purple lines indicate the change points as discovered by our method.
 We will start with general observations applicable to all the runs and will continue with individual remarks for each run.
@@
 
-Considering the overall performance and results, we have the following observations:
+Overall our impression is that the proposed method is fairly good at recognizing change points.
+Depending on the parameters for the sensitivity adjustment, almost all changes are discovered.
+With a high sensitivity it seems, initially, the method produces many false positives.
+After closer inspecting of the data, we can confirm there is a change in the data, although the performed activity stayed the same.
+This is also a weakness of the method: since the sensitivity is set globally, it can produce many semi-false positive change points for some segments of a run.
+
+An other positive observation is that the method often recognizes a change point \emph{before} our annotation.
+Again, after closer inspection we can see that the data changes shape and distribution before we annotated a change based on the video recordings.
+This effect can be observed in the case of a transition from running to walking; the last `steps' of running are already recognized of being different.
+As with the above behavior, this is subject to parameter settings.
+
+Considering the overall performance and results, we have the following observations of cases in which the method has problems with detecting change:
 \begin{itemize}
@@
 
-\TODO{Add positive remarks}
-
 When looking at the individual runs, we have the following observations for each run number:
 \begin{enumerate}
-  \item \textbf{Subject 2, straight walk and run}
+  \item \textbf{Subject 2, straight walk and run} \Cref{fig:data_with_cps_run_1}
     \begin{itemize}
@@
     \end{itemize}
-  \item \textbf{Subject 1, straight walk and run}
+  \item \textbf{Subject 1, straight walk and run} \Cref{fig:data_with_cps_run_2}
     \begin{itemize}
@@
     \end{itemize}
-  \item \textbf{Subject 2, walk around corner}
+  \item \textbf{Subject 2, walk around corner} \Cref{fig:data_with_cps_run_3}
     \begin{itemize}
@@
     \end{itemize}
-  \item \textbf{Subject 2, walk and run around fountain}
+  \item \textbf{Subject 2, walk and run around fountain} \Cref{fig:data_with_cps_run_4}
     \begin{itemize}
@@
       Due to the global parameter settings, the sensitivity is too high for this segment to recognize it as one.
+      It could also be that our used window-length is too small, to model a good representation of this segment.
       \item During the circular run, from $12s$ till $24$, there are two change points discovered.
       The difference for the rotational vectors need to accumulate to a certain value before they have enough influence to let the rotation be regarded as a change point.
-      \TODO{Beter formuleren}.
+
     \end{itemize}
-  \item \textbf{Subject 1, walk and run around fountain}
+  \item \textbf{Subject 1, walk and run around fountain} \Cref{fig:data_with_cps_run_5}
     \begin{itemize}
@@
     \end{itemize}
-  \item \textbf{Subject 2, walk and run fountain 2}
+  \item \textbf{Subject 2, walk and run fountain 2} \Cref{fig:data_with_cps_run_6}
     \begin{itemize}
@@
     \end{itemize}
-  \item \textbf{Subject 3, indoor stairs}
+  \item \textbf{Subject 3, indoor stairs} \Cref{fig:data_with_cps_run_7}
     \begin{itemize}
@@
 
-\TODO{Samenvatten observations en remarks, brugje naar possible improvements maken}
+\begin{figure}
+  \centering
+  \begin{subfigure}{1\textwidth}
+    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/run-1-walk-run-roemer/data_plot_acc_with_discovered_cps.eps}
+    \caption{Run 1}
+    \label{fig:data_with_cps_run_1}
+  \end{subfigure}
+
+  \begin{subfigure}{1\textwidth}
+    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/run-2-walk-run-jos/data_plot_acc_with_discovered_cps.eps}
+    \caption{Run 2}
+    \label{fig:data_with_cps_run_2}
+  \end{subfigure}
+
+  \begin{subfigure}{1\textwidth}
+    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/run-3-walk-turn-roemer/data_plot_acc_with_discovered_cps.eps}
+    \caption{Run 3}
+    \label{fig:data_with_cps_run_3}
+  \end{subfigure}
+
+  \begin{subfigure}{1\textwidth}
+    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/run-4-run-fountain-roemer/data_plot_acc_with_discovered_cps.eps}
+    \caption{Run 4}
+    \label{fig:data_with_cps_run_4}
+  \end{subfigure}
+
+  \begin{subfigure}{1\textwidth}
+    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/run-5-run-fountain-jos/data_plot_acc_with_discovered_cps.eps}
+    \caption{Run 5}
+    \label{fig:data_with_cps_run_5}
+  \end{subfigure}
+
+  \begin{subfigure}{1\textwidth}
+    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/run-6-walk-run-roemer/data_plot_acc_with_discovered_cps.eps}
+    \caption{Run 6}
+    \label{fig:data_with_cps_run_6}
+  \end{subfigure}
+
+  \begin{subfigure}{1\textwidth}
+    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/stairs-1-marc/data_plot_acc_with_discovered_cps.eps}
+    \caption{Run 7}
+    \label{fig:data_with_cps_run_7}
+  \end{subfigure}
+
+  \caption[Results run 1/8]{Annotated plots of all runs.
+The black vertical lines are manually determined change points, the purple dashes lines are the discovered change points.
+The label above each segment indicates the current performed activity.}\label{fig:plots_all_runs_results}
+\end{figure}
 

### 2014-01-10 main.tex
Added plots for chapter 6
Put captions above tables

@@
 \emph{Author:}\\
-\href{http://rvlasveld.github.io}{\authornames}, \href{http://www.uu.nl}\textsc{UU} % Author name - remove the \href bracket to remove the link
+\href{http://rvlasveld.github.io}{\authornames}, \href{http://www.uu.nl}{\textsc{UU}} % Author name - remove the \href bracket to remove the link
 \end{flushleft}
@@
 \href{http://www.cs.uu.nl/staff/gv.html}{Dr. G.A.W. \textsc{Vreeswijk}, \href{http://www.uu.nl}\textsc{UU}}  \\ % Supervisor name - remove the \href bracket to remove the link
-\href{http://www.annevanrossum.com}{Anne \textsc{van Rossum}, \href{http://www.almende.com}\textsc{Almende B.V.}} % Supervisor name - remove the \href bracket to remove the link
+\href{http://www.annevanrossum.com}{A.C. \textsc{van Rossum}, \href{http://www.almende.com}\textsc{Almende B.V.}} % Supervisor name - remove the \href bracket to remove the link
 \end{flushright}
@@
 }
+\TODO{Write some nice stuff here}
 \clearpage % Start a new page

### 2014-01-13 Chapters/Chapter1.tex
Feedback Chapter 1 (not everything)

@@
 Many approaches to recognize the performed activity rely on classification of recorded data \cite{devaul2001real,ward2006activity,yang2008using,anguita2012human,bao2004activity,bernecker2012activity,he2008activity}, which can be done online or in batches after the performance \cite{duque2012offline}.
-Often a time-windowed based approach is used, which feeds short consecutive segments of data through a model construction and recognition phase.
+Often time windows are used, which form short consecutive segments of data and are processed in a model construction and recognition phase.
 The constructed model is used to determine which (earlier learned) activity is performed.
@@
 In this thesis the goal is to find the temporal segmentation of time series data explicitly, prior to the classification of the activities.
+
 This is done under the assumption that it could be beneficial for classification methods to possess this explicit segmentation, since the model construction phase can use more data than the time-windowed approach would allow.
-To find a temporal segmentation, we rely on change detection in time series data.
-In the context of this research, this time series data consists of recordings from inertial sensors found in smartphones, such as the accelerometer signal, during the performance of human activities.
-We are interested in both in- and outdoor activities in an uncontrolled environment and in a continuous manner, which can be regarded as real-world data.
+
+To find a temporal segmentation, we rely on change detection in time series.
+In the context of this research, these time series consists of recordings from inertial sensors found in smartphones, such as the accelerometer signal, during the performance of human activities.
+We are interested in real-world data from both in- and outdoor activities in an uncontrolled environment and in a continuous manner.
 Although the number of performed activities will be limited, we assume the temporal segmentation must be more robust than in the case of data from a controlled environment.
-
 For the change detection algorithm we will use a special form of a \gls{svm}, the \gls{occ} based \gls{svdd}, as introduced by Tax~\cite{tax2001one}.
-This method models the data under consideration in the shape of a high-dimensional hypersphere.
+This method models the data in the shape of a high-dimensional hypersphere.
 From this hypersphere the radius is extracted, which can be used as an indication of change and be processed by one-dimensional change detection algorithms.
-Since a change in data characteristics, resulting from a change in performed activity, alters the data distribution, the model and thus the radius should reflect this change as well.
-In the reverse relation, a change in the radius indicates a change in the data distribution and thus a change in the performed activity.
+The radius of the hypersphere is used in the following relation.
+When the performed activity changes, so will the data distribution and characteristics.
+This results in a change of the constructed model, especially in the radius of the hypersphere.
+Thus, a change in the radius indicates a changes in the performed activity.
 
@@
 This relation is illustrated in \Cref{fig:thesis_goal}.
-The figure shows that we can apply classification over windows of the time series data.
-In many approaches this is constructed with a (partially overlapping) sliding window.
-In the context of \gls{har}, the window length commonly consists of around $2$ seconds of data points.
-In our setup, we create a temporal segmentation instead of a fixed window length.
+The figure visualizes that we can apply classification on time series data.
+In many approaches it is applied to (partially overlapping) windows of data.
+In the context of \gls{har}, the window length commonly consists of around two seconds of data.
+In our setup we perform temporal segmentation, instead of using a fixed window length.
 Each segment will consist of homogeneous data, in the sense that the underlying data generating activity should not change within a segment.
-Likewise, the previous and next segments (if present) should be generated by a different performed activity.
+Likewise, the previous and next segments (if present) should be generated by a different activity.
 
-To find these segments of homogeneous data, we employ a model construction phase.
-The \gls{oc-svm} model transforms the data points to a high-dimensional hyperspherical shape from which we extract the radius as characterizing metric.
-In this phase the data is processed by a fixed window length and a change in data characteristics should reflect in a (significant) change of the hypersphere's radius.
-Such a sudden change in the radius thus indicates a change in the time series data and indicates the end of the current and beginning of a new segment.
-In \Cref{sec:literature_review_temporal_segmentation} a more detailed overview of temporal segmentation methods is discussed.
-The information about segments can be used in the classification phase of \gls{har}, which applies classification to windows (or in this case: segments) of data.
-The final result, which is outside the scope is this thesis, will be a full classification of activities over the time series data.
-
-\begin{figure}
+\begin{figure}[h]
   \centering
@@
 
+To find these segments of homogeneous data, we employ a model construction phase.
+The \gls{oc-svm} model transforms the data points to a high-dimensional hyperspherical shape from which we extract the radius as characterizing metric.
+The model processes the data in windows of fixed length.
+A change in data characteristics should then result in a change of the hypersphere's radius.
+Such a (sudden) change in the radius thus reflects a change in the time series data.
+It indicates the end of the current and the beginning of a new segment.
+In \Cref{sec:literature_review_temporal_segmentation} a detailed overview of temporal segmentation methods is given.
+Information about segments can be used in the classification phase of \gls{har}.
+Instead of using relatively small windows, it can apply classification to the full segments of data.
+The final result, which is outside the scope is this thesis, will be a full classification of activities over the time series data.
+
 \section{One-Class Support Vector Machines}
@@
 In earlier research multi-class \glspl{svm} are used for direct classification of inertial sensor data \cite{he2008activity,mountrakis2011support,anguita2012human}.
-Others have used \gls{oc-svm} classifiers for novelty, or outlier, detection \cite{scholkopf1999support,camci2010change,li2003improving,ma2003time,tax1999support}.
-The research by Yin \etal  \cite{yin2008sensor} creates a \gls{oc-svm} classifier to filter normal activity traces, using training data.
+Others have used \gls{oc-svm} classifiers for novelty or outlier detection \cite{scholkopf1999support,camci2010change,li2003improving,ma2003time,tax1999support}.
+The method by Yin \etal  \cite{yin2008sensor} creates a \gls{oc-svm} classifier to filter normal activity traces, using training data.
 However, to the best of our knowledge, \gls{oc-svm} methods have not yet been applied to inertial sensor time series data directly to find changes in behavior and create a temporal segmentation.
-This algorithm transforms the time series data, which can be of any high dimension, to an higher dimensional feature space.
-When the data is assumed to be of the same class, it can be characterized by an enclosing boundary, in the shape of a hypersphere.
-Using this \gls{occ} method, we are able to test new data points of being a member of this class.
-When a new data point, after transformation to the feature space, lies within the boundary, it is a member of the class, and otherwise it is regarded as an outlier.
 
-In case the data distribution changes, the data points will have a larger variance which results in a larger hypersphere in the feature space.
-This means that with continuous model updating, a change in the underlying process of a time series data means the hypersphere will have a larger radius.
-Thus the \gls{oc-svm} method allows us to generate a one-dimensional time series data, reduced from a complex high-dimensional original signal.
-The final radius time series data can be inspected by simpler change detection methods, in a number of ways.
+The \glspl{oc-svm} algorithms transforms the time series data, which can be of any high dimension, to a higher dimensional feature space.
+When the data is assumed to be of the same class, it can be characterized by an enclosing boundary.
+In case of the \gls{svdd} algorithm this boundary is in the shape of a hypersphere.
+Using this model, we are able to test the membership of a new data point to the described class.
+The new data point is transformed to the feature space and its location is compared to the boundary.
+If it lies within the boundary then it is a member of the class and otherwise it is regarded as an outlier.
+
+In case the data distribution changes, the data points will have a larger variance.
+It will result in a larger hypersphere in the feature space.
+This means that a change in the underlying process of a time series data results in a larger hypersphere radius.
+Thus the \gls{oc-svm} method allows us to reduce a complex high-dimensional signal to an one-dimensional time series data.
+The reduced time series can be inspected by simpler change detection methods, in a number of ways.
 \Cref{Chapter3} will discuss the \gls{oc-svm} methods in detail and \Cref{Chapter4} elaborates on the overall change detection method.
@@
 \section{Data sets}
-In recent previous research two data sets for the context of \gls{har} have been made public.
-In order to benchmark our proposed method we tried to apply our method to these data sets.
+In recent research two data sets for \gls{har} have been made public.
+In order to benchmark our proposed method we applied our method to these data sets.
 However, it turned out the data sets were not appropriate for continuous temporal segmentation.
-The first set, the \gls{wisdm} from \cite{kwapisz2011activity}, consists of non-continuous recordings of activities.
-Obviously, this voids the usefulness of the data set of our research.
-The second set, the \gls{uci-har} from \cite{anguita2012human}, also lacks continuous recordings and furthermore exhibits patterns and labeling which seems to be incorrect.
-Besides these problems, since there are no visual images of the performed activities, there is no ground truth which can be consulted in case of ambiguity.
-To overcome these problems, we have created our own real-world data sets.
+In the \gls{wisdm} set \cite{kwapisz2011activity} the activities are not performed continuously.
+There is no transition period between activities.
+Since we are interested in the detection changes between consecutive activities, we can not use this data set.
+The \gls{uci-har} from \cite{anguita2012human} also lacks continuous recordings.
+Furthermore, it seems to have incorrect labels.
+Finally, for both data sets there are no visual images of the performed activities.
+So in case of ambiguity there is no ground truth to consult.
+To overcome these problems, we have created our own real-world data set.
+\TODO{Create name for our own data sets}
 
-Our own data sets are recorded by placing a smartphone with inertial sensors in the front right pants pocket of a subject.
-The subject was asked to perform common and simple activities in an uncontrolled environment, both in- and outdoor situated.
-During the performance the subjects were also recorded with a video camera from the third-person point of view.
+Our own data set is recorded by placing a smartphone with inertial sensors in the front right pants pocket of a subject.
+The subject was asked to perform common and simple activities in an uncontrolled environment, both in- and outdoors.
+During the performance the subjects were also recorded with a video camera from a third-person point of view.
 This allows us to create a ground truth for the temporal segmentation of performed activities.
@@
   This data set can be grounded by using the video recordings of the subject while performing the activities.
-  The data set, consisting of raw inertial sensor recordings, manual change point and activity labeling, and video recordings, are used for testing the method and are made publicly available \cite{vlasveld2013continuous}.
-  \item \textbf{Application of \gls{oc-svm} methods to inertial sensor data:} to the best of our knowledge, this thesis is the first application of \gls{oc-svm} methods, especially the \gls{svdd} method, to inertial sensor signals in the context of creating a temporal segmentation.
-  Furthermore, this thesis contributes to the small amount of research that has a temporal segmentation of inertial sensor data as its primary objective.
-  \item \textbf{Adaptation of \acrshort{svcpd}:} we have adopted and slightly adapted the \acrshort{svcpd} method by Camci~\cite{camci2010change} in order to find change points in time series data.
-  In our simplification we have not encountered noticeably decrease in performance.
+  The data set, consisting of raw inertial sensor recordings, manual change point and activity labeling, and video recordings, is used for testing the method and is made available to the public \cite{vlasveld2013continuous}.
+  \item \textbf{Application of \gls{oc-svm} methods to inertial sensor data:} to the best of our knowledge, this thesis is the first application of \gls{oc-svm} methods, especially the \gls{svdd} method, to inertial sensor signals in the context of temporal segmentation.
+  Furthermore, this thesis contributes to research that has temporal segmentation of inertial sensor data as its primary objective.
+  \item \textbf{Adaptation of \acrshort{svcpd}:} we have adopted and simplified the \acrshort{svcpd} method by Camci~\cite{camci2010change} in order to find change points in time series data.
+  In our simplification we have not encountered a noticeably decrease in performance.
 \end{enumerate}
@@
 Previous work in the field of \gls{har} is discussed and we end with existing applications of \glspl{svm} to change detection.
-The next chapter further analyses \glspl{svm} and two implementations of \glspl{oc-svm}.
+\Cref{Chapter3} further analyses \glspl{svm} and two implementations of \glspl{oc-svm}.
 It relates the properties of the constructed \gls{oc-svm} models to change detection in time series data.
 That relation is applied to our problem formulation in \Cref{Chapter4}, where we construct our change detection method \TODO{add name}, based on the work of Camci~\cite{camci2010change}.
-The two following chapters apply the method to both artificial and real-world data.
-\Cref{Chapter5} show the objective performance of the method to artificial constructed data.
-In \Cref{Chapter6} we apply the method to our self-recorded real-world data.
+In \Cref{Chapter5} we apply the method to artificial data and show the objective performance.
+In \Cref{Chapter6} the method is applied to our real-world data set \TODO{Add name}.
 In that chapter we show the ability of detecting change in \gls{har} data, recorded by inertial sensors.

### 2014-01-13 Chapters/Chapter6/data_sets.tex
Feedback Chapter 1 (not everything)

@@
 The manually determined change points are used to draw quality conclusion over the algorithm's performance, by comparing it to the discovered change points.
-All of the used data sets, including the video material and manual labeling, are made publicly available \cite{vlasveld2013continuous} for further research.
+All of the used data sets, including the video material and manual labeling, are made available to the public \cite{vlasveld2013continuous} for further research.
 \TODO{Publish web page with data sets}

### 2014-01-13 Chapters/Chapter2.tex
Added plots for chapter 6
Feedback Chapter 2 (not everything

@@
 
-\TODO{Give examples of temporal classification techniques}
+In recent years a lot of is research is performed to study classification of human activities, recorded by on-body sensors.
+Especially since smartphones with accelerometer sensors became widely available, research has focused on recordings from these devices.
 
-In recent years a lot is research is performed in the classification of human activities, recorded by on-body sensors.
-Especially since smartphones with accelerometer sensors became more available, research has focused on recordings from these devices.
-
-\TODO{Include highlights from old-classification text}
-
-In the following of this chapter we will start with an overview of statistical models used in this field of study.
+In \Cref{sec:literature_review_statistical_models} we will start with an overview of statistical models used in this field of study.
 That section will act as a starting point to look at the different types, applications and methods of temporal segmentation.
-The next section, \Cref{sec:literature_review_temporal_segmentation} follows the distinction made and looks at a collection of temporal segmentation and change detection methods.
+\Cref{sec:literature_review_temporal_segmentation} follows the distinction made and looks at a collection of temporal segmentation and change detection methods.
 It will also relate the concept of \gls{occ} to temporal segmentation and change detection.
+In \Cref{sec:density-ratio} methods using density-ratio estimation of \glspl{pdf} for change detection are discussed.
 The final section of this chapter, \Cref{sec:literature_review_svm}, will discuss a specific class of algorithms for \gls{occ}, namely the methods that use \glspl{svm} for the classification task.

### 2014-01-13 Chapters/Chapter2/density_ratio.tex
WIP: Working on organization of Chapter 2
Feedback Chapter 2 (not everything

@@
 % !TEX root = ../../main.tex
-\section{Change-detection by Density-Ratio Estimation}\label{density-ratio}
+\section{Change-detection by Density-Ratio Estimation}\label{sec:density-ratio}
 
@@
 Many approaches to detect change points monitor the logarithm of the likelihood ratio between two consecutive intervals.
-A change is point is regarded to be the moment in time when the underlying probabisitic generation function changes.
+A change point is regarded to be the moment in time when the underlying probabilistic generation function changes.
 Some methods which rely on this are novelty detection, maximum-likelihood estimation and online learning of autoregressive models \cite{kawahara2009change}.
 A limitation of these methods is that they rely on pre-specified parametric models.
-Non-parametric models for density estimation have been proposed, but it is said to be a hard problem \cite{hardle2004nonparametric, sugiyama2012density}.
-A solutions to this is to estimate the \emph{ratio} of probabilities instead of the probabilities themselves.
+Nonparametric models, for which the number and nature of the parameters are undetermined, for density estimation have been proposed, but it is said to be a hard problem \cite{hardle2004nonparametric, sugiyama2012density}.
+A solution to this is to estimate the \emph{ratio} of probabilities instead of the probabilities themselves.
 One of the recent methods to achieve this is the \gls{kliep} by Sugiyama \etal \cite{sugiyama2008direct}.
 
-The method proposed by Kawahara and Sugiyama \cite{kawahara2009change} is composed of an online version of the \gls{kliep} algorithm.
+The method proposed by Kawahara and Sugiyama~\cite{kawahara2009change} is composed of an online version of the \gls{kliep} algorithm.
 The method also considers \emph{sequences} of samples (rather than samples directly) because the time series samples are generally not independent over time.
-An advantage over other non-parametric approaches, such as sequential one-class support vector machines, is that the model has an natural cross-validation procedure.
+An advantage over other nonparametric approaches, such as sequential one-class support vector machines, is that the model has a natural cross-validation procedure.
 This makes that the value of tuning parameters, such as the kernel bandwidth, can be objectively obtained.
@@
 \begin{equation}
-  S = \sum_{i=1}^{n_\text{te}} \text{ln} \frac{p_{\text{te}}(\matrixsym{Y}_\text{te}(i))}{p_\text{rf}(\matrixsym{Y}_\text{te}(i))}
+  S = \sum_{i=1}^{n_\text{te}} \text{ln} \frac{p_{\text{te}}(\matrixsym{Y}_\text{te}(i)}{p_\text{rf}(\matrixsym{Y}_\text{te}(i))}
 \end{equation}
@@
 Improvements in this line of research by Liu \etal \cite{liu2013change} has led the application of improved density-ratio estimation methods to the problem of change detection.
-Such an improvement is the Unconstrained Least-Squares Importance Fitting (uLSIF) method \cite{kanamori2009least} and an extension which possesses a superior non-parametric convergence property: Relative uLSIF (RuLSIF) \cite{yamada2013relative}.
+Such an improvement is the Unconstrained Least-Squares Importance Fitting (uLSIF) method \cite{kanamori2009least} and an extension which possesses a superior nonparametric convergence property: Relative uLSIF (RuLSIF) \cite{yamada2013relative}.

### 2014-01-13 Chapters/Chapter2/statistical_models.tex
Feedback Chapter 2 (not everything

@@
 Many applications require the detection of time points at which the underlying properties of a system change.
-Often this problem is formulated in a statistical framework, by inspecting the underlying data generating \gls{pdf} of the time series data.
+Often this problem is formulated in a statistical framework, by inspecting the \gls{pdf} of the time series data at specific time windows.
 A change point is then defined as a significant change in the properties of the \gls{pdf}, such as the mean and variance.
@@
 The widely used \gls{cusum} method by Basseville \etal \cite{basseville1993detection} follows this approach.
-It originates from control methods for detection from benchmarks.
+It originates from quality control and benchmarking methods for manufacturing processes.
 This method, and some derivatives, are discussed and analyzed in \Cref{subsec:literature_review_temporal_segmentation_cusum}.
 
-Many methods rely on pre-specified parametric model assumptions and considers the data be independent over time, which makes it less flexible to real-world applications.
-The methods proposed by Kawahara \etal \cite{kawahara2009change} and Lui \etal \cite{liu2013change} try to overcome these problems by estimating the \emph{ratio} between the \glspl{pdf}, instead of estimating each \gls{pdf}.
+Many methods rely on pre-specified parametric model assumptions and consider the data to be independent over time, which makes it less flexible to real-world applications.
+The methods proposed by Kawahara \etal \cite{kawahara2009change} and Lui \etal \cite{liu2013change} try to overcome these problems by estimating the \emph{ratio} between the \glspl{pdf} of time windows, instead of estimating each \gls{pdf}.
 This approach is discussed and analyzed in \Cref{sec:density-ratio}.
@@
 One problem it tries to overcome is the (claimed) weakness of many methods to detect a decrease in variance.
-The method represents the distribution over the data points as a hypersphere in a higher dimension using the kernel trick.
+The method represents the distribution over the data points as a hypersphere in a higher dimension.
 A change in the \gls{pdf} is represented by a change in the radius of this sphere.
@@
 
-In the search for the change in properties, temporal segmentation and change point detection methods can roughly be categorized in three methods in the way the data is processed, as discussed by Avci \etal \cite{avci2010activity}:
-\begin{itemize}
-  \item \textbf{Top-Down} methods iteratively divide the signal in segments by splitting at the best location.
+In the search for the change in properties, temporal segmentation and change point detection methods can roughly be categorized in four methods in the way the data is processed, as discussed by Avci \etal \cite{avci2010activity}:
+\begin{enumerate}
+  \item \textbf{Top-Down} methods iteratively divide the signal in segments by splitting at the most optimal location.
   The algorithm starts with two segments and completes when a certain condition is met, such as when an error value or number of segments $k$ is reached.
@@
   They start with creating $n/2$ segments and iteratively join adjacent segments while the value of a cost function for that operation is below a certain value.
-  \item \textbf{Sliding-window} methods are simple and intuitive for online segmenting purposes.
+  \item \textbf{Sliding Window} methods are simple and intuitive for online segmentation.
   It starts with a small initial subsequence of the time series.
-  New data points are joined in the segment until the fit-error is above a threshold.
-  Since the data is only processed very locally, these methods can yield in poor results \cite{keogh2001online}.
-  \item \textbf{\acrlong{swab}}, as introduced by Keogh \etal \cite{keogh2001online}, combines the ability of the sliding window mechanism to process time series online and the bottom-up approach the create superior segments in terms of fit-error.
+  New data points are added to the current segment until the fit-error is above a threshold.
+  % Since the data is only processed very locally, these methods can yield in poor results \cite{keogh2001online}.
+  \item \textbf{\acrlong{swab}}, as introduced by Keogh \etal \cite{keogh2001online}, combines the ability of the sliding window mechanism to process time series online and the bottom-up approach to create superior segments in terms of fit-error.
   The algorithm processes the data in two stages.
   The first stage is to join new data points in the current segment created by a sliding window.
-  The second stage processes the data using Bottom-Up and return the first segment as the final result.
-  Since this second stage retains some (semi-)global view of the data, the results are comparative with normal Bottom-Up.
-\end{itemize}
-It is clear that for the application of this research sliding-window and preferably \gls{swab}-based algorithms should be considered.
-In the following we will dicuss classes of algorithms grouped by the type of decision function, assuming a \gls{swab}-based data processing order.
+  The second stage processes the data using Bottom-Up and returns the first segment as the final result.
+  Since this second stage retains some (semi-)global view of the data, the results are comparative with normal Bottom-Up while being able to process the data in an online manner.
+\end{enumerate}
+For the application of this research Sliding Window and preferably \gls{swab}-based algorithms will be considered.
+In the following sections we discuss classes of algorithms grouped by the type of decision function, assuming a \gls{swab}-based data processing order.
 
 % CUSUM
-A widely used class of methods is based on the \gls{cusum} approach by Basseville \etal \cite{basseville1993detection}.
-This approach originates from control methods for detection and faults and benchmarking.
-In \Cref{subsec:literature_review_temporal_segmentation_cusum} a further overview of this class of methods is discussed.
+% A widely used class of methods is based on the \gls{cusum} approach by Basseville \etal \cite{basseville1993detection}.
+% This approach originates from control methods for detection and faults and benchmarking.
+% In \Cref{subsec:literature_review_temporal_segmentation_cusum} a further overview of this class of methods is discussed.

### 2014-01-13 Chapters/Chapter2/svm.tex
Feedback Chapter 2 (not everything

@@
 
-Many proposals in the field of remote sensing and \gls{har} make use of \glspl{svm} as a (supervised) model learning method.
+Many proposals in the field of \gls{har} make use of \glspl{svm} as a (supervised) model learning method.
 An elaborate overview of applications is given in \cite{mountrakis2011support}.
@@
 Sch{\"o}lkopf \etal \cite{scholkopf1999support} applies Vapnik's principle, to never solve a problem which is more general than the one that one is actually interested in, to novelty detection.
-In the case of novelty detection, they argue there is no need for density estimation of the distribution.
+In the case of novelty detection, they argue there is no need for a full density estimation of the distribution.
 Simple algorithms estimate the density by considering how many data points fall in a region of interest.
@@
 The \gls{nu-svm} method adds the kernel method, allowing non-linear decision functions, and incorporates `softness' by the $\nu$-parameter.
-Whereas the method in \cite{vapnik1963pattern} focuses on two-class problems, the \gls{nu-svm} method introduces the method the one-class problems.
+Whereas the method in \cite{vapnik1963pattern} focuses on two-class problems, the \gls{nu-svm} method solves one-class problems.
 
 The method introduced by Ma and Perkins \cite{ma2003time} creates a \emph{projected phase} of a time series data, which is intuitively equal to applying a high-pass filter to the time series.
-The projected phase of the time series combines a history or data points to a vector, which are then classified by the \gls{nu-svm} method.
+The projected phase of the time series combines a history of data points to a vector, which are then classified by the \gls{nu-svm} method.
 The method is applied to a simple synthetic sinusoidal signal with small additional noise and a small segment with large additional noise.
@@
 The method exploits fixed-point arithmetic to create a hardware-friendly algorithm, which can be executed on a smartphone.\footnote{Smartphones have limited resources and thus require energy-efficient algorithms.
-In \cite{anguita2007hardware} a Hardware-Friendly \gls{svm} is introduced.
-It uses less memory, processor time and power consumption, with a loss of precision.}
+Anguita \etal \cite{anguita2007hardware} introduce a Hardware-Friendly \gls{svm}.
+It uses less memory, processor time, and power consumption, with a loss of precision.}
 
@@
 The \gls{svcpd} method of Camci \cite{camci2010change} uses \gls{svdd} applies it to time-series data.
-By using an \gls{oc-svm} the method is able to detect changes in mean variance.
-Especially the detection of variance \emph{decrease} is an improvement over other methods that rely in the \gls{kl} divergence, since these are unable to detect decreases in variance \cite{takeuchi2006unifying}.
+By using an \gls{oc-svm} the method is able to detect changes in mean and variance.
+Especially the detection of variance \emph{decrease} is an improvement over other methods that
+%rely in the \gls{kl} divergence, since these
+are unable to detect decreases in variance \cite{takeuchi2006unifying}.
 This main research subject of this thesis is to apply the method of Camci \cite{camci2010change} to sensor data (such as accelerometer time series) obtained by on-body smartphones.

### 2014-01-13 Chapters/Chapter2/temporal_segmentation.tex
Feedback Chapter 2 (not everything

@@
 Methods that apply temporal segmentation on time series data can be roughly categorized in three different methods, as discussed by Avci \etal \cite{avci2010activity}.
-In that survey the distinction between \emph{Top-Down}, \emph{Bottom-Up}, and \emph{Sliding-Window} approaches is made, based on the way data is processed by the algorithm.
+In that survey the distinction between \emph{Top-Down}, \emph{Bottom-Up}, and \emph{Sliding-Window} approaches is based on the way data is processed by the algorithm.
 Since we are interested in on-line change detection, the literature discussed in this section will mainly be forms of sliding-window algorithms.
@@
 The research also introduces the \gls{swab} method, which combines the simple and intuitive approach of the sliding-window approach with the (semi-)global view over the data from the bottom-up method.
-This proposed method is dependent on user provided parameters, setting the maximum error when performing the algorithm.
-Each segment is approximated by using a \gls{plr}.
-The provided error threshold controls the granularity and number of segments.
+% This proposed method depends on user provided parameters, setting the maximum error when performing the algorithm.
+Each segment is approximated by a \gls{plr}, within a certain error.
+The user-provided error threshold controls the granularity and number of segments.
 
-Other methods have been proposed, such as an adaptive threshold based on the signal energy by Guenterberg \etal \cite{guenterberg2009automatic}, the adaptive \gls{cusum}-based test by Alippi \etal \cite{alippi2006adaptive} and the \gls{mosum} by Hsu \cite{hsu2007mosum} in order to eliminate this user-dependency.
+Other methods have been proposed, such as an adaptive threshold based on the signal energy by Guenterberg \etal \cite{guenterberg2009automatic}, the adaptive \gls{cusum}-based test by Alippi \etal \cite{alippi2006adaptive} and the \gls{mosum} by Hsu \cite{hsu2007mosum} in order to eliminate this user dependency.
 The latter of these methods is able to process the accelerometer values directly, although better results are obtained when features of the signal are processed, as done in the first method \cite{guenterberg2009automatic}.
@@
 By extending the \gls{swab} method, for every frame a feature set is selected, using an enhanced version of \gls{pca}.
-The research also considered the (Stepwise) Feasable Space Window as introduced by \cite{liu2008novel}, but since it results in a higher error rate than \gls{swab} the latter was chosen to extend.
+The research also considered the (Stepwise) Feasable Space Window as introduced by \cite{liu2008novel}, but since it results in a higher error rate than \gls{swab}, the latter was chosen to extend.
 Whereas the above mentioned algorithms use a linear representation, this methods considers linear, quadratic and cubical representations for each segment.
-This differs from other methods where the model is fixed for the whole time series, such as \cite{fuchs2010online}, which is stated to perform inferior on non-stationary time series such as daily life.
+This differs from other methods where the model is fixed for the whole time series, such as \cite{fuchs2010online}, which is stated to perform inferior on non-stationary time series, such as those from daily life.
 
 The time series data from a sensor can be considered as being drawn from a certain stochastic process.
-Probabilistic models can be constructed on that signal, yielding in probabilistic and Bayesian based segmentation methods.
-The \gls{cusum}-methods follows a statistical approach and relies on the log-likelihood ratio \cite{gustafsson1996marginalized} to measure the difference between two distributions.
+% Probabilistic models can be constructed on that signal, yielding in probabilistic and Bayesian based segmentation methods.
+The \gls{cusum}-methods instead follows a statistical approach and relies on the log-likelihood ratio \cite{gustafsson1996marginalized} to measure the difference between two distributions.
 To calculate the ratio, the probability density functions need to be calculated.
@@
 The method by Liu \etal \cite{liu2013change} uses a comparable dissimilarity measure using the \gls{kliep} algorithm.
-They claim this results in a robuster approach for real-world scenarios.
-Although this is a model-based method, no strong assumptions (parameter settings) are made on the models.
+They claim this results in a robuster method for real-world scenarios.
+Although this is a model-based method, no strong assumptions (parameter settings) are made about the models.
 
 The method of Adams and MacKay \cite{adams2007bayesian} builds a probabilistic model on the segment run length, given the observed data so far.
-Instead of modeling the values of the data points as a probabilistic distribution, the length of segments as a function of time is modeled by calculating the posterior probability.
+Instead of modeling the values of the data points, the length of segments as a function of time is modeled by calculating its posterior probability.
 It uses a prior estimate for the run length and a predictive distribution for newly-observed data, given the data since the last change point.
-This method contrasts with the approach of Guralnik and Srivastava \cite{guralnik1999event} in which change points are detected by a change in the (parameters of an) underlying, observed, model.
-For each new data point, the likelihoods of being a change point and part of the current segment are calculated, without a prior model (and thus is a non-Bayesian approach).
+This method differs from the approach of Guralnik and Srivastava \cite{guralnik1999event} in which change points are detected by a change in the (parameters of an) fitted model.
+For each new data point, the likelihoods of being a change point and part of the current segment are calculated, without a prior model (and thus this is a non-Bayesian approach).
 It is observed that when no change point is detected for a long period of time, the computational complexity increases significantly.
@@
 With the observation that more dimensions are needed to keep the error below the threshold when transitions between actions occur, cut-points can be located and segments will be created.
-The superior extension of their approach uses a Probabilistic \gls{pca} algorithm to incorporate the dimensions outside the selected set as noise.
+The superior extension of their approach uses a Probabilistic \gls{pca} algorithm to model the data from dimensions outside the selected set with noise.
 
-In the method by Himberg \etal \cite{himberg2001time} a cost function is defined over segments of data which is to be minimized.
-The cost functions thereby searches for internally homogeneous segments of data, reflecting states in which the devices and the user are.
+In the method by Himberg \etal \cite{himberg2001time} a cost function is defined over segments of data.
+By minimizing the cost function it creates internally homogeneous segments of data.
+A segment reflects a state in which the devices, and eventually its user, are.
 The cost function can be any arbitrary function and in the implementation the sum of variances over the segments is used.
@@
 
-As stated before, oten methods obtain an implicit segmentation as a result of classification over a sliding window.
+As stated before, often methods obtain an implicit segmentation as a result of classification over a sliding window.
 The method of Yang \etal \cite{yang2008distributed} explicitly performs segmentation and classification simultaneously.
@@
 Originally used for quality control in production environments, its main function is to detect change in the mean of measurements \cite{basseville1993detection}.
-It is a non-Bayesian method and thus makes no assumptions, in the form of prior belief distributions, for the change points.
+It is a non-Bayesian method and thus has not explicit prior belief for the change points.
 Many extensions to this method have been proposed.
@@
   \end{inparaenum}
-As we will see in \Cref{sec:one_class_classification}, in this research we are interested in methods from Type 3, from which \gls{occ} is a member.
-More theoretical background and foundation on outliers in statistical data can be found in \cite{barnett1994outliers}.
+As we will see in \Cref{sec:one_class_classification}, in this research we are interested in methods from Type 3, of which \gls{occ} is a member.
+More theoretical background on outliers in statistical data is provided in the work by Barnett and Lewis~\cite{barnett1994outliers}.

### 2014-01-13 Chapters/Chapter3/change_detection_time_series.tex
WIP: feedback chapter 3

@@
 % !TEX root = ../../main.tex
-\section{Problem Formulation Change Detection}\label{sec:change_detection_time_series}
+\section{Problem Formulation of Change Detection}\label{sec:change_detection_time_series}
 % \TODO{Write section introduction} \\
@@
 In other words, slowly changing properties in the data are not considered to be changes.
-Considered a time series $x_1 x_1 \dots$, which is drawn from a stochastic process $p$.
-Each $x_t$ (t = 1, 2, \dots) is a $d$-dimensional real valued vector and $p$ a probability density function of the sequence $x_1 x_2 \dots$.
-Assume $p$ can be decomposed in two different \gls{iid} stationary stochastic processes $p^1$ and $p^2$ and are one-dimensional Gaussian density functions.
-For a time point $a$ data points for which $t < a$ are drawn from $p^1 = N(\mu_1, \sigma_1^2)$ and for $t \geq a$ from $p^2 = N(\mu_2, \sigma_a^2)$.
-If $p^1$ and $p^2$ are different, then the time point $t = a$ is a \emph{change point}.
-In \cite{takeuchi2006unifying} the similarity between the stochastic processes are expressed by the \gls{kl} divergence $D(p^2||p^1)$.
-It is observed that this measure is not able to detect a change by decrease in variance \cite{takeuchi2006unifying,camci2010change}.
+Considered a time series $x_1, x_2, \dots$, which is drawn from a stochastic process $p$.
+Each $x_t$ (t = 1, 2, \dots) is a $d$-dimensional real valued vector.
+Assume $p$ can be ``decomposed'' in two different \gls{iid} stationary one-dimensional Gaussian processes $p^{(1)}$ and $p^{(2)}$.
+For a time point $a$ data points for which $t < a$ are drawn from $p^{(1)} = N(\mu_1, \sigma_1^2)$ and for $t \geq a$ from $p^{(2)} = N(\mu_2, \sigma_2^2)$.
+If $p^{(1)}$ and $p^{(2)}$ are different, then the time point $t = a$ is a \emph{change point}.
+Takeuchi and Yamanishi~\cite{takeuchi2006unifying} expess the similarity between the stochastic processes with the \gls{kl} divergence $D(p^2||p^1)$.
+It is observed that their method is not able to detect a change by decrease in variance \cite{takeuchi2006unifying,camci2010change}.
 This problem is the motivation for Camci \cite{camci2010change} to create the \gls{svcpd} algorithm.

### 2014-01-13 Chapters/Chapter3/one_class_classification.tex
WIP: feedback chapter 3

@@
 \subsection{Problem formulation}\label{subsec:occ-problem-formulation}
-The problem of \gls{occ} is closely related to the (traditional) two-class classification situation\footnote{Two-class problems are considered as to be the basic problem, since multi-class classification problems can be decomposed into multiple two-class problems \cite{fukunaga1990introduction}.}.
-In the case of traditional classification algorithms, the problem is to classify an unknown object to one of the pre-defined categories.
-Every object $i$ is represented as a vector $\vectorsym{x}_i = (x_{i,1},\dots,x_{i,d}), x_{i,j} \in \mathbb{R}$ with $d$ real-valued measurements.
-Using this notation, an object $\vectorsym{x}_i$ thus represents one point in a feature space: $\vectorsym{x}_i \in \mathcal{X} = \mathbb{R}^d$.
+The problem of \acrlong{occ} is closely related to the (traditional) two-class classification situation\footnote{Two-class problems are considered the basic problem, since multi-class classification problems can be decomposed into multiple two-class problems \cite{fukunaga1990introduction}.}.
+In the case of traditional classification algorithms, the problem is to assign an unknown object to one of the pre-defined categories.
+Every object $i$ is represented as a d-dimensional vector $\vectorsym{x}_i = (x_{i,1},\dots,x_{i,d}), x_{i,j} \in \mathbb{R}$.
+Using this notation, an object $\vectorsym{x}_i$ thus represents one point in a feature space $\mathcal{X} \in \mathbb{R}^d$.
 The two classes of objects, $\omega_1$ and $\omega_2$, are labeled $-1$ and $+1$ respectively.
-The objects with a label $y_i \in \{-1, +1\}$ attached are in the training set $\mathcal{X}^{tr}$ (note that it can be both positive and negative example objects).
+The objects with a label $y_i \in \{-1, +1\}$ are in the training set (note that it can be both positive and negative example objects).
 This problem is solved by determining a decision boundary in the feature space of the data objects and label the new data object based on the location relative to this boundary.
@@
 In case of the \gls{occ} problem, only one class (often referred as the target class, or positive examples) of training data is used to create a decision boundary.
-The goal is to determine whether a new data object belongs to the target class or to the unknown class and thus is an outlier.
+The goal is to determine whether a new data object belongs to the target class.
+If it does not belong to the class it is an outlier.
 One could argue that this problem is equal to the traditional two-class problem by considering all other classes as negative examples, although there are important differences.
@@
 It can be a goal by itself and can be used as a filtering mechanism for other data processing methods.
-Often methods that rely on data-characteristics are sensitive for remote regions in the data set.
+Often methods that rely on data characteristics are sensitive for remote regions in the data set.
 Using \gls{occ} these remote regions can be removed from the data set.
@@
 By constructing a \gls{occ}-classifier using a training data set, one can compare that set to a new data set.
-This will result in a similarity-measure expressed in the properties of outliers.
-This measure is related to other similarity measures, such as \gls{pdf} estimation and the \gls{kl} divergence as discussed in \Cref{sec:change_detection_time_series}.
+This will result in a similarity measure expressed in the properties of outliers.
+This is related to other methods of expressing similarity, such as density-ratio estimation and the \gls{kl} divergence as discussed in \Cref{sec:change_detection_time_series}.
 
@@
 This can cause a large bias on the data.
-When a good probability model is assumed, these methods work very well, since when one threshold is optimized, a minimum volume is automatically found for the given probability density model \cite{tax2001one}.
+When a good probability model is postulated, these methods work very well.
+%, since when one threshold is optimized, a minimum volume is automatically found for the given probability density model \cite{tax2001one}.
 
-Boundary methods are based on Vapnik's principle\footnote{With a limited amount of data available, one should avoid solving a more general problem as an intermediate step to solve the original problem \cite{vapnik1998statistical}.} which imply in this case that estimating the complete data density for a \gls{occ} may be too complex, in case one is only interested in the closed boundary.
-Examples of methods that focus on the boundary (a direct threshold) of the training data distribution are K-centers, Nearest-neighborhood and \gls{svdd} or a combination of those methods \cite{hempstalk2008one}.
-Especially the latter has a strong bias towards minimal volume solutions.
-These type of methods are sensitive to scaling of features, since they rely on a well-defined distance measure.
-The number of objects that is required is smaller than in the case of density methods.
+Boundary methods are based on Vapnik's principle\footnote{With a limited amount of data available, one should avoid solving a more general problem as an intermediate step to solve the original problem \cite{vapnik1998statistical}.} which imply in this case that estimating the complete data density for a \gls{occ} may be too complex, if one is only interested in the closed boundary.
+Examples of methods that focus on the boundary (a direct threshold) of the training data distribution are K-centers, Nearest-neighborhood, and \gls{svdd}, or a combination of those methods \cite{hempstalk2008one}.
+Especially \gls{svdd} has a strong bias towards minimal volume solutions.
+These type of methods are sensitive to the scale and range of features, since they rely on a well-defined distance measure.
+The number of objects that is required, is smaller than in the case of density methods.
 The boundary method \gls{svdd}, constructed by Tax and which has shown good performance \cite{khan2010survey}, will be further discussed in \Cref{subsec:oc-svm-svdd}.
@@
 This results in a compact representation of the target data and for any object a reconstruction error can be defined.
-Outliers are data objects with a high reconstruction error, since they are worse represented by the constructed model.
+Outliers are data objects with a high error, since they are not well reconstructed by the model.
 Examples of reconstruction methods are K-means, \gls{pca} and different kind of neural network implementations.
@@
     \includegraphics[width=0.5\textwidth,keepaspectratio]{./Figures/chapter3/occ_methods.pdf}
-  \caption[\gls{occ} methods]{Overview of \gls{occ} methods categorized in Density Estimators, Reconstruction methods and Boundary methods.
+  \caption[\gls{occ} methods]{Overview of \gls{occ} methods categorized in Density Estimation, Reconstruction methods, and Boundary methods.
 This categorization follows the definition of Tax in \cite{tax2001one}.}

### 2014-01-13 Chapters/Chapter3/one_class_svm.tex
WIP: feedback chapter 3

@@
 The classical implementation of a \gls{svm} is to classify a dataset in two distinct classes.
-This is a common usecase, although sometimes there is no training data for both classes available.
+This is a common use case, although sometimes there is no training data for both classes available.
 Still, one would like to classify new data points as regular, in-class, or out-of-class, \eg in the case of a novelty detection.
-With that problem only data examples from one class is available and the objective is to recognize new data points that are not part of that class.
+With that problem only data examples from one class are available and the objective is to recognize new data points that are not part of that class.
 This unsupervised learning problem is closely related to density estimation.
 In that context, the problem can be the following.
-Given an underlying probabilty distribution $P$ the goal is to find a subset $S$ for which the probability that a data point from $P$ lies outside $S$ equals some predetermined value $\nu$ between $0$ and $1$ \cite{scholkopf1999support}.
+Assume an underlying probability distribution $P$ and a data point $\vectorsym{x}$ drawn from this distibution.
+The goal is to find a subset $S$ of the input space, such that the probability that $\vectorsym{x}$ lies outside of $S$ equals some predetermined value $\nu$ between $0$ and $1$ \cite{scholkopf1999support}.
 
@@
 We will first discuss the traditional two-class \gls{svm} before we consider the one-class variant, as introduced by Cortes and Vapnik in \cite{cortes1995support}.
-Consider a data set $\Omega = \{ (x_1, y_1),\allowbreak (x_2, y_2), \dots , (x_n, y_n) \}$; points $x_i \in \mathbb{R}^d$ in a (for instance two-dimensional) space where $x_i$ is the $i$-th input data point and $y_i \in \{-1, 1\}$ is the $i$-th output pattern, indicating the class membership.
+Consider a labeled data set $\Omega = \{ (x_1, y_1),\allowbreak (x_2, y_2), \dots , (x_n, y_n) \}$; points $x_i \in \mathbb{I}$ in a (for instance two-dimensional) space where $x_i$ is the $i$-th input data point and $y_i \in \{-1, 1\}$ is the $i$-th output value, indicating the class membership.
 
-A \gls{svm} can create a boundary between linear-separable data points, making it a non-probabilistic binary linear classifier.
+A \gls{svm} can create a boundary between linear-separable data points, making it a binary linear classifier.
 More flexible non-linear boundaries can be obtained by the use of a non-linear function $\phi(x)$, as illustrated in \Cref{fig:kernel_mapping}.
@@
 The \gls{svm} can create a linear separating hyperplane in the space $\mathcal{F}$ that separates the data points from the two classes.
-When the hyperplane is projected to the (lower) original input space $\mathcal{I}$ it creates a non-linear separating curve.
+When the hyperplane is projected on to the (lower) original input space $\mathcal{I}$ it creates a non-linear separating curve.
 A schematic overview of this process is illustrated in \Cref{fig:svm_mapping_spaces}.
-The original data points $x \in I$ are through $\phi(x)$ projected to the points $z \in F$.
+The original data points $x \in \mathcal{I}$ are projected by $\phi(x)$ to the points $z \in \mathcal{F}$.
 In that space the linear hyperplane is constructed by weights $w$.
-This results in a non-linear hyperplane $y$ in input space $I$.
+This results in a non-linear hyperplane $y$ in input space $\mathcal{I}$.
 The mapping and projection of data points can be efficient (and implicit) performed by using the kernel trick, which is discussed in \Cref{subsec:kernels}.
@@
   \includegraphics[width=0.8\textwidth]{./Figures/chapter3/svm_kernel_mapping.png}
-  \caption[Kernel mapping]{The non-linear boundary in the input space $\mathcal{I}$ (left) is transformed to a linear boundary in the feature space $\mathcal{F}$ (right) by mapping the data points with the function $\phi$.
-The kernel trick uses a function $K$ which performs an implicit mapping.
-Image from Wikipedia.org}
+  \caption[Kernel mapping]{Visualization of a \gls{svm} in 2D.
+  The non-linear boundary in the input space $\mathcal{I}$ (left) is transformed to a linear boundary in the feature space $\mathcal{F}$ (right) by mapping the data points with the function $\phi$.
+  The kernel trick uses a function $K$ which performs an implicit mapping. Image from Wikipedia.org}
   \label{fig:kernel_mapping}
@@
   \includegraphics[width=0.8\textwidth]{./Figures/chapter3/svm_mapping_spaces.pdf}
-  \caption[Mapping spaces in \gls{svm}]{The input vector $\vectorsym{x}$ from input space $\mathcal{I}$ is mapped by a non-linear function $\phi(\vectorsym{x})$ to a feature vector $\vectorsym{z}$ in the high dimensional feature space $\mathcal{F}$.
+  \caption[Mapping spaces in \gls{svm}]{Schematic visualization of the non-linear mapping of \glspl{svm}.
+  The input vector $\vectorsym{x}$ from input space $\mathcal{I}$ is mapped by a non-linear function $\phi(\vectorsym{x})$ to a feature vector $\vectorsym{z}$ in the high dimensional feature space $\mathcal{F}$.
   The weights of $\vectorsym{w}$ create a linear separating hypeplace, which maps the high dimensional vector to the predicted outcome $\vectorsym{y}$.}
@@
 \end{equation}
-with $w \in F$ and $b \in R$.
-The hyperplane that is created determines the \emph{margin} between the classes; the minimal distance from one of the data points to the hyperplane.
-In geometric sense, $w$ is the normal vector indication the direction of the hyperplane and $\frac{b}{\lVert{w}\rVert}$ determines the offset of the hyperplane to the origin.
-Since the distance between the two margins is equal to $\frac{2}{\lVert{w}\rVert}$, the maximum-margin hyperplane is found by minimizing $\lVert{w}\rVert$.
-The data points which lie on the margin are the \emph{support vectors}.
+with $w \in \mathcal{F}$ and $b \in \mathbb{R}$.
+The hyperplane that is created determines the \emph{margin} between the classes; the minimal distance from the data points to the hyperplane.
+In geometric sense, $w$ is the normal vector indicating the direction of the hyperplane and $\frac{b}{\lVert{w}\rVert}$ determines the offset of the hyperplane from the origin.
+Since the size of the margin is equal to $\frac{2}{\lVert{w}\rVert}$, the maximum-margin hyperplane is found by minimizing $\lVert{w}\rVert$.
+The data points which lie on the boundary of the margin are the \emph{support vectors}.
 This geometrical interpretation is illustrated in \Cref{fig:svm_hyperplane}.
 All data points for which $y_i = -1$ are on one side of the hyperplane and all other data points (for which $y_i = 1$) are on the other side.
-The minimal distance from a data point to the hyperplane is for both classes equal.
-This results in a \emph{maximal margin} between the two classes.
+The minimal distance from a data point to the hyperplane is equal for both classes.
+Minimizing $\lVert{w}\rVert$ results in a \emph{maximal margin} between the two classes.
 Thus, the \gls{svm} searches for a maximal separating hyperplane.
@@
   \caption[\gls{svm} and the separating hyperplane]{Illustration of the separating hyperplane of a \gls{svm}.
-Here $w$ is the normal vector for the separating hyperplane and the distance between the two margins is $\frac{2}{\lVert{w}\rVert}$.
+  Here $w$ is the normal vector for the separating hyperplane and size of the margin is $\frac{2}{\lVert{w}\rVert}$.
   Image from Wikipedia.org}
@@
 The \gls{svm} classifier can use a \emph{soft margin} by allowing some data points to lie within the margin, instead of on the margin or farther away from the hyperplane.
-For this it introduces \emph{slack variables} $\xi_i$ for each data point and the constant $C > 0$ the determines the trade-off between maximizing the margin and the number of data points within that margin (and thus the training errors).
-The slack variables minimize the \emph{sum of deviations} rather than the \emph{number} of incorrect data points \cite{cherkassky2007learning}.
+For this it introduces \emph{slack variables} $\xi_i$ for each data point and the constant $C > 0$ determines the trade-off between maximizing the margin and the number of data points within that margin (and thus the training errors).
+The slack variables are a generalization to minimize the \emph{sum of deviations}, rather than the \emph{number} of incorrect data points \cite{cherkassky2007learning}\footnote{if $\xi_i$ is chosen to be an indicator function, it still minimizes the \emph{number} of incorrect data points.}.
 The objective function for a \gls{svm} is the following minimization function:
@@
 In the dual formulation the problem scales with the number of training examples $n$ instead of the dimensionality $d$ of the samples.
-Solving this problem directly in the high dimensional feature space $\mathcal{F}$ makes it untractable.
+Solving this problem directly in the high dimensional feature space $\mathcal{F}$ makes it intractable.
 The linear approximation function corresponds to the kernel function in the dual formulation.
@@
 
-Using this formulation two imporant properties arise \cite{flach2012machine}:
+Using this formulation two important properties arise \cite{flach2012machine}:
 \begin{enumerate}
-  \item Searching for the maximum margin decision boundary is equivalent to searching for the support vectors; they are the training examples with non-zero Lagrange mulitpliers.
+  \item Searching for the maximum margin decision boundary is equivalent to searching for the support vectors; they are the training examples with non-zero Lagrange multipliers.
   \item The optimization problem is entirely defined by pairwise dot products between training examples: the entries of the kernel matrix $K$.
 \end{enumerate}
-An effect of the first property, combined with the equality to sparse represenations, is that \glspl{svm} often have good results, even in the case of high dimensional data (also known as the curse of dimensionality) or limited training examples \cite{cherkassky2007learning}.
-The second property is what enables powerful adaption for \glspl{svm} to learn non-linear decision boundaries.
-The workings of the kernel matrix $K$ and the non-linear boundaries is discussed in the following section.
+An effect of the first property, combined with the equality to sparse representations, is that \glspl{svm} often have good results, even in the case of high dimensional data or limited training examples \cite{cherkassky2007learning}.
+The second property is what enables an powerful adaptation of \glspl{svm} to learn non-linear decision boundaries.
+The workings of the kernel matrix $K$ and the non-linear boundaries are discussed in the following section.
 
@@
 The decision function in \Cref{eq:svm_lagrange} only relies on the dot products of mapped data points in the feature space $\mathcal{F}$ (\ie all pairwise distances between the data points in that space).
-It shows \cite{flach2012machine} that as long as any function has the same result, without an explicit mapping to the higher dimension $\mathcal{F}$, the dot products can be substituted by the kernel function $K$, as introduced by Aizerman~\etal \cite{aizerman1964theoretical} and applied to \glspl{svc} by Vapnik \cite{vapnik1998statistical}:
+It shows \cite{flach2012machine} that for any function $K$ that has the same result in feature space $\mathcal{F}$, without an explicit mapping to the higher dimension $\mathcal{F}$, the dot products can be substituted by the kernel function $K$, as introduced by Aizerman~\etal \cite{aizerman1964theoretical} and applied to \glspl{svc} by Vapnik \cite{vapnik1998statistical}:
 \begin{equation}

### 2014-01-14 Chapters/Chapter3/one_class_svm.tex
More feedback Chapter 3. Still lefT

@@
 In that context, the problem can be the following.
-Assume an underlying probability distribution $P$ and a data point $\vectorsym{x}$ drawn from this distibution.
-The goal is to find a subset $S$ of the input space, such that the probability that $\vectorsym{x}$ lies outside of $S$ equals some predetermined value $\nu$ between $0$ and $1$ \cite{scholkopf1999support}.
+Assume an underlying probability distribution $P$ and a data point $\vectorsym{x}$ drawn from this distribution.
+The goal is to find a subset $S$ of the input space, such that the probability that $\vectorsym{x}$ lies inside of $S$ equals some predetermined value $\nu$ between $0$ and $1$ \cite{scholkopf1999support}.
 
@@
 That is followed by two different implementations of \gls{oc-svm}: \gls{nu-svm} in \Cref{subsec:nu-svm} by Sch\"olkopf \etal \cite{scholkopf1999support}, which closely follows the above problem statement regarding density estimation, and \gls{svdd} by Tax and Duin \cite{tax1999support} in \Cref{subsec:oc-svm-svdd}.
-The final part of this section discusses the influende of model parameters on the performance of \glspl{svm}
+The final part of this section discusses the influence of model parameters on the performance of \glspl{svm}
 
@@
   The input vector $\vectorsym{x}$ from input space $\mathcal{I}$ is mapped by a non-linear function $\phi(\vectorsym{x})$ to a feature vector $\vectorsym{z}$ in the high dimensional feature space $\mathcal{F}$.
-  The weights of $\vectorsym{w}$ create a linear separating hypeplace, which maps the high dimensional vector to the predicted outcome $\vectorsym{y}$.}
+  The weights of $\vectorsym{w}$ create a linear separating hyperplane, which maps the high dimensional vector to the predicted outcome $\vectorsym{y}$.}
   \label{fig:svm_mapping_spaces}
@@
 \end{equation}
-where vectors $\vectorsym{z}$ and $\vectorsym{z}'$ are projections of data objects $\vectorsym{x}$ and $\vectorsym{x}'$ through $\phi(\vectorsym{x})$ on the features space $\mathcal{F}$ and the dot product kernel $K$ is determined by the sum
+where vectors $\vectorsym{z}$ and $\vectorsym{z}'$ are projections of data objects $\vectorsym{x}$ and $\vectorsym{x}'$ through $\phi(\vectorsym{x})$ on the features space $\mathcal{F}$.
+The dot product kernel $K$ is determined by the sum
 \begin{equation}
@@
 
-The number of basis functions, the center parameters that correspond with support vectors, and the weights in the ouput layer are all automatically determined via the optimal hyperplane in features space $\mathcal{F}$ \cite{cherkassky2007learning}.
+The number of basis functions, the center parameters that correspond with support vectors, and the weights in the output layer are all automatically determined via the optimal hyperplane in features space $\mathcal{F}$ \cite{cherkassky2007learning}.
 The width parameter $\sigma$ is equal for all basis functions and is set a priori and determines the flexibility and complexity of the boundary.
@@
 
-The mapping from input space $\mathcal{I}$ to $\mathcal{F}$ via $\phi(x)$ is subject to the continuity assumptions.
-This general assumption in pattern recogition state that two near objects in feature space should also resemble each other in real life \cite{tax2001one}.
+The mapping from input space $\mathcal{I}$ to $\mathcal{F}$ via $\phi(x)$ is subject to some continuity assumptions.
+This general assumption in pattern recognition, states that two near objects in feature space should also resemble each other in ``real life'' \cite{tax2001one}.
 Thus, objects which are close in feature space should be close in the original input space.
-When this assumption would not hold, the example objects would be scatter through the feature space and finding a decision function becomes very complex.
+When this assumption does not hold, the example objects would be scatter through the feature space and finding a decision function becomes very complex.
 
@@
 The first of the \gls{oc-svm} methods we will discuss is often referred to as \gls{nu-svm} and introduced by Sch\"olkopf \etal \cite{scholkopf1999support}.
-Instead of estimating the density function of an distribution $P$, it focuses on an easier problem: the algorithm find regions in the input where the ``probability density lives''.
+Instead of estimating the density function of an distribution $P$, it focuses on an easier problem: the algorithm finds regions in the input where the ``probability density lives''.
 This results in a function such that most of the data is in the region where the function is nonzero.
 
-The constructed decision function $\mathcal{F}$ resembles the function discussed in \Cref{subsec:svm}.
+The constructed decision function $f(x)$ resembles the function discussed in \Cref{subsec:svm}.
 It returns the value $+1$ in a (possibly small) region capturing most of the data points, and $-1$ elsewhere.
 The method maps the data points from input space $\mathcal{I}$ to a feature space $\mathcal{F}$ (following classical \glspl{svm}).
-In that space $\mathcal{F}$ it separates the data points with maximal margin from the origin, with a hyperplane.
-For a new data points $x$, the function value $f(x)$ determines wheter the data point is part of the distribution (\ie the value is $+1$) or a novelty (\ie the value is $-1$).
+In that space the data points are separated from the origin by a hyperplane, with maximal margin.
+Since the mapping to the feature space is based on dot products of the data points, outlier objects (which are dissimilar from the training set) will be closer to the origin.
+Thus, maximizing the distance from the hyperplane to the origin increases the discriminating ability of the decision function.
+Furthermore, it holds an intuitive relationship with the classical two-class \gls{svm}.
+
+For a new data points $x$, the function value $f(x)$ determines whether the data point is part of the distribution (\ie the value is $+1$) or a novelty (\ie the value is $-1$).
 The hyperplane is represented by $g(x) = w \cdot \phi(x) + \rho = 0$ and the decision function is $f(x) = \operatorname{sgn}(g(x))$.
@@
 This method creates a \emph{hyperplane}, characterized by $w$ and $\rho$, that separates the data with maximal margin from the origin in the feature space $\mathcal{F}$.
-In the following section we will discuss an alternative method, which uses an circumscribing \emph{hypersphere} to characterize the data (and the region of distribution $P$ where the density (or: support) lives).
+In the following section we will discuss an alternative method, which uses an circumscribing \emph{hypersphere} to characterize the training data.
+The region inside the hypersphere indicates the region $S$ where the probability that a data point drawn from $P$ is equal to $\nu$.
 
@@
 The constructed hypersphere is characterized by a center $\mathbf{a}$ and a radius $R > 0$ as distance from the center to (any data point that is a \gls{sv} on) the boundary, for which the volume, and thus the radius $R$, will be minimized.
-The center $\mathbf{a}$ is a linear combination of of the support vectors.
+The center $\mathbf{a}$ is a linear combination of the support vectors.
 Like the classical \gls{svm} and \gls{svdd} it can be required that all the distances from the data points $x_i$ to the center $\mathbf{a}$ are strict less than $R$ (or equivalent measure).
@@
 This principle is illustrated in the right image or \Cref{fig:nu-vs-svdd}.
-Instead of a separating hyperplane, constructed by \gls{nu-svm} and illustrated on the left of the Figure, the \gls{svdd} creates a hypersphere (in the illustration a cricle) around the data points.
+Instead of a separating hyperplane, constructed by \gls{nu-svm} and illustrated on the left of the Figure, the \gls{svdd} creates a hypersphere (in the illustration a circle) around the data points.
 By using kernel functions (\eg the \gls{rbf}) the hyperspheres in the high dimensional feature space $\mathcal{F}$ corresponds to a flexible and tight enclosing boundary in input space $\mathcal{I}$.
@@
 
-Because the Gaussian \gls{rbf} often yields good (\ie tight) boundaries, this set of kernels fuctions is commonly used:
+Because the Gaussian \gls{rbf} often yields good (\ie tight) boundaries, this set of kernels functions is commonly used:
 \begin{equation}
@@
 \subsection{SVM model parameters}\label{subsec:svm_model_parameters}
-\gls{svm}-model selectiong and tuning depends on two type of parameters \cite{cherkassky2007learning}:
+\gls{svm}-model selecting and tuning depends on two type of parameters \cite{cherkassky2007learning}:
 \begin{enumerate}
@@
 With a small value for the kernel width $\sigma$, each data point will tend to be used as a support vector (for almost all $\alpha_i > 0$) and the \gls{svdd} solution resembles a Parzen density estimation.
-For large values of $\sigma$, the solution will resemble the original hypersphere solution is obtained (in contrast with a tight boundary around the data).
-With a large value for the width $\sigma$, the boudary approximates the spherical boundary.
+For large values of $\sigma$, the solution will resemble the original hypersphere solution (in contrast with a tight boundary around the data).
+With a large value for the width $\sigma$, the boundary approximates the spherical boundary.
 The influence of the $\sigma$ parameter on the \gls{svdd} solution is illustrated in \Cref{fig:svdd-boundary-sigma}.
@@
 The following chapter will discuss our proposed method, which incorporates the \gls{svdd} algorithm.
-It relates the \gls{oc-svm} model constuction to outlier detection and eventually change detection, leading to finding a temporal segmentation of time series data.
+It relates the \gls{oc-svm} model construction to outlier detection and eventually change detection, leading to finding a temporal segmentation of time series data.
 

### 2014-01-14 Chapters/Chapter3/one_class_svm.tex
Feedback Chapter 4

@@
 \section{One-Class Support Vector Machine}\label{sec:one_class_svm}
-In this section we will discuss the details of a \gls{svm} and \gls{oc-svm} implementations.
-The classical implementation of a \gls{svm} is to classify a dataset in two distinct classes.
+In this section we will discuss the details of an \gls{svm} and \gls{oc-svm} implementations.
+The classical implementation of an \gls{svm} is to classify a dataset in two distinct classes.
 This is a common use case, although sometimes there is no training data for both classes available.
@@
 
-A \gls{svm} can create a boundary between linear-separable data points, making it a binary linear classifier.
+An \gls{svm} can create a boundary between linear-separable data points, making it a binary linear classifier.
 More flexible non-linear boundaries can be obtained by the use of a non-linear function $\phi(x)$, as illustrated in \Cref{fig:kernel_mapping}.
@@
   \includegraphics[width=0.8\textwidth]{./Figures/chapter3/svm_kernel_mapping.png}
-  \caption[Kernel mapping]{Visualization of a \gls{svm} in 2D.
+  \caption[Kernel mapping]{Visualization of an \gls{svm} in 2D.
   The non-linear boundary in the input space $\mathcal{I}$ (left) is transformed to a linear boundary in the feature space $\mathcal{F}$ (right) by mapping the data points with the function $\phi$.
@@
   \includegraphics[width=0.5\textwidth]{./Figures/chapter3/svm_separating_plane_with_margin.png}
-  \caption[\gls{svm} and the separating hyperplane]{Illustration of the separating hyperplane of a \gls{svm}.
+  \caption[\gls{svm} and the separating hyperplane]{Illustration of the separating hyperplane of an \gls{svm}.
   Here $w$ is the normal vector for the separating hyperplane and size of the margin is $\frac{2}{\lVert{w}\rVert}$.
@@
 The slack variables are a generalization to minimize the \emph{sum of deviations}, rather than the \emph{number} of incorrect data points \cite{cherkassky2007learning}\footnote{if $\xi_i$ is chosen to be an indicator function, it still minimizes the \emph{number} of incorrect data points.}.
-The objective function for a \gls{svm} is the following minimization function:
+The objective function for an \gls{svm} is the following minimization function:
 
@@
 The width parameter $\sigma$ is equal for all basis functions and is set a priori and determines the flexibility and complexity of the boundary.
-In \Cref{subsec:svm_model_parameters} this (hyper)parameter for a \gls{svm} is further discussed.
+In \Cref{subsec:svm_model_parameters} this (hyper)parameter for an \gls{svm} is further discussed.
 
@@
 This small fraction, for which $\alpha_i$ is non-zero, are called the \glspl{sv} and these objects lie on the boundary of the description.
-The center of the hypersphere only depends on this small number of \glspl{sv} and the objects for which $\alpha_i = 0$ can be disregarded from the solution.
+The center of the hypersphere only depends on this small number of \glspl{sv} and the objects for which $\alpha_i = 0$ can be discarded from the solution.
 Testing the membership of a (new) object $\vectorsym{z}$ is done by determining if the distance to the center $\vectorsym{a}$ of the sphere is equal or smaller to the radius $R$:

### 2014-01-14 Chapters/Chapter4/change_detection.tex
Put captions above tables
Feedback Chapter 4

@@
 This metric is used to discover changes in the underlying data generation process.
-We argued that in the case of the data objects set becomes more heterogeneous, we expect an increase in the radius $R$.
-In the same manner we expect that when the change is in the oldest data objects of the working set the radius $R$ will decrease, since the data becomes more homogeneous.
+We argued that in the case the data objects set becomes more heterogeneous, we expect an increase in the radius $R$.
+In the same manner we expect that when the change is in the oldest data objects of the working set, the radius $R$ will decrease, since the data becomes more homogeneous.
 An abstract representation of these expectations is illustrated in \Cref{fig:radius_expectation}.
@@
 In this section we discuss the methods applicable to the extracted radius $R$ of the hypersphere, in order to discover a change in the underlying distribution properties.
-This employs the second stage of our algorithm, analogue to the second stage of the unifying framework by Takeuchi and Yamanishi~\cite{takeuchi2006unifying}.
-It is also based on the \gls{svcpd} method by Camci~\cite{camci2010change}, although we have slightly simplified the algorithm.
+It is the second stage of our algorithm, analogue to the second stage of the unifying framework by Takeuchi and Yamanishi~\cite{takeuchi2006unifying}.
+It is also based on the \gls{svcpd} method by Camci~\cite{camci2010change}, although we have simplified the algorithm.
+
+\begin{table}
+\begin{center}
+  \caption[Proposed algorithm]{Algorithm of the proposed change detection method, using the Ratio-Thresholding method of \ref{subsec:ratio_thresholding} for change detection.}
+  \begin{tabular}{ l p{12cm} }
+    \hline
+    Step & Action \\
+    \hline
+    & \textbf{Input:} time series $\vectorsym{x}$, window size $n$ \\
+    & \textbf{Output:} collection of change points $\vectorsym{t}$ \\
+    1 & Start with $n$ data objects as working set and construct hypersphere \\
+    2 & Add next data object $x_t$ and drop first one \\
+    3 & Identify new hyper-sphere and its \emph{approximate radius} \\
+    4 & Calculate radius average of hyperspheres since last change point \\
+    5 & Calculate radius ratio $\hbar$. \newline
+    If $\hbar$ is  lower than ${th}_{low}$ or greater than ${th}_{high}$ then mark $t$ as change point in $\vectorsym{t}$ \\
+    6 & Continue from step 2, until no more data objects available \\
+    7 & Apply post-processing: merge close change points \\
+    \hline
+  \end{tabular}
+  \label{tab:algorithm_proposed_method}
+\end{center}
+\end{table}
+
+\TODO{Algoritme meer in pseudo-code zetten. Prominenter maken.}
 
 In the \gls{svcpd} the first step for a new data point $\vectorsym{z}$ is to check whether the data point lies within the enclosing (hyperspherical) boundary of the constructed model.
-This will label the new data point as an in- or outlier, realtive to the current data set.
+This will label the new data point as an in- or outlier, relative to the current data set.
 Depending on that outcome, the model is updated to include the new data point.
@@
 In our method we only employ the second step: the inspection of the radius $R$.
-We regard the first step (determine if $\vectorsym{z}$ is an outlier) to be superfluent: in case it is an outlier, the model update step will change the radius $R$ in order to let the boudary include the new data point $\vectorsym{z}$.
+We regard the first step (determine if $\vectorsym{z}$ is an outlier) to be superfluent: in case it is an outlier, the model update step will change the radius $R$ in order to let the boundary include the new data point $\vectorsym{z}$.
 Would the new data object $\vectorsym{z}$ be already an inlier, then the updated model would not have a significant change in radius $R$.
@@
 
-\begin{center}\begin{table}
-  \begin{tabular}{ l p{12cm} }
-    \hline
-    Step & Action \\
-    \hline
-    1 & Start with $n$ data objects as working set and construct hypersphere \\
-    2 & Add next data object $x_t$ and drop first one \\
-    3 & Identify new hyper-sphere and its \emph{approximate radius} \\
-    4 & calculate radius average of hyperspheres since last change point \\
-    5 & calculate radius ratio $\hbar$. If lower than ${th}_{low}$ or greater than ${th}_{high}$ then mark $t$ as change point \\
-    6 & continue from step 2, until no more data objects available \\
-    7 & Apply post-processing: merge close change points \\
-    \hline
-  \end{tabular}
-  \caption[Proposed algorithm]{Algorithm of the proposed change detection method, using the Ratio-Thresholding method of \ref{subsec:ratio_thresholding} for change detection.}
-  \label{tab:algorithm_proposed_method}
-\end{table}\end{center}
-
 \subsection{CUSUM based methods}
@@
 \begin{equation}
-  S_n = \sum_{k=1}^n r_k
+  S_n = \sum_{k=0}^n R_k
 \end{equation}
@@
 \begin{equation}
-  S_n - \operatorname*{min}_{0 \le i < n} S_i \ge h.
+  \left\{ S_n - \operatorname*{min}_{0 \le i < n} S_i \right\} \ge h.
 \end{equation}
@@
 Extensions of the \gls{cusum} method have been proposed.
-Amongst others, extensions which extend its use to two-threshold methods, and methods appropiate for changes in mean or variance \cite{inclan1994use}.
+Amongst others, extensions which extend its use to two-threshold methods, and methods appropriate for changes in mean or variance \cite{inclan1994use}.
 Whilst adaptive methods have been proposed \cite{alippi2006adaptive}, most methods require manual determination of the threshold parameter $h$.
@@
 
-The appropriate values for the thresholds are highly depended on the characteristics of the data and the trade-off parameter $C$ (as discussed in \Cref{subsec:oc-svm-svdd}).
-This trade-off parameters regulates the fraction of data objects that will be rejected from the constructed model.
+The appropriate values for the thresholds strongly depend on the characteristics of the data and the parameter $C$ (as discussed in \Cref{subsec:oc-svm-svdd}).
+This trade-off parameter regulates the fraction of data objects that will be rejected from the constructed model.
 Since a high value of $C$ assigns a high penalty to outliers, data objects are more likely to be incorporated into the hypersphere by increasing the radius.
@@
 \subsection{Post-processing}
-In our method we use the ratio based thresholding, since it is also used by Camci~\cite{camci2010change}.
+In our method we use the ratio based thresholding, also used by Camci~\cite{camci2010change}.
 From preliminary experiments we observed that a single change in the data can cause many detected change points, using the ratio based method.

### 2014-01-14 Chapters/Chapter4/data_gathering.tex
Put captions above tables
Feedback Chapter 4

--- 2013-12-18 Chapters/Chapter4/data_gathering.tex
@@
 \subsection{Artificial data}\label{subsec:data_gathering_artificial}
-In order to provide an objective comparison to other methods, we will use artificial data sets which are also used in the researches on which our method is based.
+In order to provide an objective comparison to other methods, we will use artificial data sets which are also used in the earlier work on which our method is based.
 These are the data sets used by Takeuchi and Yamanishi \cite{takeuchi2006unifying} and Camci \cite{camci2010change}.
@@
 In the second type of data sets we apply our method of change detection and temporal segmentation to real-world data sets.
-For our setup we record the activities of humans performed both in- and outdoor in an uncontrolled environment.
+For our setup we record the activities of humans performed both in- and outdoor in an unknown environment.
 Activities performed include sitting, standing, walking, running in a straight and curved line, and walking up- and downstairs.
@@
 A video recording from the performed activity is used to annotate the time series with real change points.
-The discovered change point are compared with these annotated change points to give a subjective quality measure.
+The discovered change points are compared with these annotated change points to give a subjective quality measure.
 In \Cref{Chapter6} we give a detailed analysis of the performed activities and the recorded data sets.
 
-For the experiments we used a HTC Sensation XE smartphone as recording device.
-The activities were recorded using a free Android application \cite{sensorlogger}.
+For the experiments we used a smartphone with inertial sensors as recording device.
+The activities were recorded using a free \textsc{Android} application \cite{sensorlogger}.
 This application was chosen for its convenient data format of the sensor recording and its regularity of the sampling interval.
 Table~\ref{tab:recorded_metrics} lists all the recorded metrics.
-For our expiments we used the data for the accelerometer, magnatief field and rotation.
+For our experiments we used the data for the accelerometer, magnetic field and rotation.
+
+We implemented our algorithm in \textsc{MATLAB}.
+For the \gls{svdd} model construction phase we used the \textsc{dd\_tools} library by Tax \cite{Ddtools2013}, which depends on the \textsc{PR\_tools} package, originating from the book by Van Der Heijden \etal \cite{van2005classification}.
+Alternatively, the widely used \textsc{LibSVM} \cite{chang2011libsvm} add-on for the \gls{svdd} algorithm by Chang \etal \cite{changrevisit} can be used.
 
 \begin{center}\begin{table}
+  \caption[Measured metrics]{Measured sensor metrics. The set of axis is always the triple (x, y, z) direction.}
   \begin{tabulary}{\textwidth}{|l|L|c|c|}
     \hline
-    Metric & Description & Units of measure & Typical range \\
+    Sensor metric & Description & Units of measure & Typical range \\
     \hline \hline
@@
     \hline
-    Gyroscope & Rate of rotationg around each axis. & $rad/s$ & $-15$ -- $15$\\
+    Gyroscope & Rate of rotation around each axis. & $rad/s$ & $-15$ -- $15$\\
     \hline
@@
     \hline
-    Rotation & Measre of rotation around the device's rotation axis. & Unitless & $-1$ -- $1$\\
+    Rotation & Measure of rotation around the device's rotation axis. & Unitless & $-1$ -- $1$\\
     \hline
   \end{tabulary}
-  \caption[Measured metrics]{Measured metrics. The set of axis is always the triple (x, y, z) direction.}
+
   \label{tab:recorded_metrics}

### 2014-01-14 Chapters/Chapter4/method_overview.tex
Finalizing Chapter 2, literatur review...
Feedback Chapter 4

@@
 The first stage determines the outliers in a time series by giving a score based on the deviation from a learned model, and thereby creates a new time series.
-The second stage runs on that new created time series and calculates a average over a window of the outlier scores.
+The second stage runs on that new created time series and calculates an average over a window of the outlier scores.
 The problem of change detection is then reduced to outlier detection over that average-scored time series.
@@
 
-Whereas \gls{changeFinder} uses double probability estimation algorithm, our approach follows \gls{svcpd} by constructing a \gls{svm} over a sliding window.
+Whereas \gls{changeFinder} uses a two-stage probability based algorithm, our approach follows \gls{svcpd} by constructing an \gls{svm} over a sliding window.
 The \gls{svcpd} algorithm uses the location of new data points in the feature space $\mathcal{F}$ with respect to the hypersphere and the hypersphere's radius $R$ to determine whether the new data point represents a change point.
 Our approach is a slight simplification of the algorithm: we only use the updated model properties (\ie the radius $R$) to detect a change.
+The \gls{svcpd} also tests every new data point with the current modeled hypersphere, to indicate whether it is an outlier.
 This difference is further discussed and justified in \Cref{sec:method_change_detection}.
@@
 A schematic overview is given in \Cref{fig:method_overview} and shows the steps of the method.
-A more detailed explanation of the ``Update model'' step follows.
+A more detailed explanation of the ``Update model'' step is given in the remainder of this section.
 
-As graphically represented in \Cref{fig:method_overview}, the change detection method starts by processing the data from sensor, such as the accelerometer, magnetic orientation and rotation metrics.
+As graphically represented in \Cref{fig:method_overview}, the change detection method starts by processing the data from sensor, such as the accelerometer, magnetic orientation, and rotation metrics.
 
@@
 Since the \gls{svdd} method requires all the data stream at every timestamp and can not handle missing data on one of the timestamps, all the unique timestamps are filtered out.
-Whilst this results in an overall filtering effect, in practice between $1\%$ and $5\%$ of each data stream is disregarded.
+Whilst this results in an overall filtering effect, in practice between $1\%$ and $5\%$ of each data stream is discarded.
 The effect of this filtering is not significant and the data is not modified.
@@
 Due to the nature of the sensor signals, a normalization step is required in order to set the weight for all the data streams equal.
-The range of the accelerometer signal typically spans $-20$ to $20$, the magnetic field from $-60$ to $60$ and the rotations range is from $-1$ to $1$.
+The range of the accelerometer signal typically spans $-20$ to $20~m/s^2$, the magnetic field from $-60$ to $60~\mu T$ and the rotations value range is from $-1$ to $1$.
 This means that a relative small change in the accelerometer stream could have a much larger impact on the model than the same (absolute) change in the rotation stream, whilst the latter has a larger relative impact.
@@
 This final step of this method, step $5$, is the interpretation of the model properties.
-Many algorithms can be used for this process, all which take a one-dimensional time series as input and determine where change has occured.
+Many algorithms can be used for this process, all which take a one-dimensional time series as input and determine where change has occurred.
 In our setup we used the \gls{rt} and \gls{cusum} methods, to show the modularity of this step.
-This final step is further discussed in \Cref{sec:change_detection_methods}.
+This final step is further discussed in \Cref{sec:method_change_detection}.

### 2014-01-14 Chapters/Chapter4/model_construction.tex
WIP: Chapter 4
Feedback Chapter 4

@@
 \section{Model Construction: Incremental SVDD}\label{sec:method_model_construction}
-After the data is collected and pre-processed (or in the case of the artificial data sets: generated), we construct an online incremental sliding window algorithm.
+After the data is collected (or in the case of the artificial data sets: generated), we construct an online incremental sliding window algorithm.
 We follow the \gls{svcpd} method by Camci \cite{camci2010change} and the \gls{isvdd} implementation by Tax and Laskov \cite{tax2003online}.
 The latter algorithm combines the techniques of online, unsupervised, and incremental learning methods with the earlier introduced \gls{oc-svm} algorithm \gls{svdd}.
-The method is initialized with a fixed window length and then in every step a new data object is added to and the oldest data object is removed from the working set.
+The method is initialized with a fixed window length and then at every step a new data object is added to and the oldest data object is removed from this working set.
+The model construction phase can be regarded as the first stage of the unifying framework by Takeuchi and Yamanishi \cite{takeuchi2006unifying}.
 
@@
 These two operations are the basis for \gls{isvdd}.
-With respect to the timed index, new data objects are added to and old data objects are removed from the working set.
+With the data points ordered by timestamp, new data objects are added to and old data objects are removed from the working set.
 

### 2014-01-14 Chapters/Chapter4/model_features.tex
Feedback Chapter 4

@@
 \end{inparaenum}
-Since in \gls{occ} situations typically there are (almost) no examples of outlier objects, Tax and Duin construct a method to generate outliers based on the assumption that the outliers are uniformly distributed around the target set.
+Since in \gls{occ} situations typically there are (almost) no examples of outlier objects, Tax and Duin construct a method to generate outliers based on the assumption that the the location of (potential) outliers are uniformly distributed around the target set.
 To minimize the error, calculated by the fractions $f_{T-}$ and $f_{O+}$, we should minimize the volume of the target data description (\ie the boundary of \gls{svdd}).
@@
 Five possible time windows are illustrated.
-The left, middle, and right windows cover an area of homogeneous signal.
+Windows $a$, $c$ and $e$ cover an area of homogeneous signal.
 The expected value of $R$ is low.
-The other two windows cover a change entering and leaving the window, respectively.
+The other two windows, $b$ and $d$ cover a change entering and leaving the window, respectively.
 At these locations the radius $R$ is expected to increase temporarily.}
@@
 Since in our method the parameters $C$ and $\sigma$ are kept constant, and thereby also the fraction of target objects being rejected, the only free parameter is the radius $R$.
-During the \gls{isvdd} algorithm the volume of the hypersphere is being minimized.
+During the \gls{isvdd} algorithm the volume of the hypersphere is minimized.
 This means that only the radius of the sphere will change during the analysis over the time series data.
 We will interpret changes in the radius $R$ with the same continuity assumptions as mentioned in \Cref{subsec:kernels}.
-This means that when two objects (in this case radius sizes) are close to each other, their original objects are also close.
-In other words, when the underlying distribution of time series is changed, and thus the characteristics of the data objects changes, the corresponding radius $R$ will also change.
-Since in that case the data objects in the working set become more heterogeneous, the radius of the hypersphere will increase.
+
+For instance, consider two hyperspheres that model two different, but partially overlapping, working sets of objects.
+If the radius size of these two hyperspheres are close to each other, that means the distribution of objects in the sets are also similar.
+In other words, if the distribution over objects in a working set changes, so does the characteristics of the set, and the radius $R$ of the hypersphere will also change.
+If in the case of a change in distribution the data objects in the working set become more heterogeneous, the radius of the hypersphere will increase.
 When, instead, the data objects change from a heterogeneous set to a more homogeneous, we expect the radius to decrease in value, since the data objects are closer to each other in the feature space.

### 2014-01-14 Chapters/Chapter5/artificial_data_generation.tex
WIP: Chapter 5
WIP: chapter 5 and 6

--- 2013-12-12 Chapters/Chapter5/artificial_data_generation.tex
@@
 \section{Artificial data generation}\label{sec:artificial_data_generation}
-In this section we present four data sets as used by \cite{camci2010change,takeuchi2006unifying} to provide for a objective performance comparison.
+In this section we present four data sets, as used by Camci~\cite{camci2010change} and Takeuchi and Yamanishi \cite{takeuchi2006unifying}, to provide for a objective performance comparison.
 All the data sets are modeled by a second-order \gls{ar}-model as in \Cref{eq:artificial_data_sets_model}, where $\epsilon_t$ is a Gaussian distribution modeling the noise with mean $0$ and variance $\sigma^2 = 1$, $a_1 = 0.6$ and $a_2 = -0.5$.
@@
 \end{equation}
-The length of $x_t$ is $10000$ and change points are generated at each $y \times 10000^\text{th}$ data point, with $y = (1, 2, 3, \dots 9)$.
+The length of $x_t$ is $10000$ and change points are generated at each $y \times 10000^\text{th}$ data point, with $y = (1, 2, 3, \dots, 9)$.
 
-Using this general \gls{ar}-model, we create four different data sets.
-Each data set is characterized by a change in mean, variance, or both of the Gaussian term $\epsilon_t$:
+Using this general \gls{ar}-model, we create four different data sets, all visualized in \Cref{fig:plots_artificial_data_sets}.
+Each data set is characterized by a change in mean, variance, or both in the Gaussian term $\epsilon_t$:
 \begin{enumerate}
   \item \textbf{Fixed increasing mean:} at each change point $y$ the mean is increased with $\Delta(y) = 5$.
-  This set origins as the first data set of Camci~\cite{camci2010change}.
+  This set is the first data set of Camci~\cite{camci2010change}.
   The main characteristic is that the difference between the means of the Gaussian term $\epsilon_t$ is fixed and thus the relative distance becomes smaller.
@@
   \item \textbf{Reduced increasing mean:} at each change point $y$ the mean is increased with $\Delta(y) = 10 - y$.
-  This set origins as the first data set of Takeuchi \etal \cite{takeuchi2006unifying}.
+  This set is the first data set of Takeuchi and Yamanishi~\cite{takeuchi2006unifying}.
   In contrast with the previous set, in this set the absolute difference in mean is reduced over time.
@@
   \item \textbf{Reduced increasing mean, increasing variance:} at each change point $y$ the mean is increased with $\Delta(y) = 10 - y$ and the standard deviation is of $\epsilon_t$ is $0.1 / (0.01 + (10000 - x)/1000)$.
-  This set origins as the second set of Camci~\cite{camci2010change}.
+  This set is the second set of Camci~\cite{camci2010change}.
   This set combines the reduced increasing mean property of the previous set with an increasing variance, causing the change points to become more vague over time.
@@
   \item \textbf{Alternating variance:} at each change point $y$ the variance $\sigma^2$ is set to $9.0$ when $y$ is odd, and to $1.0$ otherwise.
-  This set origins as the thirds sets from Camci~\cite{camci2010change} and Takeuchi \etal \cite{takeuchi2006unifying}.
+  This set is the thirds sets from Camci~\cite{camci2010change} and Takeuchi and Yamanishi~\cite{takeuchi2006unifying}.
   This data set is used to measure the ability of detecting changes in decrease of variance, as Camci~\cite{camci2010change} claims is not trivial.
@@
   \centering
+  \begin{subfigure}[b]{1\textwidth}
    \includegraphics[width=1\textwidth]{./Figures/chapter5/set_1_data.eps}
@@
    \label{fig:camci_fixed_increasing_mean}
-\end{figure}
+  \end{subfigure} \\
 
-\begin{figure}
-\centering
+  \begin{subfigure}[b]{1\textwidth}
    \includegraphics[width=1\textwidth]{./Figures/chapter5/set_2_data.eps}
@@
    \label{fig:takeuchi_reduced_increasing_mean}
-\end{figure}
+  \end{subfigure} \\
 
-\begin{figure}
-\centering
+  \begin{subfigure}[b]{1\textwidth}
     \includegraphics[width=1\textwidth]{./Figures/chapter5/set_3_data.eps}
@@
     \label{fig:camci_reduced_increasing_mean_increasing_variance}
-\end{figure}
+   \end{subfigure} \\
 
-\begin{figure}
-\centering
+  \begin{subfigure}[b]{1\textwidth}
    \includegraphics[width=1\textwidth]{./Figures/chapter5/set_4_data.eps}
@@
    \label{fig:camci_takeuchi_alternating_variance}
+  \end{subfigure} \\
+
+  \caption[Artificial data sets]{Plots of the artificial data sets used.}
+  \label{fig:plots_artificial_data_sets}
 \end{figure}
+
+% \begin{figure}
+% \centering
+%   \includegraphics[width=1\textwidth]{./Figures/chapter5/set_1_data.eps}
+%   \caption[Fixed increasing mean]{Data set 1, \emph{Fixed increasing mean}.}
+%   \label{fig:camci_fixed_increasing_mean}
+% \end{figure}
+
+% \begin{figure}
+% \centering
+%   \includegraphics[width=1\textwidth]{./Figures/chapter5/set_2_data.eps}
+%   \caption[Reduced increasing mean]{Data set 2, \emph{Reduced increasing mean}.}
+%   \label{fig:takeuchi_reduced_increasing_mean}
+% \end{figure}
+
+% \begin{figure}
+% \centering
+%   \includegraphics[width=1\textwidth]{./Figures/chapter5/set_3_data.eps}
+%   \caption[Reduced increasing mean, increasing variance]{Data set 3, \emph{Reduced increasing mean, increasing variance}.}
+%   \label{fig:camci_reduced_increasing_mean_increasing_variance}
+% \end{figure}
+
+% \begin{figure}
+% \centering
+%   \includegraphics[width=1\textwidth]{./Figures/chapter5/set_4_data.eps}
+%   \caption[Alternating variance]{Data set 4, \emph{Alternating variance}.}
+%   \label{fig:camci_takeuchi_alternating_variance}
+% \end{figure}

### 2014-01-14 Chapters/Chapter5/quality_metrics.tex
WIP: chapter 5 and 6

@@
 \section{Quality metrics}\label{sec:artificial_data_quality_metrics}
-For an objective measure of quality of the proposed method applied to the four aforementioned data sets, we employ two metrics, inspired by and slighty altered from Takeuchi and Yamanishi~\cite{takeuchi2006unifying}.
-The first metric is the $\operatorname*{average \_ delay}$, for which we first need to define the notion of $\operatorname*{delay}$ for each change:
+For an objective measure of quality of the proposed method applied to the four aforementioned data sets, we employ two metrics, inspired by and slightly altered from Takeuchi and Yamanishi~\cite{takeuchi2006unifying}.
+The first metric is the average delay $d_{avg}$, for which we first need to define the notion of delay $d$ for each change:
 \begin{equation}\label{eq:delay}
-  \operatorname*{delay}(y) = |t_y - t_y^*|,
+  d(y) = |t_y - t_y^*|,
 \end{equation}
 where $t_y$ is the time of the closest discovered change point and $t_y^*$ is the true change time.
-We can then define the $\operatorname*{average \_ delay}$ over all the discovered change points $Y$ as:
+We can then define the $d_{avg}$ over all the discovered change points $Y$ as:
 \begin{equation}\label{eq:average_delay}
-  \operatorname*{average \_ delay}(Y) = \frac{\sum_{y=1}^Y \operatorname*{delay}(y)}{Y}.
+  d_{avg}(Y) = \frac{\sum_{y=1}^Y d(y)}{Y}.
 \end{equation}
 Besides the delay, which should be low to indicate discovered change points near the actual change point, we need to penalize discovered change points which do not relate to a true change point.
-For this we use the \acrlong{far}, expressed as
+For this we use the \acrlong{far} $\gamma$, expressed as
 \begin{equation}\label{eq:false_alarm_rate}
-  \operatorname*{far}(Y) = \frac{Y_{\operatorname*{false}}}{Y}.
+  \gamma(Y) = \frac{Y_{\operatorname*{false}}}{Y}.
 \end{equation}
 
-For a high quality method the $\operatorname*{average\_delay}$ and the $\operatorname*{far}(Y)$ should both be low.
+For a high quality method the average delay $d_{avg}$ and \gls{far} $\gamma$ should both be low.
 In our results we will also list the parameters used, since these need to be user-defined and are part of the (manual) optimization process.
@@
   \item \textbf{Window length}: The window length that is processed by the \gls{svdd} algorithm. This is the number of data points that are continuously modeled in the shape in a hypersphere.
-  \item \textbf{Sigma of \gls{rbf}}: The value of $\sigma$ for the \gls{rbf} kernel. The effect of this parameter is discussed in \Cref{subsec:svm_model_parameters}.
+  \item \textbf{Sigma of \gls{rbf}}: The value of $\sigma$ for the \gls{rbf} kernel. The effect of this parameter has been discussed in \Cref{subsec:svm_model_parameters}.
   \item \textbf{High threshold}: The $th_\text{high}$ threshold value for the \gls{rt} method described in \Cref{subsec:ratio_thresholding}. A higher value indicates a more precise and robust setting.
@@
 \end{itemize}
-For all the sets we have used a value of $C = 0.1$, which indicated that $10\%$ of the data points are used as outliers.
+For all the sets we have used a value of $C = 0.1$, which indicates that $10\%$ of the data points were considered to be outliers.
 This is a normal setting for the \gls{svdd} method.

### 2014-01-14 Chapters/Chapter6.tex
WIP: chapter 5 and 6

@@
 In this chapter we discuss the real-world data sets used for this research and the performance of our proposed algorithm on these data sets.
-In \Cref{sec:data_sets} we discuss common available data sets and argue that these are inappropriate for our research.
-It is followed by an overview of the setup we have created to record our own continuous data sets in an uncontrolled environment, both for in- and outdoor activities.
+In \Cref{sec:data_sets} we discuss available data sets and argue that these are inappropriate for our research.
+It is followed by an overview of the setup we have created to record our own continuous data sets in an unknown environment, both for in- and outdoor activities.
 The series of activities performed are discussed in \Cref{sec:data_gathering}.

### 2014-01-14 Chapters/Chapter6/data_gathering.tex
WIP: chapter 5 and 6

@@
 \subsection{Outdoor free}\label{subsec:outdoor_free}
-The second series of outdoor recordings consist of more free-form activities.
+The second series of outdoor recordings consist of more free form activities.
 The performed activities consist, as with the first set, of standing, walking, and running.
@@
 \subsection{Indoor stairs}\label{subsec:indoor_stairs}
-The final series of activities recorded consist of walking up and down the stairs, indoor.
+The final series of activities recorded, consists of walking up and down the stairs, indoor.
 The stairs are shaped in an semi-circular form, with a small horizontal segment on each floor.

### 2014-01-14 Chapters/Chapter6/data_sets.tex
WIP: chapter 5 and 6

@@
 \section{Data Sets}\label{sec:data_sets}
-In the previous chapter we have applied our method to the artificial data sets from \cite{camci2010change,takeuchi2006unifying}.
+In the previous chapter we have applied our method to the artificial data sets from Camci~\cite{camci2010change} and Takeuchi and Yamanishi~\cite{takeuchi2006unifying}.
 In this chapter we apply the method to real-world data sets.
@@
   \item The environment should be natural and uncontrolled, both in- and outdoor,
-  \item There data should be clearly annotated, as objective as possible,
+  \item The data should be clearly annotated, as objective as possible,
   \item The required hardware (and software) should be widely available.
@@
 The used phones are the \textsc{HTC Sensation XE} and \textsc{HTC Desire}, running the \textsc{Android} smartphone operating system.
-During the activities the intertial sensor data is recorded using a free application, \textsc{Sensor Logger}~\cite{sensorlogger}.
+During the activities the inertial sensor data is recorded using a free application, \textsc{Sensor Logger}~\cite{sensorlogger}.
 This application records the sensor data to \textsc{CSV} files, which are then transformed to \textsc{MATLAB} compatible files.

### 2014-01-14 Chapters/Chapter6/results.tex
Put captions above tables
WIP: chapter 5 and 6

@@
 In \Cref{tab:results_real_world} for each run the results are listed, together with the used parameters.
-We have used a window length of 50 frames, which represents around $1.16$ seconds of data (the frame-rate of the recordings is not fixed, and we apply a slight filtering).
-The other parameter values are the result of an educated guess and manual optimization.
+We have used a window length of 50 frames, which represents around $1.16$ seconds of data (the frame-rate of the recordings is not fixed).
+The other parameter values, \ie the \gls{rbf} $\sigma$, high and low threshold and closeness factor, are the result of an educated guess and manual optimization.
 In general, a high value for the upper threshold and low value for the lower threshold indicates a robust method, since it indicates that change points are significantly different from homogeneous segments.
@@
   \centering
+  \caption[Results real world runs]{Parameter settings and results of the real-world data sets.}
   \begin{tabulary}{\textwidth}{|l|c|c|c|c|c|c|c|}
@@
   \end{tabulary}
-  \caption[Results real world runs]{Parameter settings and results of the real-world data sets.}
   \label{tab:results_real_world}
@@
 \subsection{Subjective observations and remarks}
-Since the manual annotation of change points is ambiguous, the objective quality metrics can give a distorted view of the performance.
+It can be ambiguous where to place the (manually annotated) change points, in the case of real world data.
+The objective quality measure can give a distorted view on the performance.
 To overcome this, we will state observations and remarks found after close (visual) inspection of the annotated and discovered change points.
 In \Cref{fig:plots_all_runs_results} all the results are plotted.
-The black solid lines indicate our manual determined change points.
+The black solid lines indicate our manually determined change points.
 The dashed purple lines indicate the change points as discovered by our method.
@@
 With a high sensitivity it seems, initially, the method produces many false positives.
-After closer inspecting of the data, we can confirm there is a change in the data, although the performed activity stayed the same.
+After closer inspecting of the data, we can confirm that there is a change in the data, although the performed activity stayed the same.
 This is also a weakness of the method: since the sensitivity is set globally, it can produce many semi-false positive change points for some segments of a run.
@@
 Currently, our method employs a global setting for the thresholds.
-A possible improvement would be to use locally optimized parameters, although that would require an reflective data processing approach or more a priori about the data.
+A possible improvement would be to use locally optimized parameters, although that would require an reflective data processing approach or more a priori information about the data.
 With our approach, the lack of a priori knowledge about the data distribution is one of the design decisions.

### 2014-01-14 Chapters/Chapter7.tex
Added plots for chapter 6
WIP: chapter 5 and 6

@@
 In this research we have applied \gls{occ} methods to accelerometer data, recorded during the performance of various human activities, in order to create a temporal segmentation of those recordings.
-The goal was to find the change points between activities, assuming that it will aid better understanding of the sensor values.
+The goal was to find the change points between activities, assuming that it will aid our understanding of the sensor values.
 To do so, we used the \gls{svdd} algorithm by Tax~\cite{tax1999support}, following the method of applying it to a sliding window of data, as done by Camci~\cite{camci2010change}.
@@
 The limitations of our method and research is subject of \Cref{sec:limitations}.
-These two sections are followed by our suggestions for future research.
+\Cref{sec:further_research} provides suggestions for future research.
 \Cref{sec:final_remarks} is the final section of this thesis.
 
-\section{What is done}
-The scope of this research was the temporal segmentation of recordings obtained from human activities.
+\section{What is done?}
+The goal of this research was to find a temporal segmentation of recordings obtained from human activities.
 This was performed in the wider context of activity classification.
@@
 We have used the accelerometer data for measurement of speed, the gyroscope data for measurement of rotation, and the magnetometer to measure the direction and orientation of the performed activities.
-The proposed method, as discussed in \Cref{Chapter4}, reduces this 9-dimensional signal to a single property, obtained from the constructed model, which can be interpreted to indicate change.
+The proposed method, as discussed in \Cref{Chapter4}, reduces a high dimensional signal to a single property, obtained from the constructed model, which can be interpreted to indicate change.
 
@@
 
-\section{Main findings / Contributions}\label{sec:main_findings}
-\TODO{Vraag aan Anne: wat is een goede titel voor deze section? Of moeten we Findings en Contributions splitsen?}
+\section{Contributions}\label{sec:main_findings}
 While temporal segmentation of time series data is not uncommon, the application to inertial signals is currently not under much research attention.
@@
 
-\TODO{Toevoegen: samenvatting van resultaten uit \Cref{Chapter6}: de subjectieve waarnemingen?}
+\TODO{Vraag aan Anne: toevoegen samenvatting van resultaten uit \Cref{Chapter6}: de subjectieve waarnemingen?}
 
@@
 This occurs mainly in situations where a rotation is part of the change (\eg walking around a corner).
-Thus depending on the desired granularity of the segmentation, it can be beneficial to disregard some of the time series data.
+Thus depending on the desired granularity of the segmentation, it can be beneficial to discard some of the time series data.
 Determining (dynamically) which sensors should be, and which should not be, used in the model construction phase (known as \emph{feature selection}) could be a subject of further research.
@@
 We believe that the temporal segmentation of time series data can aid the process of activity classification, which is a logical follow-up on this research.
-\TODO{Hier nog meer samenvatten?}
+\TODO{Vraag aan Anne: Hier nog meer samenvatten? Wat is een goede uitsmijter?}

### 2014-01-15 Chapters/Chapter7.tex
WIP: feedback chapter 7

@@
 For a better and more robust analysis of the algorithm it can be applied to more real-world data.
-Although we have recorded and publicized our used data, application to other commonly used data sets would enable a fair comparison.
-Furthermore, since the manual annotation of change points is an ambiguous task, the results are very subjective.
+Although we have recorded and publicized our data, application to other commonly used data sets would enable a fair comparison.
+Furthermore, since the manual annotation of change points is an ambiguous task, the results remain subjective.
 
 \section{Future research}\label{sec:further_research}
-The motivation for this research was, amonst others, the assumption that an explicit temporal segmentation of inertial signals can be used to obtain a better classification of the performed activities.
+The motivation for this research was, amongst others, the assumption that an explicit temporal segmentation of inertial signals can be used to obtain a better classification of the performed activities.
 There are two reasons for that.
@@
 This enables a better mathematical model to be constructed.
-A future research in this line of assumptions is to compare the performance of classification methods using this obtained temporal segmentation and those that do not.
-This would confirm of reject the assumption that a temporal segmentation enables a better classification.
+In future research the performance of classification methods that use temporal segmentation can be compared with those that do not.
+This would confirm or reject the assumption that temporal segmentation enables better classification.
 
@@
 
-Since the temporal segmentation of inertial signal is not yet a very active field of study, more studies can be performed to obtain it.
-In our approach we have used (all) the raw data directly.
+Since temporal segmentation of inertial signal is not yet a very active field of study, more studies can be performed on this subject.
+In our approach we have used (all) the raw data directly, to create a method independent of the domain.
 In the field of activity recognition many methods make use of pre-processing steps, from which one step often is feature extraction from the raw data.
-That, and other pre-processing steps, could be incorporated in the explicit temporal segmentation of inertial signals.
-Future research could focus on extracted features that aid the detection of change points.
+That, and other pre-processing steps, could be incorporated in explicit temporal segmentation of inertial signals.
+Future research could focus on using those extracted features to aid the detection of change points.
 
-In that same line, we have encountered situations in which the exclusion of some sensors, \eg the accelerometer, yields better results.
+Similar, feature selection if often used to use only the best dimensions of a signal.
+We have encountered situations in which the exclusion of some sensors, \eg the accelerometer, yields better results.
 This occurs mainly in situations where a rotation is part of the change (\eg walking around a corner).
 Thus depending on the desired granularity of the segmentation, it can be beneficial to discard some of the time series data.
-Determining (dynamically) which sensors should be, and which should not be, used in the model construction phase (known as \emph{feature selection}) could be a subject of further research.
+Applying feature selection to the sensors to restrict the data used for model construction could be a subject of further research.
 
@@
 \section{Final words}\label{sec:final_remarks}
-In this research we have constructed a method to create a temporal segmentation of inertial sensor time series data, originating from performed human activities.
+In this research we have constructed a method to segment inertial sensor time series data, from performed human activities.
 To this end we have applied \gls{occ} and \gls{oc-svm} based method to our own recorded data sets.

### 2014-01-15 Chapters/Chapter1.tex
Finished feedback Chapter 1

@@
 
-In this thesis the goal is to find the temporal segmentation of time series data explicitly, prior to the classification of the activities.
+The goal of this thesis is to create an method for online temporal segmentation of time series data.
+This will be performed explicitly, prior to the classification of data segments to learned activities.
 
-This is done under the assumption that it could be beneficial for classification methods to possess this explicit segmentation, since the model construction phase can use more data than the time-windowed approach would allow.
+Our assumption is that a temporal segmentation of time series can be beneficial information for classification methods.
+Instead of using a fixed window approach, the classification methods can use a larger set of homogeneous data.
+Thereby, the methods have more information to create a solid model of the activities.
 
 To find a temporal segmentation, we rely on change detection in time series.
-In the context of this research, these time series consists of recordings from inertial sensors found in smartphones, such as the accelerometer signal, during the performance of human activities.
-We are interested in real-world data from both in- and outdoor activities in an uncontrolled environment and in a continuous manner.
-Although the number of performed activities will be limited, we assume the temporal segmentation must be more robust than in the case of data from a controlled environment.
+In the context of this research the time series consists of recordings from inertial sensors during the performance of human activities.
+The sensors, such as the accelerometer, are nowadays widely available in smartphones.
+We are interested in real-world data from both in- and outdoor activities in an unknown environment and in a continuous manner.
+
+Compared to recordings in a controlled laboratory environment, we expect the real-world data to be prone to unexpected characteristics.
+The constructed method must be robust in order to handle these.
+
 For the change detection algorithm we will use a special form of a \gls{svm}, the \gls{occ} based \gls{svdd}, as introduced by Tax~\cite{tax2001one}.
@@
 
-\section{Scope}
-The scope of this thesis is to find a temporal segmentation of time series data from inertial sensors embedded in smartphones, obtained during the performance of human activities.
-As mentioned above, the segmentation is often obtained as a by-product from direct windowed classification.
-Other approaches have a very rigid form of segmentation, \eg by limiting the number of possible segment types \cite{himberg2001time,chamroukhi2013joint}, or have applied it to different type of time series data \cite{li2007segmentation,fuchs2010online,guenterberg2009automatic}.
+The remainder of this introduction is organized as follows.
+In \Cref{sec:intro_scope} the scope of this thesis is discussed.
+A short discussing on the workings of an \acrlong{oc-svm} is provided in \Cref{sec:intro_ocsvm}.
+The data sets we have used in this work are discussed in \Cref{sec:intro_data_sets}.
+\Cref{sec:intro_contributions} gives a short overview of our contributions.
+Finally, \Cref{sec:intro_thesis_structure} outlines the structure of this thesis.
+
+\section{Scope}\label{sec:intro_scope}
+The scope of this thesis are segmentation methods of inertial sensor data in the field of machine learning.
+The type of data is inertial sensor time series, collected by on-body worn sensors as found in smartphones.
+These time series are recorded during the performance of real-world human activities.
+The setting of the activities are real-world unknown environments.
+
+As mentioned in the previous section, a segmentation is often obtained as a by-product from direct windowed classification.
+Other approaches have a very rigid form of segmentation, \eg by limiting the number of possible segment types \cite{himberg2001time,chamroukhi2013joint}, or have applied it to different type of time series data.
+In Fuchs \etal \cite{fuchs2010online} artificial sinusoidal data is processed.
+Li and Zheng \cite{li2007segmentation} use visual motion data streams to find patterns of activities.
+The method of Guenterberg \etal \cite{guenterberg2009automatic} is able to distinguish periods of rest and activity in inertial sensor data streams.
 Others have extracted features from windows of data which are used for the segmentation \cite{guo2012adaptive}.
 These activities include sitting, standing, walking, running, and ascending and descending staircases.
+
 The problem of finding a temporal segmentation can be seen as a sub-problem to the problem of Activity Classification.
@@
 To find these segments of homogeneous data, we employ a model construction phase.
-The \gls{oc-svm} model transforms the data points to a high-dimensional hyperspherical shape from which we extract the radius as characterizing metric.
+The \acrlong{oc-svm} transforms the data to a higher dimensional space.
+Non-linear mapping functions are used to add flexibility to the classification decision boundary.
+In the higher dimensional space an enclosing boundary around the data points, in the shape of a hypersphere, is constructed.
+The size, expressed as the radius, of the hypersphere depends on the distribution of the data points in the model.
+We will use the radius of this hypersphere as a characterizing metric and base our change detection on it.
+
 The model processes the data in windows of fixed length.
-A change in data characteristics should then result in a change of the hypersphere's radius.
+A change in data distribution and characteristics should then result in a change of the hypersphere's radius.
 Such a (sudden) change in the radius thus reflects a change in the time series data.
@@
 
-\section{One-Class Support Vector Machines}
+\section{One-Class Support Vector Machines}\label{sec:intro_ocsvm}
 In the setup, as described in the previous section and visualized in \Cref{fig:thesis_goal}, we employ a model construction phase.
@@
 
-\section{Data sets}
+\section{Data sets}\label{sec:intro_data_sets}
 In recent research two data sets for \gls{har} have been made public.
@@
 So in case of ambiguity there is no ground truth to consult.
-To overcome these problems, we have created our own real-world data set.
-\TODO{Create name for our own data sets}
+To overcome these problems, we have created our own real-world data set: \acrlong{almende-data} \cite{vlasveld2014acras}.
 
@@
 
-\section{Contributions}
+\section{Contributions}\label{sec:intro_contributions}
 As a result from our proposed method and application to recorded real-world data sets, this thesis will provide the following contributions:
@@
   This data set can be grounded by using the video recordings of the subject while performing the activities.
-  The data set, consisting of raw inertial sensor recordings, manual change point and activity labeling, and video recordings, is used for testing the method and is made available to the public \cite{vlasveld2013continuous}.
+  The data set, consisting of raw inertial sensor recordings, manual change point and activity labeling, and video recordings, is used for testing the method and is made available to the public \cite{vlasveld2014acras}.
   \item \textbf{Application of \gls{oc-svm} methods to inertial sensor data:} to the best of our knowledge, this thesis is the first application of \gls{oc-svm} methods, especially the \gls{svdd} method, to inertial sensor signals in the context of temporal segmentation.
@@
 
-\section{Thesis structure}
+\section{Thesis structure}\label{sec:intro_thesis_structure}
 The structure of this thesis is as follows.
@@
 In \Cref{Chapter5} we apply the method to artificial data and show the objective performance.
-In \Cref{Chapter6} the method is applied to our real-world data set \TODO{Add name}.
+In \Cref{Chapter6} the method is applied to our real-world \gls{almende-data} \cite{vlasveld2014acras}.
 In that chapter we show the ability of detecting change in \gls{har} data, recorded by inertial sensors.

### 2014-01-15 Chapters/Chapter6/data_sets.tex
Finished feedback Chapter 1

@@
 The manually determined change points are used to draw quality conclusion over the algorithm's performance, by comparing it to the discovered change points.
-All of the used data sets, including the video material and manual labeling, are made available to the public \cite{vlasveld2013continuous} for further research.
-\TODO{Publish web page with data sets}
+All of the used data sets, including the video material and manual labeling, are made available to the public as \gls{almende-data} \cite{vlasveld2014acras} for further research.
 The following section discusses the performed activities during the recordings.

### 2014-01-15 Chapters/Chapter6/results.tex
Finished feedback Chapter 1

@@
 \subsection{Subjective observations and remarks}
-It can be ambiguous where to place the (manually annotated) change points, in the case of real world data.
+It can be ambiguous where to place the (manually annotated) change points, in the case of real-world data.
 The objective quality measure can give a distorted view on the performance.

### 2014-01-15 glossary-acronyms.tex
WIP: Working on organization of Chapter 2
WIP: kernels
Add equations
'Refactoring'
Cleanup of sections
WIP
Finished feedback Chapter 1

@@
 % A C R O N Y M S
+
+\newacronym{almende-data}{ACRAS}{Almende Continuous Real-world Activities with on-body smartphones Sensors Data Set}
 
@@
 
+\newacronym{dr}{DR}{Dimensionality Reduction}
+
+\newacronym{far}{FAR}{False Alarm Rate}
+
 \newacronym{har}{HAR}{Human Activity Recognition}
+
+\newacronym{uci-har}{UCI HAR}{Human Activity Recognition Using Smartphones Data Set}
 
@@
 
+\newacronym{plr}{PLR}{Piecewise Linear Representation}
+
 \newacronym{qp}{QP}{Quadratic Programming}
@@
 \newacronym{rbf}{RBF}{Radial Base Function}
+
+\newacronym{rkhs}{RKHS}{Reproducing Kernel Hilbert Space}
 
@@
 
+\newacronym{svr}{SVC}{Support Vector Regression}
+
 \newacronym{svcpd}{SVCPD}{Support Vector based Change Point Detection}
@@
 \newacronym{swab}{SWAB}{Sliding Window And Bottom-up}
+
+\newacronym{wisdm}{WISDM}{Wireless Sensor Data Mining}

### 2014-01-15 Chapters/Chapter2/density_ratio.tex
Finished feedback Chapter 2

@@
 
-The method proposed by Kawahara and Sugiyama~\cite{kawahara2009change} is composed of an online version of the \gls{kliep} algorithm.
-The method also considers \emph{sequences} of samples (rather than samples directly) because the time series samples are generally not independent over time.
-An advantage over other nonparametric approaches, such as sequential one-class support vector machines, is that the model has a natural cross-validation procedure.
-This makes that the value of tuning parameters, such as the kernel bandwidth, can be objectively obtained.
+The proposed method by Kawahara and Sugiyama~\cite{kawahara2009change} uses an online version of the \gls{kliep} algorithm.
+It considers \emph{sequences} of samples (rather than samples directly) because the time series samples are generally not independent over time.
+The method detects change by monitoring the logarithm of the likelihood ratio between densities of reference (past) and test (current) time intervals.
+If it exceeds a predetermined threshold value, the beginning of the test interval is marked as a change point.
 
-In their formulation change is detected by monitoring the logarithm of the likelihood ratio between the reference (past) and test (current) time intervals
-\begin{equation}
-  S = \sum_{i=1}^{n_\text{te}} \text{ln} \frac{p_{\text{te}}(\matrixsym{Y}_\text{te}(i)}{p_\text{rf}(\matrixsym{Y}_\text{te}(i))}
-\end{equation}
-Where $\matrixsym{Y}_\text{te}(i)$ is a sequence of samples from the test interval.
-A change is detected when $S > \mu$, for some predetermined threshold $\mu$.
-The question is then how to calculate the density ratio
-\begin{equation}
-  w(\matrixsym{Y}) := \frac{p_\text{te}(\matrixsym{Y)})}{p_\text{rf}(\matrixsym{Y})}
-\end{equation}
-because this ratio is unknown and should be estimated.
-The naive approach is to estimate the ratio by taking the ratio of the estimated densities.
+Since the density ratio is unknown, it needs to be estimated.
+The naive approach is to estimate it using estimated densities of the two intervals.
 Since this is known to be a hard problem and sensitive for errors, the solution would be to estimate the ratio directly.
 
-The procedure of the method proposed by Kawahara and Sugiyama \cite{kawahara2009change} is to first apply the batch \gls{kliep} algorithm with model selection for initial parameter $\vectorsym{\alpha}$ and kernel width calculation.
-Then for every new sample the reference and test intervals are shifted and the calculated parameters $\vectorsym{\alpha}$ are updated.
-Finally the logarithm of the likelihood ratio is evaluated.
-If it is above the threshold $\mu$ the current time is reported as a change point.
-
-Improvements in this line of research by Liu \etal \cite{liu2013change} has led the application of improved density-ratio estimation methods to the problem of change detection.
-Such an improvement is the Unconstrained Least-Squares Importance Fitting (uLSIF) method \cite{kanamori2009least} and an extension which possesses a superior nonparametric convergence property: Relative uLSIF (RuLSIF) \cite{yamada2013relative}.
+The method by Liu \etal \cite{liu2013change} estimates the ratio of probabilities directly, instead of estimating the densities explicitly.
+Other methods using ratio-estimation are the Unconstrained Least-Squares Importance Fitting (uLSIF) method \cite{kanamori2009least}, and an extension which possesses a superior nonparametric convergence property: Relative uLSIF (RuLSIF) \cite{yamada2013relative}.

### 2014-01-15 Chapters/Chapter2/svm.tex
Finished feedback Chapter 2

@@
 In this section we review methods of change detection based on the applications of \glspl{svm} as model construction method.
-A number of proposals have been made using this method.
-Sch{\"o}lkopf \etal \cite{scholkopf1999support} applies Vapnik's principle, to never solve a problem which is more general than the one that one is actually interested in, to novelty detection.
+% A number of proposals have been made using this method.
+Sch{\"o}lkopf \etal \cite{scholkopf1999support} applies Vapnik's principle never to solve a problem which is more general than the one that one is actually interested in.
 In the case of novelty detection, they argue there is no need for a full density estimation of the distribution.
-Simple algorithms estimate the density by considering how many data points fall in a region of interest.
+Some algorithms estimate the density by considering how many data points fall in a region of interest.
 The \gls{nu-svm} method instead starts with the number of data points that should be within the region and estimates a region with that desired property.

### 2014-01-15 Chapters/Chapter3/change_detection_time_series.tex
Finished feedback Chapter 3

@@
 Assume $p$ can be ``decomposed'' in two different \gls{iid} stationary one-dimensional Gaussian processes $p^{(1)}$ and $p^{(2)}$.
-For a time point $a$ data points for which $t < a$ are drawn from $p^{(1)} = N(\mu_1, \sigma_1^2)$ and for $t \geq a$ from $p^{(2)} = N(\mu_2, \sigma_2^2)$.
-If $p^{(1)}$ and $p^{(2)}$ are different, then the time point $t = a$ is a \emph{change point}.
+% For a time point $a$ the data points for which $t < a$ are drawn from $p^{(1)} = N(\mu_1, \sigma_1^2)$ and for $t \geq a$ from $p^{(2)} = N(\mu_2, \sigma_2^2)$.
+Consider a potential change point at time $a$.
+Data points $x_1, \dots, x_{a-1} \sim \mathcal{N}(\mu_1, \sigma_1^2) = p^{(1)}$ and $x_a, \dots, x_t \sim \mathcal{N}(\mu_2, \sigma_2^2 ) = p^{(2)}$.
+If $p^{(1)}$ and $p^{(2)}$ are different (enough), then the time point $t = a$ is marked as a change point.
+
 Takeuchi and Yamanishi~\cite{takeuchi2006unifying} expess the similarity between the stochastic processes with the \gls{kl} divergence $D(p^2||p^1)$.

### 2014-01-15 Chapters/Chapter3/one_class_svm.tex
Finished feedback Chapter 3

@@
 
-This minimization problem can be solved (using \gls{qp}) and transformed to its (Lagrange) dual formulation.
-In the dual formulation the problem scales with the number of training examples $n$ instead of the dimensionality $d$ of the samples.
+This minimization problem can be solved using \gls{qp} and scales in the number of dimensions $d$, since $w \in \mathbb{R}^d$.
+In its dual formulation, as shown in \Cref{eq:svm_lagrange}, the non-negative Lagrange multipliers $\alpha_i$ are introduced.
+Using this formulation the problem scales in the number of data points $n$, since $\vectorsym{\alpha} \in \mathbb{R}^n$.
+% In the dual formulation the problem scales with the number of training examples $n$ instead of the dimensionality $d$ of the samples.
 Solving this problem directly in the high dimensional feature space $\mathcal{F}$ makes it intractable.
-The linear approximation function corresponds to the kernel function in the dual formulation.
-Solving this dual formulation is equivalent to solving the primal formulation \cite{cherkassky2007learning}.
-In the dual formulation the Lagrange multipliers $a_i >= 0$ are introduced and the decision function becomes:
+% The linear approximation function corresponds to the kernel function in the dual formulation.
+% Solving this dual formulation is equivalent to solving the primal formulation \cite{cherkassky2007learning}.
+It is shown that solving the dual formulation is equivalent to solving the primal form \cite{cherkassky2007learning}.
+% In the dual formulation the non-negative Lagrange multipliers $a_i$ are introduced and the decision function becomes:
+The linear classifier decision function in the dual form is optimized over $\alpha_i$:
 \begin{equation}\label{eq:svm_lagrange}
-  f(x) = \operatorname{sgn}( \sum_{i=1}^n \alpha_i y_i K(x, x_i) + b),
+  f(x) = \operatorname{sgn}( \sum_{i=1}^n \left\{ \alpha_i y_i K(x, x_i) + b \right\} ),
 \end{equation}
@@
 Here every data point in $\mathcal{I}$ for which $a_i > 0$ is weighted in the decision function and thus ``supports'' the classification machine: hence the name ``\acrlong{svm}''.
-Since it is shown that under certain circumstances \glspl{svm} show an equality to sparse representations \cite{girosi1998equivalence,smola1998connection}, there will often be relatively few Lagrange multipliers with a non-zero value.
+Since it is shown that under certain circumstances \glspl{svm} resembles sparse representations \cite{girosi1998equivalence,smola1998connection}, there will often be relatively few Lagrange multipliers with a non-zero value.
 
@@
 The decision function in \Cref{eq:svm_lagrange} only relies on the dot products of mapped data points in the feature space $\mathcal{F}$ (\ie all pairwise distances between the data points in that space).
-It shows \cite{flach2012machine} that for any function $K$ that has the same result in feature space $\mathcal{F}$, without an explicit mapping to the higher dimension $\mathcal{F}$, the dot products can be substituted by the kernel function $K$, as introduced by Aizerman~\etal \cite{aizerman1964theoretical} and applied to \glspl{svc} by Vapnik \cite{vapnik1998statistical}:
+It shows \cite{flach2012machine} that for any function $K$ that corresponds to a dot product in feature space $\mathcal{F}$, without an explicit mapping to the higher dimension $\mathcal{F}$, the dot products can be substituted by that kernel function $K$.
+This is introduced by Aizerman~\etal \cite{aizerman1964theoretical} and applied to \glspl{svc} by Vapnik~\cite{vapnik1998statistical}, known as the `kernel trick':
 \begin{equation}
@@
 where vectors $\vectorsym{z}$ and $\vectorsym{z}'$ are projections of data objects $\vectorsym{x}$ and $\vectorsym{x}'$ through $\phi(\vectorsym{x})$ on the features space $\mathcal{F}$.
-The dot product kernel $K$ is determined by the sum
-\begin{equation}
-  K(\vectorsym{x}, \vectorsym{x}') = \sum_{j=1}^m g_j(\vectorsym{x}) g_j(\vectorsym{x}')
-\end{equation}
-of basis functions $g_j(\vectorsym{x})$.
+% The dot product $\phi(\vectorsym{x}) \cdot \phi(\vectorsym{x}')$ can be replaced by $K(\vectorsym{x}, \vectorsym{x}')$, which is known as the `kernel trick' and found by Vapnik~\cite{vapnik1998statistical}.
+% \begin{equation}
+  % K(\vectorsym{x}, \vectorsym{x}') = \sum_{j=1}^n g_j(\vectorsym{x}) g_j(\vectorsym{x}')
+  % K(\vectorsym{x}, \vectorsym{x}') = \sum_{j=1}^n k(\vectorsym{x}, \vectorsym{x}_j)
+% \end{equation}
+% of $n$ basis functions $k(\vectorsym{x}, \vectorsym{x}')$.
 Note that the evaluation of dot products in the feature space $\mathcal{F}$ between vectors is performed indirectly via the evaluation of the kernel $K$ between (support) vectors in the input space $\mathcal{I}$.
-This is known as the \emph{kernel trick} and gives the \gls{svm} the ability to create non-linear decision function without high computational complexity.
+% This is known as the \emph{kernel trick} and gives the \gls{svm} the ability to create non-linear decision function without high computational complexity.
+This is gives the \gls{svm} the ability to create non-linear decision function without high computational complexity.
 
 The kernel function $K$ can have different forms, such as linear, polynomial and sigmoidal but the most used (and flexible) form is the Gaussian \gls{rbf}.
-The basis functions have the form
+The inner product kernel with \gls{rbf} basis functions have the form
 \begin{equation}
-  g(x) = \operatorname*{sign} \left(  \sum_{i=1}^{n} \alpha_i \operatorname*{exp} \left\{ \frac{\rVert \vectorsym{x} - \vectorsym{x}_i \rVert ^2}{\sigma^2} \right\} \right),
+  % k(x, x') = \operatorname*{sign} \left(  \sum_{i=1}^{n} \alpha_i \operatorname*{exp} \left\{ \frac{\rVert \vectorsym{x}_j - \vectorsym{x} \rVert ^2}{\sigma^2} \right\} \right),
+  K(\vectorsym{x}, \vectorsym{x}') = \operatorname*{exp} \left\{ - \frac{\rVert \vectorsym{x} - \vectorsym{x}' \rVert ^2}{2 \sigma^2} \right\},
 \end{equation}
-where $\sigma$ defines the width and $\rVert \vectorsym{x} - \vectorsym{x}' \rVert ^2$ is the dissimilarity measure expressed as Euclidian distance.
-The inner product kernel $K$ then becomes
-\begin{equation}
-  K(\vectorsym{x}, \vectorsym{x}') = \operatorname*{exp} \left\{ - \frac{ \rVert \vectorsym{x} - \vectorsym{x}' \rVert ^2}{\sigma^2} \right\}.
-\end{equation}
+where $\sigma$ defines the Gaussian width and $\rVert \vectorsym{x} - \vectorsym{x}' \rVert ^2$ is the dissimilarity in Euclidean distance.
+% The inner product kernel $K$ then becomes
+% \begin{equation}
+  % K(\vectorsym{x}, \vectorsym{x}') = \operatorname*{exp} \left\{ - \frac{ \rVert \vectorsym{x} - \vectorsym{x}' \rVert ^2}{\sigma^2} \right\}.
+% \end{equation}
+
 The kernel $K$ maps input space $\mathcal{I}$ to the feature space $\mathcal{F}$ which is a \gls{rkhs} of (theoretically) infinite dimensions.
 As Smola \etal \cite{smola1998connection} state, this Gaussian kernel yields good performance, especially when no assumptions can be made about the data.
-As an explanation, they show a correspondence between learning \glspl{svm} with \gls{rbf} kernels and good regularization operators.
+As an explanation, they show a correspondence between learning \glspl{svm} with \gls{rbf} kernels and regularization operators.
 This may give insights in why \glspl{svm} have been found to exhibit high generalization ability (by learning with few training objects).
 
-The number of basis functions, the center parameters that correspond with support vectors, and the weights in the output layer are all automatically determined via the optimal hyperplane in features space $\mathcal{F}$ \cite{cherkassky2007learning}.
-The width parameter $\sigma$ is equal for all basis functions and is set a priori and determines the flexibility and complexity of the boundary.
+% The number of basis functions, the center parameters that correspond with support vectors, and the weights in the output layer are all automatically determined via the optimal hyperplane in features space $\mathcal{F}$ \cite{cherkassky2007learning}.
+% The center parameters that correspond with support vectors and the weights in the output layer are automatically determined via the optimal hyperplane in features space $\mathcal{F}$ \cite{cherkassky2007learning}.
+The width parameter $\sigma$ is equal for all kernel functions and is set a priori and determines the flexibility and complexity of the boundary.
 In \Cref{subsec:svm_model_parameters} this (hyper)parameter for an \gls{svm} is further discussed.
@@
 
-The constructed hypersphere is characterized by a center $\mathbf{a}$ and a radius $R > 0$ as distance from the center to (any data point that is a \gls{sv} on) the boundary, for which the volume, and thus the radius $R$, will be minimized.
+The constructed hypersphere is characterized by the center $\mathbf{a}$ and radius $R > 0$ as distance from the center to (any data point that is a \gls{sv} on) the boundary, for which the volume, and thus the radius $R$, will be minimized.
 The center $\mathbf{a}$ is a linear combination of the support vectors.
@@
 \begin{equation}\label{eq:svdd_objective}
-  L(R, \vectorsym{a}, \vectorsym{\xi}) = R^2 + C \sum_{i=1}^n \xi_i
+  L(R, \mathbf{a}, \vectorsym{\xi}) = R^2 + C \sum_{i=1}^n \xi_i
 \end{equation}
@@
 
-In the maximization of \Cref{eq:svdd_lagrange} a large fraction of the multipliers $\alpha_i$ become zero and for a small fraction $\alpha_i > 0$.
-This small fraction, for which $\alpha_i$ is non-zero, are called the \glspl{sv} and these objects lie on the boundary of the description.
+In the maximization of \Cref{eq:svdd_lagrange} a large fraction of the multipliers $\alpha_i$ become zero.
+The small fraction for which $\alpha_i > 0$ are called the \glspl{sv} and these objects lie on the boundary of the description.
 The center of the hypersphere only depends on this small number of \glspl{sv} and the objects for which $\alpha_i = 0$ can be discarded from the solution.
-Testing the membership of a (new) object $\vectorsym{z}$ is done by determining if the distance to the center $\vectorsym{a}$ of the sphere is equal or smaller to the radius $R$:
+Testing the membership of a (new) object $\vectorsym{z}$ is done by determining if the distance to the center $\mathbf{a}$ of the sphere is equal or smaller to the radius $R$:
 \begin{equation}\label{eq:svdd_test_object}
-  \lVert \vectorsym{z} - \vectorsym{a} \rVert ^2 = (\vectorsym{z} \cdot \vectorsym{z}) - 2 \sum_{i} \alpha_i(\vectorsym{z} \cdot \vectorsym{x}_i) + \sum_{i,j}(\vectorsym{x}_i \cdot \vectorsym{x}_j) \le R^2
+  \lVert \vectorsym{z} - \mathbf{a} \rVert ^2 = (\vectorsym{z} \cdot \vectorsym{z}) - 2 \sum_{i} \alpha_i(\vectorsym{z} \cdot \vectorsym{x}_i) + \sum_{i,j} \alpha_i \alpha_j (\vectorsym{x}_i \cdot \vectorsym{x}_j) \le R^2
 \end{equation}

### 2014-01-15 Chapters/Chapter4.tex
Decide on OCS-HTS. Working on algorithm in chapter 4

@@
 \lhead{Chapter 4. \emph{Application to Human Activity Time Series Data}} % This is for the header on each page - perhaps a shortened title
-
-\TODO{Decide on name}
-\begin{itemize}
-  \item Human Activity Support Vector Change Detection
-  \item Inertial Sensor Time Series Svm Temporal Segmentation
-  \item Support Vector bases Human Activity Temporal Segmentation
-  \item Human Activity Temporal Segmentation by Support Vector Machines: HATS-SVM
-  \item HATS with One-Class SVM: HATS-OCS, HATS-OS, HATSOS
-  \item OCS-HATS
-  \item HATS-SVDD
-\end{itemize}
 
@@
 
-This chapter will introduce our method and setup for the temporal segmentation of human activities, using an \gls{oc-svm} approach.
+This chapter will introduce our method, which we named \gls{ocs-hats}, and setup for the temporal segmentation of human activities, using an \gls{oc-svm} approach.
 It shows our method to apply a \gls{oc-svm} based temporal segmentation method to inertial sensor time series data, often used in the context of \acrlong{har}.

### 2014-01-15 Chapters/Chapter4/change_detection.tex
Decide on OCS-HTS. Working on algorithm in chapter 4

@@
 
+\begin{algorithm}
+\caption{\acrlong{ocs-hats} algorithm.}
+\label{alg:ocs-hats}
+\begin{algorithmic}[1]
+\Require time series $\vectorsym{x}$ and window size $n$
+\Ensure collection of change points $\vectorsym{cp}$
+
+\State $w \leftarrow \vectorsym{x}[1 \dots n-1]$ \Comment{Initialize working set}
+
+\While{$\vectorsym{x}_t$ is available}
+  \State $w[] \leftarrow \vectorsym{x}_t$ \Comment{Add new data to working set}
+
+  \State $s \leftarrow \operatorname*{SVDD}(\vectorsym{x})$ \Comment{Create hypersphere with SVDD method}
+  \State $r[] \leftarrow s.radius$ \Comment{Extract radius from hypersphere}
+
+  \State $r_{avg} \leftarrow \operatorname*{average}(r)$
+  \State $\hbar \leftarrow s.radius / r_{avg}$ \Comment{Radius ratio $\hbar$}
+
+  \If{ $\hbar$ \textless $th_{low}$ \textbf{or} $\hbar$ \textgreater $th_{high}$} \Comment{Compare $\hbar$ with thresholds}
+    \State $cp[] \leftarrow t$ \Comment{Mark $t$ as change point}
+  \EndIf
+
+  \State $w[0] \leftarrow []$ \Comment{Drop oldest data point from working set}
+\EndWhile
+
+\State $cp \leftarrow \operatorname*{merge}(cp)$ \Comment {Post-processing, change point merging}
+\end{algorithmic}
+\end{algorithm}
+
+\TODO{Make algorithm more readable}
+
 In the \gls{svcpd} the first step for a new data point $\vectorsym{z}$ is to check whether the data point lies within the enclosing (hyperspherical) boundary of the constructed model.
@@
 The second step is to inspect the (changed) model properties, \ie the radius $R$.
-In our method we only employ the second step: the inspection of the radius $R$.
+In \gls{ocs-hats} we only employ the second step: the inspection of the radius $R$.
 We regard the first step (determine if $\vectorsym{z}$ is an outlier) to be superfluent: in case it is an outlier, the model update step will change the radius $R$ in order to let the boundary include the new data point $\vectorsym{z}$.

### 2014-01-15 Chapters/Chapter4/data_gathering.tex
Decide on OCS-HTS. Working on algorithm in chapter 4

@@
 \subsection{Artificial data}\label{subsec:data_gathering_artificial}
-In order to provide an objective comparison to other methods, we will use artificial data sets which are also used in the earlier work on which our method is based.
+In order to provide an objective comparison to other methods, we will use artificial data sets which are also used in the earlier work on which \gls{ocs-hats} is based.
 These are the data sets used by Takeuchi and Yamanishi \cite{takeuchi2006unifying} and Camci \cite{camci2010change}.
@@
 \subsection{Real-world data}\label{subsec:data_gathering_real_world}
-In the second type of data sets we apply our method of change detection and temporal segmentation to real-world data sets.
+In the second type of data sets we apply our method for change detection and temporal segmentation to real-world data sets.
 For our setup we record the activities of humans performed both in- and outdoor in an unknown environment.
 Activities performed include sitting, standing, walking, running in a straight and curved line, and walking up- and downstairs.
-Our method uses the signals from the accelerometer, magnetic field, and rotation sensors.
+\gls{ocs-hats} uses the signals from the accelerometer, magnetic field, and rotation sensors.
 These time series data are used to detect change points.

### 2014-01-15 Chapters/Chapter6/data_sets.tex
Decide on OCS-HTS. Working on algorithm in chapter 4

@@
 \section{Data Sets}\label{sec:data_sets}
-In the previous chapter we have applied our method to the artificial data sets from Camci~\cite{camci2010change} and Takeuchi and Yamanishi~\cite{takeuchi2006unifying}.
+In the previous chapter we have applied our method \gls{ocs-hats} to the artificial data sets from Camci~\cite{camci2010change} and Takeuchi and Yamanishi~\cite{takeuchi2006unifying}.
 In this chapter we apply the method to real-world data sets.
 The data sets come from inertial sensor data recorded during human activities, such as walking, running, ascending and descending stairs, and standing still.
-To compare our results with other research, we have tried to apply our method to two commonly used data sets.
+To compare our results with other research, we have tried to apply \gls{ocs-hats} to two commonly used data sets.
 The first is the data from the \gls{wisdm} lab \cite{kwapisz2011activity}.

### 2014-01-15 Chapters/Chapter6/results.tex
Decide on OCS-HTS. Working on algorithm in chapter 4

@@
 \section{Results}\label{sec:real_world_results}
-In this section we will provide the results of our methods applied to the real-world data sets and discuss the quality of the solution.
+In this section we will provide the results of our methods \gls{ocs-hats} applied to the real-world data sets and discuss the quality of the solution.
 As stated above, we will use the objective quality metrics as used in \Cref{sec:artificial_data_quality_metrics}.

### 2014-01-15 Chapters/Chapter7.tex
Finished feedback Chapter 1
Decide on OCS-HTS. Working on algorithm in chapter 4

@@
 Since manual labeling (the annotation of change points on the time series data) of human activities is ambiguous we have put emphasis on the subjective inspection of the results.
-The proposed method gives fairly good results, although parameter tuning is a subject of furter research.
+The proposed method gives fairly good results, although parameter tuning is a subject of further research.
 
-Our method is based on and a slightly simplification of the \gls{svcpd} by Camci~\cite{camci2010change}.
+We have constructed \gls{ocs-hats} as a slightly simplification of \gls{svcpd} by Camci~\cite{camci2010change} application to real-world inertial sensor data.
 We have tested our proposed method with the same artificial data and did not find a deterioration in the results.
@@
 This allows for better interpretation of the results.
-All the recorded data is made publicity available~\cite{vlasveld2013continuous}.
+All the recorded data is made publicity available under the \gls{almende-data} data set \cite{vlasveld2014acras}.
 

### 2014-01-16 Chapters/Chapter4/change_detection.tex
Algorithm Chapter 4

--- 2014-01-14 Chapters/Chapter4/change_detection.tex
@@
 
-\begin{table}
-\begin{center}
-  \caption[Proposed algorithm]{Algorithm of the proposed change detection method, using the Ratio-Thresholding method of \ref{subsec:ratio_thresholding} for change detection.}
-  \begin{tabular}{ l p{12cm} }
-    \hline
-    Step & Action \\
-    \hline
-    & \textbf{Input:} time series $\vectorsym{x}$, window size $n$ \\
-    & \textbf{Output:} collection of change points $\vectorsym{t}$ \\
-    1 & Start with $n$ data objects as working set and construct hypersphere \\
-    2 & Add next data object $x_t$ and drop first one \\
-    3 & Identify new hyper-sphere and its \emph{approximate radius} \\
-    4 & Calculate radius average of hyperspheres since last change point \\
-    5 & Calculate radius ratio $\hbar$. \newline
-    If $\hbar$ is  lower than ${th}_{low}$ or greater than ${th}_{high}$ then mark $t$ as change point in $\vectorsym{t}$ \\
-    6 & Continue from step 2, until no more data objects available \\
-    7 & Apply post-processing: merge close change points \\
-    \hline
-  \end{tabular}
-  \label{tab:algorithm_proposed_method}
-\end{center}
-\end{table}
-
-\TODO{Algoritme meer in pseudo-code zetten. Prominenter maken.}
-
 In the \gls{svcpd} the first step for a new data point $\vectorsym{z}$ is to check whether the data point lies within the enclosing (hyperspherical) boundary of the constructed model.
@@
 The second step is to inspect the (changed) model properties, \ie the radius $R$.
-In our method we only employ the second step: the inspection of the radius $R$.
+In \gls{ocs-hats} we only employ the second step: the inspection of the radius $R$.
 We regard the first step (determine if $\vectorsym{z}$ is an outlier) to be superfluent: in case it is an outlier, the model update step will change the radius $R$ in order to let the boundary include the new data point $\vectorsym{z}$.
 Would the new data object $\vectorsym{z}$ be already an inlier, then the updated model would not have a significant change in radius $R$.
-Our algorithm is outlined in \Cref{tab:algorithm_proposed_method}.
+Our algorithm is outlined in \Cref{alg:ocs-hats}.
+
+\begin{algorithm}
+\caption{\acrlong{ocs-hats} algorithm.}
+\label{alg:ocs-hats}
+\begin{algorithmic}[1]
+\Require time series $X$, window size $n$, thresholds $th_{low}$ and $th_{high}$, proximity factor $p$
+\Ensure collection $change\_points$
+\item[]
+\State $change\_points \leftarrow$ empty collection
+\State $radii \leftarrow$ empty collection
+\State Initialize $w \leftarrow incremental\_svdd$ with first $n$ samples from $\vectorsym{x}$
+\item[]
+\While{ new sample $x$ from $X$ at time $t$}
+  \item[]
+  \State Add $x$ to $w$
+  \State $sphere \leftarrow update\_hypersphere(w)$
+  \State $radius \leftarrow extract\_radius(sphere)$
+  \State $radii[t] \leftarrow radius$
+  \item[]
+  \State $\hbar \leftarrow$ calculate ratio from previous average $radii$ and current $radius$
+  \item[]
+  \If{ $\hbar$ \textless \, $th_{low}$ \textbf{or} $\hbar$ \textgreater \, $th_{high}$} %\Comment{Compare $\hbar$ with thresholds}
+    \State Add current time $t$ to $change\_points$
+    \State Empty previous $radii$
+  \EndIf
+  \item[]
+  \State Remove oldest $x$ from $w$
+\EndWhile
+
+\item[]
+
+\State Merge $\vectorsym{cp}$ based on proximity $p$
+\end{algorithmic}
+\end{algorithm}
 

### 2014-01-16 Chapters/Chapter5/quality_metrics.tex
Change lines in results plots

@@
 \end{equation}
-Besides the delay, which should be low to indicate discovered change points near the actual change point, we need to penalize discovered change points which do not relate to a true change point.
+Besides the delay, and the standard deviation $std(d_{avg})$ to measure the spread, which should be low to indicate discovered change points near the actual change point, we need to penalize discovered change points which do not relate to a true change point.
 For this we use the \acrlong{far} $\gamma$, expressed as

### 2014-01-16 Chapters/Chapter5/results.tex
Put captions above tables
WIP: chapter 5 and 6
Change lines in results plots

@@
   \centering
+  \caption[Results artificial data sets]{Parameter settings and results of the artificial data sets.}
   \begin{tabulary}{\textwidth}{|l|c|c|c|c|}
@@
     \hline
-    $\operatorname*{far}(Y)$ & 0 & 0.1 & 0.1 & 0.1 \\
+    $\gamma(Y)$ & 0 & 0.1 & 0.1 & 0.1 \\
     \hline
-    $\operatorname*{Average\_delay}$ & 2.27 & 6.18 & 0.64 & 15.36 \\
+    $d_{avg}$ & 2.27 & 6.18 & 0.64 & 15.36 \\
     \hline
-    Standard deviation Delay & 1.19 & 7.94 & 0.92 & 17.70 \\
+    $std(d_{avg})$ & 1.19 & 7.94 & 0.92 & 17.70 \\
     \hline
   \end{tabulary}
-  \caption[Results artificial data sets]{Parameter settings and results of the artificial data sets.}
   \label{tab:results_artificial}
@@
 With each data set embodying different characteristics, different observations can be drawn from the data sets.
-In the references figures, the blue lines represent the radius of the constructed hypersphere.
-The vertical red lines are the discovered change points.
+In the references figures, the solid blue line represent the radius of the constructed hypersphere.
+The vertical dashed and thick red lines are the discovered change points.
 For each data set the following is observed:
@@
   \item \textbf{Fixed increasing mean:} In \Cref{fig:camci_fixed_increasing_mean_thresholds} we see that each change point is almost equally fast and accurate detected.
-  This implies that the relative difference between the means of the Gaussian distribution has no effect on the detectability of the algorithm.
+  This implies that the relative difference between the means of the Gaussian distribution has little effect on the detectability of the algorithm.
   \item \textbf{Reduced increasing mean:} \Cref{fig:takeuchi_reduced_increasing_mean_thresholds} shows that the absolute difference between the means of the Gaussian distribution is a certain influence to the detectability.
@@
 
-  \caption[Artificial data set results]{Hypersphere radius sizes for the four artificial data sets. The detected change points are displayed as red vertical bars.}
+  \caption[Artificial data set results]{Hypersphere radius sizes for the four artificial data sets. The detected change points are displayed as dashed thick red vertical bars.}
   \label{fig:plots_results_artificial_data_sets}

### 2014-01-16 Chapters/Chapter6/results.tex
Change lines in results plots

@@
 As stated above, we will use the objective quality metrics as used in \Cref{sec:artificial_data_quality_metrics}.
-That will be in the form of a tabular overview of the (manually set) best meta-parameters and results, expressed in \gls{far} and $\operatorname*{Average\_delay}$ (following \Cref{sec:artificial_data_results}).
+That will be in the form of a tabular overview of the (manually set) best meta-parameters and results, expressed in \gls{far} $\gamma$ and average delay $d_{avg}$ (following \Cref{sec:artificial_data_results}).
 Furthermore, we will discuss the results from an empirical point of view, since the problem of finding change points in human activities is ambiguous.
@@
     \hline
-    $\operatorname*{far}(Y)$ & 0.1 & 0.05 & 0 & 0 & 0.2 & 0 & 0.48 \\
+    $\gamma(Y)$ & 0.1 & 0.05 & 0 & 0 & 0.2 & 0 & 0.48 \\
     \hline
-    $\operatorname*{Average\_delay}$ & 0.52 & 0.83 & 0.47 & 0.92 & 1.26 & 1.05 & 0.92 \\
+    $d_{avg}$ & 0.52 & 0.83 & 0.47 & 0.92 & 1.26 & 1.05 & 0.92 \\
     \hline
-    STD Delay & 0.29 & 0.55 & 0.52 & 0.75 & 1.67 & 0.78 & 0.92 \\
+    $std(d_{avg})$ & 0.29 & 0.55 & 0.52 & 0.75 & 1.67 & 0.78 & 0.92 \\
     \hline

### 2014-01-16 Chapters/Chapter6/data_gathering.tex
Chapter 6 run order

@@
 The first type (Run 1 and Run 2) consists of various outdoor recordings with segments of walking, running and standing.
-For the second type (Run 4, 5, and 6) some of the same activities are performed in a straight line, whilst others are performed around corners or in a wide circular direction.
-The third type (Run 8) consists mainly of walking up and down circular shaped stairs, with small horizontal parts between the segments.
-The last type is for Run 3, in which we walked in a straight line, took a $90^{\circ}$ counter-clockwise turn (around the corner) and continued walking.
+For the second type (Run 3, 4, and 5) some of the same activities are performed in a straight line, whilst others are performed around corners or in a wide circular direction.
+The third type (Run 6) consists mainly of walking up and down circular shaped stairs, with small horizontal parts between the segments.
+The last type is for Run 7, in which we walked in a straight line, took a $90^{\circ}$ counter-clockwise turn (around the corner) and continued walking.
 A more detailed description of the performed activities follows.
@@
 Between the stairs the subject walked in a hallway with a $180^{\circ}$ counter-clockwise turn, as shown in \Cref{fig:stills_subject_3_walking}.
-The recorded sensor data of the performed activities are visible in the plots of \Cref{fig:data_gathering_run_3_rot,fig:data_gathering_run_3_mag,fig:data_gathering_run_3_acc}.
+The recorded sensor data of the performed activities are visible in the plots of \Cref{fig:data_gathering_run_7_rot,fig:data_gathering_run_7_mag,fig:data_gathering_run_7_acc}.
 
@@
     \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/stairs-1-marc/data_plot_rot_annotated.eps}
-    \caption{Run 3: Indoor stairs Marc, rotation}
-    \label{fig:data_gathering_run_3_rot}
+    \caption{Run 7: Indoor stairs subject 3, rotation}
+    \label{fig:data_gathering_run_7_rot}
   \end{subfigure} \\
@@
     \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/stairs-1-marc/data_plot_mag_annotated.eps}
-    \caption{Run 3: Indoor stairs Marc, Mag}
-    \label{fig:data_gathering_run_3_mag}
+    \caption{Run 7: Indoor stairs subject 3, Mag}
+    \label{fig:data_gathering_run_7_mag}
   \end{subfigure} \\
@@
     \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/stairs-1-marc/data_plot_acc_annotated.eps}
-    \caption{Run 3: Indoor stairs Marc, accelerometer}
-    \label{fig:data_gathering_run_3_acc}
+    \caption{Run 7: Indoor stairs subject 3, accelerometer}
+    \label{fig:data_gathering_run_7_acc}
   \end{subfigure}

### 2014-01-16 Chapters/Chapter6/data_sets.tex
Chapter 6 run order

@@
 The manually determined change points are used to draw quality conclusion over the algorithm's performance, by comparing it to the discovered change points.
-All of the used data sets, including the video material and manual labeling, are made available to the public as \gls{almende-data} \cite{vlasveld2014acras} for further research.
+All of the used data sets, including the video material and manual labeling, are made available to the public as the \acrlong{almende-data} \cite{vlasveld2014acras} for further research.
 The following section discusses the performed activities during the recordings.

### 2014-01-16 Chapters/Chapter6/results.tex
Chapter 6 run order

@@
 We have used a window length of 50 frames, which represents around $1.16$ seconds of data (the frame-rate of the recordings is not fixed).
-The other parameter values, \ie the \gls{rbf} $\sigma$, high and low threshold and closeness factor, are the result of an educated guess and manual optimization.
+The other parameter values, \ie the \gls{rbf} $\sigma$, high and low threshold, and proximity parameter $\delta$, are the result of an educated guess and manual optimization.
 In general, a high value for the upper threshold and low value for the lower threshold indicates a robust method, since it indicates that change points are significantly different from homogeneous segments.
-The closeness time frame should be low, since that indicates there are few false positives.
+The proximity $\delta$ should be low, since that indicates there are few false positives.
 
@@
     \hline
-    Sigma of \gls{rbf} & 13 & 13 & 13 & 13 & 13 & 13 & 4 \\
+    Sigma of \gls{rbf} & 13 & 13 & 13 & 13 & 13 & 4 & 13 \\
     \hline
-    High threshold & 1.2 & 1.5 & 1.3 & 1.1 & 1.3 & 1.2 & 1.7 \\
+    High threshold & 1.2 & 1.5 & 1.1 & 1.3 & 1.2 & 1.7 & 1.3 \\
     \hline
-    Low threshold & 0.8 & 0.8 & 0.6 & 0.7 & 0.7 & 0.8 & 0.6 \\
+    Low threshold & 0.8 & 0.8 & 0.7 & 0.7 & 0.8 & 0.6 & 0.6 \\
     \hline
-    Closeness (s) & 0.7 & 0.85 & 1 & 0.6 & 0.7 & 1 & 0.8 \\
+    $\delta$ (s) & 0.7 & 0.85 & 0.6 & 0.7 & 1 & 0.8 & 1 \\
     \hline
     \hline
-    $\gamma(Y)$ & 0.1 & 0.05 & 0 & 0 & 0.2 & 0 & 0.48 \\
+    $\gamma(Y)$ & 0.1 & 0.05 & 0 & 0.2 & 0 & 0.48 & 0 \\
     \hline
-    $d_{avg}$ & 0.52 & 0.83 & 0.47 & 0.92 & 1.26 & 1.05 & 0.92 \\
+    $d_{avg}$ & 0.52 & 0.83 & 0.92 & 1.26 & 1.05 & 0.92 & 0.47 \\
     \hline
-    $std(d_{avg})$ & 0.29 & 0.55 & 0.52 & 0.75 & 1.67 & 0.78 & 0.92 \\
+    $std(d_{avg})$ & 0.29 & 0.55 & 0.75 & 1.67 & 0.78 & 0.92 & 0.52 \\
     \hline
@@
 \begin{itemize}
-  \item \textbf{Merging:} in our method we use a \emph{closeness} time period $t_c$ (usually up to $0.5$ - $1.0$ seconds) to merge discovered change points that have a very small distance between them.
-  Multiple merging-strategies can be applied, and in our method we use the naive implementation by simply ignoring all the discovered change points that occur less than $t_c$ seconds after the previous change point.
+  \item \textbf{Merging:} in our method we use a proximity time period $\delta$ (usually up to $0.5$ - $1.0$ seconds) to merge discovered change points that have a very small distance between them.
+  Multiple merging-strategies can be applied, and in our method we use the naive implementation by simply ignoring all the discovered change points that occur less than $\delta$ seconds after the previous change point.
   This works well when there is a block of noisy data with a high amount of (falsely) discovered change points.
@@
     \end{itemize}
-  \item \textbf{Subject 2, walk around corner} \Cref{fig:data_with_cps_run_3}
-    \begin{itemize}
-      \item The $90^{\circ}$ counter-clockwise turn during the walking activity is hard to discover when the accelerometer sensor data is included.
-      When only the magnetic field and rotation sensors are used, the turn requires a lower sensitivity.
-      With only these two sensors all the other change points in this run are also successfully discovered.
-    \end{itemize}
-  \item \textbf{Subject 2, walk and run around fountain} \Cref{fig:data_with_cps_run_4}
+
+  \item \textbf{Subject 2, walk and run around fountain} \Cref{fig:data_with_cps_run_3}
     \begin{itemize}
@@
       The difference for the rotational vectors need to accumulate to a certain value before they have enough influence to let the rotation be regarded as a change point.
-
     \end{itemize}
-  \item \textbf{Subject 1, walk and run around fountain} \Cref{fig:data_with_cps_run_5}
+  \item \textbf{Subject 1, walk and run around fountain} \Cref{fig:data_with_cps_run_4}
     \begin{itemize}
@@
     \end{itemize}
-  \item \textbf{Subject 2, walk and run fountain 2} \Cref{fig:data_with_cps_run_6}
+  \item \textbf{Subject 2, walk and run fountain 2} \Cref{fig:data_with_cps_run_5}
     \begin{itemize}
@@
     \end{itemize}
-  \item \textbf{Subject 3, indoor stairs} \Cref{fig:data_with_cps_run_7}
+  \item \textbf{Subject 3, indoor stairs} \Cref{fig:data_with_cps_run_6}
     \begin{itemize}
       \item The walking segment around $22s$, between two segments of walking downstairs, shows little difference in the data.
-      To recognize it as a change point a high sensitivity and low closeness time period $t_c$ is required.
+      To recognize it as a change point a high sensitivity and low proximity time period $\delta$ is required.
       \item During some segments (downstairs from $24s$, upstairs from $42s$ and $54s$) the method discovers more change points than our annotation.
@@
       The delay between these segments seems to be larger, as illustrated around $33s$.
+    \end{itemize}
+  \item \textbf{Subject 2, walk around corner} \Cref{fig:data_with_cps_run_7}
+    \begin{itemize}
+      \item The $90^{\circ}$ counter-clockwise turn during the walking activity is hard to discover when the accelerometer sensor data is included.
+      When only the magnetic field and rotation sensors are used, the turn requires a lower sensitivity.
+      With only these two sensors all the other change points in this run are also successfully discovered.
     \end{itemize}
@@
   \begin{subfigure}{1\textwidth}
-    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/run-3-walk-turn-roemer/data_plot_acc_with_discovered_cps.eps}
+    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/run-4-run-fountain-roemer/data_plot_acc_with_discovered_cps.eps}
     \caption{Run 3}
@@
   \begin{subfigure}{1\textwidth}
-    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/run-4-run-fountain-roemer/data_plot_acc_with_discovered_cps.eps}
+    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/run-5-run-fountain-jos/data_plot_acc_with_discovered_cps.eps}
     \caption{Run 4}
@@
   \begin{subfigure}{1\textwidth}
-    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/run-5-run-fountain-jos/data_plot_acc_with_discovered_cps.eps}
+    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/run-6-walk-run-roemer/data_plot_acc_with_discovered_cps.eps}
     \caption{Run 5}
@@
   \begin{subfigure}{1\textwidth}
-    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/run-6-walk-run-roemer/data_plot_acc_with_discovered_cps.eps}
+    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/stairs-1-marc/data_plot_acc_with_discovered_cps.eps}
     \caption{Run 6}
@@
   \begin{subfigure}{1\textwidth}
-    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/stairs-1-marc/data_plot_acc_with_discovered_cps.eps}
+    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/run-3-walk-turn-roemer/data_plot_acc_with_discovered_cps.eps}
     \caption{Run 7}

### 2014-01-17 Appendices/AppendixA-Results.tex
Move results Chapter 6 to Appendix'

--- 2014-01-16 Chapters/Chapter6/results.tex
@@
-% !TEX root = ../../main.tex
-\section{Results}\label{sec:real_world_results}
-In this section we will provide the results of our methods \gls{ocs-hats} applied to the real-world data sets and discuss the quality of the solution.
-As stated above, we will use the objective quality metrics as used in \Cref{sec:artificial_data_quality_metrics}.
-That will be in the form of a tabular overview of the (manually set) best meta-parameters and results, expressed in \gls{far} $\gamma$ and average delay $d_{avg}$ (following \Cref{sec:artificial_data_results}).
-Furthermore, we will discuss the results from an empirical point of view, since the problem of finding change points in human activities is ambiguous.
-By stating our observations and remarks and providing fragments of visual results, we will discuss the characterizing strengths and weaknesses of our method.
+% !TEX root = ../main.tex
+% Appendix A
 
-\subsection{Objective metrics}
-The employed two-stage algorithm requires the setting of user defined parameters.
-In \Cref{tab:results_real_world} for each run the results are listed, together with the used parameters.
-We have used a window length of 50 frames, which represents around $1.16$ seconds of data (the frame-rate of the recordings is not fixed).
-The other parameter values, \ie the \gls{rbf} $\sigma$, high and low threshold, and proximity parameter $\delta$, are the result of an educated guess and manual optimization.
-In general, a high value for the upper threshold and low value for the lower threshold indicates a robust method, since it indicates that change points are significantly different from homogeneous segments.
-The proximity $\delta$ should be low, since that indicates there are few false positives.
+\chapter{Observations for Real-World Results} % Main appendix title
+
+\label{AppendixA} % For referencing this appendix elsewhere, use~\ref{AppendixA}
+
+\lhead{Appendix A. \emph{Real-World Observations}} % This is for the header on each page - perhaps a shortened title
 
@@
 
-To give a better impression on the quality of the $\operatorname*{Average\_delay}$, \Cref{fig:boxplot_real_world_runs} shows a box plot for each run.
-It shows the spread of the $\operatorname*{Average\_delay}$ in seconds over all the change points.
-A lower and more compact box plot is better.
-
 \begin{figure}
@@
   \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/boxplot_results_real_world_runs.eps}
-  \caption[Box plot results real-world runs]{Box plot of the results for the real-world runs, indicating the number of data points between the actual and closest detected change points.
+  \caption[Box plot results real-world runs individually optimized]{Box plot of the results for the real-world runs, indicating the number of data points between the actual and closest detected change points.
 A lower and more compact box plot is better.}
@@
 
-\subsection{Subjective observations and remarks}
-It can be ambiguous where to place the (manually annotated) change points, in the case of real-world data.
-The objective quality measure can give a distorted view on the performance.
-To overcome this, we will state observations and remarks found after close (visual) inspection of the annotated and discovered change points.
-In \Cref{fig:plots_all_runs_results} all the results are plotted.
-The black solid lines indicate our manually determined change points.
-The dashed purple lines indicate the change points as discovered by our method.
-We will start with general observations applicable to all the runs and will continue with individual remarks for each run.
-Where possible, a graphical illustration will support our observations.
+This appendix provides observations and remarks which we made from inspecting the real-world results from \Cref{Chapter6}.
+In \Cref{subsec:subjective_results} we have discussed a few overall effects.
+In the following we discuss the performance of each run, using the (manually determined) specific parameters as listed in \Cref{tab:results_real_world}.
 
-Overall our impression is that the proposed method is fairly good at recognizing change points.
-Depending on the parameters for the sensitivity adjustment, almost all changes are discovered.
-With a high sensitivity it seems, initially, the method produces many false positives.
-After closer inspecting of the data, we can confirm that there is a change in the data, although the performed activity stayed the same.
-This is also a weakness of the method: since the sensitivity is set globally, it can produce many semi-false positive change points for some segments of a run.
+To give a better impression on the quality of the $d_{avg}$, \Cref{fig:boxplot_real_world_runs} shows a box plot for each run.
+It shows the spread of the $d_{avg}$ in seconds over all the change points.
+A lower and more compact box plot is better.
 
-An other positive observation is that the method often recognizes a change point \emph{before} our annotation.
-Again, after closer inspection we can see that the data changes shape and distribution before we annotated a change based on the video recordings.
-This effect can be observed in the case of a transition from running to walking; the last `steps' of running are already recognized of being different.
-As with the above behavior, this is subject to parameter settings.
+When looking at the individual runs, we have the following observations for each run:
 
-Considering the overall performance and results, we have the following observations of cases in which the method has problems with detecting change:
-\begin{itemize}
-  \item \textbf{Merging:} in our method we use a proximity time period $\delta$ (usually up to $0.5$ - $1.0$ seconds) to merge discovered change points that have a very small distance between them.
-  Multiple merging-strategies can be applied, and in our method we use the naive implementation by simply ignoring all the discovered change points that occur less than $\delta$ seconds after the previous change point.
-  This works well when there is a block of noisy data with a high amount of (falsely) discovered change points.
-  On the flip side, this method will ignore real new change points which occur during the noisy period.
-  This problem on itself can be formulated as a change detection problem.
-  \item \textbf{Masking:} Since the parameters for the detection method are set globally, it is difficult to discover all the change points (and only the change points) without a high \gls{far}.
-  If a change point is proceeded by a noisy block, then the real change point will be merged with the previous change points.
-  In an other case, where a change point is very clear represented in the data but the next, close by, other change point is more subtle, the latter change point is \emph{masked} by the former.
-  This masking effect is also discussed in \cite{inclan1994use}, where an iterated approach is applied.
-  \item \textbf{False heterogeneity:} For our video-data synchronization, we started and ended each recording with a few seconds in which the recording smartphone was kept still in the air.
-  During this period, the data variance, and thus the constructed hypersphere, becomes very small.
-  As a results, even small movements are considered to be changes and in the case of the mid-air still smartphone a lot of false change points are detected.
-  Due to the merging effect described above, the final real change point (often shaking the smartphone) is not discovered.
-  \item \textbf{Incorrect weighting:} In our analyses we have used the data from the accelerometer, magnetic field, and rotation sensors.
-  In the segments which embodied movement in a circular manner (such as walking and running around a fountain), the turn was not (or at a different time point) discovered.
-  Alternatively, when we only used the magnetic field and rotation sensors, the turn was correctly discovered.
-  Other transitions, such as from walking to running, are harder to discover without the (linear) accelerometer sensor data.
-\end{itemize}
-
-When looking at the individual runs, we have the following observations for each run number:
 \begin{enumerate}
@@
 \end{figure}
-
-\section{Possible improvements}\label{sec:possible_improvements}
-From the above section we can extract a few possible improvements for the algorithm.
-Most of the observations and remarks are due to a different required sensitivity for local segments of the run.
-Currently, our method employs a global setting for the thresholds.
-A possible improvement would be to use locally optimized parameters, although that would require an reflective data processing approach or more a priori information about the data.
-With our approach, the lack of a priori knowledge about the data distribution is one of the design decisions.
-
-Furthermore, we discovered that sometimes, especially with movements incorporating circular motions, the addition of (linear) accelerometer data makes the discovery of change points harder.
-In such cases, each of the inertial sensor data streams should be relatively weighted (\eg using the covariance) in order to signal an overall change when it only occurs in one of the streams.

### 2014-01-17 Chapters/Chapter1.tex
Move results Chapter 6 to Appendix'

@@
 It relates the properties of the constructed \gls{oc-svm} models to change detection in time series data.
-That relation is applied to our problem formulation in \Cref{Chapter4}, where we construct our change detection method \TODO{add name}, based on the work of Camci~\cite{camci2010change}.
+That relation is applied to our problem formulation in \Cref{Chapter4}, where we construct our change detection method \gls{ocs-hats}~\cite{vlasveld2014hats}, based on the work of Camci~\cite{camci2010change}.
 In \Cref{Chapter5} we apply the method to artificial data and show the objective performance.

### 2014-01-17 Chapters/Chapter4.tex
Move results Chapter 6 to Appendix'

@@
 
-This chapter will introduce our method, which we named \gls{ocs-hats}, and setup for the temporal segmentation of human activities, using an \gls{oc-svm} approach.
+This chapter will introduce our method, which we named \gls{ocs-hats}~\cite{vlasveld2014hats}, and setup for the temporal segmentation of human activities, using an \gls{oc-svm} approach.
 It shows our method to apply a \gls{oc-svm} based temporal segmentation method to inertial sensor time series data, often used in the context of \acrlong{har}.

### 2014-01-17 Chapters/Chapter6/results.tex
Move results Chapter 6 to Appendix'

@@
 That will be in the form of a tabular overview of the (manually set) best meta-parameters and results, expressed in \gls{far} $\gamma$ and average delay $d_{avg}$ (following \Cref{sec:artificial_data_results}).
-Furthermore, we will discuss the results from an empirical point of view, since the problem of finding change points in human activities is ambiguous.
-By stating our observations and remarks and providing fragments of visual results, we will discuss the characterizing strengths and weaknesses of our method.
+Furthermore, we have anaylized the results from an emperical point of view, since the problem of finding change points in human activities is ambiguous.
+For the sake of brevity we excluded those observations from this chapter and instead xthey can be found in \Cref{AppendixA}.
 
@@
 The employed two-stage algorithm requires the setting of user defined parameters.
-In \Cref{tab:results_real_world} for each run the results are listed, together with the used parameters.
+In \Cref{tab:results_final_real_world} for each run the results are listed.
+The right-most column summarizes the overall performance by taking the average over all the runs.
 We have used a window length of 50 frames, which represents around $1.16$ seconds of data (the frame-rate of the recordings is not fixed).
@@
 
+To give a better impression on the quality of the $d_{avg}$, \Cref{fig:boxplot_final_real_world_runs} shows a box plot for each run.
+It shows the spread of the $d_{avg}$ in seconds over all the change points.
+A lower and more compact box plot is better.
+
 \begin{table}
   \centering
-  \caption[Results real world runs]{Parameter settings and results of the real-world data sets.}
-  \begin{tabulary}{\textwidth}{|l|c|c|c|c|c|c|c|}
-    \cline{2-8}
-    \multicolumn{1}{l|}{} & Run 1 & Run 2 & Run 3 & Run 4 & Run 5 & Run 6 & Run 7 \\
+  \caption[Results real world runs]{Results of the real-world data sets. The used parameters are: window length = $50$, $\sigma = 13$, $th_{high} = 1.2$, $th_{low} = 0.8$}
+  \begin{tabulary}{\textwidth}{|l|c|c|c|c|c|c|c||c|}
+    \cline{2-9}
+    \multicolumn{1}{l|}{} & Run 1 & Run 2 & Run 3 & Run 4 & Run 5 & Run 6 & Run 7 & Average \\
     \hline
-    Window length & 50 & 50 & 50 & 50 & 50 & 50 & 50 \\
+    $\gamma(Y)$ & 0.1 & 0.2 & 0.14 & 0.37 & 0.29 & 0.64 & 0.57 & 0.33 \\
     \hline
-    Sigma of \gls{rbf} & 13 & 13 & 13 & 13 & 13 & 4 & 13 \\
+    $d_{avg}$ & 0.52 & 0.85 & 0.93 & 0.54 & 0.69 & 0.51 & 0.32 & 0.62\\
     \hline
-    High threshold & 1.2 & 1.5 & 1.1 & 1.3 & 1.2 & 1.7 & 1.3 \\
-    \hline
-    Low threshold & 0.8 & 0.8 & 0.7 & 0.7 & 0.8 & 0.6 & 0.6 \\
-    \hline
-    $\delta$ (s) & 0.7 & 0.85 & 0.6 & 0.7 & 1 & 0.8 & 1 \\
-    \hline
-    \hline
-    $\gamma(Y)$ & 0.1 & 0.05 & 0 & 0.2 & 0 & 0.48 & 0 \\
-    \hline
-    $d_{avg}$ & 0.52 & 0.83 & 0.92 & 1.26 & 1.05 & 0.92 & 0.47 \\
-    \hline
-    $std(d_{avg})$ & 0.29 & 0.55 & 0.75 & 1.67 & 0.78 & 0.92 & 0.52 \\
+    $std(d_{avg})$ & 0.71 & 0.31 & 0.68 & 0.37 & 0.50 & 0.44 & 0.35 & 0.48 \\
     \hline
   \end{tabulary}
-  \label{tab:results_real_world}
+  \label{tab:results_final_real_world}
 \end{table}
-
-To give a better impression on the quality of the $\operatorname*{Average\_delay}$, \Cref{fig:boxplot_real_world_runs} shows a box plot for each run.
-It shows the spread of the $\operatorname*{Average\_delay}$ in seconds over all the change points.
-A lower and more compact box plot is better.
 
@@
 \centering
-  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/boxplot_results_real_world_runs.eps}
+  \includegraphics[width=1\textwidth]{./Figures/chapter6/data_collection/boxplot_final_runs.eps}
   \caption[Box plot results real-world runs]{Box plot of the results for the real-world runs, indicating the number of data points between the actual and closest detected change points.
 A lower and more compact box plot is better.}
-  \label{fig:boxplot_real_world_runs}
+  \label{fig:boxplot_final_real_world_runs}
 \end{figure}
 
-\subsection{Subjective observations and remarks}
+\subsection{Subjective observations and remarks}\label{subsec:subjective_results}
 It can be ambiguous where to place the (manually annotated) change points, in the case of real-world data.
@@
 The dashed purple lines indicate the change points as discovered by our method.
-We will start with general observations applicable to all the runs and will continue with individual remarks for each run.
+We will start with general observations applicable to all the runs.
+A more detailed discussion for each run can be found in \Cref{AppendixA}.
 Where possible, a graphical illustration will support our observations.
@@
 
-When looking at the individual runs, we have the following observations for each run number:
-\begin{enumerate}
-  \item \textbf{Subject 2, straight walk and run} \Cref{fig:data_with_cps_run_1}
-    \begin{itemize}
-      \item The first transition from running to walking, around $14s$ is harder to discover than the transition for the same activities around $33s$.
-      This shows is that in real-world applications there is a diversity between the same transitions and activities.
-      \item Around $8s$, $9s$, and $16s$ a few steps (while walking) are regarded as change points.
-      During the running segments from $17s$ and $28s$ there is a lower probability of change for each step.
-    \end{itemize}
-  \item \textbf{Subject 1, straight walk and run} \Cref{fig:data_with_cps_run_2}
-    \begin{itemize}
-      \item Following the video recordings, we have annotated a change point from running to walking around $37s$.
-      Our method discovers a change point almost a second before.
-      In retrospect, we can see that the data distribution indeed changes from the discovered change point on.
-      This shows us two important principles.
-      The first is that the annotations are very subjective.
-      The second is that between different activities the transition period is longer than we would think.
-      Looking at the data, we can see that the body slows down, even before we visually notice it on the video recordings.
-    \end{itemize}
-
-  \item \textbf{Subject 2, walk and run around fountain} \Cref{fig:data_with_cps_run_3}
-    \begin{itemize}
-      \item Like in the first run, the walking segment from $24s$ results in a change point for each step.
-      Further inspection of the data reveals that each step is indeed different from the other.
-      Due to the global parameter settings, the sensitivity is too high for this segment to recognize it as one.
-      It could also be that our used window-length is too small, to model a good representation of this segment.
-      \item During the circular run, from $12s$ till $24$, there are two change points discovered.
-      The difference for the rotational vectors need to accumulate to a certain value before they have enough influence to let the rotation be regarded as a change point.
-    \end{itemize}
-  \item \textbf{Subject 1, walk and run around fountain} \Cref{fig:data_with_cps_run_4}
-    \begin{itemize}
-      \item As with the other runs, the accelerometer data makes it harder to detect turns.
-      It requires a higher sensitivity, which results in a higher \gls{far}.
-    \end{itemize}
-  \item \textbf{Subject 2, walk and run fountain 2} \Cref{fig:data_with_cps_run_5}
-    \begin{itemize}
-      \item During the standing segment around $38s$ there are a lot of false positives.
-      It seems to be the \emph{false heterogeneity} problem described above.
-    \end{itemize}
-  \item \textbf{Subject 3, indoor stairs} \Cref{fig:data_with_cps_run_6}
-    \begin{itemize}
-      \item The walking segment around $22s$, between two segments of walking downstairs, shows little difference in the data.
-      To recognize it as a change point a high sensitivity and low proximity time period $\delta$ is required.
-      \item During some segments (downstairs from $24s$, upstairs from $42s$ and $54s$) the method discovers more change points than our annotation.
-      A closer inspecting of the raw data reveals indeed changes in behavior.
-      To exclude these (semi) false positives, a better tuning of parameters is required.
-      \item Because of the circular shape of the stairs, the magnetic field sensors constantly differs.
-      Although our method is build to exclude slow shifting changes (because we are only interested in sudden changes), with our used window width it still eventually results in change points.
-      \item The difference between taking the stairs and walking is smaller than, \eg, walking and running.
-      The delay between these segments seems to be larger, as illustrated around $33s$.
-    \end{itemize}
-  \item \textbf{Subject 2, walk around corner} \Cref{fig:data_with_cps_run_7}
-    \begin{itemize}
-      \item The $90^{\circ}$ counter-clockwise turn during the walking activity is hard to discover when the accelerometer sensor data is included.
-      When only the magnetic field and rotation sensors are used, the turn requires a lower sensitivity.
-      With only these two sensors all the other change points in this run are also successfully discovered.
-    \end{itemize}
-\end{enumerate}
-
-\begin{figure}
-  \centering
-  \begin{subfigure}{1\textwidth}
-    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/run-1-walk-run-roemer/data_plot_acc_with_discovered_cps.eps}
-    \caption{Run 1}
-    \label{fig:data_with_cps_run_1}
-  \end{subfigure}
-
-  \begin{subfigure}{1\textwidth}
-    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/run-2-walk-run-jos/data_plot_acc_with_discovered_cps.eps}
-    \caption{Run 2}
-    \label{fig:data_with_cps_run_2}
-  \end{subfigure}
-
-  \begin{subfigure}{1\textwidth}
-    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/run-4-run-fountain-roemer/data_plot_acc_with_discovered_cps.eps}
-    \caption{Run 3}
-    \label{fig:data_with_cps_run_3}
-  \end{subfigure}
-
-  \begin{subfigure}{1\textwidth}
-    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/run-5-run-fountain-jos/data_plot_acc_with_discovered_cps.eps}
-    \caption{Run 4}
-    \label{fig:data_with_cps_run_4}
-  \end{subfigure}
-
-  \begin{subfigure}{1\textwidth}
-    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/run-6-walk-run-roemer/data_plot_acc_with_discovered_cps.eps}
-    \caption{Run 5}
-    \label{fig:data_with_cps_run_5}
-  \end{subfigure}
-
-  \begin{subfigure}{1\textwidth}
-    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/stairs-1-marc/data_plot_acc_with_discovered_cps.eps}
-    \caption{Run 6}
-    \label{fig:data_with_cps_run_6}
-  \end{subfigure}
-
-  \begin{subfigure}{1\textwidth}
-    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/run-3-walk-turn-roemer/data_plot_acc_with_discovered_cps.eps}
-    \caption{Run 7}
-    \label{fig:data_with_cps_run_7}
-  \end{subfigure}
-
-  \caption[Results run 1/8]{Annotated plots of all runs.
-The black vertical lines are manually determined change points, the purple dashes lines are the discovered change points.
-The label above each segment indicates the current performed activity.}\label{fig:plots_all_runs_results}
-\end{figure}
+An more elaborate discussion with observations and remarks is provided in \Cref{AppendixA}.
 

### 2014-01-17 Chapters/Chapter7.tex
Move results Chapter 6 to Appendix'

@@
 
-In this research we have applied \gls{occ} methods to accelerometer data, recorded during the performance of various human activities, in order to create a temporal segmentation of those recordings.
+This research has applied \gls{occ} methods to inertial sensor data, recorded during the performance of various human activities, in order to create a temporal segmentation of those recordings.
 The goal was to find the change points between activities, assuming that it will aid our understanding of the sensor values.
 To do so, we used the \gls{svdd} algorithm by Tax~\cite{tax1999support}, following the method of applying it to a sliding window of data, as done by Camci~\cite{camci2010change}.
-Instead of using both the information whether a new data point is an outlier and the radius of the hypersphere to indicate change, we only used the latter model properties.
-The increase of the radius of the constructed hypersphere, which encloses the data objects and increases with heterogeneous segments of data, is extracted and used for indication of change.
+From these methods we have created our own method \gls{ocs-hats} and made it puclibly available \cite{vlasveld2014hats}, which is a slightly modified implemention of \gls{svcpd} by Camci~\cite{camci2010change} applied to \gls{har} data.
+% Instead of using both the information whether a new data point is an outlier and the radius of the hypersphere to indicate change, we only used the latter model properties.
+An increase of the radius of the constructed hypersphere, which encloses the data objects and increases with heterogeneous segments of data, is extracted and used for indication of change.
 
@@
 We have constructed \gls{ocs-hats} as a slightly simplification of \gls{svcpd} by Camci~\cite{camci2010change} application to real-world inertial sensor data.
+The \textsc{Matlab} implementation is free to use for the public \cite{vlasveld2014hats}.
 We have tested our proposed method with the same artificial data and did not find a deterioration in the results.
+Because of the noisy nature of the recorded activities and the ability to segment the time series, we have shown that \gls{ocs-hats} is a robust method for real-world applications.
 
@@
 
+The proposed \gls{ocs-hats} method processes the raw inertial sensor data, originating from sensors as accelerometers and magnetic field sensors.
+This in contrast with many other approaches that extract features in a pre-processing stage.
+The method processes the data in an online manner, which makes it suitable for applications that need direct segmentation of time series data.
+
 Finally, due to encountered problems with commonly used data sets \cite{kwapisz2011activity,anguita2012human}, we have recorded our real-world data sets by ourselves.
@@
 Furthermore, we have recorded the performing subjects from the third person point of view with a video camera.
-This allows for better interpretation of the results.
-All the recorded data is made publicity available under the \gls{almende-data} data set \cite{vlasveld2014acras}.
-
-\TODO{Vraag aan Anne: toevoegen samenvatting van resultaten uit \Cref{Chapter6}: de subjectieve waarnemingen?}
+This allows for better interpretation of the results, since it provides a ground truth which can be consulted in case of ambiguity.
+All the recorded data is made publicly available as the \gls{almende-data} data set \cite{vlasveld2014acras}.
+Using our results as a benchmark, it is possible for others to use our data set and improve on it.
 
 \section{Limitations}\label{sec:limitations}
-Our proposed method was constructed under the assumption that an explicit temporal segmentation of inertial sensor data can aid the process of activity classification.
+Our proposed method \gls{ocs-hats} was constructed under the assumption that an explicit temporal segmentation of inertial sensor data can aid the process of activity classification.
 We have expressed the problem of temporal segmentation as finding change points in the time series data.
-Since this is the first application of \gls{oc-svm} based algorithms to real-world accelerometer data, we have limited ourselves to creating a first method.
+Since this is the first application of \gls{oc-svm} based algorithms to real-world accelerometer data, we have limited ourselves to creating an initial method, without optimization.
 The (results of the) proposed method is subject to parameters, which we have set manually based on empirical results and experience.
@@
 We have applied our method to real-world data, recorded in 7 runs.
-For a better and more robust analysis of the algorithm it can be applied to more real-world data.
-Although we have recorded and publicized our data, application to other commonly used data sets would enable a fair comparison.
-Furthermore, since the manual annotation of change points is an ambiguous task, the results remain subjective.
+For a better and more robust analysis of the algorithm it can be applied to other real-world data sets.
+Although we have recorded and publicized our data, it would be nice if others also record continuous data sets with ground truth labeling, for fair comparison.
+However, due to the application domain and ambiguity, the results remain subjective.
+
+For the implementation of \gls{ocs-hats} we have created and used a \textsc{Matlab} application.
+Because of that, we can not make reliable claims on the speed and memory usage of the algorithm.
+With the online nature and applications in mind and different targeted hardware platforms, this is something that could be of importance.
 
@@
 We believe that the temporal segmentation of time series data can aid the process of activity classification, which is a logical follow-up on this research.
-\TODO{Vraag aan Anne: Hier nog meer samenvatten? Wat is een goede uitsmijter?}
+
+We have created a continuous recorded data set \gls{almende-data}~\cite{vlasveld2014acras} from inertial sensors found in typical modern smartphones.
+The transition periods between the activities are part of the time series data.
+This makes the data set useful for research on segmentation.
+Besides the manually annotated change points can the ground truth of the performed activity be determined by using the provided video recordings of the activities.
+
+With our method for temporal segmentation of human activities, \gls{ocs-hats}~\cite{vlasveld2014hats}, we believe to have contributed to the field of machine learning in the context of \acrlong{har}.

### 2014-01-17 main.tex
Feedback Chapter 1 (not everything)
Finished feedback Chapter 1
Decide on OCS-HTS. Working on algorithm in chapter 4
Move results Chapter 6 to Appendix'

@@
 \DeclarePairedDelimiter\norm{\lVert}{\rVert}%
+
+\usepackage{algorithm}
+\usepackage{algpseudocode}
+\renewcommand{\algorithmicrequire}{\textbf{Input:}}
+\renewcommand{\algorithmicensure}{\textbf{Output:}}
 
@@
 
-\textit{``All truly great thoughts are conceived by walking."}
-
+\textit{``All truly great thoughts are conceived while walking."}
 \begin{flushright}
@@
 
-The Thesis Abstract is written here (and usually kept to just this page). The page is kept centered vertically so can expand into the blank space above the title too\ldots
+% The Thesis Abstract is written here (and usually kept to just this page). The page is kept centered vertically so can expand into the blank space above the title too\ldots
+
+In this thesis we construct a method to create a temporal segmentation on time series data recorded during human activities by inertial sensors in smartphones.
+Our assumption is that information about changes between activities can aid activity classification methods, since it adds information about sequential data points.
+We explore current temporal segmentation methods and relate the problem to change detection.
+A change in the performed activity results in a change in the data distribution of a time window.
+Using a \acrlong{oc-svm} algorithm we model the data distribution in the shape of a high dimensional hypersphere.
+A change in the distribution can then be discovered by analyzing the radius of the hypersphere, which will increase in such events.
+
+We implement an algorithm to process real-world time series data and find changes in performed activities.
+The data needs to be recorded in a continuous nature, with the transition periods between activities present.
+Since common available data sets do not meet this requirement, we have recorded our own data set and made it publicly available.
+A ground truth is provided both as manually annotated change points and video recordings of the activities.
+
+The constructed method, \acrlong{ocs-hats}, is able to find changes between performed activities in an unknown environment.
 }
-
-\TODO{Write abstract}
 
@@
 
-\acknowledgements{\addtocontents{toc}{\vspace{1em}} % Add a gap in the Contents, for aesthetics
+\acknowledgements{\addtocontents{toc}{\vspace{0.8em}} % Add a gap in the Contents, for aesthetics
 
-The acknowledgements and the people to thank go here, don't forget to include your project advisor\ldots
+% The acknowledgements and the people to thank go here, don't forget to include your project advisor\ldots
+Writing a thesis is hard.
+It starts with the moment you are thinking about a subject and ends with the last part, where one should write the acknowledgments to any who deserves it and express thanks to those you owe it.
+Besides all the authors who went before me in this subject and inspired me to continue in their line of research, there are a few people who I would like to say thanks to explicitly.
+
+First, I am thankful for the attention from my supervisor at \textsc{Almende}, Anne van Rossum.
+We had great discussions, in which I often needed to convince him of my ideas, which helped me with finding a domain and problem for my research.
+I think it is great that \textsc{Almende} and \textsc{Sense} facilitate so much research for master students, in a very pleasant environment.
+Everybody contributes to a atmosphere of research on new possibilities and applications, heavily relying on artificial intelligence.
+
+Less frequent visited, but of no less importance, is my supervisor at the \textsc{Utrecht University}, Gerard Vreeswijk.
+During our meetings he gave me great advise and feedback on my writing skills and techniques.
+I am thankful for his feedback and guidance and allowing me to find a subject within my full interest.
+
+Full immersion in a subject can narrow the field of vision.
+Without reflecting on the subject from a small distance and a different point of view, the primary goals can easily become obfuscated.
+The (for both sides) explanatory conversations with, amongst others, my partner Inge Dalm, my parents, my brother Floris, my sister Esmee, their partners, my housemate Yvo Zollner, those who are gathered under the heading of Lemnisco, those who call themselves `Vlassies', those who enjoy golden liquids, and those who are fond of lunacy, helped me understand my research problem at different levels.
+
+Already mentioned, my parents (in the widest sense) have been of great support in any way.
+It is their encouragement to explore the world and question everything that has led me to my passion for computer science, artificial intelligence,  and machine learning.
+
+Finally, but certainly not least, I want to thank my partner Inge Dalm.
+Without her and the time together I could not have completed this thesis.
+I am very thankful for her and her vision on this matter.
+She encouraged me during the process and bared with me during the harder moments.
+
+With all the people mentioned above, and those who are not mentioned but know they should have, this research has become possible.
+Without them, writing a thesis is even harder.
 }
-\TODO{Write some nice stuff here}
+% \TODO{Write some nice stuff here}
 \clearpage % Start a new page
@@
 
+\input{./Appendices/AppendixA-Results}
 % \input{./Appendices/AppendixA-Summaries}

### 2014-01-20 Appendices/AppendixA-Results.tex
Updates

@@
   \centering
-  \caption[Results real world runs]{Parameter settings and results of the real-world data sets.}
+  \caption[Results real world runs]{Parameter settings and results of the real-world data sets.
+  The table shows the \gls{far} $\gamma$, average delay $d_{avg}$ and its spread, using the standard deviation.
+  Low values for the results are better.}
   \begin{tabulary}{\textwidth}{|l|c|c|c|c|c|c|c|}

### 2014-01-20 Chapters/Chapter4/data_gathering.tex
Move results Chapter 6 to Appendix'
Updates

@@
 For the experiments we used a smartphone with inertial sensors as recording device.
-The activities were recorded using a free \textsc{Android} application \cite{sensorlogger}.
+The activities were recorded using a free \textsc{Android} application \textsc{Sensor Logger}\footnote{https://play.google.com/store/apps/details?id=
+com.kzs6502.sensorlogger}.
 This application was chosen for its convenient data format of the sensor recording and its regularity of the sampling interval.
@@
 
-We implemented our algorithm in \textsc{MATLAB}.
+We implemented our algorithm in \textsc{Matlab}.
 For the \gls{svdd} model construction phase we used the \textsc{dd\_tools} library by Tax \cite{Ddtools2013}, which depends on the \textsc{PR\_tools} package, originating from the book by Van Der Heijden \etal \cite{van2005classification}.

### 2014-01-20 Chapters/Chapter5/results.tex
Chapter 6 run order
Updates

@@
 For each data set it shows the high and low thresholds used in the \gls{rt} based change detection method.
-During the post-processing all detected change points within a certain \emph{closeness} range are merged, for which the value differs per data set.
+During the post-processing all detected change points within a certain proximity $\delta$ are merged, for which the value differs per data set.
 The difference between the actual and final detected change points are expressed as an offset.
@@
   \centering
-  \caption[Results artificial data sets]{Parameter settings and results of the artificial data sets.}
+  \caption[Results artificial data sets]{Parameter settings and results of the artificial data sets.
+  The table shows the \gls{far} $\gamma$, average delay $d_{avg}$ and its spread, using the standard deviation.
+  Low values for the results are better.}
   \begin{tabulary}{\textwidth}{|l|c|c|c|c|}
@@
     \hline
-    Closeness & 10 & 10 & 10 & 50 \\
+    $\delta$ & 10 & 10 & 10 & 50 \\
     \hline

### 2014-01-20 Chapters/Chapter6/results.tex
Updates

@@
 That will be in the form of a tabular overview of the (manually set) best meta-parameters and results, expressed in \gls{far} $\gamma$ and average delay $d_{avg}$ (following \Cref{sec:artificial_data_results}).
-Furthermore, we have anaylized the results from an emperical point of view, since the problem of finding change points in human activities is ambiguous.
-For the sake of brevity we excluded those observations from this chapter and instead xthey can be found in \Cref{AppendixA}.
+Furthermore, we have analyzed the results from an empirical point of view, since the problem of finding change points in human activities is ambiguous.
+For the sake of brevity we excluded those observations from this chapter; they can be found in \Cref{AppendixA}.
 
@@
   \centering
-  \caption[Results real world runs]{Results of the real-world data sets. The used parameters are: window length = $50$, $\sigma = 13$, $th_{high} = 1.2$, $th_{low} = 0.8$}
+  \caption[Results real world runs]{Results of the real-world data sets. The used parameters are: window length = $50$, $\sigma = 13$, $th_{high} = 1.2$, $th_{low} = 0.8$, $\delta = 0.7$.
+  The table shows the \gls{far} $\gamma$, average delay $d_{avg}$ and its spread, using the standard deviation.
+  Low values are better.
+  }
   \begin{tabulary}{\textwidth}{|l|c|c|c|c|c|c|c||c|}

### 2014-01-20 Chapters/Chapter7.tex
Updates

@@
 To do so, we used the \gls{svdd} algorithm by Tax~\cite{tax1999support}, following the method of applying it to a sliding window of data, as done by Camci~\cite{camci2010change}.
-From these methods we have created our own method \gls{ocs-hats} and made it puclibly available \cite{vlasveld2014hats}, which is a slightly modified implemention of \gls{svcpd} by Camci~\cite{camci2010change} applied to \gls{har} data.
+From these methods we have created our own method \gls{ocs-hats} and made it publicly available \cite{vlasveld2014hats}, which is a slightly modified implementation of \gls{svcpd} by Camci~\cite{camci2010change} applied to \gls{har} data.
 % Instead of using both the information whether a new data point is an outlier and the radius of the hypersphere to indicate change, we only used the latter model properties.

### 2014-01-20 Chapters/Chapter1.tex
Changed objective/subjective to quantitative/qualitative

@@
 
-Besides the recorded data sets we have used artificial data sets to obtain an objective performance measure of the method.
+Besides the recorded data sets we have used artificial data sets to obtain an objective, quantitative, performance measure of the method.
 These artificial data sets are modeled to the sets used in \cite{camci2010change,takeuchi2006unifying}.
@@
 That relation is applied to our problem formulation in \Cref{Chapter4}, where we construct our change detection method \gls{ocs-hats}~\cite{vlasveld2014hats}, based on the work of Camci~\cite{camci2010change}.
-In \Cref{Chapter5} we apply the method to artificial data and show the objective performance.
+In \Cref{Chapter5} we apply the method to artificial data and show the quantitative performance.
 In \Cref{Chapter6} the method is applied to our real-world \gls{almende-data} \cite{vlasveld2014acras}.
-In that chapter we show the ability of detecting change in \gls{har} data, recorded by inertial sensors.
+In that chapter we show, by qualitative analysis, the ability of detecting change in \gls{har} data, recorded by inertial sensors.
 This thesis is concluded in \Cref{Chapter7}, in which we reflect on the performed research and state possibilities for feature research.

### 2014-01-20 Chapters/Chapter4/data_gathering.tex
Changed objective/subjective to quantitative/qualitative

@@
 \subsection{Artificial data}\label{subsec:data_gathering_artificial}
-In order to provide an objective comparison to other methods, we will use artificial data sets which are also used in the earlier work on which \gls{ocs-hats} is based.
+In order to provide an objective, quantitative, comparison to other methods, we will use artificial data sets which are also used in the earlier work on which \gls{ocs-hats} is based.
 These are the data sets used by Takeuchi and Yamanishi \cite{takeuchi2006unifying} and Camci \cite{camci2010change}.
@@
 Over the different data series the mean and variance of the Gaussian random variable $\epsilon_t$ differs and changes at pre-determined change points.
-Using this data set an objective quality measure over the change detection methods can be obtained and compared.
+Using this data set an quantitative quality measure over the change detection methods can be obtained and compared.
 All the used data sets are listed and analyzed in \Cref{Chapter5}.
@@
 A video recording from the performed activity is used to annotate the time series with real change points.
-The discovered change points are compared with these annotated change points to give a subjective quality measure.
+The discovered change points are compared with these annotated change points to give a qualitative quality measure.
 In \Cref{Chapter6} we give a detailed analysis of the performed activities and the recorded data sets.

### 2014-01-20 Chapters/Chapter5/artificial_data_generation.tex
Changed objective/subjective to quantitative/qualitative

@@
 \section{Artificial data generation}\label{sec:artificial_data_generation}
-In this section we present four data sets, as used by Camci~\cite{camci2010change} and Takeuchi and Yamanishi \cite{takeuchi2006unifying}, to provide for a objective performance comparison.
+In this section we present four data sets, as used by Camci~\cite{camci2010change} and Takeuchi and Yamanishi \cite{takeuchi2006unifying}, to provide for a quantitative performance comparison.
 All the data sets are modeled by a second-order \gls{ar}-model as in \Cref{eq:artificial_data_sets_model}, where $\epsilon_t$ is a Gaussian distribution modeling the noise with mean $0$ and variance $\sigma^2 = 1$, $a_1 = 0.6$ and $a_2 = -0.5$.

### 2014-01-20 Chapters/Chapter5/quality_metrics.tex
Chapter 6 run order
Changed objective/subjective to quantitative/qualitative

@@
 \section{Quality metrics}\label{sec:artificial_data_quality_metrics}
-For an objective measure of quality of the proposed method applied to the four aforementioned data sets, we employ two metrics, inspired by and slightly altered from Takeuchi and Yamanishi~\cite{takeuchi2006unifying}.
+For an quantitative measure of quality of the proposed method applied to the four aforementioned data sets, we employ two metrics, inspired by and slightly altered from Takeuchi and Yamanishi~\cite{takeuchi2006unifying}.
 The first metric is the average delay $d_{avg}$, for which we first need to define the notion of delay $d$ for each change:
@@
 \begin{itemize}
-  \item \textbf{Window length}: The window length that is processed by the \gls{svdd} algorithm. This is the number of data points that are continuously modeled in the shape in a hypersphere.
+  \item \textbf{Window length}: The window length that is processed by the \gls{svdd} algorithm.
+  This is the number of data points that are continuously modeled in the shape in a hypersphere.
   \item \textbf{Sigma of \gls{rbf}}: The value of $\sigma$ for the \gls{rbf} kernel. The effect of this parameter has been discussed in \Cref{subsec:svm_model_parameters}.
-  \item \textbf{High threshold}: The $th_\text{high}$ threshold value for the \gls{rt} method described in \Cref{subsec:ratio_thresholding}. A higher value indicates a more precise and robust setting.
-  \item \textbf{Low threshold}: The $th_\text{low}$ threshold value for the \gls{rt} method described in \Cref{subsec:ratio_thresholding}. A lower value indicates a more precise and robust setting.
-  \item \textbf{Closeness}: In the post-processing phase the detected change points which are within a certain \emph{closeness} range will be merged together.
+  \item \textbf{High threshold}: The $th_\text{high}$ threshold value for the \gls{rt} method described in \Cref{subsec:ratio_thresholding}.
+  A higher value indicates a more precise and robust setting.
+  \item \textbf{Low threshold}: The $th_\text{low}$ threshold value for the \gls{rt} method described in \Cref{subsec:ratio_thresholding}.
+  A lower value indicates a more precise and robust setting.
+  \item \textbf{Proximity}: In the post-processing phase the detected change points which are within a certain range $\delta$ will be merged.
   This parameter indicates the minimal number of data points that change points must differ in order to be considered separate detections.

### 2014-01-20 Chapters/Chapter6.tex
Updates
Changed objective/subjective to quantitative/qualitative

@@
 That will illustrate the setting in which the activities are performed and the characteristics of continuous data.
-Finally, the results of the algorithm are discussed in \Cref{sec:real_world_results}.
-We will use the objective measures, introduced in \Cref{sec:artificial_data_quality_metrics}, and also give subjective claims on the performance of the algorithm applied to the real-world data.
-Since the data sets used are real-world recordings, we will put emphasis on the latter type of quality measures \ie the subjective claims.
+The results of the algorithm are discussed in \Cref{sec:real_world_results}.
+Finally, we provide a list of possible improvements in \Cref{sec:possible_improvements}.
+
+The results are measured with both quantitative measures, as introduced in \Cref{sec:artificial_data_quality_metrics}, and qualitative claims on the performance of the algorithm applied to the real-world data.
+Since the data sets used are real-world recordings, we will put emphasis on the latter type of quality measures \ie the qualitative claims.
 

### 2014-01-20 Chapters/Chapter6/data_sets.tex
Updates
Changed objective/subjective to quantitative/qualitative

@@
 The used phones are the \textsc{HTC Sensation XE} and \textsc{HTC Desire}, running the \textsc{Android} smartphone operating system.
-During the activities the inertial sensor data is recorded using a free application, \textsc{Sensor Logger}~\cite{sensorlogger}.
+During the activities the inertial sensor data is recorded using a free application, \textsc{Sensor Logger}.
 This application records the sensor data to \textsc{CSV} files, which are then transformed to \textsc{MATLAB} compatible files.
@@
 In \Cref{fig:stills_subject_2_change_point} a transition between walking and running is shown.
-It is impossible to determine the exact location of the change point and thus objective quality measures (such as the delay) are of less importance.
+It is impossible to determine the exact location of the change point and thus quantitative quality measures (such as the delay) are of less importance.
 

### 2014-01-20 Chapters/Chapter6/results.tex
Changed objective/subjective to quantitative/qualitative

@@
 In this section we will provide the results of our methods \gls{ocs-hats} applied to the real-world data sets and discuss the quality of the solution.
-As stated above, we will use the objective quality metrics as used in \Cref{sec:artificial_data_quality_metrics}.
+As stated above, we will use the quantitative quality metrics as used in \Cref{sec:artificial_data_quality_metrics}.
 That will be in the form of a tabular overview of the (manually set) best meta-parameters and results, expressed in \gls{far} $\gamma$ and average delay $d_{avg}$ (following \Cref{sec:artificial_data_results}).
@@
 
-\subsection{Objective metrics}
+\subsection{Quantitative metrics}
 The employed two-stage algorithm requires the setting of user defined parameters.
@@
 
-\subsection{Subjective observations and remarks}\label{subsec:subjective_results}
+\subsection{Qualitative observations and remarks}\label{subsec:subjective_results}
 It can be ambiguous where to place the (manually annotated) change points, in the case of real-world data.
-The objective quality measure can give a distorted view on the performance.
+The quantitative quality measure can give a distorted view on the performance.
 To overcome this, we will state observations and remarks found after close (visual) inspection of the annotated and discovered change points.

### 2014-01-20 Chapters/Chapter7.tex
Changed objective/subjective to quantitative/qualitative

@@
 This was performed in a real-world setting, showing the robustness of the method for noisy data.
-Since manual labeling (the annotation of change points on the time series data) of human activities is ambiguous we have put emphasis on the subjective inspection of the results.
+Since manual labeling (the annotation of change points on the time series data) of human activities is ambiguous we put emphasis on the qualitative inspection of the results.
 The proposed method gives fairly good results, although parameter tuning is a subject of further research.

### 2014-01-20 Chapters/Chapter2/density_ratio.tex
Consistent `non-` usage

@@
 A limitation of these methods is that they rely on pre-specified parametric models.
-Nonparametric models, for which the number and nature of the parameters are undetermined, for density estimation have been proposed, but it is said to be a hard problem \cite{hardle2004nonparametric, sugiyama2012density}.
+Non-parametric models, for which the number and nature of the parameters are undetermined, for density estimation have been proposed, but it is said to be a hard problem \cite{hardle2004nonparametric, sugiyama2012density}.
 A solution to this is to estimate the \emph{ratio} of probabilities instead of the probabilities themselves.
@@
 The method by Liu \etal \cite{liu2013change} estimates the ratio of probabilities directly, instead of estimating the densities explicitly.
-Other methods using ratio-estimation are the Unconstrained Least-Squares Importance Fitting (uLSIF) method \cite{kanamori2009least}, and an extension which possesses a superior nonparametric convergence property: Relative uLSIF (RuLSIF) \cite{yamada2013relative}.
+Other methods using ratio-estimation are the Unconstrained Least-Squares Importance Fitting (uLSIF) method \cite{kanamori2009least}, and an extension which possesses a superior non-parametric convergence property: Relative uLSIF (RuLSIF) \cite{yamada2013relative}.

### 2014-01-20 Chapters/Chapter3/one_class_svm.tex
Consistent `non-` usage

@@
 Instead of estimating the density function of an distribution $P$, it focuses on an easier problem: the algorithm finds regions in the input where the ``probability density lives''.
-This results in a function such that most of the data is in the region where the function is nonzero.
+This results in a function such that most of the data is in the region where the function is non-zero.
 
@@
   \item $\nu$ is an upper bound on the fraction of outliers, \ie training examples regarded as out-of-class.
-  \item $\nu$ is a lower bound on the fraction of \glspl{sv}, \ie training examples with a nonzero Lagrange multiplier $\alpha_i$.
+  \item $\nu$ is a lower bound on the fraction of \glspl{sv}, \ie training examples with a non-zero Lagrange multiplier $\alpha_i$.
 \end{enumerate}
-When $\nu$ approaches $0$, the penalty factor for nonzero Lagrange multipliers ($\frac{1}{\nu n}$) becomes infinite, and thus the solution resembles a \emph{hard margin} boundary.
+When $\nu$ approaches $0$, the penalty factor for non-zero Lagrange multipliers ($\frac{1}{\nu n}$) becomes infinite, and thus the solution resembles a \emph{hard margin} boundary.
 

### 2014-01-20 Chapters/Chapter1.tex
Remove `unknown` as environment description

@@
 The sensors, such as the accelerometer, are nowadays widely available in smartphones.
-We are interested in real-world data from both in- and outdoor activities in an unknown environment and in a continuous manner.
+We are interested in real-world data from both in- and outdoor activities performed in a continuous manner.
 
@@
 These time series are recorded during the performance of real-world human activities.
-The setting of the activities are real-world unknown environments.
+The setting of the activities are real-world environments, both in- and outdoor.
 
@@
 Our own data set is recorded by placing a smartphone with inertial sensors in the front right pants pocket of a subject.
-The subject was asked to perform common and simple activities in an uncontrolled environment, both in- and outdoors.
+The subject was asked to perform common and simple activities in a natural environment, both in- and outdoors.
 During the performance the subjects were also recorded with a video camera from a third-person point of view.

### 2014-01-20 Chapters/Chapter6.tex
Remove `unknown` as environment description

@@
 In \Cref{sec:data_sets} we discuss available data sets and argue that these are inappropriate for our research.
-It is followed by an overview of the setup we have created to record our own continuous data sets in an unknown environment, both for in- and outdoor activities.
+It is followed by an overview of the setup we have created to record our continuous data sets in a real-world environment, both for in- and outdoor activities.
 The series of activities performed are discussed in \Cref{sec:data_gathering}.

### 2014-01-20 Chapters/Chapter6/data_sets.tex
Remove `unknown` as environment description

@@
   \caption[WISDM Excerpt]{Excerpt from the WISDM data set \cite{kwapisz2011activity}.
-First five activities for subject $33$ are displayed.
+The first five activities of subject $33$ are displayed.
 Due to the discontinuous nature of the recordings, this data set can not be used for change detection.}
@@
   \caption[UCI HAR Excerpt]{Excerpt from the UCI HAR data set \cite{anguita2012human}.
-The recorded training activities for subject $25$ are displayed.
+The recorded training activities of subject $25$ are displayed.
 Due to the imprecise (and possibly incorrect) labeling and the short duration of activities, the data set can not be used for the change detection method as described in this thesis.}
@@
 Because of the above described shortcomings of commonly used data sets, we have decided to record activities by ourselves.
-The following characterizing requirements were set for our data sets:
+The following requirements were set for our data sets:
 \begin{enumerate}
   \item The activities need to be performed and recorded in a continuous manner,
-  \item The environment should be natural and uncontrolled, both in- and outdoor,
+  \item The environment should be natural and both in- and outdoor,
   \item The data should be clearly annotated, as objective as possible,

### 2014-01-20 Chapters/Chapter3/one_class_svm.tex
Final 1.0

@@
   \includegraphics[width=0.8\textwidth]{./Figures/chapter3/svm_mapping_spaces.pdf}
-  \caption[Mapping spaces in \gls{svm}]{Schematic visualization of the non-linear mapping of \glspl{svm}.
+  \caption[Mapping spaces in Support Vector Machines]{Schematic visualization of the non-linear mapping of \glspl{svm}.
   The input vector $\vectorsym{x}$ from input space $\mathcal{I}$ is mapped by a non-linear function $\phi(\vectorsym{x})$ to a feature vector $\vectorsym{z}$ in the high dimensional feature space $\mathcal{F}$.
@@
   \includegraphics[width=0.5\textwidth]{./Figures/chapter3/svm_separating_plane_with_margin.png}
-  \caption[\gls{svm} and the separating hyperplane]{Illustration of the separating hyperplane of an \gls{svm}.
+  \caption[Support Vector Machines and the separating hyperplane]{Illustration of the separating hyperplane of an \gls{svm}.
   Here $w$ is the normal vector for the separating hyperplane and size of the margin is $\frac{2}{\lVert{w}\rVert}$.
@@
 
-\subsection{\acrlong{nu-svm}}\label{subsec:nu-svm}
+\subsection{\texorpdfstring{$\nu$-Support Vector Machine}{nu-Support Vector Machine}}\label{subsec:nu-svm}
 The first of the \gls{oc-svm} methods we will discuss is often referred to as \gls{nu-svm} and introduced by Sch\"olkopf \etal \cite{scholkopf1999support}.
@@
     \includegraphics[width=0.5\textwidth,keepaspectratio]{./Figures/chapter3/nu-svm.pdf}
-  \caption[\gls{nu-svm}]{Graphical representation of \gls{nu-svm}.
+  \caption[$\nu$-Support Vector Machine illustration]{Graphical representation of \gls{nu-svm}.
 The separating hyperplane $w \cdot \phi(x_i) + \rho = 0$ create a maximal margin, in the feature space, between the data points and the origin.
@@
 
-\subsection{\acrlong{svdd}}\label{subsec:oc-svm-svdd}
+\subsection{\texorpdfstring{\acrlong{svdd}}{SVDD}}\label{subsec:oc-svm-svdd}
 The method introduced by Tax and Duin \cite{tax1999support}, known as \acrlong{svdd}, follows a spherical instead of planar approach.
@@
     \includegraphics[width=0.8\textwidth,keepaspectratio]{./Figures/chapter3/nu-vs-svdd.pdf}
-  \caption[Difference \gls{nu-svm} and \gls{svdd}]{Graphical representation of the difference between \gls{nu-svm} (left) and \gls{svdd} (right).
+  \caption[Relation between $\nu$-SVM and SVDD]{Graphical representation of the difference between \gls{nu-svm} (left) and \gls{svdd} (right).
 Note that for the sake of simplicity the kernel functions are not applied.}
@@
     \includegraphics[width=0.5\textwidth,keepaspectratio]{./Figures/chapter3/svdd-boundary.pdf}
-  \caption[\gls{svdd} boundary]{The \gls{svdd} method trained on a banana-shaped data set with different sigma-values for the \gls{rbf} kernel.
+  \caption[SVDD boundary with sigma parameter]{The \gls{svdd} method trained on a banana-shaped data set with different sigma-values for the \gls{rbf} kernel.
 Solid circles are support vectors, the dashed line is the boundary.
@@
     \includegraphics[width=0.8\textwidth,keepaspectratio]{./Figures/chapter3/svdd-parameter-sigma.pdf}
-  \caption[\gls{svdd} boundary]{The \gls{svdd} method trained on a banana-shaped data set with different $\sigma$-values for the \gls{rbf} kernel.
+  \caption[SVDD boundary with sigma parameter]{The \gls{svdd} method trained on a banana-shaped data set with different $\sigma$-values for the \gls{rbf} kernel.
 Solid circles are support vectors, the dashed line is the boundary.}

### 2014-01-20 Chapters/Chapter6/data_gathering.tex
Final 1.0

@@
 The performed activities consist, as with the first set, of standing, walking, and running.
-Furthermore, the segments are performed in a circular manner.
-\eg the running activities are performed around a fountain with an estimated diameter of 10 meters.
+Furthermore, some segments are performed in a circular manner.
+For example, the running activities are performed around a fountain with an estimated diameter of 10 meters.
 The plots of the recordings of Subject 2, for the accelerometer, rotation, and magnetometer are shown in \Cref{fig:plots_subject_2}.
@@
   \begin{subfigure}[b]{1\textwidth}
+    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/stairs-1-marc/data_plot_acc_annotated.eps}
+    \caption{Accelerometer}
+    \label{fig:data_gathering_run_7_acc}
+  \end{subfigure} \\
+
+  \begin{subfigure}[b]{1\textwidth}
     \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/stairs-1-marc/data_plot_rot_annotated.eps}
-    \caption{Run 7: Indoor stairs subject 3, rotation}
+    \caption{Rotation}
     \label{fig:data_gathering_run_7_rot}
@@
     \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/stairs-1-marc/data_plot_mag_annotated.eps}
-    \caption{Run 7: Indoor stairs subject 3, Mag}
+    \caption{Magnetic field}
     \label{fig:data_gathering_run_7_mag}
-  \end{subfigure} \\
+  \end{subfigure}
 
-  \begin{subfigure}[b]{1\textwidth}
-    \includegraphics[width=\textwidth]{./Figures/chapter6/data_collection/stairs-1-marc/data_plot_acc_annotated.eps}
-    \caption{Run 7: Indoor stairs subject 3, accelerometer}
-    \label{fig:data_gathering_run_7_acc}
-  \end{subfigure}
-  \caption[Plots subject 3]{Annotated plots of a full recording performed by Subject~$3$ while walking the stairs.
+  \caption[Plots subject 3]{Annotated plots of a full recording performed by Subject~$3$ while walking the stairs, indoor.
 The vertical lines are manually determined change points.

### 2014-01-20 Chapters/Chapter6/data_sets.tex
Final 1.0

@@
 Using the video recordings, the data sets are manually annotated, as shown in the plots of \Cref{fig:plots_subject_1,fig:plots_subject_2,fig:plots_subject_3}.
-The graphs display each from a single recording the inertial sensor values (see \Cref{tab:recorded_metrics} for explanation of the metrics), with manually annotated activities above the data.
-The manually determined change points are used to draw quality conclusion over the algorithm's performance, by comparing it to the discovered change points.
-All of the used data sets, including the video material and manual labeling, are made available to the public as the \acrlong{almende-data} \cite{vlasveld2014acras} for further research.
+The graphs display each from a single recording the inertial sensor values (see \Cref{tab:recorded_metrics} for an explanation of the metrics), with manually annotated activities above the data.
+The manually and discovered change points are compared to draw quality conclusions over the algorithm's performance.
+All of the used data sets, including the video material and manual labeling, are made available to the public as the \glsfirst{almende-data} \cite{vlasveld2014acras} for further research.
 The following section discusses the performed activities during the recordings.

### 2014-01-20 Chapters/Chapter6/results.tex
Final 1.0

@@
 In general, a high value for the upper threshold and low value for the lower threshold indicates a robust method, since it indicates that change points are significantly different from homogeneous segments.
-The proximity $\delta$ should be low, since that indicates there are few false positives.
+The chosen proximity $\delta$ should be low, since that indicates there will be few false positives.
 
@@
 
-\begin{table}
+\begin{table}[H]
   \centering
@@
 
-\begin{figure}
+\begin{figure}[H]
 \centering
@@
 Depending on the parameters for the sensitivity adjustment, almost all changes are discovered.
-With a high sensitivity it seems, initially, the method produces many false positives.
-After closer inspecting of the data, we can confirm that there is a change in the data, although the performed activity stayed the same.
-This is also a weakness of the method: since the sensitivity is set globally, it can produce many semi-false positive change points for some segments of a run.
 
-An other positive observation is that the method often recognizes a change point \emph{before} our annotation.
+Our first observation is that with a high sensitivity it seems that the method produces many false positives.
+Although the performed activity did not change around those false positives, there was a change in the recorded data.
+This is a drawback of the method: since the sensitivity is set globally, it can produce false positive change points for some segments of a run.
+
+The second observation is that the method often recognizes a change point \emph{before} our annotation.
 Again, after closer inspection we can see that the data changes shape and distribution before we annotated a change based on the video recordings.
-This effect can be observed in the case of a transition from running to walking; the last `steps' of running are already recognized of being different.
+This effect can be observed in the case of a transition from running to walking; the last `steps' of running are already labeled as change points.
 As with the above behavior, this is subject to parameter settings.
 
-Considering the overall performance and results, we have the following observations of cases in which the method has problems with detecting change:
+Considering the overall performance and results, we have the following observations:
 \begin{itemize}
-  \item \textbf{Merging:} in our method we use a proximity time period $\delta$ (usually up to $0.5$ - $1.0$ seconds) to merge discovered change points that have a very small distance between them.
+  \item \textbf{Merging:} in our method we use a proximity time period $\delta$ (usually up to $0.5$ - $1.0$ seconds) to merge discovered change points that are close together.
   Multiple merging-strategies can be applied, and in our method we use the naive implementation by simply ignoring all the discovered change points that occur less than $\delta$ seconds after the previous change point.
@@
   This problem on itself can be formulated as a change detection problem.
-  \item \textbf{Masking:} Since the parameters for the detection method are set globally, it is difficult to discover all the change points (and only the change points) without a high \gls{far}.
+  \item \textbf{Masking:} Since the parameters for the detection method are set globally, it is difficult to discover all the change points (and only the change points) without a high false alarm rate.
   If a change point is proceeded by a noisy block, then the real change point will be merged with the previous change points.
-  In an other case, where a change point is very clear represented in the data but the next, close by, other change point is more subtle, the latter change point is \emph{masked} by the former.
+  In an other case, where a change point is very clear represented in the data but the next near change point is more subtle, the latter change point is \emph{masked} by the former.
   This masking effect is also discussed in \cite{inclan1994use}, where an iterated approach is applied.
-  \item \textbf{False heterogeneity:} For our video-data synchronization, we started and ended each recording with a few seconds in which the recording smartphone was kept still in the air.
-  During this period, the data variance, and thus the constructed hypersphere, becomes very small.
-  As a results, even small movements are considered to be changes and in the case of the mid-air still smartphone a lot of false change points are detected.
-  Due to the merging effect described above, the final real change point (often shaking the smartphone) is not discovered.
+  \item \textbf{Variance and threshold levels:} For our video synchronization, we started and ended each recording with a few seconds in which the recording smartphone was kept still in the air.
+  Then we shake the smartphone to get a sudden change in the inertial sensor data, so the video and sensor data stream can easily be aligned.
+  During the still period, the data variance, and thus the constructed hypersphere, becomes very small.
+  As a results, even small movements are considered to be changes and in the case of a mid-air kept smartphone a lot of false change points are detected.
+  Due to the merging effect described above, the final real change point (shaking the smartphone) is not discovered.
   \item \textbf{Incorrect weighting:} In our analyses we have used the data from the accelerometer, magnetic field, and rotation sensors.
-  In the segments which embodied movement in a circular manner (such as walking and running around a fountain), the turn was not (or at a different time point) discovered.
-  Alternatively, when we only used the magnetic field and rotation sensors, the turn was correctly discovered.
+  In the segments which embodied movement in a circular manner (such as walking and running around a fountain), the corner was not (or at a different time point) discovered.
+  However, when we only used the magnetic field and rotation sensors, the corner was correctly discovered.
   Other transitions, such as from walking to running, are harder to discover without the (linear) accelerometer sensor data.
@@
 
-An more elaborate discussion with observations and remarks is provided in \Cref{AppendixA}.
+A more elaborate discussion with observations and remarks is provided in \Cref{AppendixA}.
 
@@
 A possible improvement would be to use locally optimized parameters, although that would require an reflective data processing approach or more a priori information about the data.
-With our approach, the lack of a priori knowledge about the data distribution is one of the design decisions.
+The design decision for our approach was to create a general method, which requires no a priori knowledge about the data distribution.
 
 Furthermore, we discovered that sometimes, especially with movements incorporating circular motions, the addition of (linear) accelerometer data makes the discovery of change points harder.
-In such cases, each of the inertial sensor data streams should be relatively weighted (\eg using the covariance) in order to signal an overall change when it only occurs in one of the streams.
+This problem can be handled by, for example, assigning weights to the sensor data streams.
+Change can then be detected when it only occurs in one of the streams.

### 2014-01-20 Chapters/Chapter7.tex
Final 1.0

@@
 The goal was to find the change points between activities, assuming that it will aid our understanding of the sensor values.
-To do so, we used the \gls{svdd} algorithm by Tax~\cite{tax1999support}, following the method of applying it to a sliding window of data, as done by Camci~\cite{camci2010change}.
+To do so, we used the \gls{svdd} algorithm by Tax~\cite{tax1999support} and applied it to a sliding window of data, as done by Camci~\cite{camci2010change}.
 From these methods we have created our own method \gls{ocs-hats} and made it publicly available \cite{vlasveld2014hats}, which is a slightly modified implementation of \gls{svcpd} by Camci~\cite{camci2010change} applied to \gls{har} data.
@@
 
-In the following section we will further discuss the proposed method for temporal segmentation of inertial sensor signals from human activities.
+In \Cref{sec:proposed_method} we will further discuss the proposed method for temporal segmentation of inertial sensor signals from human activities.
 Our main findings and contributions are discussed in \Cref{sec:main_findings}.
@@
 
-\section{What is done?}
+\section{Proposed method}\label{sec:proposed_method}
 The goal of this research was to find a temporal segmentation of recordings obtained from human activities.
@@
 
-In this research we have shown that it is possible to explicitly obtain a temporal segmentation, expressed as change points, of activities performed by humans.
+We have shown that it is possible to explicitly obtain a temporal segmentation, expressed as change points, of activities performed by humans.
 This was performed in a real-world setting, showing the robustness of the method for noisy data.
@@
 
-This research is, as far as we know, the first application of \gls{oc-svm} based methods to on-body worn inertial sensor signals, such as accelerometer data.
+The proposed method is, as far as we know, the first application of \gls{oc-svm} based methods to on-body worn inertial sensor signals, such as accelerometer data.
 Furthermore, it applies the algorithms, amongst others of Tax and Duin~\cite{tax1999support}, to real-world data.
@@
 The proposed \gls{ocs-hats} method processes the raw inertial sensor data, originating from sensors as accelerometers and magnetic field sensors.
-This in contrast with many other approaches that extract features in a pre-processing stage.
-The method processes the data in an online manner, which makes it suitable for applications that need direct segmentation of time series data.
+This contrasts with many other approaches that extract features in a pre-processing stage.
 
-Finally, due to encountered problems with commonly used data sets \cite{kwapisz2011activity,anguita2012human}, we have recorded our real-world data sets by ourselves.
-We have recorded all the activities with a smartphone worn in the right front pocket of the subjects.
-Furthermore, we have recorded the performing subjects from the third person point of view with a video camera.
-This allows for better interpretation of the results, since it provides a ground truth which can be consulted in case of ambiguity.
+Furthermore, because the it is an online method, it is suitable for applications that need direct segmentation of time series data.
+
+The commonly used data sets \cite{kwapisz2011activity,anguita2012human} are recorded in a non-continuous manner.
+The proposed method requires data streams of continuous activities.
+Annotating change points on this kind of time series can be ambigious, so we needed to have a ground truth.
+Because of these two shortcomings, we have recorded continuous real-world data sets, together with a third persion point of view video stream.
 All the recorded data is made publicly available as the \gls{almende-data} data set \cite{vlasveld2014acras}.
@@
 
-For the implementation of \gls{ocs-hats} we have created and used a \textsc{Matlab} application.
-Because of that, we can not make reliable claims on the speed and memory usage of the algorithm.
+We have impplemented \gls{ocs-hats} in \textsc{Matlab}.
+Because of that, we can not make reliable claims on the speed and memory usage of the algorithm on real-world devices such as smartphones.
 With the online nature and applications in mind and different targeted hardware platforms, this is something that could be of importance.

### 2014-01-20 main.tex
Final 1.0

@@
 
-\textsc{\LARGE \univname}\\[1.5cm] % University name
+% \begin{figure}[h]
+  % \centering
+    \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{./Figures/UU-logo2011_PANTONE.eps}
+% \end{figure}
+
+\textsc{\LARGE \univname}\\[0.5cm] % University name
 \textsc{\Large Master's Degree Thesis}\\[0.5cm] % Thesis type
@@
 {\huge \bfseries \ttitle}\\[0.4cm] % Thesis title
-\HRule \\[1.5cm] % Horizontal line
+\HRule \\[1cm] % Horizontal line
 
-\begin{minipage}{0.4\textwidth}
+\begin{minipage}[t]{0.4\textwidth}
 \begin{flushleft} \large
@@
 \end{minipage}
-\begin{minipage}{0.5\textwidth}
+\begin{minipage}[t]{0.5\textwidth}
 \begin{flushright} \large
@@
 \end{flushright}
-\end{minipage}\\[3cm]
+\end{minipage}\\[1cm]
 
@@
 \textit{in the}\\[0.4cm]
-\facname\\\deptname\\[2cm] % Research group name and department name
+\facname\\\deptname\\[1cm] % Research group name and department name
 
-{\large \today}\\[4cm] % Date
-%\includegraphics{Logo} % University/department logo - uncomment to place it
+{\large \today}\\[0.5cm] % Date
+\includegraphics{./Figures/faculty_of_science.png} % University/department logo - uncomment to place it
 
